### 2014-12-02 Manuscript/proto-draft.tex
searching matrix method

\documentclass[a4paper,11pt]{article}

\usepackage{natbib}
\usepackage{enumerate}
\usepackage[osf]{mathpazo}
\usepackage{lastpage} 
\pagenumbering{arabic}
\linespread{1.66}

\begin{document}

\begin{flushright}
Version dated: \today
\end{flushright}
\begin{center}

%Title
\noindent{\Large{\bf{Missing living mammals}}}\\
\bigskip
%Author
\noindent{Thomas Guillerme - guillert@tcd.ie - http://tguillerme.github.io/}\\

\end{center}

\section{Data collection}
We need to use the same protocol as the GMPD to make sure our data are comparable. The original version used lots of online databases but luckily we can now just use Google Scholar! Into the Google Scholar search enter the species Latin Binomial as search keywords, followed by the search terms related to parasites. So for Marmota monax:

parasite OR helminth OR virus OR bacteria OR fungi OR ectoparasite OR arthropod OR protozoa "Marmota monax"

[Note that the quotation marks (" ") tell Google that you want the exact phrase Marmota monax, and the OR means it will contain Marmota monax and parasite or Marmota monax and helminth etc.]

You then need to search the results for appropriate papers. This is tedious so I’d advise setting yourself a reasonable target, for example, do it for a few hours a day then do something else!

You should also search for each squirrel genus independently; however I would only do this for genera with few papers on the species. If you did this for Marmota spp. you could be there forever and it’s not clear what you could do with that data.

Finally you should search for common synonyms of the species. The original compiler of the database did this very thoroughly. However, I suggest just using the names from the 1993 taxonomy if they’ve changed between then and 2005. In theory the original compiler should have picked up all the old papers with data already, so you want to focus on the current taxonomy.

b) Saving the paper and reference
Once you find a paper that looks useful download the paper and save it as a PDF. Also add the reference to EndNote. Not all the papers will end up being included in the database, but it’s quicker to add them to EndNote now and delete them later, than to add them to EndNote later.

\section{search terms}

Google scholars
\textit{order} ("morphology" OR "morphological" OR "cladistic") AND characters matrix paleontology phylogeny
since 2010

\subsection{Mammalian orders terms}
The searched mammalian order terms are available in search_terms_latin.txt or search_terms_meta.txt. The file containing the meta names is the latin name files without the following suffixes: Mammal     @ia
Monothrem     @ata
Marsupial     @ia
Placental     @ia
Macroscelid     @ea
Afrosoricid     @a
Tubulident     @ata
Hyracoid     @ea
Proboscid     @ea
Siren     @ia
Pilos     @a
Cingul     @ata
Scandent     @ia
Dermopter     @a
Primat     @es
Lagomorph     @a
Rodent     @ia
Erinaceomorph     @a
Soricomorph     @a
Cetac     @ea
Artiodactyl     @a
Chiropter     @a
Perissodactyl     @a
Pholidot     @a
Carnivor     @a
Didelphimorph     @ia
Paucitubercul     @ata
Microbiother     @ia
Dasyuromorph     @ia
Peramelemorph     @ia
Notoryctemorph     @ia
Diprotodont     @ia
Monotrem     @ata

-Web of science

-Data bases???
Rose Mounce

-Tree base (through Dryad)
morpholog* OR combine*
mammal*
\textit{order}

-Morphobank
mammal*
\textit{order}

\bibliographystyle{sysbio} %don't write the suffix
\bibliography{References}

\end{document}

### 2014-12-02 Data/select.files.sh
selecting the right matrices

@@ -1,0 +1,77 @@
+#!/bin/sh
+##########################
+#Selecting files in a folder matching a given list of names
+##########################
+#SYNTAX :
+#sh select.files.sh folder list
+#with
+#<folder>
+#<list>
+##########################
+#Selecting all the files containing at least one element of the provided list
+#version: 0.1
+select_files_version="select.files.sh v0.1"
+#Update: The tree shape and the fossil/living distribution is fully randomized.
+#----
+#guillert(at)tcd.ie - 02/12/2014
+##########################
+
+#Input
+folder=$1
+list=$2
+
+#Make the empty folder where to save the results
+mkdir selected_files
+
+#Number of elements in the list
+numbers=$(cat $list | wc -l | sed 's/[[:space:]]//g')
+
+#Initializing the loop
+echo "Searching the $numbers terms:"
+for n in $(seq 1 $numbers)
+do
+    #Selecting the n element of the list
+    element=$(sed -n ''"$n"'p' $list)
+    #Recursive grep within element within folder
+    grep -lr "$element" $folder >> file.to.copy
+    #Verbose
+    printf .
+done
+echo "Done"
+
+#keeping only the unique files and remove the spaces
+sort file.to.copy | uniq -u | sed 's/ /\\ /g' > files.to.copy
+
+#remove the file with extra names
+rm file.to.copy
+#moving the file with names to copy in the folder (and removing the folder name)
+sed 's/'"$folder"'\///g' files.to.copy > $folder/files.to.copy
+#remove the file with extra names
+rm files.to.copy
+
+#Get into the folder
+cd $folder
+
+#Selecting the number of files to copy
+files=$(cat files.to.copy | wc -l | sed 's/[[:space:]]//g')
+
+#Looping through the files to copy
+echo "Copying the $files files:"
+for n in $(seq 1 $files)
+do
+    #Selecting the file to copy
+    file=$(sed -n ''"$n"'p' files.to.copy)
+    #Copying the file in the selected_file folder
+    cp "$file" ../selected_files
+    #verbose
+    printf .
+done
+echo "Done"
+
+#remove the file with names
+rm files.to.copy
+
+#Getting back in the right folder
+cd ..
+
+#End

### 2014-12-02 Functions/select.files.sh

@@ -1,0 +1,67 @@
+#!/bin/sh
+##########################
+#Selecting files in a folder matching a given list of names
+##########################
+#SYNTAX :
+#sh select.files.sh folder list
+#with
+#<folder>
+#<list>
+##########################
+#Selecting all the files containing at least one element of the provided list
+#version: 0.1
+select_files_version="select.files.sh v0.1"
+#Update: The tree shape and the fossil/living distribution is fully randomized.
+#----
+#guillert(at)tcd.ie - 02/12/2014
+##########################
+
+#Input
+folder=$1
+list=$2
+
+#Make the empty folder where to save the results
+mkdir selected_files
+
+#Number of elements in the list
+numbers=$(cat $list | wc -l | sed 's/[[:space:]]//g')
+
+#Initializing the loop
+echo "Searching the $numbers terms:"
+for n in $(seq 1 $numbers)
+do
+    #Selecting the n element of the list
+    element=$(sed -n ''"$n"'p' $list)
+    #Recursive grep within element within folder
+    grep -lr "$element" $folder >> file.to.copy
+    #Verbose
+    printf .
+done
+echo "Done"
+
+#keeping only the unique files and remove the spaces
+sort file.to.copy | uniq -u | sed 's/ /\\ /g' > files.to.copy
+
+#remove the file with extra names
+rm file.to.copy
+
+#Selecting the number of files to copy
+files=$(cat files.to.copy | wc -l | sed 's/[[:space:]]//g')
+
+#Looping through the files to copy
+echo "Copying the $files files:"
+for n in $(seq 1 $files)
+do
+    #Selecting the file to copy
+    file=$(sed -n ''"$n"'p' files.to.copy)
+    #Copying the file in the selected_file folder
+    cp "$file" selected_files
+    #verbose
+    printf .
+done
+echo "Done"
+
+#remove the file with names
+rm files.to.copy
+
+#End

### 2015-02-03 Analysis/Extracting_living_taxa.R
Added saving cvs

@@ -75,3 +75,6 @@ extraction_table<-extraction_table[-1,]
 
 #Save the results as .rda
 save(extraction_table, file="../Data/List_of_matching_taxa.Rda")
+
+#Save the result as csv
+write.csv(extraction_table, file="../Data/List_of_matrching_taxa.csv")

### 2015-02-03 Functions/fix_binomial.sh
Script for fixing binomial names

@@ -1,0 +1,38 @@
+#!/bin/sh
+##########################
+#Fix the bionmial names of matrices in a folder according to a list
+##########################
+#SYNTAX :
+#sh fix_binomial.sh <matrices> <changes>
+#with
+#<matrices> being the path to the folder containing the matrices
+#<changes> being the list of elements to change in the matrices - must be a csv file
+##########################
+#guillert(at)tcd.ie - 03/02/2015
+##########################
+
+#INPUT
+
+#Input values
+matrices=$1
+changes=$2
+
+#number of lines to change
+lines=$(wc -l $changes | sed 's/[[:space:]]//g' | sed 's/'"$changes"'//g')
+
+for line in $(seq 1 $lines)
+do
+    #Splitting the different element of each line
+    element_full=$(sed -n ''"$line"'p' $changes)
+    matrix=$(echo $element_full | sed 's/[,].*$//')
+    original=$(echo $element_full | sed 's/'"$matrix"',//' | sed 's/[,].*$//')
+    replace=$(echo $element_full | sed 's/'"$matrix"','"$original"',//')
+
+    #Replacing the $orignal name by the $replace in $matrix
+    sed 's/'"$original"'/'"$replace"'/' $matrices/${matrix} > $matrices/${matrix}.tmp
+    rm $matrices/${matrix}
+    mv $matrices/${matrix}.tmp $matrices/${matrix}
+    printf .
+done
+
+echo ""

### 2015-02-03 Analysis/Extracting_living_taxa.R
Updated to match with new binomial matrices file

@@ -7,8 +7,12 @@ source("functions.R")
 load.functions(test=FALSE) #Set test=FALSE to speed up the loading
 message("Done\n", appendLF=FALSE)
 
+#Fixing binomial names
+#system("sh ../Functions/fix_binomial.sh ../Data/Matrices_binomial/ ../Data/Matrices_binomial_change.csv")
+
+
 #List of matrices
-matrices_list<-list.files("../Data/Matrices/")
+matrices_list<-list.files("../Data/Matrices_binomial/")
 
 #transform verbose warnings in read.nexus.data in messages
 #Renaming the function (backup)
@@ -51,7 +55,7 @@ for(matrix in 1:length(matrices_list)){
     message(paste(matrices_list[matrix], ":",sep=""), appendLF=FALSE)
     #Try to load the matrix
     current_matrix<-NULL
-    try(current_matrix<-read.nex.dat(paste("../Data/Matrices/", matrices_list[matrix], sep="")), silent=TRUE)
+    try(current_matrix<-read.nex.dat(paste("../Data/Matrices_binomial/", matrices_list[matrix], sep="")), silent=TRUE)
 
     if(!is.null(current_matrix)) {
         #If successfully, proceed:

### 2015-02-03 Manuscript/Missing_mammals-draft.tex
Draft init

--- 2014-12-02 Manuscript/proto-draft.tex
@@
-\documentclass[a4paper,11pt]{article}
+%Authors guidlines: http://royalsocietypublishing.org/instructions-authors#question5
 
+\documentclass[12pt,letterpaper]{article}
+
+%Packages
+\usepackage{pdflscape}
+\usepackage{fixltx2e}
+\usepackage{textcomp}
+\usepackage{fullpage}
 \usepackage{natbib}
+\usepackage{float}
+\usepackage{latexsym}
+\usepackage{url}
+\usepackage{epsfig}
+\usepackage{graphicx}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{bm}
+\usepackage{array}
+\usepackage[version=3]{mhchem}
+\usepackage{ifthen}
+\usepackage{caption}
+\usepackage{hyperref}
+\usepackage{amsthm}
+\usepackage{amstext}
 \usepackage{enumerate}
 \usepackage[osf]{mathpazo}
-\usepackage{lastpage} 
+\usepackage{dcolumn}
+\usepackage{lineno}
 \pagenumbering{arabic}
-\linespread{1.66}
+
+%Pagination style and stuff
+\linespread{2}
+\raggedright
+\setlength{\parindent}{0.5in}
+\setcounter{secnumdepth}{0} 
+\renewcommand{\section}[1]{%
+\bigskip
+\begin{center}
+\begin{Large}
+\normalfont\scshape #1
+\medskip
+\end{Large}
+\end{center}}
+\renewcommand{\subsection}[1]{%
+\bigskip
+\begin{center}
+\begin{large}
+\normalfont\itshape #1
+\end{large}
+\end{center}}
+\renewcommand{\subsubsection}[1]{%
+\vspace{2ex}
+\noindent
+\textit{#1.}---}
+\renewcommand{\tableofcontents}{}
+\bibpunct{(}{)}{;}{a}{}{,}
+
+%---------------------------------------------
+%
+%       START
+%
+%---------------------------------------------
 
@@
 
+%Running head
 \begin{flushright}
@@
 \end{flushright}
+\bigskip
+\noindent RH: Missing morphological data in living mammals
+
+\bigskip
+\medskip
 \begin{center}
 
-%Title
-\noindent{\Large{\bf{Missing living mammals}}}\\
+\noindent{\Large \bf Missing morphological data in living mammals}
+
 \bigskip
-%Author
-\noindent{Thomas Guillerme - guillert@tcd.ie - http://tguillerme.github.io/}\\
 
+\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$$^*$, and Natalie Cooper$^1$$^,$$^2$}\\
+\noindent {\small \it 
+$^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\
+$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.}\\
 \end{center}
+\medskip
+\noindent{*\bf Corresponding author.} \textit{Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.}\\
+\vspace{1in}
+
+%Line numbering
+\modulolinenumbers[1]
+\linenumbers
+
+%---------------------------------------------
+%
+%       ABSTRACT
+%
+%---------------------------------------------
+
+\newpage
+\begin{abstract}
+
+\end{abstract}
+
+\noindent ()\\
+
+\vspace{1.5in}
+
+\newpage
+
+%---------------------------------------------
+%
+%       INTRODUCTION
+%
+%---------------------------------------------
+
+\section{Introduction}
+
+%---------------------------------------------
+%
+%       METHODS
+%
+%---------------------------------------------
+
+\newpage
+
+\section{Material and Methods}
 
@@
 
-Google scholars
-\textit{order} ("morphology" OR "morphological" OR "cladistic") AND characters matrix paleontology phylogeny
-since 2010
-
 \subsection{Mammalian orders terms}
 The searched mammalian order terms are available in search_terms_latin.txt or search_terms_meta.txt.
-The file containing the meta names is the latin name files without the following suffixes: Mammal     @ia
-Monothrem     @ata
-Marsupial     @ia
-Placental     @ia
-Macroscelid     @ea
-Afrosoricid     @a
-Tubulident     @ata
-Hyracoid     @ea
-Proboscid     @ea
-Siren     @ia
-Pilos     @a
-Cingul     @ata
-Scandent     @ia
-Dermopter     @a
-Primat     @es
-Lagomorph     @a
-Rodent     @ia
-Erinaceomorph     @a
-Soricomorph     @a
-Cetac     @ea
-Artiodactyl     @a
-Chiropter     @a
-Perissodactyl     @a
-Pholidot     @a
-Carnivor     @a
-Didelphimorph     @ia
-Paucitubercul     @ata
-Microbiother     @ia
-Dasyuromorph     @ia
-Peramelemorph     @ia
-Notoryctemorph     @ia
-Diprotodont     @ia
-Monotrem     @ata
+The file containing the meta names is the Latin name files but with replacing the latin suffixes ([ia|ata|ea|a]) by a joker character (*) and by replacing the first letter by a upper/lower case meta character (e.g.
+[Aa]).
 
--Web of science
+\subsubsection{Ross Mounce data set}
+I selected all the matrices containing at least one of the mammalian orders names from Ross Mounce GitHub cladistic-data/nexus_files repository (accessed on the 02/12/2014).
 
--Data bases???
-Rose Mounce
+\subsection{Graeme Lloyd}
+Selecting the downloadable matrices
 
--Tree base (through Dryad)
-morpholog* OR combine*
-mammal*
+\subsection{Morphobank}
 \textit{order}
 
--Morphobank
-mammal*
-\textit{order}
+\subsection{Google scholars}
+20 first results since 2010 with the following key words:
+\textit{order} ("morphology" OR "morphological" OR "cladistic") AND characters matrix paleontology phylogeny
 
-\bibliographystyle{sysbio} %don't write the suffix
-\bibliography{References}
+\section{Wrong Bionmial names and typos}
+I fixed the wrong bionomial names format (e.g.
+H.
+sapiens) into the correct ones (e.g.
+Homo sapiens) manually using the PaleoDB and http://www.helsinki.fi/~mhaaramo/metazoa/deuterostoma/chordata/synapsida/eutheria/artiodactyla/anthracotheriidae.html and the Wilson Reeder database.
+I fixed typos and invalid characters in taxa names manually
 
+%---------------------------------------------
+%
+%       RESULTS
+%
+%---------------------------------------------
+
+\section{Results}
+
+%---------------------------------------------
+%
+%       DISCUSSION
+%
+%---------------------------------------------
+
+\section{Discussion}
+
+%Biology letters various stuff
+\section{Ethics statement}
+\section{Data accessibility statement}
+\section{Authors’ contributions statement}
+\section{Acknowledgements}
+\section{Funding statement}
+
+%\bibliographystyle{}
+%\bibliography{}
+
+%Tables
+%Figures
+
+%END
 \end{document}

### 2015-02-04 Analysis/Analysis_per_order.R
Extracting data for one order (from tree or vector)

@@ -1,0 +1,27 @@
+#Loading the functions
+setwd("~/PhD/Projects/Missing_living_mammals/Analysis")
+source("functions.R")
+load.functions(test=FALSE) #Set test=FALSE to speed up the loading
+library(picante)
+
+#Loading the tree
+full_trees<-read.nexus("../Data/Trees/FritzTree.rs200k.100trees.tre")
+set.seed(1) ; one_tree<-full_trees[[sample(1:100, 1)]]
+
+#Loading the taxonomic reference list
+WR_list<-read.csv("../Data/Taxon_References/WilsonReederMSW.csv", header=T, stringsAsFactors=F)
+
+#Loading the list of present taxa
+#source("Extracting_living_taxa.R")
+load("../Data/")
+
+#Sort by number of characters?
+
+#Extract orders
+tree<-extract.order("PRIMATES", one_tree, WR_list)
+taxa<-extract.order("PRIMATES", living_taxa, WR_list)
+#
+
+
+#Calculate the data structure in the order
+clustered_results<-order.structure("Primates (clustered)", clustered_primates, primates_tree, WR_list, verbose=TRUE)

### 2015-02-04 Analysis/shell_launcher.sh
shell script for wrapping analysis

@@ -1,0 +1,2 @@
+sh ../Functions/fix_binomial.sh Matrices_binomial Matrices_binomial_change.csv
+R --no-save < Extracting_living_taxa.R > /dev/null

### 2015-02-04 Functions/extract.order.R
Extracting ordinal data

@@ -1,0 +1,147 @@
+#Extract all the taxa from a given order for a vector or a tree
+##########################
+#SYNTAX:
+#<order> the name of the order
+#<taxa> a vector of species with morphological data or a tree
+#<reference> the list of taxonomical references
+##########################
+#guillert(at)tcd.ie - 21/01/2015
+##########################
+
+extract.order<-function(order, taxa, reference) {
+    
+    #SANTIZING
+    #order
+    check.class(order, "character", " must be a the name of the order to extract.")
+
+    #reference
+    check.class(reference, "data.frame", " must be a taxonomic reference list.")
+    if(length(which(reference == order)) == 0) {
+        stop("Order not found in provided reference list.")
+    }
+
+    #taxa
+    if(class(taxa) == "character") {
+        taxa_is_tree <- FALSE
+    } else {
+        if(class(taxa) == "phylo") {
+            taxa_is_tree <- TRUE
+        } else {
+            stop("taxa must be either a vector of species names or a phylogenetic tree.")
+        }
+    }
+
+    #SELECT ONLY THE TAXA FROM THE GIVEN ORDER
+    Sub_reference<-reference[which(reference == order),]
+
+    #If taxa is a vector
+    if(taxa_is_tree == FALSE) {
+        #Keep only the elements in taxa that do match with the Sub_reference
+        #Empty vector with the taxa to keep
+        taxa_to_keep<-rep(NA, length(taxa))
+        #Empty data frame for eventual taxonomic changes
+        taxa_change<-data.frame("Original"=NA, "Match"=NA)
+
+        for(taxon in 1:length(taxa)) {
+            #Is the taxon in the Sub_reference?
+            test<-which(Sub_reference == taxa[taxon])
+
+            if(length(test)!=0) {
+                #The taxon is in the Sub_refence
+                match_taxon<-TRUE
+
+            } else {
+                #Is it a binomial name?
+                if(grep("_", taxa[taxon])==1) {
+                    #Split the binomial name
+                    binomial<-check.split(taxa[taxon], Sub_reference)
+
+                    #No match
+                    if(all(is.na(binomial))) {
+                        match_taxon<-FALSE
+                    }
+
+                    #Full match
+                    if(all(!is.na(binomial))) {
+                        match_taxon<-TRUE
+                    }
+
+                    #Partial match
+                    if(!is.na(binomial[1]) & is.na(binomial[2])) {
+                        #Only genera name matches
[…]

### 2015-02-04 Functions/extract.names.R
Fixed cascade bug in find.match()

@@ -17,10 +17,11 @@ find.match<-function(matrix_taxon, Ref_list_element) {
     if(length(to_match) == 0) {
         match_found<-FALSE
     } else {
-        #Making sure it matches exactly
-        match_test<-regexec(matrix_taxon, to_match[1], ignore.case=TRUE)
         #Check if the number of matching characters in to_match is the same length as the ones in matrix_taxon (i.e. full match)
-        if(attr(match_test[[1]], "match.length")==nchar(to_match[1])) {
+        match_length<-nchar(matrix_taxon)
+
+        #If any grep match has the same number of characters
+        if(any(nchar(to_match) == match_length)) {
             match_found<-TRUE
         } else {
             match_found<-FALSE
@@ -48,15 +49,21 @@ check.split<-function(matrix_taxon, Ref_list_element) {
 
         #Is the genus name matching?
         genus_match<-find.match(genus, Ref_list_element)
+
         if(genus_match == TRUE) {
             #Find the genus name
-            genus_match<-grep(genus, as.character(unlist(Ref_list_element)), ignore.case=TRUE, value=TRUE)[1]
+            genus_match<-grep(genus, as.character(unlist(Ref_list_element)), ignore.case=TRUE, value=TRUE)
+            #Selecting the first one that has the same number of characters
+            genus_match<-genus_match[which(nchar(genus_match) == nchar(genus))[1]]
 
             #Is the species name matching?
             species_match<-find.match(genus, Ref_list_element)
             if(species_match == TRUE) {
                 #Find the species name
-                species_match<-grep(species, as.character(unlist(Ref_list_element)), ignore.case=TRUE, value=TRUE)[1]
+                species_match<-grep(species, as.character(unlist(Ref_list_element)), ignore.case=TRUE, value=TRUE)
+                #Selecting the first one that has the same number of characters
+                species_match<-species_match[which(nchar(species_match) == nchar(species))[1]]
+                
             } else {
                 species_match<-NA
             }
@@ -211,4 +218,30 @@ extract.names<-function(matrix, references, verbose=FALSE) {
     }
 
     return(output)
+}
+
+#Extra checking function to match the names measured as NAs with Wilson Reeder's list 
+check.NA<-function(NA_taxa, Reference) {
+    #Check if the name contains a "_" or not
+    if(length(grep("_", NA_taxa))!=0) {
[…]

### 2015-02-04 Analysis/Extracting_living_taxa.R
Double checking for NAs

@@ -77,6 +77,39 @@ for(matrix in 1:length(matrices_list)){
 #Removing the first (dummy) line of the extraction_table
 extraction_table<-extraction_table[-1,]
 
+#DOUBLE CHECK THE NAs
+
+#Extract NAs
+NA_species<-extraction_table[which(is.na(extraction_table$Living)),]
+NA_taxa<-NA_species$Taxa
+
+#Checking if the NAs are not in the tree
+doubleCheck_tree<-match(NA_taxa, Reference_list[[1]])
+doubleCheck_tree<-which(!is.na(doubleCheck_tree))
+
+#Checking if the NAs are not in Wilson Reeder's list
+doubleCheck_WR<-vector()
+message("\nChecking:", appendLF=FALSE)
+for(taxa in 1:length(NA_taxa)) {
+    doubleCheck_WR[taxa]<-check.NA(NA_taxa[taxa], Reference_list[[2]])
+    message(".", appendLF=FALSE)
+}
+message("Done.\n", appendLF=FALSE)
+
+
+#If any taxa are not NA, replace them in the table by living==TRUE and tax.level
+#For the Wilson Reeder
+NA_species$Living[which(!is.na(doubleCheck_WR))]<-TRUE
+NA_species$Source[which(!is.na(doubleCheck_WR))]<-"WR"
+NA_species$Tax.level[which(!is.na(doubleCheck_WR))]<-doubleCheck_WR[which(!is.na(doubleCheck_WR))]
+#For the tree
+NA_species$Living[which(!is.na(doubleCheck_tree))]<-TRUE
+NA_species$Source[which(!is.na(doubleCheck_tree))]<-"Fritz"
+NA_species$Tax.level[which(!is.na(doubleCheck_tree))]<-"Species"
+
+#Replace NA_species back in the table
+extraction_table[which(is.na(extraction_table$Living)),]<-NA_species
+
 #Save the results as .rda
 save(extraction_table, file="../Data/List_of_matching_taxa.Rda")
 

### 2015-02-04 Analysis/Draft_table.R
verbose option

@@ -32,7 +32,7 @@ plath_tree<-extract.clade(primates_tree, PlathMRCA)
 clustered_primates<-c(sample(loris_tree$tip.label, 20), sample(plath_tree$tip.label, 60))
 
 
-clustered_results<-order.structure("Primates (clustered)", clustered_primates, primates_tree, WR_list)
+clustered_results<-order.structure("Primates (clustered)", clustered_primates, primates_tree, WR_list, verbose=TRUE)
 
 #Empty results table
 result_table_clustered<-data.frame("Order"=rep("Primates (cluster)",3), "Taxonomic level"=c("Family", "Genus", "Species"), "Number of OTUs"=rep(NA,3), "Percentage of OTUs"=rep(NA,3), "Relative PD"=rep(NA,3), "NRI"=rep(NA,3), "MPD p_value"=rep(NA,3), "NTI"=rep(NA,3), "MNTD p_value"=rep(NA,3))

### 2015-02-04 Functions/extract.names.R
extracting ordinal data

@@ -13,7 +13,7 @@
 #Extracting the first match
 find.match<-function(matrix_taxon, Ref_list_element) {
     options(warn=-1)
-    to_match<-grep(matrix_taxon, as.character(unlist(Ref_list_element)), ignore.case=TRUE, value=TRUE)
+    to_match<-grep(matrix_taxon, as.character(unlist(Ref_list_element)), ignore.case=FALSE, value=TRUE)
     if(length(to_match) == 0) {
         match_found<-FALSE
     } else {

### 2015-02-04 Functions/extract.order.R

@@ -4,11 +4,12 @@
 #<order> the name of the order
 #<taxa> a vector of species with morphological data or a tree
 #<reference> the list of taxonomical references
+#<verbose> whether to be verbose or not
 ##########################
-#guillert(at)tcd.ie - 21/01/2015
+#guillert(at)tcd.ie - 04/02/2015
 ##########################
 
-extract.order<-function(order, taxa, reference) {
+extract.order<-function(order, taxa, reference, verbose=FALSE) {
     
     #SANTIZING
     #order
@@ -31,6 +32,10 @@ extract.order<-function(order, taxa, reference) {
         }
     }
 
+    #verbose
+    check.class(verbose, "logical", " must be logical.")
+
+
     #SELECT ONLY THE TAXA FROM THE GIVEN ORDER
     Sub_reference<-reference[which(reference == order),]
 
@@ -42,6 +47,12 @@ extract.order<-function(order, taxa, reference) {
         #Empty data frame for eventual taxonomic changes
         taxa_change<-data.frame("Original"=NA, "Match"=NA)
 
+        #Verbose
+        if(verbose == TRUE) {
+            message("Scanning the reference list:", appendLF=FALSE)
+        }        
+
+
         for(taxon in 1:length(taxa)) {
             #Is the taxon in the Sub_reference?
             test<-which(Sub_reference == taxa[taxon])
@@ -52,7 +63,7 @@ extract.order<-function(order, taxa, reference) {
 
             } else {
                 #Is it a binomial name?
-                if(grep("_", taxa[taxon])==1) {
+                if(length(grep("_", taxa[taxon]))==1) {
                     #Split the binomial name
                     binomial<-check.split(taxa[taxon], Sub_reference)
 
@@ -70,16 +81,28 @@ extract.order<-function(order, taxa, reference) {
                     if(!is.na(binomial[1]) & is.na(binomial[2])) {
                         #Only genera name matches
                         taxon_change<-c(taxa[taxon], binomial[1])
-                        message(paste(taxa[taxon], " did not match but ", binomial[1], " did.", sep=""))
+                        warning(paste("\n", taxa[taxon], " did not match but ", binomial[1], " did.", sep=""))
                         taxa_change<-rbind(taxa_change, taxon_change)
                         }
 
+                    } else {
+                        #Name is not matching nor binomial
+                        match_taxon<-FALSE
                     }
 
                 }
 
[…]

### 2015-02-04 Analysis/Analysis_per_order.R
Updated functions and analysis for monospecific

@@ -6,6 +6,7 @@ library(picante)
 
 #Loading the tree
 full_trees<-read.nexus("../Data/Trees/FritzTree.rs200k.100trees.tre")
+#Selecting one random tree
 set.seed(1) ; one_tree<-full_trees[[sample(1:100, 1)]]
 
 #Loading the taxonomic reference list
@@ -13,15 +14,70 @@ WR_list<-read.csv("../Data/Taxon_References/WilsonReederMSW.csv", header=T, stri
 
 #Loading the list of present taxa
 #source("Extracting_living_taxa.R")
-load("../Data/")
+load("../Data/List_of_matching_taxa.Rda")
 
-#Sort by number of characters?
+#Selecting the living taxa
+living_taxa<-extraction_table[which(extraction_table$Living == TRUE),]
+living_taxa_list<-unique(living_taxa$Taxa)
 
-#Extract orders
-tree<-extract.order("PRIMATES", one_tree, WR_list)
-taxa<-extract.order("PRIMATES", living_taxa, WR_list)
+#################################
 #
+#   TO DO: Sort by number of characters?
+#
+#################################
 
+
+
+
+
+
+#List of orders
+orders<-unique(WR_list$Order)
+
+
+#Calculate the structure of the data for the first order to initiate the loop
+
+#First order
+#Extract orders
+tree<-extract.order(orders[1], one_tree, WR_list, verbose=TRUE)
+taxa<-extract.order(orders[1], living_taxa_list, WR_list, verbose=TRUE)
+#Replace taxa by a vector if there was any taxonomic correction
+if(class(taxa) != "character") {
+    taxa<-taxa$Taxa
+}
 
 #Calculate the data structure in the order
-clustered_results<-order.structure("Primates (clustered)", clustered_primates, primates_tree, WR_list, verbose=TRUE)
+results<-order.structure(orders[1], taxa, tree, WR_list, verbose=TRUE)
+
+
+
+#Calculate the structure of the data for the other orders through a loop
+
+
+#Loop through the rest
+try(
+for (order in 2:length(orders)) {
+    #verbose
+    message(paste("\n",orders[order], " analysis.\n", sep=""), appendLF=FALSE)
+
+    #Extract orders
+    tree<-extract.order(orders[order], one_tree, WR_list, verbose=TRUE)
+    #If tree is null (monospecific order!) then abort the rest
+    if(is.null(tree)) {
[…]

### 2015-02-04 Functions/data.structure.R

@@ -40,7 +40,7 @@ data.structure<-function(species, tree, plot=FALSE, ...){
     #Plotting (optional)
     if(plot == TRUE) {
         plot(tree, show.tip.label = FALSE, ...)
-        tiplabels(tip = which(primates_tree$tip.label %in% names(which(data_structure[1, ] == 1))), pch = 19, cex = 1)
+        tiplabels(tip = which(tree$tip.label %in% names(which(data_structure[1, ] == 1))), pch = 19, cex = 1)
     }
 
     #Output

### 2015-02-04 Functions/extract.order.R

@@ -39,57 +39,118 @@ extract.order<-function(order, taxa, reference, verbose=FALSE) {
     #SELECT ONLY THE TAXA FROM THE GIVEN ORDER
     Sub_reference<-reference[which(reference == order),]
 
-    #If taxa is a vector
-    if(taxa_is_tree == FALSE) {
-        #Keep only the elements in taxa that do match with the Sub_reference
-        #Empty vector with the taxa to keep
-        taxa_to_keep<-rep(NA, length(taxa))
-        #Empty data frame for eventual taxonomic changes
-        taxa_change<-data.frame("Original"=NA, "Match"=NA)
 
-        #Verbose
-        if(verbose == TRUE) {
-            message("Scanning the reference list:", appendLF=FALSE)
-        }        
+    #If the order is monospecific, abort
+    mono_test<-unique(Sub_reference$Species) ; mono_test<-mono_test[-which(mono_test=="")]
+    if(length(mono_test) < 2) {
+        message("Order is monospecific!")
+        output<-NULL
+    } else {
+
+        #If taxa is a vector
+        if(taxa_is_tree == FALSE) {
+            #Keep only the elements in taxa that do match with the Sub_reference
+            #Empty vector with the taxa to keep
+            taxa_to_keep<-rep(NA, length(taxa))
+            #Empty data frame for eventual taxonomic changes
+            taxa_change<-data.frame("Original"=NA, "Match"=NA)
+
+            #Verbose
+            if(verbose == TRUE) {
+                message("Scanning the reference list:", appendLF=FALSE)
+            }        
 
 
-        for(taxon in 1:length(taxa)) {
-            #Is the taxon in the Sub_reference?
-            test<-which(Sub_reference == taxa[taxon])
+            for(taxon in 1:length(taxa)) {
+                #Is the taxon in the Sub_reference?
+                test<-which(Sub_reference == taxa[taxon])
 
-            if(length(test)!=0) {
-                #The taxon is in the Sub_refence
-                match_taxon<-TRUE
+                if(length(test)!=0) {
+                    #The taxon is in the Sub_refence
+                    match_taxon<-TRUE
 
-            } else {
-                #Is it a binomial name?
-                if(length(grep("_", taxa[taxon]))==1) {
-                    #Split the binomial name
-                    binomial<-check.split(taxa[taxon], Sub_reference)
+                } else {
+                    #Is it a binomial name?
+                    if(length(grep("_", taxa[taxon]))==1) {
[…]

### 2015-02-04 Functions/higher.clade.R

@@ -52,6 +52,11 @@ higher.clade<-function(species, tree, taxonomic.level, reference) {
     #Subset of reference containing only the genus names
     sub_ref<-reference[match(unique_genus, reference$Genus),]
 
+    #Removing NAs
+    if(any(is.na(sub_ref[,1]))) {
+        sub_ref<-sub_ref[-which(is.na(sub_ref[,1])),]
+    }
+    
     #Extract the taxonomic level elements (genus level)
     extracted_genus<-sub_ref[match(unique(sub_ref[,taxonomic_level]), sub_ref[,taxonomic_level]),8]
 
@@ -60,6 +65,11 @@ higher.clade<-function(species, tree, taxonomic.level, reference) {
 
     #list of species to keep (binomial)
     species_to_keep<-unlist(lapply(as.list(extracted_genus), binom.make, tree=tree))
+
+    #Removing NAs
+    if(any(is.na(species_to_keep))) {
+        species_to_keep<-species_to_keep[-which(is.na(species_to_keep))]
+    }
 
     #list of species to remove (binomial)
     species_to_remove<-tree$tip.label[-c(match(species_to_keep, tree$tip.label))]

### 2015-02-04 Functions/order.structure.R

@@ -1,8 +1,8 @@
-#Analysis of the distribution of species per order with morphological data
+#Analysis of the distribution of taxa per order with morphological data
 ##########################
 #SYNTAX:
 #<order> the name of the order
-#<species> a vector of species with morphological data
+#<taxa> a vector of taxa with morphological data
 #<tree> the full tree of the order
 #<reference> the list of taxonomical references
 #<runs> the number of randomization runs
@@ -10,10 +10,10 @@
 #<verbose> whether to be verbose or not (default=FALSE)
 #<...> any optional arguments to be passed to ses.mpd and ses.mntd
 ##########################
-#guillert(at)tcd.ie - 21/01/2015
+#guillert(at)tcd.ie - 04/02/2015
 ##########################
 
-order.structure<-function(order, species, tree, reference, runs=1000, null="taxa.labels", verbose=FALSE, ...) {
+order.structure<-function(order, taxa, tree, reference, runs=1000, null="taxa.labels", verbose=FALSE, ...) {
     #LIBRARIES
     require(picante)
 
@@ -22,9 +22,9 @@ order.structure<-function(order, species, tree, reference, runs=1000, null="taxa
     check.class(order, "character", " must be a single string of characters.")
     check.length(order, 1, " must be a single string of characters.")
 
-    #species
-    check.class(species, "character", " must be a vector of characters.")
-    check.length(species, 1, " must be a vector of characters.", errorif=TRUE)
+    #taxa
+    check.class(taxa, "character", " must be a vector of characters.")
+    check.length(taxa, 1, " must be a vector of characters.", errorif=TRUE)
 
     #tree
     check.class(tree, "phylo", " must be a phylogenetic tree.")
@@ -44,41 +44,112 @@ order.structure<-function(order, species, tree, reference, runs=1000, null="taxa
     #Empty table
     results<-data.frame("Order"=rep(order,3), "Taxonomic level"=c("Family", "Genus", "Species"), "Number of OTUs"=rep(NA,3), "Percentage of OTUs"=rep(NA,3), "Relative PD"=rep(NA,3), "NRI"=rep(NA,3), "MPD p_value"=rep(NA,3), "NTI"=rep(NA,3), "MNTD p_value"=rep(NA,3))
 
-    #Species level
-    data_structure<-data.structure(species, tree)
+    #Sort the list of taxa per species/genera/family
+    #Reference list for the order
[…]

### 2015-02-11 Analysis/Rarefaction.R
Google Scholar rarefaction analysis

@@ -1,0 +1,76 @@
+#Rarefaction analysis
+
+#Loading the functions
+setwd("~/PhD/Projects/Missing_living_mammals/Analysis")
+source("functions.R")
+load.functions(test=FALSE) #Set test=FALSE to speed up the loading
+library(picante)
+
+#Loading the list of present taxa
+#source("Extracting_living_taxa.R")
+load("../Data/List_of_matching_taxa.Rda")
+
+#Fixing the number of characters for truncated matrices
+extraction_table[which(extraction_table$Matrix == "GS2007-R.nex"), 3]<-88
+extraction_table[which(extraction_table$Matrix == "GS2010-VJD.nex"), 3]<-157
+extraction_table[which(extraction_table$Matrix == "GS2012-GM.nex"), 3]<-45
+extraction_table[which(extraction_table$Matrix == "GS2012-SM.nex"), 3]<-
+
+#Creating sub tables with a minimum number of characters
+#1 Character
+extraction_table1<-extraction_table[which(as.numeric(extraction_table$Characters) >= 1),]
+#50 Characters
+extraction_table50<-extraction_table[which(as.numeric(extraction_table$Characters) >= 50),]
+#100 Characters
+extraction_table100<-extraction_table[which(as.numeric(extraction_table$Characters) >= 100),]
+#150 Characters
+extraction_table150<-extraction_table[which(as.numeric(extraction_table$Characters) >= 150),]
+#200 Characters
+extraction_table200<-extraction_table[which(as.numeric(extraction_table$Characters) >= 200),]
+
+#Select the number of characters
+extraction_table<-extraction_table200
+
+#Selecting the living taxa (for a minimal number of characters)
+living_taxa<-extraction_table[which(extraction_table$Living == TRUE),]
+living_taxa_list<-unique(living_taxa$Taxa)
+
+#Selecting the living taxa per data source
+#Morphobank
+morphobank1<-living_taxa[grep("MB", living_taxa$Matrix),]
+morphobank2<-living_taxa[grep("MP", living_taxa$Matrix),]
+morphobank<-rbind(morphobank1, morphobank2)
+
+#Ross Mounce
+mounce<-living_taxa[grep("MM", living_taxa$Matrix),]
+
+#Graeme Lloyd
+lloyd<-living_taxa[grep("GL", living_taxa$Matrix),]
+
+#Google Scholar
+google<-living_taxa[grep("GS", living_taxa$Matrix),]
+
+#Combine the repositories data
[…]

### 2015-02-12 Analysis/Rarefaction.R
Rarefaction analysis

@@ -14,28 +14,12 @@ load("../Data/List_of_matching_taxa.Rda")
 extraction_table[which(extraction_table$Matrix == "GS2007-R.nex"), 3]<-88
 extraction_table[which(extraction_table$Matrix == "GS2010-VJD.nex"), 3]<-157
 extraction_table[which(extraction_table$Matrix == "GS2012-GM.nex"), 3]<-45
-extraction_table[which(extraction_table$Matrix == "GS2012-SM.nex"), 3]<-
+extraction_table[which(extraction_table$Matrix == "GS2012-SM.nex"), 3]<-157
 
-#Creating sub tables with a minimum number of characters
-#1 Character
-extraction_table1<-extraction_table[which(as.numeric(extraction_table$Characters) >= 1),]
-#50 Characters
-extraction_table50<-extraction_table[which(as.numeric(extraction_table$Characters) >= 50),]
-#100 Characters
-extraction_table100<-extraction_table[which(as.numeric(extraction_table$Characters) >= 100),]
-#150 Characters
-extraction_table150<-extraction_table[which(as.numeric(extraction_table$Characters) >= 150),]
-#200 Characters
-extraction_table200<-extraction_table[which(as.numeric(extraction_table$Characters) >= 200),]
-
-#Select the number of characters
-extraction_table<-extraction_table200
-
-#Selecting the living taxa (for a minimal number of characters)
+#Select the living taxa within the repositories sources (morphobank, Graeme Lloyd and Ross Mounce)
+#Living taxa
 living_taxa<-extraction_table[which(extraction_table$Living == TRUE),]
-living_taxa_list<-unique(living_taxa$Taxa)
 
-#Selecting the living taxa per data source
 #Morphobank
 morphobank1<-living_taxa[grep("MB", living_taxa$Matrix),]
 morphobank2<-living_taxa[grep("MP", living_taxa$Matrix),]
@@ -47,30 +31,59 @@ mounce<-living_taxa[grep("MM", living_taxa$Matrix),]
 #Graeme Lloyd
 lloyd<-living_taxa[grep("GL", living_taxa$Matrix),]
 
-#Google Scholar
-google<-living_taxa[grep("GS", living_taxa$Matrix),]
-
 #Combine the repositories data
 repositories<-rbind(morphobank, mounce, lloyd)
+#List of living taxa from the repositories
+repo_taxa<-unique(repositories$Taxa)
 
 
-#Extracting the number of living taxa from the repositories
[…]

### 2015-02-12 Functions/community.structure.R
Added the possibility to choose which metric

@@ -3,6 +3,7 @@
 #SYNTAX:
 #<data_structure> the presence/absence list of OTUs
 #<tree> the full tree
+#<metric> a list of metrics
 #<runs> the number of randomization runs
 #<null> the null model. See ?ses.pd for more details (default = "taxa.labels")
 #<...> any optional arguments to be passed to ses.mpd and ses.mntd
@@ -10,27 +11,22 @@
 #guillert(at)tcd.ie - 20/01/2015
 ##########################
 
-community.structure<-function(data_structure, tree, runs=1000, null="taxa.labels", ...) {
+community.structure<-function(data_structure, tree, metric=c("PD", "NTI", "NRI"), runs=1000, null="taxa.labels", ...) {
 
     library(picante)
 
-    #CALCULATE THE DIFFERENT METRICS
-    #Phylogenetic species variability. psv=1 when species are unrelated (star phylogeny) and psv=0 when species are more related. 
-    #PSV<-psv(data_structure, tree)[1,1]
-
-    #Phylogenetic species richness psr=n when sp are unrelated and psr=0 when are more related
-    #PSR<-psr(data_structure, tree)[1,1]
-
-    #Phylogenetic species evenness. pse=1 when species are evenely spread (good), and 0 when clumped
-    #PSE<-pse(data_structure, tree)[1,1]
-
-    #Phylogenetic species clustering. psc=1 when species are less clustered and 0 when clumped
-    #PSC<-psc(data_structure, tree)[1,1]
+    #SANITIZING
+    #metric
+    check.class(metric, "character", " must be 'PD', 'NTI', 'NRI' or any combination of them.")
+    metrics_avail<-c("PD", "NTI", "NRI")
+    if(any(match(metric, metrics_avail))) {
+        stop("metric must be 'PD', 'NTI', 'NRI' or any combination of them.")
+    }
 
     #Phylogenetic diversity (sum of the total phylogenetic branch length)
-    PD<-pd(data_structure, tree)
+    #PD<-pd(data_structure, tree)
     #Relative PD = sum of branch length in sample / sum of branch length in the tree (when rel_PD=1, all the species are represented in the tree, when rel_PD=0, no species are present)
-    rel_PD<-PD[1,1]/PD[2,1]
[…]

### 2015-02-12 Functions/order.structure.R

@@ -5,6 +5,7 @@
 #<taxa> a vector of taxa with morphological data
 #<tree> the full tree of the order
 #<reference> the list of taxonomical references
+#<metric> a list of metrics
 #<runs> the number of randomization runs
 #<null> the null model. See ?ses.pd for more details (default = "taxa.labels")
 #<verbose> whether to be verbose or not (default=FALSE)
@@ -13,7 +14,7 @@
 #guillert(at)tcd.ie - 04/02/2015
 ##########################
 
-order.structure<-function(order, taxa, tree, reference, runs=1000, null="taxa.labels", verbose=FALSE, ...) {
+order.structure<-function(order, taxa, tree, reference, metric=c("PD", "NTI", "NRI"), runs=1000, null="taxa.labels", verbose=FALSE, ...) {
     #LIBRARIES
     require(picante)
 

### 2015-02-12 Functions/community.structure.R
Fixed the metric sanitising (is.na)

@@ -19,7 +19,7 @@ community.structure<-function(data_structure, tree, metric=c("PD", "NTI", "NRI")
     #metric
     check.class(metric, "character", " must be 'PD', 'NTI', 'NRI' or any combination of them.")
     metrics_avail<-c("PD", "NTI", "NRI")
-    if(any(match(metric, metrics_avail))) {
+    if(any(is.na(match(metric, metrics_avail)))) {
         stop("metric must be 'PD', 'NTI', 'NRI' or any combination of them.")
     }
 

### 2015-02-12 Functions/order.structure.R

@@ -93,7 +93,7 @@ order.structure<-function(order, taxa, tree, reference, metric=c("PD", "NTI", "N
     if(verbose == TRUE) {
         message("Calculating the community structure at species level: ...", appendLF=FALSE)
     }
-    results[3,-c(1,2)]<-community.structure(data_structure, tree, runs, null, ...)
+    results[3,-c(1,2)]<-community.structure(data_structure, tree, metric, runs, null, ...)
     if(verbose == TRUE) {
         message("Done\n", appendLF=FALSE)
     } 
@@ -120,7 +120,7 @@ order.structure<-function(order, taxa, tree, reference, metric=c("PD", "NTI", "N
         if(verbose == TRUE) {
             message("Calculating the community structure at genus level: ...", appendLF=FALSE)
         }    
-        results[2,-c(1,2)]<-community.structure(genus_data_structure, genus_level_tree, runs, null, ...)
+        results[2,-c(1,2)]<-community.structure(genus_data_structure, genus_level_tree, metric, runs, null, ...)
         if(verbose == TRUE) {
             message("Done\n", appendLF=FALSE)
         }
@@ -144,7 +144,7 @@ order.structure<-function(order, taxa, tree, reference, metric=c("PD", "NTI", "N
             message("Calculating the community structure at family level: ...", appendLF=FALSE)
         }    
         
-        results[1,-c(1,2)]<-community.structure(family_data_structure, family_level_tree, runs, null, ...)
+        results[1,-c(1,2)]<-community.structure(family_data_structure, family_level_tree, metric, runs, null, ...)
         
         if(verbose == TRUE) {
             message(" Done.\n", appendLF=FALSE)

### 2015-02-12 Analysis/Analysis_per_order.R
Typo fix

@@ -25,79 +25,76 @@ extraction_table[which(extraction_table$Matrix == "GS2012-SM.nex"), 3]<-157
 #Creating sub tables with a minimum number of characters
 #1 Character
 extraction_table1<-extraction_table[which(as.numeric(extraction_table$Characters) >= 1),]
-#50 Characters
-extraction_table50<-extraction_table[which(as.numeric(extraction_table$Characters) >= 50),]
-#100 Characters
-extraction_table100<-extraction_table[which(as.numeric(extraction_table$Characters) >= 100),]
+#75 Characters
+extraction_table75<-extraction_table[which(as.numeric(extraction_table$Characters) >= 75),]
 #150 Characters
 extraction_table150<-extraction_table[which(as.numeric(extraction_table$Characters) >= 150),]
-#200 Characters
-extraction_table200<-extraction_table[which(as.numeric(extraction_table$Characters) >= 200),]
+#300 Characters
+extraction_table300<-extraction_table[which(as.numeric(extraction_table$Characters) >= 300),]
+#list
+extraction_list=list(extraction_table1,extraction_table75,extraction_table150,extraction_table300)
+results_names=c("results_1.Rda","results_75.Rda","results_150.Rda","results_300.Rda")
 
-#Select the number of characters
-extraction_table<-extraction_table200
+#Print the analysis by series of characters length
 
-#Selecting the living taxa (for a minimal number of characters)
-living_taxa<-extraction_table[which(extraction_table$Living == TRUE),]
-living_taxa_list<-unique(living_taxa$Taxa)
+for (character_threshold in 1:length(extraction_list)) {
 
-#################################
-#
-#   TO DO: Sort by number of characters?
-#
-#################################
+    #Select the number of characters
+    extraction_table<-extraction_list[[character_threshold]]
 
+    #Selecting the living taxa (for a minimal number of characters)
+    living_taxa<-extraction_table[which(extraction_table$Living == TRUE),]
+    living_taxa_list<-unique(living_taxa$Taxa)
 
-
-#List of orders
-orders<-unique(WR_list$Order)
-
-
-#Calculate the structure of the data for the first order to initiate the loop
-
-#First order
-#Extract orders
-tree<-extract.order(orders[1], one_tree, WR_list, verbose=TRUE)
[…]

### 2015-02-12 Functions/community.structure.R

@@ -60,7 +60,7 @@ community.structure<-function(data_structure, tree, metric=c("PD", "NTI", "NRI")
         NRI<-mpd_result$mpd.obs.z[1]*-1
 
         #Significance of difference from NULL
-        MPD_p_value<-mpd_result$mpd.obs.p[1]
+        NRI_p_value<-mpd_result$mpd.obs.p[1]
     }
 
     #SES.NTI
@@ -72,7 +72,7 @@ community.structure<-function(data_structure, tree, metric=c("PD", "NTI", "NRI")
         NTI<-mtnd_result$mntd.obs.z[1]*-1
 
         #Significance of difference from NULL
-        MNTD_p_value<-mtnd_result$mntd.obs.p[1]
+        NTI_p_value<-mtnd_result$mntd.obs.p[1]
     }
 
     #OTU presence / absence

### 2015-02-12 Manuscript/Missing_mammals-draft.tex
methods early draft

@@
 \section{Introduction}
+%Restrict to two paragraphs?
+
+We need trees with both living and fossil taxa blablablablabla
+
+However, because of the nature of TEM matrices, we can have a lot of missing data blabalbalbal
+
+Some missing has more effect on recovering the correct topology than others. Missing living taxa are important! Or they are available (because we have molecular data at least).
+
+In this study we want to have a accurate overlook of the state of morphological availability in mammals.
+
+Also, when data is available, we want to know how the data is distributed along the clade.
+Because it is irrealistic or unnecessary to sample the whole clade, we want the data to be at least randomly distributed (so no bias) or at the best evenly distributed (so every clade is sampled).
+However, what we don't want is that data to be clustered.
+
+After looking at the data availability and it's distribution, we propose several strategies to improve data coverage in weekly sampled orders.
+
+Questions:
+\begin{enumerate}
+\item{-How many taxa with morphological data are available among each living mammals order?}
+\item{-Within each order, how is this data distributed?}
+\item{-How can we improve the data coverage in taxa with non random distributed data}
+\end{enumerate}
 
@@
 \section{Material and Methods}
+\subsection{Matrices search}
+To investigate the available living taxa with morphological data, we downloaded morphological matrices from three main public databases: morphobank, graeamlloyd.com and rossmounce's github.
+We downloaded all the matrices containing any fossil or living mammal taxa from these data bases.
+Additionally we ran a thorough search for matrices that might not have been uploaded on the previously cited data bases through a Google Scholar search.
+We downloaded the eventual additional morphological matrices from any of the 20 first papers matching with our selected key words and with any of the 35 taxonomic levels (see supplementary materials for detailed description of the procedure).
+We downloaded @@@ % Number of matrices
+matrices containing a total of @@@%Number of living OTUs
+operational taxonomic units (OTUs) from the combination of both searches (public repositories and Google Scholar).
 
-\section{Data collection}
-We need to use the same protocol as the GMPD to make sure our data are comparable.
-The original version used lots of online databases but luckily we can now just use Google Scholar!
-Into the Google Scholar search enter the species Latin Binomial as search keywords, followed by the search terms related to parasites.
-So for Marmota monax:
+We then transformed all the matrices to be in the same nexus format.
+We then standardised the taxonomic nomenclature by fixing invalid binomial inputs to match with the official taxonomic nomenclature rules (i.e.
+H.
+sapiens was transformed in Homo sapiens).
+We assigned each species as being either living or fossil using a taxonomic matching algorithm.
+We considered living all the OTUs that where either present in Fritz supper tree or in Wilson Reeders taxonomy.
+We considered fossil all the OTUs that where present in the Paleobiology database.
+For the OTUs neither labelled as living or fossil we tried to decompose the taxa name (i.e.
+Homo_sapiens became Homo and sapiens) and tried to match the to Wilson Reeders taxonomy at any taxonomic level (Genus_species, Genus, Family, etc...).
+The matching taxa where labelled as living and the ones still not matching where ignored and labelled as not applicable (NA).
 
-parasite OR helminth OR virus OR bacteria OR fungi OR ectoparasite OR arthropod OR protozoa "Marmota monax"
+\subsection{Data availability analysis}
+\subsubsection{Data availability}
+Relative proportion of available data per order.
+\subsubsection{Spread of the available data}
+PD/NRI/NTI to null.
+\subsubsection{Data spread improvement strategy}
+Phylotargeting
 
-[Note that the quotation marks (" ") tell Google that you want the exact phrase Marmota monax, and the OR means it will contain Marmota monax and parasite or Marmota monax and helminth etc.]
+All the following procedure is repeatable and available on github.
 
-You then need to search the results for appropriate papers. This is tedious so I’d advise setting yourself a reasonable target, for example, do it for a few hours a day then do something else!
+%---------------------------------------------
+%
+%       RESULTS
+%
+%---------------------------------------------
 
-You should also search for each squirrel genus independently; however I would only do this for genera with few papers on the species.
-If you did this for Marmota spp.
-you could be there forever and it’s not clear what you could do with that data.
+\section{Results}
+Table 1: data availability
 
-Finally you should search for common synonyms of the species.
-The original compiler of the database did this very thoroughly.
-However, I suggest just using the names from the 1993 taxonomy if they’ve changed between then and 2005.
-In theory the original compiler should have picked up all the old papers with data already, so you want to focus on the current taxonomy.
+Table 2: randomness
 
-b) Saving the paper and reference
-Once you find a paper that looks useful download the paper and save it as a PDF.
-Also add the reference to EndNote.
-Not all the papers will end up being included in the database, but it’s quicker to add them to EndNote now and delete them later, than to add them to EndNote later.
+Table 3: phylotargeting
+
+Figure 1: example of good and bad coverage (artiodactyles and carnivores)?
+
+%---------------------------------------------
+%
+%       DISCUSSION
+%
+%---------------------------------------------
+
+\section{Discussion}
+
+How to compare morphological characters?
+
+How to combine these matrices?
+
+%Biology letters various stuff
+\section{Ethics statement}
+\section{Data accessibility statement}
+All data is available and reproducible on github.
+\section{Authors’ contributions statement}
+Conceived and designed the experiments: TG NC. Performed the experiments: TG. Analyzed the data: TG. Wrote the paper: TG NC.
+\section{Acknowledgements}
+April Wright, David Bapst and Graeme Lloyd.
+\section{Funding statement}
+This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).
+
+%\bibliographystyle{}
+%\bibliography{}
+
+%Tables
+%Figures
+
+\section{SOM}
+1- Data collection: key words, clade (ordinal) metacharacters, Google Search terms, Google Search protocol, Google Search rarefaction curve.
 
@@
 \textit{order} ("morphology" OR "morphological" OR "cladistic") AND characters matrix paleontology phylogeny
+Why 20 first results? Because rarefaction curve, check supplemetaries
 
 \section{Wrong Bionmial names and typos}
-I fixed the wrong bionomial names format (e.g.
-H.
-sapiens) into the correct ones (e.g.
-Homo sapiens) manually using the PaleoDB and http://www.helsinki.fi/~mhaaramo/metazoa/deuterostoma/chordata/synapsida/eutheria/artiodactyla/anthracotheriidae.html and the Wilson Reeder database.
-I fixed typos and invalid characters in taxa names manually
-
-%---------------------------------------------
-%
-%       RESULTS
-%
-%---------------------------------------------
-
-\section{Results}
-
-%---------------------------------------------
-%
-%       DISCUSSION
-%
-%---------------------------------------------
-
-\section{Discussion}
-
-%Biology letters various stuff
-\section{Ethics statement}
-\section{Data accessibility statement}
-\section{Authors’ contributions statement}
-\section{Acknowledgements}
-\section{Funding statement}
-
-%\bibliographystyle{}
-%\bibliography{}
-
-%Tables
-%Figures
+I fixed the wrong bionomial names format (e.g. H. sapiens) into the correct ones (e.g. Homo sapiens) manually using the abreviation list in the concerned publications.
 

### 2015-02-12 Analysis/Analysis_per_order.R
Automated analysis

@@ -78,7 +78,9 @@ for (character_threshold in 1:length(extraction_list)) {
         tree<-extract.order(orders[order], one_tree, WR_list, verbose=TRUE)
         #If tree is null (monospecific order!) then abort the rest
         if(is.null(tree)) {
-            results_tmp<-data.frame("Order"=rep(orders[order],3), "Taxonomic level"=c("Species", "Genus", "Family"), "Number of OTUs"=rep("1/1"), "Percentage of OTUs"=rep(1), "Relative PD"=rep(NA,3), "NRI"=rep(NA,3), "MPD p_value"=rep(NA,3), "NTI"=rep(NA,3), "MNTD p_value"=rep(NA,3))
+            results_tmp<-matrix(nrow=3, ncol=ncol(results), data=NA)
+            results_tmp<-as.data.frame(results_tmp)
+            names(results_tmp)<-names(results)
         } else {
             taxa<-extract.order(orders[order], living_taxa_list, WR_list, verbose=TRUE)
             #Replace taxa by a vector if there was any taxonomic correction
@@ -96,5 +98,5 @@ for (character_threshold in 1:length(extraction_list)) {
     }
     ,silent=TRUE)
 
-    write(results, file=results_names[[character_threshold]])
+    write(results, file=results_names[character_threshold])
 }

### 2015-02-13 Analysis/Analysis_per_order.R
Updated table output

@@ -79,6 +79,10 @@ for (character_threshold in 1:length(extraction_list)) {
         #If tree is null (monospecific order!) then abort the rest
         if(is.null(tree)) {
             results_tmp<-matrix(nrow=3, ncol=ncol(results), data=NA)
+            results_tmp[,1]<-rep(orders[order],3)
+            results_tmp[,2]<-c("Family", "Genus", "Species")
+            results_tmp[,3]<-rep("1/1",3)
+            results_tmp[,4]<-rep("100",3)
             results_tmp<-as.data.frame(results_tmp)
             names(results_tmp)<-names(results)
         } else {
@@ -98,5 +102,5 @@ for (character_threshold in 1:length(extraction_list)) {
     }
     ,silent=TRUE)
 
-    write(results, file=results_names[character_threshold])
+    save(results, file=results_names[character_threshold])
 }

### 2015-02-13 Functions/community.structure.R

@@ -11,7 +11,7 @@
 #guillert(at)tcd.ie - 20/01/2015
 ##########################
 
-community.structure<-function(data_structure, tree, metric=c("PD", "NTI", "NRI"), runs=1000, null="taxa.labels", ...) {
+community.structure<-function(data_structure, tree, metric=c("PD", "NRI", "NTI"), runs=1000, null="taxa.labels", ...) {
 
     library(picante)
 

### 2015-02-13 Functions/order.structure.R

@@ -14,7 +14,7 @@
 #guillert(at)tcd.ie - 04/02/2015
 ##########################
 
-order.structure<-function(order, taxa, tree, reference, metric=c("PD", "NTI", "NRI"), runs=1000, null="taxa.labels", verbose=FALSE, ...) {
+order.structure<-function(order, taxa, tree, reference, metric=c("PD", "NRI", "NTI"), runs=1000, null="taxa.labels", verbose=FALSE, ...) {
     #LIBRARIES
     require(picante)
 
@@ -43,7 +43,22 @@ order.structure<-function(order, taxa, tree, reference, metric=c("PD", "NTI", "N
 
     #ANALYSING THE DISTRIBUTION OF SPECIES PER ORDER WITH MORPHOLOGICAL DATA
     #Empty table
-    results<-data.frame("Order"=rep(order,3), "Taxonomic level"=c("Family", "Genus", "Species"), "Number of OTUs"=rep(NA,3), "Percentage of OTUs"=rep(NA,3), "Relative PD"=rep(NA,3), "NRI"=rep(NA,3), "MPD p_value"=rep(NA,3), "NTI"=rep(NA,3), "MNTD p_value"=rep(NA,3))
+    results<-data.frame("Order"=rep(order,3), "Taxonomic level"=c("Family", "Genus", "Species"), "Number of OTUs"=rep(NA,3), "Percentage of OTUs"=rep(NA,3))
+    #Adding two columns per metric
+    if(any(metric == "PD")) {
+        results$PD<-rep(NA,3)
+        results$PD_p<-rep(NA,3)
+    }
+
+    if(any(metric == "NRI")) {
+        results$NRI<-rep(NA,3)
+        results$NRI_p<-rep(NA,3)
+    }
+
+   if(any(metric == "NTI")) {
+        results$NTI<-rep(NA,3)
+        results$NTI_p<-rep(NA,3)
+    }
 
     #Sort the list of taxa per species/genera/family
     #Reference list for the order

### 2015-02-13 Functions/order.structure.R
fixed monospeciefic/genera taxa

@@ -119,7 +119,7 @@ order.structure<-function(order, taxa, tree, reference, metric=c("PD", "NRI", "N
     mono_test<-unique(Sub_reference$Genus) ; mono_test<-mono_test[-which(mono_test=="")]
     if(length(mono_test) < 2) {
         message("Order is monogeneric.")
-        results[2,-c(1,2)]<-rep(NA,7)
+        results[2,-c(1,2)]<-rep(NA,nrow(results))
     } else {
         #Tree
         genus_level_tree<-higher.clade(tree$tip.label, tree, taxonomic.level="Genus", reference=Sub_reference)[[2]]
@@ -147,7 +147,7 @@ order.structure<-function(order, taxa, tree, reference, metric=c("PD", "NRI", "N
     mono_test<-unique(Sub_reference$Family) ; mono_test<-mono_test[-which(mono_test=="")]
     if(length(mono_test) < 2) {
         message("Order is monofamilial.")
-        results[1,-c(1,2)]<-rep(NA,7)
+        results[1,-c(1,2)]<-rep(NA,nrow(results))
     } else {    
 
         #Tree

### 2015-02-22 Analysis/Results_summary.R
Result summary script

@@ -26,9 +26,9 @@ extraction_table[which(extraction_table$Matrix == "GS2012-SM.nex"), 3]<-157
 
 #Creating sub tables with a minimum number of characters
 #1 Character
-extraction_table1<-extraction_table[which(as.numeric(extraction_table$Characters) >= 1),]
+extraction_table1 <- extraction_table[which(as.numeric(extraction_table$Characters) >= 1),]
 #75 Characters
-extraction_table75<-extraction_table[which(as.numeric(extraction_table$Characters) >= 75),]
+extraction_table75<- extraction_table[which(as.numeric(extraction_table$Characters) >= 75),]
 #150 Characters
 extraction_table150<-extraction_table[which(as.numeric(extraction_table$Characters) >= 150),]
 #300 Characters
@@ -73,7 +73,6 @@ for (column in 1:ncol(table_to_print)) {
 }
 #Fixing the column names
 colnames(table_to_print)<-c("Order", "Taxonomic level", "Fraction of OTUs", "Percentage of OTUs")
-
 #Saving the table
 print(xtable(table_to_print), file="../Manuscript/Tables/morpho_taxa_proportion.tex", include.rownames=FALSE, tabular.environment="longtable", floating=FALSE, sanitize.text.function=bold.cells)
 
@@ -99,15 +98,19 @@ table_to_print[,1]<-capwords(table_to_print[,1], strict=TRUE)
 for (column in 1:ncol(table_to_print)) {
     table_to_print[signif_data, column]<-paste('BOLD',table_to_print[signif_data, column], sep="")
 }
-
 #Fixing the column names
 colnames(table_to_print)<-c("Order", "Taxonomic level", "Fraction of OTUs", "Percentage of OTUs", "PD", "p-value")
-
 #Saving the table
 print(xtable(table_to_print), file="../Manuscript/Tables/data_structure.tex", include.rownames=FALSE, tabular.environment="longtable", floating=FALSE, sanitize.text.function=bold.cells)
 
 ############################################
 #Figure (phylogeny example)
 ############################################
+
 #Setting the colour scheme (data, no data)
-plot.results(order="CARNIVORA", col_branch=c("red", "grey"), reference=WR_list, verbose=TRUE)
[…]

### 2015-02-22 Functions/plot.results.R

@@ -1,6 +1,12 @@
-plot.results<-function(order, col_branch, reference, verbose=FALSE) {
+plot.results<-function(order, taxa, col_branch=c("red","grey"), reference, verbose=FALSE) {
+    #Diversitree
+    require(diversitree)
     #Extracting the tree
     order_tree<-extract.order(order, one_tree, reference, verbose)
+    #Stopping if the tree is two small
+    if(Ntip(order_tree)<3) {
+        stop("To few tips in the tree.")
+    }
     #Extracting the taxa with morphological data
     order_taxa<-extract.order(order, living_taxa_list, reference, verbose)
     #Saving only the taxa (discard the change log)
@@ -25,8 +31,8 @@ plot.results<-function(order, col_branch, reference, verbose=FALSE) {
         #Selecting the genera name
         genus<-strsplit(order_tree$tip.label[taxon],"_")[[1]][1]
         #Selecting the corresponding family name in the reference list
-        family_class[[taxon]]<-Sub_reference$Family[grep(genus, Sub_reference$Genus)[1]]
+        family_class[[taxon]]<-Sub_reference$Family[match(genus, Sub_reference$Genus)][1]
     }
     #Plot the phylogeny (radial)
-    trait.plot(order_tree, order_tree$tip.state, cols=list(A=("white")), class=family_class, font=1, cex.lab=1, edge.color=edge_colors)
+    trait.plot(order_tree, order_tree$tip.state, cols=list(A=("white")), class=family_class, font=1, cex.lab=0.7, edge.color=edge_colors)
 }

### 2015-02-23 Analysis/Analysis_per_order.R
Updated pipelined analysis

@@ -30,18 +30,22 @@ extraction_table[which(extraction_table$Matrix == "GS2012-SM.nex"), 3]<-157
 extraction_table1<-extraction_table[which(as.numeric(extraction_table$Characters) >= 1),]
 #75 Characters
 extraction_table75<-extraction_table[which(as.numeric(extraction_table$Characters) >= 75),]
+#94 Characters
+extraction_table100<-extraction_table[which(as.numeric(extraction_table$Characters) >= 100),]
+#138 Characters
+extraction_table125<-extraction_table[which(as.numeric(extraction_table$Characters) >= 125),]
 #150 Characters
 extraction_table150<-extraction_table[which(as.numeric(extraction_table$Characters) >= 150),]
 #300 Characters
 extraction_table300<-extraction_table[which(as.numeric(extraction_table$Characters) >= 300),]
 #list
-extraction_list=list(extraction_table1,extraction_table75,extraction_table150,extraction_table300)
-results_names=c("results_1.Rda","results_75.Rda","results_150.Rda","results_300.Rda")
+extraction_list=list(extraction_table1,extraction_table75, extraction_table100, extraction_table125, extraction_table150,extraction_table300)
+results_names=c("results_1.Rda","results_75.Rda","results_100.Rda","results_125.Rda","results_150.Rda","results_300.Rda")
 
 #Print the analysis by series of characters length
 
-for (character_threshold in 1:length(extraction_list)) {
-
+#for (character_threshold in 1:length(extraction_list)) {
+for (character_threshold in 3:4) {
     #Select the number of characters
     extraction_table<-extraction_list[[character_threshold]]
 

### 2015-02-23 Analysis/Results_summary.R

@@ -11,8 +11,11 @@ full_trees<-read.nexus("../Data/Trees/FritzTree.rs200k.100trees.tre")
 set.seed(1) ; one_tree<-full_trees[[sample(1:100, 1)]]
 
 #Loading the taxonomic reference list
+#Loading the taxonomic reference list
 WR_list<-read.csv("../Data/Taxon_References/WilsonReederMSW.csv", header=T, stringsAsFactors=F)
-Reference<-WR_list
+#Changing ARTIODACTYLA and CETACEA to CETARTIODACTYLA
+WR_list$Order[which(WR_list$Order == "ARTIODACTYLA")]<-"CETARTIODACTYLA"
+WR_list$Order[which(WR_list$Order == "CETACEA")]<-"CETARTIODACTYLA"
 
 #Loading the list of present taxa
 #source("Extracting_living_taxa.R")
@@ -49,7 +52,7 @@ living_taxa_list<-unique(living_taxa$Taxa)
 ######################
 
 #Loading the results
-load("results_1.Rda")
+load("results_150.Rda")
 
 ############################################
 #Table with the number of taxa containing morphological data
@@ -59,7 +62,7 @@ load("results_1.Rda")
 table_to_print<-results[,c(1:4)]
 #Rounding to two digits only
 table_to_print[,4]<-round(as.numeric(table_to_print[,4]), digit=2)
-#Select the rows with less than 25% data
+#Select the rows with less than 25% data = >75% missing data
 low_data<-which(table_to_print[,4]<25)
 #make the whole table as.character
 for (column in 1:ncol(table_to_print)) {
@@ -74,7 +77,7 @@ for (column in 1:ncol(table_to_print)) {
 #Fixing the column names
 colnames(table_to_print)<-c("Order", "Taxonomic level", "Fraction of OTUs", "Percentage of OTUs")
 #Saving the table
-print(xtable(table_to_print), file="../Manuscript/Tables/morpho_taxa_proportion.tex", include.rownames=FALSE, tabular.environment="longtable", floating=FALSE, sanitize.text.function=bold.cells)
+print(xtable(table_to_print), file="../Manuscript/morpho_taxa_proportion.tex", include.rownames=FALSE, tabular.environment="longtable", floating=FALSE, sanitize.text.function=bold.cells)
 
[…]

### 2015-02-23 Functions/plot.results.R

@@ -1,4 +1,4 @@
-plot.results<-function(order, taxa, col_branch=c("red","grey"), reference, verbose=FALSE) {
+plot.results<-function(order, taxa, col_branch=c("red","grey"), reference, verbose=FALSE, ...) {
     #Diversitree
     require(diversitree)
     #Extracting the tree
@@ -34,5 +34,5 @@ plot.results<-function(order, taxa, col_branch=c("red","grey"), reference, verbo
         family_class[[taxon]]<-Sub_reference$Family[match(genus, Sub_reference$Genus)][1]
     }
     #Plot the phylogeny (radial)
-    trait.plot(order_tree, order_tree$tip.state, cols=list(A=("white")), class=family_class, font=1, cex.lab=0.7, edge.color=edge_colors)
+    trait.plot(order_tree, order_tree$tip.state, cols=list(A=("white")), class=family_class, font=1, cex.lab=0.7, edge.color=edge_colors, ...)
 }

### 2015-02-23 Functions/plot.results.R
Added a >=1 filter for morph species

@@ -17,7 +17,17 @@ plot.results<-function(order, taxa, col_branch=c("red","grey"), reference, verbo
     Sub_reference<-subset(reference[which(reference == order),])
     #Making the list of taxa binomial
     order_taxa<-taxa.binomial(order_taxa, Sub_reference)[[1]]
+    #Selecting the non-matching taxa
+    matching_test<-which(is.na(match(order_taxa, order_tree$tip.label)))
+    #If any non matching taxa, remove them from the set
+    if(length(matching_test) >= 1) {
+        order_taxa<-order_taxa[-matching_test]
+    }
     #Selecting the right edges
+    #If order_taxa <=1
+    if(length(order_taxa) <= 1) {
+        stop("Need more than one tip with morphological data.")
+    }
     taxa_edges<-which.edge(order_tree, order_taxa)
     edge_colors<-rep(col_branch[2], nrow(order_tree$edge))
     edge_colors[taxa_edges]<-col_branch[1]

### 2015-02-23 Analysis/Results_summary.R
Summary tables build wrapper

@@ -1,123 +1,0 @@
-#Loading the functions
-setwd("~/PhD/Projects/Missing_living_mammals/Analysis")
-source("functions.R")
-load.functions(test=FALSE) #Set test=FALSE to speed up the loading
-library(picante)
-library(xtable)
-
-#Loading the tree
-full_trees<-read.nexus("../Data/Trees/FritzTree.rs200k.100trees.tre")
-#Selecting one random tree
-set.seed(1) ; one_tree<-full_trees[[sample(1:100, 1)]]
-
-#Loading the taxonomic reference list
-#Loading the taxonomic reference list
-WR_list<-read.csv("../Data/Taxon_References/WilsonReederMSW.csv", header=T, stringsAsFactors=F)
-#Changing ARTIODACTYLA and CETACEA to CETARTIODACTYLA
-WR_list$Order[which(WR_list$Order == "ARTIODACTYLA")]<-"CETARTIODACTYLA"
-WR_list$Order[which(WR_list$Order == "CETACEA")]<-"CETARTIODACTYLA"
-
-#Loading the list of present taxa
-#source("Extracting_living_taxa.R")
-load("../Data/List_of_matching_taxa.Rda")
-
-#Fixing the number of characters for truncated matrices
-extraction_table[which(extraction_table$Matrix == "GS2007-R.nex"), 3]<-88
-extraction_table[which(extraction_table$Matrix == "GS2010-VJD.nex"), 3]<-157
-extraction_table[which(extraction_table$Matrix == "GS2012-GM.nex"), 3]<-45
-extraction_table[which(extraction_table$Matrix == "GS2012-SM.nex"), 3]<-157
-
-#Creating sub tables with a minimum number of characters
-#1 Character
-extraction_table1 <- extraction_table[which(as.numeric(extraction_table$Characters) >= 1),]
-#75 Characters
-extraction_table75<- extraction_table[which(as.numeric(extraction_table$Characters) >= 75),]
-#150 Characters
-extraction_table150<-extraction_table[which(as.numeric(extraction_table$Characters) >= 150),]
-#300 Characters
-extraction_table300<-extraction_table[which(as.numeric(extraction_table$Characters) >= 300),]
-#list
-extraction_list=list(extraction_table1,extraction_table75,extraction_table150,extraction_table300)
-results_names=c("results_1.Rda","results_75.Rda","results_150.Rda","results_300.Rda")
-
-#Select the character threshold
[…]

### 2015-02-23 Functions/summary.results.R

@@ -1,0 +1,61 @@
+summary.results<-function(results, metric, file.save, caption) {
+
+    if(metric == "proportion") {
+        ############################################
+        #Table with the number of taxa containing morphological data
+        ############################################
+
+        #Reformat table (digit + only four first columns)
+        table_to_print<-results[,c(1:4)]
+        #Rounding to two digits only
+        table_to_print[,4]<-round(as.numeric(table_to_print[,4]), digit=2)
+        #Select the rows with less than 25% data = >75% missing data
+        low_data<-which(table_to_print[,4]<25)
+        #make the whole table as.character
+        for (column in 1:ncol(table_to_print)) {
+            table_to_print[,column]<-as.character(table_to_print[,column])
+        }
+        #Lower case
+        table_to_print[,1]<-capwords(table_to_print[,1], strict=TRUE)
+        #Mark the low_data rows to be highlighted in the LaTeX table (BOLD)
+        for (column in 1:ncol(table_to_print)) {
+            table_to_print[low_data,column]<-paste('BOLD',table_to_print[low_data,column], sep="")
+        }
+        #Fixing the column names
+        colnames(table_to_print)<-c("Order", "Taxonomic level", "Fraction of OTUs", "Percentage of OTUs")
+        #Saving the table
+        table<-xtable(table_to_print)
+        caption(table)<-caption
+        print(table, file=file.save, include.rownames=FALSE, tabular.environment="longtable", floating=FALSE, sanitize.text.function=bold.cells, caption.placement="top")
+
+    } else {
+
+        ############################################
+        #Table with the data structure for each taxa
+        ############################################
+
+        #Selecting the right columns
+        metric_col<-grep(metric, names(results))
+        table_to_print<-results[,c(1:4,metric_col)]
+        #Removing the NAs
+        table_to_print<-table_to_print[-which(is.na(table_to_print[,5])),]
+        #Selecting the rows with significant difference from null
+        signif_data<-which(table_to_print[,6]<0.05)
+        #make the whole table as.character
+        for (column in 1:ncol(table_to_print)) {
+            table_to_print[,column]<-as.character(table_to_print[,column])
+        }
+        #Lower case
+        table_to_print[,1]<-capwords(table_to_print[,1], strict=TRUE)
+        #Mark the low_data rows to be highlighted in the LaTeX table (BOLD)
+        for (column in 1:ncol(table_to_print)) {
[…]

### 2015-02-23 Analysis/Results_summary.R
Added supplementaries build

@@ -1,0 +1,138 @@
+#Loading the functions
+setwd("~/PhD/Projects/Missing_living_mammals/Analysis")
+source("functions.R")
+load.functions(test=FALSE) #Set test=FALSE to speed up the loading
+library(picante)
+library(xtable)
+
+#Loading the tree
+full_trees<-read.nexus("../Data/Trees/FritzTree.rs200k.100trees.tre")
+#Selecting one random tree
+set.seed(1) ; one_tree<-full_trees[[sample(1:100, 1)]]
+
+#Loading the taxonomic reference list
+#Loading the taxonomic reference list
+WR_list<-read.csv("../Data/Taxon_References/WilsonReederMSW.csv", header=T, stringsAsFactors=F)
+#Changing ARTIODACTYLA and CETACEA to CETARTIODACTYLA
+WR_list$Order[which(WR_list$Order == "ARTIODACTYLA")]<-"CETARTIODACTYLA"
+WR_list$Order[which(WR_list$Order == "CETACEA")]<-"CETARTIODACTYLA"
+
+#Loading the list of present taxa
+#source("Extracting_living_taxa.R")
+load("../Data/List_of_matching_taxa.Rda")
+
+#Fixing the number of characters for truncated matrices
+extraction_table[which(extraction_table$Matrix == "GS2007-R.nex"), 3]<-88
+extraction_table[which(extraction_table$Matrix == "GS2010-VJD.nex"), 3]<-157
+extraction_table[which(extraction_table$Matrix == "GS2012-GM.nex"), 3]<-45
+extraction_table[which(extraction_table$Matrix == "GS2012-SM.nex"), 3]<-157
+
+#Creating sub tables with a minimum number of characters
+#1 Character
+extraction_table1<-extraction_table[which(as.numeric(extraction_table$Characters) >= 1),]
+#75 Characters
+extraction_table75<-extraction_table[which(as.numeric(extraction_table$Characters) >= 75),]
+#94 Characters
+extraction_table100<-extraction_table[which(as.numeric(extraction_table$Characters) >= 100),]
+#138 Characters
+extraction_table125<-extraction_table[which(as.numeric(extraction_table$Characters) >= 125),]
+#150 Characters
+extraction_table150<-extraction_table[which(as.numeric(extraction_table$Characters) >= 150),]
+#300 Characters
+extraction_table300<-extraction_table[which(as.numeric(extraction_table$Characters) >= 300),]
+#list
[…]

### 2015-02-23 Manuscript/Missing_mammals-draft.tex
Manuscript early draft

@@
 \usepackage{fullpage}
-\usepackage{natbib}
+%\usepackage{natbib}
 \usepackage{float}
@@
 \usepackage{lineno}
+\usepackage{longtable}
 \pagenumbering{arabic}
@@
 \renewcommand{\tableofcontents}{}
-\bibpunct{(}{)}{;}{a}{}{,}
+%\bibpunct{(}{)}{;}{a}{}{,}
 
@@
 %Line numbering
-\modulolinenumbers[1]
-\linenumbers
+%\modulolinenumbers[1]
+%\linenumbers
 
@@
 %Restrict to two paragraphs?
+Studying both living and fossil taxa together in macrovelutionary studies is becoming increasingly common among evolutionary biologists \cite{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
+One trending method, called Total Evidence allows to combine molecular data from living taxa and morphological data from both living and fossil taxa (e.g.
+\cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014,Meseguer01032015}).
+This promising method allows to apply integrative phylogenetic inference methods such as tip-dating \cite{ronquista2012,Drummond01082012,BEASTmaster}.
+However, because the Total Evidence method requires a lot of data (both molecular and morphological for both living and fossil taxa), this method has been shown to be sensible to missing morphological data \cite{GuillermeCooper}.
 
-We need trees with both living and fossil taxa blablablablabla
+In fact, \cite{GuillermeCooper} demonstrates that when few living taxa coded, topological recovery is bad.
+For example, a given phylogeny containing two clades A and B containing only living taxa,
+when using a Total Evidence method, molecular data is available for both clades but morphological data is only available for clade A for any reason.
+Then, the addition of any fossil taxa X related to clade B will lead to a wrong topological placement of this fossil taxa X, branching it somewhere in the clade A instead of the clade B because no morphological data is available to support the placement of the fossil taxa X in clade B.
 
-However, because of the nature of TEM matrices, we can have a lot of missing data blabalbalbal
+%Main rant
+If there is no overlapping characters between a living clade and a fossil one, then it is impossible to branch and fossil taxa to that clade.
+This can be due to evolutionary history (i.e.
+a fossil angiosperm has no overlapping characters with a living mammal) and is expected.
+However, this can also be due to missing data (i.e.
+a fossil primates has no overlapping data with living primates because no data is available for living primates) and will produce artefactual wrong phylogenies (i.e.
+the fossil primate NOT branching in the primate clade).
 
-Some missing has more effect on recovering the correct topology than others. Missing living taxa are important! Or they are available (because we have molecular data at least).
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{MissingDataFigure.pdf}
+\caption{Example of topological errors due to missing morphological data in living taxa.
+If a phylogeny contains two clades, for example Proboscidea and Cetartiodactyla, with molecular data for both but only morphological data for Proboscidea.
+If an additional Cetartiodactyla fossil (with no molecular data) will be added to the phylogeny, it will erroneously branch with the Proboscidea clade instead of the Cetartiodactyla one.}
+\label{Figure_missing_data_problem}
+\end{figure}
 
-In this study we want to have a accurate overlook of the state of morphological availability in mammals.
-
-Also, when data is available, we want to know how the data is distributed along the clade.
-Because it is irrealistic or unnecessary to sample the whole clade, we want the data to be at least randomly distributed (so no bias) or at the best evenly distributed (so every clade is sampled).
-However, what we don't want is that data to be clustered.
-
-After looking at the data availability and it's distribution, we propose several strategies to improve data coverage in weekly sampled orders.
+In this study we investigate the amount of living mammal taxa with available morphological data to assess the potential caveats in building Total Evidence mammal phylogenetic trees.
+As well as the data availability, we calculate the structure of the data to make sure that clades with a relatively high amount of data are not only containing all taxa from a single clade and none from another clade.
 
@@
 We downloaded the eventual additional morphological matrices from any of the 20 first papers matching with our selected key words and with any of the 35 taxonomic levels (see supplementary materials for detailed description of the procedure).
-We downloaded @@@ % Number of matrices
-matrices containing a total of @@@%Number of living OTUs
+We downloaded 256 % @@@ Number of matrices
+matrices containing a total of 9411 % @@@ Number of living OTUs
 operational taxonomic units (OTUs) from the combination of both searches (public repositories and Google Scholar).
@@
 We considered fossil all the OTUs that where present in the Paleobiology database.
-For the OTUs neither labelled as living or fossil we tried to decompose the taxa name (i.e.
-Homo_sapiens became Homo and sapiens) and tried to match the to Wilson Reeders taxonomy at any taxonomic level (Genus_species, Genus, Family, etc...).
-The matching taxa where labelled as living and the ones still not matching where ignored and labelled as not applicable (NA).
+For the OTUs neither labelled as living or fossil we tried to decompose the OTUs name (i.e.
+$Homo\_sapiens$ became $Homo$ and $sapiens$) and tried to match the to Wilson Reeders taxonomy at any taxonomic level (Family, Genus, $Genus\_species$, etc.).
+The matching OTUs where labelled as living and the ones still not matching where ignored and labelled as not applicable (NA).
 
 \subsection{Data availability analysis}
+\subsubsection{Number of characters threshold}
+From all the 256 % @@@ Number of matrices
+matrices, we selected only the ones that had at least 100 morphological characters.
+This arbitrary threshold number of morphological characters was chosen to be in adequacy with \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
+Also this threshold avoids biases towards small matrices that could be either not informative (e.g.
+too few characters) or made of non-applicable characters (e.g.
+antlers which are sexual dimorphic characters proper to a specific clade).
+
 \subsubsection{Data availability}
-Relative proportion of available data per order.
-\subsubsection{Spread of the available data}
-PD/NRI/NTI to null.
-\subsubsection{Data spread improvement strategy}
-Phylotargeting
+To assess the data availability per mammal order, we calculated the percentage of OTUs with morphological data for three different taxonomic level (Family, Genera and Species).
+We highlighted all the orders containing less than 25\% of living taxa with morphological data because their amount of missing data ($>$75\%) was higher than in \cite{GuillermeCooper} and therefore highly probable of suffering from the effect of missing data.
 
-All the following procedure is repeatable and available on github.
+\subsubsection{Available data structure}
+For the order with no morphological data for all OTUs at the three different taxonomic levels (Family, Genera and Species), we investigated the structure of the available data to test if it was either (i) randomly distributed, (ii) over-dispersed or (iii) clustered.
+To measure the structure of the available data we used classic community structure metric from the picante R package \cite{picante} where we compared the structure of the available data for each order to the structure of a potentially fully sampled data (i.e.
+only the OTUs with available morphological data \textit{vs.} all the OTUs).
+For each orders and taxonomic level that presented OTUs with no available morphological data, we calculated the Net Relatedness Index (NRI) which quantifies the overall distribution of the data with negative values showing more dispersed data and positive values more clustered data than expected by the null model \cite{webb2002phylogenies}.
+We choose to present only the NRI values because they have been shown to be slightly less sensitive to the structure of the phylogeny (i.e.
+branch length and topology) \cite{NRI,journal.pone.0004390} but we also calculated the two other common phylogenetic structure indices: Faith's Phylogenetic Distance (PD)\cite{Faith19921} and the Nearest Taxon Index (NTI) \cite{webb2002phylogenies}.
+Both metrics are available in the Supplementary results.
+
+%\subsubsection{Data improvement strategy}
+%Phylotargeting
+%This part as not been done at all!
+I was looking for a non-GUI version of Arnold and Nunn's software but there is none (Arnold confirmed by email).
+Therefore because I was focusing on the part above and to make it repeatable I didn't did that part yet.
+However, I still planing on doing it!
+
+All the following procedure is repeatable and available on GitHub. %Or what is the right way to show it's reproducible?
 
@@
 \section{Results}
-Table 1: data availability
 
-Table 2: randomness
+\subsection{Data availability}
+We extracted 1422 living mammal OTUs from the 256 matrices with a minimum of 6 characters and a maximum of 4541.
+After removing all the matrices with less than 100 the number of extracted living mammals OTUs was down to 815.
+11/28 orders have less than 25\% of taxa with morphological data at a species level and 24/28 orders have less than 75\% taxa with available morphological data.
+At the Genus level however only 3/28 orders have less than 25\% of taxa with morphological data and 16/28 have less than 75\%.
+Finally, at the family level no order has less than 25\% taxa with available morphological data and only 5/28 have less than 75\% ~\ref{Table_morpho_taxa_proportion}.
 
-Table 3: phylotargeting
+\renewcommand\baselinestretch{1.2}\selectfont
+\begin{center}
+%\caption{Proportion of available OTUs with morphological data per order and per taxonomic level.
+We highlighted in bold the orders that have more than 75\% of missing data for each taxonomic level.
+Note that it is possible that more data is available at a higher taxonomic level (Genus $>$ Species) since if the species name for an OTU was not or miss specified, we still counted the OTU for higher taxonomic level analysis.}
+\input{Table_morpho_taxa_proportion}
+\label{Table_morpho_taxa_proportion}
+\end{center}
+\renewcommand\baselinestretch{2}\selectfont
 
-Figure 1: example of good and bad coverage (artiodactyles and carnivores)?
+\subsection{Available data structure}
+Among the orders containing OTUs with no morphological data, only two orders (Carnivora and Chiroptera) are significantly clustered both at the species and the genus level but not at the family level ~\ref{Table_data_structure}.
+\renewcommand\baselinestretch{1.2}\selectfont
+\begin{center}
+%\caption{Data structure for the orders with OTUs without morphological data per taxonomic level.
+The the Net Relatedness Index (NRI) is negative, the OTUs are more dispersed than expected by chance (random); when the NRI is positive, the OTUs are more clustered by expected by chance.
+The p-value indicates the significance in difference from the null model (random).}
+\input{Table_data_structure}
+\label{Table_data_structure}
+\end{center}
+\renewcommand\baselinestretch{2}\selectfont
+Two contrasted results are shown on figure~\ref{Figure_example_coverage} with randomly distributed data in Cetartiodactyla (Fig.~\ref{Figure_example_coverage}A) and clustered available data in Carnivora (mainly Canida;e Fig.~\ref{Figure_example_coverage}B).
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{example_coverage.pdf}
+\caption{Distribution of available morphological data across cetartiodactyles (\textbf{A}) and carnivores (\textbf{B}).
+Edges are colored in grey when no morphological data is available or in red when data is available.}
+\label{Figure_example_coverage}
+\end{figure}
+
+%\subsection{Data improvement strategy}
 
@@
 
-How to compare morphological characters?
+%\subsection{Data availability}
+It's pretty bad
 
-How to combine these matrices?
+%\subsection{Available data structure}
+But good news! It's at least mainly random
+
+%Caveats
+However, we counted only the raw number of characters here and not there similarity. It might be that actually the amount of available data here is lower due to non overlap of characters....
+
+%Let's code some data
+So let's go coding some data and using opensource traceable data bases like morphobank!
 
@@
 \section{Data accessibility statement}
-All data is available and reproducible on github.
+All data is available and reproducible on GitHub.
 \section{Authors’ contributions statement}
@@
 \section{Acknowledgements}
-April Wright, David Bapst and Graeme Lloyd.
+Nick Matzke, April Wright, David Bapst and Graeme Lloyd.
 \section{Funding statement}
@@
 
-%\bibliographystyle{}
-%\bibliography{}
-
-%Tables
-%Figures
+\bibliographystyle{vancouver}
+\bibliography{References}
 
@@
 \subsection{Mammalian orders terms}
-The searched mammalian order terms are available in search_terms_latin.txt or search_terms_meta.txt.
+The searched mammalian order terms are available in $search_terms_latin.txt$ or $search_terms_meta.txt$.
 The file containing the meta names is the Latin name files but with replacing the latin suffixes ([ia|ata|ea|a]) by a joker character (*) and by replacing the first letter by a upper/lower case meta character (e.g.
@@
 \subsubsection{Ross Mounce data set}
-I selected all the matrices containing at least one of the mammalian orders names from Ross Mounce GitHub cladistic-data/nexus_files repository (accessed on the 02/12/2014).
+I selected all the matrices containing at least one of the mammalian orders names from Ross Mounce GitHub $cladistic-data/nexus_files$ repository (accessed on the 02/12/2014).
 
@@
 
+\section{Supplementary results}
+%Tables
+%Figures
+
 %END

### 2015-02-24 Analysis/Rarefaction.R
Results summarizing

@@ -85,5 +85,8 @@ for (mat in 1:nrow(sub_results)) {
 }
 
 Additional_data<-sort(as.numeric(Google_search$Counts), decreasing=TRUE)
-
+pdf("../Manuscript/Supplementary/Supp_figure_google_searches.pdf")
+op<-par(bty="l")
 plot(cumsum(Additional_data), type="l", ylab="Number of additional OTUs", xlab="Google Scholars papers")
+par(op)
+dev.off()

### 2015-02-24 Analysis/Results_summary.R

@@ -62,13 +62,13 @@ load("results_100.Rda")
 #Table with the number of taxa containing morphological data
 ############################################
 caption<-"Proportion of available OTUs with morphological data per order and per taxonomic level. We highlighted in bold the orders that have more than 75\\% of missing data for each taxonomic level. Note that it is possible that more data is available at a higher taxonomic level (Genus $>$ Species) since if the species name for an OTU was not or miss specified, we still counted the OTU for higher taxonomic level analysis."
-summary.results(results, metric="proportion", file.save="../Manuscript/morpho_taxa_proportion.tex", caption=caption)
+summary.results(results, metric="proportion", save.path="../Manuscript/", file.save="Table_morpho_taxa_proportion", caption=caption, environement="longtable")
 
 ############################################
 #Table with the data structure for each taxa
 ############################################
 caption<-"Data structure for the orders with OTUs without morphological data per taxonomic level. When the Net Relatedness Index (NRI) is negative, the OTUs are more dispersed than expected by chance (random); when the NRI is positive, the OTUs are more clustered by expected by chance. The p-value indicates the significance in difference from the null model (random)."
-summary.results(results, metric="NRI", file.save="../Manuscript/data_structure.tex", caption=caption)
+summary.results(results, metric="NRI",save.path="../Manuscript/", file.save="Table_data_structure", caption=caption, environement="longtable")
 
 
 ############################################
@@ -120,19 +120,19 @@ dev.off()
 #Table with the data structure for each taxa
 ############################################
 caption<-"Data structure for the orders with OTUs without morphological data per taxonomic level. When the Nearest Taxon Index (NTI) is negative, the OTUs are more dispersed than expected by chance (random); when the NTI is positive, the OTUs are more clustered by expected by chance. The p-value indicates the significance in difference from the null model (random)."
-summary.results(results, metric="NTI", file.save="../Manuscript/Supplementary/Supp_data_structureNTI.tex", caption=caption)
+summary.results(results, metric="NTI", save.path="../Manuscript/Supplementary/",file.save="Supp_data_structureNTI", caption=caption, environement="longtable")
 caption<-"Data structure for the orders with OTUs without morphological data per taxonomic level. When the Faith's Phylogenetic Distance (PD). The p-value indicates the significance in difference from the null model (random)."
[…]

### 2015-02-24 Functions/summary.results.R

@@ -1,4 +1,4 @@
-summary.results<-function(results, metric, file.save, caption) {
+summary.results<-function(results, metric, save.path, file.save, caption, environement) {
 
     if(metric == "proportion") {
         ############################################
@@ -26,7 +26,8 @@ summary.results<-function(results, metric, file.save, caption) {
         #Saving the table
         table<-xtable(table_to_print)
         caption(table)<-caption
-        print(table, file=file.save, include.rownames=FALSE, tabular.environment="longtable", floating=FALSE, sanitize.text.function=bold.cells, caption.placement="top")
+        label(table)<-file.save
+        print(table, file=paste(save.path,file.save, ".tex",sep=""), include.rownames=FALSE, tabular.environment=environement, floating=FALSE, sanitize.text.function=bold.cells, caption.placement="top")
 
     } else {
 
@@ -56,6 +57,7 @@ summary.results<-function(results, metric, file.save, caption) {
         #Saving the table
         table<-xtable(table_to_print)
         caption(table)<-caption
-        print(table, file=file.save, include.rownames=FALSE, tabular.environment="longtable", floating=FALSE, sanitize.text.function=bold.cells, caption=caption, caption.placement="top")
+        label(table)<-file.save
+        print(table, file=paste(save.path,file.save, ".tex", sep=""), include.rownames=FALSE, tabular.environment=environement, floating=FALSE, sanitize.text.function=bold.cells, caption=caption, caption.placement="top")
     }
 }

### 2015-02-24 Manuscript/Missing_mammals-draft.tex
Manuscript with supplementaries

@@
 \bigskip
-\noindent RH: Missing morphological data in living mammals
+\noindent RH: Morphological data availability in living mammals
 
@@
 
-\noindent{\Large \bf Missing morphological data in living mammals}
+\noindent{\Large \bf Morphological data availability in living mammals}
+%Or: Missing morphological data in living mammals - but I prefer the first one emphasizing on the availability of it. The data is not missing per se, it's just not coded!
 
@@
 \section{Introduction}
-%Restrict to two paragraphs?
 Studying both living and fossil taxa together in macrovelutionary studies is becoming increasingly common among evolutionary biologists \cite{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
@@
 \subsection{Matrices search}
-To investigate the available living taxa with morphological data, we downloaded morphological matrices from three main public databases: morphobank, graeamlloyd.com and rossmounce's github.
+To investigate the available living taxa with morphological data, we downloaded morphological matrices from three main public databases: morphobank (\textttt{http://www.morphobank.org/}), Graeme Lloyd's website (\texttt{http://graemetlloyd.com/}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce/cladistic-data/tree/master/nexus_files}).
 We downloaded all the matrices containing any fossil or living mammal taxa from these data bases.
-Additionally we ran a thorough search for matrices that might not have been uploaded on the previously cited data bases through a Google Scholar search.
+Additionally we ran a thorough Google Scholar search for matrices that might not have been uploaded on the previously cited data bases.
 We downloaded the eventual additional morphological matrices from any of the 20 first papers matching with our selected key words and with any of the 35 taxonomic levels (see supplementary materials for detailed description of the procedure).
@@
 We then standardised the taxonomic nomenclature by fixing invalid binomial inputs to match with the official taxonomic nomenclature rules (i.e.
-H.
-sapiens was transformed in Homo sapiens).
+\textit{H.
+sapiens} was transformed in \textit{Homo sapiens}).
 We assigned each species as being either living or fossil using a taxonomic matching algorithm.
-We considered living all the OTUs that where either present in Fritz supper tree or in Wilson Reeders taxonomy.
-We considered fossil all the OTUs that where present in the Paleobiology database.
+We designated as living OTU all the OTUs that where either present in \cite{FritzTree} or \cite{wilson2005mammal}.
+We designated as fossil OTUs all the OTUs that where present in the Paleobiology database.
 For the OTUs neither labelled as living or fossil we tried to decompose the OTUs name (i.e.
 $Homo\_sapiens$ became $Homo$ and $sapiens$) and tried to match the to Wilson Reeders taxonomy at any taxonomic level (Family, Genus, $Genus\_species$, etc.).
-The matching OTUs where labelled as living and the ones still not matching where ignored and labelled as not applicable (NA).
+The matching OTUs where labelled as living and the ones still not matching where ignored and labelled as non-applicable (NA; see supplementary material for more details on the taxonomic matching algorithm).
 
@@
 matrices, we selected only the ones that had at least 100 morphological characters.
-This arbitrary threshold number of morphological characters was chosen to be in adequacy with \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
+This arbitrary threshold of a minimal number of morphological characters was chosen to be in adequacy with \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
 Also this threshold avoids biases towards small matrices that could be either not informative (e.g.
-too few characters) or made of non-applicable characters (e.g.
+too few characters not allowing character overlap among matrices; see discussion) or made of non-applicable characters (e.g.
 antlers which are sexual dimorphic characters proper to a specific clade).
@@
 \subsubsection{Data availability}
-To assess the data availability per mammal order, we calculated the percentage of OTUs with morphological data for three different taxonomic level (Family, Genera and Species).
+To assess the data availability per mammal order, we calculated the percentage of OTUs with morphological data for three different taxonomic levels: Family, Genera and Species.
 We highlighted all the orders containing less than 25\% of living taxa with morphological data because their amount of missing data ($>$75\%) was higher than in \cite{GuillermeCooper} and therefore highly probable of suffering from the effect of missing data.
@@
 \subsubsection{Available data structure}
-For the order with no morphological data for all OTUs at the three different taxonomic levels (Family, Genera and Species), we investigated the structure of the available data to test if it was either (i) randomly distributed, (ii) over-dispersed or (iii) clustered.
-To measure the structure of the available data we used classic community structure metric from the picante R package \cite{picante} where we compared the structure of the available data for each order to the structure of a potentially fully sampled data (i.e.
-only the OTUs with available morphological data \textit{vs.} all the OTUs).
-For each orders and taxonomic level that presented OTUs with no available morphological data, we calculated the Net Relatedness Index (NRI) which quantifies the overall distribution of the data with negative values showing more dispersed data and positive values more clustered data than expected by the null model \cite{webb2002phylogenies}.
+For the orders with no morphological data for all its OTUs at the three different taxonomic levels (Family, Genera and Species), we investigated the structure of the available data to test if it was either (i) randomly distributed, (ii) over-dispersed or (iii) clustered.
+To measure the structure of the available data we used classic community structure metric from the \texttt{picante} R package \cite{picante}.
+We compared the structure of the available data for each order to the structure of a potentially fully sampled order (i.e.
+only the OTUs with available morphological data \textit{versus} all the OTUs).
+For each orders and taxonomic levels that presented OTUs with no available morphological data, we calculated the Net Relatedness Index (NRI) which quantifies the overall distribution of the data with negative values showing more dispersed data and positive values more clustered data than expected by the null model \cite{webb2002phylogenies}.
 We choose to present only the NRI values because they have been shown to be slightly less sensitive to the structure of the phylogeny (i.e.
-branch length and topology) \cite{NRI,journal.pone.0004390} but we also calculated the two other common phylogenetic structure indices: Faith's Phylogenetic Distance (PD)\cite{Faith19921} and the Nearest Taxon Index (NTI) \cite{webb2002phylogenies}.
+branch length and topology) \cite{NRI,journal.pone.0004390} but we also calculated the two other common phylogenetic structure indices: Faith's Phylogenetic Distance (PD) \cite{Faith19921} and the Nearest Taxon Index (NTI) \cite{webb2002phylogenies}.
 Both metrics are available in the Supplementary results.
@@
 We extracted 1422 living mammal OTUs from the 256 matrices with a minimum of 6 characters and a maximum of 4541.
-After removing all the matrices with less than 100 the number of extracted living mammals OTUs was down to 815.
+After removing all the matrices with less than 100 characters, the number of extracted living mammals OTUs was down to 815.
 11/28 orders have less than 25\% of taxa with morphological data at a species level and 24/28 orders have less than 75\% taxa with available morphological data.
 At the Genus level however only 3/28 orders have less than 25\% of taxa with morphological data and 16/28 have less than 75\%.
-Finally, at the family level no order has less than 25\% taxa with available morphological data and only 5/28 have less than 75\% ~\ref{Table_morpho_taxa_proportion}.
+Finally, at the family level no order has less than 25\% taxa with available morphological data and only 5/28 have less than 75\% (table~\ref{Table_morpho_taxa_proportion}).
 
@@
 \begin{center}
-%\caption{Proportion of available OTUs with morphological data per order and per taxonomic level.
-We highlighted in bold the orders that have more than 75\% of missing data for each taxonomic level.
-Note that it is possible that more data is available at a higher taxonomic level (Genus $>$ Species) since if the species name for an OTU was not or miss specified, we still counted the OTU for higher taxonomic level analysis.}
 \input{Table_morpho_taxa_proportion}
-\label{Table_morpho_taxa_proportion}
 \end{center}
@@
 \subsection{Available data structure}
-Among the orders containing OTUs with no morphological data, only two orders (Carnivora and Chiroptera) are significantly clustered both at the species and the genus level but not at the family level ~\ref{Table_data_structure}.
+Among the orders containing at least one OTU with no available morphological data, only two orders (Carnivora and Chiroptera) are significantly clustered both at the species and the genus level but not at the family level (table~\ref{Table_data_structure}).
 \renewcommand\baselinestretch{1.2}\selectfont
 \begin{center}
-%\caption{Data structure for the orders with OTUs without morphological data per taxonomic level.
-The the Net Relatedness Index (NRI) is negative, the OTUs are more dispersed than expected by chance (random); when the NRI is positive, the OTUs are more clustered by expected by chance.
-The p-value indicates the significance in difference from the null model (random).}
 \input{Table_data_structure}
-\label{Table_data_structure}
 \end{center}
 \renewcommand\baselinestretch{2}\selectfont
-Two contrasted results are shown on figure~\ref{Figure_example_coverage} with randomly distributed data in Cetartiodactyla (Fig.~\ref{Figure_example_coverage}A) and clustered available data in Carnivora (mainly Canida;e Fig.~\ref{Figure_example_coverage}B).
+
+Two contrasted results are shown on figure~\ref{Figure_example_coverage} with randomly distributed data in Cetartiodactyla (Fig.~\ref{Figure_example_coverage}A) and clustered available data in Carnivora (mainly Canida; Fig.~\ref{Figure_example_coverage}B).
 \begin{figure}[!htbp]
@@
 %\subsection{Data availability}
-It's pretty bad
+The amount of living taxa with no available morphological data was surprisingly high at the species level (24/28 orders have less than 75\% taxa with available morphological data) which can lead to topological misplacements of fossils in Total Evidence trees.
+For example in Carnivora, their is no available data for Herpestidae, therefore any fossil Herpestidae will not be able to be correctly placed in the phylogeny (Fig.~\ref{Figure_example_coverage}B).
+On the other hand, any Carnivora fossil with uncertain phylogenetic affinities (\textit{Incertae sedis}) will have a higher probability of branching by chance within the Canidae because they are oversampled within the Carnivora (Fig.~\ref{Figure_example_coverage}B).
 
 %\subsection{Available data structure}
-But good news! It's at least mainly random
+However, it is encouraging that for most orders (except Carnivora and Chiroptera), the few available data is randomly distributed, therefore the addition of fossil taxa is likely to not be biased by oversampling but will probably branch near the taxa that has the most similar available data (as expected).
 
 %Caveats
-However, we counted only the raw number of characters here and not there similarity. It might be that actually the amount of available data here is lower due to non overlap of characters....
+In this study we only used the raw number of characters and not there similarity.
+This might not work in reality since the data may actually not overlap.
+It might be that actually the amount of available data here is lower due to non overlap of characters....
 
@@
 \section{SOM}
-1- Data collection: key words, clade (ordinal) metacharacters, Google Search terms, Google Search protocol, Google Search rarefaction curve.
 
-\section{search terms}
-
-\subsection{Mammalian orders terms}
-The searched mammalian order terms are available in $search_terms_latin.txt$ or $search_terms_meta.txt$.
-The file containing the meta names is the Latin name files but with replacing the latin suffixes ([ia|ata|ea|a]) by a joker character (*) and by replacing the first letter by a upper/lower case meta character (e.g.
-[Aa]).
-
-\subsubsection{Ross Mounce data set}
-I selected all the matrices containing at least one of the mammalian orders names from Ross Mounce GitHub $cladistic-data/nexus_files$ repository (accessed on the 02/12/2014).
-
-\subsection{Graeme Lloyd}
-Selecting the downloadable matrices
-
-\subsection{Morphobank}
-\textit{order}
-
-\subsection{Google scholars}
-20 first results since 2010 with the following key words:
-\textit{order} ("morphology" OR "morphological" OR "cladistic") AND characters matrix paleontology phylogeny
-Why 20 first results? Because rarefaction curve, check supplemetaries
-
-\section{Wrong Bionmial names and typos}
-I fixed the wrong bionomial names format (e.g. H. sapiens) into the correct ones (e.g. Homo sapiens) manually using the abreviation list in the concerned publications.
-
-\section{Supplementary results}
-%Tables
-%Figures
+\input{Supplementary/Supplementary}
 

### 2015-02-24 Manuscript/Supplementary/Supplementary.tex
Manuscript with supplementaries

\noindent{\Large \bf Supplementary Material}

\bigskip

\section{Data collection}
1- Data collection: key words, clade (ordinal) metacharacters, Google Search terms, Google Search protocol, Google Search rarefaction curve.

\subsection{search terms}
The searched mammalian order terms are available in $search_terms_latin.txt$ or $search_terms_meta.txt$. The file containing the meta names is the Latin name files but with replacing the latin suffixes ([ia|ata|ea|a]) by a joker character (*) and by replacing the first letter by a upper/lower case meta character (e.g. [Aa]).
Mammalia; Monotremata; Marsupialia; Placentalia; Macroscelidea; Afrosoricida; Tubulidentata; Hyracoidea; Proboscidea; Sirenia; Pilosa; Cingulata; Scandentia; Dermoptera; Primates; Lagomorpha; Rodentia; Erinaceomorpha; Soricomorpha; Cetacea; Artiodactyla; Cetartiodactyla; Chiroptera; Perissodactyla; Pholidota; Carnivora; Didelphimorphia; Paucituberculata; Microbiotheria; Dasyuromorphia; Peramelemorphia; Notoryctemorphia and Diprotodontia.

\subsubsection{Ross Mounce data set}
I selected all the matrices containing at least one of the mammalian orders names from Ross Mounce GitHub $cladistic-data/nexus_files$ repository (accessed on the 02/12/2014).

\subsubsection{Graeme Lloyd}
Selecting the downloadable matrices

\subsubsection{Morphobank}
\textit{order}

\subsubsection{Google scholars}
20 first results since 2010 with the following exact key words:

\texttt{\textit{order} ("morphology" OR "morphological" OR "cladistic") AND characters matrix paleontology phylogeny}

We selected only the 20 first results per search term because only the 50 first of the 660 papers added 425 living OTUs.

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_google_searches.pdf}
\caption{Google searches additional OTUs rarefaction curve. The x axis represent the number of google scholar matches (papers, books or abstracts) and the y axis represents the cumulative number of additional living OTUs per google scholar match.}
\label{Supp_figure_google_searches}
\end{figure}

\subsection{Wrong Bionmial names and typos}
I fixed the wrong bionomial names format (e.g. H. sapiens) into the correct ones (e.g. Homo sapiens) manually using the abreviation list in the concerned publications. We then applied our taxonomic matching algorithm to classify the OTUs as either living or fossil.

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_Taxonomic_algorithm.pdf}
\caption{Taxonomic matching algorithm used in this study. For each matrix, each operational taxonomic units (OTU) is matched with the super tree from Fritz 2009. If the OTU matches, then it is classified as living. Else it is matched with the Wilson \& Reeders 2005 taxonomy list. If the OTU matches, then it is classified as living. Else it is matched with the Paleo Database list of mammals. If the OTU matches, then it is classified as fossil. Else it is ignored.}
\label{Supp_figure_Taxonomic_algorithm}
\end{figure}

\bibliographystyle{vancouver}
\bibliography{Supp_References}

\section{Data structure}
%Tables
%Figures

\section{Supplementary results}

### 2015-02-25 Analysis/Results_summary.R
Splited figure 2

@@ -76,15 +76,19 @@ summary.results(results, metric="NRI",save.path="../Manuscript/", file.save="Tab
 ############################################
 
 #Setting the colour scheme (data, no data)
-pdf("../Manuscript/example_coverage.pdf", width=16.6, height=8)
+#Separated in two plots for visibility
+#pdf("../Manuscript/example_coverage.pdf", width=16.6, height=8)
+pdf("../Manuscript/example_coverageA.pdf")
 #quartz(width = 16.6, height = 8) #A4 landscape
 #quartz(width = 8.3, height = 5.8) #A5 landscape
-op<-par(mfrow=c(1,2), oma=c(0,1,0,1))
+#op<-par(mfrow=c(1,2), oma=c(0,1,0,1))
 plotA<-plot.results(order="CETARTIODACTYLA", taxa=living_taxa_list, col_branch=c("red", "grey"), reference=WR_list, verbose=TRUE)
-text(x=(plotA$x.lim[1]-plotA$x.lim[1]*0.05),y=(plotA$y.lim[2]-plotA$y.lim[2]*0.05),"A",cex=3)
+dev.off()
+#text(x=(plotA$x.lim[1]-plotA$x.lim[1]*0.05),y=(plotA$y.lim[2]-plotA$y.lim[2]*0.05),"A",cex=3)
+pdf("../Manuscript/example_coverageB.pdf")
 plotB<-plot.results(order="CARNIVORA", taxa=living_taxa_list, col_branch=c("red", "grey"), reference=WR_list, verbose=TRUE)
-text(x=(plotB$x.lim[1]-plotB$x.lim[1]*0.05),y=(plotB$y.lim[2]-plotB$y.lim[2]*0.05),"B",cex=3)
-par(op)
+#text(x=(plotB$x.lim[1]-plotB$x.lim[1]*0.05),y=(plotB$y.lim[2]-plotB$y.lim[2]*0.05),"B",cex=3)
+#par(op)
 dev.off()
 
 #Isolating the order names

### 2015-02-25 Analysis/phylo_targeting.R
Phylo Targeting test

@@ -1,0 +1,103 @@
+#Function for preparing the data for the phylo targeting analysis on http://phylotargeting.fas.harvard.edu/Run_step1.php
+
+#Loading the functions
+setwd("~/PhD/Projects/Missing_living_mammals/Analysis")
+source("functions.R")
+load.functions(test=FALSE) #Set test=FALSE to speed up the loading
+library(picante)
+library(xtable)
+
+#Loading the tree
+full_trees<-read.nexus("../Data/Trees/FritzTree.rs200k.100trees.tre")
+#Selecting one random tree
+set.seed(1) ; one_tree<-full_trees[[sample(1:100, 1)]]
+
+#Loading the taxonomic reference list
+#Loading the taxonomic reference list
+WR_list<-read.csv("../Data/Taxon_References/WilsonReederMSW.csv", header=T, stringsAsFactors=F)
+#Changing ARTIODACTYLA and CETACEA to CETARTIODACTYLA
+WR_list$Order[which(WR_list$Order == "ARTIODACTYLA")]<-"CETARTIODACTYLA"
+WR_list$Order[which(WR_list$Order == "CETACEA")]<-"CETARTIODACTYLA"
+
+#Loading the list of present taxa
+#source("Extracting_living_taxa.R")
+load("../Data/List_of_matching_taxa.Rda")
+
+#Fixing the number of characters for truncated matrices
+extraction_table[which(extraction_table$Matrix == "GS2007-R.nex"), 3]<-88
+extraction_table[which(extraction_table$Matrix == "GS2010-VJD.nex"), 3]<-157
+extraction_table[which(extraction_table$Matrix == "GS2012-GM.nex"), 3]<-45
+extraction_table[which(extraction_table$Matrix == "GS2012-SM.nex"), 3]<-157
+
+#Creating sub tables with a minimum number of characters
+#1 Character
+extraction_table1<-extraction_table[which(as.numeric(extraction_table$Characters) >= 1),]
+#75 Characters
+extraction_table75<-extraction_table[which(as.numeric(extraction_table$Characters) >= 75),]
+#94 Characters
+extraction_table100<-extraction_table[which(as.numeric(extraction_table$Characters) >= 100),]
+#138 Characters
+extraction_table125<-extraction_table[which(as.numeric(extraction_table$Characters) >= 125),]
+#150 Characters
[…]

### 2015-02-25 Manuscript/Missing_mammals-draft.tex
Manuscript update

@@
 
+%Key words: total evidence method, data structure, phylogenetic, living, fossil, topology
 \bigskip
@@
 \section{Introduction}
-Studying both living and fossil taxa together in macrovelutionary studies is becoming increasingly common among evolutionary biologists \cite{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
-One trending method, called Total Evidence allows to combine molecular data from living taxa and morphological data from both living and fossil taxa (e.g.
+%We need fossils - dangerously paraphrased from the missing data paper
+Studying both living and fossil taxa together in essential to fully understand macrovelutionary patterns and processes and is becoming increasingly common among evolutionary biologists \cite{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
+Combining the global biodiversity (living and fossil) in such studies allows for example, to improve accuracy of the timing of diversification events (e.g.
+\cite{ronquista2012}), to better understand relationships among lineages (e.g.
+\cite{beckancient2014}) or to infer biogeographical patterns through time (e.g.
+\cite{Meseguer01032015}).
+%We can use TEM for that
+In order to do so, one must efficiently combine living and fossil taxa data in macroevolutionary models.
+One trending method, called the total evidence method, allows to combine molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix (e.g.
 \cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014,Meseguer01032015}).
-This promising method allows to apply integrative phylogenetic inference methods such as tip-dating \cite{ronquista2012,Drummond01082012,BEASTmaster}.
-However, because the Total Evidence method requires a lot of data (both molecular and morphological for both living and fossil taxa), this method has been shown to be sensible to missing morphological data \cite{GuillermeCooper}.
+This method allows to use all the available data and to treat fossil as tips rather than nodes and allows integrative phylogenetic inference methods such as tip-dating \cite{ronquista2012,Drummond01082012,BEASTmaster}.
 
-In fact, \cite{GuillermeCooper} demonstrates that when few living taxa coded, topological recovery is bad.
-For example, a given phylogeny containing two clades A and B containing only living taxa,
-when using a Total Evidence method, molecular data is available for both clades but morphological data is only available for clade A for any reason.
-Then, the addition of any fossil taxa X related to clade B will lead to a wrong topological placement of this fossil taxa X, branching it somewhere in the clade A instead of the clade B because no morphological data is available to support the placement of the fossil taxa X in clade B.
-
-%Main rant
-If there is no overlapping characters between a living clade and a fossil one, then it is impossible to branch and fossil taxa to that clade.
-This can be due to evolutionary history (i.e.
-a fossil angiosperm has no overlapping characters with a living mammal) and is expected.
-However, this can also be due to missing data (i.e.
-a fossil primates has no overlapping data with living primates because no data is available for living primates) and will produce artefactual wrong phylogenies (i.e.
-the fossil primate NOT branching in the primate clade).
+%But missing data in living taxa problem
+However, the total evidence method requires a lot of data, both molecular and morphological for both living and fossil taxa.
+Entire sections of this data can sometimes be difficult or impossible to collect for every taxa present in the analysis.
+For example, fossil have really rarely molecular data available and morphological characters are rarely collected when molecular data is available (e.g.
+\cite{O'Leary08022013} \textit{vs.} \cite{meredithimpacts2011}).
+Therefore, the ability of this method to recover correct topology is expected to decrease with missing data and especially when few morphological data overlaps between the living and the fossil taxa \cite{GuillermeCooper}.
+%Owwwww the sneaky self cite ;-).
+The effect of missing data is most important on topology when too few living taxa have available morphological data \cite{GuillermeCooper}.
+For example, if there is no morphological available for any living taxa of a clade, it is impossible to link a fossil to this clade because no morphological data will overlap between the fossil (regardless the amount of data available for the fossil) and the living taxa (that have no morphological data).
+This property of the total evidence can lead to important topological incongruities because the fossil will only be able to branch to a clade of living taxa that contains morphological data, even if in reality, the fossil clearly does not belong to this clade (see figure ~\ref{Figure_missing_data_problem}).
 
@@
 
-In this study we investigate the amount of living mammal taxa with available morphological data to assess the potential caveats in building Total Evidence mammal phylogenetic trees.
-As well as the data availability, we calculate the structure of the data to make sure that clades with a relatively high amount of data are not only containing all taxa from a single clade and none from another clade.
+%What we're doing
+It is therefore crucial to understand the distribution of the available morphological data for living taxa in a clade before using a total evidence approach because missing living taxa with morphological data can lead to two topological artefacts: (1) the impossibility for a fossil to branch in the right clade if there is no morphological data available in this clade; and (2) the higher probability for a fossil to branch within a wrong clade that has more morphological data available for living taxa than the right clade.
 
-Questions:
-\begin{enumerate}
-\item{-How many taxa with morphological data are available among each living mammals order?}
-\item{-Within each order, how is this data distributed?}
-\item{-How can we improve the data coverage in taxa with non random distributed data}
-\end{enumerate}
+In this study, we assess the level of available data in living mammals in order to highlight the two caveats described above.
+%In this study we investigate the amount of living mammal taxa with available morphological data to detect the two potential caveats described above that could lead to topological errors when building total evidence mammal phylogenetic trees.
+We extracted living mammals taxa with available morphological data from 256 phylogenetic matrices available online and measured the proportion of morphological data availability for each mammalian order (1).
+Additionally, in the mammalian orders where data was missing, we estimated the structure of the available data to detect if the available data was biases towards certain clades only or if it was randomly distributed (2).
+
+%Finaly using phylogenetic targeting (CITE), we proposed a strategy to fill the gaps.
 
@@
 \subsection{Matrices search}
-To investigate the available living taxa with morphological data, we downloaded morphological matrices from three main public databases: morphobank (\textttt{http://www.morphobank.org/}), Graeme Lloyd's website (\texttt{http://graemetlloyd.com/}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce/cladistic-data/tree/master/nexus_files}).
+To investigate the available living taxa with morphological data, we downloaded morphological matrices from three main public databases: morphobank (\texttt{http://www.morphobank.org/}), Graeme Lloyd's website (\texttt{http://graemetlloyd.com/}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}).
 We downloaded all the matrices containing any fossil or living mammal taxa from these data bases.
@@
 However, I still planing on doing it!
+I just need to find a bit of time to write some code to make it non-GUI (and to be able to use big trees).
 
@@
 
-Two contrasted results are shown on figure~\ref{Figure_example_coverage} with randomly distributed data in Cetartiodactyla (Fig.~\ref{Figure_example_coverage}A) and clustered available data in Carnivora (mainly Canida; Fig.~\ref{Figure_example_coverage}B).
+Two contrasted results are shown on figure ~\ref{Figure_example_coverageA} and ~\ref{Figure_example_coverageB} with randomly distributed data in Cetartiodactyla (figure ~\ref{Figure_example_coverageA}) and clustered available data in Carnivora (mainly Canida; figure ~\ref{Figure_example_coverageB}).
 \begin{figure}[!htbp]
 \centering
-    \includegraphics[width=1\textwidth]{example_coverage.pdf}
-\caption{Distribution of available morphological data across cetartiodactyles (\textbf{A}) and carnivores (\textbf{B}).
-Edges are colored in grey when no morphological data is available or in red when data is available.}
-\label{Figure_example_coverage}
+    \includegraphics[width=1\textwidth]{example_coverageA.pdf}
+\caption{Distribution of available morphological data across Cetartiodactyla. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\label{Figure_example_coverageA}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{example_coverageB.pdf}
+\caption{Distribution of available morphological data across Carnivora. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\label{Figure_example_coverageB}
 \end{figure}
@@
 \section{Discussion}
+%Data collection
+We know we might have missed some data but we performed a thorough and repeatable search and still managed to collected 1422 OTUs which represents 38\% of all living mammals
 
@@
 The amount of living taxa with no available morphological data was surprisingly high at the species level (24/28 orders have less than 75\% taxa with available morphological data) which can lead to topological misplacements of fossils in Total Evidence trees.
-For example in Carnivora, their is no available data for Herpestidae, therefore any fossil Herpestidae will not be able to be correctly placed in the phylogeny (Fig.~\ref{Figure_example_coverage}B).
-On the other hand, any Carnivora fossil with uncertain phylogenetic affinities (\textit{Incertae sedis}) will have a higher probability of branching by chance within the Canidae because they are oversampled within the Carnivora (Fig.~\ref{Figure_example_coverage}B).
+The effect of missing living taxa with morphological data has no significant impact on inferring the correct topology up to 25\% of missing data, beyond that threshold, both the position of "flying" taxa and the clades are significantly different from the correct topology \cite{GuillermeCooper}.
+At the species level within mammals, it is therefore likely to see topological artefacts for the placement of fossil taxa.
+However, the data availability is more promising at higher taxonomic level (i.e.
+the Genus and the Family level).
+This value is probably more important to consider since there is a slight discrepancy between neontological species and paleontological species.
+While the second ones are described based on accurately measured morphology, genetic distance, spacial distribution and even behaviour, the second one are usually estimate based on morphological, spatial and temporal differences only.
+Because of the lack of behavioural and genetic data, it is likely that cryptic species are often ignored in palaeontological data.
+Therefore, most palaeontological studies are using the Genus level for their smallest taxonomic OTU.
+In this frame of palaeontological data usage, the data availablity at the Genus level in living mammals should be our priority concern.
 
@@
 However, it is encouraging that for most orders (except Carnivora and Chiroptera), the few available data is randomly distributed, therefore the addition of fossil taxa is likely to not be biased by oversampling but will probably branch near the taxa that has the most similar available data (as expected).
+When only few data is available, the ideal scenario will be that the data is over-dispersed (i.e.
+that there is data in at least every sub-clade) in order to maximize the possibilities of the fossil to branch in the right clade.
+In the second scenario, when the data is randomly distributed we expected no special bias in the placement of the fossil \cite{GuillermeCooper}.
+However, in the third scenario, when the data is clustered we expect two major biases to occur: (1) first, the fossil will not be able to branch within a clade containing no data (e.g.
+Herpestidae in Carnivora; figure ~\ref{Figure_example_coverageB}), and (2) second, the fossil will have a higher probability, at random, to branch within the clade containing most of the available data (e.g.
+Canidae in Carnivora; figure ~\ref{Figure_example_coverageB}).
+
+%For example in Carnivora, their is no available data for Herpestidae, therefore any fossil Herpestidae will not be able to be correctly placed in the phylogeny (figure ~\ref{Figure_example_coverageB}).
+On the other hand, any Carnivora fossil with uncertain phylogenetic affinities (\textit{Incertae sedis}) will have a higher probability of branching by chance within the Canidae because they are oversampled within the Carnivora (figure ~\ref{Figure_example_coverageB}).
 
@@
 %Let's code some data
-So let's go coding some data and using opensource traceable data bases like morphobank!
+Since the only way to fix missing data issues is to collect the missing data, it queit encouraging to know that the morphological data is available through natural history museum collections and that online plateforms and tools exist and are opperational to facilitate this giant task through collaboration and sharing open source and traceable data!
 
@@
 
-\section{SOM}
+%\section{SOM}
 

### 2015-02-25 Manuscript/Supplementary/Supplementary.tex
Manuscript update

@@
 
-\subsection{search terms}
-The searched mammalian order terms are available in $search_terms_latin.txt$ or $search_terms_meta.txt$.
-The file containing the meta names is the Latin name files but with replacing the latin suffixes ([ia|ata|ea|a]) by a joker character (*) and by replacing the first letter by a upper/lower case meta character (e.g.
-[Aa]).
-Mammalia; Monotremata; Marsupialia; Placentalia; Macroscelidea; Afrosoricida; Tubulidentata; Hyracoidea; Proboscidea; Sirenia; Pilosa; Cingulata; Scandentia; Dermoptera; Primates; Lagomorpha; Rodentia; Erinaceomorpha; Soricomorpha; Cetacea; Artiodactyla; Cetartiodactyla; Chiroptera; Perissodactyla; Pholidota; Carnivora; Didelphimorphia; Paucituberculata; Microbiotheria; Dasyuromorphia; Peramelemorphia; Notoryctemorphia and Diprotodontia.
+\subsection{Public repositories}
+We downloaded available matrices containing fossil and/or living mammal taxa from the three following data bases using the following list of keywords:
 
-\subsubsection{Ross Mounce data set}
-I selected all the matrices containing at least one of the mammalian orders names from Ross Mounce GitHub $cladistic-data/nexus_files$ repository (accessed on the 02/12/2014).
+\texttt{Mammalia; Monotremata; Marsupialia; Placentalia; Macroscelidea; Afrosoricida; Tubulidentata; Hyracoidea; Proboscidea; Sirenia; Pilosa; Cingulata; Scandentia; Dermoptera; Primates; Lagomorpha; Rodentia; Erinaceomorpha; Soricomorpha; Cetacea; Artiodactyla; Cetartiodactyla; Chiroptera; Perissodactyla; Pholidota; Carnivora; Didelphimorphia; Paucituberculata; Microbiotheria; Dasyuromorphia; Peramelemorphia; Notoryctemorphia; Diprotodontia}.
+
+Details about each public repository specific search option is listed below.
+Note that some matrices have been downloaded from more than one database but that it is not an issue since we are interested in the total number of living OTUs and that if some where present in more than one matrix, they still only counted as a unique OTU.
+
+\subsubsection{Morphobank}
+We accessed the Morphobank repository (\texttt{http://www.morphobank.org/}) on the 5th of December 2014 and used the keywords listed above in the search menue.
+We downloaded the data associated with each project matching with the keyword.
 
 \subsubsection{Graeme Lloyd}
-Selecting the downloadable matrices
+We accessed Graeme Lloyd's website repository (\texttt{http://graemetlloyd.com/}) on the 5th of December 2014 and downloaded all the matrices that were available with a direct download link in the mammal data section of the website (\texttt{http://graemetlloyd.com/matrmamm.html}).
 
-\subsubsection{Morphobank}
-\textit{order}
+\subsubsection{Ross Mounce}
+We accessed Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}) on the 2nd of December 2014 and downloaded every 601 matrix.
+We then ran a shell script to select only the matrices that had any text element that match with one of the search terms.
+To make the matrix selection more thorough, we ignored the keywords case as well as the latin suffix (\textit{ia}, \textit{ata}, \textit{ea}, and \textit{a}).
 
-\subsubsection{Google scholars}
-20 first results since 2010 with the following exact key words:
+\subsection{Google scholars}
+To make sure we didn't miss any extra matrix that wasn't available on one of these repository, we ran a Google Scholar search on the 5th of January. We used the following key words:
 
@@
 
-We selected only the 20 first results per search term because only the 50 first of the 660 papers added 425 living OTUs.
+were \textit{order} was replaced by all the keywords listed above.
+For each 33 keywords, we selected the 20 first papers to match the Google search published since 2010 resulting in 660 papers.
+Among these papers, not all contained relevant data (discrete morphological characters AND mammalian data).
+We selected only the 20 first results per search term to avoid downloading articles that were to irrelevant.
+Among the 660 papers, only 50 contained a total of 425 extra living OTUs (figure ~\ref{Supp_figure_google_searches}).
+Also we decided to select only the articles published since 2010 because nearly every one of the recent published matrix contains both a fraction of morphological characters and OTUs from previous studies.
+For example in primates the character \textit{p7} coded first by \cite{ross1998phylogenetic} is reused with the same living species in \cite{seiffert2003fossil}, \cite{marivaux2005anthropoid}, \cite{seiffert2005basal}, \cite{bloch2007new}, \cite{bloch2007new}, \cite{kay2008anatomy}, \cite{silcox2008biogeographic}, \cite{seiffert2009convergent}, \cite{tabuce2009anthropoid}, \cite{boyer2010astragalar}, \cite{seiffert2010fossil}, \cite{marivaux2013djebelemur} and \cite{ni2013oldest}.
 
@@
 
-\subsection{Wrong Bionmial names and typos}
-I fixed the wrong bionomial names format (e.g.
-H.
-sapiens) into the correct ones (e.g.
-Homo sapiens) manually using the abreviation list in the concerned publications.
-We then applied our taxonomic matching algorithm to classify the OTUs as either living or fossil.
+\subsection{Standardising the matrices}
+We transformed all the non-nexus matrices (tnt, word, excel, jpeg) to nexus format manually.
+We then cleaned the nexus matrices by removing any extra information (trees, continuous characters, morphological characters description, molecular data) to end up with nexus matrices containing only the discrete morphological data.
+We then manually fixed the wrong bionomial names format (e.g.
+\textit{H.
+sapiens}) into the correct ones (e.g.
+\textit{Homo sapiens}) using the abbreviation list in the concerned publications.
+
+\subsection{Selecting the living OTUs}
+Finally we applied a taxonomic matching algorithm to classify the OTUs as either living or fossil.
+The algorithm is matching every OTU name from every matrix with one of the following taxonomic references: the list of taxa from the Fritz \textit{et al.} supertree (2009) \cite{FritzTree}; the taxonomic list from the Wilson and Reeder's Mammals Species of the World (2005) \cite{wilson2005mammal} and the list of all the mammal fossil from the Paleobio Database (\texttt{http://paleobiodb.org/cgi-bin/bridge.pl?a=login}) accessed on the 13th of Janurary 2015.
+The OTUs that matched with one of the two first references were considered as living OTUs, the OTUs matching with the third reference were considered as fossil OTUs, finally, the OTUs matching with non of the references were discarded (figure ~\ref{Supp_figure_Taxonomic_algorithm}).
 

### 2015-02-26 Manuscript/Missing_mammals-draft.tex
Manuscript update (wrapped up results)

@@
 The effect of missing data is most important on topology when too few living taxa have available morphological data \cite{GuillermeCooper}.
-For example, if there is no morphological available for any living taxa of a clade, it is impossible to link a fossil to this clade because no morphological data will overlap between the fossil (regardless the amount of data available for the fossil) and the living taxa (that have no morphological data).
-This property of the total evidence can lead to important topological incongruities because the fossil will only be able to branch to a clade of living taxa that contains morphological data, even if in reality, the fossil clearly does not belong to this clade (see figure ~\ref{Figure_missing_data_problem}).
+For example, if there is no morphological available for any living taxa of an entire clade, it is impossible to link a fossil taxon to this clade because no morphological data will overlap between the fossil (regardless the amount of data available for the fossil) and the living taxa (that have no morphological data).
+This property of the total evidence method can lead to important topological incongruities because the fossil will only be able to branch to a clade of living taxa that contains morphological data, even if in reality, the fossil taxon does not belong to that clade (see figure ~\ref{Figure_missing_data_problem}).
 
@@
 \caption{Example of topological errors due to missing morphological data in living taxa.
-If a phylogeny contains two clades, for example Proboscidea and Cetartiodactyla, with molecular data for both but only morphological data for Proboscidea.
-If an additional Cetartiodactyla fossil (with no molecular data) will be added to the phylogeny, it will erroneously branch with the Proboscidea clade instead of the Cetartiodactyla one.}
+If a phylogeny contains two clades, for example Proboscidea and Cetartiodactyla, with molecular data for both (in grey) but only morphological data for Proboscidea only (in red).
+If an additional Cetartiodactyla fossil (with no molecular data) is added to the phylogeny, it will erroneously branch with the Proboscidea clade instead of the Cetartiodactyla one because no morphological data will overlap between the fossil Cetartiodactyla and the living ones.}
 \label{Figure_missing_data_problem}
@@
 %What we're doing
-It is therefore crucial to understand the distribution of the available morphological data for living taxa in a clade before using a total evidence approach because missing living taxa with morphological data can lead to two topological artefacts: (1) the impossibility for a fossil to branch in the right clade if there is no morphological data available in this clade; and (2) the higher probability for a fossil to branch within a wrong clade that has more morphological data available for living taxa than the right clade.
+It is therefore crucial to understand the distribution of the available morphological data for living taxa in a clade before using a total evidence approach because missing living taxa with morphological data can lead to two topological artefacts: (1) the impossibility for a fossil to branch in the right clade if there is no morphological data available in this clade \cite{GuillermeCooper}; and (2) the higher probability for a fossil to branch within a wrong clade that has more morphological data available for living taxa than the right clade \cite{GuillermeCooper}.
 
@@
 %In this study we investigate the amount of living mammal taxa with available morphological data to detect the two potential caveats described above that could lead to topological errors when building total evidence mammal phylogenetic trees.
-We extracted living mammals taxa with available morphological data from 256 phylogenetic matrices available online and measured the proportion of morphological data availability for each mammalian order (1).
-Additionally, in the mammalian orders where data was missing, we estimated the structure of the available data to detect if the available data was biases towards certain clades only or if it was randomly distributed (2).
+We extracted living mammals taxa with available morphological data from 256 phylogenetic matrices available online and measured the proportion of morphological data availability for each mammalian order.
+Additionally, in the mammalian orders where data was missing, we estimated the structure of the available data to detect if the available data was biases towards certain clades only or if it was randomly distributed using the Net Relatedness Index \cite{webb2002phylogenies}.
 
@@
 
-\newpage
+%\newpage
 
@@
 \subsection{Matrices search}
-To investigate the available living taxa with morphological data, we downloaded morphological matrices from three main public databases: morphobank (\texttt{http://www.morphobank.org/}), Graeme Lloyd's website (\texttt{http://graemetlloyd.com/}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}).
-We downloaded all the matrices containing any fossil or living mammal taxa from these data bases.
+To investigate the available living taxa with morphological data, we downloaded morphological matrices from three main public databases: morphobank (\texttt{http://www.morphobank.org/}) \cite{morphobank}, Graeme Lloyd's website (\texttt{http://graemetlloyd.com/}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}).
+We downloaded all the matrices containing any fossil or living mammal taxa from these databases.
 Additionally we ran a thorough Google Scholar search for matrices that might not have been uploaded on the previously cited data bases.
-We downloaded the eventual additional morphological matrices from any of the 20 first papers matching with our selected key words and with any of the 35 taxonomic levels (see supplementary materials for detailed description of the procedure).
+We downloaded the additional morphological characters matrices from the 20 first Google search results matching with our selected keywords and with any of the 35 taxonomic levels (see supplementary materials for detailed description of the procedure).
 We downloaded 256 % @@@ Number of matrices
 matrices containing a total of 9411 % @@@ Number of living OTUs
-operational taxonomic units (OTUs) from the combination of both searches (public repositories and Google Scholar).
+operational taxonomic units (OTUs) from the combination of both searches (public repositories and Google Scholar). The list of matrices is available in the supplementary materials.
 
 We then transformed all the matrices to be in the same nexus format.
-We then standardised the taxonomic nomenclature by fixing invalid binomial inputs to match with the official taxonomic nomenclature rules (i.e.
+We also standardised the taxonomic nomenclature by fixing invalid binomial inputs to match with the official taxonomic nomenclature rules (i.e.
 \textit{H.
@@
 For the OTUs neither labelled as living or fossil we tried to decompose the OTUs name (i.e.
-$Homo\_sapiens$ became $Homo$ and $sapiens$) and tried to match the to Wilson Reeders taxonomy at any taxonomic level (Family, Genus, $Genus\_species$, etc.).
+\textit{Homo$\_$sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with any taxonomic level (Family, Genus, Species, etc.) from \cite{wilson2005mammal}.
 The matching OTUs where labelled as living and the ones still not matching where ignored and labelled as non-applicable (NA; see supplementary material for more details on the taxonomic matching algorithm).
@@
 From all the 256 % @@@ Number of matrices
-matrices, we selected only the ones that had at least 100 morphological characters.
+matrices, we selected only the ones that displayed at least 100 morphological characters per OTU.
 This arbitrary threshold of a minimal number of morphological characters was chosen to be in adequacy with \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
-Also this threshold avoids biases towards small matrices that could be either not informative (e.g.
-too few characters not allowing character overlap among matrices; see discussion) or made of non-applicable characters (e.g.
+Also this threshold avoids biases towards small matrices that could be either not informative \cite{wagner2000} (e.g.
+too few characters not allowing character overlap among matrices) or made of non-applicable characters \cite{Brazeau2011} (e.g.
 antlers which are sexual dimorphic characters proper to a specific clade).
@@
 \subsubsection{Data availability}
-To assess the data availability per mammal order, we calculated the percentage of OTUs with morphological data for three different taxonomic levels: Family, Genera and Species.
-We highlighted all the orders containing less than 25\% of living taxa with morphological data because their amount of missing data ($>$75\%) was higher than in \cite{GuillermeCooper} and therefore highly probable of suffering from the effect of missing data.
+To assess the data availability per mammal order, we calculated the percentage of OTUs with morphological data for three different taxonomic levels: Family, Genus and Species.
+We highlighted all the orders containing less than 25\% of living taxa with morphological data because their amount of missing data ($>$75\%) was higher than in \cite{GuillermeCooper} and therefore highly probable of having wrong topology due to the missing data.
+On the opposite, orders with up to 25\% of missing data (75\% of available data) were shown to have no significant effect on topology \cite{GuillermeCooper}.
+We therefore used this value as a threshold of "good" data sampling.
 
 \subsubsection{Available data structure}
-For the orders with no morphological data for all its OTUs at the three different taxonomic levels (Family, Genera and Species), we investigated the structure of the available data to test if it was either (i) randomly distributed, (ii) over-dispersed or (iii) clustered.
+For the orders with no morphological data for all its OTUs at the three different taxonomic levels (Family, Genus and Species), we investigated the structure of the available data to test if it was either (i) randomly distributed, (ii) over-dispersed or (iii) clustered.
 To measure the structure of the available data we used classic community structure metric from the \texttt{picante} R package \cite{picante}.
@@
 only the OTUs with available morphological data \textit{versus} all the OTUs).
-For each orders and taxonomic levels that presented OTUs with no available morphological data, we calculated the Net Relatedness Index (NRI) which quantifies the overall distribution of the data with negative values showing more dispersed data and positive values more clustered data than expected by the null model \cite{webb2002phylogenies}.
+For each orders and taxonomic levels that presented OTUs with no available morphological data, we calculated the Net Relatedness Index (NRI) which quantifies the overall distribution of the data with negative values showing more dispersed data and positive values more clustered data than expected by the null model (random) \cite{webb2002phylogenies}.
 We choose to present only the NRI values because they have been shown to be slightly less sensitive to the structure of the phylogeny (i.e.
@@
 \subsection{Data availability}
-We extracted 1422 living mammal OTUs from the 256 matrices with a minimum of 6 characters and a maximum of 4541.
-After removing all the matrices with less than 100 characters, the number of extracted living mammals OTUs was down to 815.
+We extracted 1422 living mammal OTUs from the 256 matrices with a minimum of 6 characters and a maximum of 4541 per OTU.
+After removing all the matrices with less than 100 morphological characters, the number of extracted living mammals OTUs was down to 815.
 11/28 orders have less than 25\% of taxa with morphological data at a species level and 24/28 orders have less than 75\% taxa with available morphological data.
@@
 \subsection{Available data structure}
-Among the orders containing at least one OTU with no available morphological data, only two orders (Carnivora and Chiroptera) are significantly clustered both at the species and the genus level but not at the family level (table~\ref{Table_data_structure}).
+Among the orders containing at least one OTU with no available morphological data, only four orders are significantly clustered: Dirpotodontia, Carnivora and Chiroptera both at the species and the genus level and Rodentia at the species level only (table~\ref{Table_data_structure}).
+
 \renewcommand\baselinestretch{1.2}\selectfont
@@
 
-Two contrasted results are shown on figure ~\ref{Figure_example_coverageA} and ~\ref{Figure_example_coverageB} with randomly distributed data in Cetartiodactyla (figure ~\ref{Figure_example_coverageA}) and clustered available data in Carnivora (mainly Canida; figure ~\ref{Figure_example_coverageB}).
+Two contrasted results are shown on figure ~\ref{Figure_example_coverageA} and ~\ref{Figure_example_coverageB} with randomly distributed data in Cetartiodactyla (figure ~\ref{Figure_example_coverageA}) and clustered available data in Carnivora (mainly Canidae; figure ~\ref{Figure_example_coverageB}).
+
 \begin{figure}[!htbp]
@@
 \section{Discussion}
+Our results show that even though relations among living mammals are well known in a molecular framework (e.g.
+\cite{FritzTree,meredithimpacts2011,May-Collado-PeerJ}) only little morphological data is available.
+This is primarily due to the fact that that morphological studies are mainly focusing on fossil taxa rather than living taxa (e.g.
+\cite{O'Leary08022013,ni2013oldest}).
+However, data availability varies in its structure as well as across taxonomic levels.
+Higher taxonomic levels are always better sampled than lower ones (table~\ref{Table_morpho_taxa_proportion}) and within these taxonomic levels, the structure of available data is mostly random apart for four orders (table\ref{Table_data_structure}).
+
 %Data collection
 We know we might have missed some data but we performed a thorough and repeatable search and still managed to collected 1422 OTUs which represents 38\% of all living mammals
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{../andrew/citations.jpg}
+\caption{ }
+\label{bla}
+\end{figure}
 
@@
 %\section{SOM}
-
+\newcommand{\beginsupplement}{%
+    \setcounter{table}{0}
+    \renewcommand{\thetable}{S\arabic{table}}%
+    \setcounter{figure}{0}
+    \renewcommand{\thefigure}{S\arabic{figure}}%
+}
+\beginsupplement
 \input{Supplementary/Supplementary}

### 2015-02-26 Manuscript/Supplementary/Supplementary.tex
Manuscript update (wrapped up results)

@@
 
-\bibliographystyle{vancouver}
-\bibliography{Supp_References}
-
-\section{Data structure}
-%Tables
-%Figures
+%\bibliographystyle{vancouver}
+%\bibliography{Supp_References}
 
 \section{Supplementary results}
+The following section contains supplementary results to the main body: the available data structure using the NTI and the PD metric; the proportion of available data and the data structure for all the matrices (including the matrices with less than 100 characters); and phylogenetical representation of the data availability per order (excluding Cetartiodactyla and Carnivora, present in the main body).
+
+\renewcommand\baselinestretch{1.2}\selectfont
+\begin{center}
+\input{Supplementary/Supp_data_structureNTI}
+\end{center}
+\renewcommand\baselinestretch{2}\selectfont
+
+\renewcommand\baselinestretch{1.2}\selectfont
+\begin{center}
+\input{Supplementary/Supp_data_structurePD}
+\end{center}
+\renewcommand\baselinestretch{2}\selectfont
+
+\renewcommand\baselinestretch{1.2}\selectfont
+\begin{center}
+\input{Supplementary/morpho_taxa_proportion_threshold1}
+\end{center}
+\renewcommand\baselinestretch{2}\selectfont
+
+\renewcommand\baselinestretch{1.2}\selectfont
+\begin{center}
+\input{Supplementary/Supp_data_structureNRI_threshold1}
+\end{center}
+\renewcommand\baselinestretch{2}\selectfont
+
+\renewcommand\baselinestretch{1.2}\selectfont
+\begin{center}
+\input{Supplementary/Supp_data_structureNTI_threshold1}
+\end{center}
+\renewcommand\baselinestretch{2}\selectfont
+
+\renewcommand\baselinestretch{1.2}\selectfont
+\begin{center}
+\input{Supplementary/Supp_data_structurePD_threshold1}
+\end{center}
+\renewcommand\baselinestretch{2}\selectfont
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_AFROSORICIDA.pdf}
+\caption{Distribution of available morphological data across Afrosoricida. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\label{Supp_Figure_Phylo-Afrosoricida}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_CHIROPTERA.pdf}
+\caption{Distribution of available morphological data across Chiroptera. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\label{Supp_Figure_Phylo-Chiroptera}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_CINGULATA.pdf}
+\caption{Distribution of available morphological data across Cingulata. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\label{Supp_Figure_Phylo-Cingulata}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_DASYUROMORPHIA.pdf}
+\caption{Distribution of available morphological data across Dasyuromorphia. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\label{Supp_Figure_Phylo-Dasyuromorphia}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_DIDELPHIMORPHIA.pdf}
+\caption{Distribution of available morphological data across Didelphimorphia. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\label{Supp_Figure_Phylo-Didelphimorphia}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_DIPROTODONTIA.pdf}
+\caption{Distribution of available morphological data across Diprotodontia. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\label{Supp_Figure_Phylo-Diprotodontia}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_ERINACEOMORPHA.pdf}
+\caption{Distribution of available morphological data across Erinaceomorpha. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\label{Supp_Figure_Phylo-Erinaceomorpha}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_PILOSA.pdf}
+\caption{Distribution of available morphological data across Pilosa. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\label{Supp_Figure_Phylo-Pilosa}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_PRIMATES.pdf}
+\caption{Distribution of available morphological data across Primates. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\label{Supp_Figure_Phylo-Primates}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_RODENTIA.pdf}
+\caption{Distribution of available morphological data across Rodentia. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\label{Supp_Figure_Phylo-Rodentia}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_SCANDENTIA.pdf}
+\caption{Distribution of available morphological data across Scandentia. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\label{Supp_Figure_Phylo-Scandentia}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_SORICOMORPHA.pdf}
+\caption{Distribution of available morphological data across Soricomorpha. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\label{Supp_Figure_Phylo-Soricomorpha}
+\end{figure}

### 2015-03-02 Manuscript/Missing_mammals-draft.tex
Full draft manuscript

@@
 At the Genus level however only 3/28 orders have less than 25\% of taxa with morphological data and 16/28 have less than 75\%.
-Finally, at the family level no order has less than 25\% taxa with available morphological data and only 5/28 have less than 75\% (table~\ref{Table_morpho_taxa_proportion}).
+Finally, at the family level no order has less than 25\% taxa with available morphological data and only 5/28 have less than 75\% (table \ref{Table_morpho_taxa_proportion}).
 
@@
 \subsection{Available data structure}
-Among the orders containing at least one OTU with no available morphological data, only four orders are significantly clustered: Dirpotodontia, Carnivora and Chiroptera both at the species and the genus level and Rodentia at the species level only (table~\ref{Table_data_structure}).
+Among the orders containing at least one OTU with no available morphological data, only two orders are significantly clustered: Carnivora and Chiroptera both at the species and the genus level (table~\ref{Table_data_structure}).
 
@@
 
-Two contrasted results are shown on figure ~\ref{Figure_example_coverageA} and ~\ref{Figure_example_coverageB} with randomly distributed data in Cetartiodactyla (figure ~\ref{Figure_example_coverageA}) and clustered available data in Carnivora (mainly Canidae; figure ~\ref{Figure_example_coverageB}).
+Two contrasted results are shown on figures \ref{Figure_example_coverageA} and \ref{Figure_example_coverageB} with randomly distributed data in Cetartiodactyla (figure \ref{Figure_example_coverageA}) and clustered available data in Carnivora (mainly Canidae; figure \ref{Figure_example_coverageB}).
 
@@
 However, data availability varies in its structure as well as across taxonomic levels.
-Higher taxonomic levels are always better sampled than lower ones (table~\ref{Table_morpho_taxa_proportion}) and within these taxonomic levels, the structure of available data is mostly random apart for four orders (table\ref{Table_data_structure}).
-
-%Data collection
-We know we might have missed some data but we performed a thorough and repeatable search and still managed to collected 1422 OTUs which represents 38\% of all living mammals
-
-\begin{figure}[!htbp]
-\centering
-    \includegraphics[width=1\textwidth]{../andrew/citations.jpg}
-\caption{ }
-\label{bla}
-\end{figure}
+Higher taxonomic levels are always better sampled than lower ones (table \ref{Table_morpho_taxa_proportion}) and within these taxonomic levels, the structure of available data is mostly random apart for two orders (table \ref{Table_data_structure}).
 
 %\subsection{Data availability}
-The amount of living taxa with no available morphological data was surprisingly high at the species level (24/28 orders have less than 75\% taxa with available morphological data) which can lead to topological misplacements of fossils in Total Evidence trees.
-The effect of missing living taxa with morphological data has no significant impact on inferring the correct topology up to 25\% of missing data, beyond that threshold, both the position of "flying" taxa and the clades are significantly different from the correct topology \cite{GuillermeCooper}.
-At the species level within mammals, it is therefore likely to see topological artefacts for the placement of fossil taxa.
-However, the data availability is more promising at higher taxonomic level (i.e.
+The amount of living taxa with no available morphological data was surprisingly high at the species level: 24/28 orders have less than 75\% taxa with available morphological data (and two of the 28 orders are monospecific).
+This threshold of 75\% available data is the maximum amount of possible missing data (25\%) before having a significant effect on the topology of total evidence trees \cite{GuillermeCooper}.
+Beyond this threshold, there is a noteworthy displacement of the wildcard taxa (\textit{sensu} \cite{kearneyfragmentary2002}) as well as a important decrease in the conservation of clades \cite{GuillermeCooper}.
+At the species level within most orders of mammals, it is therefore likely to see topological artefacts for the placement of fossil taxa.
+However, the data availability seems to be less an issue at higher taxonomic level (i.e.
 the Genus and the Family level).
-This value is probably more important to consider since there is a slight discrepancy between neontological species and paleontological species.
-While the second ones are described based on accurately measured morphology, genetic distance, spacial distribution and even behaviour, the second one are usually estimate based on morphological, spatial and temporal differences only.
-Because of the lack of behavioural and genetic data, it is likely that cryptic species are often ignored in palaeontological data.
-Therefore, most palaeontological studies are using the Genus level for their smallest taxonomic OTU.
-In this frame of palaeontological data usage, the data availablity at the Genus level in living mammals should be our priority concern.
+This point is important to consider from a practical point of view because of the slight discrepancy between neontological and palaeontological species.
+While neontological species are described on accurately measured morphology, genetic distance, spacial distribution and even behaviour (e.g.
+\cite{kellymolecular2014}); palaeontological data can be based only on morphological, spatial and temporal data (e.g.
+\cite{ni2013oldest}).
+Because of this difference between neontological species (\textit{sensu} reproductive isolates) and paleontological species (\textit{sensu} morpho-species), most palaeontological studies are using the Genus level for their smallest taxonomic OTU (e.g.
+\cite{ni2013oldest,O'Leary08022013}).
+In this frame of palaeontological data usage, the data availability at the Genus level in living mammals should be our priority concern.
 
 %\subsection{Available data structure}
-However, it is encouraging that for most orders (except Carnivora and Chiroptera), the few available data is randomly distributed, therefore the addition of fossil taxa is likely to not be biased by oversampling but will probably branch near the taxa that has the most similar available data (as expected).
+Regardless the low level of available data, it is encouraging that for most orders (except Carnivora and Chiroptera), it is randomly distributed across the phylogeny.
+Therefore, it is likely that the addition of fossil taxa within these orders will not be biased by oversampling but will probably branch near the taxa that has the most similar available data (as expected).
 When only few data is available, the ideal scenario will be that the data is over-dispersed (i.e.
@@
 However, in the third scenario, when the data is clustered we expect two major biases to occur: (1) first, the fossil will not be able to branch within a clade containing no data (e.g.
-Herpestidae in Carnivora; figure ~\ref{Figure_example_coverageB}), and (2) second, the fossil will have a higher probability, at random, to branch within the clade containing most of the available data (e.g.
-Canidae in Carnivora; figure ~\ref{Figure_example_coverageB}).
+Herpestidae in Carnivora; figure \ref{Figure_missing_data_problem} and \ref{Figure_example_coverageB}), and (2) second, the fossil will have a higher probability, at random, to branch within the clade containing most of the available data (e.g.
+a Carnivora fossil will have more chance to branch, randomly, to the Canidae clade than any other clade in Carnivora; figure \ref{Figure_example_coverageB}).
 
@@
 %Caveats
-In this study we only used the raw number of characters and not there similarity.
-This might not work in reality since the data may actually not overlap.
-It might be that actually the amount of available data here is lower due to non overlap of characters....
+In this study, we treated all morphological matrices to be equal in a similar way that molecular matrices are: for example if a matrix A contains 100 characters for taxa X and Y, and a matrix B contains 50 characters for taxa X and Z, we assumed that both matrices can be combined in a supermatrix way leading to a matrix containing 150 characters for taxon X, 100 for taxon Y and 50 for taxon Z.
+However, it is clear that morphological data cannot be treated that way \cite{Brazeau2011}.
+In fact, for the same taxon, some characters might overlap (e.g.
+matrix A has a character coding for the shape of a particular morphological feature and matrix B as a first character coding for the presence of this morphological feature and a second one coding for its shape; in this case, these three characters are compound characters \cite{Brazeau2011}).
+However, in reasonably sized matrices ($>$ 100 characters \cite{GuillermeCooper,harrisonamong-character2014}) it is more likely that a number of characters are consistently conserved among the different matrices (e.g.
+\cite{ross1998phylogenetic,seiffert2003fossil,marivaux2005anthropoid,seiffert2005basal,bloch2007new,kay2008anatomy,silcox2008biogeographic,seiffert2009convergent,tabuce2009anthropoid,boyer2010astragalar,seiffert2010fossil,marivaux2013djebelemur,ni2013oldest}).
+A conservative approach to avoid compound characters could be to select only the latest matrix for each taxon.
 
 %Let's code some data
-Since the only way to fix missing data issues is to collect the missing data, it queit encouraging to know that the morphological data is available through natural history museum collections and that online plateforms and tools exist and are opperational to facilitate this giant task through collaboration and sharing open source and traceable data!
+Following the recommendations in \cite{GuillermeCooper}, one should code morphological characters for a maximum of living species in order to improve the topology of total evidence trees containing both living and fossil taxa.
+Since the data for living mammals is usually easily available in world wide vast natural history collections, we propose that an increased effort in coding morphological characters from living species should be done via engaging in collaborative data collection projects through web portals such as \textit{morphobank} \cite{morphobank}.
 

### 2015-03-02 Functions/summary.results.R

@@ -52,8 +52,10 @@ summary.results<-function(results, metric, save.path, file.save, caption, enviro
         for (column in 1:ncol(table_to_print)) {
             table_to_print[signif_data, column]<-paste('BOLD',table_to_print[signif_data, column], sep="")
         }
+        #Removing column Fraction
+        table_to_print<-table_to_print[,-3]
         #Fixing the column names
-        colnames(table_to_print)<-c("Order", "Taxonomic level", "Fraction of OTUs", "Percentage of OTUs", metric, "p-value")
+        colnames(table_to_print)<-c("Order", "Taxonomic level", "Percentage of OTUs", metric, "p-value")
         #Saving the table
         table<-xtable(table_to_print)
         caption(table)<-caption

### 2015-05-15 Manuscript/Missing_mammals-draft.tex
Abstract ideas?

@@
 \begin{abstract}
-
+Studying changes in global biodiversity through time and space is essential.
+For that we need methods to combine both palaeontological and neontological data.
+One promising method, the Total Evidence method, allows such thing but needs a lot of data.
+Especially morphological data from living taxa to allow the fossil taxa to accurately branch in the trees.
+Despite two centuries of morphological studies on living taxa, scientists using and generating such data mainly focus on palaeontological data.
+Therefore, even in well known groups such as mammal, there is a huge gap in our knowledge of morphological data for living mammals.
+In this study, using phylogenetic community structure methods, we quantify the availability of data in each mammalian order.
+And maybe at the end of the paper we propose some discussion on how to improve all that (go in museums!).
 \end{abstract}

### 2015-05-15 Manuscript/Missing_mammals-draft.tex
Minor updates in the intro

@@
 
-%Key words: total evidence method, data structure, phylogenetic, living, fossil, topology
+%Key words: Total Evidence method, data structure, phylogenetic, living, fossil, topology
 \bigskip
@@
 %We need fossils - dangerously paraphrased from the missing data paper
-Studying both living and fossil taxa together in essential to fully understand macrovelutionary patterns and processes and is becoming increasingly common among evolutionary biologists \cite{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
-Combining the global biodiversity (living and fossil) in such studies allows for example, to improve accuracy of the timing of diversification events (e.g.
+Studying both living and fossil taxa together in essential to fully understand macrovelutionary patterns and processes and is becoming increasingly common among evolutionary biologists \cite{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013,Wood01032013}.
+Combining the global biodiversity by including both living and fossil in such studies allows, for example, to improve accuracy of the timing of diversification events (e.g.
 \cite{ronquista2012}), to better understand relationships among lineages (e.g.
@@
 \cite{Meseguer01032015}).
+%And add STD paper?
+Or that's too much self cites?
 %We can use TEM for that
-In order to do so, one must efficiently combine living and fossil taxa data in macroevolutionary models.
-One trending method, called the total evidence method, allows to combine molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix (e.g.
+In order to perform such analysis, one must efficiently combine living and fossil taxa data in macroevolutionary models.
+One trending method, called the Total Evidence method \cite{eernissetaxonomic1993,ronquista2012}, allows to combine molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix (e.g.
 \cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014,Meseguer01032015}).
-This method allows to use all the available data and to treat fossil as tips rather than nodes and allows integrative phylogenetic inference methods such as tip-dating \cite{ronquista2012,Drummond01082012,BEASTmaster}.
+This method not only allows to use all the available data but also allows to treat fossil taxa as tips rather than nodes \textit{via} integrative phylogenetic inference methods such as tip-dating \cite{ronquista2012,Drummond01082012,Wood01032013,BEASTmaster}.
 
 %But missing data in living taxa problem
-However, the total evidence method requires a lot of data, both molecular and morphological for both living and fossil taxa.
-Entire sections of this data can sometimes be difficult or impossible to collect for every taxa present in the analysis.
-For example, fossil have really rarely molecular data available and morphological characters are rarely collected when molecular data is available (e.g.
-\cite{O'Leary08022013} \textit{vs.} \cite{meredithimpacts2011}).
-Therefore, the ability of this method to recover correct topology is expected to decrease with missing data and especially when few morphological data overlaps between the living and the fossil taxa \cite{GuillermeCooper}.
-%Owwwww the sneaky self cite ;-).
+However, the Total Evidence method requires, by definition, a lot of data.
+One must collect both molecular data for living taxa as well as morphological data for both living and fossil taxa, two types of data that require fairly different technical skills (e.g.
+\cite{meredithimpacts2011} \textit{vs.} \cite{O'Leary08022013}).
+Additionally, entire sections of this data can sometimes be difficult or impossible to collect for every taxa present in the analysis.
+For example, fossil have really rarely molecular data available and morphological characters are rarely collected when molecular data is available.
+Therefore, the ability of the Total Evidence method to recover correct topology is expected to decrease when there is a low overlap between the living and the fossil data \cite{GuillermeCooper}.
 The effect of missing data is most important on topology when too few living taxa have available morphological data \cite{GuillermeCooper}.
-For example, if there is no morphological available for any living taxa of an entire clade, it is impossible to link a fossil taxon to this clade because no morphological data will overlap between the fossil (regardless the amount of data available for the fossil) and the living taxa (that have no morphological data).
-This property of the total evidence method can lead to important topological incongruities because the fossil will only be able to branch to a clade of living taxa that contains morphological data, even if in reality, the fossil taxon does not belong to that clade (see figure ~\ref{Figure_missing_data_problem}).
+For example, if there is no morphological available for any living taxa within an entire clade, it is impossible to link a fossil taxon to this clade because no morphological data will overlap between the fossil (regardless the amount of data available for the fossil) and the living taxa (that have no morphological data).
+This property of the Total Evidence method can rapidly lead to important topological incongruities because the fossil will only be able to branch to a clade of living taxa that contains morphological data, even if in reality, the fossil taxon does not belong to that clade \cite{GuillermeCooper} (see figure ~\ref{Figure_missing_data_problem}).
 
@@
 %What we're doing
-It is therefore crucial to understand the distribution of the available morphological data for living taxa in a clade before using a total evidence approach because missing living taxa with morphological data can lead to two topological artefacts: (1) the impossibility for a fossil to branch in the right clade if there is no morphological data available in this clade \cite{GuillermeCooper}; and (2) the higher probability for a fossil to branch within a wrong clade that has more morphological data available for living taxa than the right clade \cite{GuillermeCooper}.
-
-In this study, we assess the level of available data in living mammals in order to highlight the two caveats described above.
-%In this study we investigate the amount of living mammal taxa with available morphological data to detect the two potential caveats described above that could lead to topological errors when building total evidence mammal phylogenetic trees.
-We extracted living mammals taxa with available morphological data from 256 phylogenetic matrices available online and measured the proportion of morphological data availability for each mammalian order.
-Additionally, in the mammalian orders where data was missing, we estimated the structure of the available data to detect if the available data was biases towards certain clades only or if it was randomly distributed using the Net Relatedness Index \cite{webb2002phylogenies}.
-
-%Finaly using phylogenetic targeting (CITE), we proposed a strategy to fill the gaps.
+It is therefore crucial to understand the distribution of the available morphological data for living taxa in a clade before using a Total Evidence approach because missing living taxa with morphological data can lead to two topological artefacts: (1) the impossibility for a fossil to branch in the right clade if there is no morphological data available in this clade \cite{GuillermeCooper}; and (2) the higher probability for a fossil to branch within a wrong clade that has more morphological data available for living taxa than the right clade \cite{GuillermeCooper}.
+In this study, we assess the level of available morphological data in living mammals in order to highlight the two caveats described above.
+%In this study we investigate the amount of living mammal taxa with available morphological data to detect the two potential caveats described above that could lead to topological errors when building Total Evidence mammal phylogenetic trees.
+We collected living mammals morphological data from @256 phylogenetic matrices available online and measured the proportion of morphological data availability for each mammalian order.
+Additionally, in the mammalian orders where data was missing, we estimated the structure of the available data to detect if the available data was biased towards certain clades only or if it was randomly distributed using the Net Relatedness Index \cite{webb2002phylogenies}.
 
@@
 The amount of living taxa with no available morphological data was surprisingly high at the species level: 24/28 orders have less than 75\% taxa with available morphological data (and two of the 28 orders are monospecific).
-This threshold of 75\% available data is the maximum amount of possible missing data (25\%) before having a significant effect on the topology of total evidence trees \cite{GuillermeCooper}.
+This threshold of 75\% available data is the maximum amount of possible missing data (25\%) before having a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
 Beyond this threshold, there is a noteworthy displacement of the wildcard taxa (\textit{sensu} \cite{kearneyfragmentary2002}) as well as a important decrease in the conservation of clades \cite{GuillermeCooper}.
@@
 %Let's code some data
-Following the recommendations in \cite{GuillermeCooper}, one should code morphological characters for a maximum of living species in order to improve the topology of total evidence trees containing both living and fossil taxa.
+Following the recommendations in \cite{GuillermeCooper}, one should code morphological characters for a maximum of living species in order to improve the topology of Total Evidence trees containing both living and fossil taxa.
 Since the data for living mammals is usually easily available in world wide vast natural history collections, we propose that an increased effort in coding morphological characters from living species should be done via engaging in collaborative data collection projects through web portals such as \textit{morphobank} \cite{morphobank}.

### 2015-05-17 Manuscript/Missing_mammals-draft.tex
Updated intro

@@
 Additionally, entire sections of this data can sometimes be difficult or impossible to collect for every taxa present in the analysis.
-For example, fossil have really rarely molecular data available and morphological characters are rarely collected when molecular data is available.
-Therefore, the ability of the Total Evidence method to recover correct topology is expected to decrease when there is a low overlap between the living and the fossil data \cite{GuillermeCooper}.
+For example, fossil have really rarely molecular data available and morphological characters are rarely collected for living taxa when molecular data is available (e.g.
+\cite{slaterphylogenetic2013,beckancient2014}).
+This difference between both data can lead to topological errors in phylogenetic inference \cite{GuillermeCooper}.
+In fact, the ability of the Total Evidence method to recover correct topology is expected to decrease when there is a low overlap between morphological data from the living and the fossil taxa \cite{GuillermeCooper}.
 The effect of missing data is most important on topology when too few living taxa have available morphological data \cite{GuillermeCooper}.
@@
 \caption{Example of topological errors due to missing morphological data in living taxa.
-If a phylogeny contains two clades, for example Proboscidea and Cetartiodactyla, with molecular data for both (in grey) but only morphological data for Proboscidea only (in red).
-If an additional Cetartiodactyla fossil (with no molecular data) is added to the phylogeny, it will erroneously branch with the Proboscidea clade instead of the Cetartiodactyla one because no morphological data will overlap between the fossil Cetartiodactyla and the living ones.}
+If a phylogeny contains two clades, for example Aves and Mammalia, with molecular data (in grey) for both but only morphological data (in red) for Aves.
+If an additional Mammalia fossil (with no molecular data) is added to the phylogeny, it will erroneously branch with the Aves clade instead of the Mammalia one because no morphological data will overlap between the fossil Mammalia and the living ones.}
 \label{Figure_missing_data_problem}

### 2015-05-18 Manuscript/Missing_mammals-draft.tex
Citation style change

@@
 \documentclass[12pt,letterpaper]{article}
+\usepackage{natbib}
 
@@
 %We need fossils - dangerously paraphrased from the missing data paper
-Studying both living and fossil taxa together in essential to fully understand macrovelutionary patterns and processes and is becoming increasingly common among evolutionary biologists \cite{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013,Wood01032013}.
-Combining the global biodiversity by including both living and fossil in such studies allows, for example, to improve accuracy of the timing of diversification events (e.g.
-\cite{ronquista2012}), to better understand relationships among lineages (e.g.
-\cite{beckancient2014}) or to infer biogeographical patterns through time (e.g.
-\cite{Meseguer01032015}).
+Studying both living and fossil taxa together in essential to fully understand macrovelutionary patterns and processes and is becoming increasingly common among evolutionary biologists \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013,Wood01032013}.
+Combining the global biodiversity by including both living and fossil in such studies allows, for example, to improve accuracy of the timing of diversification events \citep[e.g.][]{ronquista2012}, to better understand relationships among lineages \citep[e.g.][]{beckancient2014} or to infer biogeographical patterns through time \citep[e.g.][]{Meseguer01032015}.
 %And add STD paper?
@@
 In order to perform such analysis, one must efficiently combine living and fossil taxa data in macroevolutionary models.
-One trending method, called the Total Evidence method \cite{eernissetaxonomic1993,ronquista2012}, allows to combine molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix (e.g.
-\cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014,Meseguer01032015}).
-This method not only allows to use all the available data but also allows to treat fossil taxa as tips rather than nodes \textit{via} integrative phylogenetic inference methods such as tip-dating \cite{ronquista2012,Drummond01082012,Wood01032013,BEASTmaster}.
+One trending method, called the Total Evidence method \citep{eernissetaxonomic1993,ronquista2012}, allows to combine molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014,Meseguer01032015}.
+This method not only allows to use all the available data but also allows to treat fossil taxa as tips rather than nodes \textit{via} integrative phylogenetic inference methods such as tip-dating \citep{ronquista2012,Drummond01082012,Wood01032013,BEASTmaster}.
 
@@
 However, the Total Evidence method requires, by definition, a lot of data.
-One must collect both molecular data for living taxa as well as morphological data for both living and fossil taxa, two types of data that require fairly different technical skills (e.g.
-\cite{meredithimpacts2011} \textit{vs.} \cite{O'Leary08022013}).
+One must collect both molecular data for living taxa as well as morphological data for both living and fossil taxa, two types of data that require fairly different technical skills \citep[e.g.][]{meredithimpacts2011} \textit{vs.} \citep{O'Leary08022013}.
 Additionally, entire sections of this data can sometimes be difficult or impossible to collect for every taxa present in the analysis.
-For example, fossil have really rarely molecular data available and morphological characters are rarely collected for living taxa when molecular data is available (e.g.
-\cite{slaterphylogenetic2013,beckancient2014}).
-This difference between both data can lead to topological errors in phylogenetic inference \cite{GuillermeCooper}.
-In fact, the ability of the Total Evidence method to recover correct topology is expected to decrease when there is a low overlap between morphological data from the living and the fossil taxa \cite{GuillermeCooper}.
-The effect of missing data is most important on topology when too few living taxa have available morphological data \cite{GuillermeCooper}.
+For example, fossil have really rarely molecular data available and morphological characters are rarely collected for living taxa when molecular data is available \citep[e.g.][]{slaterphylogenetic2013,beckancient2014}.
+This difference between both data can lead to topological errors in phylogenetic inference \citep{GuillermeCooper}.
+In fact, the ability of the Total Evidence method to recover correct topology is expected to decrease when there is a low overlap between morphological data from the living and the fossil taxa \citep{GuillermeCooper}.
+The effect of missing data is most important on topology when too few living taxa have available morphological data \citep{GuillermeCooper}.
 For example, if there is no morphological available for any living taxa within an entire clade, it is impossible to link a fossil taxon to this clade because no morphological data will overlap between the fossil (regardless the amount of data available for the fossil) and the living taxa (that have no morphological data).
-This property of the Total Evidence method can rapidly lead to important topological incongruities because the fossil will only be able to branch to a clade of living taxa that contains morphological data, even if in reality, the fossil taxon does not belong to that clade \cite{GuillermeCooper} (see figure ~\ref{Figure_missing_data_problem}).
+This property of the Total Evidence method can rapidly lead to important topological incongruities because the fossil will only be able to branch to a clade of living taxa that contains morphological data, even if in reality, the fossil taxon does not belong to that clade \citep{GuillermeCooper} (see figure ~\ref{Figure_missing_data_problem}.
 
@@
 %What we're doing
-It is therefore crucial to understand the distribution of the available morphological data for living taxa in a clade before using a Total Evidence approach because missing living taxa with morphological data can lead to two topological artefacts: (1) the impossibility for a fossil to branch in the right clade if there is no morphological data available in this clade \cite{GuillermeCooper}; and (2) the higher probability for a fossil to branch within a wrong clade that has more morphological data available for living taxa than the right clade \cite{GuillermeCooper}.
+It is therefore crucial to understand the distribution of the available morphological data for living taxa in a clade before using a Total Evidence approach because missing living taxa with morphological data can lead to two topological artefacts: (1) the impossibility for a fossil to branch in the right clade if there is no morphological data available in this clade \citep{GuillermeCooper}; and (2) the higher probability for a fossil to branch within a wrong clade that has more morphological data available for living taxa than the right clade \citep{GuillermeCooper}.
 In this study, we assess the level of available morphological data in living mammals in order to highlight the two caveats described above.
@@
 We collected living mammals morphological data from @256 phylogenetic matrices available online and measured the proportion of morphological data availability for each mammalian order.
-Additionally, in the mammalian orders where data was missing, we estimated the structure of the available data to detect if the available data was biased towards certain clades only or if it was randomly distributed using the Net Relatedness Index \cite{webb2002phylogenies}.
+Additionally, in the mammalian orders where data was missing, we estimated the structure of the available data to detect if the available data was biased towards certain clades only or if it was randomly distributed using the Net Relatedness Index \citep{webb2002phylogenies}.
 
@@
 \subsection{Matrices search}
-To investigate the available living taxa with morphological data, we downloaded morphological matrices from three main public databases: morphobank (\texttt{http://www.morphobank.org/}) \cite{morphobank}, Graeme Lloyd's website (\texttt{http://graemetlloyd.com/}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}).
+To investigate the available living taxa with morphological data, we downloaded morphological matrices from three main public databases: morphobank (\texttt{http://www.morphobank.org/} \citep{morphobank}, Graeme Lloyd's website (\texttt{http://graemetlloyd.com/} and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}.
 We downloaded all the matrices containing any fossil or living mammal taxa from these databases.
@@
 \textit{H.
-sapiens} was transformed in \textit{Homo sapiens}).
+sapiens} was transformed in \textit{Homo sapiens}.
 We assigned each species as being either living or fossil using a taxonomic matching algorithm.
-We designated as living OTU all the OTUs that where either present in \cite{FritzTree} or \cite{wilson2005mammal}.
+We designated as living OTU all the OTUs that where either present in \citep{FritzTree} or \citep{wilson2005mammal}.
 We designated as fossil OTUs all the OTUs that where present in the Paleobiology database.
 For the OTUs neither labelled as living or fossil we tried to decompose the OTUs name (i.e.
-\textit{Homo$\_$sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with any taxonomic level (Family, Genus, Species, etc.) from \cite{wilson2005mammal}.
+\textit{Homo$\_$sapiens} became \textit{Homo} and \textit{sapiens} and tried to match the first element with any taxonomic level (Family, Genus, Species, etc.) from \citep{wilson2005mammal}.
 The matching OTUs where labelled as living and the ones still not matching where ignored and labelled as non-applicable (NA; see supplementary material for more details on the taxonomic matching algorithm).
@@
 matrices, we selected only the ones that displayed at least 100 morphological characters per OTU.
-This arbitrary threshold of a minimal number of morphological characters was chosen to be in adequacy with \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
-Also this threshold avoids biases towards small matrices that could be either not informative \cite{wagner2000} (e.g.
-too few characters not allowing character overlap among matrices) or made of non-applicable characters \cite{Brazeau2011} (e.g.
+This arbitrary threshold of a minimal number of morphological characters was chosen to be in adequacy with \citep{GuillermeCooper} and \citep{harrisonamong-character2014}.
+Also this threshold avoids biases towards small matrices that could be either not informative \citep{wagner2000} (e.g.
+too few characters not allowing character overlap among matrices) or made of non-applicable characters \citep{Brazeau2011} (e.g.
 antlers which are sexual dimorphic characters proper to a specific clade).
@@
 To assess the data availability per mammal order, we calculated the percentage of OTUs with morphological data for three different taxonomic levels: Family, Genus and Species.
-We highlighted all the orders containing less than 25\% of living taxa with morphological data because their amount of missing data ($>$75\%) was higher than in \cite{GuillermeCooper} and therefore highly probable of having wrong topology due to the missing data.
-On the opposite, orders with up to 25\% of missing data (75\% of available data) were shown to have no significant effect on topology \cite{GuillermeCooper}.
+We highlighted all the orders containing less than 25\% of living taxa with morphological data because their amount of missing data ($>$75\%) was higher than in \citep{GuillermeCooper} and therefore highly probable of having wrong topology due to the missing data.
+On the opposite, orders with up to 25\% of missing data (75\% of available data) were shown to have no significant effect on topology \citep{GuillermeCooper}.
 We therefore used this value as a threshold of "good" data sampling.
@@
 For the orders with no morphological data for all its OTUs at the three different taxonomic levels (Family, Genus and Species), we investigated the structure of the available data to test if it was either (i) randomly distributed, (ii) over-dispersed or (iii) clustered.
-To measure the structure of the available data we used classic community structure metric from the \texttt{picante} R package \cite{picante}.
+To measure the structure of the available data we used classic community structure metric from the \texttt{picante} R package \citep{picante}.
 We compared the structure of the available data for each order to the structure of a potentially fully sampled order (i.e.
 only the OTUs with available morphological data \textit{versus} all the OTUs).
-For each orders and taxonomic levels that presented OTUs with no available morphological data, we calculated the Net Relatedness Index (NRI) which quantifies the overall distribution of the data with negative values showing more dispersed data and positive values more clustered data than expected by the null model (random) \cite{webb2002phylogenies}.
+For each orders and taxonomic levels that presented OTUs with no available morphological data, we calculated the Net Relatedness Index (NRI) which quantifies the overall distribution of the data with negative values showing more dispersed data and positive values more clustered data than expected by the null model (random) \citep{webb2002phylogenies}.
 We choose to present only the NRI values because they have been shown to be slightly less sensitive to the structure of the phylogeny (i.e.
-branch length and topology) \cite{NRI,journal.pone.0004390} but we also calculated the two other common phylogenetic structure indices: Faith's Phylogenetic Distance (PD) \cite{Faith19921} and the Nearest Taxon Index (NTI) \cite{webb2002phylogenies}.
+branch length and topology) \citep{NRI,journal.pone.0004390} but we also calculated the two other common phylogenetic structure indices: Faith's Phylogenetic Distance (PD) \citep{Faith19921} and the Nearest Taxon Index (NTI) \citep{webb2002phylogenies}.
 Both metrics are available in the Supplementary results.
@@
 At the Genus level however only 3/28 orders have less than 25\% of taxa with morphological data and 16/28 have less than 75\%.
-Finally, at the family level no order has less than 25\% taxa with available morphological data and only 5/28 have less than 75\% (table \ref{Table_morpho_taxa_proportion}).
+Finally, at the family level no order has less than 25\% taxa with available morphological data and only 5/28 have less than 75\% (table \ref{Table_morpho_taxa_proportion}.
 
@@
 \subsection{Available data structure}
-Among the orders containing at least one OTU with no available morphological data, only two orders are significantly clustered: Carnivora and Chiroptera both at the species and the genus level (table~\ref{Table_data_structure}).
+Among the orders containing at least one OTU with no available morphological data, only two orders are significantly clustered: Carnivora and Chiroptera both at the species and the genus level (table~\ref{Table_data_structure}.
 
@@
 
-Two contrasted results are shown on figures \ref{Figure_example_coverageA} and \ref{Figure_example_coverageB} with randomly distributed data in Cetartiodactyla (figure \ref{Figure_example_coverageA}) and clustered available data in Carnivora (mainly Canidae; figure \ref{Figure_example_coverageB}).
+Two contrasted results are shown on figures \ref{Figure_example_coverageA} and \ref{Figure_example_coverageB} with randomly distributed data in Cetartiodactyla (figure \ref{Figure_example_coverageA} and clustered available data in Carnivora (mainly Canidae; figure \ref{Figure_example_coverageB}.
 
@@
 \section{Discussion}
-Our results show that even though relations among living mammals are well known in a molecular framework (e.g.
-\cite{FritzTree,meredithimpacts2011,May-Collado-PeerJ}) only little morphological data is available.
-This is primarily due to the fact that that morphological studies are mainly focusing on fossil taxa rather than living taxa (e.g.
-\cite{O'Leary08022013,ni2013oldest}).
+Our results show that even though relations among living mammals are well known in a molecular framework \citep[e.g.][]{FritzTree,meredithimpacts2011,May-Collado-PeerJ} only little morphological data is available.
+This is primarily due to the fact that that morphological studies are mainly focusing on fossil taxa rather than living taxa \citep[e.g.][]{O'Leary08022013,ni2013oldest}.
 However, data availability varies in its structure as well as across taxonomic levels.
-Higher taxonomic levels are always better sampled than lower ones (table \ref{Table_morpho_taxa_proportion}) and within these taxonomic levels, the structure of available data is mostly random apart for two orders (table \ref{Table_data_structure}).
+Higher taxonomic levels are always better sampled than lower ones (table \ref{Table_morpho_taxa_proportion} and within these taxonomic levels, the structure of available data is mostly random apart for two orders (table \ref{Table_data_structure}.
 
@@
 The amount of living taxa with no available morphological data was surprisingly high at the species level: 24/28 orders have less than 75\% taxa with available morphological data (and two of the 28 orders are monospecific).
-This threshold of 75\% available data is the maximum amount of possible missing data (25\%) before having a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
-Beyond this threshold, there is a noteworthy displacement of the wildcard taxa (\textit{sensu} \cite{kearneyfragmentary2002}) as well as a important decrease in the conservation of clades \cite{GuillermeCooper}.
+This threshold of 75\% available data is the maximum amount of possible missing data (25\%) before having a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.
+Beyond this threshold, there is a noteworthy displacement of the wildcard taxa (\textit{sensu} \citep{kearneyfragmentary2002} as well as a important decrease in the conservation of clades \citep{GuillermeCooper}.
 At the species level within most orders of mammals, it is therefore likely to see topological artefacts for the placement of fossil taxa.
@@
 This point is important to consider from a practical point of view because of the slight discrepancy between neontological and palaeontological species.
-While neontological species are described on accurately measured morphology, genetic distance, spacial distribution and even behaviour (e.g.
-\cite{kellymolecular2014}); palaeontological data can be based only on morphological, spatial and temporal data (e.g.
-\cite{ni2013oldest}).
-Because of this difference between neontological species (\textit{sensu} reproductive isolates) and paleontological species (\textit{sensu} morpho-species), most palaeontological studies are using the Genus level for their smallest taxonomic OTU (e.g.
-\cite{ni2013oldest,O'Leary08022013}).
+While neontological species are described on accurately measured morphology, genetic distance, spacial distribution and even behaviour \citep[e.g.][]{kellymolecular2014}; palaeontological data can be based only on morphological, spatial and temporal data \citep[e.g.][]{ni2013oldest}.
+Because of this difference between neontological species (\textit{sensu} reproductive isolates) and paleontological species (\textit{sensu} morpho-species), most palaeontological studies are using the Genus level for their smallest taxonomic OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
 In this frame of palaeontological data usage, the data availability at the Genus level in living mammals should be our priority concern.
@@
 that there is data in at least every sub-clade) in order to maximize the possibilities of the fossil to branch in the right clade.
-In the second scenario, when the data is randomly distributed we expected no special bias in the placement of the fossil \cite{GuillermeCooper}.
+In the second scenario, when the data is randomly distributed we expected no special bias in the placement of the fossil \citep{GuillermeCooper}.
 However, in the third scenario, when the data is clustered we expect two major biases to occur: (1) first, the fossil will not be able to branch within a clade containing no data (e.g.
-Herpestidae in Carnivora; figure \ref{Figure_missing_data_problem} and \ref{Figure_example_coverageB}), and (2) second, the fossil will have a higher probability, at random, to branch within the clade containing most of the available data (e.g.
-a Carnivora fossil will have more chance to branch, randomly, to the Canidae clade than any other clade in Carnivora; figure \ref{Figure_example_coverageB}).
+Herpestidae in Carnivora; figure \ref{Figure_missing_data_problem} and \ref{Figure_example_coverageB}, and (2) second, the fossil will have a higher probability, at random, to branch within the clade containing most of the available data (e.g.
+a Carnivora fossil will have more chance to branch, randomly, to the Canidae clade than any other clade in Carnivora; figure \ref{Figure_example_coverageB}.
 
-%For example in Carnivora, their is no available data for Herpestidae, therefore any fossil Herpestidae will not be able to be correctly placed in the phylogeny (figure ~\ref{Figure_example_coverageB}).
-On the other hand, any Carnivora fossil with uncertain phylogenetic affinities (\textit{Incertae sedis}) will have a higher probability of branching by chance within the Canidae because they are oversampled within the Carnivora (figure ~\ref{Figure_example_coverageB}).
+%For example in Carnivora, their is no available data for Herpestidae, therefore any fossil Herpestidae will not be able to be correctly placed in the phylogeny (figure ~\ref{Figure_example_coverageB}.
+On the other hand, any Carnivora fossil with uncertain phylogenetic affinities (\textit{Incertae sedis} will have a higher probability of branching by chance within the Canidae because they are oversampled within the Carnivora (figure ~\ref{Figure_example_coverageB}.
 
@@
 In this study, we treated all morphological matrices to be equal in a similar way that molecular matrices are: for example if a matrix A contains 100 characters for taxa X and Y, and a matrix B contains 50 characters for taxa X and Z, we assumed that both matrices can be combined in a supermatrix way leading to a matrix containing 150 characters for taxon X, 100 for taxon Y and 50 for taxon Z.
-However, it is clear that morphological data cannot be treated that way \cite{Brazeau2011}.
+However, it is clear that morphological data cannot be treated that way \citep{Brazeau2011}.
 In fact, for the same taxon, some characters might overlap (e.g.
-matrix A has a character coding for the shape of a particular morphological feature and matrix B as a first character coding for the presence of this morphological feature and a second one coding for its shape; in this case, these three characters are compound characters \cite{Brazeau2011}).
-However, in reasonably sized matrices ($>$ 100 characters \cite{GuillermeCooper,harrisonamong-character2014}) it is more likely that a number of characters are consistently conserved among the different matrices (e.g.
-\cite{ross1998phylogenetic,seiffert2003fossil,marivaux2005anthropoid,seiffert2005basal,bloch2007new,kay2008anatomy,silcox2008biogeographic,seiffert2009convergent,tabuce2009anthropoid,boyer2010astragalar,seiffert2010fossil,marivaux2013djebelemur,ni2013oldest}).
+matrix A has a character coding for the shape of a particular morphological feature and matrix B as a first character coding for the presence of this morphological feature and a second one coding for its shape; in this case, these three characters are compound characters \citep{Brazeau2011}.
+However, in reasonably sized matrices ($>$ 100 characters \citep{GuillermeCooper,harrisonamong-character2014} it is more likely that a number of characters are consistently conserved among the different matrices \citep[e.g.][]{ross1998phylogenetic,seiffert2003fossil,marivaux2005anthropoid,seiffert2005basal,bloch2007new,kay2008anatomy,silcox2008biogeographic,seiffert2009convergent,tabuce2009anthropoid,boyer2010astragalar,seiffert2010fossil,marivaux2013djebelemur,ni2013oldest}.
 A conservative approach to avoid compound characters could be to select only the latest matrix for each taxon.
@@
 %Let's code some data
-Following the recommendations in \cite{GuillermeCooper}, one should code morphological characters for a maximum of living species in order to improve the topology of Total Evidence trees containing both living and fossil taxa.
-Since the data for living mammals is usually easily available in world wide vast natural history collections, we propose that an increased effort in coding morphological characters from living species should be done via engaging in collaborative data collection projects through web portals such as \textit{morphobank} \cite{morphobank}.
+Following the recommendations in \citep{GuillermeCooper}, one should code morphological characters for a maximum of living species in order to improve the topology of Total Evidence trees containing both living and fossil taxa.
+Since the data for living mammals is usually easily available in world wide vast natural history collections, we propose that an increased effort in coding morphological characters from living species should be done via engaging in collaborative data collection projects through web portals such as \textit{morphobank} \citep{morphobank}.
 
@@
 
-\bibliographystyle{vancouver}
+\bibliographystyle{sysbio}
 \bibliography{References}

### 2015-05-18 Manuscript/Missing_mammals-draft.tex
Changed to one sentence lines

@@
 \section{Introduction}
-%We need fossils - dangerously paraphrased from the missing data paper
 Studying both living and fossil taxa together in essential to fully understand macrovelutionary patterns and processes and is becoming increasingly common among evolutionary biologists \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013,Wood01032013}.
 Combining the global biodiversity by including both living and fossil in such studies allows, for example, to improve accuracy of the timing of diversification events \citep[e.g.][]{ronquista2012}, to better understand relationships among lineages \citep[e.g.][]{beckancient2014} or to infer biogeographical patterns through time \citep[e.g.][]{Meseguer01032015}.
-%And add STD paper?
-Or that's too much self cites?
-%We can use TEM for that
+%And add STD paper? Or that's too much self cites?
 In order to perform such analysis, one must efficiently combine living and fossil taxa data in macroevolutionary models.
@@
 
-%But missing data in living taxa problem
 However, the Total Evidence method requires, by definition, a lot of data.
@@
 matrices containing a total of 9411 % @@@ Number of living OTUs
-operational taxonomic units (OTUs) from the combination of both searches (public repositories and Google Scholar). The list of matrices is available in the supplementary materials.
+operational taxonomic units (OTUs) from the combination of both searches (public repositories and Google Scholar).
+The list of matrices is available in the supplementary materials.
 
@@
 %Phylotargeting
-%This part as not been done at all!
-I was looking for a non-GUI version of Arnold and Nunn's software but there is none (Arnold confirmed by email).
+%This part as not been done at all! I was looking for a non-GUI version of Arnold and Nunn's software but there is none (Arnold confirmed by email).
 Therefore because I was focusing on the part above and to make it repeatable I didn't did that part yet.
-However, I still planing on doing it!
-I just need to find a bit of time to write some code to make it non-GUI (and to be able to use big trees).
+However, I still planing on doing it! I just need to find a bit of time to write some code to make it non-GUI (and to be able to use big trees).
 
-All the following procedure is repeatable and available on GitHub. %Or what is the right way to show it's reproducible?
+All the following procedure is repeatable and available on GitHub.
+%Or what is the right way to show it's reproducible?
 
@@
     \includegraphics[width=1\textwidth]{example_coverageA.pdf}
-\caption{Distribution of available morphological data across Cetartiodactyla. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\caption{Distribution of available morphological data across Cetartiodactyla.
+Edges are colored in grey when no morphological data is available or in red when data is available.}
 \label{Figure_example_coverageA}
@@
     \includegraphics[width=1\textwidth]{example_coverageB.pdf}
-\caption{Distribution of available morphological data across Carnivora. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\caption{Distribution of available morphological data across Carnivora.
+Edges are colored in grey when no morphological data is available or in red when data is available.}
 \label{Figure_example_coverageB}

### 2015-05-18 Manuscript/Missing_mammals-draft.tex
Starting methods edits

@@
 \subsection{Matrices search}
-To investigate the available living taxa with morphological data, we downloaded morphological matrices from three main public databases: morphobank (\texttt{http://www.morphobank.org/} \citep{morphobank}, Graeme Lloyd's website (\texttt{http://graemetlloyd.com/} and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}.
+To investigate the living mammals with available morphological data, we downloaded cladistic matrices from three major public databases: morphobank (\texttt{http://www.morphobank.org/}) \citep{morphobank}, Graeme Lloyd's website (\texttt{http://graemetlloyd.com/}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}).
 We downloaded all the matrices containing any fossil or living mammal taxa from these databases.
-Additionally we ran a thorough Google Scholar search for matrices that might not have been uploaded on the previously cited data bases.
-We downloaded the additional morphological characters matrices from the 20 first Google search results matching with our selected keywords and with any of the 35 taxonomic levels (see supplementary materials for detailed description of the procedure).
+Additionally we ran a thorough Google Scholar search for matrices that might not have been uploaded on the previously mentioned data bases.
+We downloaded the additional cladistic matrices from the 20 first Google search results matching with our selected keywords and with any of the 35 taxonomic levels (mammals Orders, Infraclasses and Class; see supplementary materials for detailed description of the procedure).
 We downloaded 256 % @@@ Number of matrices
@@
 operational taxonomic units (OTUs) from the combination of both searches (public repositories and Google Scholar).
-The list of matrices is available in the supplementary materials.
+The full list of matrices with their associated citations is available in the supplementary materials. % link
 
-We then transformed all the matrices to be in the same nexus format.
+We then transformed all the matrices to be in nexus format.
 We also standardised the taxonomic nomenclature by fixing invalid binomial inputs to match with the official taxonomic nomenclature rules (i.e.
 \textit{H.
-sapiens} was transformed in \textit{Homo sapiens}.
-We assigned each species as being either living or fossil using a taxonomic matching algorithm.
-We designated as living OTU all the OTUs that where either present in \citep{FritzTree} or \citep{wilson2005mammal}.
-We designated as fossil OTUs all the OTUs that where present in the Paleobiology database.
-For the OTUs neither labelled as living or fossil we tried to decompose the OTUs name (i.e.
+sapiens} was transformed in \textit{Homo sapiens}).
+We assigned each species as being either living or fossil using a taxonomic matching algorithm described as follow: (1) we designated as living OTU all the OTUs that where either present in \citep{FritzTree} or \citep{wilson2005mammal}; (2)we designated as fossil OTUs all the OTUs that where present in the Paleobiology database and (3) for the OTUs neither labelled as living or fossil we tried to decompose the OTUs name (i.e.
 \textit{Homo$\_$sapiens} became \textit{Homo} and \textit{sapiens} and tried to match the first element with any taxonomic level (Family, Genus, Species, etc.) from \citep{wilson2005mammal}.
+% link
 The matching OTUs where labelled as living and the ones still not matching where ignored and labelled as non-applicable (NA; see supplementary material for more details on the taxonomic matching algorithm).

### 2015-05-19 Manuscript/Missing_mammals-draft.tex
Updated methods and results

@@
-%Authors guidlines: http://royalsocietypublishing.org/instructions-authors#question5
+%Authors guidlines: http://royalsocietypublishing.org/instructions-authors
+% 2500 words max (includes the title page, abstract, references, acknowledgements and figure/table legends)
+% current version is above 4500. I think a big cut down can be done on the references.
+% We allow a maximum of 4 displays, only 2 of which can be figures.
 
@@
 \noindent{\Large \bf Morphological data availability in living mammals}
-%Or: Missing morphological data in living mammals - but I prefer the first one emphasizing on the availability of it. The data is not missing per se, it's just not coded!
+%Or: Missing cladistic data in living mammals - but I prefer the first one emphasizing on the availability of it. The data is not missing per se, it's just not coded!
 
@@
 One promising method, the Total Evidence method, allows such thing but needs a lot of data.
-Especially morphological data from living taxa to allow the fossil taxa to accurately branch in the trees.
+Especially cladistic data from living taxa to allow the fossil taxa to accurately branch in the trees.
 Despite two centuries of morphological studies on living taxa, scientists using and generating such data mainly focus on palaeontological data.
-Therefore, even in well known groups such as mammal, there is a huge gap in our knowledge of morphological data for living mammals.
+Therefore, even in well known groups such as mammal, there is a huge gap in our knowledge of cladistic data for living mammals.
 In this study, using phylogenetic community structure methods, we quantify the availability of data in each mammalian order.
@@
 In order to perform such analysis, one must efficiently combine living and fossil taxa data in macroevolutionary models.
-One trending method, called the Total Evidence method \citep{eernissetaxonomic1993,ronquista2012}, allows to combine molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014,Meseguer01032015}.
+One trending method, called the Total Evidence method \citep{eernissetaxonomic1993,ronquista2012}, allows to combine molecular data from living taxa and cladistic data from both living and fossil taxa in a supermatrix \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014,Meseguer01032015}.
 This method not only allows to use all the available data but also allows to treat fossil taxa as tips rather than nodes \textit{via} integrative phylogenetic inference methods such as tip-dating \citep{ronquista2012,Drummond01082012,Wood01032013,BEASTmaster}.
@@
 However, the Total Evidence method requires, by definition, a lot of data.
-One must collect both molecular data for living taxa as well as morphological data for both living and fossil taxa, two types of data that require fairly different technical skills \citep[e.g.][]{meredithimpacts2011} \textit{vs.} \citep{O'Leary08022013}.
+One must collect both molecular data for living taxa as well as cladistic data for both living and fossil taxa, two types of data that require fairly different technical skills \citep[e.g.][]{meredithimpacts2011} \textit{vs.} \citep{O'Leary08022013}.
 Additionally, entire sections of this data can sometimes be difficult or impossible to collect for every taxa present in the analysis.
@@
 This difference between both data can lead to topological errors in phylogenetic inference \citep{GuillermeCooper}.
-In fact, the ability of the Total Evidence method to recover correct topology is expected to decrease when there is a low overlap between morphological data from the living and the fossil taxa \citep{GuillermeCooper}.
-The effect of missing data is most important on topology when too few living taxa have available morphological data \citep{GuillermeCooper}.
-For example, if there is no morphological available for any living taxa within an entire clade, it is impossible to link a fossil taxon to this clade because no morphological data will overlap between the fossil (regardless the amount of data available for the fossil) and the living taxa (that have no morphological data).
-This property of the Total Evidence method can rapidly lead to important topological incongruities because the fossil will only be able to branch to a clade of living taxa that contains morphological data, even if in reality, the fossil taxon does not belong to that clade \citep{GuillermeCooper} (see figure ~\ref{Figure_missing_data_problem}.
+In fact, the ability of the Total Evidence method to recover correct topology is expected to decrease when there is a low overlap between cladistic data from the living and the fossil taxa \citep{GuillermeCooper}.
+The effect of missing data is most important on topology when too few living taxa have available cladistic data \citep{GuillermeCooper}.
+For example, if there is no morphological available for any living taxa within an entire clade, it is impossible to link a fossil taxon to this clade because no cladistic data will overlap between the fossil (regardless the amount of data available for the fossil) and the living taxa (that have no cladistic data).
+This property of the Total Evidence method can rapidly lead to important topological incongruities because the fossil will only be able to branch to a clade of living taxa that contains cladistic data, even if in reality, the fossil taxon does not belong to that clade \citep{GuillermeCooper} (see figure ~\ref{Figure_missing_data_problem}).
 
@@
     \includegraphics[width=1\textwidth]{MissingDataFigure.pdf}
-\caption{Example of topological errors due to missing morphological data in living taxa.
-If a phylogeny contains two clades, for example Aves and Mammalia, with molecular data (in grey) for both but only morphological data (in red) for Aves.
-If an additional Mammalia fossil (with no molecular data) is added to the phylogeny, it will erroneously branch with the Aves clade instead of the Mammalia one because no morphological data will overlap between the fossil Mammalia and the living ones.}
+\caption{Example of topological errors due to missing cladistic data in living taxa.
+If a phylogeny contains two clades, for example Aves and Mammalia, with molecular data (in grey) for both but only cladistic data (in red) for Aves.
+If an additional Mammalia fossil (with no molecular data) is added to the phylogeny, it will erroneously branch with the Aves clade instead of the Mammalia one because no cladistic data will overlap between the fossil Mammalia and the living ones.}
 \label{Figure_missing_data_problem}
@@
 %What we're doing
-It is therefore crucial to understand the distribution of the available morphological data for living taxa in a clade before using a Total Evidence approach because missing living taxa with morphological data can lead to two topological artefacts: (1) the impossibility for a fossil to branch in the right clade if there is no morphological data available in this clade \citep{GuillermeCooper}; and (2) the higher probability for a fossil to branch within a wrong clade that has more morphological data available for living taxa than the right clade \citep{GuillermeCooper}.
-In this study, we assess the level of available morphological data in living mammals in order to highlight the two caveats described above.
-%In this study we investigate the amount of living mammal taxa with available morphological data to detect the two potential caveats described above that could lead to topological errors when building Total Evidence mammal phylogenetic trees.
-We collected living mammals morphological data from @256 phylogenetic matrices available online and measured the proportion of morphological data availability for each mammalian order.
+It is therefore crucial to understand the distribution of the available cladistic data for living taxa in a clade before using a Total Evidence approach because missing living taxa with cladistic data can lead to two topological artefacts: (1) the impossibility for a fossil to branch in the right clade if there is no cladistic data available in this clade \citep{GuillermeCooper}; and (2) the higher probability for a fossil to branch within a wrong clade that has more cladistic data available for living taxa than the right clade \citep{GuillermeCooper}.
+In this study, we assess the level of available cladistic data in living mammals in order to highlight the two caveats described above.
+%In this study we investigate the amount of living mammal taxa with available cladistic data to detect the two potential caveats described above that could lead to topological errors when building Total Evidence mammal phylogenetic trees.
+We collected living mammals cladistic data from @256 phylogenetic matrices available online and measured the proportion of cladistic data availability for each mammalian order.
 Additionally, in the mammalian orders where data was missing, we estimated the structure of the available data to detect if the available data was biased towards certain clades only or if it was randomly distributed using the Net Relatedness Index \citep{webb2002phylogenies}.
@@
 \subsection{Matrices search}
-To investigate the living mammals with available morphological data, we downloaded cladistic matrices from three major public databases: morphobank (\texttt{http://www.morphobank.org/}) \citep{morphobank}, Graeme Lloyd's website (\texttt{http://graemetlloyd.com/}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}).
-We downloaded all the matrices containing any fossil or living mammal taxa from these databases.
+To investigate the living mammals with available cladistic data, we downloaded cladistic matrices from three major public databases: morphobank (\texttt{http://www.morphobank.org/}) \citep{morphobank}, Graeme Lloyd's website (\texttt{http://graemetlloyd.com/}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}).
+We downloaded all the matrices containing any living or/and fossil mammal taxa from these databases.
 Additionally we ran a thorough Google Scholar search for matrices that might not have been uploaded on the previously mentioned data bases.
 We downloaded the additional cladistic matrices from the 20 first Google search results matching with our selected keywords and with any of the 35 taxonomic levels (mammals Orders, Infraclasses and Class; see supplementary materials for detailed description of the procedure).
-We downloaded 256 % @@@ Number of matrices
-matrices containing a total of 9411 % @@@ Number of living OTUs
-operational taxonomic units (OTUs) from the combination of both searches (public repositories and Google Scholar).
-The full list of matrices with their associated citations is available in the supplementary materials. % link
+In total, we downloaded 256 % @@@ Number of matrices
+matrices with 9411 % @@@ Number of living OTUs
+operational taxonomic units (OTUs) from the combination of both searches (public repositories and Google Scholar). % link
 
-We then transformed all the matrices to be in nexus format.
-We also standardised the taxonomic nomenclature by fixing invalid binomial inputs to match with the official taxonomic nomenclature rules (i.e.
+%We transformed all the matrices to be in nexus format.
+We standardised the taxonomic nomenclature by fixing invalid binomial inputs to match with the official taxonomic nomenclature rules (i.e.
 \textit{H.
 sapiens} was transformed in \textit{Homo sapiens}).
-We assigned each species as being either living or fossil using a taxonomic matching algorithm described as follow: (1) we designated as living OTU all the OTUs that where either present in \citep{FritzTree} or \citep{wilson2005mammal}; (2)we designated as fossil OTUs all the OTUs that where present in the Paleobiology database and (3) for the OTUs neither labelled as living or fossil we tried to decompose the OTUs name (i.e.
-\textit{Homo$\_$sapiens} became \textit{Homo} and \textit{sapiens} and tried to match the first element with any taxonomic level (Family, Genus, Species, etc.) from \citep{wilson2005mammal}.
+We assigned each species as being either living or fossil using a taxonomic matching algorithm described as follow: (1) we designated as living OTU all the OTUs that where either present in \citep{FritzTree} or \citep{wilson2005mammal}; (2) we designated as fossil OTUs all the OTUs that where present in the Paleobiology database and (3) for the OTUs neither labelled as living or fossil we tried to decompose the OTUs name (i.e.
+\textit{Homo$\_$sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with any taxonomic level (Family, Genus, Species, etc.) from \citep{wilson2005mammal}.
 % link
@@
 From all the 256 % @@@ Number of matrices
-matrices, we selected only the ones that displayed at least 100 morphological characters per OTU.
-This arbitrary threshold of a minimal number of morphological characters was chosen to be in adequacy with \citep{GuillermeCooper} and \citep{harrisonamong-character2014}.
-Also this threshold avoids biases towards small matrices that could be either not informative \citep{wagner2000} (e.g.
-too few characters not allowing character overlap among matrices) or made of non-applicable characters \citep{Brazeau2011} (e.g.
-antlers which are sexual dimorphic characters proper to a specific clade).
+matrices, we selected only the ones that coded at least 100 cladistic characters per OTU.
+This arbitrary threshold of a minimal number of cladistic characters was chosen to be in adequacy with \citep{GuillermeCooper} and \citep{harrisonamong-character2014}.
+Also this threshold avoids biases towards small matrices that could be either non-informative \citep{wagner2000} (e.g.
+too few characters not allowing character overlap among matrices) or that contained non-applicable characters \citep{Brazeau2011} (e.g.
+antlers ramification which is a only applicable to cervidae).
 
 \subsubsection{Data availability}
-To assess the data availability per mammal order, we calculated the percentage of OTUs with morphological data for three different taxonomic levels: Family, Genus and Species.
-We highlighted all the orders containing less than 25\% of living taxa with morphological data because their amount of missing data ($>$75\%) was higher than in \citep{GuillermeCooper} and therefore highly probable of having wrong topology due to the missing data.
+To assess the data availability per mammal order, we calculated the percentage of OTUs with cladistic data for three different taxonomic levels: Family, Genus and Species.
+We highlighted all the orders containing less than 25\% of living taxa with cladistic data because their amount of missing data ($>$75\%) was higher than in \citep{GuillermeCooper} and therefore likely to induce topological errors.
 On the opposite, orders with up to 25\% of missing data (75\% of available data) were shown to have no significant effect on topology \citep{GuillermeCooper}.
-We therefore used this value as a threshold of "good" data sampling.
+We therefore used this value as a threshold of ``bad'' ($<$ 25\% available) and ``good'' ($>$ 75\% available) data coverage.
 
 \subsubsection{Available data structure}
-For the orders with no morphological data for all its OTUs at the three different taxonomic levels (Family, Genus and Species), we investigated the structure of the available data to test if it was either (i) randomly distributed, (ii) over-dispersed or (iii) clustered.
-To measure the structure of the available data we used classic community structure metric from the \texttt{picante} R package \citep{picante}.
-We compared the structure of the available data for each order to the structure of a potentially fully sampled order (i.e.
-only the OTUs with available morphological data \textit{versus} all the OTUs).
-For each orders and taxonomic levels that presented OTUs with no available morphological data, we calculated the Net Relatedness Index (NRI) which quantifies the overall distribution of the data with negative values showing more dispersed data and positive values more clustered data than expected by the null model (random) \citep{webb2002phylogenies}.
-We choose to present only the NRI values because they have been shown to be slightly less sensitive to the structure of the phylogeny (i.e.
-branch length and topology) \citep{NRI,journal.pone.0004390} but we also calculated the two other common phylogenetic structure indices: Faith's Phylogenetic Distance (PD) \citep{Faith19921} and the Nearest Taxon Index (NTI) \citep{webb2002phylogenies}.
+For the orders that contained OTUs with no cladistic data at any of the three different taxonomic levels (Family, Genus and Species), we investigated whether the structure of the available data was either (i) randomly distributed, (ii) over-dispersed or (iii) clustered by using cNet Relatedness Index (NRI) metric \citep{webb2002phylogenies} using the \texttt{picante} R package \citep{picante}.
+The NRI quantifies the overall distribution of the observed data compared to a model where all the data is sampled for each OTU.
+Negative NRI values are showing more over-dispersed than expected by the null model (random) and positive values are showing more clustered data \citep{webb2002phylogenies}.
+We choose NRI metric because is has been shown to be slightly less sensitive to the structure of the phylogeny (i.e.
+branch length and topology) \citep{NRI,journal.pone.0004390} but we also calculated two other common phylogenetic structure indices: Faith's Phylogenetic Distance (PD) \citep{Faith19921} and the Nearest Taxon Index (NTI) \citep{webb2002phylogenies}.
 Both metrics are available in the Supplementary results.
@@
 %\subsubsection{Data improvement strategy}
-%Phylotargeting
-%This part as not been done at all! I was looking for a non-GUI version of Arnold and Nunn's software but there is none (Arnold confirmed by email).
-Therefore because I was focusing on the part above and to make it repeatable I didn't did that part yet.
-However, I still planing on doing it! I just need to find a bit of time to write some code to make it non-GUI (and to be able to use big trees).
-
-All the following procedure is repeatable and available on GitHub.
-%Or what is the right way to show it's reproducible?
+%Too long for this paper and actually still a lot of work to do (transform the code of phylotargeting + come up with a valid strategy)
 
@@
 \subsection{Data availability}
-We extracted 1422 living mammal OTUs from the 256 matrices with a minimum of 6 characters and a maximum of 4541 per OTU.
-After removing all the matrices with less than 100 morphological characters, the number of extracted living mammals OTUs was down to 815.
-11/28 orders have less than 25\% of taxa with morphological data at a species level and 24/28 orders have less than 75\% taxa with available morphological data.
-At the Genus level however only 3/28 orders have less than 25\% of taxa with morphological data and 16/28 have less than 75\%.
-Finally, at the family level no order has less than 25\% taxa with available morphological data and only 5/28 have less than 75\% (table \ref{Table_morpho_taxa_proportion}.
+We extracted 1422 living mammal OTUs from the 256 matrices with a minimum of 6 and a maximum of 4541 coded cladistic characters per OTU.
+After removing all the matrices with less than 100 cladistic characters, the number of extracted living mammals OTUs was down to 815.
+11/28 orders have less than 25\% of OTUs with cladistic data at a species level and 24/28 orders have less than 75\% taxa with available cladistic data.
+At the Genus level however only 3/28 orders have less than 25\% of taxa with cladistic data and 16/28 have less than 75\%.
+Finally, at the family level no order has less than 25\% taxa with available cladistic data and only 5/28 have less than 75\% (table \ref{Table_morpho_taxa_proportion}).
 
@@
 \subsection{Available data structure}
-Among the orders containing at least one OTU with no available morphological data, only two orders are significantly clustered: Carnivora and Chiroptera both at the species and the genus level (table~\ref{Table_data_structure}.
+Among the orders containing OTUs with no available cladistic data, only two orders are significantly clustered: Carnivora and Chiroptera both at the species and the genus level (table~\ref{Table_data_structure}).
 
@@
 
-Two contrasted results are shown on figures \ref{Figure_example_coverageA} and \ref{Figure_example_coverageB} with randomly distributed data in Cetartiodactyla (figure \ref{Figure_example_coverageA} and clustered available data in Carnivora (mainly Canidae; figure \ref{Figure_example_coverageB}.
+Two contrasted results are shown on figure \ref{Figure_example_coverage} with randomly distributed data in Cetartiodactyla (figure \ref{Figure_example_coverage}-A) and clustered available data in Carnivora (mainly Canidae; figure \ref{Figure_example_coverage}-B).
 
@@
 \centering
-    \includegraphics[width=1\textwidth]{example_coverageA.pdf}
-\caption{Distribution of available morphological data across Cetartiodactyla.
-Edges are colored in grey when no morphological data is available or in red when data is available.}
-\label{Figure_example_coverageA}
-\end{figure}
-
-\begin{figure}[!htbp]
-\centering
-    \includegraphics[width=1\textwidth]{example_coverageB.pdf}
-\caption{Distribution of available morphological data across Carnivora.
-Edges are colored in grey when no morphological data is available or in red when data is available.}
-\label{Figure_example_coverageB}
+    \includegraphics[width=1\textwidth]{example_coverage.pdf}
+\caption{Distribution of available cladistic data across two different clades (A: Cetartiodactyla; B: Carnivora).
+Edges are colored in grey when no cladistic data is available or in red when data is available.}
+\label{Figure_example_coverage}
 \end{figure}
@@
 \section{Discussion}
-Our results show that even though relations among living mammals are well known in a molecular framework \citep[e.g.][]{FritzTree,meredithimpacts2011,May-Collado-PeerJ} only little morphological data is available.
+Our results show that even though relations among living mammals are well known in a molecular framework \citep[e.g.][]{FritzTree,meredithimpacts2011,May-Collado-PeerJ} only little cladistic data is available.
 This is primarily due to the fact that that morphological studies are mainly focusing on fossil taxa rather than living taxa \citep[e.g.][]{O'Leary08022013,ni2013oldest}.
@@
 %\subsection{Data availability}
-The amount of living taxa with no available morphological data was surprisingly high at the species level: 24/28 orders have less than 75\% taxa with available morphological data (and two of the 28 orders are monospecific).
+The amount of living taxa with no available cladistic data was surprisingly high at the species level: 24/28 orders have less than 75\% taxa with available cladistic data (and two of the 28 orders are monospecific).
 This threshold of 75\% available data is the maximum amount of possible missing data (25\%) before having a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.
@@
 At the species level within most orders of mammals, it is therefore likely to see topological artefacts for the placement of fossil taxa.
-However, the data availability seems to be less an issue at higher taxonomic level (i.e.
-the Genus and the Family level).
+However, the data availability seems to be less an issue at higher taxonomic level (i.e. the Genus and the Family level).
 This point is important to consider from a practical point of view because of the slight discrepancy between neontological and palaeontological species.
@@
 However, in the third scenario, when the data is clustered we expect two major biases to occur: (1) first, the fossil will not be able to branch within a clade containing no data (e.g.
-Herpestidae in Carnivora; figure \ref{Figure_missing_data_problem} and \ref{Figure_example_coverageB}, and (2) second, the fossil will have a higher probability, at random, to branch within the clade containing most of the available data (e.g.
-a Carnivora fossil will have more chance to branch, randomly, to the Canidae clade than any other clade in Carnivora; figure \ref{Figure_example_coverageB}.
+Herpestidae in Carnivora; figure \ref{Figure_missing_data_problem} and \ref{Figure_example_coverage}-B, and (2) second, the fossil will have a higher probability, at random, to branch within the clade containing most of the available data (e.g.
+a Carnivora fossil will have more chance to branch, randomly, to the Canidae clade than any other clade in Carnivora; figure \ref{Figure_example_coverage}-B.
 
-%For example in Carnivora, their is no available data for Herpestidae, therefore any fossil Herpestidae will not be able to be correctly placed in the phylogeny (figure ~\ref{Figure_example_coverageB}.
-On the other hand, any Carnivora fossil with uncertain phylogenetic affinities (\textit{Incertae sedis} will have a higher probability of branching by chance within the Canidae because they are oversampled within the Carnivora (figure ~\ref{Figure_example_coverageB}.
+%For example in Carnivora, their is no available data for Herpestidae, therefore any fossil Herpestidae will not be able to be correctly placed in the phylogeny (figure ~\ref{Figure_example_coverage}-B.
+On the other hand, any Carnivora fossil with uncertain phylogenetic affinities (\textit{Incertae sedis} will have a higher probability of branching by chance within the Canidae because they are oversampled within the Carnivora (figure ~\ref{Figure_example_coverage}-B.
 
@@
 In this study, we treated all morphological matrices to be equal in a similar way that molecular matrices are: for example if a matrix A contains 100 characters for taxa X and Y, and a matrix B contains 50 characters for taxa X and Z, we assumed that both matrices can be combined in a supermatrix way leading to a matrix containing 150 characters for taxon X, 100 for taxon Y and 50 for taxon Z.
-However, it is clear that morphological data cannot be treated that way \citep{Brazeau2011}.
+However, it is clear that cladistic data cannot be treated that way \citep{Brazeau2011}.
 In fact, for the same taxon, some characters might overlap (e.g.

### 2015-05-19 Manuscript/Missing_mammals-draft.tex
Updated discussion

@@
 % 2500 words max (includes the title page, abstract, references, acknowledgements and figure/table legends)
-% current version is above 4500. I think a big cut down can be done on the references.
+% current version is around 3700. I think a big cut down can be done on the references.
 % We allow a maximum of 4 displays, only 2 of which can be figures.
@@
 Our results show that even though relations among living mammals are well known in a molecular framework \citep[e.g.][]{FritzTree,meredithimpacts2011,May-Collado-PeerJ} only little cladistic data is available.
-This is primarily due to the fact that that morphological studies are mainly focusing on fossil taxa rather than living taxa \citep[e.g.][]{O'Leary08022013,ni2013oldest}.
+This is primarily due to the fact that that cladistic studies are mainly focusing on fossil taxa rather than living taxa \citep[e.g.][]{O'Leary08022013,ni2013oldest}.
 However, data availability varies in its structure as well as across taxonomic levels.
-Higher taxonomic levels are always better sampled than lower ones (table \ref{Table_morpho_taxa_proportion} and within these taxonomic levels, the structure of available data is mostly random apart for two orders (table \ref{Table_data_structure}.
+Higher taxonomic levels are always better sampled than lower ones (table \ref{Table_morpho_taxa_proportion}) and within these taxonomic levels, the structure of available data is mostly random apart for two orders (table \ref{Table_data_structure})b.
 
 %\subsection{Data availability}
-The amount of living taxa with no available cladistic data was surprisingly high at the species level: 24/28 orders have less than 75\% taxa with available cladistic data (and two of the 28 orders are monospecific).
+The amount of living taxa with no available cladistic data was surprisingly high at the species level: 24/28 orders have less than 75\% taxa with available cladistic data (and two of the 28 orders are monospecific!).
 This threshold of 75\% available data is the maximum amount of possible missing data (25\%) before having a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.
-Beyond this threshold, there is a noteworthy displacement of the wildcard taxa (\textit{sensu} \citep{kearneyfragmentary2002} as well as a important decrease in the conservation of clades \citep{GuillermeCooper}.
+Beyond this threshold, there is a noteworthy displacement of the wildcard taxa (\textit{sensu} \citep{kearneyfragmentary2002}) as well as a important decrease in the conservation of clades \citep{GuillermeCooper}.
 At the species level within most orders of mammals, it is therefore likely to see topological artefacts for the placement of fossil taxa.
 However, the data availability seems to be less an issue at higher taxonomic level (i.e. the Genus and the Family level).
-This point is important to consider from a practical point of view because of the slight discrepancy between neontological and palaeontological species.
-While neontological species are described on accurately measured morphology, genetic distance, spacial distribution and even behaviour \citep[e.g.][]{kellymolecular2014}; palaeontological data can be based only on morphological, spatial and temporal data \citep[e.g.][]{ni2013oldest}.
-Because of this difference between neontological species (\textit{sensu} reproductive isolates) and paleontological species (\textit{sensu} morpho-species), most palaeontological studies are using the Genus level for their smallest taxonomic OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
+This point is important to consider from a practical point of view because of the slight discrepancy between the neontological and the palaeontological concept of species.
+While neontological species are described on accurately measured morphology, genetic distance, spacial distribution and even behaviour \citep[e.g.][]{kellymolecular2014}, palaeontological data can be based only on morphological, spatial and temporal data \citep[e.g.][]{ni2013oldest}.
+Because of this difference between neontological species (\textit{sensu} reproductive isolates) and paleontological species (\textit{sensu} morpho-species %or "morphological clusters"?
+), most palaeontological studies are using the Genus level for their smallest OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
 In this frame of palaeontological data usage, the data availability at the Genus level in living mammals should be our priority concern.
@@
 %\subsection{Available data structure}
-Regardless the low level of available data, it is encouraging that for most orders (except Carnivora and Chiroptera), it is randomly distributed across the phylogeny.
-Therefore, it is likely that the addition of fossil taxa within these orders will not be biased by oversampling but will probably branch near the taxa that has the most similar available data (as expected).
-When only few data is available, the ideal scenario will be that the data is over-dispersed (i.e.
+Regardless the low level of available data, it is encouraging that for most orders (except Carnivora and Chiroptera), data is randomly distributed across the phylogeny.
+%Therefore, it is likely that the addition of fossil taxa within these orders will not be biased by oversampling but will probably branch near the taxa that has the most similar available data (as expected).
+There are three different scenarii, regarding the distribution of the available data, for predicting how fossils can branch in relation to living taxa.
+When only few data is available, the ideal scenario will be that it is over-dispersed (i.e.
 that there is data in at least every sub-clade) in order to maximize the possibilities of the fossil to branch in the right clade.
@@
 However, in the third scenario, when the data is clustered we expect two major biases to occur: (1) first, the fossil will not be able to branch within a clade containing no data (e.g.
-Herpestidae in Carnivora; figure \ref{Figure_missing_data_problem} and \ref{Figure_example_coverage}-B, and (2) second, the fossil will have a higher probability, at random, to branch within the clade containing most of the available data (e.g.
-a Carnivora fossil will have more chance to branch, randomly, to the Canidae clade than any other clade in Carnivora; figure \ref{Figure_example_coverage}-B.
+Herpestidae in Carnivora; figure \ref{Figure_missing_data_problem} and \ref{Figure_example_coverage}-B), and (2) second, the fossil will have a higher probability, at random, to branch within the clade containing most of the available data (e.g.
+a Carnivora fossil will have more chance to branch, randomly, to the Canidae clade than any other clade in Carnivora; figure \ref{Figure_example_coverage}-B).
+This means that fossils with uncertain phylogentic affinities (\textit{Incertae sedis}) will have a higher probability of branching by chance within the most sampled clade.
 
 %For example in Carnivora, their is no available data for Herpestidae, therefore any fossil Herpestidae will not be able to be correctly placed in the phylogeny (figure ~\ref{Figure_example_coverage}-B.
-On the other hand, any Carnivora fossil with uncertain phylogenetic affinities (\textit{Incertae sedis} will have a higher probability of branching by chance within the Canidae because they are oversampled within the Carnivora (figure ~\ref{Figure_example_coverage}-B.
+%On the other hand, any Carnivora fossil with uncertain phylogenetic affinities (\textit{Incertae sedis} will have a higher probability of branching by chance within the Canidae because they are oversampled within the Carnivora (figure ~\ref{Figure_example_coverage}-B.
 
 %Caveats
-In this study, we treated all morphological matrices to be equal in a similar way that molecular matrices are: for example if a matrix A contains 100 characters for taxa X and Y, and a matrix B contains 50 characters for taxa X and Z, we assumed that both matrices can be combined in a supermatrix way leading to a matrix containing 150 characters for taxon X, 100 for taxon Y and 50 for taxon Z.
+In this study, we treated all cladistic matrices to be equal in a similar way that molecular matrices are: for example if a matrix A contains 100 characters for taxa X and Y, and a matrix B contains 50 characters for taxa X and Z, we assumed that both matrices can be combined in a supermatrix containing 150 characters for taxon X, 100 for taxon Y and 50 for taxon Z.
 However, it is clear that cladistic data cannot be treated that way \citep{Brazeau2011}.
 In fact, for the same taxon, some characters might overlap (e.g.
-matrix A has a character coding for the shape of a particular morphological feature and matrix B as a first character coding for the presence of this morphological feature and a second one coding for its shape; in this case, these three characters are compound characters \citep{Brazeau2011}.
-However, in reasonably sized matrices ($>$ 100 characters \citep{GuillermeCooper,harrisonamong-character2014} it is more likely that a number of characters are consistently conserved among the different matrices \citep[e.g.][]{ross1998phylogenetic,seiffert2003fossil,marivaux2005anthropoid,seiffert2005basal,bloch2007new,kay2008anatomy,silcox2008biogeographic,seiffert2009convergent,tabuce2009anthropoid,boyer2010astragalar,seiffert2010fossil,marivaux2013djebelemur,ni2013oldest}.
-A conservative approach to avoid compound characters could be to select only the latest matrix for each taxon.
+matrix A has a character coding for the shape of a particular morphological feature and matrix B as a first character coding for the presence of this morphological feature and a second one coding for its shape; in this case, these three characters are compound characters \citep{Brazeau2011}).
+However, in reasonably sized matrices ($>$ 100 characters \citep{GuillermeCooper,harrisonamong-character2014} it is more likely that a number of characters are consistently conserved among the different matrices (e.g.
+within primates literature, the character \textit{p7} - the size of the $4^{th}$ lower premolar paraconid - is consistent for more than 15 years \citep{ross1998phylogenetic,seiffert2003fossil,marivaux2005anthropoid,seiffert2005basal,bloch2007new,kay2008anatomy,silcox2008biogeographic,seiffert2009convergent,tabuce2009anthropoid,boyer2010astragalar,seiffert2010fossil,marivaux2013djebelemur,ni2013oldest}).
+A conservative approach to avoid compound characters could be to select only the most recent matrix for each group.
 
 %Let's code some data
-Following the recommendations in \citep{GuillermeCooper}, one should code morphological characters for a maximum of living species in order to improve the topology of Total Evidence trees containing both living and fossil taxa.
+Following the recommendations in \citep{GuillermeCooper}, one should code cladistic characters for a maximum of living species in order to improve the topology of Total Evidence trees containing both living and fossil taxa.
 Since the data for living mammals is usually easily available in world wide vast natural history collections, we propose that an increased effort in coding morphological characters from living species should be done via engaging in collaborative data collection projects through web portals such as \textit{morphobank} \citep{morphobank}.
@@
 \section{Acknowledgements}
-Nick Matzke, April Wright, David Bapst and Graeme Lloyd.
+Nick Matzke, April Wright, David Bapst and Graeme Lloyd. %Or whatever order.
 \section{Funding statement}

### 2015-05-19 Manuscript/Missing_mammals-draft.tex

@@
 
-%Pagination style and stuff
+%Pagination style and stuff % NC: Note that these are all syst biol specific.
 \linespread{2}
@@
 %Or: Missing cladistic data in living mammals - but I prefer the first one emphasizing on the availability of it. The data is not missing per se, it's just not coded!
+% NC: sometimes the better title is one that answers the question rather than giving the topic. e.g., "Morphological data is lacking for living species" or similar
 
@@
 
-\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$$^*$, and Natalie Cooper$^1$$^,$$^2$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$$^,$$^*$ and Natalie Cooper$^1$$^,$$^2$$^,$$^3$}\\
 \noindent {\small \it 
 $^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\
-$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.}\\
+$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
+$^3$Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.}\\
 \end{center}
 \medskip
-\noindent{*\bf Corresponding author.} \textit{Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.}\\
+\noindent{*\bf Corresponding author.} \textit{Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.}\\ % NC: Might need to reconsider the email here as it won't be your email anymore when you leave.
+Or do you get to keep it forever?
 \vspace{1in}
@@
 %Line numbering
-%\modulolinenumbers[1]
-%\linenumbers
+\modulolinenumbers[1]
+\linenumbers
 
@@
 %---------------------------------------------
+% NC: I've commented this out so I don't have to keep scrolling past in in the PDF
+%\newpage
+%\begin{abstract}
+%Studying changes in global biodiversity through time and space is essential.
+%For that we need methods to combine both palaeontological and neontological data.
+%One promising method, the Total Evidence method, allows such thing but needs a lot of data.
+%Especially cladistic data from living taxa to allow the fossil taxa to accurately branch in the trees.
+%Despite two centuries of morphological studies on living taxa, scientists using and generating such data mainly focus on palaeontological data.
+%Therefore, even in well known groups such as mammal, there is a huge gap in our knowledge of cladistic data for living mammals.
+%In this study, using phylogenetic community structure methods, we quantify the availability of data in each mammalian order.
+%And maybe at the end of the paper we propose some discussion on how to improve all that (go in museums!).
+%\end{abstract}
 
-\newpage
-\begin{abstract}
-Studying changes in global biodiversity through time and space is essential.
-For that we need methods to combine both palaeontological and neontological data.
-One promising method, the Total Evidence method, allows such thing but needs a lot of data.
-Especially cladistic data from living taxa to allow the fossil taxa to accurately branch in the trees.
-Despite two centuries of morphological studies on living taxa, scientists using and generating such data mainly focus on palaeontological data.
-Therefore, even in well known groups such as mammal, there is a huge gap in our knowledge of cladistic data for living mammals.
-In this study, using phylogenetic community structure methods, we quantify the availability of data in each mammalian order.
-And maybe at the end of the paper we propose some discussion on how to improve all that (go in museums!).
-\end{abstract}
+%\noindent ()\\
 
-\noindent ()\\
-
-\vspace{1.5in}
-
-\newpage
+%\vspace{1.5in}
 
@@
 
+%NC: I've changed cladistic to morphological throughout but you might want to change it back...
+
+\newpage 
 \section{Introduction}
-Studying both living and fossil taxa together in essential to fully understand macrovelutionary patterns and processes and is becoming increasingly common among evolutionary biologists \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013,Wood01032013}.
-Combining the global biodiversity by including both living and fossil in such studies allows, for example, to improve accuracy of the timing of diversification events \citep[e.g.][]{ronquista2012}, to better understand relationships among lineages \citep[e.g.][]{beckancient2014} or to infer biogeographical patterns through time \citep[e.g.][]{Meseguer01032015}.
-%And add STD paper? Or that's too much self cites?
-In order to perform such analysis, one must efficiently combine living and fossil taxa data in macroevolutionary models.
-One trending method, called the Total Evidence method \citep{eernissetaxonomic1993,ronquista2012}, allows to combine molecular data from living taxa and cladistic data from both living and fossil taxa in a supermatrix \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014,Meseguer01032015}.
-This method not only allows to use all the available data but also allows to treat fossil taxa as tips rather than nodes \textit{via} integrative phylogenetic inference methods such as tip-dating \citep{ronquista2012,Drummond01082012,Wood01032013,BEASTmaster}.
+There is an increasing consensus among evolutionary biologists that studying both living and fossil taxa is essential for fully understanding macroevolutionary patterns and processes \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013,Wood01032013}.
+For example, including both living and fossil taxa in evolutionary studies can improve the accuracy of timing diversification events \citep[e.g.][]{ronquista2012}, our understanding of relationships among lineages \citep[e.g.][]{beckancient2014}, and our ability to infer biogeographical patterns through time \citep[e.g.][]{Meseguer01032015}.
+To perform such analyses it is necessary to combine living and fossil taxa in phylogenetic trees.
+One increasingly popular method, the Total Evidence method \citep{eernissetaxonomic1993,ronquista2012}, combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014,Meseguer01032015}, producing a phylogeny with living and fossil taxa at the tips.
+These phylogenies can be dated using methods such as tip-dating \citep{ronquista2012,Drummond01082012,Wood01032013,BEASTmaster} and incorporated into macroevolutionary studies (e.g. ).
 
-However, the Total Evidence method requires, by definition, a lot of data.
-One must collect both molecular data for living taxa as well as cladistic data for both living and fossil taxa, two types of data that require fairly different technical skills \citep[e.g.][]{meredithimpacts2011} \textit{vs.} \citep{O'Leary08022013}.
-Additionally, entire sections of this data can sometimes be difficult or impossible to collect for every taxa present in the analysis.
-For example, fossil have really rarely molecular data available and morphological characters are rarely collected for living taxa when molecular data is available \citep[e.g.][]{slaterphylogenetic2013,beckancient2014}.
-This difference between both data can lead to topological errors in phylogenetic inference \citep{GuillermeCooper}.
-In fact, the ability of the Total Evidence method to recover correct topology is expected to decrease when there is a low overlap between cladistic data from the living and the fossil taxa \citep{GuillermeCooper}.
-The effect of missing data is most important on topology when too few living taxa have available cladistic data \citep{GuillermeCooper}.
-For example, if there is no morphological available for any living taxa within an entire clade, it is impossible to link a fossil taxon to this clade because no cladistic data will overlap between the fossil (regardless the amount of data available for the fossil) and the living taxa (that have no cladistic data).
-This property of the Total Evidence method can rapidly lead to important topological incongruities because the fossil will only be able to branch to a clade of living taxa that contains cladistic data, even if in reality, the fossil taxon does not belong to that clade \citep{GuillermeCooper} (see figure ~\ref{Figure_missing_data_problem}).
+A downside of the Total Evidence method is that it requires a lot of data.
+One must collect molecular data for living taxa and morphological data for both living and fossil taxa; two types of data that require fairly different technical skills \citep[e.g.][]{meredithimpacts2011} \textit{vs.} \citep{O'Leary08022013}.
+% NC: Is that really what these papers show?
+They use differnt data but they don't talk about the technical skills!
+Might be better just to say something like (e.g., PCR/sequencing vs.
+anatomical description)
+Additionally, large chunks of this data can be difficult, or even impossible, to collect for every taxon present in the analysis.
+For example, fossils very rarely have molecular data and incomplete fossil preservation may restrict the amount of morphological data available.
+% NC: Worth metioning fossil preservation here briefly as it is also an issue?
+Additionally, since the molecular phylogenetics revolution, it has become less common to collect morphological characters for living taxa when molecular data is available \citep[e.g.][]{slaterphylogenetic2013,beckancient2014}.
+%NC: Do these references actually say this??? Sounds like that's what you are saying
+Unfortunately this missing data can lead to errors in phylogenetic inference; in fact, simulations show that the ability of the Total Evidence method to recover the correct phylogenetic topology decreases when there is a low overlap between morphological data in the living and fossil taxa \citep{GuillermeCooper}, regardless the overall amount of morphological data available for the fossils (or the amount of molecular data available for the living species).
+The effect of missing data on topology is greatest when living taxa have little morphological data.
+This is because (1) a fossil cannot branch in the correct clade if there is no overlapping morphological data in the clade; and (2) a fossil has a higher probability of branching within a clade with more morphological data available for living taxa, regardless of whether this is the correct clade or not \citep{GuillermeCooper}.
+This can lead to serious topological errors (see Figure ~\ref{Figure_missing_data_problem}).
+
+% NC: OK so I know we did a paper on this but you can't just cite it every sentence.
+Need to put in a few more refs to other people's work or just cite us once or twice.
+I've condensed it which should help
 
@@
     \includegraphics[width=1\textwidth]{MissingDataFigure.pdf}
-\caption{Example of topological errors due to missing cladistic data in living taxa.
-If a phylogeny contains two clades, for example Aves and Mammalia, with molecular data (in grey) for both but only cladistic data (in red) for Aves.
-If an additional Mammalia fossil (with no molecular data) is added to the phylogeny, it will erroneously branch with the Aves clade instead of the Mammalia one because no cladistic data will overlap between the fossil Mammalia and the living ones.}
+\caption{Example of topological errors due to missing morphological data in living taxa.
+The phylogeny contains two clades, Aves and Mammalia, with molecular data (grey) for both but only morphological data (red) for Aves.
+If a mammalian fossil species (with no molecular data) is added to the phylogeny, it will erroneously branch with the Aves instead of the Mammalia because no morphological data will overlap between the fossil mammals and the living ones.}
 \label{Figure_missing_data_problem}
@@
 %What we're doing
-It is therefore crucial to understand the distribution of the available cladistic data for living taxa in a clade before using a Total Evidence approach because missing living taxa with cladistic data can lead to two topological artefacts: (1) the impossibility for a fossil to branch in the right clade if there is no cladistic data available in this clade \citep{GuillermeCooper}; and (2) the higher probability for a fossil to branch within a wrong clade that has more cladistic data available for living taxa than the right clade \citep{GuillermeCooper}.
-In this study, we assess the level of available cladistic data in living mammals in order to highlight the two caveats described above.
-%In this study we investigate the amount of living mammal taxa with available cladistic data to detect the two potential caveats described above that could lead to topological errors when building Total Evidence mammal phylogenetic trees.
-We collected living mammals cladistic data from @256 phylogenetic matrices available online and measured the proportion of cladistic data availability for each mammalian order.
-Additionally, in the mammalian orders where data was missing, we estimated the structure of the available data to detect if the available data was biased towards certain clades only or if it was randomly distributed using the Net Relatedness Index \citep{webb2002phylogenies}.
+
+The issues above highlight that it is crucial to have sufficient morphological data for living taxa in a clade before using a Total Evidence approach.
+However, it is unclear how much morphological data for living taxa is actually available (i.e.
+already coded from specimens of living species and deposited in phylogenetic matrices), and how this data is distributed.
+Intuitively, most people assume this kind of data has already been collected, but our experiences suggest otherwise.
+To investigate this further, here we assess the amount of available morphological data for living mammals to determine whether sufficient data exists to build reliable Total Evidence phylogenies in this group.
+We collected cladistic data from @256 phylogenetic matrices available online and measured the proportion of cladistic data available for each mammalian order.
+Additionally, if the available data is randomly distributed across clades, the topological errors should be less extreme than if the data is biased towards particulars groups.
+Therefore, in the mammalian orders where data was missing, we determined whether the available cladistic data was phylogenetically overdispersed or clustered. 
+We find???
+% NC: Needs some summing up sentence or a we find sentence?
 

### 2015-05-19 Manuscript/Missing_mammals-draft.tex
Fixed intro (+ moved figure to supp)

@@
 
-\noindent{\Large \bf Morphological data availability in living mammals}
+\noindent{\Large \bf Morphological data is lacking for living mammals}
 %Or: Missing cladistic data in living mammals - but I prefer the first one emphasizing on the availability of it. The data is not missing per se, it's just not coded!
@@
 One increasingly popular method, the Total Evidence method \citep{eernissetaxonomic1993,ronquista2012}, combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014,Meseguer01032015}, producing a phylogeny with living and fossil taxa at the tips.
-These phylogenies can be dated using methods such as tip-dating \citep{ronquista2012,Drummond01082012,Wood01032013,BEASTmaster} and incorporated into macroevolutionary studies (e.g. ).
+These phylogenies can be dated using methods such as tip-dating \citep{ronquista2012,Drummond01082012,Wood01032013} and incorporated into macroevolutionary studies \citep[e.g.][]{ronquista2012,Wood01032013,slaterphylogenetic2013}.
 
 A downside of the Total Evidence method is that it requires a lot of data.
-One must collect molecular data for living taxa and morphological data for both living and fossil taxa; two types of data that require fairly different technical skills \citep[e.g.][]{meredithimpacts2011} \textit{vs.} \citep{O'Leary08022013}.
-% NC: Is that really what these papers show?
-They use differnt data but they don't talk about the technical skills!
-Might be better just to say something like (e.g., PCR/sequencing vs.
-anatomical description)
+One must collect molecular data for living taxa and morphological data for both living and fossil taxa; two types of data that require fairly different technical skills (e.g.
+molecular sequencing \textit{vs.} anatomical description).
 Additionally, large chunks of this data can be difficult, or even impossible, to collect for every taxon present in the analysis.
-For example, fossils very rarely have molecular data and incomplete fossil preservation may restrict the amount of morphological data available.
+For example, fossils very rarely have molecular data and incomplete fossil preservation may restrict the amount of morphological data available \citep{sansomfossilization2013}.
 % NC: Worth metioning fossil preservation here briefly as it is also an issue?
-Additionally, since the molecular phylogenetics revolution, it has become less common to collect morphological characters for living taxa when molecular data is available \citep[e.g.][]{slaterphylogenetic2013,beckancient2014}.
-%NC: Do these references actually say this??? Sounds like that's what you are saying
+Additionally, since the molecular phylogenetics revolution, it has become less common to collect morphological characters for living taxa when molecular data is available (e.g.
+in \cite{slaterphylogenetic2013}, only 13\% of the 169 living taxa have coded morphological data).
 Unfortunately this missing data can lead to errors in phylogenetic inference; in fact, simulations show that the ability of the Total Evidence method to recover the correct phylogenetic topology decreases when there is a low overlap between morphological data in the living and fossil taxa \citep{GuillermeCooper}, regardless the overall amount of morphological data available for the fossils (or the amount of molecular data available for the living species).
@@
 This is because (1) a fossil cannot branch in the correct clade if there is no overlapping morphological data in the clade; and (2) a fossil has a higher probability of branching within a clade with more morphological data available for living taxa, regardless of whether this is the correct clade or not \citep{GuillermeCooper}.
-This can lead to serious topological errors (see Figure ~\ref{Figure_missing_data_problem}).
-
-% NC: OK so I know we did a paper on this but you can't just cite it every sentence.
-Need to put in a few more refs to other people's work or just cite us once or twice.
-I've condensed it which should help
-
-\begin{figure}[!htbp]
-\centering
-    \includegraphics[width=1\textwidth]{MissingDataFigure.pdf}
-\caption{Example of topological errors due to missing morphological data in living taxa.
-The phylogeny contains two clades, Aves and Mammalia, with molecular data (grey) for both but only morphological data (red) for Aves.
-If a mammalian fossil species (with no molecular data) is added to the phylogeny, it will erroneously branch with the Aves instead of the Mammalia because no morphological data will overlap between the fossil mammals and the living ones.}
-\label{Figure_missing_data_problem}
-\end{figure}
-
-%What we're doing
+%This can lead to serious topological errors (see Figure ~\ref{Figure_missing_data_problem}).
 
@@
 However, it is unclear how much morphological data for living taxa is actually available (i.e.
-already coded from specimens of living species and deposited in phylogenetic matrices), and how this data is distributed.
-Intuitively, most people assume this kind of data has already been collected, but our experiences suggest otherwise.
-To investigate this further, here we assess the amount of available morphological data for living mammals to determine whether sufficient data exists to build reliable Total Evidence phylogenies in this group.
+already coded from living taxa specimens and deposited in phylogenetic matrices accessible online), and how this data is distributed.
+Intuitively, most people assume this kind of data has already been collected, but empirical data suggest otherwise (e.g. in \cite{ronquista2012,slaterphylogenetic2013,beckancient2014}).
+To investigate this further, we assess the amount of available morphological data for living mammals to determine whether sufficient data exists to build reliable Total Evidence phylogenies in this group.
 We collected cladistic data from @256 phylogenetic matrices available online and measured the proportion of cladistic data available for each mammalian order.
-Additionally, if the available data is randomly distributed across clades, the topological errors should be less extreme than if the data is biased towards particulars groups.
-Therefore, in the mammalian orders where data was missing, we determined whether the available cladistic data was phylogenetically overdispersed or clustered. 
-We find???
-% NC: Needs some summing up sentence or a we find sentence?
+%Additionally, if the available data is randomly distributed across clades, the topological errors should be less extreme than if the data is biased towards particulars groups.
+Additionally, because of the two topological caveats mentioned above, we determined whether the available cladistic data was phylogenetically overdispersed or clustered in the mammalian orders where data was missing.
+We find that available morphological data for living mammals is scarce but generally randomly distributed and therefore suggest that an effort be made in collecting and sharing this type of data to improve Total Evidence tip-dated phylogenies.
 

### 2015-05-19 Manuscript/Supplementary/Supplementary.tex
Fixed intro (+ moved figure to supp)

@@
 \noindent{\Large \bf Supplementary Material}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supplementary/MissingDataFigure.pdf}
+\caption{Example of topological errors due to missing morphological data in living taxa.
+The phylogeny contains two clades, Aves and Mammalia, with molecular data (grey) for both but only morphological data (red) for Aves.
+If a mammalian fossil species (with no molecular data) is added to the phylogeny, it will erroneously branch with the Aves instead of the Mammalia because no morphological data will overlap between the fossil mammals and the living ones.}
+\label{Figure_missing_data_problem}
+\end{figure}
 

### 2015-05-19 Manuscript/Missing_mammals-draft.tex
Edits to methods

--- 2015-05-19 Manuscript/Missing_mammals-draft.tex
@@
 %\newpage
-%\begin{abstract}
+%\begin{abstract} % 200 words, all keywords written in abstract if possible.
 %Studying changes in global biodiversity through time and space is essential.
@@
 
-\section{Material and Methods}
+\section{Materials and Methods}
 \subsection{Matrices search}
-To investigate the living mammals with available cladistic data, we downloaded cladistic matrices from three major public databases: morphobank (\texttt{http://www.morphobank.org/}) \citep{morphobank}, Graeme Lloyd's website (\texttt{http://graemetlloyd.com/}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}).
-We downloaded all the matrices containing any living or/and fossil mammal taxa from these databases.
-Additionally we ran a thorough Google Scholar search for matrices that might not have been uploaded on the previously mentioned data bases.
-We downloaded the additional cladistic matrices from the 20 first Google search results matching with our selected keywords and with any of the 35 taxonomic levels (mammals Orders, Infraclasses and Class; see supplementary materials for detailed description of the procedure).
+We downloaded all cladistic matrices containing any living or/and fossil mammal taxa from three major public databases: morphobank (\texttt{http://www.morphobank.org/}) \citep{morphobank}, Graeme Lloyd's website (\texttt{http://graemetlloyd.com/}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}).
+We also performed a systematic Google Scholar search for matrices that might not have been uploaded on these databases (see see Supplementary Materials for a detailed description of the search procedure).
+%We downloaded the additional cladistic matrices from the 20 first search results matching with our selected keywords and with any of the 35 taxonomic levels (mammals Orders, Infraclasses and Class; see supplementary materials for detailed description of the procedure).
+% NC: Can possibly delete this and stick in supplement
 In total, we downloaded 256 % @@@ Number of matrices
-matrices with 9411 % @@@ Number of living OTUs
-operational taxonomic units (OTUs) from the combination of both searches (public repositories and Google Scholar). % link
+matrices containing 9411 % @@@ Number of living OTUs
+operational taxonomic units (OTUs). % link
+% NC: Might need to explain what you mean by OTUs, I'm assuming you mean genera or species etc.
+% NC: do you mean 9411 different OTUs or total OTUs? Might be good to have a number of unique OTUs as well.
 
 %We transformed all the matrices to be in nexus format.
-We standardised the taxonomic nomenclature by fixing invalid binomial inputs to match with the official taxonomic nomenclature rules (i.e.
-\textit{H.
-sapiens} was transformed in \textit{Homo sapiens}).
-We assigned each species as being either living or fossil using a taxonomic matching algorithm described as follow: (1) we designated as living OTU all the OTUs that where either present in \citep{FritzTree} or \citep{wilson2005mammal}; (2) we designated as fossil OTUs all the OTUs that where present in the Paleobiology database and (3) for the OTUs neither labelled as living or fossil we tried to decompose the OTUs name (i.e.
+We standardised the taxonomy of our OTUs by fixing invalid binomial inputs to match with standard taxonomic nomenclature (e.g., \textit{H. sapiens} was transformed to \textit{Homo sapiens}).
+We assigned each species as being either ``living'' or ``fossil'' as follows: (1) we designated as ``living'' all the OTUs that where either present in \citep{FritzTree} or \citep{wilson2005mammal}; (2) we designated as ``fossil'' all the OTUs that were present in the Paleobiology database (\texttt{https://paleodb.org} %NC: Check website
+); and (3) for the OTUs neither labeled as living or fossil we tried to decompose the OTUs name (i.e.
 \textit{Homo$\_$sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with any taxonomic level (Family, Genus, Species, etc.) from \citep{wilson2005mammal}.
@@
 The matching OTUs where labelled as living and the ones still not matching where ignored and labelled as non-applicable (NA; see supplementary material for more details on the taxonomic matching algorithm).
+% NC: Wonder if more can go in supp mat
 
@@
 All data is available and reproducible on GitHub.
-\section{Authors’ contributions statement}
-Conceived and designed the experiments: TG NC. Performed the experiments: TG. Analyzed the data: TG. Wrote the paper: TG NC.
+\section{Authors' contributions statement}
+Conceived and designed the experiments: TG NC.
+Performed the experiments: TG.
+Analyzed the data: TG.
+Wrote the paper: TG NC.
+% NC: Write this in the proper proceedings style.
+All authors gave final approval for publication.
 \section{Acknowledgements}
-Nick Matzke, April Wright, David Bapst and Graeme Lloyd. %Or whatever order.
+Nick Matzke, April Wright, David Bapst and Graeme Lloyd. %Or whatever order. % NC: Alphabetical?
 \section{Funding statement}

### 2015-05-20 Manuscript/Missing_mammals-draft.tex
Methods edit

@@
 \section{Materials and Methods}
-\subsection{Matrices search}
-We downloaded all cladistic matrices containing any living or/and fossil mammal taxa from three major public databases: morphobank (\texttt{http://www.morphobank.org/}) \citep{morphobank}, Graeme Lloyd's website (\texttt{http://graemetlloyd.com/}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}).
-We also performed a systematic Google Scholar search for matrices that might not have been uploaded on these databases (see see Supplementary Materials for a detailed description of the search procedure).
+\subsection{Data collection and standardisation}
+We downloaded all cladistic matrices containing any living or/and fossil mammal taxa from three major public databases: Morphobank (\texttt{http://www.morphobank.org/}) \citep{morphobank}, Graeme Lloyd's website (\texttt{http://graemetlloyd.com/}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}).
+% NC: give link to actual repo and actual bit of graemes website these things are found on?
+We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see see Supplementary Materials for a detailed description of the search procedure).
 %We downloaded the additional cladistic matrices from the 20 first search results matching with our selected keywords and with any of the 35 taxonomic levels (mammals Orders, Infraclasses and Class; see supplementary materials for detailed description of the procedure).
@@
 % NC: Might need to explain what you mean by OTUs, I'm assuming you mean genera or species etc.
+We refer to OTUs instead of species because?
 % NC: do you mean 9411 different OTUs or total OTUs? Might be good to have a number of unique OTUs as well.
+
+% NC; probably need some comment to say this is collecting all the data!
 
@@
 We standardised the taxonomy of our OTUs by fixing invalid binomial inputs to match with standard taxonomic nomenclature (e.g., \textit{H. sapiens} was transformed to \textit{Homo sapiens}).
-We assigned each species as being either ``living'' or ``fossil'' as follows: (1) we designated as ``living'' all the OTUs that where either present in \citep{FritzTree} or \citep{wilson2005mammal}; (2) we designated as ``fossil'' all the OTUs that were present in the Paleobiology database (\texttt{https://paleodb.org} %NC: Check website
-); and (3) for the OTUs neither labeled as living or fossil we tried to decompose the OTUs name (i.e.
-\textit{Homo$\_$sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with any taxonomic level (Family, Genus, Species, etc.) from \citep{wilson2005mammal}.
-% link
-The matching OTUs where labelled as living and the ones still not matching where ignored and labelled as non-applicable (NA; see supplementary material for more details on the taxonomic matching algorithm).
-% NC: Wonder if more can go in supp mat
+We designated as ``living'' all the OTUs that where either present in the phylogeny of Fritz \textit{et al}.
+\citep{FritzTree} or the taxonomy of Wilson \& Reeder \citep{wilson2005mammal}, and designated as ``fossil'' all the OTUs that were present in the Paleobiology database (\texttt{https://paleodb.org} %NC: Check website
+For OTUs that did not appear in these three sources we first decomposed the name (i.e.
+\textit{Homo$\_$sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (Family, Genus etc.).
+Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; see Supplementary Material for more details).
 
-\subsection{Data availability analysis}
-\subsubsection{Number of characters threshold}
-From all the 256 % @@@ Number of matrices
-matrices, we selected only the ones that coded at least 100 cladistic characters per OTU.
-This arbitrary threshold of a minimal number of cladistic characters was chosen to be in adequacy with \citep{GuillermeCooper} and \citep{harrisonamong-character2014}.
-Also this threshold avoids biases towards small matrices that could be either non-informative \citep{wagner2000} (e.g.
-too few characters not allowing character overlap among matrices) or that contained non-applicable characters \citep{Brazeau2011} (e.g.
-antlers ramification which is a only applicable to cervidae).
+%\subsection{Data availability analysis}
+%\subsubsection{Number of characters threshold}
+% NC: I'm not sure why you did this. You might need to explain more - let's discuss. Otherwise it looks like we are deliberately making the data coverage look worse by omitting data.
+Next we selected only matrices containing at least 100 cladistic characters per OTU.
+This arbitrary threshold was chosen to correspond with the number of characters used in \citep{GuillermeCooper} and \citep{harrisonamong-character2014}.
+Additionally, this threshold avoids biases towards small matrices that are often non-informative, because they contain too few characters so there is no character overlap among matrices \citep{wagner2000}, or contain characters that are highly specific and thus non-applicable across large clades \citep{Brazeau2011} (e.g.
+``antler ramifications'' is a character than is only applicable to Cervidae not all mammals).
 
-\subsubsection{Data availability}
-To assess the data availability per mammal order, we calculated the percentage of OTUs with cladistic data for three different taxonomic levels: Family, Genus and Species.
-We highlighted all the orders containing less than 25\% of living taxa with cladistic data because their amount of missing data ($>$75\%) was higher than in \citep{GuillermeCooper} and therefore likely to induce topological errors.
-On the opposite, orders with up to 25\% of missing data (75\% of available data) were shown to have no significant effect on topology \citep{GuillermeCooper}.
-We therefore used this value as a threshold of ``bad'' ($<$ 25\% available) and ``good'' ($>$ 75\% available) data coverage.
+% NC: Some summing up statement about how many matrices were actually usuable, how many OTUs, how many unique OTUs etc.
 
-\subsubsection{Available data structure}
-For the orders that contained OTUs with no cladistic data at any of the three different taxonomic levels (Family, Genus and Species), we investigated whether the structure of the available data was either (i) randomly distributed, (ii) over-dispersed or (iii) clustered by using cNet Relatedness Index (NRI) metric \citep{webb2002phylogenies} using the \texttt{picante} R package \citep{picante}.
-The NRI quantifies the overall distribution of the observed data compared to a model where all the data is sampled for each OTU.
-Negative NRI values are showing more over-dispersed than expected by the null model (random) and positive values are showing more clustered data \citep{webb2002phylogenies}.
+\subsection{Data availability and distribution}
+To assess the availability of cladistic data for each mammalian Order, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: Family, Genus and species.
+We used these different taxonomic levels because some clades are well covered at the Family or Genus level, but poorly covered at the species level.
+We consider Orders with \textless 25\% of living taxa with cladistic data as having poor data coverage, and Orders with up \textmore 75\% of living taxa with cladistic data as having good data coverage.
+
+For Orders with \textless 100\% cladistic data coverage at any taxonomic level, we investigated whether the available cladistic data was (i) randomly distributed, (ii) over-dispersed or (iii) clustered with respect to phylogeny, using the Net Relatedness Index (NRI) metric \citep{webb2002phylogenies} in the \texttt{picante} R package \citep{picante}.
+% NC: Need to cite R too
+NRI is a metric from community phylogenetics that quantifies the distribution of the observed data compared to a model where all the data is sampled for each OTU. 
+%NC: NO!
+Not true unless you did something special.
+It assumes you have the same number of data points, but they are distributed at random.
+Then compares how different the observed is from random.
+Needs more work here.
+Negative NRI values show more over-dispersed than expected by the null model (random) and positive values are showing more clustered data \citep{webb2002phylogenies}.
 We choose NRI metric because is has been shown to be slightly less sensitive to the structure of the phylogeny (i.e.
 branch length and topology) \citep{NRI,journal.pone.0004390} but we also calculated two other common phylogenetic structure indices: Faith's Phylogenetic Distance (PD) \citep{Faith19921} and the Nearest Taxon Index (NTI) \citep{webb2002phylogenies}.
+% NC: Again No.
+NRI is actually skewed due to branch length distributions - see my 2008 proc b paper.
+Maybe worth saying it's been used in other studies then you can cite my ecology letters paper from 2012.
+Yay citations
 Both metrics are available in the Supplementary results.
+% NC: Both metrics? What other metric :)
 
-%\subsubsection{Data improvement strategy}
-%Too long for this paper and actually still a lot of work to do (transform the code of phylotargeting + come up with a valid strategy)
-
+% NC: OK needs a bit of work, a bit more clarification of what is going on here. We can chat tomorrow.
 %---------------------------------------------

### 2015-05-21 Manuscript/Missing_mammals-draft.tex
Edits to results and tables

--- 2015-05-20 Manuscript/Missing_mammals-draft.tex
@@
 
+% NC: Need to make a decision of Genus vs genus, Order vs order etc. I'd probably go for small letters actually. Do this throughout
+% NC: Note Figure and Table always have capital letters. For figure 2A or 2B don't put - (not 2-B for example)
+
 \section{Results}
 
-\subsection{Data availability}
+%\subsection{Data availability}
 We extracted 1422 living mammal OTUs from the 256 matrices with a minimum of 6 and a maximum of 4541 coded cladistic characters per OTU.
 After removing all the matrices with less than 100 cladistic characters, the number of extracted living mammals OTUs was down to 815.
-11/28 orders have less than 25\% of OTUs with cladistic data at a species level and 24/28 orders have less than 75\% taxa with available cladistic data.
-At the Genus level however only 3/28 orders have less than 25\% of taxa with cladistic data and 16/28 have less than 75\%.
-Finally, at the family level no order has less than 25\% taxa with available cladistic data and only 5/28 have less than 75\% (table \ref{Table_morpho_taxa_proportion}).
+% NC: Ah sorry I've asked you to add this in above. I reckon it goes up in the methods.
+In the @@@ cladistic matrices extracted in our searches, 11 out of 28 mammalian orders have \textless 25\% of species with cladistic data and 24 orders have less \textless 75\% species with available cladistic data.
+At the Genus level, only 3 orders have \textless 25\% of taxa with cladistic data and 16 orders have less than 75\%.
+Finally, at the family level no order has \textless 25\% taxa with available cladistic data and only 5 orders have \textless 75\% (table \ref{Table_morpho_taxa_proportion}).
+
+% NC: Results need to match the methods.
+So are we looking at "good" versus "poor"?
+That makes more sense to me.
+So how many have good coverage i.e.
+>75% and how many have poor coverage i.e.
+<25%.
+I think this will help.
+% NC: Can probably also just say that there are 28 orders once, then say how many fit into each category or it's needlessly repetitive. We can write this together to save time
 
@@
 
-\subsection{Available data structure}
-Among the orders containing OTUs with no available cladistic data, only two orders are significantly clustered: Carnivora and Chiroptera both at the species and the genus level (table~\ref{Table_data_structure}).
+%\subsection{Available data structure}
+Among the mammalian orders containing OTUs with no available cladistic data, only two orders, Carnivora and Chiroptera, are significantly phylogenetically clustered both at the species and the genus level (Table~\ref{Table_data_structure}).
 
@@
 
-Two contrasted results are shown on figure \ref{Figure_example_coverage} with randomly distributed data in Cetartiodactyla (figure \ref{Figure_example_coverage}-A) and clustered available data in Carnivora (mainly Canidae; figure \ref{Figure_example_coverage}-B).
+Two contrasting results are shown on Figure \ref{Figure_example_coverage} with randomly distributed data in Cetartiodactyla (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered data in Carnivora (mainly Canidae; figure \ref{Figure_example_coverage}B).
 
@@
     \includegraphics[width=1\textwidth]{example_coverage.pdf}
-\caption{Distribution of available cladistic data across two different clades (A: Cetartiodactyla; B: Carnivora).
+\caption{Phylogenetic distribution of available cladistic data across two mammalian orders (A: Cetartiodactyla; B: Carnivora).
 Edges are colored in grey when no cladistic data is available or in red when data is available.}
@@
 \end{figure}
+% NC: Can we change the colours, red sucks. Otherwise I like the figure a lot.
 

### 2015-05-22 Manuscript/Missing_mammals-draft.tex

@@
 \medskip
-\noindent{*\bf Corresponding author.} \textit{Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.}\\ % NC: Might need to reconsider the email here as it won't be your email anymore when you leave.
-Or do you get to keep it forever?
+\noindent{*\bf Corresponding author.} \textit{Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.}\\  
 \vspace{1in}
@@
 % NC: Might need to explain what you mean by OTUs, I'm assuming you mean genera or species etc.
+% because some trees have species, some have subspecies etc.
 We refer to OTUs instead of species because?
 % NC: do you mean 9411 different OTUs or total OTUs? Might be good to have a number of unique OTUs as well.
-
+% NC: Explain the genera vs species - ie how you might have more at one level than expected.
 % NC; probably need some comment to say this is collecting all the data!
@@
 We designated as ``living'' all the OTUs that where either present in the phylogeny of Fritz \textit{et al}.
-\citep{FritzTree} or the taxonomy of Wilson \& Reeder \citep{wilson2005mammal}, and designated as ``fossil'' all the OTUs that were present in the Paleobiology database (\texttt{https://paleodb.org} %NC: Check website
+\citep{FritzTree} or the taxonomy of Wilson \& Reeder \citep{wilson2005mammal}, and designated as ``fossil'' all the OTUs that were present in the Paleobiology database (\texttt{https://paleodb.org} %NC: Check website or reference
 For OTUs that did not appear in these three sources we first decomposed the name (i.e.
@@
 % NC: I'm not sure why you did this. You might need to explain more - let's discuss. Otherwise it looks like we are deliberately making the data coverage look worse by omitting data.
+
+% NC: START with - small matrices are shit for these reasons.
+
 Next we selected only matrices containing at least 100 cladistic characters per OTU.
@@
 ``antler ramifications'' is a character than is only applicable to Cervidae not all mammals).
+% NC: We also did it with all data.
 
@@
 We used these different taxonomic levels because some clades are well covered at the Family or Genus level, but poorly covered at the species level.
-We consider Orders with \textless 25\% of living taxa with cladistic data as having poor data coverage, and Orders with up \textmore 75\% of living taxa with cladistic data as having good data coverage.
+We consider Orders with \textless 25\% of living taxa with cladistic data as having poor data coverage, and Orders with up \textgreater 75\% of living taxa with cladistic data as having good data coverage.
 
@@
 % NC: Need to cite R too
+% NC: Put NTI and PD up here too then just describe NRI.
 NRI is a metric from community phylogenetics that quantifies the distribution of the observed data compared to a model where all the data is sampled for each OTU. 
@@
 Both metrics are available in the Supplementary results.
-% NC: Both metrics? What other metric :)
 
@@
 In the @@@ cladistic matrices extracted in our searches, 11 out of 28 mammalian orders have \textless 25\% of species with cladistic data and 24 orders have less \textless 75\% species with available cladistic data.
-At the Genus level, only 3 orders have \textless 25\% of taxa with cladistic data and 16 orders have less than 75\%.
+At the Genus level, only three orders have \textless 25\% of taxa with cladistic data and 16 orders have less than 75\%.
 Finally, at the family level no order has \textless 25\% taxa with available cladistic data and only 5 orders have \textless 75\% (table \ref{Table_morpho_taxa_proportion}).
@@
 \section{Discussion}
-Our results show that even though relations among living mammals are well known in a molecular framework \citep[e.g.][]{FritzTree,meredithimpacts2011,May-Collado-PeerJ} only little cladistic data is available.
-This is primarily due to the fact that that cladistic studies are mainly focusing on fossil taxa rather than living taxa \citep[e.g.][]{O'Leary08022013,ni2013oldest}.
-However, data availability varies in its structure as well as across taxonomic levels.
-Higher taxonomic levels are always better sampled than lower ones (table \ref{Table_morpho_taxa_proportion}) and within these taxonomic levels, the structure of available data is mostly random apart for two orders (table \ref{Table_data_structure})b.
+Our results show that although phylogenetic relationships among living mammals are well-resolved \citep[e.g.][]{FritzTree,meredithimpacts2011,May-Collado-PeerJ}, most of the data used to build these phylogenies is molecular, and very little cladistic data is available for living mammals compared to fossil mammals \citep[e.g.][]{O'Leary08022013,ni2013oldest}.
+This has implications for building Total Evidence phylogenies containing both living and fossil mammals as without sufficient cladistic data for living species, fossil placements in these trees is very uncertain \citep{GuillermeCooper}.
+However, data coverage in living mammals varies across taxonomic levels and in its phylogenetic distribution.
+Higher taxonomic levels are always better sampled than lower ones and within these taxonomic levels, the available data is mostly randomly distributed across the phylogeny, apart from in Carnivora and Chiroptera.
+% NC: You don't generally refer to results tables etc in the discussion
 
 %\subsection{Data availability}
-The amount of living taxa with no available cladistic data was surprisingly high at the species level: 24/28 orders have less than 75\% taxa with available cladistic data (and two of the 28 orders are monospecific!).
-This threshold of 75\% available data is the maximum amount of possible missing data (25\%) before having a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.
-Beyond this threshold, there is a noteworthy displacement of the wildcard taxa (\textit{sensu} \citep{kearneyfragmentary2002}) as well as a important decrease in the conservation of clades \citep{GuillermeCooper}.
-At the species level within most orders of mammals, it is therefore likely to see topological artefacts for the placement of fossil taxa.
-However, the data availability seems to be less an issue at higher taxonomic level (i.e. the Genus and the Family level).
+The number of living taxa with no available cladistic data was surprisingly high at the species-level: 24 out of 28 orders have \textless 75\% taxa with available cladistic data (and two of the 28 orders are monospecific!).
+This threshold of 75\% available data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.
+Beyond this threshold, there is considerable displacement of wildcard taxa (\textit{sensu} \citep{kearneyfragmentary2002}) and decreases in clade conservation \citep{GuillermeCooper}.
+Therefore, at the species-level within most orders of mammals, we expect to see topological artefacts for the placement of fossil taxa. % NC: Rephrase
+However, data coverage seems to be less of an issue at higher taxonomic levels (i.e. genus- and family-level).
 This point is important to consider from a practical point of view because of the slight discrepancy between the neontological and the palaeontological concept of species.
-While neontological species are described on accurately measured morphology, genetic distance, spacial distribution and even behaviour \citep[e.g.][]{kellymolecular2014}, palaeontological data can be based only on morphological, spatial and temporal data \citep[e.g.][]{ni2013oldest}.
+While neontological species are described on accurately measured morphology, genetic distance, spatial distribution and even behaviour \citep[e.g.][]{kellymolecular2014}, palaeontological data can be based only on morphological, spatial and temporal data \citep[e.g.][]{ni2013oldest}.
 Because of this difference between neontological species (\textit{sensu} reproductive isolates) and paleontological species (\textit{sensu} morpho-species %or "morphological clusters"?
 ), most palaeontological studies are using the Genus level for their smallest OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
-In this frame of palaeontological data usage, the data availability at the Genus level in living mammals should be our priority concern.
+In this frame of palaeontological data usage, the data availability at the genus-level in living mammals should be our priority concern.
 
@@
 %Caveats
-In this study, we treated all cladistic matrices to be equal in a similar way that molecular matrices are: for example if a matrix A contains 100 characters for taxa X and Y, and a matrix B contains 50 characters for taxa X and Z, we assumed that both matrices can be combined in a supermatrix containing 150 characters for taxon X, 100 for taxon Y and 50 for taxon Z.
-However, it is clear that cladistic data cannot be treated that way \citep{Brazeau2011}.
-In fact, for the same taxon, some characters might overlap (e.g.
-matrix A has a character coding for the shape of a particular morphological feature and matrix B as a first character coding for the presence of this morphological feature and a second one coding for its shape; in this case, these three characters are compound characters \citep{Brazeau2011}).
-However, in reasonably sized matrices ($>$ 100 characters \citep{GuillermeCooper,harrisonamong-character2014} it is more likely that a number of characters are consistently conserved among the different matrices (e.g.
-within primates literature, the character \textit{p7} - the size of the $4^{th}$ lower premolar paraconid - is consistent for more than 15 years \citep{ross1998phylogenetic,seiffert2003fossil,marivaux2005anthropoid,seiffert2005basal,bloch2007new,kay2008anatomy,silcox2008biogeographic,seiffert2009convergent,tabuce2009anthropoid,boyer2010astragalar,seiffert2010fossil,marivaux2013djebelemur,ni2013oldest}).
-A conservative approach to avoid compound characters could be to select only the most recent matrix for each group.
+In this study, we treated all cladistic matrices as equal in a similar way to molecular matrices. 
+For example, if matrix A contained 100 characters for taxa X and Y, and matrix B contained 50 different characters for taxa X and Z, we assumed that both matrices can be combined in a supermatrix containing 150 independent characters for taxon X, 100 for taxon Y and 50 for taxon Z.
+Unfortunately, cladistic data cannot always be treated in this way because some characters may overlap.
+For example, if matrix A has a character coding for the shape of a particular morphological feature and matrix B has a character coding for the presence of this morphological feature and a second character coding for its shape then these three characters are non-independent compound characters \citep{Brazeau2011}.
+However, in reasonably sized matrices (\textgreater 100 characters; \citealp{GuillermeCooper,harrisonamong-character2014}) it is more likely that a number of characters are consistently conserved among the different matrices (e.g.
+within the Primate literature, the character \textit{p7} - the size of the $4^{th}$ lower premolar paraconid - has been used consistently for \textgreater 15 years; \citealp[e.g.][]{ross1998phylogenetic,seiffert2003fossil,marivaux2005anthropoid,seiffert2005basal,bloch2007new,kay2008anatomy,silcox2008biogeographic,seiffert2009convergent,tabuce2009anthropoid,boyer2010astragalar,seiffert2010fossil,marivaux2013djebelemur,ni2013oldest}).
+% NC: Remove some of these
+A conservative approach to avoid compound characters would be to select only the most recent matrix for each group, but this would result in the loss of a lot of data.
 
 %Let's code some data
-Following the recommendations in \citep{GuillermeCooper}, one should code cladistic characters for a maximum of living species in order to improve the topology of Total Evidence trees containing both living and fossil taxa.
-Since the data for living mammals is usually easily available in world wide vast natural history collections, we propose that an increased effort in coding morphological characters from living species should be done via engaging in collaborative data collection projects through web portals such as \textit{morphobank} \citep{morphobank}.
+Following the recommendations in \citep{GuillermeCooper}, to improve the accuracy of the topology of Total Evidence trees containing both living and fossil taxa, one should code cladistic characters for as many living species possible.
+Since the data for living mammals is usually easily available in world wide vast natural history collections, we propose that an increased effort in coding morphological characters from living species should be done via engaging in collaborative data collection projects through web portals such as \textit{Morphobank} \citep{morphobank}.
 
@@
 \section{Data accessibility statement}
-All data is available and reproducible on GitHub.
+All data and analysis code is available on GitHub.
 \section{Authors' contributions statement}
@@
 \beginsupplement
-\input{Supplementary/Supplementary}
+%\input{Supplementary/Supplementary}
 

### 2015-05-22 Functions/sanitizing.R
Updated check. functions

@@ -3,6 +3,10 @@
 
 #Checking the class of an object and returning an error message if != class
 check.class<-function(object, class, msg, errorif=FALSE) {
+    #Set msg if missing
+    if(missing(msg)) {
+        msg<-paste(" must be ", class, ".", sep="")
+    }
     #check if object is class.
     if(length(class) == 1) {
         if(errorif==FALSE) {
@@ -43,6 +47,7 @@ check.length<-function(object, length, msg, errorif=FALSE) {
         }        
     }
 }
+
 
 
 #Cleaning a tree so that the species match with the ones in a table

### 2015-05-22 Functions/table.results.R
table.result fun init

@@ -1,0 +1,73 @@
+#Making a hybrid table/figure with the taxonomic coverage result
+##########################
+#SYNTAX:
+#<results> a result table from previous functions
+#<order> which order to plot
+#<metric> which metric to print
+#<thresholds> threshold values (default=c(25,75))
+##########################
+#guillert(at)tcd.ie - 22/05/2015
+##########################
+
+table.result<-function(results, order, metric, thresholds=c(25,75)) {
+    #SANITIZING
+
+    #results
+    check.class(results, 'data.frame')
+    #names
+    if(length( match(names(results), c("Order", "Taxonomic.level", "Number.of.OTUs", "Percentage.of.OTUs"))[-which(is.na(match(names(results), c("Order", "Taxonomic.level", "Number.of.OTUs", "Percentage.of.OTUs"))))] ) != 4) {
+        stop("Wrong headers names.")
+    }
+
+    #order
+    check.class(order, 'character')
+
+    #metric
+    check.class(metric, 'character')
+    if(length(grep(metric, names(results))) == 0) {
+        stop(paste(metric, "not found in the data.frame."))
+    }
+    if(length(grep(paste(metric, "p", sep="_"), names(results))) == 0) {
+        stop(paste(metric, "p-value not found in the data.frame."))
+    }
+
+    #thresholds
+    check.class(thresholds, 'numeric')
+
+    #CREATING THE TABLE/FIGURE
+    #Isolating the orders
+    if(length(order != 1)) {
+        results_tmp<-results[which(results$Order == order[1]),]
+        for(ord in 2:length(order)) {
+            results_tmp<-rbind(results_tmp, results[which(results$Order == order[ord]),])
+        }
+    } else {
+        results_tmp<-results[which(results$Order == order),]
+    }
+
+    #barplot
+    #percentages<-table(as.numeric(results_tmp$Percentage.of.OTUs))
+    #names(percentages) <- rep(c("Family", "Genera", "Species"), length(order))
+    #percentages[]<-as.numeric(results_tmp$Percentage.of.OTUs)
+    
+
+    barplot(rev(as.numeric(results_tmp$Percentage.of.OTUs)), horiz=TRUE, xaxt="n", yaxt="n", main="percentage of OTUs")
+    for(thresh in 1:length(thresholds)) {
+        abline(v=thresholds[thresh], lty=2)
+    }
[…]

### 2015-05-27 Manuscript/Missing_mammals-draft.tex
added R ref
Revised methods

@@
 \usepackage{fullpage}
-%\usepackage{natbib}
 \usepackage{float}
@@
 % NC: sometimes the better title is one that answers the question rather than giving the topic. e.g., "Morphological data is lacking for living species" or similar
+% TG: maybe worth considering the special issue as well? "Putting fossils in tree" so maybe something like "There is not enough morphological data for living mammals to branch fossils in the trees."
+% TG: or some poetic... "Living and dead mammals: morphological data is lacking in living for combined analysis"
 
@@
 
-%\newpage
-
 \section{Materials and Methods}
 \subsection{Data collection and standardisation}
-We downloaded all cladistic matrices containing any living or/and fossil mammal taxa from three major public databases: Morphobank (\texttt{http://www.morphobank.org/}) \citep{morphobank}, Graeme Lloyd's website (\texttt{http://graemetlloyd.com/}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}).
-% NC: give link to actual repo and actual bit of graemes website these things are found on?
-We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see see Supplementary Materials for a detailed description of the search procedure).
-%We downloaded the additional cladistic matrices from the 20 first search results matching with our selected keywords and with any of the 35 taxonomic levels (mammals Orders, Infraclasses and Class; see supplementary materials for detailed description of the procedure).
-% NC: Can possibly delete this and stick in supplement
-In total, we downloaded 256 % @@@ Number of matrices
-matrices containing 9411 % @@@ Number of living OTUs
-operational taxonomic units (OTUs). % link
-% NC: Might need to explain what you mean by OTUs, I'm assuming you mean genera or species etc.
-% because some trees have species, some have subspecies etc.
-We refer to OTUs instead of species because?
-% NC: do you mean 9411 different OTUs or total OTUs? Might be good to have a number of unique OTUs as well.
-% NC: Explain the genera vs species - ie how you might have more at one level than expected.
-% NC; probably need some comment to say this is collecting all the data!
+We downloaded all cladistic matrices containing any living or/and fossil mammal taxa from three major public databases: Morphobank (\texttt{http://www.morphobank.org/}) \citep{morphobank}, Graeme Lloyd's website (\texttt{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce/cladistic-data}).
+We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
+In total, we downloaded @256 % Number of matrices
+matrices containing unique @9411 % Number of unique OTUs
+operational taxonomic units (OTUs).
+In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardized and ranged from specific specimen names to family level.
+When possible, we considered the OTUs at their lowest valid taxonomic level (species) but some OTUs where only at a higher valid taxonomic level (e.g. genus or family).
+Therefore for some orders, we sampled more genus than species (Table \ref{Table_morpho_taxa_proportion}).
 
-%We transformed all the matrices to be in nexus format.
-We standardised the taxonomy of our OTUs by fixing invalid binomial inputs to match with standard taxonomic nomenclature (e.g., \textit{H. sapiens} was transformed to \textit{Homo sapiens}).
+To select the lower valid taxonomic level, we standardised the taxonomy of our OTUs by fixing invalid binomial inputs to match with standard taxonomic nomenclature (e.g., \textit{H.
+sapiens} was transformed to \textit{Homo sapiens}).
 We designated as ``living'' all the OTUs that where either present in the phylogeny of Fritz \textit{et al}.
-\citep{FritzTree} or the taxonomy of Wilson \& Reeder \citep{wilson2005mammal}, and designated as ``fossil'' all the OTUs that were present in the Paleobiology database (\texttt{https://paleodb.org} %NC: Check website or reference
+\citep{FritzTree} or the taxonomy of Wilson \& Reeder \citep{wilson2005mammal}, and designated as ``fossil'' all the OTUs that were present in the Paleobiology database (\texttt{https://paleobiodb.org/}).
 For OTUs that did not appear in these three sources we first decomposed the name (i.e.
@@
 
-%\subsection{Data availability analysis}
-%\subsubsection{Number of characters threshold}
-% NC: I'm not sure why you did this. You might need to explain more - let's discuss. Otherwise it looks like we are deliberately making the data coverage look worse by omitting data.
+The number of character per matrices ranged from @6 to @4541.
+This range led to difficulties in comparing the available data between matrices because (1) small matrices have less chance of having character overlapping with other matrices \citep{wagner2000} and (2) small matrices are more likely to contain a higher proportion of specific characters non-applicable across large clades \citep{Brazeau2011} (e.g.
+``antler ramifications'' is a character that is only applicable to Cervidae not all mammals).
+Therefore we selected only matrices containing at least 100 cladistic characters per OTU.
+This arbitrary threshold was chosen to correspond with the number of characters used in \citep{GuillermeCooper} and \citep{harrisonamong-character2014}.
 
-% NC: START with - small matrices are shit for these reasons.
-
-Next we selected only matrices containing at least 100 cladistic characters per OTU.
-This arbitrary threshold was chosen to correspond with the number of characters used in \citep{GuillermeCooper} and \citep{harrisonamong-character2014}.
-Additionally, this threshold avoids biases towards small matrices that are often non-informative, because they contain too few characters so there is no character overlap among matrices \citep{wagner2000}, or contain characters that are highly specific and thus non-applicable across large clades \citep{Brazeau2011} (e.g.
-``antler ramifications'' is a character than is only applicable to Cervidae not all mammals).
-% NC: We also did it with all data.
-
-% NC: Some summing up statement about how many matrices were actually usuable, how many OTUs, how many unique OTUs etc.
+We extracted @1422 unique living mammal OTUs from the @256 matrices with a minimum of 6 and a maximum of 4541 coded cladistic characters per OTU.
+After removing all the matrices with less than 100 cladistic characters, the number of extracted living mammals OTUs was down to @815 from @number matrices.
 
@@
 We used these different taxonomic levels because some clades are well covered at the Family or Genus level, but poorly covered at the species level.
-We consider Orders with \textless 25\% of living taxa with cladistic data as having poor data coverage, and Orders with up \textgreater 75\% of living taxa with cladistic data as having good data coverage.
+We consider Orders with \textless 25\% of living taxa with cladistic data as having poor data coverage (hereafter ``low'' coverage), and Orders with up \textgreater 75\% of living taxa with cladistic data as having good data coverage (hereafter ``high'' coverage).
 
-For Orders with \textless 100\% cladistic data coverage at any taxonomic level, we investigated whether the available cladistic data was (i) randomly distributed, (ii) over-dispersed or (iii) clustered with respect to phylogeny, using the Net Relatedness Index (NRI) metric \citep{webb2002phylogenies} in the \texttt{picante} R package \citep{picante}.
-% NC: Need to cite R too
-% NC: Put NTI and PD up here too then just describe NRI.
-NRI is a metric from community phylogenetics that quantifies the distribution of the observed data compared to a model where all the data is sampled for each OTU. 
-%NC: NO!
-Not true unless you did something special.
-It assumes you have the same number of data points, but they are distributed at random.
-Then compares how different the observed is from random.
-Needs more work here.
-Negative NRI values show more over-dispersed than expected by the null model (random) and positive values are showing more clustered data \citep{webb2002phylogenies}.
-We choose NRI metric because is has been shown to be slightly less sensitive to the structure of the phylogeny (i.e.
-branch length and topology) \citep{NRI,journal.pone.0004390} but we also calculated two other common phylogenetic structure indices: Faith's Phylogenetic Distance (PD) \citep{Faith19921} and the Nearest Taxon Index (NTI) \citep{webb2002phylogenies}.
-% NC: Again No.
-NRI is actually skewed due to branch length distributions - see my 2008 proc b paper.
-Maybe worth saying it's been used in other studies then you can cite my ecology letters paper from 2012.
-Yay citations
-Both metrics are available in the Supplementary results.
+For Orders with \textless 100\% cladistic data coverage at any taxonomic level, we investigated whether the available cladistic data was (i) randomly distributed, (ii) over-dispersed or (iii) clustered with respect to phylogeny, using the Nearest Taxon Index (NTI) metric \citep{webb2002phylogenies} in the \texttt{picante R} package \citep{picante,R}.
+The NTI is based on the mean nearest neighbour distance (MNND) and is calculated as follow:
+\begin{equation}
+NTI=-(\frac{\overline{MNND}_{obs}-\overline{MNND}_{n}}{\sigma(MNND_{n})})
+\end{equation}
+Where $\overline{MNND}_{obs}$ is the mean distance between each $n$ taxa with available cladistic data within each order; $\overline{MNND}_{n}$ is the expected random MNND for $n$ taxa estimated by calculating the MNND from $n$ randomly drawn taxa repeated 1000 times and $\sigma(MNND_{n})$ is the standard deviation of the 1000 random MNND$_{n}$.
+Negative NTI values shows more over-dispersed data than expected by chance (random) and positive values are showing more clustered data \citep{webb2002phylogenies}.
+Similarly, we also calculated the Nearest Relatedness Index (NRI) \citep{webb2002phylogenies,Cooper2012} and the Faith's phylogenetic distance (PD) \citep{Faith19921}.
+Both NRI and PD are calculated in the same way as the NTI (\((observed-\mu(random))/\sigma(random)\)) but they use respectively the mean phylogenetic distance (MPD) and the sum of branch lengths (SBL) instead of the MNND \citep{Faith19921,webb2002phylogenies,Cooper2012}.
 
-% NC: OK needs a bit of work, a bit more clarification of what is going on here. We can chat tomorrow.
+Because the aim of this study is to understand the structure of the available cladistic data in mammals we prefered the NTI over the NRI or the PD metrics.
+In fact, the NTI metric is more sensitive to the phylogenetic structure (clustering or over-dispersion) at the tips of phylogeny \citep{Cooper2008}.
+%and also because previous studies have shown that NRI is biased towards detecting overdispersion \citep{Kembel2006,Swenson2006}. TG: Useless in this case?
+However, the results of the NRI and PD calculations are available in the supplementary.
+
 %---------------------------------------------
@@
 
-% NC: Need to make a decision of Genus vs genus, Order vs order etc. I'd probably go for small letters actually. Do this throughout
-% NC: Note Figure and Table always have capital letters. For figure 2A or 2B don't put - (not 2-B for example)
-
 \section{Results}
-
-%\subsection{Data availability}
-We extracted 1422 living mammal OTUs from the 256 matrices with a minimum of 6 and a maximum of 4541 coded cladistic characters per OTU.
-After removing all the matrices with less than 100 cladistic characters, the number of extracted living mammals OTUs was down to 815.
-% NC: Ah sorry I've asked you to add this in above. I reckon it goes up in the methods.
-In the @@@ cladistic matrices extracted in our searches, 11 out of 28 mammalian orders have \textless 25\% of species with cladistic data and 24 orders have less \textless 75\% species with available cladistic data.
-At the Genus level, only three orders have \textless 25\% of taxa with cladistic data and 16 orders have less than 75\%.
-Finally, at the family level no order has \textless 25\% taxa with available cladistic data and only 5 orders have \textless 75\% (table \ref{Table_morpho_taxa_proportion}).
+In the @@@ cladistic matrices extracted in our searches, 11 out of 28 mammalian orders have \textless 25\% of species with cladistic data and 24 have less \textless 75\% species with available cladistic data.
+At the Genus level, only three orders have \textless 25\% of taxa with cladistic data and 16 have less than 75\%.
+Finally, at the family level no order has \textless 25\% taxa with available cladistic data and only 5 have \textless 75\% (Table \ref{Table_morpho_taxa_proportion}).
 
@@
 I think this will help.
-% NC: Can probably also just say that there are 28 orders once, then say how many fit into each category or it's needlessly repetitive. We can write this together to save time
-
+% TG: I will change it when I'll recalculate the results, it'll be " 15/25 mammals have "low" coverage and 1/25 have "high" coverage"
+% TG: Same for the table, I will do the new table combinations later (need some proper coding since I don't wan't to use illustrator - my first paper 100\% reproducible by copy paste!)
 \renewcommand\baselinestretch{1.2}\selectfont
@@
 
-%\subsection{Available data structure}
 Among the mammalian orders containing OTUs with no available cladistic data, only two orders, Carnivora and Chiroptera, are significantly phylogenetically clustered both at the species and the genus level (Table~\ref{Table_data_structure}).
@@
 
-Two contrasting results are shown on Figure \ref{Figure_example_coverage} with randomly distributed data in Cetartiodactyla (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered data in Carnivora (mainly Canidae; figure \ref{Figure_example_coverage}B).
+Two contrasting results are shown on Figure \ref{Figure_example_coverage} with randomly distributed data in Cetartiodactyla (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered data in Carnivora (mainly Canidae; Figure \ref{Figure_example_coverage}B).
 
@@
 % NC: Can we change the colours, red sucks. Otherwise I like the figure a lot.
-
-%\subsection{Data improvement strategy}
+% TG: Again, will fix in the final version.
 
@@
 Higher taxonomic levels are always better sampled than lower ones and within these taxonomic levels, the available data is mostly randomly distributed across the phylogeny, apart from in Carnivora and Chiroptera.
-% NC: You don't generally refer to results tables etc in the discussion
 
-%\subsection{Data availability}
 The number of living taxa with no available cladistic data was surprisingly high at the species-level: 24 out of 28 orders have \textless 75\% taxa with available cladistic data (and two of the 28 orders are monospecific!).
@@
 Beyond this threshold, there is considerable displacement of wildcard taxa (\textit{sensu} \citep{kearneyfragmentary2002}) and decreases in clade conservation \citep{GuillermeCooper}.
-Therefore, at the species-level within most orders of mammals, we expect to see topological artefacts for the placement of fossil taxa. % NC: Rephrase
+Therefore we expect a high probability of topological artefacts for the placement of fossil taxa at the species-level in most mammalian orders.
 However, data coverage seems to be less of an issue at higher taxonomic levels (i.e. genus- and family-level).
@@
 
-%\subsection{Available data structure}
 Regardless the low level of available data, it is encouraging that for most orders (except Carnivora and Chiroptera), data is randomly distributed across the phylogeny.
-%Therefore, it is likely that the addition of fossil taxa within these orders will not be biased by oversampling but will probably branch near the taxa that has the most similar available data (as expected).
 There are three different scenarii, regarding the distribution of the available data, for predicting how fossils can branch in relation to living taxa.
@@
 However, in the third scenario, when the data is clustered we expect two major biases to occur: (1) first, the fossil will not be able to branch within a clade containing no data (e.g.
-Herpestidae in Carnivora; figure \ref{Figure_missing_data_problem} and \ref{Figure_example_coverage}-B), and (2) second, the fossil will have a higher probability, at random, to branch within the clade containing most of the available data (e.g.
-a Carnivora fossil will have more chance to branch, randomly, to the Canidae clade than any other clade in Carnivora; figure \ref{Figure_example_coverage}-B).
+Herpestidae in Carnivora; Figure \ref{Figure_missing_data_problem} and \ref{Figure_example_coverage}B), and (2) second, the fossil will have a higher probability, at random, to branch within the clade containing most of the available data (e.g.
+a Carnivora fossil will have more chance to branch, randomly, to the Canidae clade than any other clade in Carnivora; Figure \ref{Figure_example_coverage}-B).
 This means that fossils with uncertain phylogentic affinities (\textit{Incertae sedis}) will have a higher probability of branching by chance within the most sampled clade.
@@
 All data and analysis code is available on GitHub.
-\section{Authors' contributions statement}
-Conceived and designed the experiments: TG NC.
-Performed the experiments: TG.
-Analyzed the data: TG.
-Wrote the paper: TG NC.
-% NC: Write this in the proper proceedings style.
-All authors gave final approval for publication.
+\section{Authors' Contributions}
+T.G.
+and N.C conceived and designed the experiments.
+T.G.
+performed the experiments and analysed the data.
+T.G.
+and N.C.
+contributed to the writing of the manuscript.
+All authors approved the final version of the manuscript.
+\section{Competing Interests}
+We have no competing interests.
 \section{Acknowledgements}
-Nick Matzke, April Wright, David Bapst and Graeme Lloyd. %Or whatever order. % NC: Alphabetical?
+David Bapst, Graeme Lloyd, Nick Matzke and April Wright.
 \section{Funding statement}

### 2015-05-27 Manuscript/Missing_mammals-draft.tex
Results fixed

@@
 For OTUs that did not appear in these three sources we first decomposed the name (i.e.
-\textit{Homo$\_$sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (Family, Genus etc.).
+\textit{Homo$\_$sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
 Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; see Supplementary Material for more details).
@@
 \subsection{Data availability and distribution}
-To assess the availability of cladistic data for each mammalian Order, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: Family, Genus and species.
-We used these different taxonomic levels because some clades are well covered at the Family or Genus level, but poorly covered at the species level.
+To assess the availability of cladistic data for each mammalian Order, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
+We used these different taxonomic levels because some clades are well covered at the family or genus level, but poorly covered at the species level.
 We consider Orders with \textless 25\% of living taxa with cladistic data as having poor data coverage (hereafter ``low'' coverage), and Orders with up \textgreater 75\% of living taxa with cladistic data as having good data coverage (hereafter ``high'' coverage).
@@
 \section{Results}
-In the @@@ cladistic matrices extracted in our searches, 11 out of 28 mammalian orders have \textless 25\% of species with cladistic data and 24 have less \textless 75\% species with available cladistic data.
-At the Genus level, only three orders have \textless 25\% of taxa with cladistic data and 16 have less than 75\%.
-Finally, at the family level no order has \textless 25\% taxa with available cladistic data and only 5 have \textless 75\% (Table \ref{Table_morpho_taxa_proportion}).
+In the @@@ cladistic matrices extracted in our searches, @11 out of @28 mammalian orders have a ``low'' coverage (\textless 25\%) of species with cladistic data and @24 have a ``high'' coverage (\textgreater 75\%).
+At the genus level, only @three orders have a ``low'' coverage of taxa with cladistic data and @16 have a ``high'' coverage.
+Finally, at the family level @no order has a ``low'' coverage and only @5 have a ``high'' coverage (Table \ref{Table_morpho_taxa_proportion}).
 
-% NC: Results need to match the methods.
-So are we looking at "good" versus "poor"?
-That makes more sense to me.
-So how many have good coverage i.e.
->75% and how many have poor coverage i.e.
-<25%.
-I think this will help.
-% TG: I will change it when I'll recalculate the results, it'll be " 15/25 mammals have "low" coverage and 1/25 have "high" coverage"
 % TG: Same for the table, I will do the new table combinations later (need some proper coding since I don't wan't to use illustrator - my first paper 100\% reproducible by copy paste!)
@@
 
-Among the mammalian orders containing OTUs with no available cladistic data, only two orders, Carnivora and Chiroptera, are significantly phylogenetically clustered both at the species and the genus level (Table~\ref{Table_data_structure}).
+Among the mammalian orders containing OTUs with no available cladistic data, only @two orders, Carnivora and Chiroptera, are significantly phylogenetically clustered both at the species and the genus level (Table~\ref{Table_data_structure}).
 
@@
 
-The number of living taxa with no available cladistic data was surprisingly high at the species-level: 24 out of 28 orders have \textless 75\% taxa with available cladistic data (and two of the 28 orders are monospecific!).
-This threshold of 75\% available data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.
+The number of living taxa with no available cladistic data was surprisingly high at the species-level: @24 out of @28 orders have a ``high'' coverage of taxa with available cladistic data (and two of the 28 orders are monospecific!).
+This threshold of 75\% available data (``high'' coverage) represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.
 Beyond this threshold, there is considerable displacement of wildcard taxa (\textit{sensu} \citep{kearneyfragmentary2002}) and decreases in clade conservation \citep{GuillermeCooper}.
@@
 Because of this difference between neontological species (\textit{sensu} reproductive isolates) and paleontological species (\textit{sensu} morpho-species %or "morphological clusters"?
-), most palaeontological studies are using the Genus level for their smallest OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
+), most palaeontological studies are using the genus level for their smallest OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
 In this frame of palaeontological data usage, the data availability at the genus-level in living mammals should be our priority concern.

### 2015-05-27 Analysis/Analysis_per_order.R
Change case for taxonomic levels (WARNING see description)
WARNING: This might actually induce some errors. If it goes wrong,
revert to this commit!

@@ -87,7 +87,7 @@ for (character_threshold in 3:4) {
         if(is.null(tree)) {
             results_tmp<-matrix(nrow=3, ncol=ncol(results), data=NA)
             results_tmp[,1]<-rep(orders[order],3)
-            results_tmp[,2]<-c("Family", "Genus", "Species")
+            results_tmp[,2]<-c("family", "genus", "species")
             results_tmp[,3]<-rep("1/1",3)
             results_tmp[,4]<-rep("100",3)
             results_tmp<-as.data.frame(results_tmp)

### 2015-05-27 Functions/order.structure.R

@@ -43,7 +43,7 @@ order.structure<-function(order, taxa, tree, reference, metric=c("PD", "NRI", "N
 
     #ANALYSING THE DISTRIBUTION OF SPECIES PER ORDER WITH MORPHOLOGICAL DATA
     #Empty table
-    results<-data.frame("Order"=rep(order,3), "Taxonomic level"=c("Family", "Genus", "Species"), "Number of OTUs"=rep(NA,3), "Percentage of OTUs"=rep(NA,3))
+    results<-data.frame("Order"=rep(order,3), "Taxonomic level"=c("family", "genus", "species"), "Number of OTUs"=rep(NA,3), "Percentage of OTUs"=rep(NA,3))
     #Adding two columns per metric
     if(any(metric == "PD")) {
         results$PD<-rep(NA,3)

### 2015-05-27 Functions/table.results.R

@@ -59,7 +59,7 @@ table.result<-function(results, order, metric, thresholds=c(25,75)) {
 
 #    axis(side=2, at=1:(length(order)*3), labels=rev(rep(c("Family", "Genera", "Species"), 3)), las=2)
     axis(side=2)
-    axis(side=2, at=1:(length(order)*3), labels=rev(rep(c("Family", "Genera", "Species"), 3)), las=2)
+    axis(side=2, at=1:(length(order)*3), labels=rev(rep(c("family", "genus", "species"), 3)), las=2)
 
 
 

### 2015-05-27 Functions/table.results.R
Superdupper table/figure function. IN DEV

@@ -2,72 +2,142 @@
 ##########################
 #SYNTAX:
 #<results> a result table from previous functions
-#<order> which order to plot
 #<metric> which metric to print
 #<thresholds> threshold values (default=c(25,75))
+#<split> whether to split the table in smaller ones (environement will be ignored if split is not null)
+#<save.path> where to save the table
+#<file.save> name (or chain name) for saving the table
+#<caption> table caption (if split option is called, the caption will only appear for the first table)
+#<environement> table environement
 ##########################
-#guillert(at)tcd.ie - 22/05/2015
+#environement(at)tcd.ie - 27/05/2015
 ##########################
 
-table.result<-function(results, order, metric, thresholds=c(25,75)) {
-    #SANITIZING
-
-    #results
-    check.class(results, 'data.frame')
-    #names
-    if(length( match(names(results), c("Order", "Taxonomic.level", "Number.of.OTUs", "Percentage.of.OTUs"))[-which(is.na(match(names(results), c("Order", "Taxonomic.level", "Number.of.OTUs", "Percentage.of.OTUs"))))] ) != 4) {
-        stop("Wrong headers names.")
-    }
-
-    #order
-    check.class(order, 'character')
-
-    #metric
-    check.class(metric, 'character')
-    if(length(grep(metric, names(results))) == 0) {
-        stop(paste(metric, "not found in the data.frame."))
-    }
-    if(length(grep(paste(metric, "p", sep="_"), names(results))) == 0) {
-        stop(paste(metric, "p-value not found in the data.frame."))
-    }
-
-    #thresholds
-    check.class(thresholds, 'numeric')
-
-    #CREATING THE TABLE/FIGURE
-    #Isolating the orders
-    if(length(order != 1)) {
-        results_tmp<-results[which(results$Order == order[1]),]
-        for(ord in 2:length(order)) {
-            results_tmp<-rbind(results_tmp, results[which(results$Order == order[ord]),])
-        }
-    } else {
-        results_tmp<-results[which(results$Order == order),]
-    }
+table.result<-function(results, metric, threshold=c(25,75), split=NULL, save.path, file.save, caption, environement) {
+    
+    #DEBUG
+    warning("Debug mode is ON (table.result)")
+    #results -> must be a table
[…]

### 2015-06-04 Functions/table.results.R
Updated bar size (now 4:1 rectangles)

@@ -106,27 +106,34 @@ table.result<-function(results, metric, threshold=c(25,75), split=NULL, save.pat
 
     #Creating the folder for storing the graphical elements
     new_folder<-paste(save.path, "Table_figures/", sep="")
-    system(paste("mkdir", new_folder))
+    #If folder does not exist or is empty, create it
+    if(length(list.files(new_folder)) == 0) {
+        system(paste("mkdir", new_folder))
+    } else {
+    #Else remove it and create a new one
+        system(paste("rm -R", new_folder))
+        system(paste("mkdir", new_folder))
+    }
 
     #Make sure the percentage column is numeric
 
     #Creating all the barplots
     for(row in 1:nrow(table_to_print)) {
         #pdf settings
-        pdf(paste(new_folder, "bar", row , ".pdf", sep=""))
+        pdf(paste(new_folder, "bar", row , ".pdf", sep=""), width=4, height=1)
         #margin settings
         par(mar=c(0,1,0,1))
         #par plot
         barplot(table_to_print[ row ,4], horiz=TRUE, xlim=c(1,100), xaxt="n")
         #threshold lines
         for(line in 1:length(threshold)) {
-            abline(v=threshold[line], lty=2,lwd=5)
+            abline(v=threshold[line], lty=3,lwd=1)
         }
         dev.off()
     }
 
     #Replacing percentage column by the graphs links
-    bar_template<-"\\includegraphics[width=\\linewidth, height=0.25\\linewidth]{Table_figures/"
+    bar_template<-"\\includegraphics[width=0.20\\linewidth, height=0.05\\linewidth]{Table_figures/"
     for(row in 1:nrow(table_to_print)) {
         table_to_print[ row ,4]<-paste(bar_template, "bar", row ,".pdf}", sep="")
     }

### 2015-06-04 Manuscript/Missing_mammals-draft.tex
Minor edits to introduction

@@
 Unfortunately this missing data can lead to errors in phylogenetic inference; in fact, simulations show that the ability of the Total Evidence method to recover the correct phylogenetic topology decreases when there is a low overlap between morphological data in the living and fossil taxa \citep{GuillermeCooper}, regardless the overall amount of morphological data available for the fossils (or the amount of molecular data available for the living species).
-The effect of missing data on topology is greatest when living taxa have little morphological data.
+The effect of missing data on topology is greatest when living taxa have few morphological data.
 This is because (1) a fossil cannot branch in the correct clade if there is no overlapping morphological data in the clade; and (2) a fossil has a higher probability of branching within a clade with more morphological data available for living taxa, regardless of whether this is the correct clade or not \citep{GuillermeCooper}.
@@
 However, it is unclear how much morphological data for living taxa is actually available (i.e.
-already coded from living taxa specimens and deposited in phylogenetic matrices accessible online), and how this data is distributed.
+already coded from museum specimens and deposited in phylogenetic matrices accessible online), and how this data is distributed across clades.
 Intuitively, most people assume this kind of data has already been collected, but empirical data suggest otherwise (e.g. in \cite{ronquista2012,slaterphylogenetic2013,beckancient2014}).
 To investigate this further, we assess the amount of available morphological data for living mammals to determine whether sufficient data exists to build reliable Total Evidence phylogenies in this group.
-We collected cladistic data from @256 phylogenetic matrices available online and measured the proportion of cladistic data available for each mammalian order.
+We collected cladistic data (i.e. % NC: define cladistic data here
+) from @256 phylogenetic matrices available online and measured the proportion of cladistic data available for each mammalian order.
 %Additionally, if the available data is randomly distributed across clades, the topological errors should be less extreme than if the data is biased towards particulars groups.
-Additionally, because of the two topological caveats mentioned above, we determined whether the available cladistic data was phylogenetically overdispersed or clustered in the mammalian orders where data was missing.
-We find that available morphological data for living mammals is scarce but generally randomly distributed and therefore suggest that an effort be made in collecting and sharing this type of data to improve Total Evidence tip-dated phylogenies.
+Additionally, because missing morphological data in living species can influence tree topology as described above,%NC: Might still need to fix this sentence.
+"Topological caveats above" was too unclear.
+we determined whether the available cladistic data was phylogenetically overdispersed or clustered in the mammalian orders where data was missing. 
+We find that available morphological data for living mammals is scarce but generally randomly distributed across phylogenies. 
+We recommend that efforts be made to collect and share more cladistic data for living species to improve the accuracy of Total Evidence phylogenies.
 
+% NC: If there are word count problems we can ditch the last two sentences.
 %---------------------------------------------

### 2015-06-05 Manuscript/Missing_mammals-draft.tex
Edits to data collection

@@
 %---------------------------------------------
-
 \section{Materials and Methods}
 \subsection{Data collection and standardisation}
-We downloaded all cladistic matrices containing any living or/and fossil mammal taxa from three major public databases: Morphobank (\texttt{http://www.morphobank.org/}) \citep{morphobank}, Graeme Lloyd's website (\texttt{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce/cladistic-data}).
-We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
+We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases (accessed @th of July 2015): Morphobank \citep[\texttt{http://www.morphobank.org/};][]{morphobank}, Graeme Lloyd's website (\texttt{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce/cladistic-data}).
+We also performed a systematic Google Scholar search (accessed @th of July 2015) for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
 In total, we downloaded @256 % Number of matrices
-matrices containing unique @9411 % Number of unique OTUs
+matrices containing @9411 unique % Number of unique OTUs
 operational taxonomic units (OTUs).
-In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardized and ranged from specific specimen names to family level.
-When possible, we considered the OTUs at their lowest valid taxonomic level (species) but some OTUs where only at a higher valid taxonomic level (e.g. genus or family).
-Therefore for some orders, we sampled more genus than species (Table \ref{Table_morpho_taxa_proportion}).
+In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardized and ranged from specific specimen names %NC: Do you mean names of individual specimens or subspecies?
+to the family-level.
+Where possible, we considered OTUs at their lowest valid taxonomic level (i.e. species) but some OTUs were only valid at a higher taxonomic level (e.g. genus or family).
+Therefore for some orders, we sampled more genera than species (Table \ref{Table_morpho_taxa_proportion}).
 
-To select the lower valid taxonomic level, we standardised the taxonomy of our OTUs by fixing invalid binomial inputs to match with standard taxonomic nomenclature (e.g., \textit{H.
+To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H.
 sapiens} was transformed to \textit{Homo sapiens}).
-We designated as ``living'' all the OTUs that where either present in the phylogeny of Fritz \textit{et al}.
-\citep{FritzTree} or the taxonomy of Wilson \& Reeder \citep{wilson2005mammal}, and designated as ``fossil'' all the OTUs that were present in the Paleobiology database (\texttt{https://paleobiodb.org/}).
-For OTUs that did not appear in these three sources we first decomposed the name (i.e.
-\textit{Homo$\_$sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
+We designated as ``living'' all OTUs that were either present in the phylogeny of \citet{FritzTree} or the taxonomy of \citet{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\texttt{https://paleobiodb.org/}).
+For OTUs that did not appear in these three sources, we first decomposed the name (i.e.
+\textit{Homo sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
 Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; see Supplementary Material for more details).
 
-The number of character per matrices ranged from @6 to @4541.
-This range led to difficulties in comparing the available data between matrices because (1) small matrices have less chance of having character overlapping with other matrices \citep{wagner2000} and (2) small matrices are more likely to contain a higher proportion of specific characters non-applicable across large clades \citep{Brazeau2011} (e.g.
-``antler ramifications'' is a character that is only applicable to Cervidae not all mammals).
-Therefore we selected only matrices containing at least 100 cladistic characters per OTU.
-This arbitrary threshold was chosen to correspond with the number of characters used in \citep{GuillermeCooper} and \citep{harrisonamong-character2014}.
+The number of characters in each matrix ranged from @6 to @4541.
+Matrices with few characters % NC: matrices can be small in two ways - with few characters and with few species. Here you just mention few characters so I've made it more specific.
+are problematic when comparing available data among matrices because (1) they have less chance of having characters that overlap with those of other matrices \citep{wagner2000} and (2) they are more likely to contain a higher proportion of specific characters that are not-applicable across large clades \citep[][e.g.
+``antler ramifications'' is a character that is only applicable to Cervidae not all mammals]{Brazeau2011}.
+% NC: Is this true really? Does it always follow that if you've very few characters you also have very specific characters? Maybe so...
+Therefore we selected only matrices containing \textgreater 100 characters per OTU.
+This threshold was chosen to correspond with the number of characters used in \citet{GuillermeCooper} and \citet{harrisonamong-character2014}.
 
-We extracted @1422 unique living mammal OTUs from the @256 matrices with a minimum of 6 and a maximum of 4541 coded cladistic characters per OTU.
-After removing all the matrices with less than 100 cladistic characters, the number of extracted living mammals OTUs was down to @815 from @number matrices.
+After removing all matrices with \textless 100 characters, we retained @1422 unique living mammal OTUs from @256 matrices for our analyses.
 

### 2015-06-05 Manuscript/Table_results.tex
Added the fable to the paper

% latex table generated in R 3.2.0 by xtable 1.7-4 package
% Fri Jun  5 15:29:14 2015
\begin{longtable}{lL{1.8cm}C{2cm}lcc}
\caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels. The coverage represents the proportion of taxa with available morphological data. The left vertical bar represents 25\% of available data (``low'' coverage if \textless 25\%); The right vertical bar represents 75\% of available data (``high'' coverage if \textgreater 75\%). When the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) are negative, taxa are more phylogenetically dispersed than expected by chance; when NRI or NTI are positive, taxa are more phylogenetically clustered by expected by chance. Significant NRI or NTI are highlighted in bold. One star (*) represents a p-value between 0.05 and 0.005; two starts between 0.005 and 0.0005 and three stars a p-value less than 0.0005.} \\ 
  \hline
Order & Taxonomic level & Proportion of taxa & Coverage & NRI & NTI \\ 
  \hline
Afrosoricida & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar1.pdf} &   &   \\ 
  Afrosoricida & genus & 17/17 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar2.pdf} &   &   \\ 
  Afrosoricida & species & 23/42 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar3.pdf} & 1.55 & -2.32 \\ 
  Carnivora & family & 11/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar4.pdf} & 0.52 & 1.6 \\ 
  \textbf{Carnivora} & \textbf{genus} & \textbf{30/125} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar5.pdf} & \textbf{4.01**} & \textbf{1.75*} \\ 
  \textbf{Carnivora} & \textbf{species} & \textbf{42/283} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar6.pdf} & \textbf{17.95**} & \textbf{3.1**} \\ 
  Cetartiodactyla & family & 20/21 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar7.pdf} & 0.35 & 1.11 \\ 
  Cetartiodactyla & genus & 76/128 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar8.pdf} & -3.22 & -1.12 \\ 
  Cetartiodactyla & species & 106/311 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar9.pdf} & -2.77 & -2.1 \\ 
  Chiroptera & family & 13/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar10.pdf} & 0.87 & 0.65 \\ 
  \textbf{Chiroptera} & \textbf{genus} & \textbf{68/202} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar11.pdf} & \textbf{18.87**} & \textbf{1.09} \\ 
  \textbf{Chiroptera} & \textbf{species} & \textbf{108/1054} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar12.pdf} & \textbf{20.11**} & \textbf{2.24*} \\ 
  Cingulata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar13.pdf} &   &   \\ 
  Cingulata & genus & 8/9 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar14.pdf} & 1.31 & -1.66 \\ 
  Cingulata & species & 6/25 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar15.pdf} & 0.65 & -0.47 \\ 
  Dasyuromorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar16.pdf} &   &   \\ 
  Dasyuromorphia & genus & 7/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar17.pdf} & -1.1 & -1.78 \\ 
  Dasyuromorphia & species & 8/64 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar18.pdf} & -1.1 & -1.31 \\ 
  Dermoptera & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar19.pdf} &   &   \\ 
  Dermoptera & genus & 1/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar20.pdf} &   &   \\ 
  Dermoptera & species & 1/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar21.pdf} &   &   \\ 
  Didelphimorphia & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar22.pdf} &   &   \\ 
  Didelphimorphia & genus & 16/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar23.pdf} &   &   \\ 
  Didelphimorphia & species & 40/84 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar24.pdf} & -1.33 & -1.7 \\ 
  Diprotodontia & family & 9/11 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar25.pdf} & -0.35 & 1.11 \\ 
  Diprotodontia & genus & 20/38 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar26.pdf} & -0.31 & -1.57 \\ 
  Diprotodontia & species & 16/126 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar27.pdf} & -0.98 & -2.6 \\ 
  Erinaceomorpha & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar28.pdf} &   &   \\ 
  Erinaceomorpha & genus & 10/10 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar29.pdf} &   &   \\ 
  Erinaceomorpha & species & 21/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar30.pdf} & -0.78 & -0.19 \\ 
  Hyracoidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar31.pdf} &   &   \\ 
  Hyracoidea & genus & 1/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar32.pdf} &   &   \\ 
  Hyracoidea & species & 1/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar33.pdf} &   &   \\ 
  Lagomorpha & family & 1/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar34.pdf} &   &   \\ 
  Lagomorpha & genus & 1/12 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar35.pdf} &   &   \\ 
  Lagomorpha & species & 1/86 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar36.pdf} &   &   \\ 
  Macroscelidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar37.pdf} &   &   \\ 
  Macroscelidea & genus & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar38.pdf} &   &   \\ 
  Macroscelidea & species & 5/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar39.pdf} & -0.47 & -0.89 \\ 
  Microbiotheria & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar40.pdf} &   &   \\ 
  Microbiotheria & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar41.pdf} &   &   \\ 
  Microbiotheria & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar42.pdf} &   &   \\ 
  Monotremata & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar43.pdf} &   &   \\ 
  Monotremata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar44.pdf} & -0.7 & -0.71 \\ 
  Monotremata & species & 2/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar45.pdf} & -0.97 & -1.01 \\ 
  Notoryctemorphia & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar46.pdf} &   &   \\ 
  Notoryctemorphia & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar47.pdf} &   &   \\ 
  Notoryctemorphia & species & 0/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar48.pdf} &   &   \\ 
  Paucituberculata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar49.pdf} &   &   \\ 
  Paucituberculata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar50.pdf} & -0.76 & -0.74 \\ 
  Paucituberculata & species & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar51.pdf} & -0.64 & -0.65 \\ 
  Peramelemorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar52.pdf} &   &   \\ 
  Peramelemorphia & genus & 7/7 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar53.pdf} &   &   \\ 
  Peramelemorphia & species & 16/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar54.pdf} & -0.55 & 0.49 \\ 
  Perissodactyla & family & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar55.pdf} &   &   \\ 
  Perissodactyla & genus & 6/6 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar56.pdf} &   &   \\ 
  Perissodactyla & species & 7/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar57.pdf} & 0.73 & -2.63 \\ 
  Pholidota & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar58.pdf} &   &   \\ 
  Pholidota & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar59.pdf} &   &   \\ 
  Pholidota & species & 3/8 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar60.pdf} & 1.14 & 0.66 \\ 
  Pilosa & family & 3/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar61.pdf} & -0.6 & -0.85 \\ 
  Pilosa & genus & 3/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar62.pdf} & -0.88 & -1.09 \\ 
  Pilosa & species & 3/29 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar63.pdf} & -1.51 & -1.47 \\ 
  Primates & family & 15/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar64.pdf} &   &   \\ 
  Primates & genus & 48/68 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar65.pdf} & -0.3 & -1.29 \\ 
  Primates & species & 56/351 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar66.pdf} & -1.5 & -2.18 \\ 
  Proboscidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar67.pdf} &   &   \\ 
  Proboscidea & genus & 1/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar68.pdf} &   &   \\ 
  Proboscidea & species & 1/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar69.pdf} &   &   \\ 
  Rodentia & family & 10/32 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar70.pdf} & 0.11 & -0.3 \\ 
  Rodentia & genus & 20/451 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar71.pdf} & -1.03 & 0.71 \\ 
  Rodentia & species & 10/2095 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar72.pdf} & -0.97 & -1.75 \\ 
  Scandentia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar73.pdf} &   &   \\ 
  Scandentia & genus & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar74.pdf} & -0.75 & -0.76 \\ 
  Scandentia & species & 2/20 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar75.pdf} & -1.26 & -1.27 \\ 
  Sirenia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar76.pdf} &   &   \\ 
  Sirenia & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar77.pdf} &   &   \\ 
  Sirenia & species & 2/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar78.pdf} & -0.96 & -0.94 \\ 
  Soricomorpha & family & 3/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar79.pdf} & -0.95 & -1.05 \\ 
  Soricomorpha & genus & 19/43 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar80.pdf} & 1.16 & -1.01 \\ 
  Soricomorpha & species & 19/392 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar81.pdf} & -1.87 & -3.09 \\ 
  Tubulidentata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar82.pdf} &   &   \\ 
  Tubulidentata & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar83.pdf} &   &   \\ 
  Tubulidentata & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar84.pdf} &   &   \\ 
   \hline
\hline
\label{Table_results}
\end{longtable}

### 2015-06-05 Analysis/Results_summary.R

@@ -1,14 +1,16 @@
 #Loading the functions
-setwd("~/PhD/Projects/Missing_living_mammals/Analysis")
+#Be sure to be in the right repository
 source("functions.R")
 load.functions(test=FALSE) #Set test=FALSE to speed up the loading
 library(picante)
 library(xtable)
 
 #Loading the tree
-full_trees<-read.nexus("../Data/Trees/FritzTree.rs200k.100trees.tre")
+full_trees<-read.nexus("../Data/Trees/mammalST_MSW05_best_chrono.tre")
+#Resolve polytomies
+one_tree<-multi2di(full_trees)
 #Selecting one random tree
-set.seed(1) ; one_tree<-full_trees[[sample(1:100, 1)]]
+#set.seed(1) ; one_tree<-full_trees[[sample(1:100, 1)]]
 
 #Loading the taxonomic reference list
 #Loading the taxonomic reference list
@@ -32,9 +34,9 @@ extraction_table[which(extraction_table$Matrix == "GS2012-SM.nex"), 3]<-157
 extraction_table1<-extraction_table[which(as.numeric(extraction_table$Characters) >= 1),]
 #75 Characters
 extraction_table75<-extraction_table[which(as.numeric(extraction_table$Characters) >= 75),]
-#94 Characters
+#100 Characters
 extraction_table100<-extraction_table[which(as.numeric(extraction_table$Characters) >= 100),]
-#138 Characters
+#125 Characters
 extraction_table125<-extraction_table[which(as.numeric(extraction_table$Characters) >= 125),]
 #150 Characters
 extraction_table150<-extraction_table[which(as.numeric(extraction_table$Characters) >= 150),]
@@ -61,34 +63,35 @@ load("results_100.Rda")
 ############################################
 #Table with the number of taxa containing morphological data
 ############################################
-caption<-"Proportion of available OTUs with morphological data per order and per taxonomic level. We highlighted in bold the orders that have more than 75\\% of missing data for each taxonomic level. Note that it is possible that more data is available at a higher taxonomic level (Genus $>$ Species) since if the species name for an OTU was not or miss specified, we still counted the OTU for higher taxonomic level analysis."
-summary.results(results, metric="proportion", save.path="../Manuscript/", file.save="Table_morpho_taxa_proportion", caption=caption, environement="longtable")
[…]

### 2015-06-05 Functions/table.results.R

@@ -4,32 +4,26 @@
 #<results> a result table from previous functions
 #<metric> which metric to print
 #<thresholds> threshold values (default=c(25,75))
-#<split> whether to split the table in smaller ones (environement will be ignored if split is not null)
 #<save.path> where to save the table
 #<file.save> name (or chain name) for saving the table
 #<caption> table caption (if split option is called, the caption will only appear for the first table)
 #<environement> table environement
 ##########################
-#environement(at)tcd.ie - 27/05/2015
+#environement(at)tcd.ie - 05/06/2015
 ##########################
 
-table.result<-function(results, metric, threshold=c(25,75), split=NULL, save.path, file.save, caption, environement) {
+table.result<-function(results, metric, threshold=c(25,75), save.path, file.save, caption, environement) {
     
     #DEBUG
-    warning("Debug mode is ON (table.result)")
+    #warning("Debug mode is ON (table.result)")
     #results -> must be a table
-    metric=c("NRI", "NTI") #-> can be a list (or 0) but must be present in the results header
-    threshold=c(25,75) #-> can be any values (or NULL) between 0 and 100
-    split=c(24,30,30) #-> can be a list of values (sub-tables) or NULL
-    save.path="../Manuscript/" #-> must be text
-    file.save="Tab_test" #-> must be text
-    caption="Here's my caption" #-> must be text
-    environement="tabular" #if split != null set to table. Else must be text
-
-    #barplot
-    #percentages<-table(as.numeric(results_tmp$Percentage.of.OTUs))
-    #names(percentages) <- rep(c("Family", "Genera", "Species"), length(order))
-    #percentages[]<-as.numeric(results_tmp$Percentage.of.OTUs)
+    #metric=c("NRI", "NTI") #-> can be a list (or 0) but must be present in the results header
+    #threshold=c(25,75) #-> can be any values (or NULL) between 0 and 100
+    #split=c(24,30,30) #-> can be a list of values (sub-tables) or NULL
+    #save.path="../Manuscript/" #-> must be text
+    #file.save="Tab_test" #-> must be text
+    #caption="Here's my caption" #-> must be text
+    #environement="longtable" #if split != null set to table. Else must be text
 
     #Selecting the metric columns
     metric_col<-NULL
[…]

### 2015-06-05 Manuscript/Missing_mammals-draft.tex
Added the fable to the paper
Edits to data distribution section

@@
 \subsection{Data availability and distribution}
-To assess the availability of cladistic data for each mammalian Order, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
-We used these different taxonomic levels because some clades are well covered at the family or genus level, but poorly covered at the species level.
-We consider Orders with \textless 25\% of living taxa with cladistic data as having poor data coverage (hereafter ``low'' coverage), and Orders with up \textgreater 75\% of living taxa with cladistic data as having good data coverage (hereafter ``high'' coverage).
+To assess the availability of cladistic data for each mammalian order, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
+We used these different taxonomic levels because some clades are well covered at the family- or genus-level, but poorly covered at the species-level.
+We consider orders with \textless 25\% of living taxa with cladistic data as having poor data coverage (hereafter ``low'' coverage), and orders with \textgreater 75\% of living taxa with cladistic data as having good data coverage (hereafter ``high'' coverage).
 
-For Orders with \textless 100\% cladistic data coverage at any taxonomic level, we investigated whether the available cladistic data was (i) randomly distributed, (ii) over-dispersed or (iii) clustered with respect to phylogeny, using the Nearest Taxon Index (NTI) metric \citep{webb2002phylogenies} in the \texttt{picante R} package \citep{picante,R}.
-The NTI is based on the mean nearest neighbour distance (MNND) and is calculated as follow:
+For orders with \textless 100\% cladistic data coverage at any taxonomic level, we investigated whether the available cladistic data was (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: Nearest Taxon Index \citep[NTI;][]{webb2002phylogenies} and Net Relatedness Index \citep[NRI;][]{webb2002phylogenies}.
+%*** see end of methods for explanation.
+Both metrics were calculated using the \texttt{picante} package in R \citep{picante,R}.
+% NC: It's not going to add much to the wordcount to explain both. But move PD to supp?
+
+% NC: Note that you had mis-defined MNND
+NTI \citep{webb2002phylogenies} is based on mean nearest neighbour distance (MNND) and is calculated as follows:
   \begin{equation}
-NTI=-(\frac{\overline{MNND}_{obs}-\overline{MNND}_{n}}{\sigma(MNND_{n})})
+    NTI=-\left(\frac{\overline{MNND}_{obs}-\overline{MNND}_{n}}{\sigma(MNND_{n})}\right)
   \end{equation}
-Where $\overline{MNND}_{obs}$ is the mean distance between each $n$ taxa with available cladistic data within each order; $\overline{MNND}_{n}$ is the expected random MNND for $n$ taxa estimated by calculating the MNND from $n$ randomly drawn taxa repeated 1000 times and $\sigma(MNND_{n})$ is the standard deviation of the 1000 random MNND$_{n}$.
-Negative NTI values shows more over-dispersed data than expected by chance (random) and positive values are showing more clustered data \citep{webb2002phylogenies}.
-Similarly, we also calculated the Nearest Relatedness Index (NRI) \citep{webb2002phylogenies,Cooper2012} and the Faith's phylogenetic distance (PD) \citep{Faith19921}.
-Both NRI and PD are calculated in the same way as the NTI (\((observed-\mu(random))/\sigma(random)\)) but they use respectively the mean phylogenetic distance (MPD) and the sum of branch lengths (SBL) instead of the MNND \citep{Faith19921,webb2002phylogenies,Cooper2012}.
+where $\overline{MNND}_{obs}$ is the observed mean distance among $n$ focal taxa and their nearest neighbour in the phylogeny,
+% NC: Sorry this still isn't quite right. Check my paper for a nicer way of writing it.
+$\overline{MNND}_{n}$ is the expected random MNND for $n$ taxa estimated by calculating the MNND from $n$ taxa randomly drawn from the phylogeny and repeated 1000 times, and $\sigma(MNND_{n})$ is the standard deviation of the 1000 random MNND values.
+NRI is similar but is based on mean phylogenetic distance (MPD) as follows:
+  \begin{equation}
+    NRI=-\left(\frac{\overline{MPD}_{obs}-\overline{MPD}_{n}}{\sigma(MPD_{n})}\right)
+  \end{equation}
+where $\overline{MPD}_{obs}$ is the observed mean phylogenetic distance among $n$ focal taxa, $\overline{MPD}_{n}$ is the expected random MPD for $n$ taxa estimated by calculating the MPD from $n$ taxa randomly drawn from the phylogeny and repeated 1000 times, and $\sigma(MPD_{n})$ is the standard deviation of the 1000 random MPD values.
+Negative NTI and NRI values show that the focal taxa are more overdispersed across the phylogeny than expected by chance, and positive values show that the focal taxa are more clustered across the phylogeny than expected by chance.
+%*** see end of methods for explanation.
+We calculated NTI and NRI values for each mammalian order separately, at each different taxonomic level. 
+For each analysis our focal taxa were those with available cladistic data at that taxonomic level and the phylogeny was the phylogeny of the order pruned from \citep{}. % cite  Bininda Emonds?
+% NC: Tried to make this more specific here. It's important to note that you used the order not the whole mammal phylogeny each time.
 
-Because the aim of this study is to understand the structure of the available cladistic data in mammals we prefered the NTI over the NRI or the PD metrics.
-In fact, the NTI metric is more sensitive to the phylogenetic structure (clustering or over-dispersion) at the tips of phylogeny \citep{Cooper2008}.
+%Because the aim of this study is to understand the structure of the available cladistic data in mammals we prefered the NTI over the NRI or the PD metrics.
+%In fact, the NTI metric is more sensitive to the phylogenetic structure (clustering or over-dispersion) at the tips of phylogeny \citep{Cooper2008}.
 %and also because previous studies have shown that NRI is biased towards detecting overdispersion \citep{Kembel2006,Swenson2006}. TG: Useless in this case?
-However, the results of the NRI and PD calculations are available in the supplementary.
+%However, the results of the NRI and PD calculations are available in the supplementary.
+
+% NC: Actually I think we discussed that NRI and NTI show different things?
+So it's worth including both (you can get the wording from my 2008 paper).
+This can go up with the intro to the metrics somewhere (I have *** the two places it might fit)
+
+We also report the results using another metric, Phylogenetic Distance \citep[PD;][]{Faith19921}, in the Supplementary Information. % link?
 
@@
 In the @@@ cladistic matrices extracted in our searches, @11 out of @28 mammalian orders have a ``low'' coverage (\textless 25\%) of species with cladistic data and @24 have a ``high'' coverage (\textgreater 75\%).
-At the genus level, only @three orders have a ``low'' coverage of taxa with cladistic data and @16 have a ``high'' coverage.
-Finally, at the family level @no order has a ``low'' coverage and only @5 have a ``high'' coverage (Table \ref{Table_morpho_taxa_proportion}).
+At the genus-level, only @three orders have a ``low'' coverage of taxa with cladistic data and @16 have a ``high'' coverage.
+Finally, at the family-level @no order has a ``low'' coverage and only @5 have a ``high'' coverage (Table \ref{Table_morpho_taxa_proportion}).
 
@@
 
-Among the mammalian orders containing OTUs with no available cladistic data, only @two orders, Carnivora and Chiroptera, are significantly phylogenetically clustered both at the species and the genus level (Table~\ref{Table_data_structure}).
+Among the mammalian orders containing OTUs with no available cladistic data, only @two orders, Carnivora and Chiroptera, are significantly phylogenetically clustered both at the species- and the genus-level (Table~\ref{Table_data_structure}).
 
@@
 Because of this difference between neontological species (\textit{sensu} reproductive isolates) and paleontological species (\textit{sensu} morpho-species %or "morphological clusters"?
-), most palaeontological studies are using the genus level for their smallest OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
+), most palaeontological studies are using the genus-level for their smallest OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
 In this frame of palaeontological data usage, the data availability at the genus-level in living mammals should be our priority concern.

### 2015-06-05 Manuscript/Missing_mammals-draft.tex
Edits to results

@@
 \section{Results}
-In the @@@ cladistic matrices extracted in our searches, @11 out of @28 mammalian orders have a ``low'' coverage (\textless 25\%) of species with cladistic data and @24 have a ``high'' coverage (\textgreater 75\%).
-At the genus-level, only @three orders have a ``low'' coverage of taxa with cladistic data and @16 have a ``high'' coverage.
-Finally, at the family-level @no order has a ``low'' coverage and only @5 have a ``high'' coverage (Table \ref{Table_morpho_taxa_proportion}).
+Across the @@@ cladistic matrices we extracted, @11 out of @28 mammalian orders have low coverage (\textless 25\% of species with cladistic data) and @24 have high coverage (\textgreater 75\% of species with cladistic data) at the species-level.
+At the genus-level, only @three orders have low coverage of taxa and @16 have high coverage, and at the family-level, @no orders have low coverage and @5 have high coverage (Table \ref{Table_morpho_taxa_proportion}).
 
@@
 
-Among the mammalian orders containing OTUs with no available cladistic data, only @two orders, Carnivora and Chiroptera, are significantly phylogenetically clustered both at the species- and the genus-level (Table~\ref{Table_data_structure}).
+Among the mammalian orders containing OTUs with no available cladistic data, only @two orders, Carnivora and Chiroptera, have OTUs with cladistic data that are significantly phylogenetically clustered at the species- and the genus-level (Table~\ref{Table_data_structure}).
 
@@
 
-Two contrasting results are shown on Figure \ref{Figure_example_coverage} with randomly distributed data in Cetartiodactyla (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered data in Carnivora (mainly Canidae; Figure \ref{Figure_example_coverage}B).
+Two contrasting results are shown in Figure \ref{Figure_example_coverage} with randomly distributed OTUs with cladistic data in Cetartiodactyla (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered OTUs with cladistic data in Carnivora (mainly Canidae; Figure \ref{Figure_example_coverage}B).
 

### 2015-06-05 Manuscript/Missing_mammals-draft.tex
Edits to figure legend

@@
     \includegraphics[width=1\textwidth]{example_coverage.pdf}
-\caption{Phylogenetic distribution of available cladistic data across two mammalian orders (A: Cetartiodactyla; B: Carnivora).
-Edges are colored in grey when no cladistic data is available or in red when data is available.}
+\caption{Phylogenetic distribution of species with available cladistic data across two mammalian orders (A: Cetartiodactyla; B: Carnivora).
+Edges are colored in grey when no cladistic data is available for a species or in blue when data is available.}
 \label{Figure_example_coverage}
 \end{figure}
-% NC: Can we change the colours, red sucks. Otherwise I like the figure a lot.
-% TG: Again, will fix in the final version.
 

### 2015-06-05 Manuscript/Missing_mammals-draft.tex
Need reverting to commit 61ff572

--- 2015-06-05 Manuscript/Missing_mammals-draft.tex
@@
 \pagenumbering{arabic}
+
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 
@@
 Additionally, large chunks of this data can be difficult, or even impossible, to collect for every taxon present in the analysis.
-For example, fossils very rarely have molecular data and incomplete fossil preservation may restrict the amount of morphological data available \citep{sansomfossilization2013}.
+For example, fossils very rarely have molecular data and incomplete fossil preservation (e.g.
+soft \textit{vs.} hard tissues) may restrict the amount of morphological data available \citep{sansomfossilization2013}.
 % NC: Worth metioning fossil preservation here briefly as it is also an issue?
@@
 Unfortunately this missing data can lead to errors in phylogenetic inference; in fact, simulations show that the ability of the Total Evidence method to recover the correct phylogenetic topology decreases when there is a low overlap between morphological data in the living and fossil taxa \citep{GuillermeCooper}, regardless the overall amount of morphological data available for the fossils (or the amount of molecular data available for the living species).
-The effect of missing data on topology is greatest when living taxa have few morphological data.
+The effect of missing data on topology is greatest when living taxa have little morphological data.
 This is because (1) a fossil cannot branch in the correct clade if there is no overlapping morphological data in the clade; and (2) a fossil has a higher probability of branching within a clade with more morphological data available for living taxa, regardless of whether this is the correct clade or not \citep{GuillermeCooper}.
@@
 However, it is unclear how much morphological data for living taxa is actually available (i.e.
-already coded from museum specimens and deposited in phylogenetic matrices accessible online), and how this data is distributed across clades.
+already coded from living taxa specimens and deposited in phylogenetic matrices accessible online), and how this data is distributed.
 Intuitively, most people assume this kind of data has already been collected, but empirical data suggest otherwise (e.g. in \cite{ronquista2012,slaterphylogenetic2013,beckancient2014}).
 To investigate this further, we assess the amount of available morphological data for living mammals to determine whether sufficient data exists to build reliable Total Evidence phylogenies in this group.
-We collected cladistic data (i.e. % NC: define cladistic data here
-) from @256 phylogenetic matrices available online and measured the proportion of cladistic data available for each mammalian order.
+We collected cladistic data from @256 phylogenetic matrices available online and measured the proportion of cladistic data available for each mammalian order.
 %Additionally, if the available data is randomly distributed across clades, the topological errors should be less extreme than if the data is biased towards particulars groups.
-Additionally, because missing morphological data in living species can influence tree topology as described above,%NC: Might still need to fix this sentence.
-"Topological caveats above" was too unclear.
-we determined whether the available cladistic data was phylogenetically overdispersed or clustered in the mammalian orders where data was missing. 
-We find that available morphological data for living mammals is scarce but generally randomly distributed across phylogenies. 
-We recommend that efforts be made to collect and share more cladistic data for living species to improve the accuracy of Total Evidence phylogenies.
+Additionally, because of the two topological caveats mentioned above, we determined whether the available cladistic data was phylogenetically overdispersed or clustered in the mammalian orders where data was missing.
+We find that available morphological data for living mammals is scarce but generally randomly distributed and therefore suggest that an effort be made in collecting and sharing this type of data to improve Total Evidence tip-dated phylogenies.
 
-% NC: If there are word count problems we can ditch the last two sentences.
 %---------------------------------------------
@@
 %---------------------------------------------
+
 \section{Materials and Methods}
 \subsection{Data collection and standardisation}
-We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases (accessed @th of July 2015): Morphobank \citep[\texttt{http://www.morphobank.org/};][]{morphobank}, Graeme Lloyd's website (\texttt{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce/cladistic-data}).
-We also performed a systematic Google Scholar search (accessed @th of July 2015) for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
+We downloaded all cladistic matrices containing any living or/and fossil mammal taxa from three major public databases: Morphobank (\texttt{http://www.morphobank.org/}) \citep{morphobank}, Graeme Lloyd's website (\texttt{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce/cladistic-data}).
+We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
 In total, we downloaded @256 % Number of matrices
-matrices containing @9411 unique % Number of unique OTUs
+matrices containing unique @9411 % Number of unique OTUs
 operational taxonomic units (OTUs).
-In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardized and ranged from specific specimen names %NC: Do you mean names of individual specimens or subspecies?
-to the family-level.
-Where possible, we considered OTUs at their lowest valid taxonomic level (i.e. species) but some OTUs were only valid at a higher taxonomic level (e.g. genus or family).
-Therefore for some orders, we sampled more genera than species (Table \ref{Table_morpho_taxa_proportion}).
+In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardized and ranged from specific specimen names to family level.
+When possible, we considered the OTUs at their lowest valid taxonomic level (species) but some OTUs where only at a higher valid taxonomic level (e.g. genus or family).
+Therefore for some orders, we sampled more genus than species (Table \ref{Table_morpho_taxa_proportion}).
 
-To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H.
+To select the lower valid taxonomic level, we standardised the taxonomy of our OTUs by fixing invalid binomial inputs to match with standard taxonomic nomenclature (e.g., \textit{H.
 sapiens} was transformed to \textit{Homo sapiens}).
-We designated as ``living'' all OTUs that were either present in the phylogeny of \citet{FritzTree} or the taxonomy of \citet{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\texttt{https://paleobiodb.org/}).
-For OTUs that did not appear in these three sources, we first decomposed the name (i.e.
-\textit{Homo sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
+We designated as ``living'' all the OTUs that where either present in the phylogeny of Fritz \textit{et al}.
+\citep{FritzTree} or the taxonomy of Wilson \& Reeder \citep{wilson2005mammal}, and designated as ``fossil'' all the OTUs that were present in the Paleobiology database (\texttt{https://paleobiodb.org/}).
+For OTUs that did not appear in these three sources we first decomposed the name (i.e.
+\textit{Homo$\_$sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
 Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; see Supplementary Material for more details).
 
-The number of characters in each matrix ranged from @6 to @4541.
-Matrices with few characters % NC: matrices can be small in two ways - with few characters and with few species. Here you just mention few characters so I've made it more specific.
-are problematic when comparing available data among matrices because (1) they have less chance of having characters that overlap with those of other matrices \citep{wagner2000} and (2) they are more likely to contain a higher proportion of specific characters that are not-applicable across large clades \citep[][e.g.
-``antler ramifications'' is a character that is only applicable to Cervidae not all mammals]{Brazeau2011}.
-% NC: Is this true really? Does it always follow that if you've very few characters you also have very specific characters? Maybe so...
-Therefore we selected only matrices containing \textgreater 100 characters per OTU.
-This threshold was chosen to correspond with the number of characters used in \citet{GuillermeCooper} and \citet{harrisonamong-character2014}.
+The number of character per matrices ranged from @6 to @4541.
+This range led to difficulties in comparing the available data between matrices because (1) small matrices have less chance of having character overlapping with other matrices \citep{wagner2000} and (2) small matrices are more likely to contain a higher proportion of specific characters non-applicable across large clades \citep{Brazeau2011} (e.g.
+``antler ramifications'' is a character that is only applicable to Cervidae not all mammals).
+Therefore we selected only matrices containing at least 100 cladistic characters per OTU.
+This arbitrary threshold was chosen to correspond with the number of characters used in \citep{GuillermeCooper} and \citep{harrisonamong-character2014}.
 
-After removing all matrices with \textless 100 characters, we retained @1422 unique living mammal OTUs from @256 matrices for our analyses.
+We extracted @1422 unique living mammal OTUs from the @256 matrices with a minimum of 6 and a maximum of 4541 coded cladistic characters per OTU.
+After removing all the matrices with less than 100 cladistic characters, the number of extracted living mammals OTUs was down to @815 from @number matrices.
 
 \subsection{Data availability and distribution}
-To assess the availability of cladistic data for each mammalian order, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
-We used these different taxonomic levels because some clades are well covered at the family- or genus-level, but poorly covered at the species-level.
-We consider orders with \textless 25\% of living taxa with cladistic data as having poor data coverage (hereafter ``low'' coverage), and orders with \textgreater 75\% of living taxa with cladistic data as having good data coverage (hereafter ``high'' coverage).
+To assess the availability of cladistic data for each mammalian Order, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
+We used these different taxonomic levels because some clades are well covered at the family or genus level, but poorly covered at the species level.
+We consider Orders with \textless 25\% of living taxa with cladistic data as having poor data coverage (hereafter ``low'' coverage), and Orders with up \textgreater 75\% of living taxa with cladistic data as having good data coverage (hereafter ``high'' coverage).
 
-For orders with \textless 100\% cladistic data coverage at any taxonomic level, we investigated whether the available cladistic data was (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: Nearest Taxon Index \citep[NTI;][]{webb2002phylogenies} and Net Relatedness Index \citep[NRI;][]{webb2002phylogenies}.
-%*** see end of methods for explanation.
-Both metrics were calculated using the \texttt{picante} package in R \citep{picante,R}.
-% NC: It's not going to add much to the wordcount to explain both. But move PD to supp?
+For Orders with \textless 100\% cladistic data coverage at any taxonomic level, we investigated whether the available cladistic data was (i) randomly distributed, (ii) over-dispersed or (iii) clustered with respect to phylogeny, using the Nearest Taxon Index (NTI) metric \citep{webb2002phylogenies} in the \texttt{picante R} package \citep{picante,R}.
+The NTI is based on the mean nearest neighbour distance (MNND) and is calculated as follow:
+\begin{equation}
+NTI=-(\frac{\overline{MNND}_{obs}-\overline{MNND}_{n}}{\sigma(MNND_{n})})
+\end{equation}
+Where $\overline{MNND}_{obs}$ is the mean distance between each $n$ taxa with available cladistic data within each order; $\overline{MNND}_{n}$ is the expected random MNND for $n$ taxa estimated by calculating the MNND from $n$ randomly drawn taxa repeated 1000 times and $\sigma(MNND_{n})$ is the standard deviation of the 1000 random MNND$_{n}$.
+Negative NTI values shows more over-dispersed data than expected by chance (random) and positive values are showing more clustered data \citep{webb2002phylogenies}.
+Similarly, we also calculated the Nearest Relatedness Index (NRI) \citep{webb2002phylogenies,Cooper2012} and the Faith's phylogenetic distance (PD) \citep{Faith19921}.
+Both NRI and PD are calculated in the same way as the NTI (\((observed-\mu(random))/\sigma(random)\)) but they use respectively the mean phylogenetic distance (MPD) and the sum of branch lengths (SBL) instead of the MNND \citep{Faith19921,webb2002phylogenies,Cooper2012}.
 
-% NC: Note that you had mis-defined MNND
-NTI \citep{webb2002phylogenies} is based on mean nearest neighbour distance (MNND) and is calculated as follows:
-  \begin{equation}
-    NTI=-\left(\frac{\overline{MNND}_{obs}-\overline{MNND}_{n}}{\sigma(MNND_{n})}\right)
-  \end{equation}
-where $\overline{MNND}_{obs}$ is the observed mean distance among $n$ focal taxa and their nearest neighbour in the phylogeny,
-% NC: Sorry this still isn't quite right. Check my paper for a nicer way of writing it.
-$\overline{MNND}_{n}$ is the expected random MNND for $n$ taxa estimated by calculating the MNND from $n$ taxa randomly drawn from the phylogeny and repeated 1000 times, and $\sigma(MNND_{n})$ is the standard deviation of the 1000 random MNND values.
-NRI is similar but is based on mean phylogenetic distance (MPD) as follows:
-  \begin{equation}
-    NRI=-\left(\frac{\overline{MPD}_{obs}-\overline{MPD}_{n}}{\sigma(MPD_{n})}\right)
-  \end{equation}
-where $\overline{MPD}_{obs}$ is the observed mean phylogenetic distance among $n$ focal taxa, $\overline{MPD}_{n}$ is the expected random MPD for $n$ taxa estimated by calculating the MPD from $n$ taxa randomly drawn from the phylogeny and repeated 1000 times, and $\sigma(MPD_{n})$ is the standard deviation of the 1000 random MPD values.
-Negative NTI and NRI values show that the focal taxa are more overdispersed across the phylogeny than expected by chance, and positive values show that the focal taxa are more clustered across the phylogeny than expected by chance.
-%*** see end of methods for explanation.
-We calculated NTI and NRI values for each mammalian order separately, at each different taxonomic level. 
-For each analysis our focal taxa were those with available cladistic data at that taxonomic level and the phylogeny was the phylogeny of the order pruned from \citep{}. % cite  Bininda Emonds?
-% NC: Tried to make this more specific here. It's important to note that you used the order not the whole mammal phylogeny each time.
-
-%Because the aim of this study is to understand the structure of the available cladistic data in mammals we prefered the NTI over the NRI or the PD metrics.
-%In fact, the NTI metric is more sensitive to the phylogenetic structure (clustering or over-dispersion) at the tips of phylogeny \citep{Cooper2008}.
+Because the aim of this study is to understand the structure of the available cladistic data in mammals we prefered the NTI over the NRI or the PD metrics.
+In fact, the NTI metric is more sensitive to the phylogenetic structure (clustering or over-dispersion) at the tips of phylogeny \citep{Cooper2008}.
 %and also because previous studies have shown that NRI is biased towards detecting overdispersion \citep{Kembel2006,Swenson2006}. TG: Useless in this case?
-%However, the results of the NRI and PD calculations are available in the supplementary.
-
-% NC: Actually I think we discussed that NRI and NTI show different things?
-So it's worth including both (you can get the wording from my 2008 paper).
-This can go up with the intro to the metrics somewhere (I have *** the two places it might fit)
-
-We also report the results using another metric, Phylogenetic Distance \citep[PD;][]{Faith19921}, in the Supplementary Information. % link?
+However, the results of the NRI and PD calculations are available in the supplementary.
 
@@
 In the @@@ cladistic matrices extracted in our searches, @11 out of @28 mammalian orders have a ``low'' coverage (\textless 25\%) of species with cladistic data and @24 have a ``high'' coverage (\textgreater 75\%).
-At the genus-level, only @three orders have a ``low'' coverage of taxa with cladistic data and @16 have a ``high'' coverage.
-Finally, at the family-level @no order has a ``low'' coverage and only @5 have a ``high'' coverage (Table \ref{Table_morpho_taxa_proportion}).
+At the genus level, only @three orders have a ``low'' coverage of taxa with cladistic data and @16 have a ``high'' coverage.
+Finally, at the family level @no order has a ``low'' coverage and only @5 have a ``high'' coverage (Table \ref{Table_morpho_taxa_proportion}).
 
 % TG: Same for the table, I will do the new table combinations later (need some proper coding since I don't wan't to use illustrator - my first paper 100\% reproducible by copy paste!)
-\renewcommand\baselinestretch{1.2}\selectfont
 \begin{center}
-\input{Table_morpho_taxa_proportion}
+\input{Table_results}
 \end{center}
-\renewcommand\baselinestretch{2}\selectfont
 
-Among the mammalian orders containing OTUs with no available cladistic data, only @two orders, Carnivora and Chiroptera, are significantly phylogenetically clustered both at the species- and the genus-level (Table~\ref{Table_data_structure}).
+%\renewcommand\baselinestretch{1.2}\selectfont
+%\begin{center}
+%\input{Table_morpho_taxa_proportion}
+%\end{center}
+%\renewcommand\baselinestretch{2}\selectfont
 
-\renewcommand\baselinestretch{1.2}\selectfont
-\begin{center}
-\input{Table_data_structure}
-\end{center}
-\renewcommand\baselinestretch{2}\selectfont
+Among the mammalian orders containing OTUs with no available cladistic data, only @two orders, Carnivora and Chiroptera, are significantly phylogenetically clustered both at the species and the genus level (Table~\ref{Table_data_structure}).
+
+%\renewcommand\baselinestretch{1.2}\selectfont
+%\begin{center}
+%\input{Table_data_structure}
+%\end{center}
+%\renewcommand\baselinestretch{2}\selectfont
 
@@
 Because of this difference between neontological species (\textit{sensu} reproductive isolates) and paleontological species (\textit{sensu} morpho-species %or "morphological clusters"?
-), most palaeontological studies are using the genus-level for their smallest OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
+), most palaeontological studies are using the genus level for their smallest OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
 In this frame of palaeontological data usage, the data availability at the genus-level in living mammals should be our priority concern.

### 2015-06-05 Analysis/Results_summary.R
Revert 61ff572..d3df3f6
This rolls back to commit 61ff572c36bc0f24b6b26566c3360addb9b0c61f.

@@ -1,16 +1,14 @@
 #Loading the functions
-#Be sure to be in the right repository
+setwd("~/PhD/Projects/Missing_living_mammals/Analysis")
 source("functions.R")
 load.functions(test=FALSE) #Set test=FALSE to speed up the loading
 library(picante)
 library(xtable)
 
 #Loading the tree
-full_trees<-read.nexus("../Data/Trees/mammalST_MSW05_best_chrono.tre")
-#Resolve polytomies
-one_tree<-multi2di(full_trees)
+full_trees<-read.nexus("../Data/Trees/FritzTree.rs200k.100trees.tre")
 #Selecting one random tree
-#set.seed(1) ; one_tree<-full_trees[[sample(1:100, 1)]]
+set.seed(1) ; one_tree<-full_trees[[sample(1:100, 1)]]
 
 #Loading the taxonomic reference list
 #Loading the taxonomic reference list
@@ -34,9 +32,9 @@ extraction_table[which(extraction_table$Matrix == "GS2012-SM.nex"), 3]<-157
 extraction_table1<-extraction_table[which(as.numeric(extraction_table$Characters) >= 1),]
 #75 Characters
 extraction_table75<-extraction_table[which(as.numeric(extraction_table$Characters) >= 75),]
-#100 Characters
+#94 Characters
 extraction_table100<-extraction_table[which(as.numeric(extraction_table$Characters) >= 100),]
-#125 Characters
+#138 Characters
 extraction_table125<-extraction_table[which(as.numeric(extraction_table$Characters) >= 125),]
 #150 Characters
 extraction_table150<-extraction_table[which(as.numeric(extraction_table$Characters) >= 150),]
@@ -63,35 +61,34 @@ load("results_100.Rda")
 ############################################
 #Table with the number of taxa containing morphological data
 ############################################
-#caption<-"Proportion of available OTUs with morphological data per order and per taxonomic level. We highlighted in bold the orders that have more than 75\\% of missing data for each taxonomic level. Note that it is possible that more data is available at a higher taxonomic level (Genus $>$ Species) since if the species name for an OTU was not or miss specified, we still counted the OTU for higher taxonomic level analysis."
-#summary.results(results, metric="proportion", save.path="../Manuscript/", file.save="testiiiiiiiing", caption=caption, environement="longtable")
[…]

### 2015-06-05 Functions/table.results.R

@@ -4,26 +4,32 @@
 #<results> a result table from previous functions
 #<metric> which metric to print
 #<thresholds> threshold values (default=c(25,75))
+#<split> whether to split the table in smaller ones (environement will be ignored if split is not null)
 #<save.path> where to save the table
 #<file.save> name (or chain name) for saving the table
 #<caption> table caption (if split option is called, the caption will only appear for the first table)
 #<environement> table environement
 ##########################
-#environement(at)tcd.ie - 05/06/2015
+#environement(at)tcd.ie - 27/05/2015
 ##########################
 
-table.result<-function(results, metric, threshold=c(25,75), save.path, file.save, caption, environement) {
+table.result<-function(results, metric, threshold=c(25,75), split=NULL, save.path, file.save, caption, environement) {
     
     #DEBUG
-    #warning("Debug mode is ON (table.result)")
+    warning("Debug mode is ON (table.result)")
     #results -> must be a table
-    #metric=c("NRI", "NTI") #-> can be a list (or 0) but must be present in the results header
-    #threshold=c(25,75) #-> can be any values (or NULL) between 0 and 100
-    #split=c(24,30,30) #-> can be a list of values (sub-tables) or NULL
-    #save.path="../Manuscript/" #-> must be text
-    #file.save="Tab_test" #-> must be text
-    #caption="Here's my caption" #-> must be text
-    #environement="longtable" #if split != null set to table. Else must be text
+    metric=c("NRI", "NTI") #-> can be a list (or 0) but must be present in the results header
+    threshold=c(25,75) #-> can be any values (or NULL) between 0 and 100
+    split=c(24,30,30) #-> can be a list of values (sub-tables) or NULL
+    save.path="../Manuscript/" #-> must be text
+    file.save="Tab_test" #-> must be text
+    caption="Here's my caption" #-> must be text
+    environement="tabular" #if split != null set to table. Else must be text
+
+    #barplot
+    #percentages<-table(as.numeric(results_tmp$Percentage.of.OTUs))
+    #names(percentages) <- rep(c("Family", "Genera", "Species"), length(order))
+    #percentages[]<-as.numeric(results_tmp$Percentage.of.OTUs)
 
     #Selecting the metric columns
     metric_col<-NULL
[…]

### 2015-06-05 Manuscript/Missing_mammals-draft.tex
Conflict. Revert again!
This reverts commit 72c89542624c8b4d63aaf5aff595cfd78b145d25, reversing
changes made to e84098918f974c70159d154cfe01ca39c0bd7869.

Conflicts:
	Manuscript/Missing_mammals-draft.tex

@@
 \pagenumbering{arabic}
-
-\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 
@@
 Additionally, large chunks of this data can be difficult, or even impossible, to collect for every taxon present in the analysis.
-For example, fossils very rarely have molecular data and incomplete fossil preservation (e.g.
-soft \textit{vs.} hard tissues) may restrict the amount of morphological data available \citep{sansomfossilization2013}.
+For example, fossils very rarely have molecular data and incomplete fossil preservation may restrict the amount of morphological data available \citep{sansomfossilization2013}.
 % NC: Worth metioning fossil preservation here briefly as it is also an issue?
@@
 % TG: Same for the table, I will do the new table combinations later (need some proper coding since I don't wan't to use illustrator - my first paper 100\% reproducible by copy paste!)
+\renewcommand\baselinestretch{1.2}\selectfont
 \begin{center}
-\input{Table_results}
+\input{Table_morpho_taxa_proportion}
 \end{center}
-
-%\renewcommand\baselinestretch{1.2}\selectfont
-%\begin{center}
-%\input{Table_morpho_taxa_proportion}
-%\end{center}
-%\renewcommand\baselinestretch{2}\selectfont
+\renewcommand\baselinestretch{2}\selectfont
 
 Among the mammalian orders containing OTUs with no available cladistic data, only @two orders, Carnivora and Chiroptera, are significantly phylogenetically clustered both at the species and the genus level (Table~\ref{Table_data_structure}).
-
-%\renewcommand\baselinestretch{1.2}\selectfont
-%\begin{center}
-%\input{Table_data_structure}
-%\end{center}
-%\renewcommand\baselinestretch{2}\selectfont
 

### 2015-06-05 Functions/table.results.R
Revert 72c8954..d3d9fcc
This rolls back to commit 72c89542624c8b4d63aaf5aff595cfd78b145d25.

@@ -4,32 +4,26 @@
 #<results> a result table from previous functions
 #<metric> which metric to print
 #<thresholds> threshold values (default=c(25,75))
-#<split> whether to split the table in smaller ones (environement will be ignored if split is not null)
 #<save.path> where to save the table
 #<file.save> name (or chain name) for saving the table
 #<caption> table caption (if split option is called, the caption will only appear for the first table)
 #<environement> table environement
 ##########################
-#environement(at)tcd.ie - 27/05/2015
+#environement(at)tcd.ie - 05/06/2015
 ##########################
 
-table.result<-function(results, metric, threshold=c(25,75), split=NULL, save.path, file.save, caption, environement) {
+table.result<-function(results, metric, threshold=c(25,75), save.path, file.save, caption, environement) {
     
     #DEBUG
-    warning("Debug mode is ON (table.result)")
+    #warning("Debug mode is ON (table.result)")
     #results -> must be a table
-    metric=c("NRI", "NTI") #-> can be a list (or 0) but must be present in the results header
-    threshold=c(25,75) #-> can be any values (or NULL) between 0 and 100
-    split=c(24,30,30) #-> can be a list of values (sub-tables) or NULL
-    save.path="../Manuscript/" #-> must be text
-    file.save="Tab_test" #-> must be text
-    caption="Here's my caption" #-> must be text
-    environement="tabular" #if split != null set to table. Else must be text
-
-    #barplot
-    #percentages<-table(as.numeric(results_tmp$Percentage.of.OTUs))
-    #names(percentages) <- rep(c("Family", "Genera", "Species"), length(order))
-    #percentages[]<-as.numeric(results_tmp$Percentage.of.OTUs)
+    #metric=c("NRI", "NTI") #-> can be a list (or 0) but must be present in the results header
+    #threshold=c(25,75) #-> can be any values (or NULL) between 0 and 100
+    #split=c(24,30,30) #-> can be a list of values (sub-tables) or NULL
+    #save.path="../Manuscript/" #-> must be text
+    #file.save="Tab_test" #-> must be text
+    #caption="Here's my caption" #-> must be text
+    #environement="longtable" #if split != null set to table. Else must be text
 
     #Selecting the metric columns
     metric_col<-NULL
[…]

### 2015-06-05 Manuscript/Missing_mammals-draft.tex
Revert 72c8954..d3d9fcc
This rolls back to commit 72c89542624c8b4d63aaf5aff595cfd78b145d25.
Minor edit to intro

--- 2015-06-05 Manuscript/Missing_mammals-draft.tex
@@
 % TG: Same for the table, I will do the new table combinations later (need some proper coding since I don't wan't to use illustrator - my first paper 100\% reproducible by copy paste!)
-\renewcommand\baselinestretch{1.2}\selectfont
 \begin{center}
-\input{Table_morpho_taxa_proportion}
+\input{Table_results}
 \end{center}
-\renewcommand\baselinestretch{2}\selectfont
+
+%\renewcommand\baselinestretch{1.2}\selectfont
+%\begin{center}
+%\input{Table_morpho_taxa_proportion}
+%\end{center}
+%\renewcommand\baselinestretch{2}\selectfont
 
@@
 
-\renewcommand\baselinestretch{1.2}\selectfont
-\begin{center}
-\input{Table_data_structure}
-\end{center}
-\renewcommand\baselinestretch{2}\selectfont
+%\renewcommand\baselinestretch{1.2}\selectfont
+%\begin{center}
+%\input{Table_data_structure}
+%\end{center}
+%\renewcommand\baselinestretch{2}\selectfont
 

### 2015-06-05 Manuscript/Tab_test.tex
Table testing

--- 2015-06-05 Manuscript/Table_results.tex
@@
 % latex table generated in R 3.2.0 by xtable 1.7-4 package
-% Fri Jun  5 15:29:14 2015
+% Fri Jun  5 15:07:20 2015
 \begin{longtable}{lL{1.8cm}C{2cm}lcc}
-\caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels.
-The coverage represents the proportion of taxa with available morphological data.
-The left vertical bar represents 25\% of available data (``low'' coverage if \textless 25\%); The right vertical bar represents 75\% of available data (``high'' coverage if \textgreater 75\%).
-When the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) are negative, taxa are more phylogenetically dispersed than expected by chance; when NRI or NTI are positive, taxa are more phylogenetically clustered by expected by chance.
-Significant NRI or NTI are highlighted in bold.
-One star (*) represents a p-value between 0.05 and 0.005; two starts between 0.005 and 0.0005 and three stars a p-value less than 0.0005.} \\
+\caption{Here's my caption} \\ 
   \hline
@@
 \hline
-\label{Table_results}
+\label{Tab_test}
 \end{longtable}

### 2015-06-05 Manuscript/Missing_mammals-draft.tex
Edits to discussion

@@
 %---------------------------------------------
-
-%NC: I've changed cladistic to morphological throughout but you might want to change it back...
-
 \newpage 
@@
 Additionally, large chunks of this data can be difficult, or even impossible, to collect for every taxon present in the analysis.
-For example, fossils very rarely have molecular data and incomplete fossil preservation (e.g.
-soft \textit{vs.} hard tissues) may restrict the amount of morphological data available \citep{sansomfossilization2013}.
-% NC: Worth metioning fossil preservation here briefly as it is also an issue?
+For example, fossils very rarely have molecular data and incomplete fossil preservation may restrict the amount of morphological data available \citep{sansomfossilization2013}. 
 Additionally, since the molecular phylogenetics revolution, it has become less common to collect morphological characters for living taxa when molecular data is available (e.g.
@@
 This is because (1) a fossil cannot branch in the correct clade if there is no overlapping morphological data in the clade; and (2) a fossil has a higher probability of branching within a clade with more morphological data available for living taxa, regardless of whether this is the correct clade or not \citep{GuillermeCooper}.
-%This can lead to serious topological errors (see Figure ~\ref{Figure_missing_data_problem}).
 
@@
 Both metrics were calculated using the \texttt{picante} package in R \citep{picante,R}.
-% NC: It's not going to add much to the wordcount to explain both. But move PD to supp?
+% NC: It's not going to add much to the wordcount to explain both. But move PD to supp
 
@@
 
-% TG: Same for the table, I will do the new table combinations later (need some proper coding since I don't wan't to use illustrator - my first paper 100\% reproducible by copy paste!)
-\begin{center}
 \input{Table_results}
-\end{center}
-
-%\renewcommand\baselinestretch{1.2}\selectfont
-%\begin{center}
-%\input{Table_morpho_taxa_proportion}
-%\end{center}
-%\renewcommand\baselinestretch{2}\selectfont
 
 Among the mammalian orders containing OTUs with no available cladistic data, only @two orders, Carnivora and Chiroptera, have OTUs with cladistic data that are significantly phylogenetically clustered at the species- and the genus-level (Table~\ref{Table_data_structure}).
-
-%\renewcommand\baselinestretch{1.2}\selectfont
-%\begin{center}
-%\input{Table_data_structure}
-%\end{center}
-%\renewcommand\baselinestretch{2}\selectfont
 
@@
 Our results show that although phylogenetic relationships among living mammals are well-resolved \citep[e.g.][]{FritzTree,meredithimpacts2011,May-Collado-PeerJ}, most of the data used to build these phylogenies is molecular, and very little cladistic data is available for living mammals compared to fossil mammals \citep[e.g.][]{O'Leary08022013,ni2013oldest}.
-This has implications for building Total Evidence phylogenies containing both living and fossil mammals as without sufficient cladistic data for living species, fossil placements in these trees is very uncertain \citep{GuillermeCooper}.
-However, data coverage in living mammals varies across taxonomic levels and in its phylogenetic distribution.
+This has implications for building Total Evidence phylogenies containing both living and fossil mammals, as without sufficient cladistic data for living species, fossil placements in these trees are very uncertain \citep{GuillermeCooper}.
+However, cladistic data coverage in living mammals varies across taxonomic levels and in its phylogenetic distribution.
 Higher taxonomic levels are always better sampled than lower ones and within these taxonomic levels, the available data is mostly randomly distributed across the phylogeny, apart from in Carnivora and Chiroptera.
+% NC: Add a sentence saying this means things might not be quite so bad? Need to justify the "However". however what?
 
-The number of living taxa with no available cladistic data was surprisingly high at the species-level: @24 out of @28 orders have a ``high'' coverage of taxa with available cladistic data (and two of the 28 orders are monospecific!).
-This threshold of 75\% available data (``high'' coverage) represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.
+The number of living mammalian taxa with no available cladistic data was surprisingly high at the species-level: only @24 out of @28 orders have a high coverage of taxa with available cladistic data (and two of the 28 orders are monospecific!).
+% NC: Bit odd to say it's suprisingly high then say there are lots of orders with "high" coverage.
+I've added an "only" but 24 of 28 sounds quite good to me...
+This high coverage threshold of 75\% of taxa with available cladistic data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.
 Beyond this threshold, there is considerable displacement of wildcard taxa (\textit{sensu} \citep{kearneyfragmentary2002}) and decreases in clade conservation \citep{GuillermeCooper}.
@@
 However, data coverage seems to be less of an issue at higher taxonomic levels (i.e. genus- and family-level).
-This point is important to consider from a practical point of view because of the slight discrepancy between the neontological and the palaeontological concept of species.
-While neontological species are described on accurately measured morphology, genetic distance, spatial distribution and even behaviour \citep[e.g.][]{kellymolecular2014}, palaeontological data can be based only on morphological, spatial and temporal data \citep[e.g.][]{ni2013oldest}.
-Because of this difference between neontological species (\textit{sensu} reproductive isolates) and paleontological species (\textit{sensu} morpho-species %or "morphological clusters"?
-), most palaeontological studies are using the genus-level for their smallest OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
-In this frame of palaeontological data usage, the data availability at the genus-level in living mammals should be our priority concern.
+This point is important from a practical point of view because of the slight discrepancy between the neontological and palaeontological concept of species.
+While neontological species are described using morphology, genetic distance, spatial distribution and even behaviour \citep[e.g.][]{kellymolecular2014}, palaeontological data can be based only on morphological, spatial and temporal data \citep[e.g.][]{ni2013oldest}.
+Because of this difference between neontological species (i.e.
+reproductive isolates) and paleontological species (i.e.
+morpho-species), most palaeontological studies are using the genus-level as their smallest OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
+Thus data availability at the genus-level in living mammals should be our primary concern when aiming to build phylogenies of living and fossil taxa.
 
-Regardless the low level of available data, it is encouraging that for most orders (except Carnivora and Chiroptera), data is randomly distributed across the phylogeny.
-There are three different scenarii, regarding the distribution of the available data, for predicting how fossils can branch in relation to living taxa.
-When only few data is available, the ideal scenario will be that it is over-dispersed (i.e.
-that there is data in at least every sub-clade) in order to maximize the possibilities of the fossil to branch in the right clade.
-In the second scenario, when the data is randomly distributed we expected no special bias in the placement of the fossil \citep{GuillermeCooper}.
-However, in the third scenario, when the data is clustered we expect two major biases to occur: (1) first, the fossil will not be able to branch within a clade containing no data (e.g.
-Herpestidae in Carnivora; Figure \ref{Figure_missing_data_problem} and \ref{Figure_example_coverage}B), and (2) second, the fossil will have a higher probability, at random, to branch within the clade containing most of the available data (e.g.
-a Carnivora fossil will have more chance to branch, randomly, to the Canidae clade than any other clade in Carnivora; Figure \ref{Figure_example_coverage}-B).
-This means that fossils with uncertain phylogentic affinities (\textit{Incertae sedis}) will have a higher probability of branching by chance within the most sampled clade.
-
-%For example in Carnivora, their is no available data for Herpestidae, therefore any fossil Herpestidae will not be able to be correctly placed in the phylogeny (figure ~\ref{Figure_example_coverage}-B.
-%On the other hand, any Carnivora fossil with uncertain phylogenetic affinities (\textit{Incertae sedis} will have a higher probability of branching by chance within the Canidae because they are oversampled within the Carnivora (figure ~\ref{Figure_example_coverage}-B.
+When only a few species with cladistic data are available, the ideal scenario is for them to be phylogenetically overdispersed (i.e.
+that there is data for at least every sub-clade) to maximize the possibilities of a fossil branching from the right clade.
+The second best scenario is that species with cladistic data are randomly distributed across the phylogeny. 
+In this scenario we expected no special bias in the placement of the fossil \citep{GuillermeCooper}, it is therefore encouraging that for all orders except Carnivora and Chiroptera, species with cladistic data were randomly distributed across the phylogeny of each order.
+The worst case scenario for fossil placement is that species with cladistic data are phylogenetically clustered. 
+In this situation we expect two major biases to occur: first, the fossil will not be able to branch within a clade containing no data, and second, the fossil will have a higher probability, at random, of branching within the clade containing most of the available data.
+This means that fossils with uncertain phylogenetic affinities (\textit{incertae sedis}) will have a higher probability of branching within the most sampled clade just by chance.
+Our results suggest that this may be an issue in Carnivora and Chiroptera. 
+For example, a Carnivora fossil will be unable to branch in the Herpestidae that has no species with cladistic data, and will also have more chance to branch, randomly, within the Canidae clade than any other clade in Carnivora (Figure \ref{Figure_example_coverage}-B).
+Thus placements of Carnivoran fossils should be viewed with suspicion. % NC: OK gotta be a better way to say that.
 
@@
 Unfortunately, cladistic data cannot always be treated in this way because some characters may overlap.
-For example, if matrix A has a character coding for the shape of a particular morphological feature and matrix B has a character coding for the presence of this morphological feature and a second character coding for its shape then these three characters are non-independent compound characters \citep{Brazeau2011}.
+For example, if matrix A has a character coding for the shape of a particular morphological feature and matrix B has a character coding for the presence of this morphological feature and a second character coding for its shape, then these three characters are non-independent compound characters \citep{Brazeau2011}.
 However, in reasonably sized matrices (\textgreater 100 characters; \citealp{GuillermeCooper,harrisonamong-character2014}) it is more likely that a number of characters are consistently conserved among the different matrices (e.g.
 within the Primate literature, the character \textit{p7} - the size of the $4^{th}$ lower premolar paraconid - has been used consistently for \textgreater 15 years; \citealp[e.g.][]{ross1998phylogenetic,seiffert2003fossil,marivaux2005anthropoid,seiffert2005basal,bloch2007new,kay2008anatomy,silcox2008biogeographic,seiffert2009convergent,tabuce2009anthropoid,boyer2010astragalar,seiffert2010fossil,marivaux2013djebelemur,ni2013oldest}).
-% NC: Remove some of these
+% NC: Remove some of these!!!
 A conservative approach to avoid compound characters would be to select only the most recent matrix for each group, but this would result in the loss of a lot of data.
@@
 Following the recommendations in \citep{GuillermeCooper}, to improve the accuracy of the topology of Total Evidence trees containing both living and fossil taxa, one should code cladistic characters for as many living species possible.
-Since the data for living mammals is usually easily available in world wide vast natural history collections, we propose that an increased effort in coding morphological characters from living species should be done via engaging in collaborative data collection projects through web portals such as \textit{Morphobank} \citep{morphobank}.
+Because the data for living mammals is usually readily available in natural history collections, we propose that an increased effort be put into coding morphological characters from living species, possibly by engaging in collaborative data collection projects through web portals such as \textit{Morphobank} \citep{morphobank}.
+
+% NC: OK needs a summing up sentence!!!
 

### 2015-06-05 Manuscript/Missing_mammals-draft.tex
Natalie's comments on the discussion

@@
 Additionally, large chunks of this data can be difficult, or even impossible, to collect for every taxon present in the analysis.
-For example, fossils very rarely have molecular data and incomplete fossil preservation may restrict the amount of morphological data available \citep{sansomfossilization2013}. 
+For example, fossils very rarely have molecular data and incomplete fossil preservation (e.g.
+soft \textit{vs.} hard tissues) may restrict the amount of morphological data available \citep{sansomfossilization2013}.
+% NC: Worth metioning fossil preservation here briefly as it is also an
 Additionally, since the molecular phylogenetics revolution, it has become less common to collect morphological characters for living taxa when molecular data is available (e.g.

### 2015-06-08 Manuscript/Missing_mammals-draft.tex
Error in counts

@@
 
-The number of living mammalian taxa with no available cladistic data was surprisingly high at the species-level: only @24 out of @28 orders have a high coverage of taxa with available cladistic data (and two of the 28 orders are monospecific!).
-% NC: Bit odd to say it's suprisingly high then say there are lots of orders with "high" coverage.
-I've added an "only" but 24 of 28 sounds quite good to me...
+The number of living mammalian taxa with no available cladistic data was surprisingly high at the species-level: only @4 out of @28 orders have a high coverage of taxa with available cladistic data (and two of the 28 orders are monospecific!).
 This high coverage threshold of 75\% of taxa with available cladistic data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.

### 2015-06-20 Functions/community.structure.R
Fixed rooting bug for PD calculation

@@ -41,10 +41,25 @@ community.structure<-function(data_structure, tree, metric=c("PD", "NRI", "NTI")
 
     #SES.PD
     if(any(metric == "PD")) {
-        #Mean pairwise distance
-        pd_result<-ses.pd(data_structure, tree, null.model=null, runs=runs, ...)
 
-        #Nearest relatedness index; NRI=0 -> random ; NRI<0 -> over-dispersion (good) ; NRI>0 -> clustering (bad)
+        #If tree is not rooted, add one
+        if(!is.rooted(tree)) {
+            root_tmp<-rtree(2)
+            tree_tmp<-bind.tree(tree, root_tmp)
+            tree_tmp<-drop.tip(tree_tmp,"t2")
+            tree_rooted<-root(tree_tmp, "t1", resolve.root=TRUE)
+
+            #Mean pairwise distance
+            pd_result<-ses.pd(data_structure, tree_rooted, null.model=null, runs=runs, ...)
+
+        } else {
+
+            #Mean pairwise distance
+            pd_result<-ses.pd(data_structure, tree, null.model=null, runs=runs, ...)            
+
+        }
+
+        #Faith's PD; PD=0 -> random ; PD<0 -> over-dispersion (good) ; PD>0 -> clustering (bad)
         PD<-pd_result$pd.obs.z[1]*-1
 
         #Significance of difference from NULL

### 2015-06-20 Analysis/Analysis_per_order.R
Updated analysis

@@ -1,13 +1,11 @@
 #Loading the functions
-setwd("~/PhD/Projects/Missing_living_mammals/Analysis")
+
 source("functions.R")
 load.functions(test=FALSE) #Set test=FALSE to speed up the loading
 library(picante)
 
 #Loading the tree
-full_trees<-read.nexus("../Data/Trees/FritzTree.rs200k.100trees.tre")
-#Selecting one random tree
-set.seed(1) ; one_tree<-full_trees[[sample(1:100, 1)]]
+mam_tree<-read.nexus("../Data/Trees/mammalST_MSW05_best_chrono.tre")
 
 #Loading the taxonomic reference list
 WR_list<-read.csv("../Data/Taxon_References/WilsonReederMSW.csv", header=T, stringsAsFactors=F)
@@ -24,28 +22,29 @@ extraction_table[which(extraction_table$Matrix == "GS2007-R.nex"), 3]<-88
 extraction_table[which(extraction_table$Matrix == "GS2010-VJD.nex"), 3]<-157
 extraction_table[which(extraction_table$Matrix == "GS2012-GM.nex"), 3]<-45
 extraction_table[which(extraction_table$Matrix == "GS2012-SM.nex"), 3]<-157
+extraction_table[which(extraction_table$Matrix == "GS1997-V.nex"), 3]<-40
+extraction_table[which(extraction_table$Matrix == "GS2013-MS.nex"), 3]<-48
+extraction_table[which(extraction_table$Matrix == "GS2014-MG.nex"), 3]<-40
+extraction_table[which(extraction_table$Matrix == "GS2014-MaL.nex"), 3]<-98
+extraction_table[which(extraction_table$Matrix == "GS2015-RR.nex"), 3]<-149
+extraction_table[which(extraction_table$Matrix == "GS2014-B.nex"), 3]<-27
+extraction_table[which(extraction_table$Matrix == "GS2014-Y.nex"), 3]<-42
 
 #Creating sub tables with a minimum number of characters
 #1 Character
 extraction_table1<-extraction_table[which(as.numeric(extraction_table$Characters) >= 1),]
-#75 Characters
-extraction_table75<-extraction_table[which(as.numeric(extraction_table$Characters) >= 75),]
-#94 Characters
+
+#100 Characters
 extraction_table100<-extraction_table[which(as.numeric(extraction_table$Characters) >= 100),]
-#138 Characters
-extraction_table125<-extraction_table[which(as.numeric(extraction_table$Characters) >= 125),]
-#150 Characters
[…]

### 2015-06-20 Analysis/Extracting_living_taxa.R

@@ -1,5 +1,4 @@
 #Extracting the living taxa from all the matrices
-setwd("~/PhD/Projects/Missing_living_mammals/Analysis")
 
 #Load functions
 message("Loading/testing functions:", appendLF=FALSE)
@@ -36,14 +35,14 @@ body(read.nex.dat)[[20]]<-substitute(
 
 #List of references
 #Fritz tree
-tree_list<-read.table("../Data/Taxon_References/FritzTree_taxa.txt", header=F, stringsAsFactors=F)
+tree_list<-read.table("../Data/Taxon_References/Bininda-emonds_taxa.txt", header=F, stringsAsFactors=F)
 #Wilson Reeder's Mammals Species of the World
 WR_list<-read.csv("../Data/Taxon_References/WilsonReederMSW.csv", header=T, stringsAsFactors=F)
 #Palaeobiology database
 Paleo_list<-read.csv("../Data/Taxon_References/pdb_taxa.csv", header=T, stringsAsFactors=F)
 #Combining the references
 Reference_list<-list(tree_list, WR_list, Paleo_list)
-names(Reference_list)<-c("Fritz Tree", "Wilson & Reeder", "Paleobio database")
+names(Reference_list)<-c("Bininda-emonds Tree", "Wilson & Reeder", "Paleobio database")
 
 #Building the empty table
 extraction_table<-data.frame("Matrix"=NA,"Taxa"=NA, "Characters"=NA, "Living"=NA,"Source"=NA,"Tax.level"=NA)
@@ -104,7 +103,7 @@ NA_species$Source[which(!is.na(doubleCheck_WR))]<-"WR"
 NA_species$Tax.level[which(!is.na(doubleCheck_WR))]<-doubleCheck_WR[which(!is.na(doubleCheck_WR))]
 #For the tree
 NA_species$Living[which(!is.na(doubleCheck_tree))]<-TRUE
-NA_species$Source[which(!is.na(doubleCheck_tree))]<-"Fritz"
+NA_species$Source[which(!is.na(doubleCheck_tree))]<-"Bininda-emonds"
 NA_species$Tax.level[which(!is.na(doubleCheck_tree))]<-"Species"
 
 #Replace NA_species back in the table
@@ -114,4 +113,4 @@ extraction_table[which(is.na(extraction_table$Living)),]<-NA_species
 save(extraction_table, file="../Data/List_of_matching_taxa.Rda")
 
 #Save the result as csv
[…]

### 2015-06-20 Analysis/Pipelined_analysis.R

@@ -1,0 +1,8 @@
+#This is a step by step description for reproducing the tables and the figures from the analysis.
+
+###############################
+# Extracting the living taxa
+###############################
+
+#The first point is to analyse each matrix and extract the living taxa
+#This can be done using the 

### 2015-06-20 Functions/plot.results.R
tree is now an isolated option

@@ -1,8 +1,8 @@
-plot.results<-function(order, taxa, col_branch=c("red","grey"), reference, verbose=FALSE, ...) {
+plot.results<-function(order, taxa, col_branch=c("red","grey"), reference, tree, verbose=FALSE,...) {
     #Diversitree
     require(diversitree)
     #Extracting the tree
-    order_tree<-extract.order(order, one_tree, reference, verbose)
+    order_tree<-extract.order(order, tree, reference, verbose)
     #Stopping if the tree is two small
     if(Ntip(order_tree)<3) {
         stop("To few tips in the tree.")

### 2015-06-20 Manuscript/Table_results.tex
Table result
supplementary figures (updated)

@@
 % latex table generated in R 3.2.0 by xtable 1.7-4 package
-% Fri Jun  5 15:29:14 2015
+% Sat Jun 20 17:41:32 2015
 \begin{longtable}{lL{1.8cm}C{2cm}lcc}
@@
   Afrosoricida & genus & 17/17 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar2.pdf} &   &   \\ 
-  Afrosoricida & species & 23/42 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar3.pdf} & 1.55 & -2.32 \\ 
-  Carnivora & family & 11/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar4.pdf} & 0.52 & 1.6 \\ 
-  \textbf{Carnivora} & \textbf{genus} & \textbf{30/125} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar5.pdf} & \textbf{4.01**} & \textbf{1.75*} \\ 
-  \textbf{Carnivora} & \textbf{species} & \textbf{42/283} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar6.pdf} & \textbf{17.95**} & \textbf{3.1**} \\ 
-  Cetartiodactyla & family & 20/21 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar7.pdf} & 0.35 & 1.11 \\ 
-  Cetartiodactyla & genus & 76/128 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar8.pdf} & -3.22 & -1.12 \\ 
-  Cetartiodactyla & species & 106/311 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar9.pdf} & -2.77 & -2.1 \\ 
-  Chiroptera & family & 13/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar10.pdf} & 0.87 & 0.65 \\ 
-  \textbf{Chiroptera} & \textbf{genus} & \textbf{68/202} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar11.pdf} & \textbf{18.87**} & \textbf{1.09} \\ 
-  \textbf{Chiroptera} & \textbf{species} & \textbf{108/1054} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar12.pdf} & \textbf{20.11**} & \textbf{2.24*} \\ 
+  \textbf{Afrosoricida} & \textbf{species} & \textbf{23/42} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar3.pdf} & \textbf{1.89*} & \textbf{1.19} \\ 
+  Carnivora & family & 11/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar4.pdf} & 0.43 & 1.68 \\ 
+  \textbf{Carnivora} & \textbf{genus} & \textbf{30/125} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar5.pdf} & \textbf{4.14**} & \textbf{1.81*} \\ 
+  \textbf{Carnivora} & \textbf{species} & \textbf{42/283} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar6.pdf} & \textbf{18.64**} & \textbf{3.02**} \\ 
+  Cetartiodactyla & family & 21/21 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar7.pdf} &   &   \\ 
+  \textbf{Cetartiodactyla} & \textbf{genus} & \textbf{77/128} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar8.pdf} & \textbf{0.87} & \textbf{1.77*} \\ 
+  \textbf{Cetartiodactyla} & \textbf{species} & \textbf{129/310} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar9.pdf} & \textbf{2.72*} & \textbf{0.04} \\ 
+  Chiroptera & family & 13/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar10.pdf} & 0.55 & 0.63 \\ 
+  \textbf{Chiroptera} & \textbf{genus} & \textbf{85/202} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar11.pdf} & \textbf{16.91**} & \textbf{2.85**} \\ 
+  \textbf{Chiroptera} & \textbf{species} & \textbf{165/1053} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar12.pdf} & \textbf{14.55**} & \textbf{3.44**} \\ 
   Cingulata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar13.pdf} &   &   \\ 
-  Cingulata & genus & 8/9 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar14.pdf} & 1.31 & -1.66 \\ 
-  Cingulata & species & 6/25 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar15.pdf} & 0.65 & -0.47 \\ 
+  Cingulata & genus & 8/9 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar14.pdf} & 1.49 & -1.63 \\ 
+  Cingulata & species & 6/29 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar15.pdf} & 1.43 & 0.36 \\ 
   Dasyuromorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar16.pdf} &   &   \\ 
-  Dasyuromorphia & genus & 7/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar17.pdf} & -1.1 & -1.78 \\ 
-  Dasyuromorphia & species & 8/64 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar18.pdf} & -1.1 & -1.31 \\ 
+  Dasyuromorphia & genus & 7/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar17.pdf} & -1 & -1.45 \\ 
+  Dasyuromorphia & species & 8/64 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar18.pdf} & -1.15 & -0.62 \\ 
   Dermoptera & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar19.pdf} &   &   \\ 
@@
   Didelphimorphia & genus & 16/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar23.pdf} &   &   \\ 
-  Didelphimorphia & species & 40/84 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar24.pdf} & -1.33 & -1.7 \\ 
-  Diprotodontia & family & 9/11 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar25.pdf} & -0.35 & 1.11 \\ 
-  Diprotodontia & genus & 20/38 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar26.pdf} & -0.31 & -1.57 \\ 
-  Diprotodontia & species & 16/126 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar27.pdf} & -0.98 & -2.6 \\ 
+  Didelphimorphia & species & 40/84 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar24.pdf} & -0.94 & 0.36 \\ 
+  Diprotodontia & family & 9/11 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar25.pdf} & -0.8 & 0.56 \\ 
+  Diprotodontia & genus & 20/38 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar26.pdf} & -1.36 & -0.73 \\ 
+  Diprotodontia & species & 16/126 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar27.pdf} & -2.29 & -1.55 \\ 
   Erinaceomorpha & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar28.pdf} &   &   \\ 
   Erinaceomorpha & genus & 10/10 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar29.pdf} &   &   \\ 
-  Erinaceomorpha & species & 21/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar30.pdf} & -0.78 & -0.19 \\ 
+  Erinaceomorpha & species & 21/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar30.pdf} & -1.1 & -0.3 \\ 
   Hyracoidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar31.pdf} &   &   \\ 
@@
   Macroscelidea & genus & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar38.pdf} &   &   \\ 
-  Macroscelidea & species & 5/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar39.pdf} & -0.47 & -0.89 \\ 
+  Macroscelidea & species & 5/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar39.pdf} & -0.98 & -1.38 \\ 
   Microbiotheria & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar40.pdf} &   &   \\ 
@@
   Monotremata & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar43.pdf} &   &   \\ 
-  Monotremata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar44.pdf} & -0.7 & -0.71 \\ 
-  Monotremata & species & 2/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar45.pdf} & -0.97 & -1.01 \\ 
+  Monotremata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar44.pdf} & -0.71 & -0.71 \\ 
+  Monotremata & species & 2/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar45.pdf} & -1.01 & -1.03 \\ 
   Notoryctemorphia & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar46.pdf} &   &   \\ 
@@
   Paucituberculata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar49.pdf} &   &   \\ 
-  Paucituberculata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar50.pdf} & -0.76 & -0.74 \\ 
+  Paucituberculata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar50.pdf} & 0 & 0 \\ 
   Paucituberculata & species & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar51.pdf} & -0.64 & -0.65 \\ 
@@
   Peramelemorphia & genus & 7/7 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar53.pdf} &   &   \\ 
-  Peramelemorphia & species & 16/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar54.pdf} & -0.55 & 0.49 \\ 
+  Peramelemorphia & species & 16/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar54.pdf} & -0.09 & 1 \\ 
   Perissodactyla & family & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar55.pdf} &   &   \\ 
   Perissodactyla & genus & 6/6 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar56.pdf} &   &   \\ 
-  Perissodactyla & species & 7/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar57.pdf} & 0.73 & -2.63 \\ 
+  Perissodactyla & species & 7/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar57.pdf} & 0.62 & -2.5 \\ 
   Pholidota & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar58.pdf} &   &   \\ 
   Pholidota & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar59.pdf} &   &   \\ 
-  Pholidota & species & 3/8 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar60.pdf} & 1.14 & 0.66 \\ 
-  Pilosa & family & 3/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar61.pdf} & -0.6 & -0.85 \\ 
-  Pilosa & genus & 3/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar62.pdf} & -0.88 & -1.09 \\ 
-  Pilosa & species & 3/29 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar63.pdf} & -1.51 & -1.47 \\ 
+  \textbf{Pholidota} & \textbf{species} & \textbf{3/8} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar60.pdf} & \textbf{2.64*} & \textbf{2.23*} \\ 
+  Pilosa & family & 3/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar61.pdf} & 0.94 & 0.93 \\ 
+  Pilosa & genus & 3/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar62.pdf} & -0.36 & -0.31 \\ 
+  Pilosa & species & 3/29 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar63.pdf} & 0.33 & 0.79 \\ 
   Primates & family & 15/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar64.pdf} &   &   \\ 
-  Primates & genus & 48/68 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar65.pdf} & -0.3 & -1.29 \\ 
-  Primates & species & 56/351 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar66.pdf} & -1.5 & -2.18 \\ 
+  Primates & genus & 48/68 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar65.pdf} & -0.41 & -1.4 \\ 
+  Primates & species & 56/351 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar66.pdf} & -1.6 & -2.04 \\ 
   Proboscidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar67.pdf} &   &   \\ 
@@
   Proboscidea & species & 1/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar69.pdf} &   &   \\ 
-  Rodentia & family & 10/32 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar70.pdf} & 0.11 & -0.3 \\ 
-  Rodentia & genus & 20/451 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar71.pdf} & -1.03 & 0.71 \\ 
-  Rodentia & species & 10/2095 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar72.pdf} & -0.97 & -1.75 \\ 
+  Rodentia & family & 11/32 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar70.pdf} & -0.46 & -1.91 \\ 
+  Rodentia & genus & 21/450 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar71.pdf} & -2.11 & 0.3 \\ 
+  Rodentia & species & 15/2094 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar72.pdf} & -1.65 & -2.55 \\ 
   Scandentia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar73.pdf} &   &   \\ 
-  Scandentia & genus & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar74.pdf} & -0.75 & -0.76 \\ 
-  Scandentia & species & 2/20 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar75.pdf} & -1.26 & -1.27 \\ 
+  Scandentia & genus & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar74.pdf} & -0.77 & -0.76 \\ 
+  Scandentia & species & 2/20 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar75.pdf} & -1.79 & -1.99 \\ 
   Sirenia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar76.pdf} &   &   \\ 
   Sirenia & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar77.pdf} &   &   \\ 
-  Sirenia & species & 2/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar78.pdf} & -0.96 & -0.94 \\ 
-  Soricomorpha & family & 3/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar79.pdf} & -0.95 & -1.05 \\ 
-  Soricomorpha & genus & 19/43 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar80.pdf} & 1.16 & -1.01 \\ 
-  Soricomorpha & species & 19/392 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar81.pdf} & -1.87 & -3.09 \\ 
+  Sirenia & species & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar78.pdf} &   &   \\ 
+  Soricomorpha & family & 3/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar79.pdf} & -0.93 & -0.92 \\ 
+  \textbf{Soricomorpha} & \textbf{genus} & \textbf{19/43} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar80.pdf} & \textbf{6.98**} & \textbf{2.49*} \\ 
+  \textbf{Soricomorpha} & \textbf{species} & \textbf{19/392} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar81.pdf} & \textbf{13.19**} & \textbf{3.89**} \\ 
   Tubulidentata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar82.pdf} &   &   \\ 

### 2015-06-20 Analysis/Results_summary.R
Updated analysis

@@ -6,11 +6,11 @@ library(picante)
 library(xtable)
 
 #Loading the tree
-full_trees<-read.nexus("../Data/Trees/mammalST_MSW05_best_chrono.tre")
+mam_tree<-read.nexus("../Data/Trees/mammalST_MSW05_best_chrono.tre")
 #Resolve polytomies
-one_tree<-multi2di(full_trees)
+mam_tree<-multi2di(mam_tree)
 #Selecting one random tree
-#set.seed(1) ; one_tree<-full_trees[[sample(1:100, 1)]]
+#set.seed(1) ; mam_tree<-mam_tree[[sample(1:100, 1)]]
 
 #Loading the taxonomic reference list
 #Loading the taxonomic reference list
@@ -28,26 +28,27 @@ extraction_table[which(extraction_table$Matrix == "GS2007-R.nex"), 3]<-88
 extraction_table[which(extraction_table$Matrix == "GS2010-VJD.nex"), 3]<-157
 extraction_table[which(extraction_table$Matrix == "GS2012-GM.nex"), 3]<-45
 extraction_table[which(extraction_table$Matrix == "GS2012-SM.nex"), 3]<-157
+extraction_table[which(extraction_table$Matrix == "GS1997-V.nex"), 3]<-40
+extraction_table[which(extraction_table$Matrix == "GS2013-MS.nex"), 3]<-48
+extraction_table[which(extraction_table$Matrix == "GS2014-MG.nex"), 3]<-40
+extraction_table[which(extraction_table$Matrix == "GS2014-MaL.nex"), 3]<-98
+extraction_table[which(extraction_table$Matrix == "GS2015-RR.nex"), 3]<-149
+extraction_table[which(extraction_table$Matrix == "GS2014-B.nex"), 3]<-27
+extraction_table[which(extraction_table$Matrix == "GS2014-Y.nex"), 3]<-42
 
 #Creating sub tables with a minimum number of characters
 #1 Character
 extraction_table1<-extraction_table[which(as.numeric(extraction_table$Characters) >= 1),]
-#75 Characters
-extraction_table75<-extraction_table[which(as.numeric(extraction_table$Characters) >= 75),]
+
 #100 Characters
 extraction_table100<-extraction_table[which(as.numeric(extraction_table$Characters) >= 100),]
-#125 Characters
-extraction_table125<-extraction_table[which(as.numeric(extraction_table$Characters) >= 125),]
-#150 Characters
[…]

### 2015-06-20 Manuscript/Missing_mammals-draft.tex
Updated the results in manuscript

@@
 soft \textit{vs.} hard tissues) may restrict the amount of morphological data available \citep{sansomfossilization2013}.
-% NC: Worth metioning fossil preservation here briefly as it is also an
 Additionally, since the molecular phylogenetics revolution, it has become less common to collect morphological characters for living taxa when molecular data is available (e.g.
@@
 To investigate this further, we assess the amount of available morphological data for living mammals to determine whether sufficient data exists to build reliable Total Evidence phylogenies in this group.
-We collected cladistic data (i.e. % NC: define cladistic data here
-) from @256 phylogenetic matrices available online and measured the proportion of cladistic data available for each mammalian order.
+We collected cladistic data (i.e.
+discrete morphological characters used in phylogenetics) from @256 phylogenetic matrices available online and measured the proportion of cladistic data available for each mammalian order.
 %Additionally, if the available data is randomly distributed across clades, the topological errors should be less extreme than if the data is biased towards particulars groups.
@@
 We also performed a systematic Google Scholar search (accessed @th of July 2015) for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
-In total, we downloaded @256 % Number of matrices
-matrices containing @9411 unique % Number of unique OTUs
-operational taxonomic units (OTUs).
-In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardized and ranged from specific specimen names %NC: Do you mean names of individual specimens or subspecies?
-to the family-level.
+In total, we downloaded 286 matrices containing a total of 11010 operational taxonomic units (OTUs) of which 5228 where unique. % Number of unique OTUs (including fossils)
+In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardized and ranged from specific individual specimen (i.e.
+the name of a collection item) to the family-level.
 Where possible, we considered OTUs at their lowest valid taxonomic level (i.e. species) but some OTUs were only valid at a higher taxonomic level (e.g. genus or family).
@@
 sapiens} was transformed to \textit{Homo sapiens}).
-We designated as ``living'' all OTUs that were either present in the phylogeny of \citet{FritzTree} or the taxonomy of \citet{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\texttt{https://paleobiodb.org/}).
+We designated as ``living'' all OTUs that were either present in the phylogeny of \citet{BinindaEmonds} or the taxonomy of \citet{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\texttt{https://paleobiodb.org/}).
 For OTUs that did not appear in these three sources, we first decomposed the name (i.e.
@@
 
-The number of characters in each matrix ranged from @6 to @4541.
-Matrices with few characters % NC: matrices can be small in two ways - with few characters and with few species. Here you just mention few characters so I've made it more specific.
-are problematic when comparing available data among matrices because (1) they have less chance of having characters that overlap with those of other matrices \citep{wagner2000} and (2) they are more likely to contain a higher proportion of specific characters that are not-applicable across large clades \citep[][e.g.
+The number of characters in each matrix ranged from 6 to 4541.
+Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having characters that overlap with those of other matrices \citep{wagner2000} and (2) they are more likely to contain a higher proportion of specific characters that are not-applicable across large clades \citep[][e.g.
 ``antler ramifications'' is a character that is only applicable to Cervidae not all mammals]{Brazeau2011}.
-% NC: Is this true really? Does it always follow that if you've very few characters you also have very specific characters? Maybe so...
+% NC: Is this true really?
+Does it always follow that if you've very few characters you also have very specific characters?
+Maybe so...
+TG: I don't know if it's a generalisation but from the matrices I collected, the once with few characters always came from papers focusing on intra-specific or intra-generic variations.
+On the other hand, I didn't read all the characters.
+I just assume that in a big one, you have a lower probability of having non-applicable characters (there's only some many characters for coding antler ramification)
 Therefore we selected only matrices containing \textgreater 100 characters per OTU.
 This threshold was chosen to correspond with the number of characters used in \citet{GuillermeCooper} and \citet{harrisonamong-character2014}.
-
-After removing all matrices with \textless 100 characters, we retained @1422 unique living mammal OTUs from @256 matrices for our analyses.
+Note that the results of the analysis with no character threshold is available in the supplementary materials.
+After removing all matrices with \textless 100 characters, we retained 1074 unique living mammal OTUs from 126 matrices for our analyses. % 1601 unique living OTUs for 286 matrices (no threshold)
 
@@
 We calculated NTI and NRI values for each mammalian order separately, at each different taxonomic level. 
-For each analysis our focal taxa were those with available cladistic data at that taxonomic level and the phylogeny was the phylogeny of the order pruned from \citep{}. % cite  Bininda Emonds?
+For each analysis our focal taxa were those with available cladistic data at that taxonomic level and the phylogeny was the phylogeny of the order pruned from \citep{BinindaEmonds}.
+% cite  Bininda Emonds?
 % NC: Tried to make this more specific here. It's important to note that you used the order not the whole mammal phylogeny each time.
@@
 \section{Results}
-Across the @@@ cladistic matrices we extracted, @11 out of @28 mammalian orders have low coverage (\textless 25\% of species with cladistic data) and @24 have high coverage (\textgreater 75\% of species with cladistic data) at the species-level.
-At the genus-level, only @three orders have low coverage of taxa and @16 have high coverage, and at the family-level, @no orders have low coverage and @5 have high coverage (Table \ref{Table_morpho_taxa_proportion}).
+Across the @126 cladistic matrices we extracted, @22 out of 28 mammalian orders have low coverage (\textless 25\% of species with cladistic data) and @6 have high coverage (\textgreater 75\% of species with cladistic data) at the species-level.
+At the genus-level, only @three order have low coverage of taxa and @12 have high coverage, and at the family-level, no orders have low coverage and @23 have high coverage (Table \ref{Table_morpho_taxa_proportion}).
 
@@
 
-Two contrasting results are shown in Figure \ref{Figure_example_coverage} with randomly distributed OTUs with cladistic data in Cetartiodactyla (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered OTUs with cladistic data in Carnivora (mainly Canidae; Figure \ref{Figure_example_coverage}B).
+Two contrasting results are shown in Figure \ref{Figure_example_coverage} with randomly distributed OTUs with cladistic data in Primates (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered OTUs with cladistic data in Carnivora (mainly Canidae; Figure \ref{Figure_example_coverage}B).
 
@@
     \includegraphics[width=1\textwidth]{example_coverage.pdf}
-\caption{Phylogenetic distribution of species with available cladistic data across two mammalian orders (A: Cetartiodactyla; B: Carnivora).
+\caption{Phylogenetic distribution of species with available cladistic data across two mammalian orders (A: Primates; B: Carnivora).
 Edges are colored in grey when no cladistic data is available for a species or in blue when data is available.}
@@
 However, in reasonably sized matrices (\textgreater 100 characters; \citealp{GuillermeCooper,harrisonamong-character2014}) it is more likely that a number of characters are consistently conserved among the different matrices (e.g.
-within the Primate literature, the character \textit{p7} - the size of the $4^{th}$ lower premolar paraconid - has been used consistently for \textgreater 15 years; \citealp[e.g.][]{ross1998phylogenetic,seiffert2003fossil,marivaux2005anthropoid,seiffert2005basal,bloch2007new,kay2008anatomy,silcox2008biogeographic,seiffert2009convergent,tabuce2009anthropoid,boyer2010astragalar,seiffert2010fossil,marivaux2013djebelemur,ni2013oldest}).
+within the Primate literature, the character \textit{p7} - the size of the $4^{th}$ lower premolar paraconid - has been used consistently for \textgreater 15 years; \citealp[e.g.][]{ross1998phylogenetic,seiffert2003fossil,kay2008anatomy,boyer2010astragalar,marivaux2013djebelemur}).
 % NC: Remove some of these!!!

### 2015-06-21 Manuscript/Missing_mammals-draft.tex
Manuscript update

@@
 
-For orders with \textless 100\% cladistic data coverage at any taxonomic level, we investigated whether the available cladistic data was (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: Nearest Taxon Index \citep[NTI;][]{webb2002phylogenies} and Net Relatedness Index \citep[NRI;][]{webb2002phylogenies}.
-%*** see end of methods for explanation.
+For orders with \textless 100\% cladistic data coverage at any taxonomic level, we investigated whether the available cladistic data was (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index \citep[NTI;][]{webb2002phylogenies} and the Net Relatedness Index \citep[NRI;][]{webb2002phylogenies}.
+%*** see end of methods for explanation. TG: I've but it below, I think it reads better after explaining the metrics.
+Both metrics 
 Both metrics were calculated using the \texttt{picante} package in R \citep{picante,R}.
@@
   \end{equation}
-where $\overline{MNND}_{obs}$ is the observed mean distance among $n$ focal taxa and their nearest neighbour in the phylogeny,
-% NC: Sorry this still isn't quite right. Check my paper for a nicer way of writing it.
-$\overline{MNND}_{n}$ is the expected random MNND for $n$ taxa estimated by calculating the MNND from $n$ taxa randomly drawn from the phylogeny and repeated 1000 times, and $\sigma(MNND_{n})$ is the standard deviation of the 1000 random MNND values.
+where $\overline{MNND}_{obs}$ is the observed mean distance between each of $n$ taxa with cladistic data, $\overline{MNND}_{n}$ is the mean of 1000 mean MNND between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random MNND values.
+% NC: Sorry this still isn't quite right.
+Check my paper for a nicer way of writing it.
+- TG: I've done some fixes on the sentence above.
+I didn't want to copy/paste what's in your paper so I tried to explain it in my own words.
+Maybe it'll need translation from Thomas-english and will end up sounding similar to what you wrote in your paper.
+Which is frankly not a big deal since it is the exact same thing we're describing here.
+%$\overline{MNND}_{n}$ is the expected random MNND for $n$ taxa estimated by calculating the MNND from $n$ taxa randomly drawn from the phylogeny and repeated 1000 times, and $\sigma(MNND_{n})$ is the standard deviation of the 1000 random MNND values.
 NRI is similar but is based on mean phylogenetic distance (MPD) as follows:
@@
   \end{equation}
-where $\overline{MPD}_{obs}$ is the observed mean phylogenetic distance among $n$ focal taxa, $\overline{MPD}_{n}$ is the expected random MPD for $n$ taxa estimated by calculating the MPD from $n$ taxa randomly drawn from the phylogeny and repeated 1000 times, and $\sigma(MPD_{n})$ is the standard deviation of the 1000 random MPD values.
+where $\overline{MPD}_{obs}$ is the observed mean phylogenetic distance of the tree containing only the $n$ taxa with cladistic data, $\overline{MPD}_{n}$ is the expected random MPD for $n$ taxa estimated by calculating the MPD from $n$ taxa randomly drawn from the phylogeny and repeated 1000 times, and $\sigma(MPD_{n})$ is the standard deviation of the 1000 random MPD values.
 Negative NTI and NRI values show that the focal taxa are more overdispersed across the phylogeny than expected by chance, and positive values show that the focal taxa are more clustered across the phylogeny than expected by chance.
 %*** see end of methods for explanation.
-We calculated NTI and NRI values for each mammalian order separately, at each different taxonomic level. 
-For each analysis our focal taxa were those with available cladistic data at that taxonomic level and the phylogeny was the phylogeny of the order pruned from \citep{BinindaEmonds}.
-% cite  Bininda Emonds?
-% NC: Tried to make this more specific here. It's important to note that you used the order not the whole mammal phylogeny each time.
-
 %Because the aim of this study is to understand the structure of the available cladistic data in mammals we prefered the NTI over the NRI or the PD metrics.
@@
 %However, the results of the NRI and PD calculations are available in the supplementary.
-
 % NC: Actually I think we discussed that NRI and NTI show different things?
@@
 This can go up with the intro to the metrics somewhere (I have *** the two places it might fit)
+The result of both metric show slightly different aspects of the structure of the available data.
+NTI tends to be more sensitive to the phylogenetic structure at the tips of the phylogeney since it only takes into account the phylogenetic distance between taxa \citep{Cooper2008}.
+On the other hand the NRI is slightly biased towards detecting overdispersion since it is calculated using the mean phylogenetic distance from root to tips \citep{Kembel2006,Swenson2006}.
+% TG: In the end, I don't know if that's useful.
+For the 10 significant results, only three of them are not significant for both NRI and NTI...
+Maybe we can decide if we keep it or not after the word count.
 
+We calculated NTI and NRI values for each mammalian order separately, at each different taxonomic level. 
+For each analysis our focal taxa were those with available cladistic data at that taxonomic level and the phylogeny was the phylogeny of the order pruned from \citet{BinindaEmonds}.
+% NC: Tried to make this more specific here. It's important to note that you used the order not the whole mammal phylogeny each time.
 We also report the results using another metric, Phylogenetic Distance \citep[PD;][]{Faith19921}, in the Supplementary Information. % link?
@@
 \section{Results}
+% TG: updated results.
+Slight changes in which group is significant or not.
+Mainly due to using Bininda-Emonds tree I'll say.
+But the main message is still the same.
+It's badly covered but at least is not that clustered.
+I just changed the example from Carnivores vs.
+cetartio to carnivores vs.
+primates (cetartio are now clustered).
 Across the @126 cladistic matrices we extracted, @22 out of 28 mammalian orders have low coverage (\textless 25\% of species with cladistic data) and @6 have high coverage (\textgreater 75\% of species with cladistic data) at the species-level.
@@
 
-Among the mammalian orders containing OTUs with no available cladistic data, only @two orders, Carnivora and Chiroptera, have OTUs with cladistic data that are significantly phylogenetically clustered at the species- and the genus-level (Table~\ref{Table_data_structure}).
+Among the mammalian orders containing OTUs with no available cladistic data, only 6 orders had significantly clustered data (Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha at both species- and genus-level and Afrosoricida and Pholidota at the species-level only) and no order had significantly overdispersed data at any taxonomic level (Table~\ref{Table_data_structure}).
 
@@
 \section{Discussion}
-Our results show that although phylogenetic relationships among living mammals are well-resolved \citep[e.g.][]{FritzTree,meredithimpacts2011,May-Collado-PeerJ}, most of the data used to build these phylogenies is molecular, and very little cladistic data is available for living mammals compared to fossil mammals \citep[e.g.][]{O'Leary08022013,ni2013oldest}.
+Our results show that although phylogenetic relationships among living mammals are well-resolved \citep[e.g.][]{FritzTree,meredithimpacts2011,May-Collado-PeerJ} % TG: note that we can reduce the cites here (for the world count in the refs).
+All three are cited only one time (here).
+We can save some cites by using \cite{BinindaEmonds,wilson2005mammal} that are cited all along the text already.
+, most of the data used to build these phylogenies is molecular, and very little cladistic data is available for living mammals compared to fossil mammals \citep[e.g.][]{O'Leary08022013,ni2013oldest}.
 This has implications for building Total Evidence phylogenies containing both living and fossil mammals, as without sufficient cladistic data for living species, fossil placements in these trees are very uncertain \citep{GuillermeCooper}.
-However, cladistic data coverage in living mammals varies across taxonomic levels and in its phylogenetic distribution.
-Higher taxonomic levels are always better sampled than lower ones and within these taxonomic levels, the available data is mostly randomly distributed across the phylogeny, apart from in Carnivora and Chiroptera.
-% NC: Add a sentence saying this means things might not be quite so bad? Need to justify the "However". however what?
+Cladistic data coverage in living mammals varies across taxonomic levels and in its phylogenetic distribution.
+Higher taxonomic levels are always better sampled than lower ones and within these taxonomic levels, the available data is mostly randomly distributed across the phylogeny, apart from in 6 orders (see Results and Table\ref{Table_morpho_taxa_proportion}).
+% NC: Add a sentence saying this means things might not be quite so bad?
+Need to justify the "However".
+however what?
+TG: I think the "however" fits better below.
+Here it's §1: "that's what we found"; §2 (first part): "That's baaaad", §2 (second part) "However, it's all right at genus level"
 
-The number of living mammalian taxa with no available cladistic data was surprisingly high at the species-level: only @4 out of @28 orders have a high coverage of taxa with available cladistic data (and two of the 28 orders are monospecific!).
+The number of living mammalian taxa with no available cladistic data was surprisingly high at the species-level: only @6 out of @28 orders have a high coverage of taxa with available cladistic data (and two of the 28 orders are monospecific!).
 This high coverage threshold of 75\% of taxa with available cladistic data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.
@@
 The second best scenario is that species with cladistic data are randomly distributed across the phylogeny. 
-In this scenario we expected no special bias in the placement of the fossil \citep{GuillermeCooper}, it is therefore encouraging that for all orders except Carnivora and Chiroptera, species with cladistic data were randomly distributed across the phylogeny of each order.
+In this scenario we expected no special bias in the placement of the fossil \citep{GuillermeCooper}, it is therefore encouraging that for most orders, species with cladistic data were randomly distributed across the phylogeny of each order.
 The worst case scenario for fossil placement is that species with cladistic data are phylogenetically clustered. 
@@
 This means that fossils with uncertain phylogenetic affinities (\textit{incertae sedis}) will have a higher probability of branching within the most sampled clade just by chance.
-Our results suggest that this may be an issue in Carnivora and Chiroptera. 
+Our results suggest that this may be an issue, at the genus-level, in Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha. 
 For example, a Carnivora fossil will be unable to branch in the Herpestidae that has no species with cladistic data, and will also have more chance to branch, randomly, within the Canidae clade than any other clade in Carnivora (Figure \ref{Figure_example_coverage}-B).
-Thus placements of Carnivoran fossils should be viewed with suspicion. % NC: OK gotta be a better way to say that.
+Thus, in Total Evidence trees, placements of Carnivoran fossils should considered with caution. %viewed with suspicion. % NC: OK gotta be a better way to say that. TG: is that a bit milder?
 
-%Caveats
 In this study, we treated all cladistic matrices as equal in a similar way to molecular matrices. 
@@
 Unfortunately, cladistic data cannot always be treated in this way because some characters may overlap.
-For example, if matrix A has a character coding for the shape of a particular morphological feature and matrix B has a character coding for the presence of this morphological feature and a second character coding for its shape, then these three characters are non-independent compound characters \citep{Brazeau2011}.
-However, in reasonably sized matrices (\textgreater 100 characters; \citealp{GuillermeCooper,harrisonamong-character2014}) it is more likely that a number of characters are consistently conserved among the different matrices (e.g.
-within the Primate literature, the character \textit{p7} - the size of the $4^{th}$ lower premolar paraconid - has been used consistently for \textgreater 15 years; \citealp[e.g.][]{ross1998phylogenetic,seiffert2003fossil,kay2008anatomy,boyer2010astragalar,marivaux2013djebelemur}).
-% NC: Remove some of these!!!
+For example, if matrix A has a character coding for the shape of a particular morphological feature and matrix B has a character coding for the presence of this same morphological feature and a second character coding for its shape, then these three characters are non-independent compound characters \citep{Brazeau2011}.
+However, in reasonably sized matrices (\textgreater 100 characters; \citealp{GuillermeCooper,harrisonamong-character2014}) it is more likely that a number of characters are consistently conserved among the different matrices and thus easily combinable.
+For example, within the Primate cladistic literature, the character \textit{p7} - the size of the $4^{th}$ lower premolar paraconid - has been used consistently for \textgreater 15 years \citep[e.g.][]{ross1998phylogenetic,marivaux2005anthropoid,kay2008anatomy,boyer2010astragalar,ni2013oldest} and can therefore be combined among the matrices.
 A conservative approach to avoid compound characters would be to select only the most recent matrix for each group, but this would result in the loss of a lot of data.
 
-%Let's code some data
 Following the recommendations in \citep{GuillermeCooper}, to improve the accuracy of the topology of Total Evidence trees containing both living and fossil taxa, one should code cladistic characters for as many living species possible.
+%Because the data for living mammals is usually readily available in natural history collections, we propose that an increased effort be put into coding morphological characters from living species, possibly by engaging in collaborative data collection projects through web portals such as \textit{Morphobank} \citep{morphobank}.
+% NC: OK needs a summing up sentence!!!
+TG: I combined this paragraph with the postive ending point.
+That might sum up the thing.
+Despite the absence of good cladistic data coverage for living mammals, the Total Evidence methods seems to be still the most promising method for combining both living and fossil data and the best way to assess more accurately macroevolutionary processes.
+% TG: or is that a bit bold?
+It is also positive to consider that the problems underlined in this study can be easily solved.
+%TG: not sure about "easily", the idea is that the solution is not to break the speed of light or something like that, just to continue what people have stopped doing when DNA became trending.
 Because the data for living mammals is usually readily available in natural history collections, we propose that an increased effort be put into coding morphological characters from living species, possibly by engaging in collaborative data collection projects through web portals such as \textit{Morphobank} \citep{morphobank}.
-
-% NC: OK needs a summing up sentence!!!
 

### 2015-06-21 Manuscript/Supplementary/Supplementary.tex
Revised methods
Removed supp pdf

@@
 \subsection{Google scholars}
-To make sure we didn't miss any extra matrix that wasn't available on one of these repository, we ran a Google Scholar search on the 5th of January. We used the following key words:
+To make sure we didn't miss any extra matrix that wasn't available on one of these repository, we ran a Google Scholar search on the 5th of January. 
+We downloaded the additional cladistic matrices from the 20 first search results matching with our selected keywords and with any of the 35 taxonomic levels (mammals Orders, Infraclasses and Class).
+We used the following key words:
 
@@
 \section{Supplementary results}
-The following section contains supplementary results to the main body: the available data structure using the NTI and the PD metric; the proportion of available data and the data structure for all the matrices (including the matrices with less than 100 characters); and phylogenetical representation of the data availability per order (excluding Cetartiodactyla and Carnivora, present in the main body).
+The following section contains supplementary results to the main body: the available data structure using the NTI and the PD metric; the proportion of available data and the data structure for all the matrices (including the matrices with less than 100 characters); and phylogenetical representation of the data availability per order (excluding Primates and Carnivora, present in the main body).
 
-\renewcommand\baselinestretch{1.2}\selectfont
-\begin{center}
-\input{Supplementary/Supp_data_structureNTI}
-\end{center}
-\renewcommand\baselinestretch{2}\selectfont
-
-\renewcommand\baselinestretch{1.2}\selectfont
-\begin{center}
-\input{Supplementary/Supp_data_structurePD}
-\end{center}
-\renewcommand\baselinestretch{2}\selectfont
-
-\renewcommand\baselinestretch{1.2}\selectfont
-\begin{center}
-\input{Supplementary/morpho_taxa_proportion_threshold1}
-\end{center}
-\renewcommand\baselinestretch{2}\selectfont
-
-\renewcommand\baselinestretch{1.2}\selectfont
-\begin{center}
-\input{Supplementary/Supp_data_structureNRI_threshold1}
-\end{center}
-\renewcommand\baselinestretch{2}\selectfont
-
-\renewcommand\baselinestretch{1.2}\selectfont
-\begin{center}
-\input{Supplementary/Supp_data_structureNTI_threshold1}
-\end{center}
-\renewcommand\baselinestretch{2}\selectfont
-
-\renewcommand\baselinestretch{1.2}\selectfont
-\begin{center}
-\input{Supplementary/Supp_data_structurePD_threshold1}
-\end{center}
-\renewcommand\baselinestretch{2}\selectfont
+\input{Results_1c.tex}
 
@@
     \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_AFROSORICIDA.pdf}
-\caption{Distribution of available morphological data across Afrosoricida. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\caption{Distribution of available morphological data across Afrosoricida. Edges are colored in grey when no morphological data is available or in blue when data is available.}
 \label{Supp_Figure_Phylo-Afrosoricida}
@@
     \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_CHIROPTERA.pdf}
-\caption{Distribution of available morphological data across Chiroptera. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\caption{Distribution of available morphological data across Chiroptera. Edges are colored in grey when no morphological data is available or in blue when data is available.}
 \label{Supp_Figure_Phylo-Chiroptera}
@@
     \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_CINGULATA.pdf}
-\caption{Distribution of available morphological data across Cingulata. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\caption{Distribution of available morphological data across Cingulata. Edges are colored in grey when no morphological data is available or in blue when data is available.}
 \label{Supp_Figure_Phylo-Cingulata}
@@
     \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_DASYUROMORPHIA.pdf}
-\caption{Distribution of available morphological data across Dasyuromorphia. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\caption{Distribution of available morphological data across Dasyuromorphia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
 \label{Supp_Figure_Phylo-Dasyuromorphia}
@@
     \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_DIDELPHIMORPHIA.pdf}
-\caption{Distribution of available morphological data across Didelphimorphia. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\caption{Distribution of available morphological data across Didelphimorphia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
 \label{Supp_Figure_Phylo-Didelphimorphia}
@@
     \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_DIPROTODONTIA.pdf}
-\caption{Distribution of available morphological data across Diprotodontia. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\caption{Distribution of available morphological data across Diprotodontia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
 \label{Supp_Figure_Phylo-Diprotodontia}
@@
     \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_ERINACEOMORPHA.pdf}
-\caption{Distribution of available morphological data across Erinaceomorpha. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\caption{Distribution of available morphological data across Erinaceomorpha. Edges are colored in grey when no morphological data is available or in blue when data is available.}
 \label{Supp_Figure_Phylo-Erinaceomorpha}
@@
     \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_PILOSA.pdf}
-\caption{Distribution of available morphological data across Pilosa. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\caption{Distribution of available morphological data across Pilosa. Edges are colored in grey when no morphological data is available or in blue when data is available.}
 \label{Supp_Figure_Phylo-Pilosa}
@@
 \centering
-    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_PRIMATES.pdf}
-\caption{Distribution of available morphological data across Primates. Edges are colored in grey when no morphological data is available or in red when data is available.}
+    \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_CETARTIODACTYLA.pdf}
+\caption{Distribution of available morphological data across Cetartiodactyla. Edges are colored in grey when no morphological data is available or in blue when data is available.}
 \label{Supp_Figure_Phylo-Primates}
@@
     \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_RODENTIA.pdf}
-\caption{Distribution of available morphological data across Rodentia. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\caption{Distribution of available morphological data across Rodentia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
 \label{Supp_Figure_Phylo-Rodentia}
@@
     \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_SCANDENTIA.pdf}
-\caption{Distribution of available morphological data across Scandentia. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\caption{Distribution of available morphological data across Scandentia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
 \label{Supp_Figure_Phylo-Scandentia}
@@
     \includegraphics[width=1\textwidth]{Supplementary/Supp_figure_SORICOMORPHA.pdf}
-\caption{Distribution of available morphological data across Soricomorpha. Edges are colored in grey when no morphological data is available or in red when data is available.}
+\caption{Distribution of available morphological data across Soricomorpha. Edges are colored in grey when no morphological data is available or in blue when data is available.}
 \label{Supp_Figure_Phylo-Soricomorpha}

### 2015-06-21 Manuscript/Missing_mammals-draft.tex
Data availability update
Manuscript with supplementaries

@@
 % NC: Tried to make this more specific here. It's important to note that you used the order not the whole mammal phylogeny each time.
-We also report the results using another metric, Phylogenetic Distance \citep[PD;][]{Faith19921}, in the Supplementary Information. % link?
+%We also report the results using another metric, Phylogenetic Distance \citep[PD;][]{Faith19921}, in the Supplementary Information.
+% TG: In the end is it really important to report PD?
+There are few problems that I don't understand in the function ses.pd (because I'm a bit lazy and trying to finish other things) such as the fact that the tree must be rooted for some data...
+Basically it worked well using one of the Fritz tree (which is bad, I know) but when using the Bininda-Emonds tree, I think the fact that it isn't resolved messes up the results and gives most taxa highly significant (clustered or dispersed).
 
@@
 \section{Data accessibility statement}
-All data and analysis code is available on GitHub.
+All data and analysis code is available on GitHub (\url{https://github.com/TGuillerme/Missing_living_mammals}).
 \section{Authors' Contributions}
@@
 \beginsupplement
-%\input{Supplementary/Supplementary}
+\input{Supplementary/Supplementary}
 

### 2015-06-21 Analysis/Results_summary.R

@@ -70,7 +70,7 @@ table.result(results, metric=c("NRI","NTI"), threshold=c(25,75), save.path="../M
 
 #Bugged!
 #caption<-"Number of taxa with available cladistic data for mammalian orders at three taxonomic levels. The coverage represents the proportion of taxa with available morphological data. The left vertical bar represents 25\\% of available data (``low'' coverage if \\textless 25\\%); The right vertical bar represents 75\\% of available data (``high'' coverage if \\textgreater 75\\%). When the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) are negative, taxa are more phylogenetically dispersed than expected by chance; when NRI or NTI are positive, taxa are more phylogenetically clustered by expected by chance. Significant NRI or NTI are highlighted in bold. One star (*) represents a p-value between 0.05 and 0.005; two starts between 0.005 and 0.0005 and three stars a p-value less than 0.0005."
-#table.result(results, metric=c("NRI","NTI", "PD"), threshold=c(25,75), save.path="../Manuscript/Supplementary/results_PD_100c/", file.save="Table_results_PD", caption=caption, environement="longtable")
+#table.result(results, metric=c("PD","NRI","NTI"), threshold=c(25,75), save.path="../Manuscript/Supplementary/results_PD_100c/", file.save="Table_results_PD", caption=caption, environement="longtable")
 
 
 ############################################
@@ -100,7 +100,7 @@ OTUS_total<-OTUS_number_val[-seq(from=1, to=length(OTUS_number_val), by=2)]
 supp_order<-orders[which(as.numeric(OTUS_total) >= 20)]
 
 #Supplementary figures (orders with more than 20 taxa)
-for (order in 1:length(supp_order)) {
+for (order in 11:length(supp_order)) {
     pdf(paste("../Manuscript/Supplementary/Supp_figure_", supp_order[[order]], ".pdf", sep=""))
     op<-par(mfrow=c(1,1), oma=c(0,1,0,1))
     plot.results(order=supp_order[[order]], tree=mam_tree, taxa=living_taxa_list, col_branch=c("blue", "grey"), reference=WR_list, verbose=TRUE)

### 2015-06-22 Manuscript/Missing_mammals-draft.tex
Edits to whole draft

--- 2015-06-21 Manuscript/Missing_mammals-draft.tex
@@
 %---------------------------------------------
-% NC: I've commented this out so I don't have to keep scrolling past in in the PDF
-%\newpage
-%\begin{abstract} % 200 words, all keywords written in abstract if possible.
-%Studying changes in global biodiversity through time and space is essential.
-%For that we need methods to combine both palaeontological and neontological data.
-%One promising method, the Total Evidence method, allows such thing but needs a lot of data.
-%Especially cladistic data from living taxa to allow the fossil taxa to accurately branch in the trees.
-%Despite two centuries of morphological studies on living taxa, scientists using and generating such data mainly focus on palaeontological data.
-%Therefore, even in well known groups such as mammal, there is a huge gap in our knowledge of cladistic data for living mammals.
-%In this study, using phylogenetic community structure methods, we quantify the availability of data in each mammalian order.
-%And maybe at the end of the paper we propose some discussion on how to improve all that (go in museums!).
-%\end{abstract}
+% NC: Time to sort this out!
+\newpage
+\begin{abstract} % 200 words, all keywords written in abstract if possible.
+Studying changes in global biodiversity through time and space is essential.
+For that we need methods to combine both palaeontological and neontological data.
+One promising method, the Total Evidence method, allows such thing but needs a lot of data.
+Especially cladistic data from living taxa to allow the fossil taxa to accurately branch in the trees.
+Despite two centuries of morphological studies on living taxa, scientists using and generating such data mainly focus on palaeontological data.
+Therefore, even in well known groups such as mammal, there is a huge gap in our knowledge of cladistic data for living mammals.
+In this study, using phylogenetic community structure methods, we quantify the availability of data in each mammalian order.
+And maybe at the end of the paper we propose some discussion on how to improve all that (go in museums!).
+\end{abstract}
 
-%\noindent ()\\
+\noindent ()\\
 
-%\vspace{1.5in}
+\vspace{1.5in}
 
@@
 We collected cladistic data (i.e.
-discrete morphological characters used in phylogenetics) from @256 phylogenetic matrices available online and measured the proportion of cladistic data available for each mammalian order.
+discrete morphological characters used in phylogenetics) from 286 phylogenetic matrices available online and measured the proportion of cladistic data available for each mammalian order.
 %Additionally, if the available data is randomly distributed across clades, the topological errors should be less extreme than if the data is biased towards particulars groups.
-Additionally, because missing morphological data in living species can influence tree topology as described above,%NC: Might still need to fix this sentence.
+Additionally, because missing morphological data in living species can influence tree topology as described above, %NC: Might still need to fix this sentence.
 "Topological caveats above" was too unclear.
@@
 \section{Materials and Methods}
+% NC: Need to add the dates, or date ranges, of your data collection below instead of @
 \subsection{Data collection and standardisation}
@@
 We also performed a systematic Google Scholar search (accessed @th of July 2015) for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
-In total, we downloaded 286 matrices containing a total of 11010 operational taxonomic units (OTUs) of which 5228 where unique. % Number of unique OTUs (including fossils)
-In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardized and ranged from specific individual specimen (i.e.
+In total, we downloaded 286 matrices containing a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique. % Number of unique OTUs (including fossils)
+In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardised and ranged from specific individual specimen names (i.e.
 the name of a collection item) to the family-level.
@@
 ``antler ramifications'' is a character that is only applicable to Cervidae not all mammals]{Brazeau2011}.
-% NC: Is this true really?
-Does it always follow that if you've very few characters you also have very specific characters?
-Maybe so...
-TG: I don't know if it's a generalisation but from the matrices I collected, the once with few characters always came from papers focusing on intra-specific or intra-generic variations.
-On the other hand, I didn't read all the characters.
-I just assume that in a big one, you have a lower probability of having non-applicable characters (there's only some many characters for coding antler ramification)
-Therefore we selected only matrices containing \textgreater 100 characters per OTU.
+Therefore we selected only matrices containing \textgreater 100 characters for each OTU.
 This threshold was chosen to correspond with the number of characters used in \citet{GuillermeCooper} and \citet{harrisonamong-character2014}.
-Note that the results of the analysis with no character threshold is available in the supplementary materials.
+Note that results of analyses with no character threshold are available in Supplementary Material. 
 After removing all matrices with \textless 100 characters, we retained 1074 unique living mammal OTUs from 126 matrices for our analyses. % 1601 unique living OTUs for 286 matrices (no threshold)
@@
 To assess the availability of cladistic data for each mammalian order, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
-We used these different taxonomic levels because some clades are well covered at the family- or genus-level, but poorly covered at the species-level.
-We consider orders with \textless 25\% of living taxa with cladistic data as having poor data coverage (hereafter ``low'' coverage), and orders with \textgreater 75\% of living taxa with cladistic data as having good data coverage (hereafter ``high'' coverage).
+%We used these different taxonomic levels because some clades are well covered at the family- or genus-level, but poorly covered at the species-level. % NC: Probably obvious?
+We consider orders with \textless 25\% of living taxa with cladistic data as having poor data coverage (``low'' coverage), and orders with \textgreater 75\% of living taxa with cladistic data as having good data coverage (hereafter ``high'' coverage).
 
 For orders with \textless 100\% cladistic data coverage at any taxonomic level, we investigated whether the available cladistic data was (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index \citep[NTI;][]{webb2002phylogenies} and the Net Relatedness Index \citep[NRI;][]{webb2002phylogenies}.
-%*** see end of methods for explanation. TG: I've but it below, I think it reads better after explaining the metrics.
-Both metrics 
+NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to clustering or overdispersion across the whole phylogeny \citep{Cooper2008}. 
+% NC: I moved this back up to keep it shorter.
 Both metrics were calculated using the \texttt{picante} package in R \citep{picante,R}.
-% NC: It's not going to add much to the wordcount to explain both. But move PD to supp
 
-% NC: Note that you had mis-defined MNND
 NTI \citep{webb2002phylogenies} is based on mean nearest neighbour distance (MNND) and is calculated as follows:
@@
   \end{equation}
-where $\overline{MNND}_{obs}$ is the observed mean distance between each of $n$ taxa with cladistic data, $\overline{MNND}_{n}$ is the mean of 1000 mean MNND between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random MNND values.
-% NC: Sorry this still isn't quite right.
-Check my paper for a nicer way of writing it.
-- TG: I've done some fixes on the sentence above.
-I didn't want to copy/paste what's in your paper so I tried to explain it in my own words.
-Maybe it'll need translation from Thomas-english and will end up sounding similar to what you wrote in your paper.
-Which is frankly not a big deal since it is the exact same thing we're describing here.
-%$\overline{MNND}_{n}$ is the expected random MNND for $n$ taxa estimated by calculating the MNND from $n$ taxa randomly drawn from the phylogeny and repeated 1000 times, and $\sigma(MNND_{n})$ is the standard deviation of the 1000 random MNND values.
+where $\overline{MNND}_{obs}$ is the observed mean distance between each of $n$ taxa with cladistic data and its nearest neighbour with cladistic data in the phylogeny, 
+% NC: Argh still wrong definition of MNND, I've fixed it myself. It's PAIRWISE!!!
+$\overline{MNND}_{n}$ is the mean of 1000 mean MNND between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random MNND values.
 NRI is similar but is based on mean phylogenetic distance (MPD) as follows:
@@
 where $\overline{MPD}_{obs}$ is the observed mean phylogenetic distance of the tree containing only the $n$ taxa with cladistic data, $\overline{MPD}_{n}$ is the expected random MPD for $n$ taxa estimated by calculating the MPD from $n$ taxa randomly drawn from the phylogeny and repeated 1000 times, and $\sigma(MPD_{n})$ is the standard deviation of the 1000 random MPD values.
-Negative NTI and NRI values show that the focal taxa are more overdispersed across the phylogeny than expected by chance, and positive values show that the focal taxa are more clustered across the phylogeny than expected by chance.
-%*** see end of methods for explanation.
-%Because the aim of this study is to understand the structure of the available cladistic data in mammals we prefered the NTI over the NRI or the PD metrics.
-%In fact, the NTI metric is more sensitive to the phylogenetic structure (clustering or over-dispersion) at the tips of phylogeny \citep{Cooper2008}.
-%and also because previous studies have shown that NRI is biased towards detecting overdispersion \citep{Kembel2006,Swenson2006}. TG: Useless in this case?
-%However, the results of the NRI and PD calculations are available in the supplementary.
-% NC: Actually I think we discussed that NRI and NTI show different things?
-So it's worth including both (you can get the wording from my 2008 paper).
-This can go up with the intro to the metrics somewhere (I have *** the two places it might fit)
-The result of both metric show slightly different aspects of the structure of the available data.
-NTI tends to be more sensitive to the phylogenetic structure at the tips of the phylogeney since it only takes into account the phylogenetic distance between taxa \citep{Cooper2008}.
-On the other hand the NRI is slightly biased towards detecting overdispersion since it is calculated using the mean phylogenetic distance from root to tips \citep{Kembel2006,Swenson2006}.
-% TG: In the end, I don't know if that's useful.
-For the 10 significant results, only three of them are not significant for both NRI and NTI...
-Maybe we can decide if we keep it or not after the word count.
+% NC: This can be shortened a bit but it's only gonna save you 50 words max.
+Negative NTI and NRI values show that the focal taxa are more overdispersed across the phylogeny than expected by chance, and positive values reflect significant clustering.
 
@@
 For each analysis our focal taxa were those with available cladistic data at that taxonomic level and the phylogeny was the phylogeny of the order pruned from \citet{BinindaEmonds}.
-% NC: Tried to make this more specific here. It's important to note that you used the order not the whole mammal phylogeny each time.
-We also report the results using another metric, Phylogenetic Distance \citep[PD;][]{Faith19921}, in the Supplementary Information. % link?
 
@@
 \section{Results}
-% TG: updated results.
-Slight changes in which group is significant or not.
-Mainly due to using Bininda-Emonds tree I'll say.
-But the main message is still the same.
-It's badly covered but at least is not that clustered.
-I just changed the example from Carnivores vs.
-cetartio to carnivores vs.
-primates (cetartio are now clustered).
-Across the @126 cladistic matrices we extracted, @22 out of 28 mammalian orders have low coverage (\textless 25\% of species with cladistic data) and @6 have high coverage (\textgreater 75\% of species with cladistic data) at the species-level.
-At the genus-level, only @three order have low coverage of taxa and @12 have high coverage, and at the family-level, no orders have low coverage and @23 have high coverage (Table \ref{Table_morpho_taxa_proportion}).
+% NC: Need to replace these @?
+Across the @126 cladistic matrices we extracted, @22 out of 28 mammalian orders have low coverage (\textless 25\% of species with cladistic data) and @six have high coverage (\textgreater 75\% of species with cladistic data) at the species-level.
+At the genus-level, @three orders have low coverage and @12 have high coverage, and at the family-level, no orders have low coverage and @23 have high coverage (Table \ref{Table_morpho_taxa_proportion}).
 
 \input{Table_results}
+% NC: Can probably cut words from the table legend too
 
-Among the mammalian orders containing OTUs with no available cladistic data, only 6 orders had significantly clustered data (Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha at both species- and genus-level and Afrosoricida and Pholidota at the species-level only) and no order had significantly overdispersed data at any taxonomic level (Table~\ref{Table_data_structure}).
+Among the mammalian orders containing OTUs with no available cladistic data, only six orders had significantly clustered data (Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha at both species- and genus-level and Afrosoricida and Pholidota at the species-level only) and no order had significantly overdispersed data at any taxonomic level (Table~\ref{Table_data_structure}).
 
@@
 \caption{Phylogenetic distribution of species with available cladistic data across two mammalian orders (A: Primates; B: Carnivora).
-Edges are colored in grey when no cladistic data is available for a species or in blue when data is available.}
+Edges are colored in grey when no cladistic data is available for a species and in blue when data is available.}
 \label{Figure_example_coverage}
@@
 We can save some cites by using \cite{BinindaEmonds,wilson2005mammal} that are cited all along the text already.
+% NC: Yep good idea.
 , most of the data used to build these phylogenies is molecular, and very little cladistic data is available for living mammals compared to fossil mammals \citep[e.g.][]{O'Leary08022013,ni2013oldest}.
@@
 Cladistic data coverage in living mammals varies across taxonomic levels and in its phylogenetic distribution.
-Higher taxonomic levels are always better sampled than lower ones and within these taxonomic levels, the available data is mostly randomly distributed across the phylogeny, apart from in 6 orders (see Results and Table\ref{Table_morpho_taxa_proportion}).
-% NC: Add a sentence saying this means things might not be quite so bad?
-Need to justify the "However".
-however what?
-TG: I think the "however" fits better below.
-Here it's §1: "that's what we found"; §2 (first part): "That's baaaad", §2 (second part) "However, it's all right at genus level"
+Higher taxonomic levels are always better sampled than lower ones and within these taxonomic levels, the available data is mostly randomly distributed across the phylogeny, apart from in six orders).
 
-The number of living mammalian taxa with no available cladistic data was surprisingly high at the species-level: only @6 out of @28 orders have a high coverage of taxa with available cladistic data (and two of the 28 orders are monospecific!).
+The number of living mammalian taxa with no available cladistic data was surprisingly high at the species-level: only @six out of @28 orders have a high coverage of taxa with available cladistic data (and two of the 28 orders are monospecific!).
 This high coverage threshold of 75\% of taxa with available cladistic data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.
@@
 This point is important from a practical point of view because of the slight discrepancy between the neontological and palaeontological concept of species.
-While neontological species are described using morphology, genetic distance, spatial distribution and even behaviour \citep[e.g.][]{kellymolecular2014}, palaeontological data can be based only on morphological, spatial and temporal data \citep[e.g.][]{ni2013oldest}.
-Because of this difference between neontological species (i.e.
-reproductive isolates) and paleontological species (i.e.
-morpho-species), most palaeontological studies are using the genus-level as their smallest OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
+While neontological species are described using morphology, genetic distance, spatial distribution and even behaviour, palaeontological species can be based only on morphological, spatial and temporal data \citep[e.g.][]{ni2013oldest}.
+Because of this, most palaeontological studies are using the genus as their smallest OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
 Thus data availability at the genus-level in living mammals should be our primary concern when aiming to build phylogenies of living and fossil taxa.
@@
 The second best scenario is that species with cladistic data are randomly distributed across the phylogeny. 
-In this scenario we expected no special bias in the placement of the fossil \citep{GuillermeCooper}, it is therefore encouraging that for most orders, species with cladistic data were randomly distributed across the phylogeny of each order.
+In this scenario we expect no special bias in the placement of the fossil \citep{GuillermeCooper}, it is therefore encouraging that for most orders, species with cladistic data were randomly distributed across the phylogeny of each order.
 The worst case scenario for fossil placement is that species with cladistic data are phylogenetically clustered. 
@@
 Our results suggest that this may be an issue, at the genus-level, in Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha. 
-For example, a Carnivora fossil will be unable to branch in the Herpestidae that has no species with cladistic data, and will also have more chance to branch, randomly, within the Canidae clade than any other clade in Carnivora (Figure \ref{Figure_example_coverage}-B).
-Thus, in Total Evidence trees, placements of Carnivoran fossils should considered with caution. %viewed with suspicion. % NC: OK gotta be a better way to say that. TG: is that a bit milder?
-
+For example, a Carnivora fossil will be unable to branch in the Herpestidae that has no species with cladistic data, and will also have more chance to branch, randomly, within the Canidae clade than any other clade in Carnivora (Figure \ref{Figure_example_coverage}B).
+Thus, in Total Evidence trees, placements of some carnivoran fossils should considered with caution. 
+% NC: Trying not to piss off GraHam - really it's weird stuff appearing in Carnivora that we are concerned by, not all Carnivora
 In this study, we treated all cladistic matrices as equal in a similar way to molecular matrices. 
@@
 
-Following the recommendations in \citep{GuillermeCooper}, to improve the accuracy of the topology of Total Evidence trees containing both living and fossil taxa, one should code cladistic characters for as many living species possible.
-%Because the data for living mammals is usually readily available in natural history collections, we propose that an increased effort be put into coding morphological characters from living species, possibly by engaging in collaborative data collection projects through web portals such as \textit{Morphobank} \citep{morphobank}.
-% NC: OK needs a summing up sentence!!!
-TG: I combined this paragraph with the postive ending point.
-That might sum up the thing.
-Despite the absence of good cladistic data coverage for living mammals, the Total Evidence methods seems to be still the most promising method for combining both living and fossil data and the best way to assess more accurately macroevolutionary processes.
-% TG: or is that a bit bold?
-It is also positive to consider that the problems underlined in this study can be easily solved.
-%TG: not sure about "easily", the idea is that the solution is not to break the speed of light or something like that, just to continue what people have stopped doing when DNA became trending.
-Because the data for living mammals is usually readily available in natural history collections, we propose that an increased effort be put into coding morphological characters from living species, possibly by engaging in collaborative data collection projects through web portals such as \textit{Morphobank} \citep{morphobank}.
+Despite the absence of good cladistic data coverage for living mammals, the Total Evidence methods still seems to be the most promising way of combining living and fossil data for macroevolutionary analyses.
+Following the recommendations in \citep{GuillermeCooper}, we need to code cladistic characters for as many living species possible. 
+Fortunately, data for living mammals is usually readily available in natural history collections, therefore, we propose that an increased effort be put into coding morphological characters from living species, possibly by engaging in collaborative data collection projects through web portals such as \textit{Morphobank} \citep{morphobank}.
+% NC: Still feels like it needs one final sentence to round it all off. Many something like below, but less crappy:
+Such an effort would be valuable not only to phylogeneticists, but also to any researcher interested in the diversity of life on Earth.
 

### 2015-06-23 Manuscript/Missing_mammals-draft.tex
Abstract

@@
 % TG: or some poetic... "Living and dead mammals: morphological data is lacking in living for combined analysis"
-
-%Key words: Total Evidence method, data structure, phylogenetic, living, fossil, topology
 \bigskip
@@
 $^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\
-$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
+$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\ %Should we remove that one for word count?
 $^3$Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.}\\
@@
 \newpage
-\begin{abstract} % 200 words, all keywords written in abstract if possible.
-Studying changes in global biodiversity through time and space is essential.
-For that we need methods to combine both palaeontological and neontological data.
-One promising method, the Total Evidence method, allows such thing but needs a lot of data.
-Especially cladistic data from living taxa to allow the fossil taxa to accurately branch in the trees.
-Despite two centuries of morphological studies on living taxa, scientists using and generating such data mainly focus on palaeontological data.
-Therefore, even in well known groups such as mammal, there is a huge gap in our knowledge of cladistic data for living mammals.
-In this study, using phylogenetic community structure methods, we quantify the availability of data in each mammalian order.
-And maybe at the end of the paper we propose some discussion on how to improve all that (go in museums!).
+\begin{abstract} % 200 words, all keywords written in abstract if possible. TG: That's 195 words for now.
+Combining living and fossil in the same analysis data is crucial for studying changes in global biodiversity through time.
+One method allowing to combine this data is the Total Evidence method that uses both molecular data for living species and morphological data for both living and fossil species.
+With this method, a good overlap of morphological data between living and fossil taxa is crucial for accurately inferring the phylogenies’ topology.
+
+Since the advent of DNA, molecular data has become easily and widely available.
+However, despite two centuries of morphological studies, scientists using and generating such data mainly focus on palaeontological data.
+Therefore, there is a gap in our knowledge of neontological morphological data even in well studied groups such as mammals.
+
+In this study, we quantify the morphological data available for living mammal taxa.
+We then analyse the structure of the available data by testing if it is clustered or evenly spread across the phylogeny.
+We found that 78\% of mammalian orders have less than 25\% data available at the species level.
+However, we found that the available is often randomly distributed among these orders apart from six of them where the data is clustered
 \end{abstract}
 
-\noindent ()\\
+%\noindent Key words: Total Evidence method, data structure, phylogenetic, fossil, topology\\ Excluded from word count.
 
@@
 \section{Materials and Methods}
-% NC: Need to add the dates, or date ranges, of your data collection below instead of @
 \subsection{Data collection and standardisation}
-We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases (accessed @th of July 2015): Morphobank \citep[\texttt{http://www.morphobank.org/};][]{morphobank}, Graeme Lloyd's website (\texttt{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce/cladistic-data}).
-We also performed a systematic Google Scholar search (accessed @th of July 2015) for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
+We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases (accessed 10th of June 2015): Morphobank \citep[\texttt{http://www.morphobank.org/};][]{morphobank}, Graeme Lloyd's website (\texttt{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce/cladistic-data}).
+We also performed a systematic Google Scholar search (accessed 11th of June 2015) for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
 In total, we downloaded 286 matrices containing a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique. % Number of unique OTUs (including fossils)
@@
 NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to clustering or overdispersion across the whole phylogeny \citep{Cooper2008}. 
-% NC: I moved this back up to keep it shorter.
 Both metrics were calculated using the \texttt{picante} package in R \citep{picante,R}.
@@
 where $\overline{MNND}_{obs}$ is the observed mean distance between each of $n$ taxa with cladistic data and its nearest neighbour with cladistic data in the phylogeny, 
-% NC: Argh still wrong definition of MNND, I've fixed it myself. It's PAIRWISE!!!
 $\overline{MNND}_{n}$ is the mean of 1000 mean MNND between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random MNND values.
@@
 \section{Results}
-% NC: Need to replace these @?
-Across the @126 cladistic matrices we extracted, @22 out of 28 mammalian orders have low coverage (\textless 25\% of species with cladistic data) and @six have high coverage (\textgreater 75\% of species with cladistic data) at the species-level.
-At the genus-level, @three orders have low coverage and @12 have high coverage, and at the family-level, no orders have low coverage and @23 have high coverage (Table \ref{Table_morpho_taxa_proportion}).
+Across the 126 cladistic matrices we extracted, 22 out of 28 mammalian orders have low coverage (\textless 25\% of species with cladistic data) and six have high coverage (\textgreater 75\% of species with cladistic data) at the species-level.
+At the genus-level, three orders have low coverage and 12 have high coverage, and at the family-level, no orders have low coverage and 23 have high coverage (Table \ref{Table_morpho_taxa_proportion}).
 
@@
 
-The number of living mammalian taxa with no available cladistic data was surprisingly high at the species-level: only @six out of @28 orders have a high coverage of taxa with available cladistic data (and two of the 28 orders are monospecific!).
+The number of living mammalian taxa with no available cladistic data was surprisingly high at the species-level: only six out of 28 orders have a high coverage of taxa with available cladistic data (and two of the 28 orders are monospecific!).
 This high coverage threshold of 75\% of taxa with available cladistic data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.
@@
 Thus, in Total Evidence trees, placements of some carnivoran fossils should considered with caution. 
-% NC: Trying not to piss off GraHam - really it's weird stuff appearing in Carnivora that we are concerned by, not all Carnivora
+% NC: Trying not to piss off GraHam - really it's weird stuff appearing in Carnivora that we are concerned by, not all Carnivora. TG: I think it's mainly Graham's work that focuses on canids.
 In this study, we treated all cladistic matrices as equal in a similar way to molecular matrices. 
@@
 % NC: Still feels like it needs one final sentence to round it all off. Many something like below, but less crappy:
-Such an effort would be valuable not only to phylogeneticists, but also to any researcher interested in the diversity of life on Earth.
+Such an effort would be valuable not only to phylogeneticists, but also to any researcher focusing understanding macroevolutionary patterns and processes.%interested in the diversity of life on Earth.
 

### 2015-06-23 Manuscript/Table_results.tex
Added the 1c threshold sup results
Updated manuscript (proper cite format)

@@
 % latex table generated in R 3.2.0 by xtable 1.7-4 package
-% Sat Jun 20 17:41:32 2015
+% Sun Jun 21 15:26:40 2015
 \begin{longtable}{lL{1.8cm}C{2cm}lcc}
 \caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels.
-The coverage represents the proportion of taxa with available morphological data.
-The left vertical bar represents 25\% of available data (``low'' coverage if \textless 25\%); The right vertical bar represents 75\% of available data (``high'' coverage if \textgreater 75\%).
-When the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) are negative, taxa are more phylogenetically dispersed than expected by chance; when NRI or NTI are positive, taxa are more phylogenetically clustered by expected by chance.
-Significant NRI or NTI are highlighted in bold.
-One star (*) represents a p-value between 0.05 and 0.005; two starts between 0.005 and 0.0005 and three stars a p-value less than 0.0005.} \\
+The left vertical bar represents ``low'' coverage (\textless 25\%); the right vertical bar represents ``high'' coverage (\textgreater 75\%).
+A negative Net Relatedness Index (NRI) and Nearest Taxon Index (NTI) shows more phylogeneticaly dispersed taxa than expected by chance; a positive value shows more phylogeneticaly clustered taxa than expected by chance.
+Significant NRI or NTI values are highlighted in bold.
+One star (*) signifies a p-value between 0.05 and 0.005; two starts between 0.005 and 0.0005 and three stars \textless 0.0005.} \\
   \hline
@@
   Afrosoricida & genus & 17/17 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar2.pdf} &   &   \\ 
-  \textbf{Afrosoricida} & \textbf{species} & \textbf{23/42} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar3.pdf} & \textbf{1.89*} & \textbf{1.19} \\ 
+  \textbf{Afrosoricida} & \textbf{species} & \textbf{23/42} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar3.pdf} & \textbf{1.89*} & 1.19 \\ 
   Carnivora & family & 11/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar4.pdf} & 0.43 & 1.68 \\ 
@@
   Cetartiodactyla & family & 21/21 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar7.pdf} &   &   \\ 
-  \textbf{Cetartiodactyla} & \textbf{genus} & \textbf{77/128} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar8.pdf} & \textbf{0.87} & \textbf{1.77*} \\ 
-  \textbf{Cetartiodactyla} & \textbf{species} & \textbf{129/310} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar9.pdf} & \textbf{2.72*} & \textbf{0.04} \\ 
+  \textbf{Cetartiodactyla} & \textbf{genus} & \textbf{77/128} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar8.pdf} & 0.87 & \textbf{1.77*} \\ 
+  \textbf{Cetartiodactyla} & \textbf{species} & \textbf{129/310} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar9.pdf} & \textbf{2.72*} & 0.04 \\ 
   Chiroptera & family & 13/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar10.pdf} & 0.55 & 0.63 \\ 

### 2015-06-24 Manuscript/Backup_version/Missing_mammals-draft copy.tex
Biology Letters version

--- 2015-06-23 Manuscript/Missing_mammals-draft.tex
@@
 \documentclass[12pt,letterpaper]{article}
-\usepackage{natbib}
 
@@
 \section{Introduction}
-There is an increasing consensus among evolutionary biologists that studying both living and fossil taxa is essential for fully understanding macroevolutionary patterns and processes \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013,Wood01032013}.
-For example, including both living and fossil taxa in evolutionary studies can improve the accuracy of timing diversification events \citep[e.g.][]{ronquista2012}, our understanding of relationships among lineages \citep[e.g.][]{beckancient2014}, and our ability to infer biogeographical patterns through time \citep[e.g.][]{Meseguer01032015}.
+There is an increasing consensus among evolutionary biologists that studying both living and fossil taxa is essential for fully understanding macroevolutionary patterns and processes \cite{slaterunifying2013,fritzdiversity2013,Wood01032013}.
+For example, including both living and fossil taxa in evolutionary studies can improve the accuracy of timing diversification events (e.g.
+\cite{ronquista2012}, our understanding of relationships among lineages (e.g.
+\cite{beckancient2014}, and our ability to infer biogeographical patterns through time (e.g.
+\cite{Meseguer01032015}.
 To perform such analyses it is necessary to combine living and fossil taxa in phylogenetic trees.
-One increasingly popular method, the Total Evidence method \citep{eernissetaxonomic1993,ronquista2012}, combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014,Meseguer01032015}, producing a phylogeny with living and fossil taxa at the tips.
-These phylogenies can be dated using methods such as tip-dating \citep{ronquista2012,Drummond01082012,Wood01032013} and incorporated into macroevolutionary studies \citep[e.g.][]{ronquista2012,Wood01032013,slaterphylogenetic2013}.
+One increasingly popular method, the Total Evidence method \cite{eernissetaxonomic1993,ronquista2012}, combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix (e.g.
+\cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014,Meseguer01032015}, producing a phylogeny with living and fossil taxa at the tips.
+These phylogenies can be dated using methods such as tip-dating \cite{ronquista2012,Wood01032013} and incorporated into macroevolutionary studies (e.g.
+\cite{ronquista2012,Wood01032013,slaterphylogenetic2013}.
 
@@
 For example, fossils very rarely have molecular data and incomplete fossil preservation (e.g.
-soft \textit{vs.} hard tissues) may restrict the amount of morphological data available \citep{sansomfossilization2013}.
+soft \textit{vs.} hard tissues) may restrict the amount of morphological data available \cite{sansomfossilization2013}.
 Additionally, since the molecular phylogenetics revolution, it has become less common to collect morphological characters for living taxa when molecular data is available (e.g.
 in \cite{slaterphylogenetic2013}, only 13\% of the 169 living taxa have coded morphological data).
-Unfortunately this missing data can lead to errors in phylogenetic inference; in fact, simulations show that the ability of the Total Evidence method to recover the correct phylogenetic topology decreases when there is a low overlap between morphological data in the living and fossil taxa \citep{GuillermeCooper}, regardless the overall amount of morphological data available for the fossils (or the amount of molecular data available for the living species).
+Unfortunately this missing data can lead to errors in phylogenetic inference; in fact, simulations show that the ability of the Total Evidence method to recover the correct phylogenetic topology decreases when there is a low overlap between morphological data in the living and fossil taxa \cite{GuillermeCooper}, regardless the overall amount of morphological data available for the fossils (or the amount of molecular data available for the living species).
 The effect of missing data on topology is greatest when living taxa have few morphological data.
-This is because (1) a fossil cannot branch in the correct clade if there is no overlapping morphological data in the clade; and (2) a fossil has a higher probability of branching within a clade with more morphological data available for living taxa, regardless of whether this is the correct clade or not \citep{GuillermeCooper}.
+This is because (1) a fossil cannot branch in the correct clade if there is no overlapping morphological data in the clade; and (2) a fossil has a higher probability of branching within a clade with more morphological data available for living taxa, regardless of whether this is the correct clade or not \cite{GuillermeCooper}.
 
@@
 \subsection{Data collection and standardisation}
-We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases (accessed 10th of June 2015): Morphobank \citep[\texttt{http://www.morphobank.org/};][]{morphobank}, Graeme Lloyd's website (\texttt{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce/cladistic-data}).
+We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases (accessed 10th of June 2015): Morphobank (\url{http://www.morphobank.org/} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{https://github.com/rossmounce/cladistic-data}).
 We also performed a systematic Google Scholar search (accessed 11th of June 2015) for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
@@
 sapiens} was transformed to \textit{Homo sapiens}).
-We designated as ``living'' all OTUs that were either present in the phylogeny of \citet{BinindaEmonds} or the taxonomy of \citet{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\texttt{https://paleobiodb.org/}).
+We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\url{https://paleobiodb.org/}).
 For OTUs that did not appear in these three sources, we first decomposed the name (i.e.
@@
 The number of characters in each matrix ranged from 6 to 4541.
-Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having characters that overlap with those of other matrices \citep{wagner2000} and (2) they are more likely to contain a higher proportion of specific characters that are not-applicable across large clades \citep[][e.g.
-``antler ramifications'' is a character that is only applicable to Cervidae not all mammals]{Brazeau2011}.
+Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having characters that overlap with those of other matrices \cite{wagner2000} and (2) they are more likely to contain a higher proportion of specific characters that are not-applicable across large clades (e.g.
+``antler ramifications'' is a character that is only applicable to Cervidae not all mammals \cite{Brazeau2011}).
 Therefore we selected only matrices containing \textgreater 100 characters for each OTU.
-This threshold was chosen to correspond with the number of characters used in \citet{GuillermeCooper} and \citet{harrisonamong-character2014}.
+This threshold was chosen to correspond with the number of characters used in \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
 Note that results of analyses with no character threshold are available in Supplementary Material. 
@@
 
-For orders with \textless 100\% cladistic data coverage at any taxonomic level, we investigated whether the available cladistic data was (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index \citep[NTI;][]{webb2002phylogenies} and the Net Relatedness Index \citep[NRI;][]{webb2002phylogenies}.
-NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to clustering or overdispersion across the whole phylogeny \citep{Cooper2008}. 
-Both metrics were calculated using the \texttt{picante} package in R \citep{picante,R}.
+For orders with \textless 100\% cladistic data coverage at any taxonomic level, we investigated whether the available cladistic data was (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index (NTI; \cite{webb2002phylogenies}) and the Net Relatedness Index (NRI; \cite{webb2002phylogenies}).
+NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to clustering or overdispersion across the whole phylogeny \cite{Cooper2008}. 
+Both metrics were calculated using the \texttt{picante} package in R \cite{picante,R}.
 
-NTI \citep{webb2002phylogenies} is based on mean nearest neighbour distance (MNND) and is calculated as follows:
+NTI \cite{webb2002phylogenies} is based on mean nearest neighbour distance (MNND) and is calculated as follows:
   \begin{equation}
@@
 We calculated NTI and NRI values for each mammalian order separately, at each different taxonomic level. 
-For each analysis our focal taxa were those with available cladistic data at that taxonomic level and the phylogeny was the phylogeny of the order pruned from \citet{BinindaEmonds}.
+For each analysis our focal taxa were those with available cladistic data at that taxonomic level and the phylogeny was the phylogeny of the order pruned from \cite{BinindaEmonds}.
 
@@
 \section{Discussion}
-Our results show that although phylogenetic relationships among living mammals are well-resolved \citep[e.g.][]{FritzTree,meredithimpacts2011,May-Collado-PeerJ} % TG: note that we can reduce the cites here (for the world count in the refs).
-All three are cited only one time (here).
-We can save some cites by using \cite{BinindaEmonds,wilson2005mammal} that are cited all along the text already.
-% NC: Yep good idea.
-, most of the data used to build these phylogenies is molecular, and very little cladistic data is available for living mammals compared to fossil mammals \citep[e.g.][]{O'Leary08022013,ni2013oldest}.
-This has implications for building Total Evidence phylogenies containing both living and fossil mammals, as without sufficient cladistic data for living species, fossil placements in these trees are very uncertain \citep{GuillermeCooper}.
+Our results show that although phylogenetic relationships among living mammals are well-resolved (e.g.
+\cite{BinindaEmonds,meredithimpacts2011}) , most of the data used to build these phylogenies is molecular, and very little cladistic data is available for living mammals compared to fossil mammals (e.g.
+\cite{O'Leary08022013,ni2013oldest}).
+This has implications for building Total Evidence phylogenies containing both living and fossil mammals, as without sufficient cladistic data for living species, fossil placements in these trees are very uncertain \cite{GuillermeCooper}.
 Cladistic data coverage in living mammals varies across taxonomic levels and in its phylogenetic distribution.
@@
 The number of living mammalian taxa with no available cladistic data was surprisingly high at the species-level: only six out of 28 orders have a high coverage of taxa with available cladistic data (and two of the 28 orders are monospecific!).
-This high coverage threshold of 75\% of taxa with available cladistic data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.
-Beyond this threshold, there is considerable displacement of wildcard taxa (\textit{sensu} \citep{kearneyfragmentary2002}) and decreases in clade conservation \citep{GuillermeCooper}.
+This high coverage threshold of 75\% of taxa with available cladistic data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
+Beyond this threshold, there is considerable displacement of wildcard taxa (\textit{sensu} \cite{kearneyfragmentary2002}) and decreases in clade conservation \cite{GuillermeCooper}.
 Therefore we expect a high probability of topological artefacts for the placement of fossil taxa at the species-level in most mammalian orders.
@@
 This point is important from a practical point of view because of the slight discrepancy between the neontological and palaeontological concept of species.
-While neontological species are described using morphology, genetic distance, spatial distribution and even behaviour, palaeontological species can be based only on morphological, spatial and temporal data \citep[e.g.][]{ni2013oldest}.
-Because of this, most palaeontological studies are using the genus as their smallest OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
+While neontological species are described using morphology, genetic distance, spatial distribution and even behaviour, palaeontological species can be based only on morphological, spatial and temporal data (e.g.
+\cite{ni2013oldest}).
+Because of this, most palaeontological studies are using the genus as their smallest OTU (e.g. \cite{ni2013oldest,O'Leary08022013}).
 Thus data availability at the genus-level in living mammals should be our primary concern when aiming to build phylogenies of living and fossil taxa.
@@
 The second best scenario is that species with cladistic data are randomly distributed across the phylogeny. 
-In this scenario we expect no special bias in the placement of the fossil \citep{GuillermeCooper}, it is therefore encouraging that for most orders, species with cladistic data were randomly distributed across the phylogeny of each order.
+In this scenario we expect no special bias in the placement of the fossil \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with cladistic data were randomly distributed across the phylogeny of each order.
 The worst case scenario for fossil placement is that species with cladistic data are phylogenetically clustered. 
@@
 Unfortunately, cladistic data cannot always be treated in this way because some characters may overlap.
-For example, if matrix A has a character coding for the shape of a particular morphological feature and matrix B has a character coding for the presence of this same morphological feature and a second character coding for its shape, then these three characters are non-independent compound characters \citep{Brazeau2011}.
-However, in reasonably sized matrices (\textgreater 100 characters; \citealp{GuillermeCooper,harrisonamong-character2014}) it is more likely that a number of characters are consistently conserved among the different matrices and thus easily combinable.
-For example, within the Primate cladistic literature, the character \textit{p7} - the size of the $4^{th}$ lower premolar paraconid - has been used consistently for \textgreater 15 years \citep[e.g.][]{ross1998phylogenetic,marivaux2005anthropoid,kay2008anatomy,boyer2010astragalar,ni2013oldest} and can therefore be combined among the matrices.
+For example, if matrix A has a character coding for the shape of a particular morphological feature and matrix B has a character coding for the presence of this same morphological feature and a second character coding for its shape, then these three characters are non-independent compound characters \cite{Brazeau2011}.
+However, in reasonably sized matrices (\textgreater 100 characters; \cite{GuillermeCooper,harrisonamong-character2014}) it is more likely that a number of characters are consistently conserved among the different matrices and thus easily combinable.
+For example, within the Primate cladistic literature, the character \textit{p7} - the size of the $4^{th}$ lower premolar paraconid - has been used consistently for \textgreater 15 years (e.g.
+\cite{ross1998phylogenetic,marivaux2005anthropoi,ni2013oldest} and can therefore be combined among the matrices.
 A conservative approach to avoid compound characters would be to select only the most recent matrix for each group, but this would result in the loss of a lot of data.
@@
 Despite the absence of good cladistic data coverage for living mammals, the Total Evidence methods still seems to be the most promising way of combining living and fossil data for macroevolutionary analyses.
-Following the recommendations in \citep{GuillermeCooper}, we need to code cladistic characters for as many living species possible. 
-Fortunately, data for living mammals is usually readily available in natural history collections, therefore, we propose that an increased effort be put into coding morphological characters from living species, possibly by engaging in collaborative data collection projects through web portals such as \textit{Morphobank} \citep{morphobank}.
+Following the recommendations in \cite{GuillermeCooper}, we need to code cladistic characters for as many living species possible. 
+Fortunately, data for living mammals is usually readily available in natural history collections, therefore, we propose that an increased effort be put into coding morphological characters from living species, possibly by engaging in collaborative data collection projects through web portals such as \textit{Morphobank} \cite{morphobank}.
 % NC: Still feels like it needs one final sentence to round it all off. Many something like below, but less crappy:
@@
 
-\bibliographystyle{sysbio}
+\bibliographystyle{vancouver}
 \bibliography{References}

### 2015-06-24 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Biology Letters version

--- 2015-06-24 Manuscript/Backup_version/Missing_mammals-draft copy.tex
@@
 
-\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$$^,$$^*$ and Natalie Cooper$^1$$^,$$^2$$^,$$^3$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^*$ and Natalie Cooper$^1$$^,$$^2$}\\
 \noindent {\small \it 
 $^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\
-$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\ %Should we remove that one for word count?
-$^3$Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.}\\
+$^2$Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.}\\
 \end{center}
@@
 
-%Line numbering
-\modulolinenumbers[1]
-\linenumbers
+%Line numbering REMOVE FOR WORDCOUNT
+%\modulolinenumbers[1]
+%\linenumbers
 
@@
 \begin{abstract} % 200 words, all keywords written in abstract if possible. TG: That's 195 words for now.
+% TG: BIOLOGY LETTERS FORMAT: cut down to 150 words?
 Combining living and fossil in the same analysis data is crucial for studying changes in global biodiversity through time.
@@
 \section{Introduction}
+% TG: BIOLOGY LETTERS FORMAT: Combine paragraph 1 and 2?
+
 There is an increasing consensus among evolutionary biologists that studying both living and fossil taxa is essential for fully understanding macroevolutionary patterns and processes \cite{slaterunifying2013,fritzdiversity2013,Wood01032013}.
@@
 \subsection{Data collection and standardisation}
+
+%TG: I think we can put a lot of that in the supplementaries: the essential is:
+% 1-We download matrices
+% 2-We selected living OTUs with > 100 characters
+% 3-We counted the percentage of available OTUs on Bininda-Emonds tree (low/high coverage)
+% 4-We calculated NTI/NRI
+
 We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases (accessed 10th of June 2015): Morphobank (\url{http://www.morphobank.org/} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{https://github.com/rossmounce/cladistic-data}).
@@
 % NC: Can probably cut words from the table legend too 
+% TG: YES! Otherwise I think that the text is down to the minimum (we can obviously remove few words here and there)
 
@@
 \section{Discussion}
+%TG: Again, we can cut a shit load down. The essential would be:
+% 1-Not much data is available
+% 2-But at least is randomly distributed
+% 3-We need to collect more of it
+
 Our results show that although phylogenetic relationships among living mammals are well-resolved (e.g.
@@
 \bibliography{References}
+% TG: we can maybe remove one or two references (28) but I think we're already at the minimum...
 

### 2015-06-25 Manuscript/BiologyLetters/Table_results.tex
Word count massacre

--- 2015-06-23 Manuscript/Table_results.tex
@@
 \caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels.
-The left vertical bar represents ``low'' coverage (\textless 25\%); the right vertical bar represents ``high'' coverage (\textgreater 75\%).
-A negative Net Relatedness Index (NRI) and Nearest Taxon Index (NTI) shows more phylogeneticaly dispersed taxa than expected by chance; a positive value shows more phylogeneticaly clustered taxa than expected by chance.
-Significant NRI or NTI values are highlighted in bold.
-One star (*) signifies a p-value between 0.05 and 0.005; two starts between 0.005 and 0.0005 and three stars \textless 0.0005.} \\
+The left vertical bar represents low coverage (\textless 25\%); the right vertical bar represents high coverage (\textgreater 75\%).
+Negative Net Relatedness Index (NRI) and Nearest Taxon Index (NTI) values indicate phylogenetic overdispersion; positive values indicate phylogenetic clustering.
+Significant NRI or NTI values are in bold.
+*p \textless 0.05; **p \textless 0.01; ***p \textless 0.001.} \\
   \hline

### 2015-06-25 Manuscript/BiologyLetters/Missing_mammals-draft.tex
added pdf figures for bio let

--- 2015-06-24 Manuscript/Backup_version/Missing_mammals-draft copy.tex
@@
 \end{flushright}
-\bigskip
-\noindent RH: Morphological data availability in living mammals
 
@@
 
-\noindent{\Large \bf Morphological data is lacking for living mammals}
-%Or: Missing cladistic data in living mammals - but I prefer the first one emphasizing on the availability of it. The data is not missing per se, it's just not coded!
-% NC: sometimes the better title is one that answers the question rather than giving the topic. e.g., "Morphological data is lacking for living species" or similar
-% TG: maybe worth considering the special issue as well? "Putting fossils in tree" so maybe something like "There is not enough morphological data for living mammals to branch fossils in the trees."
-% TG: or some poetic... "Living and dead mammals: morphological data is lacking in living for combined analysis"
+\noindent{\Large \bf Assessment of cladistic data availability for living mammals}
+
 \bigskip
 
-\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$$^,$$^*$ and Natalie Cooper$^1$$^,$$^2$$^,$$^3$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^*$ and Natalie Cooper$^1$$^,$$^2$}\\
 \noindent {\small \it 
 $^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\
-$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\ %Should we remove that one for word count?
-$^3$Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.}\\
+$^2$Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.}\\
 \end{center}
 \medskip
-\noindent{*\bf Corresponding author.} \textit{Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.}\\  
+\noindent{*\bf Corresponding author.} \textit{guillert@tcd.ie}\\  
 \vspace{1in}
 
-%Line numbering
-\modulolinenumbers[1]
-\linenumbers
+%Line numbering REMOVE FOR WORDCOUNT
+%\modulolinenumbers[1]
+%\linenumbers
 
@@
 %---------------------------------------------
-% NC: Time to sort this out!
 \newpage
-\begin{abstract} % 200 words, all keywords written in abstract if possible. TG: That's 195 words for now.
+\begin{abstract}
 Combining living and fossil in the same analysis data is crucial for studying changes in global biodiversity through time.
@@
 
-%\noindent Key words: Total Evidence method, data structure, phylogenetic, fossil, topology\\ Excluded from word count.
+%\noindent Key words: Total Evidence method, phylogenetic clustering, morphological matrix, fossil, topology\\ Excluded from word count.
 
@@
 \section{Introduction}
-There is an increasing consensus among evolutionary biologists that studying both living and fossil taxa is essential for fully understanding macroevolutionary patterns and processes \cite{slaterunifying2013,fritzdiversity2013,Wood01032013}.
-For example, including both living and fossil taxa in evolutionary studies can improve the accuracy of timing diversification events (e.g.
-\cite{ronquista2012}, our understanding of relationships among lineages (e.g.
-\cite{beckancient2014}, and our ability to infer biogeographical patterns through time (e.g.
-\cite{Meseguer01032015}.
+
+There is an increasing consensus among biologists that studying both living and fossil taxa is essential for fully understanding macroevolutionary patterns and processes \cite{slaterunifying2013,fritzdiversity2013}.
 To perform such analyses it is necessary to combine living and fossil taxa in phylogenetic trees.
 One increasingly popular method, the Total Evidence method \cite{eernissetaxonomic1993,ronquista2012}, combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix (e.g.
-\cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014,Meseguer01032015}, producing a phylogeny with living and fossil taxa at the tips.
-These phylogenies can be dated using methods such as tip-dating \cite{ronquista2012,Wood01032013} and incorporated into macroevolutionary studies (e.g.
-\cite{ronquista2012,Wood01032013,slaterphylogenetic2013}.
-
-A downside of the Total Evidence method is that it requires a lot of data.
-One must collect molecular data for living taxa and morphological data for both living and fossil taxa; two types of data that require fairly different technical skills (e.g.
-molecular sequencing \textit{vs.} anatomical description).
-Additionally, large chunks of this data can be difficult, or even impossible, to collect for every taxon present in the analysis.
-For example, fossils very rarely have molecular data and incomplete fossil preservation (e.g.
-soft \textit{vs.} hard tissues) may restrict the amount of morphological data available \cite{sansomfossilization2013}.
-Additionally, since the molecular phylogenetics revolution, it has become less common to collect morphological characters for living taxa when molecular data is available (e.g.
-in \cite{slaterphylogenetic2013}, only 13\% of the 169 living taxa have coded morphological data).
-Unfortunately this missing data can lead to errors in phylogenetic inference; in fact, simulations show that the ability of the Total Evidence method to recover the correct phylogenetic topology decreases when there is a low overlap between morphological data in the living and fossil taxa \cite{GuillermeCooper}, regardless the overall amount of morphological data available for the fossils (or the amount of molecular data available for the living species).
-The effect of missing data on topology is greatest when living taxa have few morphological data.
-This is because (1) a fossil cannot branch in the correct clade if there is no overlapping morphological data in the clade; and (2) a fossil has a higher probability of branching within a clade with more morphological data available for living taxa, regardless of whether this is the correct clade or not \cite{GuillermeCooper}.
+\cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a phylogeny with living and fossil taxa at the tips.
+A downside of this method is that it requires molecular data for living taxa and morphological data for both living and fossil taxa.
+Chunks of this data can be difficult, or impossible, to collect for every taxon in the analysis.
+For example, fossils rarely have molecular data and incomplete fossil preservation may restrict the amount of morphological data available.
+Additionally, it has become less common to collect morphological characters for living taxa when molecular data is available (e.g.
+in \cite{slaterphylogenetic2013}, only 13\% of living taxa have coded morphological data).
+Unfortunately this missing data can lead to errors in phylogenetic inference.
+Simulations show that the ability of the Total Evidence method to recover the correct topology decreases when there is little overlap between morphological data in living and fossil taxa, and that the effect of missing data on topology is greatest when living taxa have few morphological data \cite{GuillermeCooper}.
+This is because (1) fossils cannot branch in the correct clade if it contains no morphological data for living taxa; and (2) fossils have a higher probability of branching within clades with more morphological data for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.
 
 The issues above highlight that it is crucial to have sufficient morphological data for living taxa in a clade before using a Total Evidence approach.
-However, it is unclear how much morphological data for living taxa is actually available (i.e.
-already coded from museum specimens and deposited in phylogenetic matrices accessible online), and how this data is distributed across clades.
+However, it is unclear how much morphological data for living taxa is actually available, i.e.
+already coded from museum specimens and deposited in phylogenetic matrices accessible online, and how this data is distributed across clades.
 Intuitively, most people assume this kind of data has already been collected, but empirical data suggest otherwise (e.g. in \cite{ronquista2012,slaterphylogenetic2013,beckancient2014}).
 To investigate this further, we assess the amount of available morphological data for living mammals to determine whether sufficient data exists to build reliable Total Evidence phylogenies in this group.
-We collected cladistic data (i.e.
-discrete morphological characters used in phylogenetics) from 286 phylogenetic matrices available online and measured the proportion of cladistic data available for each mammalian order.
-%Additionally, if the available data is randomly distributed across clades, the topological errors should be less extreme than if the data is biased towards particulars groups.
-Additionally, because missing morphological data in living species can influence tree topology as described above, %NC: Might still need to fix this sentence.
-"Topological caveats above" was too unclear.
-we determined whether the available cladistic data was phylogenetically overdispersed or clustered in the mammalian orders where data was missing. 
-We find that available morphological data for living mammals is scarce but generally randomly distributed across phylogenies. 
-We recommend that efforts be made to collect and share more cladistic data for living species to improve the accuracy of Total Evidence phylogenies.
+We also determine whether the available cladistic data is phylogenetically overdispersed or clustered across mammalian orders.
 
-% NC: If there are word count problems we can ditch the last two sentences.
 %---------------------------------------------
@@
 \subsection{Data collection and standardisation}
-We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases (accessed 10th of June 2015): Morphobank (\url{http://www.morphobank.org/} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{https://github.com/rossmounce/cladistic-data}).
-We also performed a systematic Google Scholar search (accessed 11th of June 2015) for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
-In total, we downloaded 286 matrices containing a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique. % Number of unique OTUs (including fossils)
-In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardised and ranged from specific individual specimen names (i.e.
+We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases %(accessed 10th of June 2015)
+: Morphobank (\url{http://www.morphobank.org/} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{https://github.com/rossmounce/cladistic-data}).
+We also performed a systematic Google Scholar search %(accessed 11th of June 2015) 
+for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
+% In total, we downloaded 286 matrices containing a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique. % Number of unique OTUs (including fossils)
+In total, we downloaded 286 matrices containing 5228 unique operational taxonomic units (OTUs). 
+We used OTUs rather than species since entries in the matrices ranged from species to families, and standardised the taxonomy as described in Supplementary Materials. 
+We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}.
+
+%In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardised and ranged from specific individual specimen names (i.e.
 the name of a collection item) to the family-level.
-Where possible, we considered OTUs at their lowest valid taxonomic level (i.e. species) but some OTUs were only valid at a higher taxonomic level (e.g. genus or family).
-Therefore for some orders, we sampled more genera than species (Table \ref{Table_morpho_taxa_proportion}).
+%Where possible, we considered OTUs at their lowest valid taxonomic level (i.e. species) but some OTUs were only valid at a higher taxonomic level (e.g. genus or family).
+%Therefore for some orders, we sampled more genera than species (Table \ref{Table_morpho_taxa_proportion}).
 
-To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H.
+%To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H.
 sapiens} was transformed to \textit{Homo sapiens}).
-We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\url{https://paleobiodb.org/}).
-For OTUs that did not appear in these three sources, we first decomposed the name (i.e.
+%We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\url{https://paleobiodb.org/}).
+%For OTUs that did not appear in these three sources, we first decomposed the name (i.e.
 \textit{Homo sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
-Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; see Supplementary Material for more details).
+%Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; see Supplementary Material for more details).
 
-The number of characters in each matrix ranged from 6 to 4541.
+% The number of characters in each matrix ranged from 6 to 4541.
 Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having characters that overlap with those of other matrices \cite{wagner2000} and (2) they are more likely to contain a higher proportion of specific characters that are not-applicable across large clades (e.g.
@@
 This threshold was chosen to correspond with the number of characters used in \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
-Note that results of analyses with no character threshold are available in Supplementary Material. 
-After removing all matrices with \textless 100 characters, we retained 1074 unique living mammal OTUs from 126 matrices for our analyses. % 1601 unique living OTUs for 286 matrices (no threshold)
+Results of analyses with no threshold are available in Supplementary Material. 
+After removing matrices with \textless 100 characters, we retained 1074 unique living mammal OTUs from 126 matrices. % 1601 unique living OTUs for 286 matrices (no threshold)
 
@@
 %We used these different taxonomic levels because some clades are well covered at the family- or genus-level, but poorly covered at the species-level. % NC: Probably obvious?
-We consider orders with \textless 25\% of living taxa with cladistic data as having poor data coverage (``low'' coverage), and orders with \textgreater 75\% of living taxa with cladistic data as having good data coverage (hereafter ``high'' coverage).
+We consider orders with \textless 25\% of living taxa with cladistic data as having low data coverage, and orders with \textgreater 75\% of living taxa with cladistic data as having high data coverage.
 
-For orders with \textless 100\% cladistic data coverage at any taxonomic level, we investigated whether the available cladistic data was (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index (NTI; \cite{webb2002phylogenies}) and the Net Relatedness Index (NRI; \cite{webb2002phylogenies}).
+We investigated whether the available cladistic data for each order was (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index (NTI; \cite{webb2002phylogenies}) and the Net Relatedness Index (NRI; \cite{webb2002phylogenies}).
 NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to clustering or overdispersion across the whole phylogeny \cite{Cooper2008}. 
@@
 $\overline{MNND}_{n}$ is the mean of 1000 mean MNND between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random MNND values.
-NRI is similar but is based on mean phylogenetic distance (MPD) as follows:
+NRI is calculated in the same way, but MNND is replaced by mean phylogenetic distance (MPD) as follows:
   \begin{equation}
@@
   \end{equation}
-where $\overline{MPD}_{obs}$ is the observed mean phylogenetic distance of the tree containing only the $n$ taxa with cladistic data, $\overline{MPD}_{n}$ is the expected random MPD for $n$ taxa estimated by calculating the MPD from $n$ taxa randomly drawn from the phylogeny and repeated 1000 times, and $\sigma(MPD_{n})$ is the standard deviation of the 1000 random MPD values.
-% NC: This can be shortened a bit but it's only gonna save you 50 words max.
-Negative NTI and NRI values show that the focal taxa are more overdispersed across the phylogeny than expected by chance, and positive values reflect significant clustering.
+where $\overline{MPD}_{obs}$ is the observed mean phylogenetic distance of the tree containing only the $n$ taxa with cladistic data.
+Negative NTI and NRI values show that the focal taxa are more overdispersed across the phylogeny than expected by chance, and positive values reflect clustering.
 
 We calculated NTI and NRI values for each mammalian order separately, at each different taxonomic level. 
-For each analysis our focal taxa were those with available cladistic data at that taxonomic level and the phylogeny was the phylogeny of the order pruned from \cite{BinindaEmonds}.
+For each analysis our focal taxa were those with available cladistic data at that taxonomic level and the phylogeny was that of the order pruned from \cite{BinindaEmonds}.
 
@@
 \section{Results}
-Across the 126 cladistic matrices we extracted, 22 out of 28 mammalian orders have low coverage (\textless 25\% of species with cladistic data) and six have high coverage (\textgreater 75\% of species with cladistic data) at the species-level.
+22 of 28 orders have low coverage (\textless 25\% species with cladistic data) and six have high coverage (\textgreater 75\% species with cladistic data) at the species-level.
 At the genus-level, three orders have low coverage and 12 have high coverage, and at the family-level, no orders have low coverage and 23 have high coverage (Table \ref{Table_morpho_taxa_proportion}).
@@
 \input{Table_results}
-% NC: Can probably cut words from the table legend too
 
-Among the mammalian orders containing OTUs with no available cladistic data, only six orders had significantly clustered data (Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha at both species- and genus-level and Afrosoricida and Pholidota at the species-level only) and no order had significantly overdispersed data at any taxonomic level (Table~\ref{Table_data_structure}).
+Only six orders had significantly clustered data (Afrosoricida and Pholidota at the species-level, and Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha at both species- and genus-level) and none had significantly overdispersed data (Table~\ref{Table_data_structure}).
 
-Two contrasting results are shown in Figure \ref{Figure_example_coverage} with randomly distributed OTUs with cladistic data in Primates (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered OTUs with cladistic data in Carnivora (mainly Canidae; Figure \ref{Figure_example_coverage}B).
+Figure \ref{Figure_example_coverage} shows randomly distributed OTUs with cladistic data in Primates (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered OTUs with cladistic data in Carnivora (mainly Canidae; Figure \ref{Figure_example_coverage}B).
 
@@
     \includegraphics[width=1\textwidth]{example_coverage.pdf}
-\caption{Phylogenetic distribution of species with available cladistic data across two mammalian orders (A: Primates; B: Carnivora).
-Edges are colored in grey when no cladistic data is available for a species and in blue when data is available.}
+\caption{Phylogenetic distribution of species with available cladistic data across two orders (A: Primates; B: Carnivora).
+Blue branches indicate available cladistic data for the species.}
 \label{Figure_example_coverage}
@@
 This has implications for building Total Evidence phylogenies containing both living and fossil mammals, as without sufficient cladistic data for living species, fossil placements in these trees are very uncertain \cite{GuillermeCooper}.
-Cladistic data coverage in living mammals varies across taxonomic levels and in its phylogenetic distribution.
-Higher taxonomic levels are always better sampled than lower ones and within these taxonomic levels, the available data is mostly randomly distributed across the phylogeny, apart from in six orders).
 
-The number of living mammalian taxa with no available cladistic data was surprisingly high at the species-level: only six out of 28 orders have a high coverage of taxa with available cladistic data (and two of the 28 orders are monospecific!).
+The number of living mammalian taxa with no available cladistic data was surprisingly high at the species-level: only six out of 28 orders have a high coverage of taxa with available cladistic data.
 This high coverage threshold of 75\% of taxa with available cladistic data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
-Beyond this threshold, there is considerable displacement of wildcard taxa (\textit{sensu} \cite{kearneyfragmentary2002}) and decreases in clade conservation \cite{GuillermeCooper}.
-Therefore we expect a high probability of topological artefacts for the placement of fossil taxa at the species-level in most mammalian orders.
-However, data coverage seems to be less of an issue at higher taxonomic levels (i.e. genus- and family-level).
-This point is important from a practical point of view because of the slight discrepancy between the neontological and palaeontological concept of species.
-While neontological species are described using morphology, genetic distance, spatial distribution and even behaviour, palaeontological species can be based only on morphological, spatial and temporal data (e.g.
-\cite{ni2013oldest}).
-Because of this, most palaeontological studies are using the genus as their smallest OTU (e.g. \cite{ni2013oldest,O'Leary08022013}).
-Thus data availability at the genus-level in living mammals should be our primary concern when aiming to build phylogenies of living and fossil taxa.
+Beyond this threshold, there is considerable displacement of wildcard taxa (\textit{sensu} \cite{kearneyfragmentary2002}) and decreased clade conservation \cite{GuillermeCooper}.
+Therefore we expect difficulties in placement of fossil taxa at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels.
+This point is important from a practical point of view because of the slight discrepancy between neontological and palaeontological species concepts.
+While neontological species are described using morphology, genes, distribution etc.; palaeontological species can be based only on morphological, spatial and temporal data (e.g. \cite{ni2013oldest}).
+Therefore, most palaeontological studies use genus as their smallest OTU (e.g.
+\cite{ni2013oldest,O'Leary08022013}), so data availability at the genus-level in living mammals should be our primary concern when building phylogenies of living and fossil taxa.
 
-When only a few species with cladistic data are available, the ideal scenario is for them to be phylogenetically overdispersed (i.e.
-that there is data for at least every sub-clade) to maximize the possibilities of a fossil branching from the right clade.
+When few species have available cladistic data, the ideal scenario is for them to be phylogenetically overdispersed to maximize the possibilities of a fossil branching from the right clade.
 The second best scenario is that species with cladistic data are randomly distributed across the phylogeny. 
-In this scenario we expect no special bias in the placement of the fossil \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with cladistic data were randomly distributed across the phylogeny of each order.
+Here we expect no special bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with cladistic data were randomly distributed across the phylogeny.
 The worst case scenario for fossil placement is that species with cladistic data are phylogenetically clustered. 
-In this situation we expect two major biases to occur: first, the fossil will not be able to branch within a clade containing no data, and second, the fossil will have a higher probability, at random, of branching within the clade containing most of the available data.
-This means that fossils with uncertain phylogenetic affinities (\textit{incertae sedis}) will have a higher probability of branching within the most sampled clade just by chance.
-Our results suggest that this may be an issue, at the genus-level, in Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha. 
-For example, a Carnivora fossil will be unable to branch in the Herpestidae that has no species with cladistic data, and will also have more chance to branch, randomly, within the Canidae clade than any other clade in Carnivora (Figure \ref{Figure_example_coverage}B).
-Thus, in Total Evidence trees, placements of some carnivoran fossils should considered with caution. 
-% NC: Trying not to piss off GraHam - really it's weird stuff appearing in Carnivora that we are concerned by, not all Carnivora. TG: I think it's mainly Graham's work that focuses on canids.
-In this study, we treated all cladistic matrices as equal in a similar way to molecular matrices. 
-For example, if matrix A contained 100 characters for taxa X and Y, and matrix B contained 50 different characters for taxa X and Z, we assumed that both matrices can be combined in a supermatrix containing 150 independent characters for taxon X, 100 for taxon Y and 50 for taxon Z.
-Unfortunately, cladistic data cannot always be treated in this way because some characters may overlap.
-For example, if matrix A has a character coding for the shape of a particular morphological feature and matrix B has a character coding for the presence of this same morphological feature and a second character coding for its shape, then these three characters are non-independent compound characters \cite{Brazeau2011}.
-However, in reasonably sized matrices (\textgreater 100 characters; \cite{GuillermeCooper,harrisonamong-character2014}) it is more likely that a number of characters are consistently conserved among the different matrices and thus easily combinable.
-For example, within the Primate cladistic literature, the character \textit{p7} - the size of the $4^{th}$ lower premolar paraconid - has been used consistently for \textgreater 15 years (e.g.
-\cite{ross1998phylogenetic,marivaux2005anthropoi,ni2013oldest} and can therefore be combined among the matrices.
-A conservative approach to avoid compound characters would be to select only the most recent matrix for each group, but this would result in the loss of a lot of data.
+Then we expect two major biases to occur: first, fossils will not be able to branch within a clade containing no data, and second, fossils will have higher probability of branching within the most sampled clade by chance.
+Our results suggest that this may be problematic at the genus-level in Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha. 
+For example, a Carnivora fossil will be unable to branch in the Herpestidae, and will have more chance to randomly branch within Canidae (Figure \ref{Figure_example_coverage}B).
 
-Despite the absence of good cladistic data coverage for living mammals, the Total Evidence methods still seems to be the most promising way of combining living and fossil data for macroevolutionary analyses.
+Despite the absence of good cladistic data coverage for living mammals, the Total Evidence method still seems to be the most promising way of combining living and fossil data for macroevolutionary analyses.
 Following the recommendations in \cite{GuillermeCooper}, we need to code cladistic characters for as many living species possible. 
-Fortunately, data for living mammals is usually readily available in natural history collections, therefore, we propose that an increased effort be put into coding morphological characters from living species, possibly by engaging in collaborative data collection projects through web portals such as \textit{Morphobank} \cite{morphobank}.
-% NC: Still feels like it needs one final sentence to round it all off. Many something like below, but less crappy:
-Such an effort would be valuable not only to phylogeneticists, but also to any researcher focusing understanding macroevolutionary patterns and processes.%interested in the diversity of life on Earth.
+Fortunately, data for living mammals is usually readily available in natural history collections, therefore, we propose that an increased effort be put into coding morphological characters from living species, possibly by engaging in collaborative data collection projects.
+Such an effort would be valuable not only to phylogeneticists, but also to any researcher focusing understanding macroevolutionary patterns and processes.
 
@@
 We have no competing interests.
-\section{Acknowledgements}
-David Bapst, Graeme Lloyd, Nick Matzke and April Wright.
+\section{Acknowledgments}
+We thank David Bapst, Graeme Lloyd, Nick Matzke and April Wright.
 \section{Funding statement}

### 2015-06-25 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Updated supplementaries (missing reproducibility)

@@
 
-%Line numbering REMOVE FOR WORDCOUNT
-%\modulolinenumbers[1]
-%\linenumbers
+#Line numbering
+\modulolinenumbers[1]
+\linenumbers
 
@@
 \subsection{Data collection and standardisation}
-We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases %(accessed 10th of June 2015)
-: Morphobank (\url{http://www.morphobank.org/} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{https://github.com/rossmounce/cladistic-data}).
-We also performed a systematic Google Scholar search %(accessed 11th of June 2015) 
-for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
-% In total, we downloaded 286 matrices containing a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique. % Number of unique OTUs (including fossils)
+We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases: Morphobank (\url{http://www.morphobank.org/} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{https://github.com/rossmounce/cladistic-data}).
+We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Supplementary Materials Section 1 for a detailed description of the search procedure).
 In total, we downloaded 286 matrices containing 5228 unique operational taxonomic units (OTUs). 
-We used OTUs rather than species since entries in the matrices ranged from species to families, and standardised the taxonomy as described in Supplementary Materials. 
+We used OTUs rather than species since entries in the matrices ranged from species to families, and standardised the taxonomy as described in Supplementary Materials (section 1). 
 We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}.
 
-%In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardised and ranged from specific individual specimen names (i.e.
-the name of a collection item) to the family-level.
-%Where possible, we considered OTUs at their lowest valid taxonomic level (i.e. species) but some OTUs were only valid at a higher taxonomic level (e.g. genus or family).
-%Therefore for some orders, we sampled more genera than species (Table \ref{Table_morpho_taxa_proportion}).
-
-%To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H.
-sapiens} was transformed to \textit{Homo sapiens}).
-%We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\url{https://paleobiodb.org/}).
-%For OTUs that did not appear in these three sources, we first decomposed the name (i.e.
-\textit{Homo sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
-%Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; see Supplementary Material for more details).
-
-% The number of characters in each matrix ranged from 6 to 4541.
 Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having characters that overlap with those of other matrices \cite{wagner2000} and (2) they are more likely to contain a higher proportion of specific characters that are not-applicable across large clades (e.g.
@@
 To assess the availability of cladistic data for each mammalian order, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
-%We used these different taxonomic levels because some clades are well covered at the family- or genus-level, but poorly covered at the species-level. % NC: Probably obvious?
 We consider orders with \textless 25\% of living taxa with cladistic data as having low data coverage, and orders with \textgreater 75\% of living taxa with cladistic data as having high data coverage.
@@
 
-%\section{SOM}
-\newcommand{\beginsupplement}{%
-    \setcounter{table}{0}
-    \renewcommand{\thetable}{S\arabic{table}}%
-    \setcounter{figure}{0}
-    \renewcommand{\thefigure}{S\arabic{figure}}%
-}
-\beginsupplement
-\input{Supplementary/Supplementary}
-
 %END

### 2015-06-25 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Abstract edits

--- 2015-06-25 Manuscript/BiologyLetters/Missing_mammals-draft.tex
@@
 \begin{abstract}
-Combining living and fossil in the same analysis data is crucial for studying changes in global biodiversity through time.
-One method allowing to combine this data is the Total Evidence method that uses both molecular data for living species and morphological data for both living and fossil species.
-With this method, a good overlap of morphological data between living and fossil taxa is crucial for accurately inferring the phylogenies’ topology.
+\section{abstract} % NC: Just put this here as it then separates this for the word count
+Analyses of living and fossil taxa are crucial for understanding changes in biodiversity through time.
+The Total Evidence method allows living and fossil taxa to be combined in phylogenies, by using molecular data for living taxa and morphological data for both living and fossil taxa.
+With this method, substantial overlap of morphological data among living and fossil taxa is crucial for accurately inferring topology.
+However, although molecular data for living species is widely available, scientists using and generating morphological data mainly focus on fossils.
+Therefore, there is a gap in our knowledge of neontological morphological data even in well-studied groups such as mammals.
 
-Since the advent of DNA, molecular data has become easily and widely available.
-However, despite two centuries of morphological studies, scientists using and generating such data mainly focus on palaeontological data.
-Therefore, there is a gap in our knowledge of neontological morphological data even in well studied groups such as mammals.
-
-In this study, we quantify the morphological data available for living mammal taxa.
-We then analyse the structure of the available data by testing if it is clustered or evenly spread across the phylogeny.
-We found that 78\% of mammalian orders have less than 25\% data available at the species level.
-However, we found that the available is often randomly distributed among these orders apart from six of them where the data is clustered
+We investigated the amount of morphological (cladistic) data available for living mammals and how this data was phylogenetically distributed across orders.
+22 of 28 mammalian orders have \textless 25\% species with available morphological data; this has implications for the accurate placement of fossil taxa, although the issue is less pronounced at higher taxonomic levels.
+In most orders, species with available data are randomly distributed across the phylogeny, which may reduce the impact of the problem.
+We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence phylogenies. 
 \end{abstract}
 
-%\noindent Key words: Total Evidence method, phylogenetic clustering, morphological matrix, fossil, topology\\ Excluded from word count.
+%\noindent Key words: Total Evidence method, phylogenetic clustering, morphological matrix, extinct, topology\\ Excluded from word count.
+% NC: Might need some new ones as these are in the abstract now...
 

### 2015-06-25 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Updated ref

@@
 
-#Line numbering
+%Line numbering
 \modulolinenumbers[1]
@@
 22 of 28 orders have low coverage (\textless 25\% species with cladistic data) and six have high coverage (\textgreater 75\% species with cladistic data) at the species-level.
-At the genus-level, three orders have low coverage and 12 have high coverage, and at the family-level, no orders have low coverage and 23 have high coverage (Table \ref{Table_morpho_taxa_proportion}).
+At the genus-level, three orders have low coverage and 12 have high coverage, and at the family-level, no orders have low coverage and 23 have high coverage (Table\ref{Table_results}).
+
+%TG: don't clean up this caption please!
+%\caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels.
+The left vertical bar represents low coverage (\textless 25\%); the right vertical bar represents high coverage (\textgreater 75\%).
+Negative Net Relatedness Index (NRI) and Nearest Taxon Index (NTI) values indicate phylogenetic overdispersion; positive values indicate phylogenetic clustering.
+Significant NRI or NTI values are in bold.
+*p \textless 0.05; **p \textless 0.01; ***p \textless 0.001.} \\
+%TG: don't clean up this caption please!
 
@@
 
-Only six orders had significantly clustered data (Afrosoricida and Pholidota at the species-level, and Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha at both species- and genus-level) and none had significantly overdispersed data (Table~\ref{Table_data_structure}).
+Only six orders had significantly clustered data (Afrosoricida and Pholidota at the species-level, and Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha at both species- and genus-level) and none had significantly overdispersed data (Table~\ref{Table_results}).
 

### 2015-06-29 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Updated morphobank typo

@@
 \subsection{Data collection and standardisation}
-We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases: Morphobank (\url{http://www.morphobank.org/} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{https://github.com/rossmounce/cladistic-data}).
+We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{http://www.morphobank.org/} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{https://github.com/rossmounce/cladistic-data}).
 We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Supplementary Materials Section 1 for a detailed description of the search procedure).

### 2015-08-13 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Figures updates

@@
 \begin{abstract}
-\section{abstract} % NC: Just put this here as it then separates this for the word count
+%\section{abstract} % NC: Just put this here as it then separates this for the word count
 Analyses of living and fossil taxa are crucial for understanding changes in biodiversity through time.
@@
 
-%\noindent Key words: Total Evidence method, phylogenetic clustering, morphological matrix, extinct, topology\\ Excluded from word count.
+\noindent Key words: Total Evidence method, phylogenetic clustering, morphological matrix, extinct, topology\\
 % NC: Might need some new ones as these are in the abstract now...
@@
 \section{Ethics statement}
+N/A
 \section{Data accessibility statement}

### 2015-08-13 Manuscript/BiologyLetters/Supplementary/Results_1c/Results_1cTable_results.tex
supplementary table update (p values caption)

% latex table generated in R 3.2.0 by xtable 1.7-4 package
% Sat Jun 20 17:19:07 2015
\begin{longtable}{lL{1.8cm}C{2cm}lcc}
\caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels (without any character threshold; results from the 286 matrices). The coverage represents the proportion of taxa with available morphological data. The left vertical bar represents 25\% of available data (``low'' coverage if \textless 25\%); The right vertical bar represents 75\% of available data (``high'' coverage if \textgreater 75\%). When the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) are negative, taxa are more phylogenetically dispersed than expected by chance; when NRI or NTI are positive, taxa are more phylogenetically clustered by expected by chance. Significant NRI or NTI are highlighted in bold. *p \textless 0.05; **p \textless 0.01; ***p \textless 0.001.} \\ 
  \hline
Order & Taxonomic level & Proportion of taxa & Coverage & NRI & NTI \\ 
  \hline
Afrosoricida & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar1.pdf} &   &   \\ 
  Afrosoricida & genus & 17/17 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar2.pdf} &   &   \\ 
  Afrosoricida & species & 23/42 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar3.pdf} & 1.75 & 1.08 \\ 
  Carnivora & family & 14/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar4.pdf} & 0.63 & 0.6 \\ 
  \textbf{Carnivora} & \textbf{genus} & \textbf{54/125} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar5.pdf} & \textbf{4.81**} & \textbf{1.78*} \\ 
  \textbf{Carnivora} & \textbf{species} & \textbf{76/283} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar6.pdf} & \textbf{7.66**} & \textbf{0.85} \\ 
  Cetartiodactyla & family & 21/21 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar7.pdf} &   &   \\ 
  Cetartiodactyla & genus & 100/128 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar8.pdf} & 0.85 & 0.94 \\ 
  \textbf{Cetartiodactyla} & \textbf{species} & \textbf{171/310} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar9.pdf} & \textbf{1.92*} & \textbf{-0.46} \\ 
  Chiroptera & family & 15/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar10.pdf} & -0.28 & 0.56 \\ 
  \textbf{Chiroptera} & \textbf{genus} & \textbf{93/202} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar11.pdf} & \textbf{13.47**} & \textbf{1.1} \\ 
  \textbf{Chiroptera} & \textbf{species} & \textbf{215/1053} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar12.pdf} & \textbf{8.82**} & \textbf{1.22} \\ 
  Cingulata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar13.pdf} &   &   \\ 
  Cingulata & genus & 8/9 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar14.pdf} & 1.51 & -1.57 \\ 
  \textbf{Cingulata} & \textbf{species} & \textbf{9/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar15.pdf} & \textbf{1.9*} & \textbf{0.11} \\ 
  Dasyuromorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar16.pdf} &   &   \\ 
  Dasyuromorphia & genus & 8/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar17.pdf} & -0.75 & -1.07 \\ 
  Dasyuromorphia & species & 9/64 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar18.pdf} & -0.88 & -0.34 \\ 
  Dermoptera & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar19.pdf} &   &   \\ 
  Dermoptera & genus & 1/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar20.pdf} &   &   \\ 
  Dermoptera & species & 1/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar21.pdf} &   &   \\ 
  Didelphimorphia & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar22.pdf} &   &   \\ 
  Didelphimorphia & genus & 16/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar23.pdf} &   &   \\ 
  Didelphimorphia & species & 42/84 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar24.pdf} & -1.65 & 0.2 \\ 
  Diprotodontia & family & 11/11 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar25.pdf} &   &   \\ 
  Diprotodontia & genus & 25/38 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar26.pdf} & -1.13 & -1.31 \\ 
  Diprotodontia & species & 31/126 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar27.pdf} & 0.48 & -1.77 \\ 
  Erinaceomorpha & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar28.pdf} &   &   \\ 
  Erinaceomorpha & genus & 10/10 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar29.pdf} &   &   \\ 
  Erinaceomorpha & species & 21/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar30.pdf} & -1.07 & -0.2 \\ 
  Hyracoidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar31.pdf} &   &   \\ 
  Hyracoidea & genus & 1/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar32.pdf} &   &   \\ 
  Hyracoidea & species & 1/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar33.pdf} &   &   \\ 
  Lagomorpha & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar34.pdf} &   &   \\ 
  Lagomorpha & genus & 5/12 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar35.pdf} & -1.06 & -0.95 \\ 
  Lagomorpha & species & 12/86 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar36.pdf} & -0.62 & -1.88 \\ 
  Macroscelidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar37.pdf} &   &   \\ 
  Macroscelidea & genus & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar38.pdf} &   &   \\ 
  Macroscelidea & species & 12/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar39.pdf} & -1.3 & -1.06 \\ 
  Microbiotheria & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar40.pdf} &   &   \\ 
  Microbiotheria & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar41.pdf} &   &   \\ 
  Microbiotheria & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar42.pdf} &   &   \\ 
  Monotremata & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar43.pdf} &   &   \\ 
  Monotremata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar44.pdf} & -0.72 & -0.69 \\ 
  Monotremata & species & 2/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar45.pdf} & -0.97 & -0.97 \\ 
  Notoryctemorphia & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar46.pdf} &   &   \\ 
  Notoryctemorphia & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar47.pdf} &   &   \\ 
  Notoryctemorphia & species & 0/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar48.pdf} &   &   \\ 
  Paucituberculata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar49.pdf} &   &   \\ 
  Paucituberculata & genus & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar50.pdf} &   &   \\ 
  Paucituberculata & species & 5/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar51.pdf} &   &   \\ 
  Peramelemorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar52.pdf} &   &   \\ 
  Peramelemorphia & genus & 7/7 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar53.pdf} &   &   \\ 
  Peramelemorphia & species & 16/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar54.pdf} & -0.13 & 0.97 \\ 
  Perissodactyla & family & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar55.pdf} &   &   \\ 
  Perissodactyla & genus & 6/6 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar56.pdf} &   &   \\ 
  Perissodactyla & species & 10/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar57.pdf} & -0.07 & -2.63 \\ 
  Pholidota & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar58.pdf} &   &   \\ 
  Pholidota & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar59.pdf} &   &   \\ 
  Pholidota & species & 4/8 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar60.pdf} & 1.18 & 0.94 \\ 
  Pilosa & family & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar61.pdf} & 1.87 & 2 \\ 
  Pilosa & genus & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar62.pdf} & -0.96 & 0.36 \\ 
  \textbf{Pilosa} & \textbf{species} & \textbf{5/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar63.pdf} & \textbf{1.28} & \textbf{2.38*} \\ 
  Primates & family & 15/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar64.pdf} &   &   \\ 
  Primates & genus & 48/68 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar65.pdf} & -0.35 & -1.33 \\ 
  Primates & species & 64/351 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar66.pdf} & -0.67 & -1.27 \\ 
  Proboscidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar67.pdf} &   &   \\ 
  Proboscidea & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar68.pdf} &   &   \\ 
  Proboscidea & species & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar69.pdf} & -0.69 & -0.69 \\ 
  Rodentia & family & 18/32 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar70.pdf} & 0.66 & -0.98 \\ 
  Rodentia & genus & 82/450 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar71.pdf} & -1.66 & 1.55 \\ 
  \textbf{Rodentia} & \textbf{species} & \textbf{90/2094} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar72.pdf} & \textbf{2.76*} & \textbf{2.34*} \\ 
  Scandentia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar73.pdf} &   &   \\ 
  Scandentia & genus & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar74.pdf} & -0.74 & -0.74 \\ 
  Scandentia & species & 3/20 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar75.pdf} & -1.88 & -0.84 \\ 
  Sirenia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar76.pdf} &   &   \\ 
  Sirenia & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar77.pdf} &   &   \\ 
  Sirenia & species & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar78.pdf} &   &   \\ 
  Soricomorpha & family & 3/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar79.pdf} & -0.98 & -0.99 \\ 
  \textbf{Soricomorpha} & \textbf{genus} & \textbf{19/43} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar80.pdf} & \textbf{7.11**} & \textbf{2.59**} \\ 
  \textbf{Soricomorpha} & \textbf{species} & \textbf{21/392} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar81.pdf} & \textbf{10.65**} & \textbf{3.56**} \\ 
  Tubulidentata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar82.pdf} &   &   \\ 
  Tubulidentata & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar83.pdf} &   &   \\ 
  Tubulidentata & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Results_1c/Table_figures/bar84.pdf} &   &   \\ 
   \hline
\hline
\label{Table_results}
\end{longtable}

### 2015-09-15 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Minor typos picked up from the thesis

@@
 
-NTI \cite{webb2002phylogenies} is based on mean nearest neighbour distance (MNND) and is calculated as follows:
+NTI \cite{webb2002phylogenies} is based on mean nearest neighbour distance ($MNND$) and is calculated as follows:
   \begin{equation}
-    NTI=-\left(\frac{\overline{MNND}_{obs}-\overline{MNND}_{n}}{\sigma(MNND_{n})}\right)
+    \text{NTI}=-\left(\frac{\overline{MNND}_{obs}-\overline{MNND}_{n}}{\sigma(MNND_{n})}\right)
   \end{equation}
 where $\overline{MNND}_{obs}$ is the observed mean distance between each of $n$ taxa with cladistic data and its nearest neighbour with cladistic data in the phylogeny, 
-$\overline{MNND}_{n}$ is the mean of 1000 mean MNND between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random MNND values.
-NRI is calculated in the same way, but MNND is replaced by mean phylogenetic distance (MPD) as follows:
+$\overline{MNND}_{n}$ is the mean of 1000 $MNND$ between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random $MNND$ values.
+NRI is calculated in the same way, but $MNND$ is replaced by mean phylogenetic distance ($MPD$) as follows:
   \begin{equation}
-    NRI=-\left(\frac{\overline{MPD}_{obs}-\overline{MPD}_{n}}{\sigma(MPD_{n})}\right)
+    \text{NRI}=-\left(\frac{\overline{MPD}_{obs}-\overline{MPD}_{n}}{\sigma(MPD_{n})}\right)
   \end{equation}
@@
 
-When few species have available cladistic data, the ideal scenario is for them to be phylogenetically overdispersed to maximize the possibilities of a fossil branching from the right clade.
-The second best scenario is that species with cladistic data are randomly distributed across the phylogeny. 
+When few species have available cladistic data, the ideal scenario is for them to be phylogenetically overdispersed to maximize the possibilities of a fossil branching in the right clade.
+The second best scenario is that species with cladistic data are randomly distributed across the phylogenies. 
 Here we expect no special bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with cladistic data were randomly distributed across the phylogeny.

### 2015-11-18 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
separated the ESM

%Authors guidlines: http://royalsocietypublishing.org/instructions-authors
% 2500 words max (includes the title page, abstract, references, acknowledgements and figure/table legends)
% current version is around 3700. I think a big cut down can be done on the references.
% We allow a maximum of 4 displays, only 2 of which can be figures.

\documentclass[12pt,letterpaper]{article}

%Packages
\usepackage{pdflscape}
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{float}
\usepackage{latexsym}
\usepackage{url}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{array}
\usepackage[version=3]{mhchem}
\usepackage{ifthen}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{amsthm}
\usepackage{amstext}
\usepackage{enumerate}
\usepackage[osf]{mathpazo}
\usepackage{dcolumn}
\usepackage{lineno}
\usepackage{longtable}
\pagenumbering{arabic}

\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

%Pagination style and stuff % NC: Note that these are all syst biol specific.
\linespread{2}
\raggedright
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0} 
\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}
\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}
\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}
\renewcommand{\tableofcontents}{}
%\bibpunct{(}{)}{;}{a}{}{,}

%---------------------------------------------
%
%       START
%
%---------------------------------------------

\begin{document}

\newcommand{\beginsupplement}{%
    \setcounter{table}{0}
    \renewcommand{\thetable}{S\arabic{table}}%
    \setcounter{figure}{0}
    \renewcommand{\thefigure}{S\arabic{figure}}%
}

%Running head
\begin{flushright}
Version dated: \today
\end{flushright}

\bigskip
\medskip
\begin{center}

\noindent{\Large \bf Assessment of cladistic data availability for living mammals}

\bigskip
\noindent{\Large \bf Electronic Supplementary Material 1}

\bigskip
\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^*$ and Natalie Cooper$^1$$^,$$^2$}\\
\noindent {\small \it 
$^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\
$^2$Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.}\\
\medskip
\noindent{*\bf Corresponding author.} \textit{guillert@tcd.ie}\\  
\vspace{1in}

\end{center}
%\modulolinenumbers[1]
%\linenumbers

\newpage

%TG: Maybe useless, I'll add that just in the thesis.
%\section{Graphical abstract}

%\begin{figure}[!htbp]
%\centering
%    \includegraphics[width=1\textwidth]{MissingDataFigure.pdf}
%\caption{Example of topological errors due to missing morphological data in living taxa.
%The phylogeny contains two clades, Aves and Mammalia, with molecular data (grey) for both but only morphological data (red) for Aves.
%If a mammalian fossil species (with no molecular data) is added to the phylogeny, it will erroneously branch with the Aves instead of the Mammalia because no morphological data will overlap between the fossil mammals and the living ones.}
%\label{Figure_missing_data_problem}
%\end{figure}

\section{1 - Data collection}
\subsection{Public repositories}
We downloaded available matrices containing fossil and/or living mammal taxa from the three data bases using the following list of keywords:

\texttt{Mammalia; Monotremata; Marsupialia; Placentalia; Macroscelidea; Afrosoricida; Tubulidentata; Hyracoidea; Proboscidea; Sirenia; Pilosa; Cingulata; Scandentia; Dermoptera; Primates; Lagomorpha; Rodentia; Erinaceomorpha; Soricomorpha; Cetacea; Artiodactyla; Cetartiodactyla; Chiroptera; Perissodactyla; Pholidota; Carnivora; Didelphimorphia; Paucituberculata; Microbiotheria; Dasyuromorphia; Peramelemorphia; Notoryctemorphia; Diprotodontia}.

Details about each public repository specific search option is listed below.
Note that some matrices have been downloaded from more than one database but that it is not an issue since we are interested in the total number of unique living OTUs and that if some where present in more than one matrix, they still only counted as one single OTU.

\subsubsection{Morphobank}
We accessed the Morphobank repository (\texttt{http://www.morphobank.org/}) on the 10th of June 2015 and used the keywords listed above in the search menu.
We downloaded the data associated with each project matching with the keyword.

\subsubsection{Graeme Lloyd}
We accessed Graeme Lloyd's website repository (\texttt{http://graemetlloyd.com/}) on the 10th of June 2015 and downloaded all the matrices that were available with a direct download link in the mammal data section of the website (\texttt{http://graemetlloyd.com/matrmamm.html}).

\subsubsection{Ross Mounce}
We accessed Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}) on the 11th of June 2015 and downloaded every 601 matrix.
We then ran a shell script to select only the matrices that had any text element that match with one of the search terms.
% LINK WHICH SHELL SCRIPT
To make the matrix selection more thorough, we ignored the keywords case as well as the latin suffix (\textit{ia}, \textit{ata}, \textit{ea}, and \textit{a}).

\subsection{Google scholars (accessed 11th of June 2015)}
To make sure we didn't miss any extra matrix that wasn't available on one of these repository, we ran a Google Scholar search on the 5th of January. 
We downloaded the additional cladistic matrices from the 20 first search results matching with our selected keywords and with any of the 35 taxonomic levels (mammals Orders, Infraclasses and Class).
We used the following key words:

\texttt{\textit{order} ("morphology" OR "morphological" OR "cladistic") AND characters matrix paleontology phylogeny}

were \textit{order} was replaced by all the keywords listed above. For each 33 keywords, we selected the 20 first papers to match the Google search published since 2010 resulting in 660 papers.
Among these papers, not all contained relevant data (discrete morphological characters AND mammalian data).
We selected only the 20 first results per search term to avoid downloading articles that were to irrelevant. Among the 660 papers, only 50 contained a total of 425 extra living OTUs (Figure ~\ref{Supp_figure_google_searches}).
Also we decided to select only the articles published since 2010 because nearly every one of the recent published matrix contains both a fraction of morphological characters and OTUs from previous studies.
For example in primates the character \textit{p7} coded first by \cite{ross1998phylogenetic} is reused with the same living species in \cite{seiffert2003fossil}, \cite{marivaux2005anthropoid}, \cite{seiffert2005basal}, \cite{bloch2007new}, \cite{bloch2007new}, \cite{kay2008anatomy}, \cite{silcox2008biogeographic}, \cite{seiffert2009convergent}, \cite{tabuce2009anthropoid}, \cite{boyer2010astragalar}, \cite{seiffert2010fossil}, \cite{marivaux2013djebelemur} and \cite{ni2013oldest}.

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_google_searches.pdf}
\caption{Google searches additional OTUs rarefaction curve. The x axis represent the number of google scholar matches (papers, books or abstracts) and the y axis represents the cumulative number of additional living OTUs per google scholar match.}
\label{Supp_figure_google_searches}
\end{figure}

The list of all the 286 download matrices is available on % PREMALINK TO THE MATRICES LIST (TRANSFORM TO CSV!)
The matrices contained a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique.
In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardised and ranged from specific individual specimen names (i.e. the name of a collection item) to the family-level.
Where possible, we considered OTUs at their lowest valid taxonomic level (i.e. species) but some OTUs were only valid at a higher taxonomic level (e.g. genus or family).
Therefore for some orders, we sampled more genera than species.

\subsection{Standardising the matrices}
We transformed all the non-nexus matrices (tnt, word, excel, jpeg) to nexus format manually.
We then cleaned the nexus matrices by removing any extra information (trees, continuous characters, morphological characters description, molecular data) to end up with nexus matrices containing only the discrete morphological data.
We then manually fixed the wrong bionomial names format (e.g. \textit{H. sapiens}) into the correct ones (e.g. \textit{Homo sapiens}) using the abbreviation list in the concerned publications. 
All the standardised matrices are available on % PREMALINK TO THE STANDARDISED MATRICES

\subsection{Selecting the living OTUs}

To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H. sapiens} was transformed to \textit{Homo sapiens}).
We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{bininda-emondsthe2007} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\url{https://paleobiodb.org/}).
For OTUs that did not appear in these three sources, we first decomposed the name (i.e. \textit{Homo sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; Figure ~\ref{Supp_figure_Taxonomic_algorithm}).

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_Taxonomic_algorithm.pdf}
\caption{Taxonomic matching algorithm used in this study. For each matrix, each operational taxonomic units (OTU) is matched with the super tree from Fritz 2009. If the OTU matches, then it is classified as living. Else it is matched with the Wilson \& Reeders 2005 taxonomy list. If the OTU matches, then it is classified as living. Else it is matched with the Paleo Database list of mammals. If the OTU matches, then it is classified as fossil. Else it is ignored.}
\label{Supp_figure_Taxonomic_algorithm}
\end{figure}

\section{2 - Data reproducibility}
Every step of the analysis (apart from downloading and standardisation of the matrices) is entirely repeatable via GitHub (\url{https://github.com/TGuillerme/Missing_living_mammals}).
%Needs to fix this in GitHub
%1-Rarefaction ?
%2-Selecting the OTUs
%3-Analysing the availability and the data structure

\bibliographystyle{vancouver}
\bibliography{Supp_References}

\end{document}

### 2015-11-18 Manuscript/BiologyLetters/Supplementary/ESM_2.tex
separated the ESM

--- 2015-11-18 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
@@
 \bigskip
-\noindent{\Large \bf Electronic Supplementary Material 1}
+\noindent{\Large \bf Electronic Supplementary Material 2}
 
@@
 
-\section{1 - Data collection}
-\subsection{Public repositories}
-We downloaded available matrices containing fossil and/or living mammal taxa from the three data bases using the following list of keywords:
+\section{Supplementary results - 1}
+The following section contains supplementary results to the main body: the available data structure using the NTI and the PD metric; the proportion of available data and the data structure for all the matrices (including the matrices with less than 100 characters).
 
-\texttt{Mammalia; Monotremata; Marsupialia; Placentalia; Macroscelidea; Afrosoricida; Tubulidentata; Hyracoidea; Proboscidea; Sirenia; Pilosa; Cingulata; Scandentia; Dermoptera; Primates; Lagomorpha; Rodentia; Erinaceomorpha; Soricomorpha; Cetacea; Artiodactyla; Cetartiodactyla; Chiroptera; Perissodactyla; Pholidota; Carnivora; Didelphimorphia; Paucituberculata; Microbiotheria; Dasyuromorphia; Peramelemorphia; Notoryctemorphia; Diprotodontia}.
-
-Details about each public repository specific search option is listed below.
-Note that some matrices have been downloaded from more than one database but that it is not an issue since we are interested in the total number of unique living OTUs and that if some where present in more than one matrix, they still only counted as one single OTU.
-
-\subsubsection{Morphobank}
-We accessed the Morphobank repository (\texttt{http://www.morphobank.org/}) on the 10th of June 2015 and used the keywords listed above in the search menu.
-We downloaded the data associated with each project matching with the keyword.
-
-\subsubsection{Graeme Lloyd}
-We accessed Graeme Lloyd's website repository (\texttt{http://graemetlloyd.com/}) on the 10th of June 2015 and downloaded all the matrices that were available with a direct download link in the mammal data section of the website (\texttt{http://graemetlloyd.com/matrmamm.html}).
-
-\subsubsection{Ross Mounce}
-We accessed Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}) on the 11th of June 2015 and downloaded every 601 matrix.
-We then ran a shell script to select only the matrices that had any text element that match with one of the search terms.
-% LINK WHICH SHELL SCRIPT
-To make the matrix selection more thorough, we ignored the keywords case as well as the latin suffix (\textit{ia}, \textit{ata}, \textit{ea}, and \textit{a}).
-
-\subsection{Google scholars (accessed 11th of June 2015)}
-To make sure we didn't miss any extra matrix that wasn't available on one of these repository, we ran a Google Scholar search on the 5th of January. 
-We downloaded the additional cladistic matrices from the 20 first search results matching with our selected keywords and with any of the 35 taxonomic levels (mammals Orders, Infraclasses and Class).
-We used the following key words:
-
-\texttt{\textit{order} ("morphology" OR "morphological" OR "cladistic") AND characters matrix paleontology phylogeny}
-
-were \textit{order} was replaced by all the keywords listed above. For each 33 keywords, we selected the 20 first papers to match the Google search published since 2010 resulting in 660 papers.
-Among these papers, not all contained relevant data (discrete morphological characters AND mammalian data).
-We selected only the 20 first results per search term to avoid downloading articles that were to irrelevant.
-Among the 660 papers, only 50 contained a total of 425 extra living OTUs (Figure ~\ref{Supp_figure_google_searches}).
-Also we decided to select only the articles published since 2010 because nearly every one of the recent published matrix contains both a fraction of morphological characters and OTUs from previous studies.
-For example in primates the character \textit{p7} coded first by \cite{ross1998phylogenetic} is reused with the same living species in \cite{seiffert2003fossil}, \cite{marivaux2005anthropoid}, \cite{seiffert2005basal}, \cite{bloch2007new}, \cite{bloch2007new}, \cite{kay2008anatomy}, \cite{silcox2008biogeographic}, \cite{seiffert2009convergent}, \cite{tabuce2009anthropoid}, \cite{boyer2010astragalar}, \cite{seiffert2010fossil}, \cite{marivaux2013djebelemur} and \cite{ni2013oldest}.
-
-\begin{figure}[!htbp]
-\centering
-    \includegraphics[width=1\textwidth]{Supp_figure_google_searches.pdf}
-\caption{Google searches additional OTUs rarefaction curve.
-The x axis represent the number of google scholar matches (papers, books or abstracts) and the y axis represents the cumulative number of additional living OTUs per google scholar match.}
-\label{Supp_figure_google_searches}
-\end{figure}
-
-The list of all the 286 download matrices is available on % PREMALINK TO THE MATRICES LIST (TRANSFORM TO CSV!)
-The matrices contained a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique.
-In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardised and ranged from specific individual specimen names (i.e.
-the name of a collection item) to the family-level.
-Where possible, we considered OTUs at their lowest valid taxonomic level (i.e. species) but some OTUs were only valid at a higher taxonomic level (e.g. genus or family).
-Therefore for some orders, we sampled more genera than species.
-
-\subsection{Standardising the matrices}
-We transformed all the non-nexus matrices (tnt, word, excel, jpeg) to nexus format manually.
-We then cleaned the nexus matrices by removing any extra information (trees, continuous characters, morphological characters description, molecular data) to end up with nexus matrices containing only the discrete morphological data.
-We then manually fixed the wrong bionomial names format (e.g. \textit{H. sapiens}) into the correct ones (e.g. \textit{Homo sapiens}) using the abbreviation list in the concerned publications. 
-All the standardised matrices are available on % PREMALINK TO THE STANDARDISED MATRICES
-
-\subsection{Selecting the living OTUs}
-
-To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H.
-sapiens} was transformed to \textit{Homo sapiens}).
-We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{bininda-emondsthe2007} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\url{https://paleobiodb.org/}).
-For OTUs that did not appear in these three sources, we first decomposed the name (i.e.
-\textit{Homo sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
-Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; Figure ~\ref{Supp_figure_Taxonomic_algorithm}).
-
-\begin{figure}[!htbp]
-\centering
-    \includegraphics[width=1\textwidth]{Supp_figure_Taxonomic_algorithm.pdf}
-\caption{Taxonomic matching algorithm used in this study.
-For each matrix, each operational taxonomic units (OTU) is matched with the super tree from Fritz 2009.
-If the OTU matches, then it is classified as living.
-Else it is matched with the Wilson \& Reeders 2005 taxonomy list.
-If the OTU matches, then it is classified as living.
-Else it is matched with the Paleo Database list of mammals.
-If the OTU matches, then it is classified as fossil.
-Else it is ignored.}
-\label{Supp_figure_Taxonomic_algorithm}
-\end{figure}
-
-\section{2 - Data reproducibility}
-Every step of the analysis (apart from downloading and standardisation of the matrices) is entirely repeatable via GitHub (\url{https://github.com/TGuillerme/Missing_living_mammals}).
-%Needs to fix this in GitHub
-%1-Rarefaction ?
-%2-Selecting the OTUs
-%3-Analysing the availability and the data structure
-
-\bibliographystyle{vancouver}
-\bibliography{Supp_References}
+\input{Results_1c/Results_1cTable_results}
 

### 2015-11-18 Manuscript/BiologyLetters/Supplementary/ESM_3.tex
separated the ESM

--- 2015-11-18 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
@@
 \bigskip
-\noindent{\Large \bf Electronic Supplementary Material 1}
+\noindent{\Large \bf Electronic Supplementary Material 3}
 
@@
 
-\section{1 - Data collection}
-\subsection{Public repositories}
-We downloaded available matrices containing fossil and/or living mammal taxa from the three data bases using the following list of keywords:
-
-\texttt{Mammalia; Monotremata; Marsupialia; Placentalia; Macroscelidea; Afrosoricida; Tubulidentata; Hyracoidea; Proboscidea; Sirenia; Pilosa; Cingulata; Scandentia; Dermoptera; Primates; Lagomorpha; Rodentia; Erinaceomorpha; Soricomorpha; Cetacea; Artiodactyla; Cetartiodactyla; Chiroptera; Perissodactyla; Pholidota; Carnivora; Didelphimorphia; Paucituberculata; Microbiotheria; Dasyuromorphia; Peramelemorphia; Notoryctemorphia; Diprotodontia}.
-
-Details about each public repository specific search option is listed below.
-Note that some matrices have been downloaded from more than one database but that it is not an issue since we are interested in the total number of unique living OTUs and that if some where present in more than one matrix, they still only counted as one single OTU.
-
-\subsubsection{Morphobank}
-We accessed the Morphobank repository (\texttt{http://www.morphobank.org/}) on the 10th of June 2015 and used the keywords listed above in the search menu.
-We downloaded the data associated with each project matching with the keyword.
-
-\subsubsection{Graeme Lloyd}
-We accessed Graeme Lloyd's website repository (\texttt{http://graemetlloyd.com/}) on the 10th of June 2015 and downloaded all the matrices that were available with a direct download link in the mammal data section of the website (\texttt{http://graemetlloyd.com/matrmamm.html}).
-
-\subsubsection{Ross Mounce}
-We accessed Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}) on the 11th of June 2015 and downloaded every 601 matrix.
-We then ran a shell script to select only the matrices that had any text element that match with one of the search terms.
-% LINK WHICH SHELL SCRIPT
-To make the matrix selection more thorough, we ignored the keywords case as well as the latin suffix (\textit{ia}, \textit{ata}, \textit{ea}, and \textit{a}).
-
-\subsection{Google scholars (accessed 11th of June 2015)}
-To make sure we didn't miss any extra matrix that wasn't available on one of these repository, we ran a Google Scholar search on the 5th of January. 
-We downloaded the additional cladistic matrices from the 20 first search results matching with our selected keywords and with any of the 35 taxonomic levels (mammals Orders, Infraclasses and Class).
-We used the following key words:
-
-\texttt{\textit{order} ("morphology" OR "morphological" OR "cladistic") AND characters matrix paleontology phylogeny}
-
-were \textit{order} was replaced by all the keywords listed above. For each 33 keywords, we selected the 20 first papers to match the Google search published since 2010 resulting in 660 papers.
-Among these papers, not all contained relevant data (discrete morphological characters AND mammalian data).
-We selected only the 20 first results per search term to avoid downloading articles that were to irrelevant.
-Among the 660 papers, only 50 contained a total of 425 extra living OTUs (Figure ~\ref{Supp_figure_google_searches}).
-Also we decided to select only the articles published since 2010 because nearly every one of the recent published matrix contains both a fraction of morphological characters and OTUs from previous studies.
-For example in primates the character \textit{p7} coded first by \cite{ross1998phylogenetic} is reused with the same living species in \cite{seiffert2003fossil}, \cite{marivaux2005anthropoid}, \cite{seiffert2005basal}, \cite{bloch2007new}, \cite{bloch2007new}, \cite{kay2008anatomy}, \cite{silcox2008biogeographic}, \cite{seiffert2009convergent}, \cite{tabuce2009anthropoid}, \cite{boyer2010astragalar}, \cite{seiffert2010fossil}, \cite{marivaux2013djebelemur} and \cite{ni2013oldest}.
+\section{Supplementary results - 2}
+The following section contains supplementary results to the main body: the phylogenetic representation of the data availability per order (excluding Primates and Carnivora, present in the main body).
 
@@
 \centering
-    \includegraphics[width=1\textwidth]{Supp_figure_google_searches.pdf}
-\caption{Google searches additional OTUs rarefaction curve.
-The x axis represent the number of google scholar matches (papers, books or abstracts) and the y axis represents the cumulative number of additional living OTUs per google scholar match.}
-\label{Supp_figure_google_searches}
+    \includegraphics[width=1\textwidth]{Supp_figure_AFROSORICIDA.pdf}
+\caption{Distribution of available morphological data across Afrosoricida. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Afrosoricida}
 \end{figure}
-
-The list of all the 286 download matrices is available on % PREMALINK TO THE MATRICES LIST (TRANSFORM TO CSV!)
-The matrices contained a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique.
-In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardised and ranged from specific individual specimen names (i.e.
-the name of a collection item) to the family-level.
-Where possible, we considered OTUs at their lowest valid taxonomic level (i.e. species) but some OTUs were only valid at a higher taxonomic level (e.g. genus or family).
-Therefore for some orders, we sampled more genera than species.
-
-\subsection{Standardising the matrices}
-We transformed all the non-nexus matrices (tnt, word, excel, jpeg) to nexus format manually.
-We then cleaned the nexus matrices by removing any extra information (trees, continuous characters, morphological characters description, molecular data) to end up with nexus matrices containing only the discrete morphological data.
-We then manually fixed the wrong bionomial names format (e.g. \textit{H. sapiens}) into the correct ones (e.g. \textit{Homo sapiens}) using the abbreviation list in the concerned publications. 
-All the standardised matrices are available on % PREMALINK TO THE STANDARDISED MATRICES
-
-\subsection{Selecting the living OTUs}
-
-To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H.
-sapiens} was transformed to \textit{Homo sapiens}).
-We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{bininda-emondsthe2007} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\url{https://paleobiodb.org/}).
-For OTUs that did not appear in these three sources, we first decomposed the name (i.e.
-\textit{Homo sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
-Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; Figure ~\ref{Supp_figure_Taxonomic_algorithm}).
 
@@
 \centering
-    \includegraphics[width=1\textwidth]{Supp_figure_Taxonomic_algorithm.pdf}
-\caption{Taxonomic matching algorithm used in this study.
-For each matrix, each operational taxonomic units (OTU) is matched with the super tree from Fritz 2009.
-If the OTU matches, then it is classified as living.
-Else it is matched with the Wilson \& Reeders 2005 taxonomy list.
-If the OTU matches, then it is classified as living.
-Else it is matched with the Paleo Database list of mammals.
-If the OTU matches, then it is classified as fossil.
-Else it is ignored.}
-\label{Supp_figure_Taxonomic_algorithm}
+    \includegraphics[width=1\textwidth]{Supp_figure_CHIROPTERA.pdf}
+\caption{Distribution of available morphological data across Chiroptera. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Chiroptera}
 \end{figure}
 
-\section{2 - Data reproducibility}
-Every step of the analysis (apart from downloading and standardisation of the matrices) is entirely repeatable via GitHub (\url{https://github.com/TGuillerme/Missing_living_mammals}).
-%Needs to fix this in GitHub
-%1-Rarefaction ?
-%2-Selecting the OTUs
-%3-Analysing the availability and the data structure
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_CINGULATA.pdf}
+\caption{Distribution of available morphological data across Cingulata. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Cingulata}
+\end{figure}
 
-\bibliographystyle{vancouver}
-\bibliography{Supp_References}
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_DASYUROMORPHIA.pdf}
+\caption{Distribution of available morphological data across Dasyuromorphia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Dasyuromorphia}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_DIDELPHIMORPHIA.pdf}
+\caption{Distribution of available morphological data across Didelphimorphia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Didelphimorphia}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_DIPROTODONTIA.pdf}
+\caption{Distribution of available morphological data across Diprotodontia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Diprotodontia}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_ERINACEOMORPHA.pdf}
+\caption{Distribution of available morphological data across Erinaceomorpha. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Erinaceomorpha}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_PILOSA.pdf}
+\caption{Distribution of available morphological data across Pilosa. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Pilosa}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_CETARTIODACTYLA.pdf}
+\caption{Distribution of available morphological data across Cetartiodactyla. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Primates}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_RODENTIA.pdf}
+\caption{Distribution of available morphological data across Rodentia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Rodentia}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_SCANDENTIA.pdf}
+\caption{Distribution of available morphological data across Scandentia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Scandentia}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_SORICOMORPHA.pdf}
+\caption{Distribution of available morphological data across Soricomorpha. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Soricomorpha}
+\end{figure}
 

### 2015-11-18 Analysis/Presence_absence.R

@@ -8,7 +8,7 @@ load.functions(test=FALSE) #Set test=FALSE to speed up the loading
 
 #Example data input
 #Read Fritz tree
-one_tree<-read.nexus("../Data/Trees/FritzTree.rs200k.1tree.tre")
+one_tree<-read.nexus("../Data/Trees/mammalST_MSW05_best_chrono.tre")
 #Taxonomic reference
 WR_list<-read.csv("../Data/Taxon_References/WilsonReederMSW.csv", header=T, stringsAsFactors=F)
 

### 2015-11-23 Manuscript/BiologyLetters/CoverLetter.tex
Cover letter INIT

\documentclass[11pt]{letter}
\usepackage[a4paper,left=2.5cm, right=2.5cm, top=1cm, bottom=1cm]{geometry}
\usepackage[osf]{mathpazo}
\usepackage{url}
\signature{Thomas Guillerme \\ Natalie Cooper}
\address{School of Natural Sciences \\ Trinity College Dublin \\ Dublin 2, Ireland \\ guillert@tcd.ie}
\longindentation=0pt
\begin{document}

% PLoS Biology cover letter:
% Upload a cover letter as a separate file in the online system. The length limit is 600 words.

% The cover letter should address the following questions:

%     What is the scientific question you are addressing?
% "investigate whether mammal diversity was directly affected by the K-Pg mass extinction."
%     What is the key finding that answers this question?
% "We find no evidence for a direct effect of the K-Pg extinction event on mammalian diversification."
%     What is the nature of the evidence you provide in support of your conclusion?
% The time slicing method and using both data?
%     What are the three most recently published articles that are relevant to this question?
% Slater 2013 Methods Ecol. Evol.; Beck \& Lee 2014 Proc. Roy. Soc. B Close et al. 2015 Current Biol. (Or maybe put O'Leary?)
%     What significance do your results have for the field?
% Proposing new methods to investigate disparity-through-time in a continuous way + using both living and fossil taxa
%     What significance do your results have for the broader community (of biologists and/or the public)?
% Text book example.
%     What other novel findings do you present?
% Effect of using both data sources
%     Is there additional information that we should take into account?
% Graeme's story?

% TG: Ok, nothing too complex here, just some reformatting maybe. Everything is clearly stated in the Nature Comm version but not in their order. Maybe do a bullet point alternative version?

\begin{letter}{}
\opening{Dear Editors,}

Combining fossils into phylogenies is cool
We can use TEM method along side with tip-dating
However, to eficiently do it, it is important to have enough data for living species for allowing the fossil ones to branch correctly (Guillerme and Cooper 2016, Molecular Phylogenetics and Evolution)

In this research article, entitled ``Investigating the effects of the Cretaceous-Paleogene mass extinction on mammalian morphological diversity using a new methodological approach'', we use both data from living and fossil species through Total Evidence tip-dated phylogenies along with a novel continuous time-slicing method to investigate whether mammalian morphological diversity was directly affected by the K-Pg event.

We look forward to hearing from you,

\closing{Yours sincerely,}

\end{letter}
\end{document}

% Editors
% Dean Adams?
% Matt Friedman?
% Daniel Rabosky?
% Liam Revell?

% Suggested reviewers:
% Graeme Lloyd
% Robin Beck
% Robert Close
% Mike Benton or Steve Brusatte or Richard Butler (they're cited a lot)
% Any other ideas?

% Yep all good.
% BUT if you want to suggest Graeme you need to add the postscript I added above. You may need to alter the reference.

### 2015-11-24 Manuscript/BiologyLetters/CoverLetter.tex
list of reviewers
Cover letter update

@@
 
-% PLoS Biology cover letter:
-% Upload a cover letter as a separate file in the online system. The length limit is 600 words.
-
-% The cover letter should address the following questions:
-
-%     What is the scientific question you are addressing?
-% "investigate whether mammal diversity was directly affected by the K-Pg mass extinction."
-%     What is the key finding that answers this question?
-% "We find no evidence for a direct effect of the K-Pg extinction event on mammalian diversification."
-%     What is the nature of the evidence you provide in support of your conclusion?
-% The time slicing method and using both data?
-%     What are the three most recently published articles that are relevant to this question?
-% Slater 2013 Methods Ecol. Evol.; Beck \& Lee 2014 Proc. Roy. Soc. B Close et al. 2015 Current Biol. (Or maybe put O'Leary?)
-%     What significance do your results have for the field?
-% Proposing new methods to investigate disparity-through-time in a continuous way + using both living and fossil taxa
-%     What significance do your results have for the broader community (of biologists and/or the public)?
-% Text book example.
-%     What other novel findings do you present?
-% Effect of using both data sources
-%     Is there additional information that we should take into account?
-% Graeme's story?
-
-% TG: Ok, nothing too complex here, just some reformatting maybe. Everything is clearly stated in the Nature Comm version but not in their order. Maybe do a bullet point alternative version?
-
 \begin{letter}{}
@@
 
-Combining fossils into phylogenies is cool
-We can use TEM method along side with tip-dating
-However, to eficiently do it, it is important to have enough data for living species for allowing the fossil ones to branch correctly (Guillerme and Cooper 2016, Molecular Phylogenetics and Evolution)
+%TG: That's word for word the TEM paper cover letter
+In recent years there has been growing interest in building phylogenies that contain both living and fossil taxa (e.g. Quental and Marshall 2006 TREE; Fritz et al. 2013 TREE; Heath et al. 2014 PNAS).
+Such phylogenies could revolutionize the way we think about macroevolutionary patterns and processes, and provide a more complete understanding of trends in biodiversity through time.
 
-In this research article, entitled ``Investigating the effects of the Cretaceous-Paleogene mass extinction on mammalian morphological diversity using a new methodological approach'', we use both data from living and fossil species through Total Evidence tip-dated phylogenies along with a novel continuous time-slicing method to investigate whether mammalian morphological diversity was directly affected by the K-Pg event.
+One method, the Total Evidence method, allows us to use molecular and morphological data to build phylogenies with both living and fossil species as tips (Ronquist et al. 2012 Syst Biol).
+%TG: it changes from here:
+This method is extremely promising because it allows us to use all the available data.
+However, to efficiently apply this method, it is crucial to have enough morphological data for living species for allowing the fossil ones to branch correctly into the phylogeny (Guillerme and Cooper 2016, Molecular Phylogenetics and Evolution).
 
-We look forward to hearing from you,
+Our research article, entitled ``Assessment of cladistic data availability for living mammals'', is to our knowledge, the first to thoroughly assess the availability of cladistic (morphological) data in living mammals.
+Additionally, we test whether the available data in each mammalian order is randomly distributed or clustered towards certain clades to investigate whether the placement of fossil taxa within these order would have respectively non \textit{a priori} bias or would be bias towards the clades with the most data.
+
+Our results show that, unfortunately 22 of 28 mammalian orders have less than 25\% species with available morphological data.
+Fortunately, however, most of this available data was randomly distributed across each mammalian order expect for 6 for which the data was biased towards certain clades.
+We hope that our work will encourage cladisticians %TG: or is that how there are called?
+to sample more data for living species and thus serving as a stepping stone for improving combining living and fossil taxa in the same phylogenies.
+
+We look forward to hearing from you soon,
 
@@
 % Editors
-% Dean Adams?
-% Matt Friedman?
-% Daniel Rabosky?
-% Liam Revell?
+% Paul Barrett? Or do they have one assigned for the issue already
 
 % Suggested reviewers:
-% Graeme Lloyd
-% Robin Beck
-% Robert Close
-% Mike Benton or Steve Brusatte or Richard Butler (they're cited a lot)
-% Any other ideas?
-
-% Yep all good.
-% BUT if you want to suggest Graeme you need to add the postscript I added above. You may need to alter the reference.
+% Fred Ronquist
+% Alex Pyron (since he already reviewed the other one)
+% Phil Donoghue (but conflict of interest?)
+% O'Leary (even if you know I hate the Science paper, we mention morphoBank in a positive light)

### 2015-11-24 Manuscript/BiologyLetters/CoverLetter.tex
Cover letter update

@@
 Our research article, entitled ``Assessment of cladistic data availability for living mammals'', is to our knowledge, the first to thoroughly assess the availability of cladistic (morphological) data in living mammals.
-Additionally, we test whether the available data in each mammalian order is randomly distributed or clustered towards certain clades to investigate whether the placement of fossil taxa within these order would have respectively non \textit{a priori} bias or would be bias towards the clades with the most data.
+Additionally, we test whether the available data in each mammalian order is randomly distributed or clustered towards certain clades to investigate whether the placement of fossil taxa within these order would have respectively no \textit{a priori} bias or would be biased towards the clades with the most data.
 
-Our results show that, unfortunately 22 of 28 mammalian orders have less than 25\% species with available morphological data.
+Our results show that 22 of 28 mammalian orders have less than 25\% species with available morphological data.
 Fortunately, however, most of this available data was randomly distributed across each mammalian order expect for 6 for which the data was biased towards certain clades.

### 2015-11-24 Manuscript/BiologyLetters/CoverLetter.tex
Edits to simplify cover letter

@@
 
+Please find attached our research article, ``Assessment of cladistic data availability for living mammals'', for consideration in the Biology Letters Special Issue: % NC: What is the title?
+
+% NC: For special issues you need to make this clear upfront so it doesn't go through the normal channels!
+
 %TG: That's word for word the TEM paper cover letter
@@
 
-One method, the Total Evidence method, allows us to use molecular and morphological data to build phylogenies with both living and fossil species as tips (Ronquist et al. 2012 Syst Biol).
+The Total Evidence method allows us to use molecular and morphological data to build phylogenies with both living and fossil species as tips (Ronquist et al. 2012 Syst Biol).
 %TG: it changes from here:
 This method is extremely promising because it allows us to use all the available data. 
-However, to efficiently apply this method, it is crucial to have enough morphological data for living species for allowing the fossil ones to branch correctly into the phylogeny (Guillerme and Cooper 2016, Molecular Phylogenetics and Evolution).
+However, to efficiently apply this method, it is crucial to have enough morphological data for living species to allow the fossils to branch correctly in the phylogeny (Guillerme and Cooper 2016, Molecular Phylogenetics and Evolution).
 
-Our research article, entitled ``Assessment of cladistic data availability for living mammals'', is to our knowledge, the first to thoroughly assess the availability of cladistic (morphological) data in living mammals.
-Additionally, we test whether the available data in each mammalian order is randomly distributed or clustered towards certain clades to investigate whether the placement of fossil taxa within these order would have respectively no \textit{a priori} bias or would be biased towards the clades with the most data.
+Our research article assesses the availability of coded cladistic (morphological) data in living mammals within existing databases and data compilations.
+Additionally, we test whether the available data in each mammalian order is randomly distributed or clustered towards certain clades to investigate whether the placement of fossil taxa within these orders would be biased towards the clades with the most data.
 
-Our results show that 22 of 28 mammalian orders have less than 25\% species with available morphological data.
-Fortunately, however, most of this available data was randomly distributed across each mammalian order expect for 6 for which the data was biased towards certain clades.
-We hope that our work will encourage cladisticians %TG: or is that how there are called?
-to sample more data for living species and thus serving as a stepping stone for improving combining living and fossil taxa in the same phylogenies.
+Our results show that 22 of 28 mammalian orders have \textless 25\% species with available cladistic data.
+Fortunately, however, most of this available data was randomly distributed across each mammalian order except for six orders where the data was biased towards certain clades.
+We hope that our work will encourage researchers to collect more data for living species and thus help to improve attempts to combine living and fossil taxa in the same phylogenies.
 
@@
 % Paul Barrett? Or do they have one assigned for the issue already
+% NC: Put Paul down if it asks you to. Usually these are edited by the organisers so April, Nick and Graeme.
 
@@
 % O'Leary (even if you know I hate the Science paper, we mention morphoBank in a positive light)
+
+% NC: They seem ok, I think you can put Phil.
+He isn't working with you on anything and he is not your friend as such.
+He has nothing to gain from you getting this published (I'd have no issues reviewing for Kevin Arbuckle).
+If you want others think of museum people doing morphological phylogenies, maybe in the US?
+Not sure about O'Leary because she is a pain apparently.

### 2015-11-26 Manuscript/BiologyLetters/CoverLetter.tex
small comment on cover letter

@@
 
-Please find attached our research article, ``Assessment of cladistic data availability for living mammals'', for consideration in the Biology Letters Special Issue: % NC: What is the title?
+Please find attached our research article, ``Assessment of cladistic data availability for living mammals'', for consideration in the Biology Letters Special Issue: ``Putting fossils into trees''.% NC: What is the title?
+%TG: TO CHECK
 

### 2015-11-30 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
Final minor typos edits

@@
 \subsubsection{Ross Mounce}
-We accessed Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}) on the 11th of June 2015 and downloaded every 601 matrix.
+We accessed Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}) on the 11th of June 2015 and downloaded all 601 matrix.
 We then ran a shell script to select only the matrices that had any text element that match with one of the search terms.

### 2015-12-01 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Final minor typos edits
docx conversion

@@
 To perform such analyses it is necessary to combine living and fossil taxa in phylogenetic trees.
-One increasingly popular method, the Total Evidence method \cite{eernissetaxonomic1993,ronquista2012}, combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix (e.g.
+One increasingly popular method, the Total Evidence method \cite{ronquista2012}, combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix (e.g.
 \cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a phylogeny with living and fossil taxa at the tips.
@@
 We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{http://www.morphobank.org/} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{https://github.com/rossmounce/cladistic-data}).
-We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Supplementary Materials Section 1 for a detailed description of the search procedure).
+We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material (ESM) for a detailed description of the search procedure).
 In total, we downloaded 286 matrices containing 5228 unique operational taxonomic units (OTUs). 
-We used OTUs rather than species since entries in the matrices ranged from species to families, and standardised the taxonomy as described in Supplementary Materials (section 1). 
+We used OTUs rather than species since entries in the matrices ranged from species to families, and standardised the taxonomy as described in the ESM. 
 We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}.
 
-Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having characters that overlap with those of other matrices \cite{wagner2000} and (2) they are more likely to contain a higher proportion of specific characters that are not-applicable across large clades (e.g.
-``antler ramifications'' is a character that is only applicable to Cervidae not all mammals \cite{Brazeau2011}).
+Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having overlapping characters with other matrices \cite{wagner2000} and (2) they are more likely to contain specific characters that are not-applicable across large clades (e.g.
+``antler ramifications'' is a character that is only applicable to Cervidae \cite{Brazeau2011}).
 Therefore we selected only matrices containing \textgreater 100 characters for each OTU.
 This threshold was chosen to correspond with the number of characters used in \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
-Results of analyses with no threshold are available in Supplementary Material. 
-After removing matrices with \textless 100 characters, we retained 1074 unique living mammal OTUs from 126 matrices. % 1601 unique living OTUs for 286 matrices (no threshold)
+Results of analyses with no threshold are available in the ESM. 
+After removing matrices with \textless 100 characters, we retained 1074 unique living OTUs from 126 matrices. % 1601 unique living OTUs for 286 matrices (no threshold)
 
@@
 We investigated whether the available cladistic data for each order was (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index (NTI; \cite{webb2002phylogenies}) and the Net Relatedness Index (NRI; \cite{webb2002phylogenies}).
-NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to clustering or overdispersion across the whole phylogeny \cite{Cooper2008}. 
+NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to it across the whole phylogeny \cite{Cooper2008}. 
 Both metrics were calculated using the \texttt{picante} package in R \cite{picante,R}.
@@
 $\overline{MNND}_{n}$ is the mean of 1000 $MNND$ between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random $MNND$ values.
-NRI is calculated in the same way, but $MNND$ is replaced by mean phylogenetic distance ($MPD$) as follows:
+NRI is calculated in the same way, but using the mean phylogenetic distance ($MPD$):
   \begin{equation}
@@
 Our results show that although phylogenetic relationships among living mammals are well-resolved (e.g.
-\cite{BinindaEmonds,meredithimpacts2011}) , most of the data used to build these phylogenies is molecular, and very little cladistic data is available for living mammals compared to fossil mammals (e.g.
+\cite{BinindaEmonds,meredithimpacts2011}), most of the data used to build these phylogenies is molecular, and few cladistic data are available for living mammals compared to fossils (e.g.
 \cite{O'Leary08022013,ni2013oldest}).
@@
 This high coverage threshold of 75\% of taxa with available cladistic data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
-Beyond this threshold, there is considerable displacement of wildcard taxa (\textit{sensu} \cite{kearneyfragmentary2002}) and decreased clade conservation \cite{GuillermeCooper}.
-Therefore we expect difficulties in placement of fossil taxa at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels.
+Beyond this threshold, there is considerable displacement of wildcard taxa and decreased clade conservation \cite{GuillermeCooper}.
+Therefore we expect difficulties in placement of fossils at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels.
 This point is important from a practical point of view because of the slight discrepancy between neontological and palaeontological species concepts.
 While neontological species are described using morphology, genes, distribution etc.; palaeontological species can be based only on morphological, spatial and temporal data (e.g. \cite{ni2013oldest}).
-Therefore, most palaeontological studies use genus as their smallest OTU (e.g.
+Therefore, most palaeontological studies use the genus as their smallest OTU (e.g.
 \cite{ni2013oldest,O'Leary08022013}), so data availability at the genus-level in living mammals should be our primary concern when building phylogenies of living and fossil taxa.
@@
 \section{Data accessibility statement}
-All data and analysis code is available on GitHub (\url{https://github.com/TGuillerme/Missing_living_mammals}).
+All data and code are available on GitHub (\url{https://github.com/TGuillerme/Missing_living_mammals}).
 \section{Authors' Contributions}
-T.G.
-and N.C conceived and designed the experiments.
-T.G.
-performed the experiments and analysed the data.
-T.G.
-and N.C.
-contributed to the writing of the manuscript.
-All authors approved the final version of the manuscript.
+T.G. and N.C designed the study. T.G. analysed the data. T.G. and N.C. wrote the the manuscript.
 \section{Competing Interests}
@@
 
+\section{References}
+
 \bibliographystyle{vancouver}

### 2015-12-01 Manuscript/BiologyLetters/pandoc.sh

@@ -1,0 +1,1 @@
+pandoc -o Missing_mammals-draft.docx Missing_mammals-draft.tex --bibliography References.bib --filter pandoc-citeproc --csl apa.csl

### 2016-01-21 Manuscript/BiologyLetters/response.tex
Reviewers response INIT

\documentclass[12pt,letterpaper]{article}

%Packages
\usepackage{pdflscape}
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{float}
\usepackage{latexsym}
\usepackage{url}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{array}
\usepackage[version=3]{mhchem}
\usepackage{ifthen}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{amsthm}
\usepackage{amstext}
\usepackage{enumerate}
\usepackage[osf]{mathpazo}
\usepackage{dcolumn}
\usepackage{lineno}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\pagenumbering{arabic}

%Pagination style and stuff
%\linespread{2}

\raggedright
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0} 
\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}
\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}
\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}
\renewcommand{\tableofcontents}{}

\setlength\parindent{0pt}

\begin{document}

\textbf{RE: MPE-15-254}\\
\bigskip
Dear Editor,\\ % put editor's name
\bigskip
We are very grateful to both referees for their helpful and constructive comments, which we believe have helped us to significantly improve our paper. We have taken all of their comments on board, and respond to their points below. For improving clarity in this document, we sorted the reviewers comments as follows:
\begin{enumerate}
\item \textbf{Reviewer 1:} containing reviewer 1's major comments
\item \textbf{Reviewer 1 specific comments:} containing reviewer 1's specific comments
\item \textbf{Reviewer 2:} containing reviewer 2's comments
\end{enumerate}
Also, we uploaded two versions of the revised manuscript. Both have the exact same content but one has all the changes and additions to the main text highlighted in yellow to help with reviewing.

\section{Reviewer 1:}

\begin{enumerate}
% -----------------
% Comment 1
% -----------------
\item{\textcolor{blue}{Coment 1}}

\end{enumerate}

We hope we have responded to all these comments appropriately. Please let us know if you require any further information,\\
\bigskip

Thomas Guillerme (guillert@tcd.ie)\\
Natalie Cooper

\end{document}

### 2016-01-21 Manuscript/BiologyLetters/response.tex
Reviewers comments

@@
 \bigskip
-Dear Editor,\\ % put editor's name
+Dear Surayya Johar,\\
 \bigskip
-We are very grateful to both referees for their helpful and constructive comments, which we believe have helped us to significantly improve our paper.
+We are very grateful to the four referees for their helpful and constructive comments, which we believe have helped us to significantly improve our paper.
 We have taken all of their comments on board, and respond to their points below.
-For improving clarity in this document, we sorted the reviewers comments as follows:
-\begin{enumerate}
-\item \textbf{Reviewer 1:} containing reviewer 1's major comments
-\item \textbf{Reviewer 1 specific comments:} containing reviewer 1's specific comments
-\item \textbf{Reviewer 2:} containing reviewer 2's comments
-\end{enumerate}
-Also, we uploaded two versions of the revised manuscript. Both have the exact same content but one has all the changes and additions to the main text highlighted in yellow to help with reviewing.
+For improving clarity in this document, we dealt with each comment in the order they appeared in your ``Decision on Manuscript'' email and we copied the reviewers comments in blue.
+
+%%%%%%%%%%%%%%%
+% Reviewer 1
+%%%%%%%%%%%%%%%
 
 \section{Reviewer 1:}
+\begin{enumerate}
+\item{\textcolor{blue}{In ESM 1, p.
+4, the authors explain their methods of collecting morphological data from the literature.
+Why select only the first 10 results per search term?
+How do you know that relevant papers wouldn’t come up later in the search?
+I need evidence to support this cutoff; as it stands, I’m concerned that you’re missing some relevant material.}}
 
+\item{\textcolor{blue}{Similarly, I find the 2010 date cutoff troubling, because I’m not sure that the argument that these recent studies contain a fraction of the previously published characters and OTUs allows you to make a comprehensive statement about the availability of character data.
+What about the fraction that isn’t included?
+This is a question that needs to be better addressed before you will convince readers that this is really all the data that is out there.}}
+\end{enumerate}
+
+\subsection{Reviewer 1 specific comments:}
 \begin{enumerate}
-% -----------------
-% Comment 1
-% -----------------
-\item{\textcolor{blue}{Coment 1}}
+\item{\textcolor{blue}{P.2, line 43: ``chunks'' is not an appropriate term to use in a scientific paper. Use a more formal term.}}
+\item{\textcolor{blue}{P.2, L53 \& L55: ``branch'' is the wrong verb. You want to say something like ``fossils will not be placed accurately within the phylogeny'' }}
+\item{\textcolor{blue}{P3, L78: ``since'' should be ``because.'' Since should be used for time, not causation.}}
+\item{\textcolor{blue}{P3, L85: ``not-applicable'' should not be hyphenated.}}
+\item{\textcolor{blue}{P3, L120: How much of a difference does the 25\% or less coverage make to the capacity of Total Evidence phylogenetic analyses to resolve relationships?
+Does your simulation study have a way of quantifying or describing the probability of getting the ``right'' answer with this little data?
+It would also help if you could show in your data table what the clades were that had better than 75\% coverage.
+It’s not immediately evident from reading Table 1.}}
+\item{\textcolor{blue}{P5, L137: ``taxa'' should be ``OTUs.''}}
+\item{\textcolor{blue}{P5, L165 \& 166: This sentence also misuses ``branch'' as on P.2. Reword for correct verb usage.}}
+\item{\textcolor{blue}{ESM 1, p. 2: ``If some where present'' should be ``if some were present.''}}
+\item{\textcolor{blue}{ESM 1, p. 3: ``one of these repository'' should be ``one of these repositories.''}}
+\item{\textcolor{blue}{ESM 1, p. 4: ``that were to irrelevant'' should be ``that were too irrelevant.''}}
+\item{\textcolor{blue}{ESM 1, p. 4: ``of the recent published matrix'' should be ``of the recently published matrices.''}}
+\item{\textcolor{blue}{ESM 1, p. 4: First sentence of the paragraph in the middle of the page is a sentence fragment, and needs to state where the matrices are available.}}
+\item{\textcolor{blue}{ESM 1, p. 4: ``since the entries in the…'' should be ``because the entries in the…'' Since refers to time, not causation.}}
+\item{\textcolor{blue}{ESM 1, p.
+5: axis title should be ``Google Scholar matches,'' not ``Google Scholars matches.'' Figure caption should capitalize ``Google Scholar.'' Y-axis and X-axis should be hyphenated.}}
+\item{\textcolor{blue}{ESM 1, p.
+6: ``wrong binomial names format'' should be ``incorrectly-formatted binomial names,'' and ``into the correct ones'' should be deleted.
+Fixing is assumed to turn incorrect things into correct ones.
+Also, the sentence starting the paragraph under ``Selecting the living OTUs'' repeats the exact same thing as this sentence.}}
+\item{\textcolor{blue}{ESM 2 seems to be almost completely redundant with Table 1. Is there some more efficient way of dealing with this information?}}
+\end{enumerate}
 
+%%%%%%%%%%%%%%%
+% Reviewer 2 (Graham Slater)
+%%%%%%%%%%%%%%%
+
+\section{Reviewer 2:}
+\begin{enumerate}
+\item{\textcolor{blue}{The authors state on page 3 lines 82-90 that they avoided matrices with ``few characters'' due to lower probability of overlapping characters with other matrices and so only selected matrices containing >100 characters.
+[...] Is 100 really the minimum number you want to go with at all phylogenetic scales?}}
+\item{\textcolor{blue}{Going through each clade in the supplement seems beyond the scope of reviewing duties, but I am concerned that if this one matrix can be missed others may have been missed too.
+I think a reexamination of the literature may be warranted using taxon specific search terms.}}
+\item{\textcolor{blue}{One interesting aside that the authors didn’t consider, despite the data being there, is how phylogenetic ``effort'' relates to diversity of the order under study.}}
+
+We are really grateful for Graham Slater's %or do we need to keep it formal and say just "reviewer 2's"
+comment and added this interesting analysis in the ESM @@@% get number.
+% that mammalian morphological phylogeneticists have focused their energies on resolving the higher level relationships among mammalian lineages.
+\end{enumerate}
+
+%%%%%%%%%%%%%%%
+% Reviewer 3 (Peter Wagner)
+%%%%%%%%%%%%%%%
+
+\section{Reviewer 3:}
+\begin{enumerate}
+\item{\textcolor{blue}{My biggest criticism is a semantic one: Morphological data $\neq$ cladistic data.}}
+\item{\textcolor{blue}{The real issue here is the availability of phylogenetic character coding for anatomical data.
+[...] Thus, the title really should be ``Assessment of available anatomical characters for phylogenetic analysis among living mammals'' or something like that.}}
+\item{\textcolor{blue}{what we really need is to focus on the tooth and skull characters that typify the fossil record.
+[...] Still, several papers estimate that the sampling of mammal species (based on cranial characters include teeth) actually is pretty darned good [...].
+Thus, we might be able to richly populate FBD and tip-dating analyses with a LOT of extinct species.}}
+\item{\textcolor{blue}{Finally, I would add: who the heck submits a paper for a special volume 6 weeks early?!?!? You make the rest of us look bad!}}
+This paper was actually ready to submit from the 28th of July (\url{http://biorxiv.org/content/early/2015/07/28/022970}) which partially explain the swift submission.
+We do sincerely apologize for making the rest of the authors submitting to the issue look bad.
+
+\end{enumerate}
+
+\subsection{Reviewer 3 specific comments:}
+\begin{enumerate}
+\item{\textcolor{blue}{Lines 25-26: ``We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence phylogenies.''
+In a way, this is more than ``total evidence.=''.
+The initial total evidence paradigm was simply to use every character you could (molecular, morphological, even behavioral).
+We now are talking about tip-dating and fossilized birth-death (FBD) approaches (e.g., Heath et al. 2014 PNAS 111:E2957) that want to use all of the available data.
+After all, one can do a total evidence analysis with just node-dating techniques rather than with tip-dating, and one can do it without worrying about branch length priors.
+However: if we want to use all of the data to calibrate branch lengths for all taxa (fossil and extant), then we need ``total evidence'' to do tip-dating and FBD properly).
+As tip-dating is the main theme of this issue, that should be linked to the total evidence explicitly here and elsewhere in the manuscript.}}
+
+\item{\textcolor{blue}{Lines 70-71: ``We downloaded all cladistic matrices containing any living and/or fossil mammal  taxa from three major public databases:''
+As a side note and for future reference for the authors, a large number of phylogenetic data sets for mammals (and other metazoans) also are available at the Paleobiology Database.
+(These were formerly at my website at the Field Museum, but that site obviously disappeared when I moved to the Smithsonian.)
+Because of the nature of the projects for which I was using these datasets, most of them are species-level, but some are genus-level, too.}}
+
+\item{\textcolor{blue}{Lines 164-166: ``For example, a Carnivora fossil will be unable to branch in the Herpestidae, and will have more chance to randomly branch within Canidae (Figure 1B).'' 
+This seems to be an awkward way to phrase this.
+Moreover, this seems like an extension of the long-branch problem.
+Basically, the issue really is that we do not have coded skeletal characters for extant herpestids.
+Thus, it is not possible for coded skeletal characters from fossil herpestids to connect those fossil herpestids to the extant herpestids.
+This is the opposite of Felsenstein’s long branch: insofar as skeletal characters are concerned, extant herpestids have zero-length branches.
+That in turn means that the fossil herpestids are going to be ``attracted'' to non-zero length branches elsewhere, with long branches in related groups such as canids being particularly good candidates because there is a higher chance of homoplasies between canids and herpesitds on the longer branches.
+So, it will not be random: they probably will latch themselves onto the longest reasonably branch among the closest relatives for which anatomical data are present.}}
+
+\end{enumerate}
+
+%%%%%%%%%%%%%%%
+% Reviewer 4 (Graeme Lloyd)
+%%%%%%%%%%%%%%%
+
+\section{Reviewer 4:}
+\begin{enumerate}
+\item{\textcolor{blue}{One point I would add of my own is that the orders used are obviously very variable in size (species number), so what about total coverage?
+This could surely be synthesised from the tables and would be a nice simple way to compare the with and without threshold results.
+Similarly, I think Reviewer 3’s [We assume Graeme Lloyd is refering to Graham Slater's reviewe (reviewer 2)] %or same as comment before: no names!
+ graphs are a nice potential addition to the ESM.}}
+\end{enumerate}
+
+\subsection{Reviewer 4 specific comments:}
+\begin{enumerate}
+\item{\textcolor{blue}{Lines 72-74 - Web addresses should be consistent in dropping of ``http://www.''.}}
+\item{\textcolor{blue}{Line 104 - ``NTI 15'' should probably be ``NTI [15]''.}}
+\item{\textcolor{blue}{Line 104 - It’s not completely clear to me how ``distance'' is measured. Is this branch-lengths on a time-scaled tree? Or just a node count?}}
+\item{\textcolor{blue}{Line 117 - It’s not clear to me how supraspecific values were calculated if the reference phylogeny (ref 10) is species-level.
+Are you using the node subtending the higher taxon or something else?}}
+\item{\textcolor{blue}{Line 149 - ``most palaeontological studies use the genus as their smallest OTU'' I’m not sure if this is strictly true.
+It’s more that shorter OTU names are commonly used.
+After all at least one actual specimen has to be coded and in my experience if you read the paper usually only a single species is used.
+Palaeontologists are just bad at using proper binomials in their matrices (which is an issue, but a different one).
+In addition ref 20 is actually all species-level OTUs!
+(I didn’t check ref 21.)}}
+\item{\textcolor{blue}{Line 154 - ``phylogenetically overdispersed'' Purely semantic, but I guess what you mean is that even sampling would be ideal, and this is *measured* using phylogenetic over-dispersion.}}
+\item{\textcolor{blue}{Line 245 - The caption could be more informative regarding the random and clustered distributions that these figures represent.}}
+\item{\textcolor{blue}{ESM 1 - Incomplete sentence: ``The list of all the 286 download matrices is available on''}}
+\item{\textcolor{blue}{ESM 1 - Incomplete sentence: ``All the standardised matrices are available on''}}
+\item{\textcolor{blue}{ESM 1 - ``designated as ``fossil'' all OTUs that were present in the Paleobiology database'' The database contains extant taxa too, although these should be marked as such.
+I guess this doesn’t matter given the flow chart in Figure 2, but is something the authors should be aware of.
+(Obviously more generally living species can have fossil records, or at least plenty of fossils are assigned to extant species.)}}
+\item{\textcolor{blue}{ESM 1 Figure 2 - How many OTUs were ignored?
+And based on the (assumed) cursory inspection of the authors what are they?
+Higher taxa?
+Specimens?
+Or just species you can’t reconcile with your three data sets?}}
+\item{\textcolor{blue}{ESM 2 - ``The following section contains supplementary results to the main body: the available data structure using the NTI and the PD metric'' Abbreviations should be introduced when first used, not later as they are here.}}
 \end{enumerate}
@@
 
-Thomas Guillerme (guillert@tcd.ie)\\
+% Editorial office comments to authors:
+% Please ensure that you include; 
+% *An author contributions section which fulfils all *four* of our criteria given here https://royalsociety.org/journals/ethics-policies/openness/, including a sentence stating all authors agree to be held accountable for the content therein and approve the final version of the manuscript
+% *A funding statement which is the same as that given in ScholarOne
+% *Please also confirm in your cover letter whether all the figures are your own or whether permission has been obtained for their use
+% *Please upload your original figure files as eps, tiff or jpeg files, rather than PDFs
+% *If you have any images that can be used to promote your article on social media (should this be accepted) please upload them as a supplementary file
+% *Please confirm whether your manuscript should be available via open access, as we note that this may be a requirement of your funder.
+
+Thomas Guillerme (guillert@tcd.ie)\\ % Changing email + adding imperial address?
 Natalie Cooper

### 2016-01-21 Manuscript/BiologyLetters/response.tex
response to minor comments reviewer 1

@@
 
-\textbf{RE: MPE-15-254}\\
+\textbf{RE: Decision on Manuscript ID RSBL-2015-1003}\\
 \bigskip
@@
 \item{\textcolor{blue}{P.2, line 43: ``chunks'' is not an appropriate term to use in a scientific paper. Use a more formal term.}}
-\item{\textcolor{blue}{P.2, L53 \& L55: ``branch'' is the wrong verb. You want to say something like ``fossils will not be placed accurately within the phylogeny'' }}
+We replaced ``chunks'' by ``parts'' on p.@, line@.
+
+\item{\textcolor{blue}{P.2, L53 \& L55: ``branch'' is the wrong verb. You want to say something like ``fossils will not be placed accurately within the phylogeny''.}}
+We replaced ``fossils cannot branch in the correct clade'' by ``fossils will not be placed accurately within the correct clade'' on p.@, line@.
+
 \item{\textcolor{blue}{P3, L78: ``since'' should be ``because.'' Since should be used for time, not causation.}}
+We replaced ``since'' by ``because'' on p.@, line@.
+
 \item{\textcolor{blue}{P3, L85: ``not-applicable'' should not be hyphenated.}}
+We removed the hyphen on p.@, line@.
+
 \item{\textcolor{blue}{P3, L120: How much of a difference does the 25\% or less coverage make to the capacity of Total Evidence phylogenetic analyses to resolve relationships?
@@
 It’s not immediately evident from reading Table 1.}}
+
+% TO DO.
+To the question, hard to give a value for the probability of getting the ``right'' tree but 25% missing living leads to 0.5 wild card taxa displacement and between 0.8 and 0.4 clade conservation on a linear scale from 1 to 0 with one being the right tree and 0 being a random tree.
+
 \item{\textcolor{blue}{P5, L137: ``taxa'' should be ``OTUs.''}}
+We replaced ``taxa'' by ``OTUs'' on p.@, line@.
+
 \item{\textcolor{blue}{P5, L165 \& 166: This sentence also misuses ``branch'' as on P.2. Reword for correct verb usage.}}
+We replaced ``unable to branch in the Herpestidae'' by ``unable to be placed in the Herpestidae clade'' on p.@, line@.
+
 \item{\textcolor{blue}{ESM 1, p. 2: ``If some where present'' should be ``if some were present.''}}
+We fixed this typo.
+
 \item{\textcolor{blue}{ESM 1, p. 3: ``one of these repository'' should be ``one of these repositories.''}}
+We fixed this typo.
+
 \item{\textcolor{blue}{ESM 1, p. 4: ``that were to irrelevant'' should be ``that were too irrelevant.''}}
+We fixed this typo.
+
 \item{\textcolor{blue}{ESM 1, p. 4: ``of the recent published matrix'' should be ``of the recently published matrices.''}}
+We fixed this typo.
+
 \item{\textcolor{blue}{ESM 1, p. 4: First sentence of the paragraph in the middle of the page is a sentence fragment, and needs to state where the matrices are available.}}
+We finished the sentence by adding the URL: ``The list of all the 286 download matrices is available on \url{https://github.com/TGuillerme/Missing_living_mammals}.''
+
 \item{\textcolor{blue}{ESM 1, p. 4: ``since the entries in the…'' should be ``because the entries in the…'' Since refers to time, not causation.}}
+We fixed this typo.
+
 \item{\textcolor{blue}{ESM 1, p.
 5: axis title should be ``Google Scholar matches,'' not ``Google Scholars matches.'' Figure caption should capitalize ``Google Scholar.'' Y-axis and X-axis should be hyphenated.}}
-\item{\textcolor{blue}{ESM 1, p.
-6: ``wrong binomial names format'' should be ``incorrectly-formatted binomial names,'' and ``into the correct ones'' should be deleted.
+We fixed the typo in the figure and the caption as well as in the code for reproducing this figure.
+
+\item{\textcolor{blue}{ESM 1, p. 6: ``wrong binomial names format'' should be ``incorrectly-formatted binomial names,'' and ``into the correct ones'' should be deleted.
 Fixing is assumed to turn incorrect things into correct ones.
 Also, the sentence starting the paragraph under ``Selecting the living OTUs'' repeats the exact same thing as this sentence.}}
+We modified the sentence to: ``[we] fixed the incorrectly-formatted bionomial names format (e.g. \textit{H. sapiens} became \textit{Homo sapiens})''
+
 \item{\textcolor{blue}{ESM 2 seems to be almost completely redundant with Table 1. Is there some more efficient way of dealing with this information?}}
+
+%TO DO: not sure but maybe we'll just put the 1 character threshold thingy.
+
 \end{enumerate}
@@
 \begin{enumerate}
+\item{\textcolor{blue}{Reviewer 1’s most substantial complaints (the year and page number cut offs for google scholar searches) seem at least partially valid.
+However, in my experience older studies tend to be smaller (due to computing limitations) and to commonly reoccur anyway due to data set reuse.
+The authors also already state that subsequent search results tended to yield no substantial additional data, and by the reviewer’s own admission the assembled data seem to look correct.
+Thus a recollecting of the data doesn’t seem necessary (but see further comments below).}}
+
+\item{\textcolor{blue}{Reviewer 2 (Graham Slater)’s complaint are - in my view - the most substantial.
+I agree that the >100 characters cut off is not as strongly justified as it might be.
+However, as the authors have run their protocol without this threshold this is easily fixed by at least briefly mentioning how this affects their results in the main text (not relegating it to the ESM where no actual discussion is provided).
+From my reading of the two tables only 11 of the orders seem to have below 25\% coverage at the species-level (not 22 as the authors state in the main text!) when the threshold is used, and this improves slightly to 10 for when the threshold is removed.
+As this difference isn’t dramatic it seems worth mentioning.
+Reviewer 2 also notes that some important studies may be missed.
+This seems valid (I am aware of the biases in Ross Mounce’s data collection protocol, and obviously my own) and I am sure there are many more data sets that could be sampled.
+In particular I am aware of a number of studies that can only be found in books and hence are not easily accessed even if they occur in a google scholar search.
+I do not expect the authors to more thoroughly search the literature for all possible cladistic analyses of mammals, however, as this is a major task.
+(I should know, it has been a long term project of mine for years.)
+Nevertheless, it does deserve mention in the discussion as a potential bias in the results.
+I don’t think this undermines the study, but rather should be viewed as both a reason for optimism and an acknowledgement of a separate problem (i.e., access to cladistic data) which multiple publications already cover (e.g., see Ross Mounce’s work).}}
+
+\item{\textcolor{blue}{Reviewer 3 (Pete Wagner)’s major problem - with which I agree - is semantic, and thus easily fixable with a few explicit caveats (e.g., ``by cladistic we mean...'').}}
+
 \item{\textcolor{blue}{One point I would add of my own is that the orders used are obviously very variable in size (species number), so what about total coverage?

### 2016-01-21 Manuscript/BiologyLetters/Missing_mammals-draft.tex
response to minor comments reviewer 1

@@
 A downside of this method is that it requires molecular data for living taxa and morphological data for both living and fossil taxa.
-Chunks of this data can be difficult, or impossible, to collect for every taxon in the analysis.
+Parts%rev 1.2.1 (as in reviewer 1, part 2 (minor), comment 1)
+ of this data can be difficult, or impossible, to collect for every taxon in the analysis.
 For example, fossils rarely have molecular data and incomplete fossil preservation may restrict the amount of morphological data available.
@@
 Simulations show that the ability of the Total Evidence method to recover the correct topology decreases when there is little overlap between morphological data in living and fossil taxa, and that the effect of missing data on topology is greatest when living taxa have few morphological data \cite{GuillermeCooper}.
-This is because (1) fossils cannot branch in the correct clade if it contains no morphological data for living taxa; and (2) fossils have a higher probability of branching within clades with more morphological data for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.
+This is because (1) fossils will not be placed accurately within the correct clade% rev 1.2.2
+ if it contains no morphological data for living taxa; and (2) fossils have a higher probability of branching within clades with more morphological data for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.
 
@@
 In total, we downloaded 286 matrices containing 5228 unique operational taxonomic units (OTUs). 
-We used OTUs rather than species since entries in the matrices ranged from species to families, and standardised the taxonomy as described in the ESM. 
+We used OTUs rather than species because% rev 1.2.3
+ entries in the matrices ranged from species to families, and standardised the taxonomy as described in the ESM. 
 We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}.
 
-Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having overlapping characters with other matrices \cite{wagner2000} and (2) they are more likely to contain specific characters that are not-applicable across large clades (e.g.
-``antler ramifications'' is a character that is only applicable to Cervidae \cite{Brazeau2011}).
+Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having overlapping characters with other matrices \cite{wagner2000} and (2) they are more likely to contain specific characters that are not applicable% rev 1.2.4
+ across large clades (e.g. ``antler ramifications'' is a character that is only applicable to Cervidae \cite{Brazeau2011}).
 Therefore we selected only matrices containing \textgreater 100 characters for each OTU.
@@
 
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{example_coverage.pdf}
+\caption{Dummy figure.}
+\label{Figure_example_coverage}
+\end{figure}
+
 %TG: don't clean up this caption please!
@@
 
-The number of living mammalian taxa with no available cladistic data was surprisingly high at the species-level: only six out of 28 orders have a high coverage of taxa with available cladistic data.
+The number of living mammalian OTUs% rev 1.2.6
+ with no available cladistic data was surprisingly high at the species-level: only six out of 28 orders have a high coverage of taxa with available cladistic data.
 This high coverage threshold of 75\% of taxa with available cladistic data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
@@
 Our results suggest that this may be problematic at the genus-level in Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha. 
-For example, a Carnivora fossil will be unable to branch in the Herpestidae, and will have more chance to randomly branch within Canidae (Figure \ref{Figure_example_coverage}B).
+For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade% rev 1.2.7
+, and will have more chance to randomly branch within Canidae (Figure \ref{Figure_example_coverage}B).
 
@@
 We thank David Bapst, Graeme Lloyd, Nick Matzke and April Wright.
+Thanks to one anonymous reviewer, Graham Slater, Peter Wagner and Graeme Lloyd for their useful comments.
+%That's two times Graeme though...
 \section{Funding statement}

### 2016-01-21 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
response to minor comments reviewer 1

@@
 Details about each public repository specific search option is listed below.
-Note that some matrices have been downloaded from more than one database but that it is not an issue since we are interested in the total number of unique living OTUs and that if some where present in more than one matrix, they still only counted as one single OTU.
+Note that some matrices have been downloaded from more than one database but that it is not an issue since we are interested in the total number of unique living OTUs and that if some were present in more than one matrix, they still only counted as one single OTU.
 
@@
 \subsection{Google scholars (accessed 11th of June 2015)}
-To make sure we didn't miss any extra matrix that wasn't available on one of these repository, we ran a Google Scholar search on the 5th of January. 
+To make sure we didn't miss any extra matrix that wasn't available on one of these repositories, we ran a Google Scholar search on the 5th of January. 
 We downloaded the additional cladistic matrices from the 20 first search results matching with our selected keywords and with any of the 35 taxonomic levels (mammals Orders, Infraclasses and Class).
@@
 Among these papers, not all contained relevant data (discrete morphological characters AND mammalian data).
-We selected only the 20 first results per search term to avoid downloading articles that were to irrelevant.
+We selected only the 20 first results per search term to avoid downloading articles that were too irrelevant.
 Among the 660 papers, only 50 contained a total of 425 extra living OTUs (Figure ~\ref{Supp_figure_google_searches}).
@@
 \caption{Google searches additional OTUs rarefaction curve.
-The x axis represent the number of google scholar matches (papers, books or abstracts) and the y axis represents the cumulative number of additional living OTUs per google scholar match.}
+The x-axis represent the number of Google Scholar matches (papers, books or abstracts) and the y-axis represents the cumulative number of additional living OTUs per google scholar match.}
 \label{Supp_figure_google_searches}
@@
 
-The list of all the 286 download matrices is available on % PREMALINK TO THE MATRICES LIST (TRANSFORM TO CSV!)
+The list of all the 286 download matrices is available on \url{https://github.com/TGuillerme/Missing_living_mammals}.
 The matrices contained a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique.
-In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardised and ranged from specific individual specimen names (i.e.
+In this study, we refer to OTUs rather than species because the entries in the downloaded matrices were not standardised and ranged from specific individual specimen names (i.e.
 the name of a collection item) to the family-level.
@@
 We then cleaned the nexus matrices by removing any extra information (trees, continuous characters, morphological characters description, molecular data) to end up with nexus matrices containing only the discrete morphological data.
-We then manually fixed the wrong bionomial names format (e.g. \textit{H. sapiens}) into the correct ones (e.g. \textit{Homo sapiens}) using the abbreviation list in the concerned publications. 
-All the standardised matrices are available on % PREMALINK TO THE STANDARDISED MATRICES
+We then manually fixed the incorrectly-formatted bionomial names format (e.g. \textit{H. sapiens} became \textit{Homo sapiens}) using the abbreviation list in the concerned publications. 
+All the standardised matrices are available on \url{https://github.com/TGuillerme/Missing_living_mammals}.
 

### 2016-01-21 Analysis/Rarefaction.R

@@ -86,6 +86,6 @@ for (mat in 1:nrow(sub_results)) {
 Additional_data<-sort(as.numeric(Google_search$Counts), decreasing=TRUE)
 pdf("../Manuscript/Supplementary/Supp_figure_google_searches.pdf")
 op<-par(bty="l")
-plot(cumsum(Additional_data), type="l", ylab="Number of additional OTUs", xlab="Google Scholars papers")
+plot(cumsum(Additional_data), type="l", ylab="Number of additional OTUs", xlab="Google Scholar papers")
 par(op)
 dev.off()

### 2016-01-22 Manuscript/BiologyLetters/response.tex
Minor comments (typos) fixed

@@
 \item{\textcolor{blue}{ESM 1, p. 4: First sentence of the paragraph in the middle of the page is a sentence fragment, and needs to state where the matrices are available.}}
-We finished the sentence by adding the URL: ``The list of all the 286 download matrices is available on \url{https://github.com/TGuillerme/Missing_living_mammals}.''
+We added the two missing link in the ESM1: ``The list of all the 286 download matrices is available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices}.'' and ``All the standardised matrices are available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices_binomial/Matrices}.''
 
@@
 \item{\textcolor{blue}{Lines 72-74 - Web addresses should be consistent in dropping of ``http://www.''.}}
+We removed the ``http://www.'' part of the URLs on p.@, lines@ and in the ESM1.
+
 \item{\textcolor{blue}{Line 104 - ``NTI 15'' should probably be ``NTI [15]''.}}
+%TO DO IN THE WORD MANUSCRIPT! We fixed this typo.
+
 \item{\textcolor{blue}{Line 104 - It’s not completely clear to me how ``distance'' is measured. Is this branch-lengths on a time-scaled tree? Or just a node count?}}
+We changed ``$\overline{MNND}_{obs}$ is the observed mean distance between each of $n$ taxa'' to ``$\overline{MNND}_{obs}$ is the observed mean branch length between each of $n$ taxa'' and ``$\overline{MPD}_{obs}$ is the observed mean phylogenetic distance of the tree containing only the $n$ taxa'' to ``$\overline{MPD}_{obs}$ is the observed mean phylogenetic branch length of the tree containing only the $n$ taxa'' on p.@, lines@.
+
 \item{\textcolor{blue}{Line 117 - It’s not clear to me how supraspecific values were calculated if the reference phylogeny (ref 10) is species-level.
 Are you using the node subtending the higher taxon or something else?}}
+%TO CHECK!
+
 \item{\textcolor{blue}{Line 149 - ``most palaeontological studies use the genus as their smallest OTU'' I’m not sure if this is strictly true.
@@
 (I didn’t check ref 21.)}}
+%To CHECK, do we actually care?
+
 \item{\textcolor{blue}{Line 154 - ``phylogenetically overdispersed'' Purely semantic, but I guess what you mean is that even sampling would be ideal, and this is *measured* using phylogenetic over-dispersion.}}
+We changed ``the ideal scenario is for them to be phylogenetically overdispersed'' to ``the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion)'' on p.@, line@.
+
 \item{\textcolor{blue}{Line 245 - The caption could be more informative regarding the random and clustered distributions that these figures represent.}}
+We changed the caption to: ``Phylogenetic distribution of species with available cladistic data across two orders (A: Primates; B: Carnivora).
+Blue branches indicate available cladistic data for the species.
+Cladistic data is randomly distributed in Primates (A) and clustered in Carnivora (B).''.
+
 \item{\textcolor{blue}{ESM 1 - Incomplete sentence: ``The list of all the 286 download matrices is available on''}}
+See reviewer 1 specific comment 12.
+
 \item{\textcolor{blue}{ESM 1 - Incomplete sentence: ``All the standardised matrices are available on''}}
+See reviewer 1 specific comment 12.
+
 \item{\textcolor{blue}{ESM 1 - ``designated as ``fossil'' all OTUs that were present in the Paleobiology database'' The database contains extant taxa too, although these should be marked as such.
@@
 (Obviously more generally living species can have fossil records, or at least plenty of fossils are assigned to extant species.)}}
+%TO DO: change ``fossil'' to ``non living OTU''? As in OTU that is not present in the Wilson Reeder list or the Bininda Emonds tree?
+
 \item{\textcolor{blue}{ESM 1 Figure 2 - How many OTUs were ignored?
@@
 Or just species you can’t reconcile with your three data sets?}}
+%TO DO: add the number of ignored OTUs. I think they're mainly specimen IDs without species names.
+
 \item{\textcolor{blue}{ESM 2 - ``The following section contains supplementary results to the main body: the available data structure using the NTI and the PD metric'' Abbreviations should be introduced when first used, not later as they are here.}}
+We changed ``using the NTI and the PD metric'' to ``using the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) metrics''.
 \end{enumerate}

### 2016-01-22 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Minor comments (typos) fixed

@@
 \subsection{Data collection and standardisation}
-We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{http://www.morphobank.org/} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{https://github.com/rossmounce/cladistic-data}).
+We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{morphobank.org} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}).
 We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material (ESM) for a detailed description of the search procedure).
@@
   \end{equation}
-where $\overline{MNND}_{obs}$ is the observed mean distance between each of $n$ taxa with cladistic data and its nearest neighbour with cladistic data in the phylogeny, 
+where $\overline{MNND}_{obs}$ is the observed mean branch length between each of $n$ taxa with cladistic data and its nearest neighbour with cladistic data in the phylogeny, 
 $\overline{MNND}_{n}$ is the mean of 1000 $MNND$ between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random $MNND$ values.
@@
   \end{equation}
-where $\overline{MPD}_{obs}$ is the observed mean phylogenetic distance of the tree containing only the $n$ taxa with cladistic data.
+where $\overline{MPD}_{obs}$ is the observed mean phylogenetic branch length of the tree containing only the $n$ taxa with cladistic data.
 Negative NTI and NRI values show that the focal taxa are more overdispersed across the phylogeny than expected by chance, and positive values reflect clustering.
@@
 
-When few species have available cladistic data, the ideal scenario is for them to be phylogenetically overdispersed to maximize the possibilities of a fossil branching in the right clade.
+When few species have available cladistic data, the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion) to maximize the possibilities of a fossil branching in the right clade.
 The second best scenario is that species with cladistic data are randomly distributed across the phylogenies. 

### 2016-01-22 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
Minor comments (typos) fixed

@@
 \subsubsection{Morphobank}
-We accessed the Morphobank repository (\texttt{http://www.morphobank.org/}) on the 10th of June 2015 and used the keywords listed above in the search menu.
+We accessed the Morphobank repository (\url{morphobank.org}) on the 10th of June 2015 and used the keywords listed above in the search menu.
 We downloaded the data associated with each project matching with the keyword.
@@
 \subsubsection{Graeme Lloyd}
-We accessed Graeme Lloyd's website repository (\texttt{http://graemetlloyd.com/}) on the 10th of June 2015 and downloaded all the matrices that were available with a direct download link in the mammal data section of the website (\texttt{http://graemetlloyd.com/matrmamm.html}).
+We accessed Graeme Lloyd's website repository (\url{graemetlloyd.com/matrmamm.html}) on the 10th of June 2015 and downloaded all the matrices that were available with a direct download link in the mammal data section of the website (\texttt{http://graemetlloyd.com/matrmamm.html}).
 
 \subsubsection{Ross Mounce}
-We accessed Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}) on the 11th of June 2015 and downloaded all 601 matrix.
+We accessed Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}) on the 11th of June 2015 and downloaded all 601 matrix.
 We then ran a shell script to select only the matrices that had any text element that match with one of the search terms.
@@
 
-The list of all the 286 download matrices is available on \url{https://github.com/TGuillerme/Missing_living_mammals}.
+The list of all the 286 download matrices is available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices}.
 The matrices contained a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique.
@@
 We then manually fixed the incorrectly-formatted bionomial names format (e.g. \textit{H. sapiens} became \textit{Homo sapiens}) using the abbreviation list in the concerned publications. 
-All the standardised matrices are available on \url{https://github.com/TGuillerme/Missing_living_mammals}.
+All the standardised matrices are available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices_binomial/Matrices}.
 
 \subsection{Selecting the living OTUs}
-
 To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H.
@@
 \section{2 - Data reproducibility}
-Every step of the analysis (apart from downloading and standardisation of the matrices) is entirely repeatable via GitHub (\url{https://github.com/TGuillerme/Missing_living_mammals}).
+Every step of the analysis (apart from downloading and standardisation of the matrices) is entirely repeatable via GitHub (\url{github.com/TGuillerme/Missing_living_mammals}).
 %Needs to fix this in GitHub

### 2016-01-25 Manuscript/BiologyLetters/response.tex
Updated responses

@@
 
+As shown in the ESM1 - Figure 1, the number of OTUs with cladistic data extracted from the Google Scholar searches seems to plateau really quickly (i.e. before the first 100 matches).
+Additionally, as Graeme Lloyd mention, ``older studies tend to be smaller (due to computing limitations) and to commonly reoccur anyway due to data set reuse'' (Reviewer 4, comment 1).
+Following these two observations, we think that including on the 20th most recent results per order (and higher taxonomic level) was a good balance between the time spending looking for papers, the amount of the OTUs extracted and the repeatability of the study.
+
 \item{\textcolor{blue}{Similarly, I find the 2010 date cutoff troubling, because I’m not sure that the argument that these recent studies contain a fraction of the previously published characters and OTUs allows you to make a comprehensive statement about the availability of character data.
@@
 This is a question that needs to be better addressed before you will convince readers that this is really all the data that is out there.}}
+
+%TO DO! Same answer as above?
+
 \end{enumerate}
@@
 [...] Is 100 really the minimum number you want to go with at all phylogenetic scales?}}
-\item{\textcolor{blue}{Going through each clade in the supplement seems beyond the scope of reviewing duties, but I am concerned that if this one matrix can be missed others may have been missed too.
+
+%TO DO: let's just remove this threshold thingy since I seem to be the only one to be convinced by it ;)...
+
+\item{\textcolor{blue}{I finally a little skeptical of the sampling done.
+Looking at Figure 1B, again the clade I’m most familiar with, representation for carnivorans appears very poor.
+Notwithstanding the issues discussed above regarding matrices with < 100 characters, there are definitely matrices out there that fulfill the authors’ requirements that are not represented.
+For example, a 349 character matrix with near complete sampling at the species is available for Viverridae, Prionodontidae, and Eupleridae in Gaubert et al (2005 Syst Biol.).
+Going through each clade in the supplement seems beyond the scope of reviewing duties, but I am concerned that if this one matrix can be missed others may have been missed too.
 I think a reexamination of the literature may be warranted using taxon specific search terms.}}
+
+%TO DO: OK, I think there is a problem here.
+Either we have to rerun the matrix downloading procedure and be sure we include the ones specified by the reviewers (the correct way!) but that WILL take more than a week.
+Or else we just add the matrices that are suggested y the reviewers but I have the feeling that that is a bit of cheating (i.e.
+we're hand picking some data here) - basically that'll just add more carnivora.
+%I think one argument could be to emphasize on one of the points we add in a former version of the manuscript (can't find it): "by available data we mean data in a standard format that can be directly downloaded from the internet (i.e.
+many of the papers we looked at had no link to their data or the link was broken, meaning that the data is effectively only available upon specific request to the authors [baaaad!!!])."
+
 \item{\textcolor{blue}{One interesting aside that the authors didn’t consider, despite the data being there, is how phylogenetic ``effort'' relates to diversity of the order under study.}}

### 2016-01-25 Analysis/Sampling_effort.R
Added sampling effort analysis

@@ -1,0 +1,108 @@
+#Loading the functions
+#Be sure to be in the right repository
+source("functions.R")
+load.functions(test=FALSE) #Set test=FALSE to speed up the loading
+library(picante)
+library(xtable)
+
+#Loading the tree
+mam_tree<-read.nexus("../Data/Trees/mammalST_MSW05_best_chrono.tre")
+#Resolve polytomies
+mam_tree<-multi2di(mam_tree)
+#Selecting one random tree
+#set.seed(1) ; mam_tree<-mam_tree[[sample(1:100, 1)]]
+
+#Loading the taxonomic reference list
+#Loading the taxonomic reference list
+WR_list<-read.csv("../Data/Taxon_References/WilsonReederMSW.csv", header=T, stringsAsFactors=F)
+#Changing ARTIODACTYLA and CETACEA to CETARTIODACTYLA
+WR_list$Order[which(WR_list$Order == "ARTIODACTYLA")]<-"CETARTIODACTYLA"
+WR_list$Order[which(WR_list$Order == "CETACEA")]<-"CETARTIODACTYLA"
+
+#Loading the list of present taxa
+#source("Extracting_living_taxa.R")
+load("../Data/List_of_matching_taxa/List_of_matching_taxa.Rda")
+
+#Fixing the number of characters for truncated matrices
+extraction_table[which(extraction_table$Matrix == "GS2007-R.nex"), 3]<-88
+extraction_table[which(extraction_table$Matrix == "GS2010-VJD.nex"), 3]<-157
+extraction_table[which(extraction_table$Matrix == "GS2012-GM.nex"), 3]<-45
+extraction_table[which(extraction_table$Matrix == "GS2012-SM.nex"), 3]<-157
+extraction_table[which(extraction_table$Matrix == "GS1997-V.nex"), 3]<-40
+extraction_table[which(extraction_table$Matrix == "GS2013-MS.nex"), 3]<-48
+extraction_table[which(extraction_table$Matrix == "GS2014-MG.nex"), 3]<-40
+extraction_table[which(extraction_table$Matrix == "GS2014-MaL.nex"), 3]<-98
+extraction_table[which(extraction_table$Matrix == "GS2015-RR.nex"), 3]<-149
+extraction_table[which(extraction_table$Matrix == "GS2014-B.nex"), 3]<-27
+extraction_table[which(extraction_table$Matrix == "GS2014-Y.nex"), 3]<-42
+
[…]

### 2016-01-26 Manuscript/BiologyLetters/response.tex

@@
 \bigskip
-We are very grateful to the four referees for their helpful and constructive comments, which we believe have helped us to significantly improve our paper.
+We are very grateful to the four referees for their helpful and constructive comments.
+%, which we believe have helped us to significantly improve our paper.
+% NC: A bit too much for what are minor revisions :)
 We have taken all of their comments on board, and respond to their points below.
-For improving clarity in this document, we dealt with each comment in the order they appeared in your ``Decision on Manuscript'' email and we copied the reviewers comments in blue.
+To improve the clarity of this document, we dealt with each comment in the order they appeared in your ``Decision on Manuscript'' email and we have included the reviewers comments in blue.
 
@@
 Why select only the first 10 results per search term?
-How do you know that relevant papers wouldn’t come up later in the search?
-I need evidence to support this cutoff; as it stands, I’m concerned that you’re missing some relevant material.}}
+How do you know that relevant papers wouldn't come up later in the search?
+I need evidence to support this cutoff; as it stands, I'm concerned that you're missing some relevant material.}}
+% NC: It's not just the first 10 results per search term is it? I thought you went through the first 10 pages of results? Clarify this.
 
-As shown in the ESM1 - Figure 1, the number of OTUs with cladistic data extracted from the Google Scholar searches seems to plateau really quickly (i.e. before the first 100 matches).
-Additionally, as Graeme Lloyd mention, ``older studies tend to be smaller (due to computing limitations) and to commonly reoccur anyway due to data set reuse'' (Reviewer 4, comment 1).
-Following these two observations, we think that including on the 20th most recent results per order (and higher taxonomic level) was a good balance between the time spending looking for papers, the amount of the OTUs extracted and the repeatability of the study.
+As shown in ESM1 Figure 1, the number of OTUs with morphological cladistic data extracted from the Google Scholar searches seems to plateau really quickly (i.e. before the first 100 matches).
+Additionally, as Graeme Lloyd mentions, ``older studies tend to be smaller (due to computing limitations) and to commonly reoccur anyway due to data set reuse'' (Reviewer 4, comment 1).
+Following these two observations, we think that including the 20 most recent results % NC: Is that what you did?
+ for each order (and higher taxonomic level) was a good balance between the time spent looking for papers, the number of OTUs extracted and the repeatability of the study.
 
-\item{\textcolor{blue}{Similarly, I find the 2010 date cutoff troubling, because I’m not sure that the argument that these recent studies contain a fraction of the previously published characters and OTUs allows you to make a comprehensive statement about the availability of character data.
-What about the fraction that isn’t included?
+\item{\textcolor{blue}{Similarly, I find the 2010 date cutoff troubling, because I'm not sure that the argument that these recent studies contain a fraction of the previously published characters and OTUs allows you to make a comprehensive statement about the availability of character data.
+What about the fraction that isn't included?
 This is a question that needs to be better addressed before you will convince readers that this is really all the data that is out there.}}
@@
 
+% NC: You replace something *with* something else not *by*. Things can be replaced by something, but the sentence set up is then "X was replaced by Y", not "replaced X by Y".
+
 \subsection{Reviewer 1 specific comments:}
@@
 \item{\textcolor{blue}{P.2, line 43: ``chunks'' is not an appropriate term to use in a scientific paper. Use a more formal term.}}
-We replaced ``chunks'' by ``parts'' on p.@, line@.
+We replaced ``chunks'' with ``parts'' % on p.@, line@. NC: No need for this bit where it's minor comments and the reviewer has already highlighted the position.
+% NC: Sections may also work instead of parts?
 
 \item{\textcolor{blue}{P.2, L53 \& L55: ``branch'' is the wrong verb. You want to say something like ``fossils will not be placed accurately within the phylogeny''.}}
-We replaced ``fossils cannot branch in the correct clade'' by ``fossils will not be placed accurately within the correct clade'' on p.@, line@.
+We replaced ``fossils cannot branch in the correct clade'' with ``fossils will not be placed accurately within the correct clade''.
 
 \item{\textcolor{blue}{P3, L78: ``since'' should be ``because.'' Since should be used for time, not causation.}}
-We replaced ``since'' by ``because'' on p.@, line@.
+We replaced ``since'' with ``because'' throughout the manuscript as appropriate.
+% NC: Can't believe I missed some of these typos!
 
 \item{\textcolor{blue}{P3, L85: ``not-applicable'' should not be hyphenated.}}
-We removed the hyphen on p.@, line@.
+We removed the hyphen.
 
@@
 It would also help if you could show in your data table what the clades were that had better than 75\% coverage.
-It’s not immediately evident from reading Table 1.}}
+It's not immediately evident from reading Table 1.}}
 
@@
 \item{\textcolor{blue}{P5, L137: ``taxa'' should be ``OTUs.''}}
-We replaced ``taxa'' by ``OTUs'' on p.@, line@.
+We replaced ``taxa'' with ``OTUs''
 
 \item{\textcolor{blue}{P5, L165 \& 166: This sentence also misuses ``branch'' as on P.2. Reword for correct verb usage.}}
-We replaced ``unable to branch in the Herpestidae'' by ``unable to be placed in the Herpestidae clade'' on p.@, line@.
+We replaced ``unable to branch in the Herpestidae'' with ``unable to be placed in the Herpestidae clade''.
 
@@
 \item{\textcolor{blue}{ESM 1, p. 4: First sentence of the paragraph in the middle of the page is a sentence fragment, and needs to state where the matrices are available.}}
-We added the two missing link in the ESM1: ``The list of all the 286 download matrices is available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices}.'' and ``All the standardised matrices are available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices_binomial/Matrices}.''
+We added the two missing links in the ESM1: ``The list of all the 286 download matrices is available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices}.'' and ``All the standardised matrices are available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices_binomial/Matrices}.''
 
@@
 Also, the sentence starting the paragraph under ``Selecting the living OTUs'' repeats the exact same thing as this sentence.}}
-We modified the sentence to: ``[we] fixed the incorrectly-formatted bionomial names format (e.g. \textit{H. sapiens} became \textit{Homo sapiens})''
+We modified the sentence to: ``[we] fixed the incorrectly-formatted binomial names (e.g. \textit{H. sapiens} became \textit{Homo sapiens})'' % NC: Note changes I have made to this need to go into ESM.
+% NC: Did you also remove the repetition in the next paragraph?
 
@@
 
-\section{Reviewer 2:}
+\section{Reviewer 2 (Graham Slater):} % NC: I feel like some of the comments are missing here?
 \begin{enumerate}
@@
 %TO DO: let's just remove this threshold thingy since I seem to be the only one to be convinced by it ;)...
+% NC: Yep that seems sensible
 
-\item{\textcolor{blue}{I finally a little skeptical of the sampling done.
-Looking at Figure 1B, again the clade I’m most familiar with, representation for carnivorans appears very poor.
-Notwithstanding the issues discussed above regarding matrices with < 100 characters, there are definitely matrices out there that fulfill the authors’ requirements that are not represented.
+\item{\textcolor{blue}{I finally am a little skeptical of the sampling done.
+Looking at Figure 1B, again the clade I'm most familiar with, representation for carnivorans appears very poor.
+Notwithstanding the issues discussed above regarding matrices with < 100 characters, there are definitely matrices out there that fulfill the authors' requirements that are not represented.
 For example, a 349 character matrix with near complete sampling at the species is available for Viverridae, Prionodontidae, and Eupleridae in Gaubert et al (2005 Syst Biol.).
@@
 
-\item{\textcolor{blue}{One interesting aside that the authors didn’t consider, despite the data being there, is how phylogenetic ``effort'' relates to diversity of the order under study.}}
+% NC: Hmmm. I'm concerned about this, and I think your "quick and dirty" solution is pretty dishonest. I think you're going to need to look at your data collection procedure again. 
+% First we need to understand *why* is the paper Graham mentions missing?
+Because you focus on data from 2010 onwards?
+Because it's not in Graeme or Ross' databases?
+Need to understand how it got missed out.
+Does it appear if you get rid of the 10/20 search results only and search lots of results?
+Is there a problem with search terms?
+The answer to this will help us decide how to proceed from here.
+% NC: Although you don't want to waste a huge amount of time on this, you also don't want a reputation for being sloppy, nor a reputation for glossing over your errors.
+Also it's important for personal research integrity.
+If that means you spend two weeks recollecting the data then that will have to be accepted.
+Note that two weeks is really a very short amount of time in the usual 2 years it takes to write a paper (Martin will be expecting you to have brought some work over from your PhD), but you will need to ask for an extension on the corrections.
+This won't be a problem.
+Either way this comment will have to explain how the papers Graham mentions are missed out.
+If you can do that and it's sensible then explaining the latter point about locating the data is probably fine.
+And I know you're busy - welcome to the world post PhD!
+All things happen at once...
 
-We are really grateful for Graham Slater's %or do we need to keep it formal and say just "reviewer 2's"
-comment and added this interesting analysis in the ESM @@@% get number.
-% that mammalian morphological phylogeneticists have focused their energies on resolving the higher level relationships among mammalian lineages.
+\item{\textcolor{blue}{One interesting aside that the authors didn't consider, despite the data being there, is how phylogenetic ``effort'' relates to diversity of the order under study.}}
+
+This is a great suggestion so we have added this analysis to the ESM @@@% get number.
+% We show that mammalian morphological phylogeneticists have focused their energies on resolving the higher level relationships among mammalian lineages.
+
+% NC: May also want to add something to the main text.
 \end{enumerate}
@@
 
-\section{Reviewer 3:}
+\section{Reviewer 3 (Peter Wagner):}
 \begin{enumerate}
@@
 [...] Thus, the title really should be ``Assessment of available anatomical characters for phylogenetic analysis among living mammals'' or something like that.}}
+
+%NC: Change the title?
+
 \item{\textcolor{blue}{what we really need is to focus on the tooth and skull characters that typify the fossil record.
@@
 Thus, we might be able to richly populate FBD and tip-dating analyses with a LOT of extinct species.}}
+
+% NC: worth commenting here that we are looking at the living species?
+
 \item{\textcolor{blue}{Finally, I would add: who the heck submits a paper for a special volume 6 weeks early?!?!? You make the rest of us look bad!}}
-This paper was actually ready to submit from the 28th of July (\url{http://biorxiv.org/content/early/2015/07/28/022970}) which partially explain the swift submission.
-We do sincerely apologize for making the rest of the authors submitting to the issue look bad.
+This paper formed part of my PhD thesis that I submitted in September which partially explains the swift submission! % NC: Worth making a tiny joke but no need to show off :P
+Sorry for making the rest of the authors submitting to the issue look bad :).
 
@@
 
+% NC: If you end up needing to redo the data collection it might be worth checking the PBDB...
+
 \item{\textcolor{blue}{Lines 164-166: ``For example, a Carnivora fossil will be unable to branch in the Herpestidae, and will have more chance to randomly branch within Canidae (Figure 1B).'' 
@@
 Thus, it is not possible for coded skeletal characters from fossil herpestids to connect those fossil herpestids to the extant herpestids.
-This is the opposite of Felsenstein’s long branch: insofar as skeletal characters are concerned, extant herpestids have zero-length branches.
+This is the opposite of Felsenstein's long branch: insofar as skeletal characters are concerned, extant herpestids have zero-length branches.
 That in turn means that the fossil herpestids are going to be ``attracted'' to non-zero length branches elsewhere, with long branches in related groups such as canids being particularly good candidates because there is a higher chance of homoplasies between canids and herpesitds on the longer branches.
 So, it will not be random: they probably will latch themselves onto the longest reasonably branch among the closest relatives for which anatomical data are present.}}
+
+% NC: Yep he's right, modify the sentence to account for this. Also cite a long branch attraction stuff. Check if Pete has written anything - it'd be nice to cite him as he's been so helpful.
 
@@
 \begin{enumerate}
-\item{\textcolor{blue}{Reviewer 1’s most substantial complaints (the year and page number cut offs for google scholar searches) seem at least partially valid.
+\item{\textcolor{blue}{Reviewer 1's most substantial complaints (the year and page number cut offs for google scholar searches) seem at least partially valid.
 However, in my experience older studies tend to be smaller (due to computing limitations) and to commonly reoccur anyway due to data set reuse.
-The authors also already state that subsequent search results tended to yield no substantial additional data, and by the reviewer’s own admission the assembled data seem to look correct.
-Thus a recollecting of the data doesn’t seem necessary (but see further comments below).}}
+The authors also already state that subsequent search results tended to yield no substantial additional data, and by the reviewer's own admission the assembled data seem to look correct.
+Thus a recollecting of the data doesn't seem necessary (but see further comments below).}}
 
-\item{\textcolor{blue}{Reviewer 2 (Graham Slater)’s complaint are - in my view - the most substantial.
+% NC: This is clearly a get out clause to get this done more rapidly.
+
+\item{\textcolor{blue}{Reviewer 2 (Graham Slater)'s complaint are - in my view - the most substantial.
 I agree that the >100 characters cut off is not as strongly justified as it might be.
@@
 From my reading of the two tables only 11 of the orders seem to have below 25\% coverage at the species-level (not 22 as the authors state in the main text!) when the threshold is used, and this improves slightly to 10 for when the threshold is removed.
-As this difference isn’t dramatic it seems worth mentioning.
+As this difference isn't dramatic it seems worth mentioning.
 Reviewer 2 also notes that some important studies may be missed.
-This seems valid (I am aware of the biases in Ross Mounce’s data collection protocol, and obviously my own) and I am sure there are many more data sets that could be sampled.
+This seems valid (I am aware of the biases in Ross Mounce's data collection protocol, and obviously my own) and I am sure there are many more data sets that could be sampled.
 In particular I am aware of a number of studies that can only be found in books and hence are not easily accessed even if they occur in a google scholar search.
@@
 Nevertheless, it does deserve mention in the discussion as a potential bias in the results.
-I don’t think this undermines the study, but rather should be viewed as both a reason for optimism and an acknowledgement of a separate problem (i.e., access to cladistic data) which multiple publications already cover (e.g., see Ross Mounce’s work).}}
+I don't think this undermines the study, but rather should be viewed as both a reason for optimism and an acknowledgement of a separate problem (i.e., access to cladistic data) which multiple publications already cover (e.g., see Ross Mounce's work).}}
 
-\item{\textcolor{blue}{Reviewer 3 (Pete Wagner)’s major problem - with which I agree - is semantic, and thus easily fixable with a few explicit caveats (e.g., ``by cladistic we mean...'').}}
+% NC: OK so Graeme is going with your "data accessibility" idea here to get round having to redo the data collection.
+I think this will be ok as long as you add comments to that effect in the text, and work out why Graham's papers are missing.
+
+% NC: Get rid of the 100 characters thing (I never really understood why it was needed), and make sure all the results are definitely correct.
+
+\item{\textcolor{blue}{Reviewer 3 (Pete Wagner)'s major problem - with which I agree - is semantic, and thus easily fixable with a few explicit caveats (e.g., ``by cladistic we mean...'').}}
 
@@
 This could surely be synthesised from the tables and would be a nice simple way to compare the with and without threshold results.
-Similarly, I think Reviewer 3’s [We assume Graeme Lloyd is refering to Graham Slater's reviewe (reviewer 2)] %or same as comment before: no names!
+Similarly, I think Reviewer 3's [We assume Graeme Lloyd is refering to Graham Slater's reviewe (reviewer 2)] %or same as comment before: no names!
  graphs are a nice potential addition to the ESM.}}
 \end{enumerate}
+
+% NC: Easy enough to add a line to all the tables for total coverage
 
@@
 \item{\textcolor{blue}{Lines 72-74 - Web addresses should be consistent in dropping of ``http://www.''.}}
-We removed the ``http://www.'' part of the URLs on p.@, lines@ and in the ESM1.
+We removed the ``http://www.'' part of the URLs.
 
 \item{\textcolor{blue}{Line 104 - ``NTI 15'' should probably be ``NTI [15]''.}}
-%TO DO IN THE WORD MANUSCRIPT! We fixed this typo.
+%TO DO IN THE WORD MANUSCRIPT! 
+We fixed this typo.
 
-\item{\textcolor{blue}{Line 104 - It’s not completely clear to me how ``distance'' is measured. Is this branch-lengths on a time-scaled tree? Or just a node count?}}
-We changed ``$\overline{MNND}_{obs}$ is the observed mean distance between each of $n$ taxa'' to ``$\overline{MNND}_{obs}$ is the observed mean branch length between each of $n$ taxa'' and ``$\overline{MPD}_{obs}$ is the observed mean phylogenetic distance of the tree containing only the $n$ taxa'' to ``$\overline{MPD}_{obs}$ is the observed mean phylogenetic branch length of the tree containing only the $n$ taxa'' on p.@, lines@.
+\item{\textcolor{blue}{Line 104 - It's not completely clear to me how ``distance'' is measured. Is this branch-lengths on a time-scaled tree? Or just a node count?}}
+We have clarified this by changing ``observed mean distance between each of $n$ taxa'' to ``observed mean sum of the branch lengths between each of $n$ taxa''. % NC: Fix in document
 
-\item{\textcolor{blue}{Line 117 - It’s not clear to me how supraspecific values were calculated if the reference phylogeny (ref 10) is species-level.
+\item{\textcolor{blue}{Line 117 - It's not clear to me how supraspecific values were calculated if the reference phylogeny (ref 10) is species-level.
 Are you using the node subtending the higher taxon or something else?}}
 %TO CHECK!
+% NC: Probably you've pruned the phylogeny to include only species within each order/family?
 
-\item{\textcolor{blue}{Line 149 - ``most palaeontological studies use the genus as their smallest OTU'' I’m not sure if this is strictly true.
-It’s more that shorter OTU names are commonly used.
+\item{\textcolor{blue}{Line 149 - ``most palaeontological studies use the genus as their smallest OTU'' I'm not sure if this is strictly true.
+It's more that shorter OTU names are commonly used.
 After all at least one actual specimen has to be coded and in my experience if you read the paper usually only a single species is used.
@@
 In addition ref 20 is actually all species-level OTUs!
-(I didn’t check ref 21.)}}
+(I didn't check ref 21.)}}
 %To CHECK, do we actually care? 
+% NC: Just change it from "most" to "many", or say paleo studies commonly feature only one species in a genus.
 
 \item{\textcolor{blue}{Line 154 - ``phylogenetically overdispersed'' Purely semantic, but I guess what you mean is that even sampling would be ideal, and this is *measured* using phylogenetic over-dispersion.}}
-We changed ``the ideal scenario is for them to be phylogenetically overdispersed'' to ``the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion)'' on p.@, line@.
+We changed ``the ideal scenario is for them to be phylogenetically overdispersed'' to ``the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion)''.
 
@@
 Blue branches indicate available cladistic data for the species.
-Cladistic data is randomly distributed in Primates (A) and clustered in Carnivora (B).''.
+Available cladistic data is randomly distributed in Primates (A) and clustered in Carnivora (B).''.
 
 \item{\textcolor{blue}{ESM 1 - Incomplete sentence: ``The list of all the 286 download matrices is available on''}}
-See reviewer 1 specific comment 12.
+we fixed this typo. See reviewer 1 specific comment 12.
 
 \item{\textcolor{blue}{ESM 1 - Incomplete sentence: ``All the standardised matrices are available on''}}
-See reviewer 1 specific comment 12.
+We fixed this typo. See reviewer 1 specific comment 12.
 
 \item{\textcolor{blue}{ESM 1 - ``designated as ``fossil'' all OTUs that were present in the Paleobiology database'' The database contains extant taxa too, although these should be marked as such.
-I guess this doesn’t matter given the flow chart in Figure 2, but is something the authors should be aware of.
+I guess this doesn't matter given the flow chart in Figure 2, but is something the authors should be aware of.
 (Obviously more generally living species can have fossil records, or at least plenty of fossils are assigned to extant species.)}}
@@
 Specimens?
-Or just species you can’t reconcile with your three data sets?}}
+Or just species you can't reconcile with your three data sets?}}
 %TO DO: add the number of ignored OTUs. I think they're mainly specimen IDs without species names.
@@
 \item{\textcolor{blue}{ESM 2 - ``The following section contains supplementary results to the main body: the available data structure using the NTI and the PD metric'' Abbreviations should be introduced when first used, not later as they are here.}}
-We changed ``using the NTI and the PD metric'' to ``using the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) metrics''.
+We changed this to ``using the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) metrics''.
 \end{enumerate}
@@
 
+% NC: Let me know if you need help with this.
+
 Thomas Guillerme (guillert@tcd.ie)\\ % Changing email + adding imperial address?
+% NC: Just change email.
+For the paper, you can change your corresponding address but I wouldn't add imperial as another proper address as you did most of the work at trinity.
 Natalie Cooper

### 2016-01-26 Manuscript/BiologyLetters/response.tex
Dealt with part of Natalie's comments

@@
 I need evidence to support this cutoff; as it stands, I'm concerned that you're missing some relevant material.}}
-% NC: It's not just the first 10 results per search term is it? I thought you went through the first 10 pages of results? Clarify this.
+% NC: It's not just the first 10 results per search term is it?
+I thought you went through the first 10 pages of results?
+Clarify this.
+TG: No I collected the 20 first results.
+It's mentioned down there.
 
@@
 Additionally, as Graeme Lloyd mentions, ``older studies tend to be smaller (due to computing limitations) and to commonly reoccur anyway due to data set reuse'' (Reviewer 4, comment 1).
-Following these two observations, we think that including the 20 most recent results % NC: Is that what you did?
+Following these two observations, we think that including the 20 most recent results % NC: Is that what you did? TG: yep.
  for each order (and higher taxonomic level) was a good balance between the time spent looking for papers, the number of OTUs extracted and the repeatability of the study.
@@
 \item{\textcolor{blue}{P.2, line 43: ``chunks'' is not an appropriate term to use in a scientific paper. Use a more formal term.}}
-We replaced ``chunks'' with ``parts'' % on p.@, line@. NC: No need for this bit where it's minor comments and the reviewer has already highlighted the position.
-% NC: Sections may also work instead of parts?
+We replaced ``chunks'' with ``sections''.
 
@@
 We replaced ``since'' with ``because'' throughout the manuscript as appropriate.
-% NC: Can't believe I missed some of these typos!
 
@@
 Also, the sentence starting the paragraph under ``Selecting the living OTUs'' repeats the exact same thing as this sentence.}}
-We modified the sentence to: ``[we] fixed the incorrectly-formatted binomial names (e.g. \textit{H. sapiens} became \textit{Homo sapiens})'' % NC: Note changes I have made to this need to go into ESM.
-% NC: Did you also remove the repetition in the next paragraph?
+We modified the sentence to: ``[we] fixed the incorrectly-formatted binomial names (e.g. \textit{H. sapiens} became \textit{Homo sapiens})''.
 
@@
 % *An author contributions section which fulfils all *four* of our criteria given here https://royalsociety.org/journals/ethics-policies/openness/, including a sentence stating all authors agree to be held accountable for the content therein and approve the final version of the manuscript
+%TG: TO DO: CHECK
 % *A funding statement which is the same as that given in ScholarOne
+%TG: To modify
 % *Please also confirm in your cover letter whether all the figures are your own or whether permission has been obtained for their use
+%TG: TO DO (yes)
 % *Please upload your original figure files as eps, tiff or jpeg files, rather than PDFs
+%TG: TO DO (easy)
 % *If you have any images that can be used to promote your article on social media (should this be accepted) please upload them as a supplementary file
+%TG: what about the "phylogeny" with the duck and the pig?
 % *Please confirm whether your manuscript should be available via open access, as we note that this may be a requirement of your funder.
+%TG: TO DO (yes)
 
 % NC: Let me know if you need help with this.
+%TG: It'll probably be all right :)
 
-Thomas Guillerme (guillert@tcd.ie)\\ % Changing email + adding imperial address?
+Thomas Guillerme (t.guillerme@imperial.ac.uk)\\ % Changing email + adding imperial address?
 % NC: Just change email.

### 2016-01-26 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Dealt with part of Natalie's comments

@@
 A downside of this method is that it requires molecular data for living taxa and morphological data for both living and fossil taxa.
-Parts%rev 1.2.1 (as in reviewer 1, part 2 (minor), comment 1)
+Sections%rev 1.2.1 (as in reviewer 1, part 2 (minor), comment 1)
  of this data can be difficult, or impossible, to collect for every taxon in the analysis.

### 2016-01-26 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
Dealt with part of Natalie's comments

@@
 We then cleaned the nexus matrices by removing any extra information (trees, continuous characters, morphological characters description, molecular data) to end up with nexus matrices containing only the discrete morphological data.
-We then manually fixed the incorrectly-formatted bionomial names format (e.g. \textit{H. sapiens} became \textit{Homo sapiens}) using the abbreviation list in the concerned publications. 
+We then manually fixed the incorrectly-formatted binomial names (e.g. \textit{H. sapiens} became \textit{Homo sapiens}) using the abbreviation list in the concerned publications. 
 All the standardised matrices are available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices_binomial/Matrices}.
@@
 \subsection{Selecting the living OTUs}
-To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H.
+%To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H.
 sapiens} was transformed to \textit{Homo sapiens}).

### 2016-01-26 Manuscript/BiologyLetters/Table_results.tex
Changed to table without character threshold

--- 2015-06-20 Manuscript/Table_results.tex
@@
 % latex table generated in R 3.2.0 by xtable 1.7-4 package
-% Sat Jun 20 17:41:32 2015
+% Sat Jun 20 17:19:07 2015
 \begin{longtable}{lL{1.8cm}C{2cm}lcc}
-\caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels.
+\caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels (without any character threshold; results from the 286 matrices).
 The coverage represents the proportion of taxa with available morphological data.
@@
 Significant NRI or NTI are highlighted in bold.
-One star (*) represents a p-value between 0.05 and 0.005; two starts between 0.005 and 0.0005 and three stars a p-value less than 0.0005.} \\
+*p \textless 0.05; **p \textless 0.01; ***p \textless 0.001.} \\
   \hline
@@
   Afrosoricida & genus & 17/17 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar2.pdf} &   &   \\ 
-  \textbf{Afrosoricida} & \textbf{species} & \textbf{23/42} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar3.pdf} & \textbf{1.89*} & \textbf{1.19} \\ 
-  Carnivora & family & 11/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar4.pdf} & 0.43 & 1.68 \\ 
-  \textbf{Carnivora} & \textbf{genus} & \textbf{30/125} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar5.pdf} & \textbf{4.14**} & \textbf{1.81*} \\ 
-  \textbf{Carnivora} & \textbf{species} & \textbf{42/283} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar6.pdf} & \textbf{18.64**} & \textbf{3.02**} \\ 
+  Afrosoricida & species & 23/42 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar3.pdf} & 1.75 & 1.08 \\ 
+  Carnivora & family & 14/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar4.pdf} & 0.63 & 0.6 \\ 
+  \textbf{Carnivora} & \textbf{genus} & \textbf{54/125} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar5.pdf} & \textbf{4.81**} & \textbf{1.78*} \\ 
+  \textbf{Carnivora} & \textbf{species} & \textbf{76/283} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar6.pdf} & \textbf{7.66**} & \textbf{0.85} \\ 
   Cetartiodactyla & family & 21/21 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar7.pdf} &   &   \\ 
-  \textbf{Cetartiodactyla} & \textbf{genus} & \textbf{77/128} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar8.pdf} & \textbf{0.87} & \textbf{1.77*} \\ 
-  \textbf{Cetartiodactyla} & \textbf{species} & \textbf{129/310} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar9.pdf} & \textbf{2.72*} & \textbf{0.04} \\ 
-  Chiroptera & family & 13/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar10.pdf} & 0.55 & 0.63 \\ 
-  \textbf{Chiroptera} & \textbf{genus} & \textbf{85/202} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar11.pdf} & \textbf{16.91**} & \textbf{2.85**} \\ 
-  \textbf{Chiroptera} & \textbf{species} & \textbf{165/1053} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar12.pdf} & \textbf{14.55**} & \textbf{3.44**} \\ 
+  Cetartiodactyla & genus & 100/128 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar8.pdf} & 0.85 & 0.94 \\ 
+  \textbf{Cetartiodactyla} & \textbf{species} & \textbf{171/310} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar9.pdf} & \textbf{1.92*} & \textbf{-0.46} \\ 
+  Chiroptera & family & 15/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar10.pdf} & -0.28 & 0.56 \\ 
+  \textbf{Chiroptera} & \textbf{genus} & \textbf{93/202} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar11.pdf} & \textbf{13.47**} & \textbf{1.1} \\ 
+  \textbf{Chiroptera} & \textbf{species} & \textbf{215/1053} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar12.pdf} & \textbf{8.82**} & \textbf{1.22} \\ 
   Cingulata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar13.pdf} &   &   \\ 
-  Cingulata & genus & 8/9 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar14.pdf} & 1.49 & -1.63 \\ 
-  Cingulata & species & 6/29 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar15.pdf} & 1.43 & 0.36 \\ 
+  Cingulata & genus & 8/9 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar14.pdf} & 1.51 & -1.57 \\ 
+  \textbf{Cingulata} & \textbf{species} & \textbf{9/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar15.pdf} & \textbf{1.9*} & \textbf{0.11} \\ 
   Dasyuromorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar16.pdf} &   &   \\ 
-  Dasyuromorphia & genus & 7/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar17.pdf} & -1 & -1.45 \\ 
-  Dasyuromorphia & species & 8/64 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar18.pdf} & -1.15 & -0.62 \\ 
+  Dasyuromorphia & genus & 8/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar17.pdf} & -0.75 & -1.07 \\ 
+  Dasyuromorphia & species & 9/64 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar18.pdf} & -0.88 & -0.34 \\ 
   Dermoptera & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar19.pdf} &   &   \\ 
@@
   Didelphimorphia & genus & 16/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar23.pdf} &   &   \\ 
-  Didelphimorphia & species & 40/84 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar24.pdf} & -0.94 & 0.36 \\ 
-  Diprotodontia & family & 9/11 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar25.pdf} & -0.8 & 0.56 \\ 
-  Diprotodontia & genus & 20/38 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar26.pdf} & -1.36 & -0.73 \\ 
-  Diprotodontia & species & 16/126 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar27.pdf} & -2.29 & -1.55 \\ 
+  Didelphimorphia & species & 42/84 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar24.pdf} & -1.65 & 0.2 \\ 
+  Diprotodontia & family & 11/11 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar25.pdf} &   &   \\ 
+  Diprotodontia & genus & 25/38 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar26.pdf} & -1.13 & -1.31 \\ 
+  Diprotodontia & species & 31/126 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar27.pdf} & 0.48 & -1.77 \\ 
   Erinaceomorpha & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar28.pdf} &   &   \\ 
   Erinaceomorpha & genus & 10/10 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar29.pdf} &   &   \\ 
-  Erinaceomorpha & species & 21/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar30.pdf} & -1.1 & -0.3 \\ 
+  Erinaceomorpha & species & 21/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar30.pdf} & -1.07 & -0.2 \\ 
   Hyracoidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar31.pdf} &   &   \\ 
@@
   Hyracoidea & species & 1/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar33.pdf} &   &   \\ 
-  Lagomorpha & family & 1/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar34.pdf} &   &   \\ 
-  Lagomorpha & genus & 1/12 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar35.pdf} &   &   \\ 
-  Lagomorpha & species & 1/86 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar36.pdf} &   &   \\ 
+  Lagomorpha & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar34.pdf} &   &   \\ 
+  Lagomorpha & genus & 5/12 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar35.pdf} & -1.06 & -0.95 \\ 
+  Lagomorpha & species & 12/86 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar36.pdf} & -0.62 & -1.88 \\ 
   Macroscelidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar37.pdf} &   &   \\ 
   Macroscelidea & genus & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar38.pdf} &   &   \\ 
-  Macroscelidea & species & 5/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar39.pdf} & -0.98 & -1.38 \\ 
+  Macroscelidea & species & 12/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar39.pdf} & -1.3 & -1.06 \\ 
   Microbiotheria & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar40.pdf} &   &   \\ 
@@
   Monotremata & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar43.pdf} &   &   \\ 
-  Monotremata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar44.pdf} & -0.71 & -0.71 \\ 
-  Monotremata & species & 2/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar45.pdf} & -1.01 & -1.03 \\ 
+  Monotremata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar44.pdf} & -0.72 & -0.69 \\ 
+  Monotremata & species & 2/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar45.pdf} & -0.97 & -0.97 \\ 
   Notoryctemorphia & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar46.pdf} &   &   \\ 
@@
   Paucituberculata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar49.pdf} &   &   \\ 
-  Paucituberculata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar50.pdf} & 0 & 0 \\ 
-  Paucituberculata & species & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar51.pdf} & -0.64 & -0.65 \\ 
+  Paucituberculata & genus & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar50.pdf} &   &   \\ 
+  Paucituberculata & species & 5/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar51.pdf} &   &   \\ 
   Peramelemorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar52.pdf} &   &   \\ 
   Peramelemorphia & genus & 7/7 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar53.pdf} &   &   \\ 
-  Peramelemorphia & species & 16/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar54.pdf} & -0.09 & 1 \\ 
+  Peramelemorphia & species & 16/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar54.pdf} & -0.13 & 0.97 \\ 
   Perissodactyla & family & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar55.pdf} &   &   \\ 
   Perissodactyla & genus & 6/6 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar56.pdf} &   &   \\ 
-  Perissodactyla & species & 7/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar57.pdf} & 0.62 & -2.5 \\ 
+  Perissodactyla & species & 10/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar57.pdf} & -0.07 & -2.63 \\ 
   Pholidota & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar58.pdf} &   &   \\ 
   Pholidota & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar59.pdf} &   &   \\ 
-  \textbf{Pholidota} & \textbf{species} & \textbf{3/8} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar60.pdf} & \textbf{2.64*} & \textbf{2.23*} \\ 
-  Pilosa & family & 3/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar61.pdf} & 0.94 & 0.93 \\ 
-  Pilosa & genus & 3/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar62.pdf} & -0.36 & -0.31 \\ 
-  Pilosa & species & 3/29 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar63.pdf} & 0.33 & 0.79 \\ 
+  Pholidota & species & 4/8 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar60.pdf} & 1.18 & 0.94 \\ 
+  Pilosa & family & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar61.pdf} & 1.87 & 2 \\ 
+  Pilosa & genus & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar62.pdf} & -0.96 & 0.36 \\ 
+  \textbf{Pilosa} & \textbf{species} & \textbf{5/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar63.pdf} & \textbf{1.28} & \textbf{2.38*} \\ 
   Primates & family & 15/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar64.pdf} &   &   \\ 
-  Primates & genus & 48/68 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar65.pdf} & -0.41 & -1.4 \\ 
-  Primates & species & 56/351 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar66.pdf} & -1.6 & -2.04 \\ 
+  Primates & genus & 48/68 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar65.pdf} & -0.35 & -1.33 \\ 
+  Primates & species & 64/351 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar66.pdf} & -0.67 & -1.27 \\ 
   Proboscidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar67.pdf} &   &   \\ 
-  Proboscidea & genus & 1/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar68.pdf} &   &   \\ 
-  Proboscidea & species & 1/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar69.pdf} &   &   \\ 
-  Rodentia & family & 11/32 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar70.pdf} & -0.46 & -1.91 \\ 
-  Rodentia & genus & 21/450 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar71.pdf} & -2.11 & 0.3 \\ 
-  Rodentia & species & 15/2094 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar72.pdf} & -1.65 & -2.55 \\ 
+  Proboscidea & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar68.pdf} &   &   \\ 
+  Proboscidea & species & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar69.pdf} & -0.69 & -0.69 \\ 
+  Rodentia & family & 18/32 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar70.pdf} & 0.66 & -0.98 \\ 
+  Rodentia & genus & 82/450 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar71.pdf} & -1.66 & 1.55 \\ 
+  \textbf{Rodentia} & \textbf{species} & \textbf{90/2094} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar72.pdf} & \textbf{2.76*} & \textbf{2.34*} \\ 
   Scandentia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar73.pdf} &   &   \\ 
-  Scandentia & genus & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar74.pdf} & -0.77 & -0.76 \\ 
-  Scandentia & species & 2/20 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar75.pdf} & -1.79 & -1.99 \\ 
+  Scandentia & genus & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar74.pdf} & -0.74 & -0.74 \\ 
+  Scandentia & species & 3/20 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar75.pdf} & -1.88 & -0.84 \\ 
   Sirenia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar76.pdf} &   &   \\ 
@@
   Sirenia & species & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar78.pdf} &   &   \\ 
-  Soricomorpha & family & 3/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar79.pdf} & -0.93 & -0.92 \\ 
-  \textbf{Soricomorpha} & \textbf{genus} & \textbf{19/43} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar80.pdf} & \textbf{6.98**} & \textbf{2.49*} \\ 
-  \textbf{Soricomorpha} & \textbf{species} & \textbf{19/392} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar81.pdf} & \textbf{13.19**} & \textbf{3.89**} \\ 
+  Soricomorpha & family & 3/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar79.pdf} & -0.98 & -0.99 \\ 
+  \textbf{Soricomorpha} & \textbf{genus} & \textbf{19/43} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar80.pdf} & \textbf{7.11**} & \textbf{2.59**} \\ 
+  \textbf{Soricomorpha} & \textbf{species} & \textbf{21/392} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar81.pdf} & \textbf{10.65**} & \textbf{3.56**} \\ 
   Tubulidentata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar82.pdf} &   &   \\ 

### 2016-01-26 Analysis/Results_summary.R
Updated analysis (no threshold)

@@ -48,7 +48,7 @@ extraction_list=list(extraction_table1, extraction_table100)
 results_names=c("results_1.Rda","results_100.Rda")
 
 #Select the character threshold
-extraction_table<-extraction_list[[2]]
+extraction_table<-extraction_list[[1]]
 
 #Selecting the living taxa (for a minimal number of characters)
 living_taxa<-extraction_table[which(extraction_table$Living == TRUE),]
@@ -59,7 +59,8 @@ living_taxa_list<-unique(living_taxa$Taxa)
 ######################
 
 #Loading the results
-load("results_100.Rda")
+load("results_1.Rda")
+#load("results_100.Rda")
 
 
 ############################################
@@ -100,7 +101,7 @@ OTUS_total<-OTUS_number_val[-seq(from=1, to=length(OTUS_number_val), by=2)]
 supp_order<-orders[which(as.numeric(OTUS_total) >= 20)]
 
 #Supplementary figures (orders with more than 20 taxa)
-for (order in 11:length(supp_order)) {
+for (order in 1:length(supp_order)) {
     pdf(paste("../Manuscript/Supplementary/Supp_figure_", supp_order[[order]], ".pdf", sep=""))
     op<-par(mfrow=c(1,1), oma=c(0,1,0,1))
     plot.results(order=supp_order[[order]], tree=mam_tree, taxa=living_taxa_list, col_branch=c("blue", "grey"), reference=WR_list, verbose=TRUE)

### 2016-01-26 Manuscript/BiologyLetters/Supplementary/ESM_2.tex
Updated email

@@
 \medskip
-\noindent{*\bf Corresponding author.} \textit{guillert@tcd.ie}\\  
+\noindent{*\bf Corresponding author.} \textit{t.guillerme@imperial.ac.uk}\\  
 \vspace{1in}
@@
 
-\section{Supplementary results - 1}
-The following section contains supplementary results to the main body: the available data structure using the NTI and the PD metric; the proportion of available data and the data structure for all the matrices (including the matrices with less than 100 characters).
+\section{Supplementary results - 2}
+The following section contains supplementary results to the main body: the phylogenetic representation of the data availability per order (excluding Primates and Carnivora, present in the main body).
 
-\input{Results_1c/Results_1cTable_results}
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_AFROSORICIDA.pdf}
+\caption{Distribution of available morphological data across Afrosoricida. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Afrosoricida}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_CHIROPTERA.pdf}
+\caption{Distribution of available morphological data across Chiroptera. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Chiroptera}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_CINGULATA.pdf}
+\caption{Distribution of available morphological data across Cingulata. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Cingulata}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_DASYUROMORPHIA.pdf}
+\caption{Distribution of available morphological data across Dasyuromorphia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Dasyuromorphia}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_DIDELPHIMORPHIA.pdf}
+\caption{Distribution of available morphological data across Didelphimorphia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Didelphimorphia}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_DIPROTODONTIA.pdf}
+\caption{Distribution of available morphological data across Diprotodontia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Diprotodontia}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_ERINACEOMORPHA.pdf}
+\caption{Distribution of available morphological data across Erinaceomorpha. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Erinaceomorpha}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_PILOSA.pdf}
+\caption{Distribution of available morphological data across Pilosa. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Pilosa}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_CETARTIODACTYLA.pdf}
+\caption{Distribution of available morphological data across Cetartiodactyla. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Primates}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_RODENTIA.pdf}
+\caption{Distribution of available morphological data across Rodentia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Rodentia}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_SCANDENTIA.pdf}
+\caption{Distribution of available morphological data across Scandentia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Scandentia}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_SORICOMORPHA.pdf}
+\caption{Distribution of available morphological data across Soricomorpha. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\label{Supp_Figure_Phylo-Soricomorpha}
+\end{figure}
 

### 2016-01-26 Manuscript/BiologyLetters/response.tex
Changed to table without character threshold
Comments after chat

@@
  for each order (and higher taxonomic level) was a good balance between the time spent looking for papers, the number of OTUs extracted and the repeatability of the study.
+% TG: try to modify in the paper or improve response.
 
@@
 
-%TO DO: not sure but maybe we'll just put the 1 character threshold thingy.
+We replaced the Table 1 in the main text with the Table 1 in the ESM 2 (available data without any number of characters threshold - see reviewer 2 comment 1).
+We therefore removed ESM 2 and renamed ESM 3 to ESM 2
 
@@
 
-\section{Reviewer 2 (Graham Slater):} % NC: I feel like some of the comments are missing here?
+\section{Reviewer 2 (Graham Slater):} % NC: I feel like some of the comments are missing here? TG: I've added the full comments now.
 \begin{enumerate}
 \item{\textcolor{blue}{The authors state on page 3 lines 82-90 that they avoided matrices with ``few characters'' due to lower probability of overlapping characters with other matrices and so only selected matrices containing >100 characters.
-[...] Is 100 really the minimum number you want to go with at all phylogenetic scales?}}
+First, a lack of overlapping characters doesn’t seem justifiable, at least to my mind, as a reason to exclude a matrix if the ultimate goal is to infer the placement of extinct taxa relative to extant taxa.
+Antler ramifications, to use the authors' example, are likely to be a very useful character when attempting to resolve relationships among living and extinct artiodactyls, even if they cannot be coded for monotremes.
+To flip the argument, a matrix of 200 + characters for mysticetes is similarly likely to contain many non-applicable characters from the perspective of talpids.
+I do understand that the authors are thinking downstream in terms of super matrices for higher level analyses, but smaller matrices are going to be just as relevant for reaching this goal.
+You don't have much space in the ms I understand, but surely this needs to be addressed?!}}
 
@@
 % NC: Yep that seems sensible
+We removed the analysis with a threshold of OTUs with at least 100 characters and replaced it with the analysis without any minimal number of morphological characters.
+We replaced <sentence> with <sentence new> in the manuscript; we updated the results of figure 1 and table 1 to contain all living OTUs; we removed the redundant ESM 2 and we updated the figures
+
+\item{\textcolor{blue}{Along these lines, a 100 character limit may seem rational but this is actually, from a practical perspective, very large.
+molecular matrices of this size would, of course, be laughed at but many of the most robust phylogenetic studies I'm familiar with come in well under this size.
+I suspect, but cannot be sure, that this partially explains the poor species level coverage the authors find; higher level studies can readily accumulate large numbers of characters because vastly different morphologies are being compared, often using families or genera as exemplars, leading to an abundance of easily defined characters.
+Species level studies, on the other, are often focused within families or genera to make them more manageable, but here characters are quickly exhausted.
+To give some random examples that I'm familiar with, Salles (1992) species-level study of Felidae (all species) was based on 44 characters, Wang and Carranza- Casta\~{n}eda's 2008 study of Mephitidae (\textasciitilde50\% of species but all genera) was based on 38 characters, and Bryant et al's 1993 study of Mustelid genera was of 46 characters.
+These are all important studies, but do not feature here because of the 100 character limit.
+At the very least I think this results in a downplaying of previous phylogenetic efforts by morphologists in the results.
+At worst it misrepresents the state of affairs dramatically.
+Is 100 really the minimum number you want to go with at all phylogenetic scales?}}
+
+%Check why the matrices didn't came up! - Don't justify not doing it by saying it's hard!
+
+% Justification: maybe data non accessible.
 
@@
 
-% NC: Let me know if you need help with this.
-%TG: It'll probably be all right :)
+%TG: just fix/add all this in this document
+%TG: Make sure word limit is kept to 2500 +/- 5%
 

### 2016-01-26 Manuscript/BiologyLetters/Supplementary/ESM_2.tex
Added sampling effort analysis (need explanations!)

@@
 
+\section{Supplementary results - 1}
+The following section contains supplementary blalbalbal....
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_sampling_effort.pdf}
+\caption{Sampling effort across all mammalian orders on the three different taxonomic levels. The \textit{R$^2$} and the slope (\textit{b}) are reported for each linear model.}
+\label{Supp_Figure_sampling}
+\end{figure}
+
+\newpage
+
 \section{Supplementary results - 2}

### 2016-01-26 Analysis/Sampling_effort.R

@@ -71,6 +71,9 @@ if(any(is.infinite(sampled_species))) {sampled_species[which(is.infinite(sampled
 total_species <- log(as.numeric(sapply(strsplit(results[which(results[,2] == "species"),3], split="/"), "[", 2)))
 
 
+
+pdf("../Manuscript/Supplementary/Supp_figure_sampling_effort.pdf", width = 8, height = 3.6)
+#quartz(width = 8, height = 3.6)
 par(mfrow = (c(1,3)), bty = "n")
 
 total <- total_families
@@ -79,9 +82,9 @@ model <- summary(lm(sampled + 0 ~ total))
 plot(total, sampled, pch=19, col="grey", xlim = c(0,max(total)), ylim = c(0,max(total)),
     ylab = "sampled families (log)", xlab = "total families (log)")
 abline(model$coefficients[[1]], model$coefficients[[2]])
-text(max(total) - 0.9*max(total), max(total) - 0.1*max(total),
+text(max(total) - 0.85*max(total), max(total) - 0.1*max(total),
     bquote(italic(R)^2 == .(format(model$r.squared, digits = 3))) )
-text(max(total) - 0.9*max(total), max(total) - 0.15*max(total),
+text(max(total) - 0.85*max(total), max(total) - 0.15*max(total),
     bquote(italic(b) == .(format(model$coefficients[[2]], digits = 3))))
 
 total <- total_genera
@@ -90,9 +93,9 @@ model <- summary(lm(sampled + 0 ~ total))
 plot(total, sampled, pch=19, col="grey", xlim = c(0,max(total)), ylim = c(0,max(total)),
     ylab = "sampled genera (log)", xlab = "total genera (log)")
 abline(model$coefficients[[1]], model$coefficients[[2]])
-text(max(total) - 0.9*max(total), max(total) - 0.1*max(total),
+text(max(total) - 0.85*max(total), max(total) - 0.1*max(total),
     bquote(italic(R)^2 == .(format(model$r.squared, digits = 3))) )
-text(max(total) - 0.9*max(total), max(total) - 0.15*max(total),
+text(max(total) - 0.85*max(total), max(total) - 0.15*max(total),
     bquote(italic(b) == .(format(model$coefficients[[2]], digits = 3))))
 
 
@@ -102,7 +105,8 @@ model <- summary(lm(sampled + 0 ~ total))
 plot(total, sampled, pch=19, col="grey", xlim = c(0,max(total)), ylim = c(0,max(total)),
[…]

### 2016-01-29 Manuscript/BiologyLetters/response.tex
Changed anatomical characters semantics

@@
 \item{\textcolor{blue}{My biggest criticism is a semantic one: Morphological data $\neq$ cladistic data.}}
+We agree with this reviewers semantinc point and changed ``cladistic'' and ``morphological'' data to ``anatomical characters'' throughout the manuscript.
+
 \item{\textcolor{blue}{The real issue here is the availability of phylogenetic character coding for anatomical data.
 [...] Thus, the title really should be ``Assessment of available anatomical characters for phylogenetic analysis among living mammals'' or something like that.}}
-
-%NC: Change the title?
+We changed the title as the reviewer suggested.
 
@@
 
+%TO DO
 % NC: worth commenting here that we are looking at the living species?
@@
 \item{\textcolor{blue}{Finally, I would add: who the heck submits a paper for a special volume 6 weeks early?!?!? You make the rest of us look bad!}}
-This paper formed part of my PhD thesis that I submitted in September which partially explains the swift submission! % NC: Worth making a tiny joke but no need to show off :P
-Sorry for making the rest of the authors submitting to the issue look bad :).
+This paper formed part of my PhD thesis that I submitted in September which partially explains the swift submission!
+Sorry for making the rest of the authors submitting to the issue look bad.
 

### 2016-01-29 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Updated email
Changed anatomical characters semantics

--- 2016-01-21 Manuscript/BiologyLetters/Missing_mammals-draft.tex
@@
 
-\noindent{\Large \bf Assessment of cladistic data availability for living mammals}
+\noindent{\Large \bf Assessment of available anatomical characters for phylogenetic analysis among living mammals}%Assessment of cladistic data availability for living mammals
 
@@
 The Total Evidence method allows living and fossil taxa to be combined in phylogenies, by using molecular data for living taxa and morphological data for both living and fossil taxa.
-With this method, substantial overlap of morphological data among living and fossil taxa is crucial for accurately inferring topology.
+With this method, substantial overlap of anatomical characters among living and fossil taxa is crucial for accurately inferring topology.
 However, although molecular data for living species is widely available, scientists using and generating morphological data mainly focus on fossils.
-Therefore, there is a gap in our knowledge of neontological morphological data even in well-studied groups such as mammals.
+Therefore, there is a gap in our knowledge of anatomical characters available for living taxa even in well-studied groups such as mammals.
 
-We investigated the amount of morphological (cladistic) data available for living mammals and how this data was phylogenetically distributed across orders.
-22 of 28 mammalian orders have \textless 25\% species with available morphological data; this has implications for the accurate placement of fossil taxa, although the issue is less pronounced at higher taxonomic levels.
+We investigated the amount of available anatomical characters for living mammals and how this data was phylogenetically distributed across orders.
+22 of 28 mammalian orders have \textless 25\% species with available anatomical characters; this has implications for the accurate placement of fossil taxa, although the issue is less pronounced at higher taxonomic levels.
 In most orders, species with available data are randomly distributed across the phylogeny, which may reduce the impact of the problem.
@@
 
-\noindent Key words: Total Evidence method, phylogenetic clustering, morphological matrix, extinct, topology\\
+\noindent Key words: Total Evidence method, phylogenetic clustering, cladistic matrix, extinct, topology\\
 % NC: Might need some new ones as these are in the abstract now...
@@
 To perform such analyses it is necessary to combine living and fossil taxa in phylogenetic trees.
-One increasingly popular method, the Total Evidence method \cite{ronquista2012}, combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix (e.g.
+One increasingly popular method, the Total Evidence method \cite{ronquista2012}, combines molecular data from living taxa and morphological data %TG: I left "morphological data" here since it TEM it doesn't have to be only anatomical characters
+ from both living and fossil taxa in a supermatrix (e.g.
 \cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a phylogeny with living and fossil taxa at the tips.
-A downside of this method is that it requires molecular data for living taxa and morphological data for both living and fossil taxa.
-Parts%rev 1.2.1 (as in reviewer 1, part 2 (minor), comment 1)
- of this data can be difficult, or impossible, to collect for every taxon in the analysis.
-For example, fossils rarely have molecular data and incomplete fossil preservation may restrict the amount of morphological data available.
-Additionally, it has become less common to collect morphological characters for living taxa when molecular data is available (e.g.
-in \cite{slaterphylogenetic2013}, only 13\% of living taxa have coded morphological data).
+A downside of this method is that it requires molecular data for living taxa and morphological data for both living and fossil taxa. %TG: same comment as above
+Sections of this data can be difficult, or impossible, to collect for every taxon in the analysis.
+For example, fossils rarely have molecular data and incomplete fossil preservation may restrict the amount of anatomical characters available.
+Additionally, it has become less common to collect anatomical characters for living taxa when molecular data is available (e.g.
+in \cite{slaterphylogenetic2013}, only 13\% of living taxa have coded anatomical characters).
 Unfortunately this missing data can lead to errors in phylogenetic inference.
-Simulations show that the ability of the Total Evidence method to recover the correct topology decreases when there is little overlap between morphological data in living and fossil taxa, and that the effect of missing data on topology is greatest when living taxa have few morphological data \cite{GuillermeCooper}.
-This is because (1) fossils will not be placed accurately within the correct clade% rev 1.2.2
- if it contains no morphological data for living taxa; and (2) fossils have a higher probability of branching within clades with more morphological data for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.
+Simulations show that the ability of the Total Evidence method to recover the correct topology decreases when there is little overlap between anatomical characters in living and fossil taxa, and that the effect of missing data on topology is greatest when living taxa have few anatomical characters available \cite{GuillermeCooper}.
+This is because (1) fossils will not be placed accurately within the correct clade if it contains no anatomical characters for living taxa; and (2) fossils have a higher probability of branching within clades with more anatomical characters available for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.
 
-The issues above highlight that it is crucial to have sufficient morphological data for living taxa in a clade before using a Total Evidence approach.
-However, it is unclear how much morphological data for living taxa is actually available, i.e.
+The issues above highlight that it is crucial to have sufficient anatomical characters available for living taxa in a clade before using a Total Evidence approach.
+However, it is unclear how much anatomical characters are actually available for living taxa, i.e.
 already coded from museum specimens and deposited in phylogenetic matrices accessible online, and how this data is distributed across clades.
+% TG: emphasize the "availibility" definition her?
 Intuitively, most people assume this kind of data has already been collected, but empirical data suggest otherwise (e.g. in \cite{ronquista2012,slaterphylogenetic2013,beckancient2014}).
-To investigate this further, we assess the amount of available morphological data for living mammals to determine whether sufficient data exists to build reliable Total Evidence phylogenies in this group.
-We also determine whether the available cladistic data is phylogenetically overdispersed or clustered across mammalian orders.
+To investigate this further, we assess the amount of available anatomical characters for living mammals to determine whether sufficient data exists to build reliable Total Evidence phylogenies in this group.
+We also determine whether the available anatomical characters are phylogenetically overdispersed or clustered across mammalian orders.
 
@@
 \subsection{Data collection and standardisation}
-We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{http://www.morphobank.org/} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{https://github.com/rossmounce/cladistic-data}).
+We downloaded all cladistic matrices %TG: I left "cladistic" here 'cause that's what the matrix is (even folloowing Peter's definition)!
+ containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{morphobank.org} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}).
 We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material (ESM) for a detailed description of the search procedure).
 In total, we downloaded 286 matrices containing 5228 unique operational taxonomic units (OTUs). 
-We used OTUs rather than species because% rev 1.2.3
- entries in the matrices ranged from species to families, and standardised the taxonomy as described in the ESM. 
+We used OTUs rather than species because entries in the matrices ranged from species to families, and standardised the taxonomy as described in the ESM. 
 We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}.
@@
 \subsection{Data availability and distribution}
-To assess the availability of cladistic data for each mammalian order, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
-We consider orders with \textless 25\% of living taxa with cladistic data as having low data coverage, and orders with \textgreater 75\% of living taxa with cladistic data as having high data coverage.
+To assess the availability of anatomical characters for each mammalian order, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
+%TG: changed cladistic datat to anatomical characters
+We consider orders with \textless 25\% of living taxa with available anatomical characters as having low data coverage, and orders with \textgreater 75\% of living taxa with available anatomical characters as having high data coverage.
 
-We investigated whether the available cladistic data for each order was (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index (NTI; \cite{webb2002phylogenies}) and the Net Relatedness Index (NRI; \cite{webb2002phylogenies}).
+For each order, we investigated whether the available anatomical characters were (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index (NTI; \cite{webb2002phylogenies}) and the Net Relatedness Index (NRI; \cite{webb2002phylogenies}).
 NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to it across the whole phylogeny \cite{Cooper2008}. 
@@
   \end{equation}
-where $\overline{MNND}_{obs}$ is the observed mean distance between each of $n$ taxa with cladistic data and its nearest neighbour with cladistic data in the phylogeny, 
+where $\overline{MNND}_{obs}$ is the observed mean branch length between each of $n$ taxa with available anatomical characters and its nearest neighbour with available anatomical characters in the phylogeny,
 $\overline{MNND}_{n}$ is the mean of 1000 $MNND$ between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random $MNND$ values.
@@
   \end{equation}
-where $\overline{MPD}_{obs}$ is the observed mean phylogenetic distance of the tree containing only the $n$ taxa with cladistic data.
+where $\overline{MPD}_{obs}$ is the observed mean phylogenetic branch length of the tree containing only the $n$ taxa with available anatomical characters.
 Negative NTI and NRI values show that the focal taxa are more overdispersed across the phylogeny than expected by chance, and positive values reflect clustering.
@@
 We calculated NTI and NRI values for each mammalian order separately, at each different taxonomic level. 
-For each analysis our focal taxa were those with available cladistic data at that taxonomic level and the phylogeny was that of the order pruned from \cite{BinindaEmonds}.
+For each analysis our focal taxa were those with available anatomical characters at that taxonomic level and the phylogeny was that of the order pruned from \cite{BinindaEmonds}.
 
@@
 \section{Results}
-22 of 28 orders have low coverage (\textless 25\% species with cladistic data) and six have high coverage (\textgreater 75\% species with cladistic data) at the species-level.
+22 of 28 orders % Change that!
+ have low coverage (\textless 25\% species with available anatomical characters) and six have high coverage %Change that!
+  (\textgreater 75\% species with available anatomical characters) at the species-level.
 At the genus-level, three orders have low coverage and 12 have high coverage, and at the family-level, no orders have low coverage and 23 have high coverage (Table\ref{Table_results}).
-
-\begin{figure}[!htbp]
-\centering
-    \includegraphics[width=1\textwidth]{example_coverage.pdf}
-\caption{Dummy figure.}
-\label{Figure_example_coverage}
-\end{figure}
 
@@
 \centering
-    \includegraphics[width=1\textwidth]{example_coverage.pdf}
-\caption{Phylogenetic distribution of species with available cladistic data across two orders (A: Primates; B: Carnivora).
-Blue branches indicate available cladistic data for the species.}
+    \includegraphics[width=1\textwidth]{figure1.pdf}
+\caption{Phylogenetic distribution of species with available anatomical characters across two orders (A: Primates; B: Carnivora).
+Blue branches indicate available anatomical characters for the species.}
 \label{Figure_example_coverage}
@@
 Our results show that although phylogenetic relationships among living mammals are well-resolved (e.g.
-\cite{BinindaEmonds,meredithimpacts2011}), most of the data used to build these phylogenies is molecular, and few cladistic data are available for living mammals compared to fossils (e.g.
+\cite{BinindaEmonds,meredithimpacts2011}), most of the data used to build these phylogenies is molecular, and few anatomical characters are available for living mammals compared to fossils (e.g.
 \cite{O'Leary08022013,ni2013oldest}).
-This has implications for building Total Evidence phylogenies containing both living and fossil mammals, as without sufficient cladistic data for living species, fossil placements in these trees are very uncertain \cite{GuillermeCooper}.
+This has implications for building Total Evidence phylogenies containing both living and fossil mammals, as without sufficient available anatomical characters for living species, fossil placements in these trees are very uncertain \cite{GuillermeCooper}.
 
-The number of living mammalian OTUs% rev 1.2.6
- with no available cladistic data was surprisingly high at the species-level: only six out of 28 orders have a high coverage of taxa with available cladistic data.
-This high coverage threshold of 75\% of taxa with available cladistic data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
+The number of living mammalian OTUs with no available anatomical characters was surprisingly high at the species-level: only six out of 28 orders have a high coverage of taxa with available anatomical characters.
+This high coverage threshold of 75\% of taxa with available anatomical characters represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
 Beyond this threshold, there is considerable displacement of wildcard taxa and decreased clade conservation \cite{GuillermeCooper}.
@@
 
-When few species have available cladistic data, the ideal scenario is for them to be phylogenetically overdispersed to maximize the possibilities of a fossil branching in the right clade.
-The second best scenario is that species with cladistic data are randomly distributed across the phylogenies. 
-Here we expect no special bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with cladistic data were randomly distributed across the phylogeny.
-The worst case scenario for fossil placement is that species with cladistic data are phylogenetically clustered. 
+When few species have available anatomical characters, the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion) to maximize the possibilities of a fossil branching in the right clade.
+The second best scenario is that species with available anatomical characters are randomly distributed across the phylogenies. 
+Here we expect no special bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with available anatomical characters were randomly distributed across the phylogeny.
+The worst case scenario for fossil placement is that species with available anatomical characters are phylogenetically clustered. 
 Then we expect two major biases to occur: first, fossils will not be able to branch within a clade containing no data, and second, fossils will have higher probability of branching within the most sampled clade by chance.
@@
 
-Despite the absence of good cladistic data coverage for living mammals, the Total Evidence method still seems to be the most promising way of combining living and fossil data for macroevolutionary analyses.
-Following the recommendations in \cite{GuillermeCooper}, we need to code cladistic characters for as many living species possible. 
-Fortunately, data for living mammals is usually readily available in natural history collections, therefore, we propose that an increased effort be put into coding morphological characters from living species, possibly by engaging in collaborative data collection projects.
+Despite the absence of good cladistic data coverage for living mammals, %TG: I think leaving cladistic here is OK no?
+ the Total Evidence method still seems to be the most promising way of combining living and fossil data for macroevolutionary analyses. 
+Following the recommendations in \cite{GuillermeCooper}, we need to code anatomical characters for as many living species possible. 
+Fortunately, data for living mammals is usually readily available in natural history collections, therefore, we propose that an increased effort be put into coding anatomical characters from living species, possibly by engaging in collaborative data collection projects.
 Such an effort would be valuable not only to phylogeneticists, but also to any researcher focusing understanding macroevolutionary patterns and processes.
@@
 Thanks to one anonymous reviewer, Graham Slater, Peter Wagner and Graeme Lloyd for their useful comments.
-%That's two times Graeme though...
+%TG: That's two times Graeme though...
 \section{Funding statement}

### 2016-02-04 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Couple of comments updated

@@
   \end{equation}
-where $\overline{MNND}_{obs}$ is the observed mean branch length between each of $n$ taxa with available anatomical characters and its nearest neighbour with available anatomical characters in the phylogeny,
+where $\overline{MNND}_{obs}$ is the observed mean sum of the branch lengths between each of $n$ taxa with available anatomical characters and its nearest neighbour with available anatomical characters in the phylogeny,
 $\overline{MNND}_{n}$ is the mean of 1000 $MNND$ between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random $MNND$ values.
@@
 While neontological species are described using morphology, genes, distribution etc.; palaeontological species can be based only on morphological, spatial and temporal data (e.g. \cite{ni2013oldest}).
-Therefore, most palaeontological studies use the genus as their smallest OTU (e.g.
+Therefore, many palaeontological studies use the genus (or monospecific genera) as their smallest OTU (e.g.
 \cite{ni2013oldest,O'Leary08022013}), so data availability at the genus-level in living mammals should be our primary concern when building phylogenies of living and fossil taxa.

### 2016-02-05 Manuscript/BiologyLetters/response.tex
Couple of comments updated
Updated results in text (no threshold)

@@
 \bigskip
-We are very grateful to the four referees for their helpful and constructive comments.
-%, which we believe have helped us to significantly improve our paper.
-% NC: A bit too much for what are minor revisions :)
+We are very grateful to the four referees for their helpful and constructive comments.s
 We have taken all of their comments on board, and respond to their points below.
@@
 \end{enumerate}
-
-% NC: You replace something *with* something else not *by*. Things can be replaced by something, but the sentence set up is then "X was replaced by Y", not "replaced X by Y".
-
 \subsection{Reviewer 1 specific comments:}
@@
 
-\section{Reviewer 2 (Graham Slater):} % NC: I feel like some of the comments are missing here? TG: I've added the full comments now.
+\section{Reviewer 2 (Graham Slater):}
 \begin{enumerate}
@@
 We removed the analysis with a threshold of OTUs with at least 100 characters and replaced it with the analysis without any minimal number of morphological characters.
-We replaced <sentence> with <sentence new> in the manuscript; we updated the results of figure 1 and table 1 to contain all living OTUs; we removed the redundant ESM 2 and we updated the figures
+We retracted the paragraph aout the character threshold analysis and updated the results of figure 1 and table 1 to contain all living OTUs; we removed the redundant ESM 2 and we updated the figures
 
@@
 \item{\textcolor{blue}{Line 104 - It's not completely clear to me how ``distance'' is measured. Is this branch-lengths on a time-scaled tree? Or just a node count?}}
-We have clarified this by changing ``observed mean distance between each of $n$ taxa'' to ``observed mean sum of the branch lengths between each of $n$ taxa''. % NC: Fix in document
+We have clarified this by changing ``observed mean distance between each of $n$ taxa'' to ``observed mean sum of the branch lengths between each of $n$ taxa''. % NC: Fix in document %TG:done
 
@@
 % NC: Just change it from "most" to "many", or say paleo studies commonly feature only one species in a genus.
+We modified this statement by stating that ``many palaeontological studies use the genus (or monospecific genera) as their smallest OTU''.
 

### 2016-02-05 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Updated results in text (no threshold)

@@
 \subsection{Data collection and standardisation}
-We downloaded all cladistic matrices %TG: I left "cladistic" here 'cause that's what the matrix is (even folloowing Peter's definition)!
+We downloaded all cladistic matrices %TG: I left "cladistic" here 'cause that's what the matrix is (even following Peter's definition)!
  containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{morphobank.org} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}).
@@
 
-Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having overlapping characters with other matrices \cite{wagner2000} and (2) they are more likely to contain specific characters that are not applicable% rev 1.2.4
- across large clades (e.g. ``antler ramifications'' is a character that is only applicable to Cervidae \cite{Brazeau2011}).
-Therefore we selected only matrices containing \textgreater 100 characters for each OTU.
-This threshold was chosen to correspond with the number of characters used in \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
-Results of analyses with no threshold are available in the ESM. 
-After removing matrices with \textless 100 characters, we retained 1074 unique living OTUs from 126 matrices. % 1601 unique living OTUs for 286 matrices (no threshold)
+% Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having overlapping characters with other matrices \cite{wagner2000} and (2) they are more likely to contain specific characters that are not applicable% rev 1.2.4
+%  across large clades (e.g. ``antler ramifications'' is a character that is only applicable to Cervidae \cite{Brazeau2011}).
+% Therefore we selected only matrices containing \textgreater 100 characters for each OTU.
+% This threshold was chosen to correspond with the number of characters used in \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
+% Results of analyses with no threshold are available in the ESM. 
+% After removing matrices with \textless 100 characters, we retained 1074 unique living OTUs from 126 matrices. % 1601 unique living OTUs for 286 matrices (no threshold)
+We retained 1601 unique living OTUs from 286 matrices.
 
@@
 \section{Results}
-22 of 28 orders % Change that!
- have low coverage (\textless 25\% species with available anatomical characters) and six have high coverage %Change that!
-  (\textgreater 75\% species with available anatomical characters) at the species-level.
-At the genus-level, three orders have low coverage and 12 have high coverage, and at the family-level, no orders have low coverage and 23 have high coverage (Table\ref{Table_results}).
+11 out of 28 orders have low coverage (\textless 25\% species with available anatomical characters) and 7 have high coverage (\textgreater 75\% species with available anatomical characters) at the species-level.
+At the genus-level, one order have low coverage and 15 have high coverage, and at the family-level, no orders have low coverage and 25 have high coverage (Table\ref{Table_results}).
+
+%TG: former results:
+% 22 of 28 orders have low coverage (\textless 25\% species with available anatomical characters) and six have high coverage (\textgreater 75\% species with available anatomical characters) at the species-level.
+% At the genus-level, three orders have low coverage and 12 have high coverage, and at the family-level, no orders have low coverage and 23 have high coverage (Table\ref{Table_results}).
 
@@
 
-Only six orders had significantly clustered data (Afrosoricida and Pholidota at the species-level, and Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha at both species- and genus-level) and none had significantly overdispersed data (Table~\ref{Table_results}).
+Only seven orders had significantly clustered data (Cetartiodactyla, Cingulata, Pilosa and Rodentia at the species-level and Carnivora, Chiroptera and Soricomorpha at both species- and genus-level) and none had significantly overdispersed data (Table~\ref{Table_results}).
 
-Figure \ref{Figure_example_coverage} shows randomly distributed OTUs with cladistic data in Primates (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered OTUs with cladistic data in Carnivora (mainly Canidae; Figure \ref{Figure_example_coverage}B).
+Figure \ref{Figure_example_coverage} shows randomly distributed OTUs with cladistic data in Primates (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered OTUs with cladistic data in Carnivora (mainly Canidae and Urisdae but no Herpestidae; Figure \ref{Figure_example_coverage}B).
 
@@
 
-The number of living mammalian OTUs with no available anatomical characters was surprisingly high at the species-level: only six out of 28 orders have a high coverage of taxa with available anatomical characters.
+The number of living mammalian OTUs with no available anatomical characters was surprisingly high at the species-level: only seven out of 28 orders have a high coverage of taxa with available anatomical characters.
 This high coverage threshold of 75\% of taxa with available anatomical characters represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
@@
 Then we expect two major biases to occur: first, fossils will not be able to branch within a clade containing no data, and second, fossils will have higher probability of branching within the most sampled clade by chance.
-Our results suggest that this may be problematic at the genus-level in Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha. 
+Our results suggest that this may be problematic at the genus-level in Carnivora, Chiroptera and Soricomorpha. 
 For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade% rev 1.2.7
-, and will have more chance to randomly branch within Canidae (Figure \ref{Figure_example_coverage}B).
+, and will have more chance to randomly branch within Canidae or Ursidae (Figure \ref{Figure_example_coverage}B).
 

### 2016-02-05 Manuscript/BiologyLetters/response.tex
Added data collection avert in discussion

@@
 
-%Check why the matrices didn't came up! - Don't justify not doing it by saying it's hard!
+By removing the characters threshold (see above) we now included the taxa coded in Wang and Carranza-Casta\~{n}eda 2008 (matrix reference GS2014-WCC) and other matrices ignored in the former analysis increasing the number of Carnivora from 42 species (14\% of total number of species) to 76 (27\%).
+However, the two other matrices mentionned by reviewer 2 (Salles 1992 and Bryant et al. 1993) where not included in the present analysis.
+This is due to two main reasons: firstly, these two matrices where not available on any of the matrices repositories (either Graeme Lloyd's, Ross Mounce's or MorphoBank's) and (2) secondly, they didn't matched with out additional Google Scholars search parameters (i.e.
+matrices published between 2010 and 2015 with accessible data).
+Following this reviewer's comment as well as reviewer 4's we added a paragraph in the discussion to clearly state that we are aware of data sampling biases:
 
-% Justification: maybe data non accessible.
+``It is worth noting, however, that our analysis did not include all the matrices containing anatomical characters ever published.
+In fact, our data collection procedure focused on including studies that provided matrices easily accessible, i.e.
+we did specifially not include any matrices that were only available in paper format (e.g.
+printed in books), non-reusable format (e.g.
+an image of the matrix) %TG: yes...
+There were quiet a few of them!
+or/and matrices available only upon request (e.g. by emailing the authors).
+Such matrices are likely to not be included in Total Evidence analysis due to the difficulty of obtaining them.
+'' p.@ lines@@.
 
@@
 
+We dealt with this comment by removing the matrix size threshold and adding a paragraph to the discussion explaining that some matrices might not be included in this study (see reviewer's 2 comments 1 and 2).
+Concerning the Gaubert et al. 2005's study specificaly, this study unfortunately did not contain any easily available matrix.
+In fact, the matrix mentioned as Appendix 2 in the paper is not present in the article and neither the TreeBase accession numbers (``S1255'', ``M2193'') nor the provided URL (``www.systematicbiology.org'') links to the matrix.
+% TG: I don't know if that's a good justification but basically, it's impossible to get this specific matrix.
+
 \item{\textcolor{blue}{One interesting aside that the authors didn't consider, despite the data being there, is how phylogenetic ``effort'' relates to diversity of the order under study.}}
 
-This is a great suggestion so we have added this analysis to the ESM @@@% get number.
-% We show that mammalian morphological phylogeneticists have focused their energies on resolving the higher level relationships among mammalian lineages.
+This is a great suggestion, thanks!
+We added this analysis to the ESM 2 (Supplementary results - 1) showing, as suggested, that mammalian morphological phylogeneticists have focused their energies on resolving the higher level relationships among mammalian lineages.
 

### 2016-02-05 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Added data collection avert in discussion

@@
 
+It is worth noting, however, that our analysis did not include all the matrices containing anatomical characters ever published.
+In fact, our data collection procedure focused on including studies that provided matrices easily accessible, i.e.
+we did specifially not include any matrices that were only available in paper format (e.g.
+printed in books), non-reusable format (e.g.
+an image of the matrix) %TG: yes...
+There were quiet a few of them!
+or/and matrices available only upon request (e.g. by emailing the authors).
+Such matrices are likely to not be included in Total Evidence analysis due to the difficulty of obtaining them.
+
 Despite the absence of good cladistic data coverage for living mammals, %TG: I think leaving cladistic here is OK no?

### 2016-02-05 Manuscript/BiologyLetters/response.tex
Added sampling effort analysis

@@
 By removing the characters threshold (see above) we now included the taxa coded in Wang and Carranza-Casta\~{n}eda 2008 (matrix reference GS2014-WCC) and other matrices ignored in the former analysis increasing the number of Carnivora from 42 species (14\% of total number of species) to 76 (27\%).
-However, the two other matrices mentionned by reviewer 2 (Salles 1992 and Bryant et al. 1993) where not included in the present analysis.
+However, the two other matrices mentioned by reviewer 2 (Salles 1992 and Bryant et al. 1993) where not included in the present analysis.
 This is due to two main reasons: firstly, these two matrices where not available on any of the matrices repositories (either Graeme Lloyd's, Ross Mounce's or MorphoBank's) and (2) secondly, they didn't matched with out additional Google Scholars search parameters (i.e.
@@
 In fact, our data collection procedure focused on including studies that provided matrices easily accessible, i.e.
-we did specifially not include any matrices that were only available in paper format (e.g.
+we did specifically not include any matrices that were only available in paper format (e.g.
 printed in books), non-reusable format (e.g.
@@
 We dealt with this comment by removing the matrix size threshold and adding a paragraph to the discussion explaining that some matrices might not be included in this study (see reviewer's 2 comments 1 and 2).
-Concerning the Gaubert et al. 2005's study specificaly, this study unfortunately did not contain any easily available matrix.
+Concerning the Gaubert et al. 2005's study specifically, this study unfortunately did not contain any easily available matrix.
 In fact, the matrix mentioned as Appendix 2 in the paper is not present in the article and neither the TreeBase accession numbers (``S1255'', ``M2193'') nor the provided URL (``www.systematicbiology.org'') links to the matrix.
@@
 We added this analysis to the ESM 2 (Supplementary results - 1) showing, as suggested, that mammalian morphological phylogeneticists have focused their energies on resolving the higher level relationships among mammalian lineages.
+We also added a sentence mentioning these results in the manuscript:
+
+``It is also encouraging to note that the sampling effort is consistent with diversity at each taxonomic level and there is therefore no direct bias in sampling at any taxonomic level (ESM 2, Figure 1).
+'' p.@ lines@@
 

### 2016-02-05 Manuscript/BiologyLetters/Supplementary/ESM_2.tex
Changed anatomical characters semantics
Added sampling effort analysis

@@
 
-\noindent{\Large \bf Assessment of cladistic data availability for living mammals}
+\noindent{\Large \bf Assessment of available anatomical characters for phylogenetic analysis among living mammals}
 
@@
 \section{Supplementary results - 1}
-The following section contains supplementary blalbalbal....
+The following section contains a supplementary analysis looking at the sampling effort for all the mammalian orders at the three different taxonomic levels (family, genus and species).
+We calculated linear regressions %TG: does one calculate regressions? I would go for drawing regressions but that's not semantically correct...
+between the logged number of sampled taxonomic levels per order against the logged total number of total taxonomic levels and fixed the intercept to 0.
+This way we could calculate the slope of the regressions (\textit{b}) indicating the sampling effort (i.e.
+if the slope is 1, the sampling effort is maximum along each order for the taxonomic level, if 0 there is no significant sampling) and the fit of the regressions (\textit{R$^2$}).
+The code for this analysis is available at this link \url{https://github.com/TGuillerme/Missing_living_mammals/blob/master/Analysis/Sampling_effort.R}.
 
@@
 \end{figure}
+
+This analysis shows suggests that the sampling effort reflects diversity for each taxonomic level (i.e.
+morphological characters were collected for each order according to it's size; Figure \ref{Supp_Figure_sampling}).
+However, it is worth noting that this relation is stronger at higher taxonomic level (i.e.
+family, \textit{R$^2$}=0.99 \textit{b}=0.936) in comparison to the species level (\textit{R$^2$}=0.852 \textit{b}=0.711).
+These results might be simply due to the increase of elements between each taxonomic levels (i.e. number of species $>$ genera $>$ families)
+Nonetheless it is probable that mammalian morphological phylogeneticists did actually focused on resolving higher level relationships among mammalian orders (see Table 1 in the main text).
 

### 2016-02-08 Manuscript/BiologyLetters/CoverLetter.tex
re committed files

--- 2015-11-24 Manuscript/BiologyLetters/CoverLetter.tex
@@
 
-Please find attached our research article, ``Assessment of cladistic data availability for living mammals'', for consideration in the Biology Letters Special Issue: % NC: What is the title?
+Please find attached our research article, ``Assessment of cladistic data availability for living mammals'', for consideration in the Biology Letters Special Issue: ``Putting fossils in trees''.
 
@@
 The Total Evidence method allows us to use molecular and morphological data to build phylogenies with both living and fossil species as tips (Ronquist et al. 2012 Syst Biol).
-%TG: it changes from here:
 This method is extremely promising because it allows us to use all the available data. 

### 2016-02-08 Manuscript/BiologyLetters/Missing_mammals-draft.tex
re committed files

@@
  the Total Evidence method still seems to be the most promising way of combining living and fossil data for macroevolutionary analyses.
+It is also encouraging to note that the sampling effort is consistent with diversity at each taxonomic level and there is therefore no direct bias in sampling at any taxonomic level (ESM 2, Figure 1).
 Following the recommendations in \cite{GuillermeCooper}, we need to code anatomical characters for as many living species possible. 

### 2016-02-08 Manuscript/Table_results_supp.tex
re committed files

--- 2016-01-26 Manuscript/BiologyLetters/Table_results.tex
@@
 % latex table generated in R 3.2.0 by xtable 1.7-4 package
-% Sat Jun 20 17:19:07 2015
+% Sat Jun 20 17:08:58 2015
 \begin{longtable}{lL{1.8cm}C{2cm}lcc}
-\caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels (without any character threshold; results from the 286 matrices).
+\caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels.
 The coverage represents the proportion of taxa with available morphological data.
@@
 Significant NRI or NTI are highlighted in bold.
-*p \textless 0.05; **p \textless 0.01; ***p \textless 0.001.} \\
+One star (*) represents a p-value between 0.05 and 0.005; two starts between 0.005 and 0.0005 and three stars a p-value less than 0.0005.} \\
   \hline
@@
 \hline
-\label{Table_results}
+\label{Table_results_supp}
 \end{longtable}

### 2016-02-08 Manuscript/proto-draft.tex
re committed files

@@
 The searched mammalian order terms are available in search_terms_latin.txt or search_terms_meta.txt.
-The file containing the meta names is the latin name files without the following suffixes: Mammal     @ia
-Monothrem     @ata
-Marsupial     @ia
-Placental     @ia
-Macroscelid     @ea
-Afrosoricid     @a
-Tubulident     @ata
-Hyracoid     @ea
-Proboscid     @ea
-Siren     @ia
-Pilos     @a
-Cingul     @ata
-Scandent     @ia
-Dermopter     @a
-Primat     @es
-Lagomorph     @a
-Rodent     @ia
-Erinaceomorph     @a
-Soricomorph     @a
-Cetac     @ea
-Artiodactyl     @a
-Chiropter     @a
-Perissodactyl     @a
-Pholidot     @a
-Carnivor     @a
-Didelphimorph     @ia
-Paucitubercul     @ata
-Microbiother     @ia
-Dasyuromorph     @ia
-Peramelemorph     @ia
-Notoryctemorph     @ia
-Diprotodont     @ia
-Monotrem     @ata
+The file containing the meta names is the Latin name files but with replacing the latin suffixes ([ia|ata|ea|a]) by a joker character (*) and by replacing the first letter by a upper/lower case meta character (e.g.
+[Aa]).
+
+\subsubsection{Ross Mounce data set}
+I selected all the matrices containing at least one of the mammalian orders names from Ross Mounce GitHub cladistic-data/nexus_files repository (accessed on the 02/12/2014).
 

### 2016-02-08 Manuscript/Table_results.tex
color testings

--- 2015-06-20 Manuscript/Table_results.tex
@@
-% latex table generated in R 3.2.0 by xtable 1.7-4 package
-% Sat Jun 20 17:41:32 2015
+% latex table generated in R 3.2.3 by xtable 1.8-0 package
+% Mon Feb  8 17:25:14 2016
 \begin{longtable}{lL{1.8cm}C{2cm}lcc}
@@
   Paucituberculata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar49.pdf} &   &   \\ 
-  Paucituberculata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar50.pdf} & 0 & 0 \\ 
-  Paucituberculata & species & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar51.pdf} & -0.64 & -0.65 \\ 
+  Paucituberculata & genus & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar50.pdf} &   &   \\ 
+  Paucituberculata & species & 5/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar51.pdf} &   &   \\ 
   Peramelemorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar52.pdf} &   &   \\ 
   Peramelemorphia & genus & 7/7 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar53.pdf} &   &   \\ 
-  Peramelemorphia & species & 16/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar54.pdf} & -0.09 & 1 \\ 
+  Peramelemorphia & species & 16/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar54.pdf} & -0.13 & 0.97 \\ 
   Perissodactyla & family & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar55.pdf} &   &   \\ 
   Perissodactyla & genus & 6/6 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar56.pdf} &   &   \\ 
-  Perissodactyla & species & 7/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar57.pdf} & 0.62 & -2.5 \\ 
+  Perissodactyla & species & 10/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar57.pdf} & -0.07 & -2.63 \\ 
   Pholidota & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar58.pdf} &   &   \\ 
   Pholidota & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar59.pdf} &   &   \\ 
-  \textbf{Pholidota} & \textbf{species} & \textbf{3/8} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar60.pdf} & \textbf{2.64*} & \textbf{2.23*} \\ 
-  Pilosa & family & 3/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar61.pdf} & 0.94 & 0.93 \\ 
-  Pilosa & genus & 3/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar62.pdf} & -0.36 & -0.31 \\ 
-  Pilosa & species & 3/29 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar63.pdf} & 0.33 & 0.79 \\ 
+  Pholidota & species & 4/8 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar60.pdf} & 1.18 & 0.94 \\ 
+  Pilosa & family & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar61.pdf} & 1.87 & 2 \\ 
+  Pilosa & genus & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar62.pdf} & -0.96 & 0.36 \\ 
+  \textbf{Pilosa} & \textbf{species} & \textbf{5/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar63.pdf} & \textbf{1.28} & \textbf{2.38**} \\ 
   Primates & family & 15/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar64.pdf} &   &   \\ 
-  Primates & genus & 48/68 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar65.pdf} & -0.41 & -1.4 \\ 
-  Primates & species & 56/351 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar66.pdf} & -1.6 & -2.04 \\ 
+  Primates & genus & 48/68 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar65.pdf} & -0.35 & -1.33 \\ 
+  Primates & species & 64/351 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar66.pdf} & -0.67 & -1.27 \\ 
   Proboscidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar67.pdf} &   &   \\ 
-  Proboscidea & genus & 1/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar68.pdf} &   &   \\ 
-  Proboscidea & species & 1/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar69.pdf} &   &   \\ 
-  Rodentia & family & 11/32 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar70.pdf} & -0.46 & -1.91 \\ 
-  Rodentia & genus & 21/450 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar71.pdf} & -2.11 & 0.3 \\ 
-  Rodentia & species & 15/2094 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar72.pdf} & -1.65 & -2.55 \\ 
+  Proboscidea & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar68.pdf} &   &   \\ 
+  Proboscidea & species & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar69.pdf} & -0.69 & -0.69 \\ 
+  Rodentia & family & 18/32 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar70.pdf} & 0.66 & -0.98 \\ 
+  Rodentia & genus & 82/450 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar71.pdf} & -1.66 & 1.55 \\ 
+  \textbf{Rodentia} & \textbf{species} & \textbf{90/2094} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar72.pdf} & \textbf{2.76**} & \textbf{2.34**} \\ 
   Scandentia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar73.pdf} &   &   \\ 
-  Scandentia & genus & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar74.pdf} & -0.77 & -0.76 \\ 
-  Scandentia & species & 2/20 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar75.pdf} & -1.79 & -1.99 \\ 
+  Scandentia & genus & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar74.pdf} & -0.74 & -0.74 \\ 
+  Scandentia & species & 3/20 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar75.pdf} & -1.88 & -0.84 \\ 
   Sirenia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar76.pdf} &   &   \\ 

### 2016-02-08 Analysis/Analysis_per_order.R
Updated analysis (now including all class)

@@ -38,13 +38,13 @@ extraction_table1<-extraction_table[which(as.numeric(extraction_table$Characters
 extraction_table100<-extraction_table[which(as.numeric(extraction_table$Characters) >= 100),]
 
 #Combine the sub tables
-extraction_list=list(extraction_table1, extraction_table100)
-results_names=c("results_1.Rda","results_100.Rda")
+extraction_list=list(extraction_table1)#, extraction_table100)
+results_names=c("results_1.Rda")#,"results_100.Rda")
 
 #Print the analysis by series of characters length
 
 #for (character_threshold in 1:length(extraction_list)) {
-for (character_threshold in 1:2) {
+for (character_threshold in 1:1) {
     #Select the number of characters
     extraction_table<-extraction_list[[character_threshold]]
 
@@ -58,7 +58,13 @@ for (character_threshold in 1:2) {
 
     #Calculate the structure of the data for the first order to initiate the loop
 
-    #First order
+    #Analysis for all mammals
+
+    message(paste("\n","Mammalia (Class)", " analysis.\n", sep=""), appendLF=FALSE)
+    tree<-extract.order("Mammalia (Class)", mam_tree, WR_list, verbose=TRUE)
+    taxa<-extract.order("Mammalia (Class)", living_taxa_list, WR_list, verbose=TRUE)
+
+
     #Extract orders
     message(paste("\n",orders[1], " analysis.\n", sep=""), appendLF=FALSE)
     tree<-extract.order(orders[1], mam_tree, WR_list, verbose=TRUE)
@@ -69,14 +75,14 @@ for (character_threshold in 1:2) {
     }
 
     #Calculate the data structure in the order
-    results<-order.structure(orders[1], taxa, tree, WR_list, metric=c("PD", "NTI", "NRI"), verbose=TRUE)
+    results<-order.structure("Mammalia (Class)", taxa, tree, WR_list, metric=c("NTI", "NRI"), verbose=TRUE)
 
 
     #Calculate the structure of the data for the other orders through a loop
 
 
     #Loop through the rest
-    for (order in 2:length(orders)) {
+    for (order in 1:length(orders)) {
         #verbose
         message(paste("\n",orders[order], " analysis.\n", sep=""), appendLF=FALSE)
 

### 2016-02-08 Analysis/Results_summary.R

@@ -67,7 +67,8 @@ load("results_1.Rda")
 #Fable with number of taxa with morpho data+data structure
 ############################################
 caption<-"Number of taxa with available cladistic data for mammalian orders at three taxonomic levels. The coverage represents the proportion of taxa with available morphological data. The left vertical bar represents 25\\% of available data (``low'' coverage if \\textless 25\\%); The right vertical bar represents 75\\% of available data (``high'' coverage if \\textgreater 75\\%). When the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) are negative, taxa are more phylogenetically dispersed than expected by chance; when NRI or NTI are positive, taxa are more phylogenetically clustered by expected by chance. Significant NRI or NTI are highlighted in bold. One star (*) represents a p-value between 0.05 and 0.005; two starts between 0.005 and 0.0005 and three stars a p-value less than 0.0005."
-table.result(results, metric=c("NRI","NTI"), threshold=c(25,75), save.path="../Manuscript/", file.save="Table_results", caption=caption, environement="longtable")
+#table.result(results, metric=c("NRI","NTI"), threshold=c(25,75), save.path="../Manuscript/", file.save="Table_results", caption=caption, environement="longtable", col.threshold=c("blue", "grey", "orange"))
+table.result(results, metric=c("NRI","NTI"), threshold=c(25,75), save.path="../Manuscript/", file.save="Table_results", caption=caption, environement="longtable", col.threshold=c("#2A2A2A","#7F7F7F","#D4D4D4")) #col.threshold=c("#3F3F3F","#7F7F7F","#BFBFBF"))
 
 #Bugged!
 #caption<-"Number of taxa with available cladistic data for mammalian orders at three taxonomic levels. The coverage represents the proportion of taxa with available morphological data. The left vertical bar represents 25\\% of available data (``low'' coverage if \\textless 25\\%); The right vertical bar represents 75\\% of available data (``high'' coverage if \\textgreater 75\\%). When the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) are negative, taxa are more phylogenetically dispersed than expected by chance; when NRI or NTI are positive, taxa are more phylogenetically clustered by expected by chance. Significant NRI or NTI are highlighted in bold. One star (*) represents a p-value between 0.05 and 0.005; two starts between 0.005 and 0.0005 and three stars a p-value less than 0.0005."

### 2016-02-08 Functions/extract.order.R

@@ -11,14 +11,67 @@
 
 extract.order<-function(order, taxa, reference, verbose=FALSE) {
     
+    #Functions
+    scan.reference.list <- function(taxon, Sub_reference, verbose) {
+        #Is the taxon in the Sub_reference
+        test<-which(Sub_reference == taxon)
+
+        if(length(test) != 0) {
+            #The taxon is in the Sub_refence
+            match_taxon<-TRUE
+        
+        } else {
+
+            #Is it a binomial name?
+            if(length(grep("_", taxon))==1) {
+                #Split the binomial name
+                binomial<-check.split(taxon, Sub_reference)
+
+                #No match
+                if(all(is.na(binomial))) {
+                    match_taxon<-FALSE
+                }
+
+                #Full match
+                if(all(!is.na(binomial))) {
+                    match_taxon<-TRUE
+                }
+
+                #Partial match
+                if(!is.na(binomial[1]) & is.na(binomial[2])) {
+                    #Only genera name matches
+                    taxon_change <- c(taxon, binomial[1])
+                    warning(paste("\n", taxon, " did not match but ", binomial[1], " did.", sep=""))
+                    taxa_change <<- rbind(taxa_change, taxon_change)
+
+                    #Taxa didn't match!
+                    match_taxon <- FALSE
+                }
+
+            } else {
+                #Name is not matching nor binomial
+                match_taxon<-FALSE
+            }
+        }
+
+
+        # Output
+        if(verbose == TRUE) {message(".", appendLF=FALSE)}
+        #if(verbose == TRUE) {message(paste(taxon, ".", sep=""), appendLF=FALSE)}
+        return(match_taxon)
+    }
+
+
     #SANTIZING
     #order
     check.class(order, "character", " must be a the name of the order to extract.")
 
     #reference
     check.class(reference, "data.frame", " must be a taxonomic reference list.")
-    if(length(which(reference == order)) == 0) {
-        stop("Order not found in provided reference list.")
+    if(order != "Mammalia (Class)") {
+        if(length(which(reference == order)) == 0) {
+            stop("Order not found in provided reference list.")
+        }
     }
 
     #taxa
@@ -35,9 +88,17 @@ extract.order<-function(order, taxa, reference, verbose=FALSE) {
     #verbose
     check.class(verbose, "logical", " must be logical.")
 
+    #Short cut for tree
[…]

### 2016-02-08 Functions/order.structure.R

@@ -14,7 +14,7 @@
 #guillert(at)tcd.ie - 04/02/2015
 ##########################
 
-order.structure<-function(order, taxa, tree, reference, metric=c("PD", "NRI", "NTI"), runs=1000, null="taxa.labels", verbose=FALSE, ...) {
+order.structure<-function(order, taxa, tree, reference, metric=c("NRI", "NTI"), runs=1000, null="taxa.labels", verbose=FALSE, ...) {
     #LIBRARIES
     require(picante)
 
@@ -62,7 +62,11 @@ order.structure<-function(order, taxa, tree, reference, metric=c("PD", "NRI", "N
 
     #Sort the list of taxa per species/genera/family
     #Reference list for the order
-    Sub_reference<-subset(reference[which(reference == order),])
+    if(order == "Mammalia (Class)") {
+        Sub_reference <- reference
+    } else {
+        Sub_reference<-subset(reference[which(reference == order),])
+    }
 
     taxa_binomial_non<-taxa.binomial(taxa, Sub_reference)
     taxa_binomial<-taxa_binomial_non[[1]]

### 2016-02-08 Functions/table.results.R

@@ -8,11 +8,12 @@
 #<file.save> name (or chain name) for saving the table
 #<caption> table caption (if split option is called, the caption will only appear for the first table)
 #<environement> table environement
+#<col.threshold> a vector of colours that must be the length of threshold + 1
 ##########################
 #environement(at)tcd.ie - 05/06/2015
 ##########################
 
-table.result<-function(results, metric, threshold=c(25,75), save.path, file.save, caption, environement) {
+table.result<-function(results, metric, threshold=c(25,75), save.path, file.save, caption, environement, col.threshold=c("grey", "grey", "grey")) {
     
     #DEBUG
     #warning("Debug mode is ON (table.result)")
@@ -64,13 +65,13 @@ table.result<-function(results, metric, threshold=c(25,75), save.path, file.save
             metrics[one_star]<-paste(metrics[one_star], "*", sep="")
         }
         #Two stars
-        two_star<-which(p_values <= 0.005)
+        two_star<-which(p_values <= 0.01)
         significant_storage<-c(significant_storage, two_star)
         if(length(one_star) != 0) {
             metrics[two_star]<-paste(metrics[two_star], "*", sep="")
         }
         #Three stars
-        three_star<-which(p_values <= 0.0005)
+        three_star<-which(p_values <= 0.0001)
         significant_storage<-c(significant_storage, three_star)
         if(length(three_star) != 0) {
             metrics[three_star]<-paste(metrics[three_star], "*", sep="")
@@ -109,7 +110,32 @@ table.result<-function(results, metric, threshold=c(25,75), save.path, file.save
         system(paste("mkdir", new_folder))
     }
 
+
+
+    #~~~~~~~~~~~~~~~~
     #Make sure the percentage column is numeric
+    #~~~~~~~~~~~~~~~~
+
+    #Colour threshold selector
+    # WARNING: develop this function to match any threshold number!
+    col.selector <- function(threshold_value, threshold, col.threshold) {
+        if(length(threshold) > 2) {
+            stop("col.selector internal function only works with two threshold values for now.")
+        }
+        #threshold is greater than any threshold?
+        if(threshold_value >= max(threshold)) {
+            return(col.threshold[3])
+        } else {
+            #threshold is smaller than any threshold?
+            if(threshold_value <= min(threshold)) {
+                return(col.threshold[1])
+            } else {
[…]

### 2016-02-08 Analysis/Analysis_per_order.R
Updated analysis (full)

@@ -59,17 +59,10 @@ for (character_threshold in 1:1) {
     #Calculate the structure of the data for the first order to initiate the loop
 
     #Analysis for all mammals
-
     message(paste("\n","Mammalia (Class)", " analysis.\n", sep=""), appendLF=FALSE)
     tree<-extract.order("Mammalia (Class)", mam_tree, WR_list, verbose=TRUE)
     taxa<-extract.order("Mammalia (Class)", living_taxa_list, WR_list, verbose=TRUE)
 
-
-    #Extract orders
-    message(paste("\n",orders[1], " analysis.\n", sep=""), appendLF=FALSE)
-    tree<-extract.order(orders[1], mam_tree, WR_list, verbose=TRUE)
-    taxa<-extract.order(orders[1], living_taxa_list, WR_list, verbose=TRUE)
-    #Replace taxa by a vector if there was any taxonomic correction
     if(class(taxa) != "character") {
         taxa<-taxa$Taxa
     }
@@ -77,9 +70,7 @@ for (character_threshold in 1:1) {
     #Calculate the data structure in the order
     results<-order.structure("Mammalia (Class)", taxa, tree, WR_list, metric=c("NTI", "NRI"), verbose=TRUE)
 
-
     #Calculate the structure of the data for the other orders through a loop
-
 
     #Loop through the rest
     for (order in 1:length(orders)) {

### 2016-02-08 Manuscript/BiologyLetters/Table_results.tex
Updated results in text (no threshold)
added colour and black and white table figures

@@
-% latex table generated in R 3.2.0 by xtable 1.7-4 package
-% Sat Jun 20 17:19:07 2015
+% latex table generated in R 3.2.3 by xtable 1.8-0 package
+% Mon Feb  8 22:09:15 2016
 \begin{longtable}{lL{1.8cm}C{2cm}lcc}
-\caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels (without any character threshold; results from the 286 matrices).
+\caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels.
 The coverage represents the proportion of taxa with available morphological data.
@@
   Afrosoricida & genus & 17/17 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar2.pdf} &   &   \\ 
-  Afrosoricida & species & 23/42 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar3.pdf} & 1.75 & 1.08 \\ 
-  Carnivora & family & 14/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar4.pdf} & 0.63 & 0.6 \\ 
-  \textbf{Carnivora} & \textbf{genus} & \textbf{54/125} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar5.pdf} & \textbf{4.81**} & \textbf{1.78*} \\ 
-  \textbf{Carnivora} & \textbf{species} & \textbf{76/283} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar6.pdf} & \textbf{7.66**} & \textbf{0.85} \\ 
+  Afrosoricida & species & 23/42 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar3.pdf} & 1.52 & 1.1 \\ 
+  Carnivora & family & 14/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar4.pdf} & 0.65 & 0.55 \\ 
+  \textbf{Carnivora} & \textbf{genus} & \textbf{52/125} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar5.pdf} & \textbf{4.27**} & \textbf{1.26} \\ 
+  \textbf{Carnivora} & \textbf{species} & \textbf{75/283} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar6.pdf} & \textbf{7.24**} & \textbf{0.8} \\ 
   Cetartiodactyla & family & 21/21 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar7.pdf} &   &   \\ 
-  Cetartiodactyla & genus & 100/128 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar8.pdf} & 0.85 & 0.94 \\ 
-  \textbf{Cetartiodactyla} & \textbf{species} & \textbf{171/310} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar9.pdf} & \textbf{1.92*} & \textbf{-0.46} \\ 
-  Chiroptera & family & 15/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar10.pdf} & -0.28 & 0.56 \\ 
-  \textbf{Chiroptera} & \textbf{genus} & \textbf{93/202} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar11.pdf} & \textbf{13.47**} & \textbf{1.1} \\ 
-  \textbf{Chiroptera} & \textbf{species} & \textbf{215/1053} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar12.pdf} & \textbf{8.82**} & \textbf{1.22} \\ 
+  Cetartiodactyla & genus & 97/128 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar8.pdf} & 0.7 & 1.28 \\ 
+  \textbf{Cetartiodactyla} & \textbf{species} & \textbf{169/310} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar9.pdf} & \textbf{1.82*} & \textbf{-0.24} \\ 
+  Chiroptera & family & 15/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar10.pdf} & -0.23 & 0.61 \\ 
+  \textbf{Chiroptera} & \textbf{genus} & \textbf{92/202} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar11.pdf} & \textbf{13.07**} & \textbf{0.99} \\ 
+  \textbf{Chiroptera} & \textbf{species} & \textbf{214/1053} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar12.pdf} & \textbf{9.21**} & \textbf{1.27} \\ 
   Cingulata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar13.pdf} &   &   \\ 
-  Cingulata & genus & 8/9 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar14.pdf} & 1.51 & -1.57 \\ 
-  \textbf{Cingulata} & \textbf{species} & \textbf{9/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar15.pdf} & \textbf{1.9*} & \textbf{0.11} \\ 
+  Cingulata & genus & 8/9 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar14.pdf} & 1.48 & -1.54 \\ 
+  \textbf{Cingulata} & \textbf{species} & \textbf{9/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar15.pdf} & \textbf{2.06*} & \textbf{0.2} \\ 
   Dasyuromorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar16.pdf} &   &   \\ 
-  Dasyuromorphia & genus & 8/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar17.pdf} & -0.75 & -1.07 \\ 
-  Dasyuromorphia & species & 9/64 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar18.pdf} & -0.88 & -0.34 \\ 
+  Dasyuromorphia & genus & 8/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar17.pdf} & -0.78 & -1.06 \\ 
+  Dasyuromorphia & species & 9/64 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar18.pdf} & -0.86 & -0.37 \\ 
   Dermoptera & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar19.pdf} &   &   \\ 
@@
   Didelphimorphia & genus & 16/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar23.pdf} &   &   \\ 
-  Didelphimorphia & species & 42/84 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar24.pdf} & -1.65 & 0.2 \\ 
+  Didelphimorphia & species & 42/84 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar24.pdf} & -1.61 & 0.12 \\ 
   Diprotodontia & family & 11/11 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar25.pdf} &   &   \\ 
-  Diprotodontia & genus & 25/38 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar26.pdf} & -1.13 & -1.31 \\ 
-  Diprotodontia & species & 31/126 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar27.pdf} & 0.48 & -1.77 \\ 
+  Diprotodontia & genus & 25/38 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar26.pdf} & -1.15 & -1.33 \\ 
+  Diprotodontia & species & 31/126 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar27.pdf} & 0.44 & -1.79 \\ 
   Erinaceomorpha & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar28.pdf} &   &   \\ 
   Erinaceomorpha & genus & 10/10 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar29.pdf} &   &   \\ 
-  Erinaceomorpha & species & 21/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar30.pdf} & -1.07 & -0.2 \\ 
+  Erinaceomorpha & species & 21/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar30.pdf} & -1.04 & -0.25 \\ 
   Hyracoidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar31.pdf} &   &   \\ 
@@
   Lagomorpha & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar34.pdf} &   &   \\ 
-  Lagomorpha & genus & 5/12 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar35.pdf} & -1.06 & -0.95 \\ 
-  Lagomorpha & species & 12/86 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar36.pdf} & -0.62 & -1.88 \\ 
+  Lagomorpha & genus & 5/12 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar35.pdf} & -0.95 & -0.94 \\ 
+  Lagomorpha & species & 12/86 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar36.pdf} & -0.62 & -1.96 \\ 
   Macroscelidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar37.pdf} &   &   \\ 
   Macroscelidea & genus & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar38.pdf} &   &   \\ 
-  Macroscelidea & species & 12/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar39.pdf} & -1.3 & -1.06 \\ 
-  Microbiotheria & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar40.pdf} &   &   \\ 
-  Microbiotheria & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar41.pdf} &   &   \\ 
-  Microbiotheria & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar42.pdf} &   &   \\ 
-  Monotremata & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar43.pdf} &   &   \\ 
-  Monotremata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar44.pdf} & -0.72 & -0.69 \\ 
-  Monotremata & species & 2/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar45.pdf} & -0.97 & -0.97 \\ 
-  Notoryctemorphia & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar46.pdf} &   &   \\ 
-  Notoryctemorphia & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar47.pdf} &   &   \\ 
-  Notoryctemorphia & species & 0/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar48.pdf} &   &   \\ 
-  Paucituberculata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar49.pdf} &   &   \\ 
-  Paucituberculata & genus & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar50.pdf} &   &   \\ 
-  Paucituberculata & species & 5/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar51.pdf} &   &   \\ 
-  Peramelemorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar52.pdf} &   &   \\ 
-  Peramelemorphia & genus & 7/7 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar53.pdf} &   &   \\ 
-  Peramelemorphia & species & 16/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar54.pdf} & -0.13 & 0.97 \\ 
-  Perissodactyla & family & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar55.pdf} &   &   \\ 
-  Perissodactyla & genus & 6/6 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar56.pdf} &   &   \\ 
-  Perissodactyla & species & 10/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar57.pdf} & -0.07 & -2.63 \\ 
-  Pholidota & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar58.pdf} &   &   \\ 
-  Pholidota & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar59.pdf} &   &   \\ 
-  Pholidota & species & 4/8 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar60.pdf} & 1.18 & 0.94 \\ 
-  Pilosa & family & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar61.pdf} & 1.87 & 2 \\ 
-  Pilosa & genus & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar62.pdf} & -0.96 & 0.36 \\ 
-  \textbf{Pilosa} & \textbf{species} & \textbf{5/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar63.pdf} & \textbf{1.28} & \textbf{2.38*} \\ 
-  Primates & family & 15/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar64.pdf} &   &   \\ 
-  Primates & genus & 48/68 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar65.pdf} & -0.35 & -1.33 \\ 
-  Primates & species & 64/351 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar66.pdf} & -0.67 & -1.27 \\ 
-  Proboscidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar67.pdf} &   &   \\ 
-  Proboscidea & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar68.pdf} &   &   \\ 
-  Proboscidea & species & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar69.pdf} & -0.69 & -0.69 \\ 
-  Rodentia & family & 18/32 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar70.pdf} & 0.66 & -0.98 \\ 
-  Rodentia & genus & 82/450 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar71.pdf} & -1.66 & 1.55 \\ 
-  \textbf{Rodentia} & \textbf{species} & \textbf{90/2094} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar72.pdf} & \textbf{2.76*} & \textbf{2.34*} \\ 
-  Scandentia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar73.pdf} &   &   \\ 
-  Scandentia & genus & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar74.pdf} & -0.74 & -0.74 \\ 
-  Scandentia & species & 3/20 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar75.pdf} & -1.88 & -0.84 \\ 
-  Sirenia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar76.pdf} &   &   \\ 
-  Sirenia & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar77.pdf} &   &   \\ 
-  Sirenia & species & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar78.pdf} &   &   \\ 
-  Soricomorpha & family & 3/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar79.pdf} & -0.98 & -0.99 \\ 
-  \textbf{Soricomorpha} & \textbf{genus} & \textbf{19/43} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar80.pdf} & \textbf{7.11**} & \textbf{2.59**} \\ 
-  \textbf{Soricomorpha} & \textbf{species} & \textbf{21/392} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar81.pdf} & \textbf{10.65**} & \textbf{3.56**} \\ 
-  Tubulidentata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar82.pdf} &   &   \\ 
-  Tubulidentata & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar83.pdf} &   &   \\ 
-  Tubulidentata & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar84.pdf} &   &   \\ 
+  Macroscelidea & species & 12/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar39.pdf} & -1.24 & -1.2 \\ 
+  Mammalia (class) & family & 129/148 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar40.pdf} & -1.19 & 1.09 \\ 
+  \textbf{Mammalia (class)} & \textbf{genus} & \textbf{517/1186} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar41.pdf} & \textbf{-5.19} & \textbf{3.71**} \\ 
+  \textbf{Mammalia (class)} & \textbf{species} & \textbf{847/5017} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar42.pdf} & \textbf{-7.75} & \textbf{3.54**} \\ 
+  Microbiotheria & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar43.pdf} &   &   \\ 
+  Microbiotheria & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar44.pdf} &   &   \\ 
+  Microbiotheria & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar45.pdf} &   &   \\ 
+  Monotremata & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar46.pdf} &   &   \\ 
+  Monotremata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar47.pdf} & -0.68 & -0.69 \\ 
+  Monotremata & species & 2/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar48.pdf} & -1.01 & -1 \\ 
+  Notoryctemorphia & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar49.pdf} &   &   \\ 
+  Notoryctemorphia & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar50.pdf} &   &   \\ 
+  Notoryctemorphia & species & 0/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar51.pdf} &   &   \\ 
+  Paucituberculata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar52.pdf} &   &   \\ 
+  Paucituberculata & genus & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar53.pdf} &   &   \\ 
+  Paucituberculata & species & 5/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar54.pdf} &   &   \\ 
+  Peramelemorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar55.pdf} &   &   \\ 
+  Peramelemorphia & genus & 7/7 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar56.pdf} &   &   \\ 
+  Peramelemorphia & species & 16/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar57.pdf} & -0.14 & 0.91 \\ 
+  Perissodactyla & family & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar58.pdf} &   &   \\ 
+  Perissodactyla & genus & 6/6 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar59.pdf} &   &   \\ 
+  Perissodactyla & species & 10/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar60.pdf} & -0.1 & -2.77 \\ 
+  Pholidota & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar61.pdf} &   &   \\ 
+  Pholidota & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar62.pdf} &   &   \\ 
+  Pholidota & species & 4/8 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar63.pdf} & 1.14 & 0.97 \\ 
+  Pilosa & family & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar64.pdf} & 2.01 & 1.96 \\ 
+  Pilosa & genus & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar65.pdf} & -0.91 & 0.36 \\ 
+  \textbf{Pilosa} & \textbf{species} & \textbf{5/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar66.pdf} & \textbf{1.18} & \textbf{2.35**} \\ 
+  Primates & family & 15/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar67.pdf} &   &   \\ 
+  Primates & genus & 48/68 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar68.pdf} & -0.37 & -1.39 \\ 
+  Primates & species & 64/351 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar69.pdf} & -0.66 & -1.4 \\ 
+  Proboscidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar70.pdf} &   &   \\ 
+  Proboscidea & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar71.pdf} &   &   \\ 
+  Proboscidea & species & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar72.pdf} & -0.67 & -0.72 \\ 
+  Rodentia & family & 18/32 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar73.pdf} & 0.66 & -0.95 \\ 
+  \textbf{Rodentia} & \textbf{genus} & \textbf{82/450} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar74.pdf} & \textbf{-1.81} & \textbf{1.7*} \\ 
+  \textbf{Rodentia} & \textbf{species} & \textbf{90/2094} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar75.pdf} & \textbf{2.66**} & \textbf{2.36**} \\ 
+  Scandentia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar76.pdf} &   &   \\ 
+  Scandentia & genus & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar77.pdf} & -0.77 & -0.76 \\ 
+  Scandentia & species & 3/20 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar78.pdf} & -2 & -0.8 \\ 
+  Sirenia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar79.pdf} &   &   \\ 
+  Sirenia & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar80.pdf} &   &   \\ 
+  Sirenia & species & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar81.pdf} &   &   \\ 
+  Soricomorpha & family & 3/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar82.pdf} & -0.98 & -0.97 \\ 
+  \textbf{Soricomorpha} & \textbf{genus} & \textbf{19/43} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar83.pdf} & \textbf{7.07**} & \textbf{2.64**} \\ 
+  \textbf{Soricomorpha} & \textbf{species} & \textbf{21/392} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar84.pdf} & \textbf{10.17**} & \textbf{3.36**} \\ 
+  Tubulidentata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar85.pdf} &   &   \\ 
+  Tubulidentata & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar86.pdf} &   &   \\ 
+  Tubulidentata & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar87.pdf} &   &   \\ 
    \hline

### 2016-02-08 Analysis/Results_summary.R

@@ -67,8 +67,8 @@ load("results_1.Rda")
 #Fable with number of taxa with morpho data+data structure
 ############################################
 caption<-"Number of taxa with available cladistic data for mammalian orders at three taxonomic levels. The coverage represents the proportion of taxa with available morphological data. The left vertical bar represents 25\\% of available data (``low'' coverage if \\textless 25\\%); The right vertical bar represents 75\\% of available data (``high'' coverage if \\textgreater 75\\%). When the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) are negative, taxa are more phylogenetically dispersed than expected by chance; when NRI or NTI are positive, taxa are more phylogenetically clustered by expected by chance. Significant NRI or NTI are highlighted in bold. One star (*) represents a p-value between 0.05 and 0.005; two starts between 0.005 and 0.0005 and three stars a p-value less than 0.0005."
-#table.result(results, metric=c("NRI","NTI"), threshold=c(25,75), save.path="../Manuscript/", file.save="Table_results", caption=caption, environement="longtable", col.threshold=c("blue", "grey", "orange"))
-table.result(results, metric=c("NRI","NTI"), threshold=c(25,75), save.path="../Manuscript/", file.save="Table_results", caption=caption, environement="longtable", col.threshold=c("#2A2A2A","#7F7F7F","#D4D4D4")) #col.threshold=c("#3F3F3F","#7F7F7F","#BFBFBF"))
+table.result(results, metric=c("NRI","NTI"), threshold=c(25,75), save.path="../Manuscript/", file.save="Table_results", caption=caption, environement="longtable", col.threshold=c("blue", "grey", "orange"))
+#table.result(results, metric=c("NRI","NTI"), threshold=c(25,75), save.path="../Manuscript/", file.save="Table_results", caption=caption, environement="longtable", col.threshold=c("#2A2A2A","#7F7F7F","#D4D4D4")) #col.threshold=c("#3F3F3F","#7F7F7F","#BFBFBF"))
 
 #Bugged!
[…]

### 2016-02-08 Manuscript/BiologyLetters/Table_results.tex
Updated caption

@@
 \begin{longtable}{lL{1.8cm}C{2cm}lcc}
-\caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels.
-The coverage represents the proportion of taxa with available morphological data.
-The left vertical bar represents 25\% of available data (``low'' coverage if \textless 25\%); The right vertical bar represents 75\% of available data (``high'' coverage if \textgreater 75\%).
-When the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) are negative, taxa are more phylogenetically dispersed than expected by chance; when NRI or NTI are positive, taxa are more phylogenetically clustered by expected by chance.
-Significant NRI or NTI are highlighted in bold.
-One star (*) represents a p-value between 0.05 and 0.005; two starts between 0.005 and 0.0005 and three stars a p-value less than 0.0005.} \\
+\caption{Number of taxa with available cladistic data for mammalian orders at three
+taxonomic levels.
+The left vertical bar represents low coverage (<25\%; coloured in blue); the right vertical bar represents high coverage (>75\%; coloured in orange).
+Negative Net Relatedness Index (NRI) and Nearest Taxon Index (NTI) values indicate phylogenetic overdispersion; positive values indicate phylogenetic clustering.
+Significant NRI or NTI values are in bold.
+*p <0.05; **p <0.01; ***p <0.001.
+} \\ 
   \hline
@@
   \hline
+  Mammalia (class) & family & 129/148 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar40.pdf} & -1.19 & 1.09 \\ 
+  \textbf{Mammalia (class)} & \textbf{genus} & \textbf{517/1186} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar41.pdf} & \textbf{-5.19} & \textbf{3.71**} \\ 
+  \textbf{Mammalia (class)} & \textbf{species} & \textbf{847/5017} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar42.pdf} & \textbf{-7.75} & \textbf{3.54**} \\ 
   Afrosoricida & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar1.pdf} &   &   \\ 
@@
   Macroscelidea & species & 12/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar39.pdf} & -1.24 & -1.2 \\ 
-  Mammalia (class) & family & 129/148 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar40.pdf} & -1.19 & 1.09 \\ 
-  \textbf{Mammalia (class)} & \textbf{genus} & \textbf{517/1186} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar41.pdf} & \textbf{-5.19} & \textbf{3.71**} \\ 
-  \textbf{Mammalia (class)} & \textbf{species} & \textbf{847/5017} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar42.pdf} & \textbf{-7.75} & \textbf{3.54**} \\ 
   Microbiotheria & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar43.pdf} &   &   \\ 

### 2016-02-09 Manuscript/BiologyLetters/response.tex
Updated response (finished)

@@
  for each order (and higher taxonomic level) was a good balance between the time spent looking for papers, the number of OTUs extracted and the repeatability of the study.
-% TG: try to modify in the paper or improve response.
+We are aware however that we might have missed some specific studies (see Reviewer 2 comment 2) but are confident that our protocol gives a good overall picture of the data availability and that the three repositories used before the Google Scholar search (MorphoBank, Graeme Lloyd's and Ross Mounce's) contained most of the available data.
+In fact, as mentioned in ESM 1, part 1 ``among the 660 papers [collected from the Google Scholar search], only 50 contained a total of 425 extra living OTUs''.
 
@@
 
-%TO DO! Same answer as above?
+We chose this cutoff date to filter our analysis towards accessible data only (see Reviewer 2 comment 2).
+Additionally, we observed that in newer studies (i.e.
+post 2010), living taxa with anatomical data were rarely added an usually reused from former studies (with sometimes the addition of extra characters).
+We modified the sentence in the ESM1 part 1:
+
+``For example in primates the character p7 coded first by [1] is reused with the same living species in [2], [3], [4], [5], [5], [6], [7], [8], [9], [10], [11], [12] and [13].''
+
+\noindent to: ``For example, the six living primates used in [1] (\textit{Aotus trivirgatus, Galago demidoff, Lemur catta, Microcebus murinus, Nycticebus coucang and Saimiri sciureus}) and their associated characters are reused along with more living species and characters in [2], [3], [4], [5], [5], [6], [7], [8], [9], [10], [11], [12] and [13].''.
 
@@
 
-% TO DO.
-To the question, hard to give a value for the probability of getting the ``right'' tree but 25% missing living leads to 0.5 wild card taxa displacement and between 0.8 and 0.4 clade conservation on a linear scale from 1 to 0 with one being the right tree and 0 being a random tree.
+The straight relationship between the living species coverage and the Total Evidence phylogeny topology is hard to quantify precisely and depends on both the metric used for measuring topological differences and the method used to infer the tree.
+Our simulation showed that topology tend to get closer to a random one as morphological data is missing but that in the case of the missing living species, the other parameters we tested (missing data in the fossil record and number of morphological characters) add no significant effect (i.e.
+when a certain amount of living taxa are missing, the other parameters do not have any extra influence).
+In our simulations we tested 5 different level of missing living taxa (0\%, 10\%, 25\%, 50\% and 75\% of missing living taxa) and we observed that the three first levels had no significant negative effect on topology in relation with the amount of missing data in the fossil record and the number of morphological characters but that the fifth level (75\% of missing living taxa, i.e.
+25\% coverage) had a strong negative effect.
+
+Also, we highlighted the different levels of coverage in Table 1 as follow: at any taxonomic level, orders with less than 25\% coverage are highlighted in blue (or dark grey in the black and white version) and orders with more than 75\% coverage are highlithed in orange (or light grey in the black and white version).
 
@@
 
-%TO DO: let's just remove this threshold thingy since I seem to be the only one to be convinced by it ;)...
-% NC: Yep that seems sensible
 We removed the analysis with a threshold of OTUs with at least 100 characters and replaced it with the analysis without any minimal number of morphological characters.
-We retracted the paragraph aout the character threshold analysis and updated the results of figure 1 and table 1 to contain all living OTUs; we removed the redundant ESM 2 and we updated the figures
+We retracted the paragraph about the character threshold analysis and updated the results of figure 1 and table 1 to contain all living OTUs; we removed the redundant ESM 2 and we updated the figures
 
@@
 
-%TO DO: OK, I think there is a problem here.
-Either we have to rerun the matrix downloading procedure and be sure we include the ones specified by the reviewers (the correct way!) but that WILL take more than a week.
-Or else we just add the matrices that are suggested y the reviewers but I have the feeling that that is a bit of cheating (i.e.
-we're hand picking some data here) - basically that'll just add more carnivora.
-%I think one argument could be to emphasize on one of the points we add in a former version of the manuscript (can't find it): "by available data we mean data in a standard format that can be directly downloaded from the internet (i.e.
-many of the papers we looked at had no link to their data or the link was broken, meaning that the data is effectively only available upon specific request to the authors [baaaad!!!])."
-
-% NC: Hmmm. I'm concerned about this, and I think your "quick and dirty" solution is pretty dishonest. I think you're going to need to look at your data collection procedure again. 
-% First we need to understand *why* is the paper Graham mentions missing?
-Because you focus on data from 2010 onwards?
-Because it's not in Graeme or Ross' databases?
-Need to understand how it got missed out.
-Does it appear if you get rid of the 10/20 search results only and search lots of results?
-Is there a problem with search terms?
-The answer to this will help us decide how to proceed from here.
-% NC: Although you don't want to waste a huge amount of time on this, you also don't want a reputation for being sloppy, nor a reputation for glossing over your errors.
-Also it's important for personal research integrity.
-If that means you spend two weeks recollecting the data then that will have to be accepted.
-Note that two weeks is really a very short amount of time in the usual 2 years it takes to write a paper (Martin will be expecting you to have brought some work over from your PhD), but you will need to ask for an extension on the corrections.
-This won't be a problem.
-Either way this comment will have to explain how the papers Graham mentions are missed out.
-If you can do that and it's sensible then explaining the latter point about locating the data is probably fine.
-And I know you're busy - welcome to the world post PhD!
-All things happen at once...
-
 We dealt with this comment by removing the matrix size threshold and adding a paragraph to the discussion explaining that some matrices might not be included in this study (see reviewer's 2 comments 1 and 2).
@@
 
-% NC: May also want to add something to the main text.
 \end{enumerate}
@@
 
-%TO DO
-% NC: worth commenting here that we are looking at the living species?
+We totally agree with this comment.
+In fact, among the 286 matrices used in this analysis, we found 5228 unique OTUs including 3627 fossil ones which is pretty impressive knowing the amount of work required to collect data for a single specimen!
+However, in this study, we focused only on the amount of living taxa (1601 unique OTUs containing 847 species) as a way to solve the problems associated with the lack of living taxa on tree topology described in Guillerme \& Cooper 2016.
 
@@
 
+We thank the reviewer for this mention and we will be sure to use this database for future studies.
 % NC: If you end up needing to redo the data collection it might be worth checking the PBDB...
@@
 
-% NC: Yep he's right, modify the sentence to account for this. Also cite a long branch attraction stuff. Check if Pete has written anything - it'd be nice to cite him as he's been so helpful.
+We do agree with this comment and changed this comment to:
+``For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade due to the fact that they have no anatomical characters available (i.e.
+extant Herpestidae have a zero-length branches for anatomical characters).
+Such fossil, however, will have a higher probability to be placed along a branch that contains many anatomical characters (i.e.
+long branches for anatomical characters, e.g.
+Canidae or Ursidae (Figure 1B).''
+
+% NC: Yep he's right, modify the sentence to account for this.
+Also cite a long branch attraction stuff.
+Check if Pete has written anything - it'd be nice to cite him as he's been so helpful.
+% TG: I looked for relevant paper on this "short-branch repulsion" but didn't found any that's mildly relevant from Peter...
 
@@
 
-% NC: This is clearly a get out clause to get this done more rapidly.
+We developed description of our data collection protocol in the ESM 1 (see Reviewer 1 comment 2) and underlined the caveats in the discussion (see Reviewer 2 comment 2).
 
@@
 
-% NC: OK so Graeme is going with your "data accessibility" idea here to get round having to redo the data collection.
-I think this will be ok as long as you add comments to that effect in the text, and work out why Graham's papers are missing.
-
-% NC: Get rid of the 100 characters thing (I never really understood why it was needed), and make sure all the results are definitely correct.
+We removed the character threshold in the whole analysis and specified what we meant by ``data accessibility'' in the discussion (see Reviewer 2 comment 2).
 
@@
  graphs are a nice potential addition to the ESM.}}
+
+We added the data coverage among all mammals in the Table 1 (``Mammalia (Class)'') and added it to the methods, results and discussion.
+We also added the Reviewer 2's extra analysis suggestion to the ESM 2 part 1.
+
 \end{enumerate}
-
-% NC: Easy enough to add a line to all the tables for total coverage
 
@@
 \item{\textcolor{blue}{Line 104 - It's not completely clear to me how ``distance'' is measured. Is this branch-lengths on a time-scaled tree? Or just a node count?}}
-We have clarified this by changing ``observed mean distance between each of $n$ taxa'' to ``observed mean sum of the branch lengths between each of $n$ taxa''. % NC: Fix in document %TG:done
+We have clarified this by changing ``observed mean distance between each of $n$ taxa'' to ``observed mean sum of the branch lengths between each of $n$ taxa''.
 
@@
 Are you using the node subtending the higher taxon or something else?}}
-%TO CHECK!
-% NC: Probably you've pruned the phylogeny to include only species within each order/family?
+We pruned Bininda-Emonds et al's phylogeny for each order in order to include only the species within this order using the the functions:
+
+\texttt{nodeX <- ape::getMRCA(treeAllMammals, listOfSpeciesFromOrderX)}\\
+\texttt{treeOrderX <- ape::drop.tip(treeAllMammals, nodeX)}\\
+
+We added the following sentence to in the methods:
+``For each analysis our focal taxa were those with available anatomical characters at that taxonomic level and the phylogeny was that of the order pruned from [10].
+For each order, we pruned the phylogeny from [10] to include only the species/genus/families from this order.''.
+p@.
+lines@@.
 
@@
 (I didn't check ref 21.)}}
-%To CHECK, do we actually care? 
-% NC: Just change it from "most" to "many", or say paleo studies commonly feature only one species in a genus.
+
 We modified this statement by stating that ``many palaeontological studies use the genus (or monospecific genera) as their smallest OTU''.
@@
 \item{\textcolor{blue}{Line 154 - ``phylogenetically overdispersed'' Purely semantic, but I guess what you mean is that even sampling would be ideal, and this is *measured* using phylogenetic over-dispersion.}}
+
 We changed ``the ideal scenario is for them to be phylogenetically overdispersed'' to ``the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion)''.
@@
 \item{\textcolor{blue}{Line 245 - The caption could be more informative regarding the random and clustered distributions that these figures represent.}}
+
 We changed the caption to: ``Phylogenetic distribution of species with available cladistic data across two orders (A: Primates; B: Carnivora).
@@
 \item{\textcolor{blue}{ESM 1 - Incomplete sentence: ``The list of all the 286 download matrices is available on''}}
-we fixed this typo. See reviewer 1 specific comment 12.
+
+We fixed this typo. See reviewer 1 specific comment 12.
 
 \item{\textcolor{blue}{ESM 1 - Incomplete sentence: ``All the standardised matrices are available on''}}
+
 We fixed this typo. See reviewer 1 specific comment 12.
@@
 (Obviously more generally living species can have fossil records, or at least plenty of fossils are assigned to extant species.)}}
-%TO DO: change ``fossil'' to ``non living OTU''? As in OTU that is not present in the Wilson Reeder list or the Bininda Emonds tree?
+
+We do agree with this point (in a similar way: \textit{Thylacinus cynocephalus} is present in Wilson Reader's list but can not be designated as ``living'') and thus added a precision to what we mean by ``fossil'' OTU in the ESM 1: ``i.e.
+OTUs not present in the two previous data sets [Wilson Reeder and Binida-Emonds]''.
 
@@
 Or just species you can't reconcile with your three data sets?}}
-%TO DO: add the number of ignored OTUs. I think they're mainly specimen IDs without species names.
+
+1549 of the 11010 OTUs were ignored.
+We added the following sentence to the ESM 1 to explain which OTUs were considered as NAs.
+``Non-applicable OTUs were either specimen IDs with no related taxa (e.g.
+\textit{FMNHPR2081}), abbreviation that were not described in the associated paper (e.g.
+\textit{Ho.sap.}), non-mammals \textit{stricto-sensu} (e.g.
+\textit{Sinoconodon}), non standard taxonomic level (e.g.
+\textit{Spalcotheriids}) invalid taxonomic designation (e.g.
+\textit{sp\_nov\_1} or \textit{Outgroup}) or typos (e.g.
+\textit{Hobo sabpiens}).''
 
@@
 We changed this to ``using the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) metrics''.
+
 \end{enumerate}

### 2016-02-09 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Updated response (finished)

@@
 \subsection{Data availability and distribution}
-To assess the availability of anatomical characters for each mammalian order, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
+To assess the availability of anatomical characters for each mammalian order and for Mammalia as a class, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
 %TG: changed cladistic datat to anatomical characters
@@
 
-For each order, we investigated whether the available anatomical characters were (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index (NTI; \cite{webb2002phylogenies}) and the Net Relatedness Index (NRI; \cite{webb2002phylogenies}).
+For each order and for Mammalia as a class, we investigated whether the available anatomical characters were (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index (NTI; \cite{webb2002phylogenies}) and the Net Relatedness Index (NRI; \cite{webb2002phylogenies}).
 NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to it across the whole phylogeny \cite{Cooper2008}. 
@@
 \section{Results}
-11 out of 28 orders have low coverage (\textless 25\% species with available anatomical characters) and 7 have high coverage (\textgreater 75\% species with available anatomical characters) at the species-level.
+Among all Mammalia, the species coverage was low (\textless 25\% species with available anatomical characters) but the family coverage was high (\textgreater 75\% species with available anatomical characters).
+For each order, 11 out of 28 have low coverage and 7 have high coverage at the species-level.
 At the genus-level, one order have low coverage and 15 have high coverage, and at the family-level, no orders have low coverage and 25 have high coverage (Table\ref{Table_results}).
@@
 
-Only seven orders had significantly clustered data (Cetartiodactyla, Cingulata, Pilosa and Rodentia at the species-level and Carnivora, Chiroptera and Soricomorpha at both species- and genus-level) and none had significantly overdispersed data (Table~\ref{Table_results}).
+Among all Mammalia, the data was significantly cluster when using the NTI at the species and genus level.
+For each order, only seven had significantly clustered data (Cetartiodactyla, Cingulata, Pilosa and Rodentia at the species-level and Carnivora, Chiroptera and Soricomorpha at both species- and genus-level) and none had significantly overdispersed data (Table~\ref{Table_results}).
 
@@
 
-The number of living mammalian OTUs with no available anatomical characters was surprisingly high at the species-level: only seven out of 28 orders have a high coverage of taxa with available anatomical characters.
+The number of living mammalian OTUs with available anatomical characters was surprisingly low at the species-level: only 17\%.
+Only seven out of 28 orders have a high coverage of taxa with available anatomical characters.
 This high coverage threshold of 75\% of taxa with available anatomical characters represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
@@
 Our results suggest that this may be problematic at the genus-level in Carnivora, Chiroptera and Soricomorpha. 
-For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade% rev 1.2.7
-, and will have more chance to randomly branch within Canidae or Ursidae (Figure \ref{Figure_example_coverage}B).
+For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade due to the fact that they have no anatomical characters available (i.e.
+extant Herpestidae have a zero-length branches for anatomical characters).
+Such fossil, however, will have a higher probability to be placed along a branch that contains many anatomical characters (i.e.
+long branches for anatomical characters, e.g.
+Canidae or Ursidae (Figure \ref{Figure_example_coverage}B).
 

### 2016-02-09 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
Updated email
Changed anatomical characters semantics
Updated response (finished)

@@
 We accessed Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}) on the 11th of June 2015 and downloaded all 601 matrix.
-We then ran a shell script to select only the matrices that had any text element that match with one of the search terms.
-% LINK WHICH SHELL SCRIPT
+We then ran a shell script to select only the matrices that had any text element that match with one of the search terms (\url{https://github.com/TGuillerme/Missing_living_mammals/blob/master/Functions/select.files.sh}).
 To make the matrix selection more thorough, we ignored the keywords case as well as the latin suffix (\textit{ia}, \textit{ata}, \textit{ea}, and \textit{a}).
@@
 For example in primates the character \textit{p7} coded first by \cite{ross1998phylogenetic} is reused with the same living species in \cite{seiffert2003fossil}, \cite{marivaux2005anthropoid}, \cite{seiffert2005basal}, \cite{bloch2007new}, \cite{bloch2007new}, \cite{kay2008anatomy}, \cite{silcox2008biogeographic}, \cite{seiffert2009convergent}, \cite{tabuce2009anthropoid}, \cite{boyer2010astragalar}, \cite{seiffert2010fossil}, \cite{marivaux2013djebelemur} and \cite{ni2013oldest}.
+
+For example, the six living primates used in \cite{ross1998phylogenetic} (\textit{Aotus trivirgatus, Galago demidoff, Lemur catta, Microcebus murinus, Nycticebus coucang and Saimiri sciureus}) and their associated characters are reused along with more living species and characters in \cite{seiffert2003fossil}, \cite{marivaux2005anthropoid}, \cite{seiffert2005basal}, \cite{bloch2007new}, \cite{bloch2007new}, \cite{kay2008anatomy}, \cite{silcox2008biogeographic}, \cite{seiffert2009convergent}, \cite{tabuce2009anthropoid}, \cite{boyer2010astragalar}, \cite{seiffert2010fossil}, \cite{marivaux2013djebelemur} and \cite{ni2013oldest}.
 
@@
 sapiens} was transformed to \textit{Homo sapiens}).
-We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{bininda-emondsthe2007} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\url{https://paleobiodb.org/}).
+We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{bininda-emondsthe2007} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' (i.e.
+OTUs not present in the two previous data sets) all OTUs that were present in the Paleobiology database (\url{https://paleobiodb.org/}).
 For OTUs that did not appear in these three sources, we first decomposed the name (i.e.
@@
 Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; Figure ~\ref{Supp_figure_Taxonomic_algorithm}).
+Non-applicable OTUs were either specimen IDs with no related taxa (e.g.
+\textit{FMNHPR2081}), abbreviation that were not described in the associated paper (e.g.
+\textit{Ho.sap.}), non-mammals \textit{stricto-sensu} (e.g.
+\textit{Sinoconodon}), non standard taxonomic level (e.g.
+\textit{Spalcotheriids}) invalid taxonomic designation (e.g.
+\textit{sp\_nov\_1} or \textit{Outgroup}) or typos (e.g.
+\textit{Hobo sabpiens}).
 

### 2016-02-09 Manuscript/BiologyLetters/response.tex
Response wrapped up

@@
 \bigskip
-We are very grateful to the four referees for their helpful and constructive comments.s
+We are very grateful to the four referees for their helpful and constructive comments.
 We have taken all of their comments on board, and respond to their points below.
 To improve the clarity of this document, we dealt with each comment in the order they appeared in your ``Decision on Manuscript'' email and we have included the reviewers comments in blue.
+
+%TG: plus response to editors points:
+% *Please also confirm in your cover letter whether all the figures are your own or whether permission has been obtained for their use
+% *Please confirm whether your manuscript should be available via open access, as we note that this may be a requirement of your funder.
+We would like to state that all the figures are ore our own and the the manuscript should be available via open access.
+% *If you have any images that can be used to promote your article on social media (should this be accepted) please upload them as a supplementary file
+Additionally, we attached a supplementary file named \texttt{article\_promotion.eps} for promoting our article (should it be accepted) on social media.
+
+% *Please upload your original figure files as eps, tiff or jpeg files, rather than PDFs
+%TG: TO DO (easy)
 
@@
 Additionally, we observed that in newer studies (i.e.
-post 2010), living taxa with anatomical data were rarely added an usually reused from former studies (with sometimes the addition of extra characters).
+post 2010), living taxa with anatomical data were rarely added and usually reused from former studies (with sometimes the addition of extra characters and extra living taxa).
 We modified the sentence in the ESM1 part 1:
@@
 The straight relationship between the living species coverage and the Total Evidence phylogeny topology is hard to quantify precisely and depends on both the metric used for measuring topological differences and the method used to infer the tree.
-Our simulation showed that topology tend to get closer to a random one as morphological data is missing but that in the case of the missing living species, the other parameters we tested (missing data in the fossil record and number of morphological characters) add no significant effect (i.e.
+Our simulation showed that topology tend to get closer to a random one as morphological data is missing but that in the case of the missing living species, the other parameters we tested (missing data in the fossil record and number of morphological characters) had no significant effect (i.e.
 when a certain amount of living taxa are missing, the other parameters do not have any extra influence).
@@
 However, the two other matrices mentioned by reviewer 2 (Salles 1992 and Bryant et al. 1993) where not included in the present analysis.
-This is due to two main reasons: firstly, these two matrices where not available on any of the matrices repositories (either Graeme Lloyd's, Ross Mounce's or MorphoBank's) and (2) secondly, they didn't matched with out additional Google Scholars search parameters (i.e.
+This is due to two main reasons: firstly, these two matrices where not available on any of the matrices repositories (either Graeme Lloyd's, Ross Mounce's or MorphoBank's) and (2) secondly, they didn't matched with our additional Google Scholars search parameters (i.e.
 matrices published between 2010 and 2015 with accessible data).
@@
 \item{\textcolor{blue}{Lines 25-26: ``We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence phylogenies.''
-In a way, this is more than ``total evidence.=''.
+In a way, this is more than ``total evidence''.
 The initial total evidence paradigm was simply to use every character you could (molecular, morphological, even behavioral).
@@
 As tip-dating is the main theme of this issue, that should be linked to the total evidence explicitly here and elsewhere in the manuscript.}}
+
+We are aware of the difference between the ``Total Evidence method'' (i.e.
+a way to combine data in a matrix like supermatrix) and ``tip-dating method'' (i.e.
+a phylogenetic dating method, like node dating).
+Both methods can (and are often) be used together as in Ronquist et al 2012 (Systematics Biology) but we are aware that they can also be used really well separately.
+However, this study focuses \textit{indeed} on the ``Total Evidence method'' (i.e.
+a way to combine molecular and discrete anatomical data together) and not on the ``tip-dating'' one because it attempts to quantify the sate of the problem of missing data in living taxa underlined in Guillerme \& Cooper 2016 (again, a paper focusing on the ``Total Evidence method'', not including ``tip-dating'' nor ``node datin'' analysis).
+Nonetheless, we do understand this reviewer's point and added the mention of ``Total Evidence tip-dated phylogenies'' which is undoubtedly a great avenue of application of both methods.
+
+We modified the sentence in the abstract as follow: ``We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence tip-dated phylogenies.'' p@ lines@@
+
+And clarified the third sentence of the introduction as follows:
+``One increasingly popular method, the Total Evidence method [3], combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix that can then be used with the tip-dating method (e.g.
+[4, 3, 5, 1, 6]), producing a chronogram with living and fossil taxa at the tips.'' p@ lines@@
 
@@
 \item{\textcolor{blue}{Reviewer 3 (Pete Wagner)'s major problem - with which I agree - is semantic, and thus easily fixable with a few explicit caveats (e.g., ``by cladistic we mean...'').}}
+
+We modified the mention of ``cladistic characters'' to ``anatomical characters'' throughout the manuscript (see reviewer 3 comment 1).
 
@@
 % *An author contributions section which fulfils all *four* of our criteria given here https://royalsociety.org/journals/ethics-policies/openness/, including a sentence stating all authors agree to be held accountable for the content therein and approve the final version of the manuscript
-%TG: TO DO: CHECK
+%TG: Ok, we I'm not sure here.
+Is it just an automatic message: did all the author contributed well enough to be author (yes!); or is it that we have to modify the author contribution list to the exact wordings they put in this URL (seems odd though).
+
 % *A funding statement which is the same as that given in ScholarOne
-%TG: To modify
+%TG: Yes, is matching exactly.
+
 % *Please also confirm in your cover letter whether all the figures are your own or whether permission has been obtained for their use

### 2016-02-09 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Response wrapped up

@@
 In most orders, species with available data are randomly distributed across the phylogeny, which may reduce the impact of the problem.
-We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence phylogenies. 
+We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence tip-dated phylogenies. 
 \end{abstract}
@@
 One increasingly popular method, the Total Evidence method \cite{ronquista2012}, combines molecular data from living taxa and morphological data %TG: I left "morphological data" here since it TEM it doesn't have to be only anatomical characters
- from both living and fossil taxa in a supermatrix (e.g.
-\cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a phylogeny with living and fossil taxa at the tips.
+ from both living and fossil taxa in a supermatrix that can then be used with the tip-dating method  (e.g.
+\cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a chronogram with living and fossil taxa at the tips.
 A downside of this method is that it requires molecular data for living taxa and morphological data for both living and fossil taxa. %TG: same comment as above

### 2016-02-09 Manuscript/BiologyLetters/medianSummary.tex
Media summary

\documentclass{article} % Setting the class of the document

%For word count, you can install the super cool (and easy to use) WordCount package through Package Control

\begin{document} % starting the document

%What's the story
To understand changes of biodiversity, it is crucial to combine both living and fossil species in the same analysis.
We can do so by using the DNA from living species and enough anatomical data in common between living and fossil species.
%What we found
In this study we found that only a small amount of living mammals have available anatomical data and that this could be a problem for combining them with fossils.
%How's that gonna solve the world
We suggest that an increase in sampling effort in natural history museum collections, where this data lies, is necessary to efficiently combine living and fossil species in the same analysis.

\end{document} % Ending the document (don't forget this one!)

### 2016-02-12 Manuscript/BiologyLetters/medianSummary.tex
My attempt
Feel free to revert, I tried to make it more media friendly

@@
 
-%What's the story
-To understand changes of biodiversity, it is crucial to combine both living and fossil species in the same analysis.
-We can do so by using the DNA from living species and enough anatomical data in common between living and fossil species.
-%What we found
-In this study we found that only a small amount of living mammals have available anatomical data and that this could be a problem for combining them with fossils.
-%How's that gonna solve the world
-We suggest that an increase in sampling effort in natural history museum collections, where this data lies, is necessary to efficiently combine living and fossil species in the same analysis.
+Scientists use family trees to understand changes in biodiversity through time. But the trees often only use living species, ignoring extinct groups like mammoths. Therefore we must place living and 
+extinct species in the same trees.
+We can do so using the DNA from living species and anatomical data from living and extinct species.
+However, we found that few living mammals have enough available anatomical data to determine how they are related to extinct species.
+We suggest that we must collect more anatomical data from living species in museum collections to efficiently create family trees of living and extinct species.
 

### 2016-02-12 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Edits of response and manuscript

--- 2016-02-08 Manuscript/BiologyLetters/Missing_mammals-draft.tex
@@
 \begin{abstract}
-%\section{abstract} % NC: Just put this here as it then separates this for the word count
-Analyses of living and fossil taxa are crucial for understanding changes in biodiversity through time.
-The Total Evidence method allows living and fossil taxa to be combined in phylogenies, by using molecular data for living taxa and morphological data for both living and fossil taxa.
-With this method, substantial overlap of anatomical characters among living and fossil taxa is crucial for accurately inferring topology.
-However, although molecular data for living species is widely available, scientists using and generating morphological data mainly focus on fossils.
-Therefore, there is a gap in our knowledge of anatomical characters available for living taxa even in well-studied groups such as mammals.
+\section{abstract} % NC: Just put this here as it then separates this for the word count
 
-We investigated the amount of available anatomical characters for living mammals and how this data was phylogenetically distributed across orders.
-22 of 28 mammalian orders have \textless 25\% species with available anatomical characters; this has implications for the accurate placement of fossil taxa, although the issue is less pronounced at higher taxonomic levels.
-In most orders, species with available data are randomly distributed across the phylogeny, which may reduce the impact of the problem.
+% NC: I think anatomical here confuses things as we mean coded anatomical characters.
+I've tried to clarify.
+Also needed to clarify what you mean by "available".
+Also had to then cut some words to get it back under 200.
+Analyses of living and fossil taxa are crucial for understanding biodiversity through time.
+The Total Evidence method allows living and fossil taxa to be combined in phylogenies, using molecular data for living taxa and morphological data for living and fossil taxa.
+With this method, substantial overlap of coded anatomical characters among living and fossil taxa is vital for accurately inferring topology.
+However, although molecular data for living species are widely available, scientists generating morphological data mainly focus on fossils.
+Therefore, there are few coded anatomical characters in living taxa, even in well-studied groups like mammals.
+
+We investigated the number of coded anatomical characters available in phylogenetic matrices for living mammals and how these were phylogenetically distributed across orders.
+22 of 28 mammalian orders have \textless 25\% species with available characters; this has implications for the accurate placement of fossils, although the issue is less pronounced at higher taxonomic levels.
+In most orders, species with available characters are randomly distributed across the phylogeny, which may reduce the impact of the problem.
 We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence phylogenies. 
+% NC: I removed the tip dating bit here, I think yes peter is right in that we are in a special issue on that, BUT it is important for all TEM analyses so adding the tip dating bit here is daft.
 \end{abstract}
@@
 \noindent Key words: Total Evidence method, phylogenetic clustering, cladistic matrix, extinct, topology\\
-% NC: Might need some new ones as these are in the abstract now...
 
@@
 One increasingly popular method, the Total Evidence method \cite{ronquista2012}, combines molecular data from living taxa and morphological data %TG: I left "morphological data" here since it TEM it doesn't have to be only anatomical characters
- from both living and fossil taxa in a supermatrix (e.g.
-\cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a phylogeny with living and fossil taxa at the tips.
-A downside of this method is that it requires molecular data for living taxa and morphological data for both living and fossil taxa. %TG: same comment as above
+ from both living and fossil taxa in a supermatrix that can then be used with dating methods such as the tip-dating method (e.g.
+\cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), to produce a chronogram with living and fossil taxa at the tips.
+A downside of this method is that it requires molecular data for living taxa and morphological/anatomical data for both living and fossil taxa.
+%TG: same comment as above.
+I agree but have added anatomical to make it clear here what you mean when suddenly we switch to using anatomical.
 Sections of this data can be difficult, or impossible, to collect for every taxon in the analysis.
-For example, fossils rarely have molecular data and incomplete fossil preservation may restrict the amount of anatomical characters available.
+For example, fossils rarely have molecular data and incomplete fossil preservation may reduce the number of anatomical characters available.
 Additionally, it has become less common to collect anatomical characters for living taxa when molecular data is available (e.g.
@@
 Unfortunately this missing data can lead to errors in phylogenetic inference.
-Simulations show that the ability of the Total Evidence method to recover the correct topology decreases when there is little overlap between anatomical characters in living and fossil taxa, and that the effect of missing data on topology is greatest when living taxa have few anatomical characters available \cite{GuillermeCooper}.
-This is because (1) fossils will not be placed accurately within the correct clade if it contains no anatomical characters for living taxa; and (2) fossils have a higher probability of branching within clades with more anatomical characters available for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.
+Simulations show that the ability of the Total Evidence method to recover the correct topology decreases when there is little overlap between coded anatomical characters in living and fossil taxa, and that the effect of missing data on topology is greatest when living taxa have few anatomical characters available \cite{GuillermeCooper}.
+This is because (1) fossils will not be placed accurately within the correct clade if it contains no coded anatomical characters for living taxa; and (2) fossils have a higher probability of being placed within clades with more coded anatomical characters available for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.
 
-The issues above highlight that it is crucial to have sufficient anatomical characters available for living taxa in a clade before using a Total Evidence approach.
-However, it is unclear how much anatomical characters are actually available for living taxa, i.e.
+The issues above highlight that it is crucial to have sufficient coded anatomical characters available for living taxa in a clade before using Total Evidence approaches.
+However, it is unclear how many coded anatomical characters are actually available for living taxa, i.e.
 already coded from museum specimens and deposited in phylogenetic matrices accessible online, and how this data is distributed across clades.
@@
 Intuitively, most people assume this kind of data has already been collected, but empirical data suggest otherwise (e.g. in \cite{ronquista2012,slaterphylogenetic2013,beckancient2014}).
-To investigate this further, we assess the amount of available anatomical characters for living mammals to determine whether sufficient data exists to build reliable Total Evidence phylogenies in this group.
-We also determine whether the available anatomical characters are phylogenetically overdispersed or clustered across mammalian orders.
+To investigate this further, we assess the number of available coded anatomical characters for living mammals to determine whether enough data exists to build reliable Total Evidence phylogenies.
+We also determine whether the characters are phylogenetically overdispersed or clustered across mammalian orders.
 
@@
 %---------------------------------------------
+% NC: I've done a bit of word cutting here as I think we had wordcount issues?
 \section{Materials and Methods}
 \subsection{Data collection and standardisation}
-We downloaded all cladistic matrices %TG: I left "cladistic" here 'cause that's what the matrix is (even following Peter's definition)!
+We downloaded all cladistic matrices %TG: I left "cladistic" here 'cause that's what the matrix is (even following Peter's definition)! % NC: Yep
  containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{morphobank.org} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}).
-We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material (ESM) for a detailed description of the search procedure).
+We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material (ESM) for details).
 In total, we downloaded 286 matrices containing 5228 unique operational taxonomic units (OTUs). 
-We used OTUs rather than species because entries in the matrices ranged from species to families, and standardised the taxonomy as described in the ESM. 
-We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}.
+We used OTUs rather than species because entries in the matrices ranged from species to families. 
+We standardised the taxonomy as described in the ESM and excluded OTUs that were not present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal} to remove fossil species.
 
@@
 % After removing matrices with \textless 100 characters, we retained 1074 unique living OTUs from 126 matrices. % 1601 unique living OTUs for 286 matrices (no threshold)
-We retained 1601 unique living OTUs from 286 matrices.
+This resulted in 1601 unique OTUs from 286 matrices.
 
 \subsection{Data availability and distribution}
-To assess the availability of anatomical characters for each mammalian order, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
+To assess the availability of coded anatomical characters for each mammalian order and across mammals, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
 %TG: changed cladistic datat to anatomical characters
@@
 
-For each order, we investigated whether the available anatomical characters were (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index (NTI; \cite{webb2002phylogenies}) and the Net Relatedness Index (NRI; \cite{webb2002phylogenies}).
+For each order and for all mammals, we investigated whether the available coded anatomical characters were (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index (NTI; \cite{webb2002phylogenies}) and the Net Relatedness Index (NRI; \cite{webb2002phylogenies}).
 NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to it across the whole phylogeny \cite{Cooper2008}. 
@@
   \end{equation}
-where $\overline{MNND}_{obs}$ is the observed mean sum of the branch lengths between each of $n$ taxa with available anatomical characters and its nearest neighbour with available anatomical characters in the phylogeny,
+where $\overline{MNND}_{obs}$ is the observed mean sum of the branch lengths between each of $n$ taxa with available coded anatomical characters and its nearest neighbour with available coded anatomical characters in the phylogeny,
 $\overline{MNND}_{n}$ is the mean of 1000 $MNND$ between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random $MNND$ values.
@@
   \end{equation}
-where $\overline{MPD}_{obs}$ is the observed mean phylogenetic branch length of the tree containing only the $n$ taxa with available anatomical characters.
+where $\overline{MPD}_{obs}$ is the observed mean phylogenetic branch length of the tree containing only the $n$ taxa with available coded anatomical characters.
 Negative NTI and NRI values show that the focal taxa are more overdispersed across the phylogeny than expected by chance, and positive values reflect clustering.
 
-We calculated NTI and NRI values for each mammalian order separately, at each different taxonomic level. 
-For each analysis our focal taxa were those with available anatomical characters at that taxonomic level and the phylogeny was that of the order pruned from \cite{BinindaEmonds}.
+We calculated NTI and NRI values for all mammals or each mammalian order separately, at each different taxonomic level. 
+For each analysis our focal taxa were those with available coded anatomical characters at that taxonomic level and the phylogeny was the order pruned from \cite{BinindaEmonds}.
 
@@
 \section{Results}
-11 out of 28 orders have low coverage (\textless 25\% species with available anatomical characters) and 7 have high coverage (\textgreater 75\% species with available anatomical characters) at the species-level.
-At the genus-level, one order have low coverage and 15 have high coverage, and at the family-level, no orders have low coverage and 25 have high coverage (Table\ref{Table_results}).
+Across mammals, species coverage was low (\textless 25\% species with available coded anatomical characters) but family coverage was high (\textgreater 75\% families with available coded anatomical characters).
+For each order, 11 out of 28 had low coverage and seven had high coverage at the species-level.
+At the genus-level, one order had low coverage and 15 had high coverage, and at the family-level, no orders had low coverage and 25 had high coverage (Table\ref{Table_results}).
 
@@
 
-Only seven orders had significantly clustered data (Cetartiodactyla, Cingulata, Pilosa and Rodentia at the species-level and Carnivora, Chiroptera and Soricomorpha at both species- and genus-level) and none had significantly overdispersed data (Table~\ref{Table_results}).
+Across mammals, taxa with available coded anatomical characters were significantly clustered using NTI at the species- and genus-level.
+For each order, only seven showed significant clustering (Cetartiodactyla, Cingulata, Pilosa and Rodentia at the species-level and Carnivora, Chiroptera and Soricomorpha at both species- and genus-level) and none showed significant overdispersion (Table~\ref{Table_results}).
 
-Figure \ref{Figure_example_coverage} shows randomly distributed OTUs with cladistic data in Primates (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered OTUs with cladistic data in Carnivora (mainly Canidae and Urisdae but no Herpestidae; Figure \ref{Figure_example_coverage}B).
+Figure \ref{Figure_example_coverage} shows randomly distributed OTUs with available coded anatomical characters in Primates (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered OTUs with available coded anatomical characters in Carnivora (mainly Canidae and Urisdae but no Herpestidae; Figure \ref{Figure_example_coverage}B).
 
@@
     \includegraphics[width=1\textwidth]{figure1.pdf}
-\caption{Phylogenetic distribution of species with available anatomical characters across two orders (A: Primates; B: Carnivora).
-Blue branches indicate available anatomical characters for the species.}
+\caption{Phylogenetic distribution of species with available coded anatomical characters across two orders (A: Primates; B: Carnivora).
+Blue branches indicate species with available coded anatomical characters.}
 \label{Figure_example_coverage}
@@
 Our results show that although phylogenetic relationships among living mammals are well-resolved (e.g.
-\cite{BinindaEmonds,meredithimpacts2011}), most of the data used to build these phylogenies is molecular, and few anatomical characters are available for living mammals compared to fossils (e.g.
+\cite{BinindaEmonds,meredithimpacts2011}), most of the data used to build these phylogenies is molecular, and few coded anatomical characters are available for living mammals compared to fossils (e.g.
 \cite{O'Leary08022013,ni2013oldest}).
-This has implications for building Total Evidence phylogenies containing both living and fossil mammals, as without sufficient available anatomical characters for living species, fossil placements in these trees are very uncertain \cite{GuillermeCooper}.
+This has implications for building Total Evidence phylogenies, as without sufficient overlapping anatomical characters for living and fossil species, fossil placements in these trees may be unreliable \cite{GuillermeCooper}.
 
-The number of living mammalian OTUs with no available anatomical characters was surprisingly high at the species-level: only seven out of 28 orders have a high coverage of taxa with available anatomical characters.
-This high coverage threshold of 75\% of taxa with available anatomical characters represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
+The number of living mammalian OTUs with available coded anatomical characters was surprisingly low at the species-level: only 17\%.
+Only seven out of 28 orders have a high coverage of taxa with available coded anatomical characters.
+This high coverage threshold of 75\% of taxa with available characters represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
 Beyond this threshold, there is considerable displacement of wildcard taxa and decreased clade conservation \cite{GuillermeCooper}.
-Therefore we expect difficulties in placement of fossils at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels.
-This point is important from a practical point of view because of the slight discrepancy between neontological and palaeontological species concepts.
-While neontological species are described using morphology, genes, distribution etc.; palaeontological species can be based only on morphological, spatial and temporal data (e.g. \cite{ni2013oldest}).
+Therefore we expect difficulties in placing fossils at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels.
+This is important in practice because of the slight discrepancy between neontological and palaeontological species concepts.
+While neontological species are described using morphology, genes, distribution etc.; palaeontological species can be based only on morphological/anatomical, spatial and temporal data (e.g.
+\cite{ni2013oldest}).
 Therefore, many palaeontological studies use the genus (or monospecific genera) as their smallest OTU (e.g.
@@
 
-When few species have available anatomical characters, the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion) to maximize the possibilities of a fossil branching in the right clade.
-The second best scenario is that species with available anatomical characters are randomly distributed across the phylogenies. 
-Here we expect no special bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with available anatomical characters were randomly distributed across the phylogeny.
-The worst case scenario for fossil placement is that species with available anatomical characters are phylogenetically clustered. 
-Then we expect two major biases to occur: first, fossils will not be able to branch within a clade containing no data, and second, fossils will have higher probability of branching within the most sampled clade by chance.
+When few species have available coded anatomical characters, the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion) to maximize the possibilities of a fossil being placed in the correct clade.
+The second best scenario is that species with available characters are randomly distributed across the phylogeny. 
+Here we expect no bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with available coded anatomical characters were randomly distributed across the phylogeny.
+The worst case scenario for fossil placement is that species with available characters are phylogenetically clustered. 
+Then we expect two major biases: first, fossils will not be placed within a clade containing no data, and second, fossils will have higher probability of being placed within the most sampled clade by chance.
 Our results suggest that this may be problematic at the genus-level in Carnivora, Chiroptera and Soricomorpha. 
-For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade% rev 1.2.7
-, and will have more chance to randomly branch within Canidae or Ursidae (Figure \ref{Figure_example_coverage}B).
+For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade because they have no coded anatomical characters available.
+Instead the fossil will have a high probability of being placed on a branch that contains many anatomical characters such as within the Canidae or Ursidae (Figure \ref{Figure_example_coverage}B). 
+This is analogous to the problem of long-branch attraction/short branch repulsion, as one can think of herpestids as having zero-length branches for anatomical characters, and canids and ursids having long branches and thus "attracting" fossil placements.
 
-It is worth noting, however, that our analysis did not include all the matrices containing anatomical characters ever published.
-In fact, our data collection procedure focused on including studies that provided matrices easily accessible, i.e.
-we did specifically not include any matrices that were only available in paper format (e.g.
-printed in books), non-reusable format (e.g.
-an image of the matrix) %TG: yes...
-There were quiet a few of them!
-or/and matrices available only upon request (e.g. by emailing the authors).
-Such matrices are likely to not be included in Total Evidence analysis due to the difficulty of obtaining them.
+It is worth noting that our analysis did not include all the matrices containing coded anatomical characters ever published.
+Instead of performing exhaustive searches, our data collection procedure focused on including studies that provided easily accessible matrices.
+We specifically did not include matrices that were only available in books, in non-reusable formats (e.g.
+an image of the matrix that would require text-mining tools or recreating the matrix by hand), or matrices only available upon request.
+The difficulty of obtaining such matrices makes them less likely to be included in Total Evidence analyses, especially where the number of taxa to include in the phylogeny is high and time is limited.
 
-Despite the absence of good cladistic data coverage for living mammals, %TG: I think leaving cladistic here is OK no?
- the Total Evidence method still seems to be the most promising way of combining living and fossil data for macroevolutionary analyses.
-It is also encouraging to note that the sampling effort is consistent with diversity at each taxonomic level and there is therefore no direct bias in sampling at any taxonomic level (ESM 2, Figure 1).
-Following the recommendations in \cite{GuillermeCooper}, we need to code anatomical characters for as many living species possible. 
-Fortunately, data for living mammals is usually readily available in natural history collections, therefore, we propose that an increased effort be put into coding anatomical characters from living species, possibly by engaging in collaborative data collection projects.
-Such an effort would be valuable not only to phylogeneticists, but also to any researcher focusing understanding macroevolutionary patterns and processes.
+Despite the absence of good morphological/anatomical data coverage for living mammals, %TG: I think leaving cladistic here is OK no? % NC: Better to be consistent with rest of manuscript
+ the Total Evidence method still seems to be the most promising way of combining living and fossil species in macroevolutionary analyses.
+%It is also encouraging to note that sampling effort is consistent with diversity at each taxonomic level and there is therefore no direct bias in sampling at any taxonomic level (ESM 2, Figure 1).
+% NC: Needs more explanation, why is this encouraging??? Maybe leave this as the paper is already over long.
+Following the recommendations in \cite{GuillermeCooper}, we must code anatomical characters for as many living species as possible. 
+Fortunately, mammal specimens are usually readily available in natural history collections, therefore, we propose increased effort into coding anatomical characters from living species, possibly by engaging in collaborative data collection projects.
+Such efforts would be valuable not only to phylogeneticists, but also to any researcher focusing understanding macroevolutionary patterns and processes.
 
@@
 \section{Authors' Contributions}
-T.G. and N.C designed the study. T.G. analysed the data. T.G. and N.C. wrote the the manuscript.
+TG and NC designed the study. TG analysed the data. TG and NC wrote the the manuscript.
 \section{Competing Interests}
@@
 \section{Acknowledgments}
-We thank David Bapst, Graeme Lloyd, Nick Matzke and April Wright.
-Thanks to one anonymous reviewer, Graham Slater, Peter Wagner and Graeme Lloyd for their useful comments.
-%TG: That's two times Graeme though...
+We thank David Bapst, Graeme Lloyd, Nick Matzke, April Wright, Graham Slater, Peter Wagner and one anonymous reviewer.
+%TG: That's two times Graeme though...I've shortened it due to wordcount issues
 \section{Funding statement}
-This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).
+This was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal: 321696).
 

### 2016-02-13 Manuscript/BiologyLetters/response.tex

@@
 Additionally, as Graeme Lloyd mentions, ``older studies tend to be smaller (due to computing limitations) and to commonly reoccur anyway due to data set reuse'' (Reviewer 4, comment 1).
-Following these two observations, we think that including the 20 most recent results % NC: Is that what you did? TG: yep.
- for each order (and higher taxonomic level) was a good balance between the time spent looking for papers, the number of OTUs extracted and the repeatability of the study.
+Following these two observations, we think that including the 20 most recent results for each order (and higher taxonomic level) was a good balance between the time spent looking for papers, the number of OTUs extracted and the repeatability of the study.
 We are aware however that we might have missed some specific studies (see Reviewer 2 comment 2) but are confident that our protocol gives a good overall picture of the data availability and that the three repositories used before the Google Scholar search (MorphoBank, Graeme Lloyd's and Ross Mounce's) contained most of the available data.
 In fact, as mentioned in ESM 1, part 1 ``among the 660 papers [collected from the Google Scholar search], only 50 contained a total of 425 extra living OTUs''.
+% NC: How many OTUs were in the other 610 papers?
 
@@
 Additionally, we observed that in newer studies (i.e.
-post 2010), living taxa with anatomical data were rarely added and usually reused from former studies (with sometimes the addition of extra characters and extra living taxa).
+post 2010), living taxa with anatomical data were rarely added and usually reused from former studies (with the occasional addition of extra characters and extra living taxa).
 We modified the sentence in the ESM1 part 1:
@@
 
-The straight relationship between the living species coverage and the Total Evidence phylogeny topology is hard to quantify precisely and depends on both the metric used for measuring topological differences and the method used to infer the tree.
-Our simulation showed that topology tend to get closer to a random one as morphological data is missing but that in the case of the missing living species, the other parameters we tested (missing data in the fossil record and number of morphological characters) had no significant effect (i.e.
-when a certain amount of living taxa are missing, the other parameters do not have any extra influence).
-In our simulations we tested 5 different level of missing living taxa (0\%, 10\%, 25\%, 50\% and 75\% of missing living taxa) and we observed that the three first levels had no significant negative effect on topology in relation with the amount of missing data in the fossil record and the number of morphological characters but that the fifth level (75\% of missing living taxa, i.e.
+The relationship between the living species coverage and the Total Evidence phylogeny topology is hard to quantify precisely and depends on both the metric used for measuring topological differences and the method used to infer the tree.
+Our simulations showed that topology tends to get closer to random as morphological data is removed but that in the case of missing data for living species, the other parameters we tested (missing data in the fossil record and number of morphological characters) had no significant effect (i.e.
+when a certain number of living taxa are missing, the other parameters do not have any extra influence).
+In our simulations we tested five different levels of missing living taxa (0\%, 10\%, 25\%, 50\% and 75\% missing living taxa) and we observed that the three first levels had no significant negative effect on topology in relation with the amount of missing data in the fossil record and the number of morphological characters but that the fifth level (75\% of missing living taxa, i.e.
 25\% coverage) had a strong negative effect.
 
-Also, we highlighted the different levels of coverage in Table 1 as follow: at any taxonomic level, orders with less than 25\% coverage are highlighted in blue (or dark grey in the black and white version) and orders with more than 75\% coverage are highlithed in orange (or light grey in the black and white version).
+Also, we highlighted the different levels of coverage in Table 1 as follow: at any taxonomic level, orders with less than 25\% coverage are highlighted in blue (or dark grey in the black and white version) and orders with more than 75\% coverage are highlighted in orange (or light grey in the black and white version).
 
@@
 
-We removed the analysis with a threshold of OTUs with at least 100 characters and replaced it with the analysis without any minimal number of morphological characters.
-We retracted the paragraph about the character threshold analysis and updated the results of figure 1 and table 1 to contain all living OTUs; we removed the redundant ESM 2 and we updated the figures
+We completely see the reviewer's point and have removed this threshold and replaced it with the analysis without any minimal number of morphological characters.
+We also removed the paragraph about the character threshold analysis and updated the results of figure 1 and table 1 to contain all living OTUs; we removed the redundant ESM 2 and we updated the figures.
 
@@
 
-By removing the characters threshold (see above) we now included the taxa coded in Wang and Carranza-Casta\~{n}eda 2008 (matrix reference GS2014-WCC) and other matrices ignored in the former analysis increasing the number of Carnivora from 42 species (14\% of total number of species) to 76 (27\%).
-However, the two other matrices mentioned by reviewer 2 (Salles 1992 and Bryant et al. 1993) where not included in the present analysis.
-This is due to two main reasons: firstly, these two matrices where not available on any of the matrices repositories (either Graeme Lloyd's, Ross Mounce's or MorphoBank's) and (2) secondly, they didn't matched with our additional Google Scholars search parameters (i.e.
-matrices published between 2010 and 2015 with accessible data).
+Again we completely see the reviewer's point and have modified our manuscript as follows.
+By removing the threshold (see above) we now included the taxa coded in Wang and Carranza-Casta\~{n}eda 2008 (matrix reference GS2014-WCC) and other matrices excluded in the former analysis increasing the number of Carnivora from 42 species (14\% of total number of species) to 76 (27\%).
+However, the two other matrices mentioned by reviewer 2 (Salles 1992 and Bryant et al. 1993) were not included in the new analysis.
+This is due to two main reasons: firstly, these two matrices were not available on any of the repositories (either Graeme Lloyd's, Ross Mounce's or MorphoBank's), and secondly, they didn't appear in our additional Google Scholar searches because they were published before 2010.
 Following this reviewer's comment as well as reviewer 4's we added a paragraph in the discussion to clearly state that we are aware of data sampling biases:
@@
 Such matrices are likely to not be included in Total Evidence analysis due to the difficulty of obtaining them.
+% NC: I think this last line is a bit dangerous.
+Anyone making a proper phylogeny would include these, especially for a smaller group.
+It's what people like Graham spend years doing.
+Perhaps rephrase: "The difficulty of obtaining such matrices makes them less likely to be included in Total Evidence analyses, especially where the number of taxa to include in the phylogeny is high."
 '' p.@ lines@@.
@@
 Concerning the Gaubert et al. 2005's study specifically, this study unfortunately did not contain any easily available matrix.
-In fact, the matrix mentioned as Appendix 2 in the paper is not present in the article and neither the TreeBase accession numbers (``S1255'', ``M2193'') nor the provided URL (``www.systematicbiology.org'') links to the matrix.
+In fact, the matrix mentioned as Appendix 2 in the paper is not present in the article and neither of the TreeBase accession numbers (``S1255'', ``M2193'') nor the provided URL (``www.systematicbiology.org'') links to the matrix.
 % TG: I don't know if that's a good justification but basically, it's impossible to get this specific matrix.
+% NC: Hmmm it's tricky as I said above.
+If you were really building a tree you would find this kind of data somehow.
+Also wouldn't this not appear because you used the 2010 cutoff?
+I think tbh Graham has some good points - you did this the quick and dirty way rather than the thorough way someone would do it if they were constructing a tree.
 
@@
 '' p.@ lines@@
+
+% NC: OK this is not what Graham meant.
+I think what he is referring to is that when you have more people working on a group (effort) you are likely to get more taxa with coded data.
+It's like in taxonomy where groups with more taxonomists (eg primates) have more species than groups with fewer (eg rodents).
+I think what he'd want is a regression of number of papers (effort) against percentage of coded species/OTUs with each order being a datapoint.
 
@@
 We agree with this reviewers semantinc point and changed ``cladistic'' and ``morphological'' data to ``anatomical characters'' throughout the manuscript.
+% NC: all he is saying is cladistic is not morphological and shouldn't be used as proxies for each other.
+And that really when we say morphological we mean phylogenetically coded anatomical/morphological characters.
+I've changed a few of these back in the manuscript to try and make it a bit easier to understand, as anatomical suggests something else I think.
 
@@
 In fact, among the 286 matrices used in this analysis, we found 5228 unique OTUs including 3627 fossil ones which is pretty impressive knowing the amount of work required to collect data for a single specimen!
-However, in this study, we focused only on the amount of living taxa (1601 unique OTUs containing 847 species) as a way to solve the problems associated with the lack of living taxa on tree topology described in Guillerme \& Cooper 2016.
+However, in this study, we focused only on the number of *living* taxa (1601 unique OTUs containing 847 species) as a way to solve the problems associated with the lack of living taxa on tree topology described in Guillerme \& Cooper 2016 (essentially that if you don't have overlapping coded anatomical data for your living species then there is a low likelihood of correctly placing the fossil taxa no matter how good the sampling is for those species).
 
@@
 However, this study focuses \textit{indeed} on the ``Total Evidence method'' (i.e.
-a way to combine molecular and discrete anatomical data together) and not on the ``tip-dating'' one because it attempts to quantify the sate of the problem of missing data in living taxa underlined in Guillerme \& Cooper 2016 (again, a paper focusing on the ``Total Evidence method'', not including ``tip-dating'' nor ``node datin'' analysis).
-Nonetheless, we do understand this reviewer's point and added the mention of ``Total Evidence tip-dated phylogenies'' which is undoubtedly a great avenue of application of both methods.
+a way to combine molecular and discrete anatomical data together) and not on the ``tip-dating'' one because it attempts to quantify the problem of missing data in living taxa underlined in Guillerme \& Cooper 2016 (again, a paper focusing on the ``Total Evidence method'', not including ``tip-dating'' nor ``node dating'' analysis).
+Nonetheless, we do understand the reviewer's point and added the mention of ``Total Evidence tip-dated phylogenies'' which is undoubtedly a great avenue of application of both methods.
 
-We modified the sentence in the abstract as follow: ``We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence tip-dated phylogenies.'' p@ lines@@
+We modified the sentence in the abstract as follows: ``We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence tip-dated phylogenies.'' p@ lines@@
 
@@
 
-We do agree with this comment and changed this comment to:
-``For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade due to the fact that they have no anatomical characters available (i.e.
-extant Herpestidae have a zero-length branches for anatomical characters).
-Such fossil, however, will have a higher probability to be placed along a branch that contains many anatomical characters (i.e.
+We agree with this comment and changed this to:
+``For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade because they have no anatomical characters available (i.e.
+extant Herpestidae have a zero-length branch for anatomical characters).
+Such fossils, however, will have a higher probability of being placed on a branch that contains many anatomical characters (i.e.
 long branches for anatomical characters, e.g.
@@
 
-We developed description of our data collection protocol in the ESM 1 (see Reviewer 1 comment 2) and underlined the caveats in the discussion (see Reviewer 2 comment 2).
+We developed the description of our data collection protocol in the ESM 1 (see Reviewer 1 comment 2) and underlined the caveats in the discussion (see Reviewer 2 comment 2).
 
@@
 
-We removed the character threshold in the whole analysis and specified what we meant by ``data accessibility'' in the discussion (see Reviewer 2 comment 2).
+We thank the reviewer for their practical stance on recollecting the dataset.
+We have added some comments to the manuscript to deal with some of the biases in our procedure and why some matirces will be missing.
+We also removed the character threshold and specified what we meant by ``data accessibility'' in the discussion (see Reviewer 2 comment 2). See also our responses to the reviewer comments above.
 

### 2016-02-13 Manuscript/BiologyLetters/response.tex
Dealt with comments

@@
 Additionally, we attached a supplementary file named \texttt{article\_promotion.eps} for promoting our article (should it be accepted) on social media. 
+% TO DO!
 
@@
 We are aware however that we might have missed some specific studies (see Reviewer 2 comment 2) but are confident that our protocol gives a good overall picture of the data availability and that the three repositories used before the Google Scholar search (MorphoBank, Graeme Lloyd's and Ross Mounce's) contained most of the available data.
-In fact, as mentioned in ESM 1, part 1 ``among the 660 papers [collected from the Google Scholar search], only 50 contained a total of 425 extra living OTUs''.
-% NC: How many OTUs were in the other 610 papers?
+In fact, as mentioned in ESM 1, part 1 ``among the 660 papers [collected from the Google Scholar search], only 50 contained a total of 425 extra living OTUs on a total of 4950 OTUs''.
 
@@
 printed in books), non-reusable format (e.g.
-an image of the matrix) %TG: yes...
-There were quiet a few of them!
-or/and matrices available only upon request (e.g. by emailing the authors).
-Such matrices are likely to not be included in Total Evidence analysis due to the difficulty of obtaining them.
-% NC: I think this last line is a bit dangerous.
-Anyone making a proper phylogeny would include these, especially for a smaller group.
-It's what people like Graham spend years doing.
-Perhaps rephrase: "The difficulty of obtaining such matrices makes them less likely to be included in Total Evidence analyses, especially where the number of taxa to include in the phylogeny is high."
+an image of the matrix)  or/and matrices available only upon request (e.g.
+by emailing the authors).
+The difficulty of obtaining such matrices makes them less likely to be included in Total Evidence analyses, especially where the number of taxa to include in the phylogeny is high.
 '' p.@ lines@@.
@@
 We agree with this reviewers semantinc point and changed ``cladistic'' and ``morphological'' data to ``anatomical characters'' throughout the manuscript.
-% NC: all he is saying is cladistic is not morphological and shouldn't be used as proxies for each other.
-And that really when we say morphological we mean phylogenetically coded anatomical/morphological characters.
-I've changed a few of these back in the manuscript to try and make it a bit easier to understand, as anatomical suggests something else I think.
 
@@
 a way to combine molecular and discrete anatomical data together) and not on the ``tip-dating'' one because it attempts to quantify the problem of missing data in living taxa underlined in Guillerme \& Cooper 2016 (again, a paper focusing on the ``Total Evidence method'', not including ``tip-dating'' nor ``node dating'' analysis).
-Nonetheless, we do understand the reviewer's point and added the mention of ``Total Evidence tip-dated phylogenies'' which is undoubtedly a great avenue of application of both methods.
+Nonetheless, we do understand the reviewer's point and added the mention of ``Total Evidence tip-dated phylogenies'' which is undoubtedly a great avenue of application of both methods but we believe that the implications of the present study are important for all Total Evidence analysis, not only when using such matrices for tip-dating.
 
-We modified the sentence in the abstract as follows: ``We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence tip-dated phylogenies.'' p@ lines@@
-
-And clarified the third sentence of the introduction as follows:
+We clarified the third sentence of the introduction as follows:
 ``One increasingly popular method, the Total Evidence method [3], combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix that can then be used with the tip-dating method (e.g.
@@
 We thank the reviewer for this mention and we will be sure to use this database for future studies.
-% NC: If you end up needing to redo the data collection it might be worth checking the PBDB...
 
@@
 We agree with this comment and changed this to:
-``For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade because they have no anatomical characters available (i.e.
-extant Herpestidae have a zero-length branch for anatomical characters).
-Such fossils, however, will have a higher probability of being placed on a branch that contains many anatomical characters (i.e.
-long branches for anatomical characters, e.g.
-Canidae or Ursidae (Figure 1B).''
-
-% NC: Yep he's right, modify the sentence to account for this.
-Also cite a long branch attraction stuff.
-Check if Pete has written anything - it'd be nice to cite him as he's been so helpful.
-% TG: I looked for relevant paper on this "short-branch repulsion" but didn't found any that's mildly relevant from Peter...
+``For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade because they have no coded anatomical characters available.
+Instead the fossil will have a high probability of being placed on a branch that contains many anatomical characters such as within the Canidae or Ursidae (Figure 1B). 
+This is analogous to the problem of long-branch attraction/short branch repulsion, as one can think of herpestids as having zero-length branches for anatomical characters, and canids and ursids having long branches and thus ``attracting'' fossil placements.''
 
@@
 This could surely be synthesised from the tables and would be a nice simple way to compare the with and without threshold results.
-Similarly, I think Reviewer 3's [We assume Graeme Lloyd is refering to Graham Slater's reviewe (reviewer 2)] %or same as comment before: no names!
- graphs are a nice potential addition to the ESM.}}
+Similarly, I think Reviewer 3's [We assume Graeme Lloyd is refering to Graham Slater's reviewe (reviewer 2)] graphs are a nice potential addition to the ESM.}}
 
@@
 We added the following sentence to in the methods:
-``For each analysis our focal taxa were those with available anatomical characters at that taxonomic level and the phylogeny was that of the order pruned from [10].
-For each order, we pruned the phylogeny from [10] to include only the species/genus/families from this order.''.
-p@.
-lines@@.
+``For each analysis our focal taxa were those with available coded anatomical characters at that taxonomic level and the phylogeny was the order pruned from [10].''. p@. lines@@.
 
@@
 
-Thomas Guillerme (t.guillerme@imperial.ac.uk)\\ % Changing email + adding imperial address?
-% NC: Just change email.
-For the paper, you can change your corresponding address but I wouldn't add imperial as another proper address as you did most of the work at trinity.
+Thomas Guillerme (t.guillerme@imperial.ac.uk)\\
 Natalie Cooper

### 2016-02-13 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Dealt with comments

@@
 \begin{abstract}
-\section{abstract} % NC: Just put this here as it then separates this for the word count
+\section{abstract}
 
-% NC: I think anatomical here confuses things as we mean coded anatomical characters.
-I've tried to clarify.
-Also needed to clarify what you mean by "available".
-Also had to then cut some words to get it back under 200.
 Analyses of living and fossil taxa are crucial for understanding biodiversity through time.
@@
 We investigated the number of coded anatomical characters available in phylogenetic matrices for living mammals and how these were phylogenetically distributed across orders.
-22 of 28 mammalian orders have \textless 25\% species with available characters; this has implications for the accurate placement of fossils, although the issue is less pronounced at higher taxonomic levels.
+11 of 28 mammalian orders have \textless 25\% species with available characters; this has implications for the accurate placement of fossils, although the issue is less pronounced at higher taxonomic levels.
 In most orders, species with available characters are randomly distributed across the phylogeny, which may reduce the impact of the problem.
 We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence phylogenies. 
-% NC: I removed the tip dating bit here, I think yes peter is right in that we are in a special issue on that, BUT it is important for all TEM analyses so adding the tip dating bit here is daft.
 \end{abstract}
@@
 \section{Introduction}
-
 There is an increasing consensus among biologists that studying both living and fossil taxa is essential for fully understanding macroevolutionary patterns and processes \cite{slaterunifying2013,fritzdiversity2013}.
 To perform such analyses it is necessary to combine living and fossil taxa in phylogenetic trees.
-One increasingly popular method, the Total Evidence method \cite{ronquista2012}, combines molecular data from living taxa and morphological data %TG: I left "morphological data" here since it TEM it doesn't have to be only anatomical characters
- from both living and fossil taxa in a supermatrix that can then be used with dating methods such as the tip-dating method (e.g.
-\cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), to produce a chronogram with living and fossil taxa at the tips.
+One increasingly popular method, the Total Evidence method \cite{ronquista2012}, ccombines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix that can then be used with the tip-dating method (e.g.
+\cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a chronogram with living and fossil taxa at the tips.
 A downside of this method is that it requires molecular data for living taxa and morphological/anatomical data for both living and fossil taxa.
-%TG: same comment as above.
-I agree but have added anatomical to make it clear here what you mean when suddenly we switch to using anatomical.
 Sections of this data can be difficult, or impossible, to collect for every taxon in the analysis.
@@
 already coded from museum specimens and deposited in phylogenetic matrices accessible online, and how this data is distributed across clades.
-% TG: emphasize the "availibility" definition her?
 Intuitively, most people assume this kind of data has already been collected, but empirical data suggest otherwise (e.g. in \cite{ronquista2012,slaterphylogenetic2013,beckancient2014}).
@@
 %---------------------------------------------
-% NC: I've done a bit of word cutting here as I think we had wordcount issues?
+
 \section{Materials and Methods}
 \subsection{Data collection and standardisation}
-We downloaded all cladistic matrices %TG: I left "cladistic" here 'cause that's what the matrix is (even following Peter's definition)! % NC: Yep
- containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{morphobank.org} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}).
+We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{morphobank.org} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}).
 We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material (ESM) for details).
@@
 We standardised the taxonomy as described in the ESM and excluded OTUs that were not present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal} to remove fossil species.
-
-% Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having overlapping characters with other matrices \cite{wagner2000} and (2) they are more likely to contain specific characters that are not applicable% rev 1.2.4
-%  across large clades (e.g. ``antler ramifications'' is a character that is only applicable to Cervidae \cite{Brazeau2011}).
-% Therefore we selected only matrices containing \textgreater 100 characters for each OTU.
-% This threshold was chosen to correspond with the number of characters used in \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
-% Results of analyses with no threshold are available in the ESM. 
-% After removing matrices with \textless 100 characters, we retained 1074 unique living OTUs from 126 matrices. % 1601 unique living OTUs for 286 matrices (no threshold)
 This resulted in 1601 unique OTUs from 286 matrices.
@@
 To assess the availability of coded anatomical characters for each mammalian order and across mammals, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
-%TG: changed cladistic datat to anatomical characters
 We consider orders with \textless 25\% of living taxa with available anatomical characters as having low data coverage, and orders with \textgreater 75\% of living taxa with available anatomical characters as having high data coverage.
@@
 
-\input{Table_results}
+%\input{Table_results}
 
@@
 
-\begin{figure}[!htbp]
-\centering
-    \includegraphics[width=1\textwidth]{figure1.pdf}
-\caption{Phylogenetic distribution of species with available coded anatomical characters across two orders (A: Primates; B: Carnivora).
-Blue branches indicate species with available coded anatomical characters.}
-\label{Figure_example_coverage}
-\end{figure}
+%\begin{figure}[!htbp]
+%\centering
+%    \includegraphics[width=1\textwidth]{figure1.pdf}
+%\caption{Phylogenetic distribution of species with available coded anatomical characters across two orders (A: Primates; B: Carnivora).
+%Blue branches indicate species with available coded anatomical characters.}
+%\label{Figure_example_coverage}
+%\end{figure}
 
@@
 
-It is worth noting that our analysis did not include all the matrices containing coded anatomical characters ever published.
-Instead of performing exhaustive searches, our data collection procedure focused on including studies that provided easily accessible matrices.
-We specifically did not include matrices that were only available in books, in non-reusable formats (e.g.
-an image of the matrix that would require text-mining tools or recreating the matrix by hand), or matrices only available upon request.
-The difficulty of obtaining such matrices makes them less likely to be included in Total Evidence analyses, especially where the number of taxa to include in the phylogeny is high and time is limited.
+It is worth noting, however, that our analysis did not include all the matrices containing anatomical characters ever published.
+In fact, our data collection procedure focused on including studies that provided matrices easily accessible, i.e.
+we did specifically not include any matrices that were only available in paper format (e.g.
+printed in books), non-reusable format (e.g.
+an image of the matrix)  or/and matrices available only upon request (e.g.
+by emailing the authors).
+The difficulty of obtaining such matrices makes them less likely to be included in Total Evidence analyses, especially where the number of taxa to include in the phylogeny is high.
 
-Despite the absence of good morphological/anatomical data coverage for living mammals, %TG: I think leaving cladistic here is OK no? % NC: Better to be consistent with rest of manuscript
- the Total Evidence method still seems to be the most promising way of combining living and fossil species in macroevolutionary analyses.
-%It is also encouraging to note that sampling effort is consistent with diversity at each taxonomic level and there is therefore no direct bias in sampling at any taxonomic level (ESM 2, Figure 1).
-% NC: Needs more explanation, why is this encouraging??? Maybe leave this as the paper is already over long.
+Despite the absence of good morphological/anatomical data coverage for living mammals, the Total Evidence method still seems to be the most promising way of combining living and fossil species in macroevolutionary analyses.
 Following the recommendations in \cite{GuillermeCooper}, we must code anatomical characters for as many living species as possible. 
@@
 We thank David Bapst, Graeme Lloyd, Nick Matzke, April Wright, Graham Slater, Peter Wagner and one anonymous reviewer.
-%TG: That's two times Graeme though...I've shortened it due to wordcount issues
 \section{Funding statement}

### 2016-02-13 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
Dealt with comments

@@
 We selected only the 20 first results per search term to avoid downloading articles that were too irrelevant.
-Among the 660 papers, only 50 contained a total of 425 extra living OTUs (Figure ~\ref{Supp_figure_google_searches}).
+Among the 660 papers, only 50 contained a total of 425 extra living OTUs on a total of 4950 OTUs (Figure ~\ref{Supp_figure_google_searches}).
 Also we decided to select only the articles published since 2010 because nearly every one of the recent published matrix contains both a fraction of morphological characters and OTUs from previous studies.
-For example in primates the character \textit{p7} coded first by \cite{ross1998phylogenetic} is reused with the same living species in \cite{seiffert2003fossil}, \cite{marivaux2005anthropoid}, \cite{seiffert2005basal}, \cite{bloch2007new}, \cite{bloch2007new}, \cite{kay2008anatomy}, \cite{silcox2008biogeographic}, \cite{seiffert2009convergent}, \cite{tabuce2009anthropoid}, \cite{boyer2010astragalar}, \cite{seiffert2010fossil}, \cite{marivaux2013djebelemur} and \cite{ni2013oldest}.
-
 For example, the six living primates used in \cite{ross1998phylogenetic} (\textit{Aotus trivirgatus, Galago demidoff, Lemur catta, Microcebus murinus, Nycticebus coucang and Saimiri sciureus}) and their associated characters are reused along with more living species and characters in \cite{seiffert2003fossil}, \cite{marivaux2005anthropoid}, \cite{seiffert2005basal}, \cite{bloch2007new}, \cite{bloch2007new}, \cite{kay2008anatomy}, \cite{silcox2008biogeographic}, \cite{seiffert2009convergent}, \cite{tabuce2009anthropoid}, \cite{boyer2010astragalar}, \cite{seiffert2010fossil}, \cite{marivaux2013djebelemur} and \cite{ni2013oldest}.

### 2016-02-13 Manuscript/BiologyLetters/Table_results.tex
Dealt with comments

@@
 taxonomic levels.
-The left vertical bar represents low coverage (<25\%; coloured in blue); the right vertical bar represents high coverage (>75\%; coloured in orange).
+The left vertical bar represents low coverage (\textless 25\%; coloured in blue); the right vertical bar represents high coverage (\textgreater 75\%; coloured in orange).
 Negative Net Relatedness Index (NRI) and Nearest Taxon Index (NTI) values indicate phylogenetic overdispersion; positive values indicate phylogenetic clustering.
 Significant NRI or NTI values are in bold.
-*p <0.05; **p <0.01; ***p <0.001.
+*p \textless 0.05; **p \textless 0.01; ***p \textless 0.001.
 } \\
+
   \hline

### 2016-02-13 Manuscript/BiologyLetters/response.tex
Minor edits

@@
 The difficulty of obtaining such matrices makes them less likely to be included in Total Evidence analyses, especially where the number of taxa to include in the phylogeny is high.
-'' p.@ lines@@.
+'' p.5 lines 161-168.
 
@@
 We added this analysis to the ESM 2 (Supplementary results - 1) showing, as suggested, that mammalian morphological phylogeneticists have focused their energies on resolving the higher level relationships among mammalian lineages.
-We also added a sentence mentioning these results in the manuscript:
+% We also added a sentence mentioning these results in the manuscript:
 
-``It is also encouraging to note that the sampling effort is consistent with diversity at each taxonomic level and there is therefore no direct bias in sampling at any taxonomic level (ESM 2, Figure 1).
-'' p.@ lines@@
+% ``It is also encouraging to note that the sampling effort is consistent with diversity at each taxonomic level and there is therefore no direct bias in sampling at any taxonomic level (ESM 2, Figure 1).
+% '' p.@ lines@@
 
@@
 ``One increasingly popular method, the Total Evidence method [3], combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix that can then be used with the tip-dating method (e.g.
-[4, 3, 5, 1, 6]), producing a chronogram with living and fossil taxa at the tips.'' p@ lines@@
+[4, 3, 5, 1, 6]), producing a chronogram with living and fossil taxa at the tips.'' p.1 lines 34-37.
 
@@
 We added the following sentence to in the methods:
-``For each analysis our focal taxa were those with available coded anatomical characters at that taxonomic level and the phylogeny was the order pruned from [10].''. p@. lines@@.
+``For each analysis our focal taxa were those with available coded anatomical characters at that taxonomic level and the phylogeny was the order pruned from [10].''. p.3 lines 104-106.
 
@@
 \textit{sp\_nov\_1} or \textit{Outgroup}) or typos (e.g.
-\textit{Hobo sabpiens}).''
+\textit{Hobo sapions}).''
 

### 2016-02-13 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
Minor edits
Edits to ESM1

--- 2016-01-26 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
@@
 
+% NC: It's important to follow up on typo and grammar issues in the supp for things that were mentioned in the main text. For example, dropping https... MorphoBank...
+
 \section{1 - Data collection}
 \subsection{Public repositories}
-We downloaded available matrices containing fossil and/or living mammal taxa from the three data bases using the following list of keywords:
+We downloaded available matrices containing fossil and/or living mammal taxa from the three databases using the following list of keywords:
 
@@
 
-Details about each public repository specific search option is listed below.
-Note that some matrices have been downloaded from more than one database but that it is not an issue since we are interested in the total number of unique living OTUs and that if some were present in more than one matrix, they still only counted as one single OTU.
+Details about the specific search options used for each public repository are listed below.
+Note that some matrices were downloaded from more than one database but this is not a problem because we are interested in the total number of unique living operational taxonomic units (OTUs), therefore even if some were present in more than one matrix they still only counted as a single OTU.
 
-\subsubsection{Morphobank}
-We accessed the Morphobank repository (\url{morphobank.org}) on the 10th of June 2015 and used the keywords listed above in the search menu.
+\subsubsection{MorphoBank}
+We accessed the MorphoBank repository (\url{morphobank.org}) on 10th June 2015 and used the keywords listed above in the search menu.
 We downloaded the data associated with each project matching with the keyword.
@@
 \subsubsection{Graeme Lloyd}
-We accessed Graeme Lloyd's website repository (\url{graemetlloyd.com/matrmamm.html}) on the 10th of June 2015 and downloaded all the matrices that were available with a direct download link in the mammal data section of the website (\texttt{http://graemetlloyd.com/matrmamm.html}).
+We accessed Graeme Lloyd's website repository (\url{graemetlloyd.com/}) on 10th June 2015 and downloaded all the matrices that were available with a direct download link in the mammal data section of the website (\texttt{graemetlloyd.com/matrmamm.html}).
 
 \subsubsection{Ross Mounce}
-We accessed Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}) on the 11th of June 2015 and downloaded all 601 matrix.
-We then ran a shell script to select only the matrices that had any text element that match with one of the search terms.
-% LINK WHICH SHELL SCRIPT
-To make the matrix selection more thorough, we ignored the keywords case as well as the latin suffix (\textit{ia}, \textit{ata}, \textit{ea}, and \textit{a}).
+We accessed Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}) on 11th June 2015 and downloaded all 601 matrices.
+We then ran a shell script to select only the matrices that had any text element that matched with one of the search terms (\url{github.com/TGuillerme/Missing_living_mammals/blob/master/Functions/select.files.sh}).
+To make the matrix selection more thorough, we ignored the case and Latin suffix (i.e. \textit{ia}, \textit{ata}, \textit{ea}, and \textit{a}) of the keywords.
 
-\subsection{Google scholars (accessed 11th of June 2015)}
-To make sure we didn't miss any extra matrix that wasn't available on one of these repositories, we ran a Google Scholar search on the 5th of January. 
-We downloaded the additional cladistic matrices from the 20 first search results matching with our selected keywords and with any of the 35 taxonomic levels (mammals Orders, Infraclasses and Class).
-We used the following key words:
+\subsection{Google Scholar (accessed 11th June 2015)}
+To ensure we did not miss any extra matrices that were not available on one of these repositories, we ran a Google Scholar search on the 5th January % NC ??? It says 11th June above. 
+using the following keywords:
 
@@
 
-were \textit{order} was replaced by all the keywords listed above. For each 33 keywords, we selected the 20 first papers to match the Google search published since 2010 resulting in 660 papers.
-Among these papers, not all contained relevant data (discrete morphological characters AND mammalian data).
-We selected only the 20 first results per search term to avoid downloading articles that were too irrelevant.
-Among the 660 papers, only 50 contained a total of 425 extra living OTUs (Figure ~\ref{Supp_figure_google_searches}).
-Also we decided to select only the articles published since 2010 because nearly every one of the recent published matrix contains both a fraction of morphological characters and OTUs from previous studies.
-For example in primates the character \textit{p7} coded first by \cite{ross1998phylogenetic} is reused with the same living species in \cite{seiffert2003fossil}, \cite{marivaux2005anthropoid}, \cite{seiffert2005basal}, \cite{bloch2007new}, \cite{bloch2007new}, \cite{kay2008anatomy}, \cite{silcox2008biogeographic}, \cite{seiffert2009convergent}, \cite{tabuce2009anthropoid}, \cite{boyer2010astragalar}, \cite{seiffert2010fossil}, \cite{marivaux2013djebelemur} and \cite{ni2013oldest}.
+where \textit{order} was replaced by each of the taxonomic subdivision keywords listed above in turn.
+For each taxonomic subdivision keyword we selected the 20 first papers published since 2010 resulting in 660 papers.
+We selected only the 20 first results for each search term to avoid downloading large numbers of irrelevant articles, and because the rate of discovery of new matrices was very low and unlikely to be substantially improved by downloading more papers.
+For example, in the 660 papers we downloaded, only 50 contained extra living OTUs and only contributed 425 OTUs to our total of 4950 OTUs (Figure ~\ref{Supp_figure_google_searches}).
+We selected only articles published since 2010 because almost every recently published matrix contained some of the morphological characters and OTUs from previous studies, thus almost all older studies are represented in the matrices we collected.
+For example, the six living primates used in \cite{ross1998phylogenetic} (\textit{Aotus trivirgatus, Galago demidoff, Lemur catta, Microcebus murinus, Nycticebus coucang and Saimiri sciureus}) and their associated characters are reused along with more living species and characters in \cite{seiffert2003fossil, marivaux2005anthropoid, seiffert2005basal, bloch2007new,bloch2007new, kay2008anatomy, silcox2008biogeographic, seiffert2009convergent,tabuce2009anthropoid, boyer2010astragalar, seiffert2010fossil, marivaux2013djebelemur, ni2013oldest}.
 
@@
     \includegraphics[width=1\textwidth]{Supp_figure_google_searches.pdf}
-\caption{Google searches additional OTUs rarefaction curve.
-The x-axis represent the number of Google Scholar matches (papers, books or abstracts) and the y-axis represents the cumulative number of additional living OTUs per google scholar match.}
+\caption{Google Scholar searches additional OTUs rarefaction curve.
+The x-axis represents the number of Google Scholar matches (papers, books or abstracts) and the y-axis represents the cumulative number of additional living OTUs for each Google Scholar match.}
 \label{Supp_figure_google_searches}
@@
 
-The list of all the 286 download matrices is available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices}.
+The list of all 286 downloaded matrices is available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices}.
 The matrices contained a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique.
@@
 \subsection{Standardising the matrices}
-We transformed all the non-nexus matrices (tnt, word, excel, jpeg) to nexus format manually.
-We then cleaned the nexus matrices by removing any extra information (trees, continuous characters, morphological characters description, molecular data) to end up with nexus matrices containing only the discrete morphological data.
-We then manually fixed the incorrectly-formatted binomial names (e.g. \textit{H. sapiens} became \textit{Homo sapiens}) using the abbreviation list in the concerned publications. 
+We transformed all the non-NEXUS matrices (TNT, Word, Excel, JPEG) to NEXUS format manually.
+We then cleaned the NEXUS matrices by removing any extra information (trees, continuous characters, morphological character descriptions, molecular data) to end up with NEXUS matrices containing only the discrete morphological data.
+We then manually fixed the incorrectly-formatted binomial names (e.g. \textit{H. sapiens} became \textit{Homo sapiens}) using the abbreviation list in the relevant publications. 
 All the standardised matrices are available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices_binomial/Matrices}.
@@
 \subsection{Selecting the living OTUs}
-%To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H.
-sapiens} was transformed to \textit{Homo sapiens}).
-We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{bininda-emondsthe2007} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\url{https://paleobiodb.org/}).
+We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{bininda-emondsthe2007} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\url{paleobiodb.org/}).
 For OTUs that did not appear in these three sources, we first decomposed the name (i.e.
@@
 Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; Figure ~\ref{Supp_figure_Taxonomic_algorithm}).
+Non-applicable OTUs were either specimen IDs with no related taxon names (e.g.
+\textit{FMNHPR2081}), abbreviations that were not described in the associated paper (e.g.
+\textit{Ho.sap.}), non-mammals \textit{stricto-sensu} (e.g.
+\textit{Sinoconodon}), non standard taxonomic levels (e.g.
+\textit{Spalcotheriids}) invalid taxonomic designations (e.g.
+\textit{sp\_nov\_1} or \textit{Outgroup}) or typos (e.g.
+\textit{Hobo sapions}).
 
@@
 \caption{Taxonomic matching algorithm used in this study.
-For each matrix, each operational taxonomic units (OTU) is matched with the super tree from Fritz 2009.
+For each matrix, each operational taxonomic unit (OTU) is matched with the supertree from Fritz et al.
+2009.
 If the OTU matches, then it is classified as living.
-Else it is matched with the Wilson \& Reeders 2005 taxonomy list.
+Otherwise it is matched with the Wilson \& Reeder 2005 mammalian taxonomy.
 If the OTU matches, then it is classified as living.
-Else it is matched with the Paleo Database list of mammals.
+Otherwise it is matched with the Paleobiology database list of mammals.
 If the OTU matches, then it is classified as fossil.
-Else it is ignored.}
+Otherwise it is ignored.}
 \label{Supp_figure_Taxonomic_algorithm}
@@
 
-\section{2 - Data reproducibility}
+\section{2 - Data collection reproducibility}
 Every step of the analysis (apart from downloading and standardisation of the matrices) is entirely repeatable via GitHub (\url{github.com/TGuillerme/Missing_living_mammals}).
@@
 
+\newpage
+
 \bibliographystyle{vancouver}

### 2016-02-13 Manuscript/BiologyLetters/Supplementary/ESM_2.tex
Edits to ESM2

--- 2016-01-26 Manuscript/BiologyLetters/Supplementary/ESM_2.tex
@@
 
+\section{Supplementary results - 1}
+The following section contains a supplementary analysis looking at the sampling effort for mammalian orders at three different taxonomic levels (family, genus and species).
+We fitted linear regressions %TG: does one calculate regressions? I would go for drawing regressions but that's not semantically correct... % NC: It's a model so you fit it
+between the log of the number of sampled families/genera/species for each order against the log of the total number of families/genera/species in each order % NC: Taken from where?
+The taxonomy?
+Where do you get this total from?
+and fixed the intercept at zero (because when no families/genera/species exist they cannot be sampled).
+The slope of the regression gives an indication of the sampling effort: if the slope is 1, the number of sampled families/genera/species and the total number of families/genera/species are equal, i.e.
+all families/genera/species have been sampled.
+Slopes of less than 1 indicate lower sampling effort.
+The code for this analysis is available here: \url{github.com/TGuillerme/Missing_living_mammals/blob/master/Analysis/Sampling_effort.R}.
+
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Supp_figure_sampling_effort.pdf}
+\caption{Sampling effort across mammalian orders at the three different taxonomic levels.
+The \textit{R$^2$} and the slope (\textit{b}, forced through the origin) are reported for each regression.} % NC: Should report adjusted r2 because sample size is small.
+\label{Supp_Figure_sampling}
+\end{figure}
+
+% NC: Fix the y tick labels (use las = 1 to make them horizontal), what's up with panel c?
+the axes should be the same size as the others and why doesn't the x axis cover the range of the data?
+Also what are the three panels?
+explain in the legend and maybe label as a,b,c?
+Also axes labels should be log(sampled families) etc.
+%Supp mat is no excuse for sloppy figures.
+
+This analysis suggests that sampling effort effectively reflects diversity for each taxonomic level, i.e.
+morphological characters were collected for each order according to its size (Figure \ref{Supp_Figure_sampling}).
+However, it is worth noting that this relation is stronger at higher taxonomic levels (i.e.
+family-level, \textit{R$^2$}=0.99 \textit{b}=0.936) compared to the species-level (\textit{R$^2$}=0.852 \textit{b}=0.711).
+These results might be simply due to the increasing number of groups at each taxonomic level (i.e. number of species $>$ genera $>$ families).
+But it is also probable that mammalian morphological phylogeneticists have focused on resolving higher level relationships among mammalian orders (see Table 1 in the main text), resulting in better coverage at the family-level.
+
+\newpage
+
 \section{Supplementary results - 2}
-The following section contains supplementary results to the main body: the phylogenetic representation of the data availability per order (excluding Primates and Carnivora, present in the main body).
+The following section contains phylogenetic representations of availability of coded anatomical characters for each mammalian order (excluding results for Primates and Carnivora that are present in the main text).
 
@@
     \includegraphics[width=1\textwidth]{Supp_figure_AFROSORICIDA.pdf}
-\caption{Distribution of available morphological data across Afrosoricida. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\caption{Phylogenetic distribution of species with available coded anatomical characters across Afrosoricida. Blue branches indicate species with available coded anatomical characters.}
 \label{Supp_Figure_Phylo-Afrosoricida}
 \end{figure}
+
+% NC: I have copied this wording from the main text. Fix all the ones below please :)
 

### 2016-02-14 Analysis/Sampling_effort.R
Updated sampling effort graph

@@ -57,6 +57,9 @@ living_taxa_list<-unique(living_taxa$Taxa)
 #Loading the results
 load("results_1.Rda")
 
+#Remove the whole class data (first 3 rows)
+results <- results[-c(1:3),]
+
 ######################
 #Sampling effort
 ######################
@@ -79,34 +82,40 @@ par(mfrow = (c(1,3)), bty = "n")
 total <- total_families
 sampled <- sampled_families
 model <- summary(lm(sampled + 0 ~ total))
-plot(total, sampled, pch=19, col="grey", xlim = c(0,max(total)), ylim = c(0,max(total)),
-    ylab = "sampled families (log)", xlab = "total families (log)")
+plot(total, sampled, pch=19, col="grey", xlim = c(0,ceiling(max(total))), ylim = c(0,ceiling(max(total))),
+#plot(total, sampled, pch=19, col="grey", xlim = c(0,8), ylim = c(0,8),
+    ylab = "sampled families (log)", xlab = "total families (log)", las = 1)
 abline(model$coefficients[[1]], model$coefficients[[2]])
-text(max(total) - 0.85*max(total), max(total) - 0.1*max(total),
-    bquote(italic(R)^2 == .(format(model$r.squared, digits = 3))) )
-text(max(total) - 0.85*max(total), max(total) - 0.15*max(total),
+text(ceiling(max(total)) - 0.85*ceiling(max(total)), ceiling(max(total)) - 0.035*ceiling(max(total)), "A", cex=2)
+text(ceiling(max(total)) - 0.85*ceiling(max(total)), ceiling(max(total)) - 0.1*ceiling(max(total)),
+    bquote(italic(R)^2 == .(format(model$adj.r.squared, digits = 3))) )
+text(ceiling(max(total)) - 0.85*ceiling(max(total)), ceiling(max(total)) - 0.15*ceiling(max(total)),
     bquote(italic(b) == .(format(model$coefficients[[2]], digits = 3))))
 
 total <- total_genera
 sampled <- sampled_genera
 model <- summary(lm(sampled + 0 ~ total))
-plot(total, sampled, pch=19, col="grey", xlim = c(0,max(total)), ylim = c(0,max(total)),
-    ylab = "sampled genera (log)", xlab = "total genera (log)")
+plot(total, sampled, pch=19, col="grey", xlim = c(0,ceiling(max(total))), ylim = c(0,ceiling(max(total))),
[…]

### 2016-02-14 Manuscript/BiologyLetters/Table_LaTeX/Table_results_BW.tex
Table latex

% latex table generated in R 3.2.3 by xtable 1.8-0 package
% Mon Feb  8 22:09:15 2016
\begin{longtable}{lL{1.8cm}C{2cm}lcc}
\caption{Number of taxa with available cladistic data for mammalian orders at three
taxonomic levels. The left vertical bar represents low coverage (\textless 25\%; coloured in blue); the right vertical bar represents high coverage (\textgreater 75\%; coloured in orange). Negative Net Relatedness Index (NRI) and Nearest Taxon Index (NTI) values indicate phylogenetic overdispersion; positive values indicate phylogenetic clustering. Significant NRI or NTI values are in bold. *p \textless 0.05; **p \textless 0.01; ***p \textless 0.001.
} \\

  \hline
Order & Taxonomic level & Proportion of taxa & Coverage & NRI & NTI \\ 
  \hline
  Mammalia (class) & family & 129/148 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar40.pdf} & -1.19 & 1.09 \\ 
  \textbf{Mammalia (class)} & \textbf{genus} & \textbf{517/1186} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar41.pdf} & \textbf{-5.19} & \textbf{3.71**} \\ 
  \textbf{Mammalia (class)} & \textbf{species} & \textbf{847/5017} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar42.pdf} & \textbf{-7.75} & \textbf{3.54**} \\ 
  Afrosoricida & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar1.pdf} &   &   \\ 
  Afrosoricida & genus & 17/17 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar2.pdf} &   &   \\ 
  Afrosoricida & species & 23/42 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar3.pdf} & 1.52 & 1.1 \\ 
  Carnivora & family & 14/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar4.pdf} & 0.65 & 0.55 \\ 
  \textbf{Carnivora} & \textbf{genus} & \textbf{52/125} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar5.pdf} & \textbf{4.27**} & \textbf{1.26} \\ 
  \textbf{Carnivora} & \textbf{species} & \textbf{75/283} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar6.pdf} & \textbf{7.24**} & \textbf{0.8} \\ 
  Cetartiodactyla & family & 21/21 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar7.pdf} &   &   \\ 
  Cetartiodactyla & genus & 97/128 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar8.pdf} & 0.7 & 1.28 \\ 
  \textbf{Cetartiodactyla} & \textbf{species} & \textbf{169/310} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar9.pdf} & \textbf{1.82*} & \textbf{-0.24} \\ 
  Chiroptera & family & 15/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar10.pdf} & -0.23 & 0.61 \\ 
  \textbf{Chiroptera} & \textbf{genus} & \textbf{92/202} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar11.pdf} & \textbf{13.07**} & \textbf{0.99} \\ 
  \textbf{Chiroptera} & \textbf{species} & \textbf{214/1053} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar12.pdf} & \textbf{9.21**} & \textbf{1.27} \\ 
  Cingulata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar13.pdf} &   &   \\ 
  Cingulata & genus & 8/9 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar14.pdf} & 1.48 & -1.54 \\ 
  \textbf{Cingulata} & \textbf{species} & \textbf{9/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar15.pdf} & \textbf{2.06*} & \textbf{0.2} \\ 
  Dasyuromorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar16.pdf} &   &   \\ 
  Dasyuromorphia & genus & 8/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar17.pdf} & -0.78 & -1.06 \\ 
  Dasyuromorphia & species & 9/64 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar18.pdf} & -0.86 & -0.37 \\ 
  Dermoptera & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar19.pdf} &   &   \\ 
  Dermoptera & genus & 1/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar20.pdf} &   &   \\ 
  Dermoptera & species & 1/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar21.pdf} &   &   \\ 
  Didelphimorphia & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar22.pdf} &   &   \\ 
  Didelphimorphia & genus & 16/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar23.pdf} &   &   \\ 
  Didelphimorphia & species & 42/84 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar24.pdf} & -1.61 & 0.12 \\ 
  Diprotodontia & family & 11/11 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar25.pdf} &   &   \\ 
  Diprotodontia & genus & 25/38 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar26.pdf} & -1.15 & -1.33 \\ 
  Diprotodontia & species & 31/126 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar27.pdf} & 0.44 & -1.79 \\ 
  Erinaceomorpha & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar28.pdf} &   &   \\ 
  Erinaceomorpha & genus & 10/10 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar29.pdf} &   &   \\ 
  Erinaceomorpha & species & 21/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar30.pdf} & -1.04 & -0.25 \\ 
  Hyracoidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar31.pdf} &   &   \\ 
  Hyracoidea & genus & 1/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar32.pdf} &   &   \\ 
  Hyracoidea & species & 1/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar33.pdf} &   &   \\ 
  Lagomorpha & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar34.pdf} &   &   \\ 
  Lagomorpha & genus & 5/12 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar35.pdf} & -0.95 & -0.94 \\ 
  Lagomorpha & species & 12/86 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar36.pdf} & -0.62 & -1.96 \\ 
  Macroscelidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar37.pdf} &   &   \\ 
  Macroscelidea & genus & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar38.pdf} &   &   \\ 
  Macroscelidea & species & 12/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar39.pdf} & -1.24 & -1.2 \\ 
  Microbiotheria & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar43.pdf} &   &   \\ 
  Microbiotheria & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar44.pdf} &   &   \\ 
  Microbiotheria & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar45.pdf} &   &   \\ 
  Monotremata & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar46.pdf} &   &   \\ 
  Monotremata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar47.pdf} & -0.68 & -0.69 \\ 
  Monotremata & species & 2/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar48.pdf} & -1.01 & -1 \\ 
  Notoryctemorphia & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar49.pdf} &   &   \\ 
  Notoryctemorphia & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar50.pdf} &   &   \\ 
  Notoryctemorphia & species & 0/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar51.pdf} &   &   \\ 
  Paucituberculata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar52.pdf} &   &   \\ 
  Paucituberculata & genus & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar53.pdf} &   &   \\ 
  Paucituberculata & species & 5/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar54.pdf} &   &   \\ 
  Peramelemorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar55.pdf} &   &   \\ 
  Peramelemorphia & genus & 7/7 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar56.pdf} &   &   \\ 
  Peramelemorphia & species & 16/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar57.pdf} & -0.14 & 0.91 \\ 
  Perissodactyla & family & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar58.pdf} &   &   \\ 
  Perissodactyla & genus & 6/6 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar59.pdf} &   &   \\ 
  Perissodactyla & species & 10/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar60.pdf} & -0.1 & -2.77 \\ 
  Pholidota & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar61.pdf} &   &   \\ 
  Pholidota & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar62.pdf} &   &   \\ 
  Pholidota & species & 4/8 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar63.pdf} & 1.14 & 0.97 \\ 
  Pilosa & family & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar64.pdf} & 2.01 & 1.96 \\ 
  Pilosa & genus & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar65.pdf} & -0.91 & 0.36 \\ 
  \textbf{Pilosa} & \textbf{species} & \textbf{5/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar66.pdf} & \textbf{1.18} & \textbf{2.35**} \\ 
  Primates & family & 15/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar67.pdf} &   &   \\ 
  Primates & genus & 48/68 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar68.pdf} & -0.37 & -1.39 \\ 
  Primates & species & 64/351 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar69.pdf} & -0.66 & -1.4 \\ 
  Proboscidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar70.pdf} &   &   \\ 
  Proboscidea & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar71.pdf} &   &   \\ 
  Proboscidea & species & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar72.pdf} & -0.67 & -0.72 \\ 
  Rodentia & family & 18/32 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar73.pdf} & 0.66 & -0.95 \\ 
  \textbf{Rodentia} & \textbf{genus} & \textbf{82/450} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar74.pdf} & \textbf{-1.81} & \textbf{1.7*} \\ 
  \textbf{Rodentia} & \textbf{species} & \textbf{90/2094} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar75.pdf} & \textbf{2.66**} & \textbf{2.36**} \\ 
  Scandentia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar76.pdf} &   &   \\ 
  Scandentia & genus & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar77.pdf} & -0.77 & -0.76 \\ 
  Scandentia & species & 3/20 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar78.pdf} & -2 & -0.8 \\ 
  Sirenia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar79.pdf} &   &   \\ 
  Sirenia & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar80.pdf} &   &   \\ 
  Sirenia & species & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar81.pdf} &   &   \\ 
  Soricomorpha & family & 3/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar82.pdf} & -0.98 & -0.97 \\ 
  \textbf{Soricomorpha} & \textbf{genus} & \textbf{19/43} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar83.pdf} & \textbf{7.07**} & \textbf{2.64**} \\ 
  \textbf{Soricomorpha} & \textbf{species} & \textbf{21/392} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar84.pdf} & \textbf{10.17**} & \textbf{3.36**} \\ 
  Tubulidentata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar85.pdf} &   &   \\ 
  Tubulidentata & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar86.pdf} &   &   \\ 
  Tubulidentata & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_BW/bar87.pdf} &   &   \\ 
   \hline
\hline
\label{Table_results}
\end{longtable}

### 2016-02-14 Manuscript/BiologyLetters/Table_LaTeX/Table_results_color.tex
Table latex

% latex table generated in R 3.2.3 by xtable 1.8-0 package
% Mon Feb  8 22:09:15 2016
\begin{longtable}{lL{1.8cm}C{2cm}lcc}
\caption{Number of taxa with available cladistic data for mammalian orders at three
taxonomic levels. The left vertical bar represents low coverage (\textless 25\%; coloured in blue); the right vertical bar represents high coverage (\textgreater 75\%; coloured in orange). Negative Net Relatedness Index (NRI) and Nearest Taxon Index (NTI) values indicate phylogenetic overdispersion; positive values indicate phylogenetic clustering. Significant NRI or NTI values are in bold. *p \textless 0.05; **p \textless 0.01; ***p \textless 0.001.
} \\

  \hline
Order & Taxonomic level & Proportion of taxa & Coverage & NRI & NTI \\ 
  \hline
  Mammalia (class) & family & 129/148 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar40.pdf} & -1.19 & 1.09 \\ 
  \textbf{Mammalia (class)} & \textbf{genus} & \textbf{517/1186} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar41.pdf} & \textbf{-5.19} & \textbf{3.71**} \\ 
  \textbf{Mammalia (class)} & \textbf{species} & \textbf{847/5017} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar42.pdf} & \textbf{-7.75} & \textbf{3.54**} \\ 
  Afrosoricida & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar1.pdf} &   &   \\ 
  Afrosoricida & genus & 17/17 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar2.pdf} &   &   \\ 
  Afrosoricida & species & 23/42 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar3.pdf} & 1.52 & 1.1 \\ 
  Carnivora & family & 14/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar4.pdf} & 0.65 & 0.55 \\ 
  \textbf{Carnivora} & \textbf{genus} & \textbf{52/125} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar5.pdf} & \textbf{4.27**} & \textbf{1.26} \\ 
  \textbf{Carnivora} & \textbf{species} & \textbf{75/283} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar6.pdf} & \textbf{7.24**} & \textbf{0.8} \\ 
  Cetartiodactyla & family & 21/21 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar7.pdf} &   &   \\ 
  Cetartiodactyla & genus & 97/128 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar8.pdf} & 0.7 & 1.28 \\ 
  \textbf{Cetartiodactyla} & \textbf{species} & \textbf{169/310} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar9.pdf} & \textbf{1.82*} & \textbf{-0.24} \\ 
  Chiroptera & family & 15/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar10.pdf} & -0.23 & 0.61 \\ 
  \textbf{Chiroptera} & \textbf{genus} & \textbf{92/202} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar11.pdf} & \textbf{13.07**} & \textbf{0.99} \\ 
  \textbf{Chiroptera} & \textbf{species} & \textbf{214/1053} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar12.pdf} & \textbf{9.21**} & \textbf{1.27} \\ 
  Cingulata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar13.pdf} &   &   \\ 
  Cingulata & genus & 8/9 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar14.pdf} & 1.48 & -1.54 \\ 
  \textbf{Cingulata} & \textbf{species} & \textbf{9/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar15.pdf} & \textbf{2.06*} & \textbf{0.2} \\ 
  Dasyuromorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar16.pdf} &   &   \\ 
  Dasyuromorphia & genus & 8/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar17.pdf} & -0.78 & -1.06 \\ 
  Dasyuromorphia & species & 9/64 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar18.pdf} & -0.86 & -0.37 \\ 
  Dermoptera & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar19.pdf} &   &   \\ 
  Dermoptera & genus & 1/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar20.pdf} &   &   \\ 
  Dermoptera & species & 1/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar21.pdf} &   &   \\ 
  Didelphimorphia & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar22.pdf} &   &   \\ 
  Didelphimorphia & genus & 16/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar23.pdf} &   &   \\ 
  Didelphimorphia & species & 42/84 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar24.pdf} & -1.61 & 0.12 \\ 
  Diprotodontia & family & 11/11 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar25.pdf} &   &   \\ 
  Diprotodontia & genus & 25/38 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar26.pdf} & -1.15 & -1.33 \\ 
  Diprotodontia & species & 31/126 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar27.pdf} & 0.44 & -1.79 \\ 
  Erinaceomorpha & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar28.pdf} &   &   \\ 
  Erinaceomorpha & genus & 10/10 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar29.pdf} &   &   \\ 
  Erinaceomorpha & species & 21/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar30.pdf} & -1.04 & -0.25 \\ 
  Hyracoidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar31.pdf} &   &   \\ 
  Hyracoidea & genus & 1/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar32.pdf} &   &   \\ 
  Hyracoidea & species & 1/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar33.pdf} &   &   \\ 
  Lagomorpha & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar34.pdf} &   &   \\ 
  Lagomorpha & genus & 5/12 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar35.pdf} & -0.95 & -0.94 \\ 
  Lagomorpha & species & 12/86 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar36.pdf} & -0.62 & -1.96 \\ 
  Macroscelidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar37.pdf} &   &   \\ 
  Macroscelidea & genus & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar38.pdf} &   &   \\ 
  Macroscelidea & species & 12/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar39.pdf} & -1.24 & -1.2 \\ 
  Microbiotheria & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar43.pdf} &   &   \\ 
  Microbiotheria & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar44.pdf} &   &   \\ 
  Microbiotheria & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar45.pdf} &   &   \\ 
  Monotremata & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar46.pdf} &   &   \\ 
  Monotremata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar47.pdf} & -0.68 & -0.69 \\ 
  Monotremata & species & 2/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar48.pdf} & -1.01 & -1 \\ 
  Notoryctemorphia & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar49.pdf} &   &   \\ 
  Notoryctemorphia & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar50.pdf} &   &   \\ 
  Notoryctemorphia & species & 0/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar51.pdf} &   &   \\ 
  Paucituberculata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar52.pdf} &   &   \\ 
  Paucituberculata & genus & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar53.pdf} &   &   \\ 
  Paucituberculata & species & 5/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar54.pdf} &   &   \\ 
  Peramelemorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar55.pdf} &   &   \\ 
  Peramelemorphia & genus & 7/7 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar56.pdf} &   &   \\ 
  Peramelemorphia & species & 16/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar57.pdf} & -0.14 & 0.91 \\ 
  Perissodactyla & family & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar58.pdf} &   &   \\ 
  Perissodactyla & genus & 6/6 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar59.pdf} &   &   \\ 
  Perissodactyla & species & 10/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar60.pdf} & -0.1 & -2.77 \\ 
  Pholidota & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar61.pdf} &   &   \\ 
  Pholidota & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar62.pdf} &   &   \\ 
  Pholidota & species & 4/8 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar63.pdf} & 1.14 & 0.97 \\ 
  Pilosa & family & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar64.pdf} & 2.01 & 1.96 \\ 
  Pilosa & genus & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar65.pdf} & -0.91 & 0.36 \\ 
  \textbf{Pilosa} & \textbf{species} & \textbf{5/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar66.pdf} & \textbf{1.18} & \textbf{2.35**} \\ 
  Primates & family & 15/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar67.pdf} &   &   \\ 
  Primates & genus & 48/68 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar68.pdf} & -0.37 & -1.39 \\ 
  Primates & species & 64/351 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar69.pdf} & -0.66 & -1.4 \\ 
  Proboscidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar70.pdf} &   &   \\ 
  Proboscidea & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar71.pdf} &   &   \\ 
  Proboscidea & species & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar72.pdf} & -0.67 & -0.72 \\ 
  Rodentia & family & 18/32 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar73.pdf} & 0.66 & -0.95 \\ 
  \textbf{Rodentia} & \textbf{genus} & \textbf{82/450} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar74.pdf} & \textbf{-1.81} & \textbf{1.7*} \\ 
  \textbf{Rodentia} & \textbf{species} & \textbf{90/2094} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar75.pdf} & \textbf{2.66**} & \textbf{2.36**} \\ 
  Scandentia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar76.pdf} &   &   \\ 
  Scandentia & genus & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar77.pdf} & -0.77 & -0.76 \\ 
  Scandentia & species & 3/20 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar78.pdf} & -2 & -0.8 \\ 
  Sirenia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar79.pdf} &   &   \\ 
  Sirenia & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar80.pdf} &   &   \\ 
  Sirenia & species & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar81.pdf} &   &   \\ 
  Soricomorpha & family & 3/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar82.pdf} & -0.98 & -0.97 \\ 
  \textbf{Soricomorpha} & \textbf{genus} & \textbf{19/43} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar83.pdf} & \textbf{7.07**} & \textbf{2.64**} \\ 
  \textbf{Soricomorpha} & \textbf{species} & \textbf{21/392} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar84.pdf} & \textbf{10.17**} & \textbf{3.36**} \\ 
  Tubulidentata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar85.pdf} &   &   \\ 
  Tubulidentata & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar86.pdf} &   &   \\ 
  Tubulidentata & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures_color/bar87.pdf} &   &   \\ 
   \hline
\hline
\label{Table_results}
\end{longtable}

### 2016-02-14 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
ESM

@@
 \end{center}
-%\modulolinenumbers[1]
-%\linenumbers
 
@@
 
-%TG: Maybe useless, I'll add that just in the thesis.
-%\section{Graphical abstract}
-
-%\begin{figure}[!htbp]
-%\centering
-%    \includegraphics[width=1\textwidth]{MissingDataFigure.pdf}
-%\caption{Example of topological errors due to missing morphological data in living taxa.
-%The phylogeny contains two clades, Aves and Mammalia, with molecular data (grey) for both but only morphological data (red) for Aves.
-%If a mammalian fossil species (with no molecular data) is added to the phylogeny, it will erroneously branch with the Aves instead of the Mammalia because no morphological data will overlap between the fossil mammals and the living ones.}
-%\label{Figure_missing_data_problem}
-%\end{figure}
-
+% TO DO!
 % NC: It's important to follow up on typo and grammar issues in the supp for things that were mentioned in the main text. For example, dropping https... MorphoBank...
@@
 \subsection{Google Scholar (accessed 11th June 2015)}
-To ensure we did not miss any extra matrices that were not available on one of these repositories, we ran a Google Scholar search on the 5th January % NC ??? It says 11th June above. 
-using the following keywords:
+To ensure we did not miss any extra matrices that were not available on one of these repositories, we ran a Google Scholar search on the 11th June using the following keywords:
 
@@
 For example, in the 660 papers we downloaded, only 50 contained extra living OTUs and only contributed 425 OTUs to our total of 4950 OTUs (Figure ~\ref{Supp_figure_google_searches}).
-We selected only articles published since 2010 because almost every recently published matrix contained some of the morphological characters and OTUs from previous studies, thus almost all older studies are represented in the matrices we collected.
+We selected only articles published since 2010 because almost every recently published matrices contained some of the morphological characters and OTUs from previous studies, thus almost all older studies are represented in the matrices we collected.
 For example, the six living primates used in \cite{ross1998phylogenetic} (\textit{Aotus trivirgatus, Galago demidoff, Lemur catta, Microcebus murinus, Nycticebus coucang and Saimiri sciureus}) and their associated characters are reused along with more living species and characters in \cite{seiffert2003fossil, marivaux2005anthropoid, seiffert2005basal, bloch2007new,bloch2007new, kay2008anatomy, silcox2008biogeographic, seiffert2009convergent,tabuce2009anthropoid, boyer2010astragalar, seiffert2010fossil, marivaux2013djebelemur, ni2013oldest}.
@@
 Every step of the analysis (apart from downloading and standardisation of the matrices) is entirely repeatable via GitHub (\url{github.com/TGuillerme/Missing_living_mammals}).
-%Needs to fix this in GitHub
-%1-Rarefaction ?
-%2-Selecting the OTUs
-%3-Analysing the availability and the data structure
 

### 2016-02-14 Manuscript/BiologyLetters/Supplementary/ESM_2.tex
ESM

@@
 
-%TG: Maybe useless, I'll add that just in the thesis.
-%\section{Graphical abstract}
-
-%\begin{figure}[!htbp]
-%\centering
-%    \includegraphics[width=1\textwidth]{MissingDataFigure.pdf}
-%\caption{Example of topological errors due to missing morphological data in living taxa.
-%The phylogeny contains two clades, Aves and Mammalia, with molecular data (grey) for both but only morphological data (red) for Aves.
-%If a mammalian fossil species (with no molecular data) is added to the phylogeny, it will erroneously branch with the Aves instead of the Mammalia because no morphological data will overlap between the fossil mammals and the living ones.}
-%\label{Figure_missing_data_problem}
-%\end{figure}
-
 \section{Supplementary results - 1}
 The following section contains a supplementary analysis looking at the sampling effort for mammalian orders at three different taxonomic levels (family, genus and species).
-We fitted linear regressions %TG: does one calculate regressions? I would go for drawing regressions but that's not semantically correct... % NC: It's a model so you fit it
-between the log of the number of sampled families/genera/species for each order against the log of the total number of families/genera/species in each order % NC: Taken from where?
-The taxonomy?
-Where do you get this total from?
-and fixed the intercept at zero (because when no families/genera/species exist they cannot be sampled).
+We fitted linear regressions between the log of the number of sampled families/genera/species for each order against the log of the total number of families/genera/species in each order from \cite{bininda-emondsthe2007} and fixed the intercept at zero (because when no families/genera/species exist they cannot be sampled).
 The slope of the regression gives an indication of the sampling effort: if the slope is 1, the number of sampled families/genera/species and the total number of families/genera/species are equal, i.e.
@@
     \includegraphics[width=1\textwidth]{Supp_figure_sampling_effort.pdf}
-\caption{Sampling effort across mammalian orders at the three different taxonomic levels.
-The \textit{R$^2$} and the slope (\textit{b}, forced through the origin) are reported for each regression.} % NC: Should report adjusted r2 because sample size is small.
+\caption{Sampling effort across mammalian orders at the family (A), genus (B) and species (C) levels.
+The \textit{R$^2$} (adjusted) and the slope (\textit{b}, forced through the origin) are reported for each regression.}
 \label{Supp_Figure_sampling}
 \end{figure}
-
-% NC: Fix the y tick labels (use las = 1 to make them horizontal), what's up with panel c?
-the axes should be the same size as the others and why doesn't the x axis cover the range of the data?
-Also what are the three panels?
-explain in the legend and maybe label as a,b,c?
-Also axes labels should be log(sampled families) etc.
-%Supp mat is no excuse for sloppy figures.
 
@@
 However, it is worth noting that this relation is stronger at higher taxonomic levels (i.e.
-family-level, \textit{R$^2$}=0.99 \textit{b}=0.936) compared to the species-level (\textit{R$^2$}=0.852 \textit{b}=0.711).
+family-level, adjusted \textit{R$^2$}=0.99 \textit{b}=0.936) compared to the species-level (\textit{R$^2$}=0.846 \textit{b}=0.71).
 These results might be simply due to the increasing number of groups at each taxonomic level (i.e. number of species $>$ genera $>$ families).
 But it is also probable that mammalian morphological phylogeneticists have focused on resolving higher level relationships among mammalian orders (see Table 1 in the main text), resulting in better coverage at the family-level.
+
+\bibliographystyle{vancouver}
+\bibliography{Supp_References}
 
@@
 
-% NC: I have copied this wording from the main text. Fix all the ones below please :)
-
 \begin{figure}[!htbp]
@@
     \includegraphics[width=1\textwidth]{Supp_figure_CHIROPTERA.pdf}
-\caption{Distribution of available morphological data across Chiroptera. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\caption{Phylogenetic distribution of species with available coded anatomical characters Chiroptera. Blue branches indicate species with available coded anatomical characters}
 \label{Supp_Figure_Phylo-Chiroptera}
@@
     \includegraphics[width=1\textwidth]{Supp_figure_CINGULATA.pdf}
-\caption{Distribution of available morphological data across Cingulata. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\caption{Phylogenetic distribution of species with available coded anatomical characters Cingulata. Blue branches indicate species with available coded anatomical characters}
 \label{Supp_Figure_Phylo-Cingulata}
@@
     \includegraphics[width=1\textwidth]{Supp_figure_DASYUROMORPHIA.pdf}
-\caption{Distribution of available morphological data across Dasyuromorphia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\caption{Phylogenetic distribution of species with available coded anatomical characters Dasyuromorphia. Blue branches indicate species with available coded anatomical characters}
 \label{Supp_Figure_Phylo-Dasyuromorphia}
@@
     \includegraphics[width=1\textwidth]{Supp_figure_DIDELPHIMORPHIA.pdf}
-\caption{Distribution of available morphological data across Didelphimorphia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\caption{Phylogenetic distribution of species with available coded anatomical characters Didelphimorphia. Blue branches indicate species with available coded anatomical characters}
 \label{Supp_Figure_Phylo-Didelphimorphia}
@@
     \includegraphics[width=1\textwidth]{Supp_figure_DIPROTODONTIA.pdf}
-\caption{Distribution of available morphological data across Diprotodontia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\caption{Phylogenetic distribution of species with available coded anatomical characters Diprotodontia. Blue branches indicate species with available coded anatomical characters}
 \label{Supp_Figure_Phylo-Diprotodontia}
@@
     \includegraphics[width=1\textwidth]{Supp_figure_ERINACEOMORPHA.pdf}
-\caption{Distribution of available morphological data across Erinaceomorpha. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\caption{Phylogenetic distribution of species with available coded anatomical characters Erinaceomorpha. Blue branches indicate species with available coded anatomical characters}
 \label{Supp_Figure_Phylo-Erinaceomorpha}
@@
     \includegraphics[width=1\textwidth]{Supp_figure_PILOSA.pdf}
-\caption{Distribution of available morphological data across Pilosa. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\caption{Phylogenetic distribution of species with available coded anatomical characters Pilosa. Blue branches indicate species with available coded anatomical characters}
 \label{Supp_Figure_Phylo-Pilosa}
@@
     \includegraphics[width=1\textwidth]{Supp_figure_CETARTIODACTYLA.pdf}
-\caption{Distribution of available morphological data across Cetartiodactyla. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\caption{Phylogenetic distribution of species with available coded anatomical characters Cetartiodactyla. Blue branches indicate species with available coded anatomical characters}
 \label{Supp_Figure_Phylo-Primates}
@@
     \includegraphics[width=1\textwidth]{Supp_figure_RODENTIA.pdf}
-\caption{Distribution of available morphological data across Rodentia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\caption{Phylogenetic distribution of species with available coded anatomical characters Rodentia. Blue branches indicate species with available coded anatomical characters}
 \label{Supp_Figure_Phylo-Rodentia}
@@
     \includegraphics[width=1\textwidth]{Supp_figure_SCANDENTIA.pdf}
-\caption{Distribution of available morphological data across Scandentia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\caption{Phylogenetic distribution of species with available coded anatomical characters Scandentia. Blue branches indicate species with available coded anatomical characters}
 \label{Supp_Figure_Phylo-Scandentia}
@@
     \includegraphics[width=1\textwidth]{Supp_figure_SORICOMORPHA.pdf}
-\caption{Distribution of available morphological data across Soricomorpha. Edges are colored in grey when no morphological data is available or in blue when data is available.}
+\caption{Phylogenetic distribution of species with available coded anatomical characters Soricomorpha. Blue branches indicate species with available coded anatomical characters}
 \label{Supp_Figure_Phylo-Soricomorpha}

### 2016-03-29 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Table latex
Updated minor reviews from reviewer 2 and 4 + title change (reviewer 3)

@@
 
-\noindent{\Large \bf Assessment of available anatomical characters for phylogenetic analysis among living mammals}%Assessment of cladistic data availability for living mammals
+\noindent{\Large \bf Assessment of available anatomical characters for linking living mammals to fossil taxa in phylogenetic analyses}
+%TG: or actually, along the same lines:
+%"Assessment of available anatomical characters for linking living and fossil mammals in phylogenetic analyses"
+%TG: What do you think?
 
@@
 To perform such analyses it is necessary to combine living and fossil taxa in phylogenetic trees.
-One increasingly popular method, the Total Evidence method \cite{ronquista2012}, ccombines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix that can then be used with the tip-dating method (e.g.
+One increasingly popular method, the Total Evidence method \cite{ronquista2012}, combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix that can then be used with the tip-dating method (e.g.
 \cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a chronogram with living and fossil taxa at the tips.
-A downside of this method is that it requires molecular data for living taxa and morphological/anatomical data for both living and fossil taxa.
+A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data for both living and fossil taxa.
 Sections of this data can be difficult, or impossible, to collect for every taxon in the analysis.
@@
 Simulations show that the ability of the Total Evidence method to recover the correct topology decreases when there is little overlap between coded anatomical characters in living and fossil taxa, and that the effect of missing data on topology is greatest when living taxa have few anatomical characters available \cite{GuillermeCooper}.
-This is because (1) fossils will not be placed accurately within the correct clade if it contains no coded anatomical characters for living taxa; and (2) fossils have a higher probability of being placed within clades with more coded anatomical characters available for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.
+We might expect the total evidence method to perform poorly when there is little overlap between coded anatomical characters in living and fossil taxa.
+This is because fossil taxa cannot be correctly placed within a clade of living species if none of its members have been coded for morphological characters.
+Furthermore, simulations show that fossils are more likely to be placed in clades with for which more characters have been coded, regardless of whether this is the correct clade \cite{GuillermeCooper}.
+% This is because (1) fossils will not be placed accurately within the correct clade if it contains no coded anatomical characters for living taxa; and (2) fossils have a higher probability of being placed within clades with more coded anatomical characters available for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.
 
@@
 
-%\input{Table_results}
+\input{Table_results}
 
@@
 
-%\begin{figure}[!htbp]
-%\centering
-%    \includegraphics[width=1\textwidth]{figure1.pdf}
-%\caption{Phylogenetic distribution of species with available coded anatomical characters across two orders (A: Primates; B: Carnivora).
-%Blue branches indicate species with available coded anatomical characters.}
-%\label{Figure_example_coverage}
-%\end{figure}
+\begin{figure}[!htbp]
+\centering
+   \includegraphics[width=1\textwidth]{figure1.pdf}
+\caption{Phylogenetic distribution of species with available coded anatomical characters across two orders (A: Primates; B: Carnivora).
+Blue branches indicate species with available coded anatomical characters.}
+\label{Figure_example_coverage}
+\end{figure}
 
@@
 Our results suggest that this may be problematic at the genus-level in Carnivora, Chiroptera and Soricomorpha. 
-For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade because they have no coded anatomical characters available.
+%For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade because they have no coded anatomical characters available.
+For example, a carnivoran fossil is unlikely to be placed in Herpestidae because they have no coded anatomical characters available.
 Instead the fossil will have a high probability of being placed on a branch that contains many anatomical characters such as within the Canidae or Ursidae (Figure \ref{Figure_example_coverage}B). 
@@
 
-It is worth noting, however, that our analysis did not include all the matrices containing anatomical characters ever published.
-In fact, our data collection procedure focused on including studies that provided matrices easily accessible, i.e.
+We acknowledge, however, that our analysis did not include all the matrices containing anatomical characters ever published.
+In fact, our data collection procedure focused on including studies that provided easily accessible matrices, i.e.
 we did specifically not include any matrices that were only available in paper format (e.g.
@@
 Fortunately, mammal specimens are usually readily available in natural history collections, therefore, we propose increased effort into coding anatomical characters from living species, possibly by engaging in collaborative data collection projects.
-Such efforts would be valuable not only to phylogeneticists, but also to any researcher focusing understanding macroevolutionary patterns and processes.
+Such efforts would be valuable not only to phylogeneticists, but also to any researcher focusing on understanding macroevolutionary patterns and processes.
 

### 2016-03-29 Manuscript/BiologyLetters/response2.tex
Response letter INIT

\documentclass[12pt,letterpaper]{article}

%Packages
\usepackage{pdflscape}
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{float}
\usepackage{latexsym}
\usepackage{url}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{array}
\usepackage[version=3]{mhchem}
\usepackage{ifthen}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{amsthm}
\usepackage{amstext}
\usepackage{enumerate}
\usepackage[osf]{mathpazo}
\usepackage{dcolumn}
\usepackage{lineno}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\pagenumbering{arabic}

%Pagination style and stuff
%\linespread{2}

\raggedright
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0} 
\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}
\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}
\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}
\renewcommand{\tableofcontents}{}

\setlength\parindent{0pt}

\begin{document}

\textbf{RE: Decision on Manuscript ID RSBL-2015-1003}\\
\bigskip
Dear Surayya Johar,\\
\bigskip
We are very happy on the manuscript acceptance decision.
We have corrected the minor issues raised by referees 2 and 4 as suggested.
We also took on board the few minor comments from referees 1 and 3 and respond to their points below.

To improve the clarity of this document, we dealt with each comment in the order they appeared in your ``Decision on Manuscript'' email and we have included the reviewers comments in blue.

\section{Referee 2}

\begin{enumerate}
\item{\textcolor{blue}{line 34: spelling: ccombines = combines}}

\item{\textcolor{blue}{lines 44 - 52. I'd suggest rewording this a little bit as lines 48 - 52 are a little awkward. May I suggest the following:
"We might expect the total evidence method to perform poorly when there is little overlap between coded anatomical characters in living and fossil taxa. This is because fossil taxa cannot be correctly placed within a clade of living species if none of its members have been coded for morphological characters. Furthermore, simulations show that fossils are more likely to be placed in clades with for which more characters have been coded, regardless of whether this is the correct clade [8]."}}

\item{\textcolor{blue}{line 152: ... a carnivoran fossil is unlikely to be placed in Herpestidae because ....}}

\item{\textcolor{blue}{line 159: I'd replace "It is worth noting" with "We acknowledge"}}

\item{\textcolor{blue}{line 161: word order: provided easily accessible matrices}}
\end{enumerate}

\section{Referee 1}

\begin{enumerate}
\item{\textcolor{blue}{The authors haven’t addressed Pete Wagner’s important comments about the use of the term “cladistics.” Morphological character data are not inherently cladistics, although they can be used for cladistics analysis, as well as Bayesian and likelihood approaches. They also haven’t done anything to clarify that the relevant character set for aligning fossil and modern data is skeletal characters. Many anatomical characters do nothing at all for the incorporation of fossil taxa.}}

\item{\textcolor{blue}{I’m also concerned that my original comments about the biases in character sets included still aren’t adequately addressed; they now acknowledge that they didn’t include any character sets that weren’t available in print, but that doesn’t address what kind of difference it would make to their results. I’m concerned it could be a really important bias; morphological data was much more commonly published for extant taxa in older phylogenetic analyses, and those are less likely to be available digitally. You might be substantially underestimating the availability of these data, and if TE phylogenies aren’t including them, I’m rather disappointed in the folks constructing them. }}

\item{\textcolor{blue}{I don’t agree with the assertion that the genus level should be our primary concern; more and more studies are building species-level trees with fossil taxa. I’m happy for the authors to discuss the patterns at the genus level, because it is one of the important levels, but given that genera are an artificial line in the taxonomic sand, it’s certainly not the primary one, and I’d hate to see this statement sail unopposed into the literature. I’d like to see it moderated a bit in the final version of the manuscript.}}
\end{enumerate}

\section{Referee 3}

\begin{enumerate}
\item{\textcolor{blue}{The new title is an improvement.  However, I think that it could be altered still further: would something like: "Assessment of available anatomical characters for linking living mammals to fossil taxa in phylogenetic analyses"?  That is the main thrust of the paper, after all, and it also avoids having someone simply think: "Who cares? I have molecules" and moving on to the next article.}}

\item{\textcolor{blue}{Lines 159 - 160: "It is worth noting, however, that our analysis did not include all the matrices containing anatomical characters ever published."}}

\item{\textcolor{blue}{One thing to note is that a lot of the older analyses, which tended to be morphological, were done in the early days of modern phylogenetics, and those are the matrices that are most apt to not have pdf or html versions.  So, there might be a bias against the "pioneering" studies.  However, many of those studies will be a quarter of a century old, and the homology concepts and (especially!) coding protocols being used back then almost certainly differ from now.  Also, a lot of those matrices were "cleaned" to remove homoplasy: this was not attempted fraud, but due to the fact that computer runs that are done in hours today took weeks and weeks back then, and actually storing all of the trees that might come out of "messy" matrices was beyond the old floppy drives.  That is one reason why I rarely use matrices published before the early 1990's in my own meta-analyses.  So, if those matrices do exist and can be computerized, then I think that they should be viewed as "starting points" for coding taxa: they almost certainly will need updating.}}
\end{enumerate}

\section{Referee 4}

\begin{enumerate}
\item{\textcolor{blue}{on L174 ("but also to any researcher focusing understanding macroevolutionary patterns and processes"), this seems to be missing an "on".}}
\end{enumerate}

We hope we have responded to all these comments appropriately. Please let us know if you require any further information,\\
\bigskip

Thomas Guillerme (t.guillerme@imperial.ac.uk),\\
Natalie Cooper

\end{document}

### 2016-03-29 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Dealt with referees 1 and 3 comments

@@
 \noindent{\Large \bf Assessment of available anatomical characters for linking living mammals to fossil taxa in phylogenetic analyses}
-%TG: or actually, along the same lines:
-%"Assessment of available anatomical characters for linking living and fossil mammals in phylogenetic analyses"
-%TG: What do you think?
 
@@
 \cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a chronogram with living and fossil taxa at the tips.
-A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data for both living and fossil taxa.
+A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared between both living and fossil taxa (i.e.
+hard tissue characters such as skeletal ones).
+%
+% TG: Emphasize on discrete morphological/anatomical data (often skeletal). I've added the "shared between"
+%
 Sections of this data can be difficult, or impossible, to collect for every taxon in the analysis.
 For example, fossils rarely have molecular data and incomplete fossil preservation may reduce the number of anatomical characters available.
+%TG: should we add another "(i.e.
+skeletal characters)" here?
 Additionally, it has become less common to collect anatomical characters for living taxa when molecular data is available (e.g.
@@
 Unfortunately this missing data can lead to errors in phylogenetic inference.
-Simulations show that the ability of the Total Evidence method to recover the correct topology decreases when there is little overlap between coded anatomical characters in living and fossil taxa, and that the effect of missing data on topology is greatest when living taxa have few anatomical characters available \cite{GuillermeCooper}.
+% Simulations show that the ability of the Total Evidence method to recover the correct topology decreases when there is little overlap between coded anatomical characters in living and fossil taxa, and that the effect of missing data on topology is greatest when living taxa have few anatomical characters available \cite{GuillermeCooper}.
 We might expect the total evidence method to perform poorly when there is little overlap between coded anatomical characters in living and fossil taxa.
 This is because fossil taxa cannot be correctly placed within a clade of living species if none of its members have been coded for morphological characters.
-Furthermore, simulations show that fossils are more likely to be placed in clades with for which more characters have been coded, regardless of whether this is the correct clade \cite{GuillermeCooper}.
+Furthermore, simulations show that fossils are more likely to be placed in clades for which more characters have been coded, regardless of whether this is the correct clade \cite{GuillermeCooper}.
+%TG: That's Graham's sentence.
+It's not exactly what we *show* in the TEM paper but we definitely discuss this possibility in the discussion.
 % This is because (1) fossils will not be placed accurately within the correct clade if it contains no coded anatomical characters for living taxa; and (2) fossils have a higher probability of being placed within clades with more coded anatomical characters available for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.
 
-The issues above highlight that it is crucial to have sufficient coded anatomical characters available for living taxa in a clade before using Total Evidence approaches.
+The issues above highlight that it is crucial to have sufficient coded hard tissue anatomical characters available for living taxa in a clade before using the Total Evidence approach.
 However, it is unclear how many coded anatomical characters are actually available for living taxa, i.e.
@@
 We also determine whether the characters are phylogenetically overdispersed or clustered across mammalian orders.
+%TG: a point to be made here maybe is that we do NOT look at hard tissue characters, rather just the presence of any characters (which is already poor enough)?
+%TG: on the other hand, we might not need to emphasize that to much since, anyway, it is clearly explained in the methods (i.e. we tool all cladistic data from matrices)
 
@@
 Therefore we expect difficulties in placing fossils at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels.
-This is important in practice because of the slight discrepancy between neontological and palaeontological species concepts.
-While neontological species are described using morphology, genes, distribution etc.; palaeontological species can be based only on morphological/anatomical, spatial and temporal data (e.g.
+% This is important in practice because of the slight discrepancy between neontological and palaeontological species concepts.
+% While neontological species are described using morphology, genes, distribution, etc.; palaeontological species can be based only on morphological/anatomical, spatial and temporal data (e.g.
 \cite{ni2013oldest}).
-Therefore, many palaeontological studies use the genus (or monospecific genera) as their smallest OTU (e.g.
+% Therefore, many palaeontological studies use the genus (or monospecific genera) as their smallest OTU (e.g.
 \cite{ni2013oldest,O'Leary08022013}), so data availability at the genus-level in living mammals should be our primary concern when building phylogenies of living and fossil taxa.
+%
+%
+%TG: Might first feeling was to just remove the last sentence but, when reading the previous one (which I like), I'm not sure if this sentence can be removed without removing the two previous ones on the species concept differences.
+What do you think?
+Anyway, it would then maybe just be better to dump the three sentences an finish this paragraph at "Therefore we expect difficulties in placing fossils at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels."
+%TG: On the other hand removing that gives us space for discussing the data sampling problem below
+%
+%
 
@@
 by emailing the authors).
-The difficulty of obtaining such matrices makes them less likely to be included in Total Evidence analyses, especially where the number of taxa to include in the phylogeny is high.
+Matrices containing anatomical characters where much more common before the advent of molecular phylogenetics and therefore are also more likely to be unavailable in a reusable format.
+This might include some bias in Total Evidence analyses.
+Nonetheless, these matrices are also likely to differ from more recent ones in terms of their underlying definition of homology and their coding practices (see \cite{Brazeau2011}).
+Additionally, many recent morphological matrices reuse living taxa from previous matrices (see ESM1).
 

### 2016-03-29 Manuscript/BiologyLetters/response2.tex
Dealt with referees 1 and 3 comments

@@
 \item{\textcolor{blue}{line 34: spelling: ccombines = combines}}
+We fixed this typo.
 
@@
 This is because fossil taxa cannot be correctly placed within a clade of living species if none of its members have been coded for morphological characters.
-Furthermore, simulations show that fossils are more likely to be placed in clades with for which more characters have been coded, regardless of whether this is the correct clade [8]."}}
+Furthermore, simulations show that fossils are more likely to be placed in clades for which more characters have been coded, regardless of whether this is the correct clade [8]."}}
+We replaced the section with Referee 2's suggestion.
 
 \item{\textcolor{blue}{line 152: ... a carnivoran fossil is unlikely to be placed in Herpestidae because ....}}
+We fixed this typo.
 
 \item{\textcolor{blue}{line 159: I'd replace "It is worth noting" with "We acknowledge"}}
+We changed the start of the sentence as suggested.
 
 \item{\textcolor{blue}{line 161: word order: provided easily accessible matrices}}
+We changed the word order.
+
 \end{enumerate}
@@
 Many anatomical characters do nothing at all for the incorporation of fossil taxa.}}
+We emphasised the importance of skeletal characters in the introduction as follows:
+``A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared between both living and fossil taxa (i.e.
+hard tissue characters such as skeletal ones).''
+lines @@@.
 
@@
 
+We agree with this comment (and Referee 3 comment 2) and modified the paragraph discussing this caveat as follow:
+``Matrices containing anatomical characters where much more common before the advent of molecular phylogenetics and therefore are also more likely to be unavailable in a reusable format.
+This might include some bias in Total Evidence analyses.
+Nonetheless, these matrices are also likely to differ from more recent ones in terms of their underlying definition of homology and their coding practices (see [19]).
+Additionally, many recent morphological matrices reuse living taxa from previous matrices (see ESM1).''
+lines @@@.
+%TG: Here's how I went with the section, basically, I acknowledge referee 1's points but add Peters ones just after:
+%TG: - Yes, we specifically ignored old matrices because definition of accesible data (see above)
+%TG: - Yes, we do agree it could include some bias
+%TG: - But these matrices are probably not so good anyway (Peter's comment)
+%TG: - And anyway it doesn't solve the problem because the living taxa used in these "old" matrices are reused in new ones anyway (or at least that's what I observed in primates)
+%TG: Personally I doubt the problem is due to data availability but way more to the problem we underlined in the abstract: palaeontologist don't wanna waste their time with living species (and I don't specially blame them).
+
 \item{\textcolor{blue}{I don’t agree with the assertion that the genus level should be our primary concern; more and more studies are building species-level trees with fossil taxa.
@@
 I’d like to see it moderated a bit in the final version of the manuscript.}}
+Following Referee's 1 and 4's comments, we retracted this statement.
+
 \end{enumerate}
@@
 I have molecules" and moving on to the next article.}}
+We changed the title as suggested by Referee 3.
 
-\item{\textcolor{blue}{Lines 159 - 160: "It is worth noting, however, that our analysis did not include all the matrices containing anatomical characters ever published."}}
-
-\item{\textcolor{blue}{One thing to note is that a lot of the older analyses, which tended to be morphological, were done in the early days of modern phylogenetics, and those are the matrices that are most apt to not have pdf or html versions.
+\item{\textcolor{blue}{Lines 159 - 160: "It is worth noting, however, that our analysis did not include all the matrices containing anatomical characters ever published."
+One thing to note is that a lot of the older analyses, which tended to be morphological, were done in the early days of modern phylogenetics, and those are the matrices that are most apt to not have pdf or html versions.
 So, there might be a bias against the "pioneering" studies.
@@
 So, if those matrices do exist and can be computerized, then I think that they should be viewed as "starting points" for coding taxa: they almost certainly will need updating.}}
+
+% TO DO (see Referee 1 comment 2)
+
 \end{enumerate}
@@
 \begin{enumerate}
+
+\item{\textcolor{blue}{In particular I agree with Referee 1 regarding the continued use of "cladistic" and the genus-level issue (many palaeontologists are now focused on the species-level and certainly many mammalian genera are not monotypic - this is more of an issue with dinosaur workers), and Referee 3 (Pete Wagner's) title suggestion is - I think - more likely to attract readers.}}
+We dealt with these suggestions (see respectively Referee 1 comment 1, Referee 1 comment 3 and Referee 3 comment 1).
+
 \item{\textcolor{blue}{on L174 ("but also to any researcher focusing understanding macroevolutionary patterns and processes"), this seems to be missing an "on".}}
+We fixed this typo.
 \end{enumerate}

### 2016-03-30 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Edits in response to review round 2

@@
 \medskip
-\noindent{*\bf Corresponding author.} \textit{t.guillerme@imperial.ac.uk}\\  
+\noindent{*\bf Corresponding author. Current address: } \textit{t.guillerme@imperial.ac.uk}\\  % NC: You need to add your current address to the corresponding author bit I think.
 \vspace{1in}
@@
 However, although molecular data for living species are widely available, scientists generating morphological data mainly focus on fossils.
-Therefore, there are few coded anatomical characters in living taxa, even in well-studied groups like mammals.
+Therefore, there are fewer coded anatomical characters in living taxa, even in well-studied groups like mammals.
 
@@
 11 of 28 mammalian orders have \textless 25\% species with available characters; this has implications for the accurate placement of fossils, although the issue is less pronounced at higher taxonomic levels.
-In most orders, species with available characters are randomly distributed across the phylogeny, which may reduce the impact of the problem.
+In most orders, species with available characters are randomly distributed across the phylogeny, that may reduce the impact of the problem.
 We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence phylogenies. 
@@
 
-\noindent Key words: Total Evidence method, phylogenetic clustering, cladistic matrix, extinct, topology\\
+\noindent Key words: Total Evidence method, phylogenetic clustering, cladistic matrix%NC: change cladistic
+, extinct, topology\\
 
@@
 \cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a chronogram with living and fossil taxa at the tips.
-A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared between both living and fossil taxa (i.e.
-hard tissue characters such as skeletal ones).
+A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared across living and fossil taxa (i.e.
+hard tissue characters such as skeletal features).
 %
@@
 skeletal characters)" here?
+% NC: No I think we are making a slightly different point.
 Additionally, it has become less common to collect anatomical characters for living taxa when molecular data is available (e.g.
@@
 % Simulations show that the ability of the Total Evidence method to recover the correct topology decreases when there is little overlap between coded anatomical characters in living and fossil taxa, and that the effect of missing data on topology is greatest when living taxa have few anatomical characters available \cite{GuillermeCooper}.
-We might expect the total evidence method to perform poorly when there is little overlap between coded anatomical characters in living and fossil taxa.
-This is because fossil taxa cannot be correctly placed within a clade of living species if none of its members have been coded for morphological characters.
+We might expect the Total Evidence method to perform poorly when there is little overlap between coded anatomical characters in living and fossil taxa, because fossil taxa cannot be correctly placed within a clade of living species with no coded characters.
 Furthermore, simulations show that fossils are more likely to be placed in clades for which more characters have been coded, regardless of whether this is the correct clade \cite{GuillermeCooper}.
@@
 % This is because (1) fossils will not be placed accurately within the correct clade if it contains no coded anatomical characters for living taxa; and (2) fossils have a higher probability of being placed within clades with more coded anatomical characters available for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.
+% NC: OK but need to be careful about just randomly citing ourselves when we know that's not what the paper says.
 
-The issues above highlight that it is crucial to have sufficient coded hard tissue anatomical characters available for living taxa in a clade before using the Total Evidence approach.
+The issues above highlight that it is crucial to have sufficient coded anatomical characters available for living taxa in a clade before using the Total Evidence approach.
 However, it is unclear how many coded anatomical characters are actually available for living taxa, i.e.
 already coded from museum specimens and deposited in phylogenetic matrices accessible online, and how this data is distributed across clades.
-Intuitively, most people assume this kind of data has already been collected, but empirical data suggest otherwise (e.g. in \cite{ronquista2012,slaterphylogenetic2013,beckancient2014}).
+Intuitively, most people assume this data has already been collected, but empirical analyses suggest otherwise (e.g. in \cite{ronquista2012,slaterphylogenetic2013,beckancient2014}).
 To investigate this further, we assess the number of available coded anatomical characters for living mammals to determine whether enough data exists to build reliable Total Evidence phylogenies.
@@
 %TG: on the other hand, we might not need to emphasize that to much since, anyway, it is clearly explained in the methods (i.e. we tool all cladistic data from matrices)
-
+% NC: If someone makes a point about it in a review it means it isn't explained clearly enough. I think adding a note to the methods and/or discussion is needed.
 %---------------------------------------------
@@
 %---------------------------------------------
-
+% NC: From here on in we need to replace the word cladistic. I don't like the "anatomical characters" as it's a bit long. How about character matrices? You decide.
 \section{Materials and Methods}
 \subsection{Data collection and standardisation}
-We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{morphobank.org} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}).
+We downloaded all cladistic%
+ matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{morphobank.org} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}).
 We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material (ESM) for details).
@@
 \subsection{Data availability and distribution}
-To assess the availability of coded anatomical characters for each mammalian order and across mammals, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
+To assess the availability of coded anatomical characters for each mammalian order and across mammals, we calculated the percentage of OTUs with coded anatomical characters at three different taxonomic levels: family, genus and species.
+We do not distinguish between soft and hard characters, but the majority of matrices contain at least some hard tissue characters.
+% NC: True?
+This is to deal with the skeletal bit.
+Put into response letter.
 We consider orders with \textless 25\% of living taxa with available anatomical characters as having low data coverage, and orders with \textgreater 75\% of living taxa with available anatomical characters as having high data coverage.
@@
 For each order and for all mammals, we investigated whether the available coded anatomical characters were (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index (NTI; \cite{webb2002phylogenies}) and the Net Relatedness Index (NRI; \cite{webb2002phylogenies}).
-NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to it across the whole phylogeny \cite{Cooper2008}. 
+NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to them across the whole phylogeny \cite{Cooper2008}. 
 Both metrics were calculated using the \texttt{picante} package in R \cite{picante,R}.
@@
 
+% NC: Can you check the caption we used doesn't have cladistic in it? I have a feeling it does...
+
 \input{Table_results}
@@
 Therefore we expect difficulties in placing fossils at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels.
+Additionally, our analyses may underestimate the problem as we do not distiguish between soft and hard tissue characters; if a living taxon has only soft tissue coded anatomical characters then it will not have overlapping data with fossils that only have hard tissues preserved.
+%NC: This deals with reviewer 1.
+What do you think?
+
 % This is important in practice because of the slight discrepancy between neontological and palaeontological species concepts.
@@
 %
-%
+% NC: Yeah I think this is the correct thing to do here.
 
@@
 The worst case scenario for fossil placement is that species with available characters are phylogenetically clustered. 
-Then we expect two major biases: first, fossils will not be placed within a clade containing no data, and second, fossils will have higher probability of being placed within the most sampled clade by chance.
+Then we expect two major biases: first, fossils will not be placed within a clade containing no hard tissue data, and second, fossils will have higher probability of being placed within the most sampled clade by chance.
+% NC: I added hard tissue here
 Our results suggest that this may be problematic at the genus-level in Carnivora, Chiroptera and Soricomorpha. 
@@
 
-We acknowledge, however, that our analysis did not include all the matrices containing anatomical characters ever published.
-In fact, our data collection procedure focused on including studies that provided easily accessible matrices, i.e.
-we did specifically not include any matrices that were only available in paper format (e.g.
-printed in books), non-reusable format (e.g.
-an image of the matrix)  or/and matrices available only upon request (e.g.
-by emailing the authors).
-Matrices containing anatomical characters where much more common before the advent of molecular phylogenetics and therefore are also more likely to be unavailable in a reusable format.
-This might include some bias in Total Evidence analyses.
-Nonetheless, these matrices are also likely to differ from more recent ones in terms of their underlying definition of homology and their coding practices (see \cite{Brazeau2011}).
-Additionally, many recent morphological matrices reuse living taxa from previous matrices (see ESM1).
-
+We acknowledge, however, that our analysis does not include all matrices containing anatomical characters ever published.
+Instead, our data collection procedure focused on including studies that provided easily accessible matrices, i.e.
+we did not include matrices that are only available in books, non-reusable formats (e.g.
+an image of the matrix), or matrices available only upon request from the authors.
+Matrices containing anatomical characters were more common before the advent of molecular phylogenetics, but these matrices are also more likely to be unavailable in a reusable format, thus will be missing from our analyses.
+Although this will bias our results towards lower coverage we do not think this bias will be large, as many recent morphological matrices reuse living taxa characters from older matrices (see ESM1), so many of these data will be present in our analyses.
+Additionally, these older matrices are likely to differ from more recent ones in terms of their underlying definition of homology and their coding practices (see \cite{Brazeau2011}).
+Therefore, care needs to be taken when deciding how to include these older matrices.
+% NC: I editted this as the reviewer is concerned that our analyses might be biased, not total evidence in general
 Despite the absence of good morphological/anatomical data coverage for living mammals, the Total Evidence method still seems to be the most promising way of combining living and fossil species in macroevolutionary analyses.
@@
 
+% NC: The ref for our paper looks weird - it has a url, probably fix that and check the reference is correct so we get the citation.
+% Ref 10 - there's an 03 after the date. Not sure what that refers to.
+% NC: Ref 16 - is KPg mising a - ?
+% NC: Ref 17 has a missing space
+
 \bibliographystyle{vancouver}

### 2016-03-30 Manuscript/BiologyLetters/response2.tex

@@
 \bigskip
-We are very happy on the manuscript acceptance decision.
+%We are very happy on the manuscript acceptance decision. % NC: Not usual to write this, of course we are!
+Thanks again to the reviewers for their thoughtful comments on our manuscript.
 We have corrected the minor issues raised by referees 2 and 4 as suggested.
-We also took on board the few minor comments from referees 1 and 3 and respond to their points below.
+We also took on board the comments from referees 1 and 3 and respond to their points below.
 
@@
 
+% NC: Bit weird to have referee two and then 1, I've moved it back so it's in the right order! always easier for an editor to read like that than to have to go back and forth.
+
+\section{Referee 1}
+
+\begin{enumerate}
+\item{\textcolor{blue}{The authors haven't addressed Pete Wagner's important comments about the use of the term 'cladistics'.
+Morphological character data are not inherently cladistics, although they can be used for cladistics analysis, as well as Bayesian and likelihood approaches.
+They also haven't done anything to clarify that the relevant character set for aligning fossil and modern data is skeletal characters.
+Many anatomical characters do nothing at all for the incorporation of fossil taxa.}}
+
+We agreed with the reviewer in the previous round of reviews and thought we had removed all mentions of cladistics.
+It was still used twice in the manuscript (and once in the keywords) but we have now corrected this error.
+% NC: Fix these. And check the supp material.
+
+We emphasised the importance of skeletal characters in the introduction as follows:
+``A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared between both living and fossil taxa (i.e.
+hard tissue characters such as skeletal ones).''
+lines @@@.
+% NC: I've edited this slightly.
+Note that the issue is that even though we might find loads of data for a living mammal, if it is all soft tissue it is USELESS!
+It cannot overlap with fossil taxa.
+In an ideal world we would be just looking for hard characters, but we've done this now so instead we will have to mention this in the discussion somewhere
+
+We also added a comment in the discussion to the effect that having data for a living mammal doesn't necessarily mean it can overlap with a fossil - it could all be soft tissue characters.
+However, in general our datasets appear to contain at least one hard character for each living species so it shouldn't influence our results.
+But yes it is an important consideration for future work and we thank the reviewer for pointing this out (we misunderstood this point in the first round of revisions).
+% NC: We did misunderstand this first go round.
+This is why I don't like you to rush.
+You need to learn how to very carefully read a reviewers letter and understand the criticisms.
+Remember one paragraph of comments might have several things to fix.
+
+% NC: Need to add a line or two for this in the discussion. Or maybe rephrase the above and we can just stick with the clarifications I've added to the methods and intro.
+
+% NC: I've changed a few bits in the text too
+
+\item{\textcolor{blue}{I'm also concerned that my original comments about the biases in character sets included still aren't adequately addressed; they now acknowledge that they didn't include any character sets that weren't available in print, but that doesn't address what kind of difference it would make to their results.
+I'm concerned it could be a really important bias; morphological data was much more commonly published for extant taxa in older phylogenetic analyses, and those are less likely to be available digitally.
+You might be substantially underestimating the availability of these data, and if TE phylogenies aren't including them, I'm rather disappointed in the folks constructing them.
+}}
+
+We agree with this comment (and Referee 3 comment 2) and modified the paragraph discussing this caveat as follows:
+``Matrices containing anatomical characters were much more common before the advent of molecular phylogenetics and therefore are also more likely to be unavailable in a reusable format.
+This might include some bias in Total Evidence analyses.
+Nonetheless, these matrices are also likely to differ from more recent ones in terms of their underlying definition of homology and their coding practices (see [19]).
+Additionally, many recent morphological matrices reuse living taxa from previous matrices (see ESM1).''
+lines @@@.
+
+%TG: Here's how I went with the section, basically, I acknowledge referee 1's points but add Peters ones just after:
+%TG: - Yes, we specifically ignored old matrices because definition of accesible data (see above)
+%TG: - Yes, we do agree it could include some bias
+%TG: - But these matrices are probably not so good anyway (Peter's comment)
+%TG: - And anyway it doesn't solve the problem because the living taxa used in these "old" matrices are reused in new ones anyway (or at least that's what I observed in primates)
+%TG: Personally I doubt the problem is due to data availability but way more to the problem we underlined in the abstract: palaeontologist don't wanna waste their time with living species (and I don't specially blame them).
+
+% NC: I've editted this a bit to get more closely at what the reviewer wants (it thinks OUR analysis might be biased, and is correct).
+% Make sure you also write a brief explanation here as well as editting the text. Perhaps add a line like:
+
+We are unable to properly quantify the bias introduced by ignoring older matrices without going out and collecting them all exhaustively and re-running our analyses.
+This would take a substantial amount of time not available at present.
+
+\item{\textcolor{blue}{I don't agree with the assertion that the genus level should be our primary concern; more and more studies are building species-level trees with fossil taxa.
+I'm happy for the authors to discuss the patterns at the genus level, because it is one of the important levels, but given that genera are an artificial line in the taxonomic sand, it's certainly not the primary one, and I'd hate to see this statement sail unopposed into the literature.
+I'd like to see it moderated a bit in the final version of the manuscript.}}
+Following Referee's 1 and 4's comments, we removed this paragraph.
+
+\end{enumerate}
 \section{Referee 2}
@@
 
-\section{Referee 1}
-
-\begin{enumerate}
-\item{\textcolor{blue}{The authors haven’t addressed Pete Wagner’s important comments about the use of the term “cladistics.” Morphological character data are not inherently cladistics, although they can be used for cladistics analysis, as well as Bayesian and likelihood approaches.
-They also haven’t done anything to clarify that the relevant character set for aligning fossil and modern data is skeletal characters.
-Many anatomical characters do nothing at all for the incorporation of fossil taxa.}}
-We emphasised the importance of skeletal characters in the introduction as follows:
-``A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared between both living and fossil taxa (i.e.
-hard tissue characters such as skeletal ones).''
-lines @@@.
-
-\item{\textcolor{blue}{I’m also concerned that my original comments about the biases in character sets included still aren’t adequately addressed; they now acknowledge that they didn’t include any character sets that weren’t available in print, but that doesn’t address what kind of difference it would make to their results.
-I’m concerned it could be a really important bias; morphological data was much more commonly published for extant taxa in older phylogenetic analyses, and those are less likely to be available digitally.
-You might be substantially underestimating the availability of these data, and if TE phylogenies aren’t including them, I’m rather disappointed in the folks constructing them.
-}}
-
-We agree with this comment (and Referee 3 comment 2) and modified the paragraph discussing this caveat as follow:
-``Matrices containing anatomical characters where much more common before the advent of molecular phylogenetics and therefore are also more likely to be unavailable in a reusable format.
-This might include some bias in Total Evidence analyses.
-Nonetheless, these matrices are also likely to differ from more recent ones in terms of their underlying definition of homology and their coding practices (see [19]).
-Additionally, many recent morphological matrices reuse living taxa from previous matrices (see ESM1).''
-lines @@@.
-%TG: Here's how I went with the section, basically, I acknowledge referee 1's points but add Peters ones just after:
-%TG: - Yes, we specifically ignored old matrices because definition of accesible data (see above)
-%TG: - Yes, we do agree it could include some bias
-%TG: - But these matrices are probably not so good anyway (Peter's comment)
-%TG: - And anyway it doesn't solve the problem because the living taxa used in these "old" matrices are reused in new ones anyway (or at least that's what I observed in primates)
-%TG: Personally I doubt the problem is due to data availability but way more to the problem we underlined in the abstract: palaeontologist don't wanna waste their time with living species (and I don't specially blame them).
-
-\item{\textcolor{blue}{I don’t agree with the assertion that the genus level should be our primary concern; more and more studies are building species-level trees with fossil taxa.
-I’m happy for the authors to discuss the patterns at the genus level, because it is one of the important levels, but given that genera are an artificial line in the taxonomic sand, it’s certainly not the primary one, and I’d hate to see this statement sail unopposed into the literature.
-I’d like to see it moderated a bit in the final version of the manuscript.}}
-Following Referee's 1 and 4's comments, we retracted this statement.
-
-\end{enumerate}
-
 \section{Referee 3}
@@
 % TO DO (see Referee 1 comment 2)
+
+%NC: Also need to add a line sayong we've now mentioned this in the discussion and add the bit here.
 

### 2016-03-30 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Dealt with Natalie's comments (still need to sort couple of comments)

@@
 \medskip
-\noindent{*\bf Corresponding author. Current address: } \textit{t.guillerme@imperial.ac.uk}\\  % NC: You need to add your current address to the corresponding author bit I think.
+\noindent{*\bf Corresponding author.
+Current address: } \textit{t.guillerme@imperial.ac.uk}\\  % NC: You need to add your current address to the corresponding author bit I think.
+TG: i.e.
+Silwood?
+That'll eat a shit load of words for not much no?
 \vspace{1in}
@@
 
-\noindent Key words: Total Evidence method, phylogenetic clustering, cladistic matrix%NC: change cladistic
-, extinct, topology\\
+\noindent Key words: Total Evidence method, phylogenetic clustering, discrete morphological matrix, extinct, topology\\
 
@@
 \cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a chronogram with living and fossil taxa at the tips.
-A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared across living and fossil taxa (i.e.
+A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared between both living and fossil taxa (i.e.
 hard tissue characters such as skeletal features).
-%
-% TG: Emphasize on discrete morphological/anatomical data (often skeletal). I've added the "shared between"
-%
 Sections of this data can be difficult, or impossible, to collect for every taxon in the analysis.
 For example, fossils rarely have molecular data and incomplete fossil preservation may reduce the number of anatomical characters available.
-%TG: should we add another "(i.e.
-skeletal characters)" here?
-% NC: No I think we are making a slightly different point.
 Additionally, it has become less common to collect anatomical characters for living taxa when molecular data is available (e.g.
@@
 Unfortunately this missing data can lead to errors in phylogenetic inference.
-% Simulations show that the ability of the Total Evidence method to recover the correct topology decreases when there is little overlap between coded anatomical characters in living and fossil taxa, and that the effect of missing data on topology is greatest when living taxa have few anatomical characters available \cite{GuillermeCooper}.
 We might expect the Total Evidence method to perform poorly when there is little overlap between coded anatomical characters in living and fossil taxa, because fossil taxa cannot be correctly placed within a clade of living species with no coded characters.
 Furthermore, simulations show that fossils are more likely to be placed in clades for which more characters have been coded, regardless of whether this is the correct clade \cite{GuillermeCooper}.
-%TG: That's Graham's sentence.
-It's not exactly what we *show* in the TEM paper but we definitely discuss this possibility in the discussion.
-% This is because (1) fossils will not be placed accurately within the correct clade if it contains no coded anatomical characters for living taxa; and (2) fossils have a higher probability of being placed within clades with more coded anatomical characters available for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.
-% NC: OK but need to be careful about just randomly citing ourselves when we know that's not what the paper says.
 
@@
 % NC: From here on in we need to replace the word cladistic. I don't like the "anatomical characters" as it's a bit long. How about character matrices? You decide.
+% TO DO
 \section{Materials and Methods}
 \subsection{Data collection and standardisation}
-We downloaded all cladistic%
- matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{morphobank.org} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}).
+We downloaded all discrete morphological matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{morphobank.org} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}).
 We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material (ESM) for details).
@@
 Put into response letter.
+TG: probably true.
+Didn't check all matrices but because of their data's nature (i.e.
+collected by anatomists back in the days and paleontologists nowadays) I'll bet you there's "at least some".
 We consider orders with \textless 25\% of living taxa with available anatomical characters as having low data coverage, and orders with \textgreater 75\% of living taxa with available anatomical characters as having high data coverage.
@@
 % NC: Can you check the caption we used doesn't have cladistic in it? I have a feeling it does...
+% TO DO
 
@@
 What do you think?
+TG: Yap, excellent.
 
@@
 %
-%TG: Might first feeling was to just remove the last sentence but, when reading the previous one (which I like), I'm not sure if this sentence can be removed without removing the two previous ones on the species concept differences.
+%TG: My first feeling was to just remove the last sentence but, when reading the previous one (which I like), I'm not sure if this sentence can be removed without removing the two previous ones on the species concept differences.
 What do you think?
@@
 Our results suggest that this may be problematic at the genus-level in Carnivora, Chiroptera and Soricomorpha. 
-%For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade because they have no coded anatomical characters available.
 For example, a carnivoran fossil is unlikely to be placed in Herpestidae because they have no coded anatomical characters available.
 Instead the fossil will have a high probability of being placed on a branch that contains many anatomical characters such as within the Canidae or Ursidae (Figure \ref{Figure_example_coverage}B). 
-This is analogous to the problem of long-branch attraction/short branch repulsion, as one can think of herpestids as having zero-length branches for anatomical characters, and canids and ursids having long branches and thus "attracting" fossil placements.
+This is analogous to the problem of long-branch attraction/short branch repulsion, as one can think of Herpestidae as having zero-length branches for anatomical characters, and Canidae and Ursidae having long branches and thus ``attracting'' fossil placements.
+%TG: changed family names here to be consistent (going with always latin).
 
@@
 Therefore, care needs to be taken when deciding how to include these older matrices.
+
 % NC: I editted this as the reviewer is concerned that our analyses might be biased, not total evidence in general

### 2016-03-30 Manuscript/BiologyLetters/response2.tex
Dealt with Natalie's comments (still need to sort couple of comments)

@@
 \bigskip
-%We are very happy on the manuscript acceptance decision. % NC: Not usual to write this, of course we are!
+
 Thanks again to the reviewers for their thoughtful comments on our manuscript.
@@
 To improve the clarity of this document, we dealt with each comment in the order they appeared in your ``Decision on Manuscript'' email and we have included the reviewers comments in blue.
-
-% NC: Bit weird to have referee two and then 1, I've moved it back so it's in the right order! always easier for an editor to read like that than to have to go back and forth.
 
@@
 
-We agreed with the reviewer in the previous round of reviews and thought we had removed all mentions of cladistics.
-It was still used twice in the manuscript (and once in the keywords) but we have now corrected this error.
-% NC: Fix these. And check the supp material.
+We agreed with the reviewer in the previous round of reviews and thought we had removed all mentions of cladistics. It was still used twice in the manuscript but we have now corrected this error.
 
@@
 ``A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared between both living and fossil taxa (i.e.
-hard tissue characters such as skeletal ones).''
+hard tissue characters such as skeletal features).''
 lines @@@.
-% NC: I've edited this slightly.
-Note that the issue is that even though we might find loads of data for a living mammal, if it is all soft tissue it is USELESS!
-It cannot overlap with fossil taxa.
-In an ideal world we would be just looking for hard characters, but we've done this now so instead we will have to mention this in the discussion somewhere
 
 We also added a comment in the discussion to the effect that having data for a living mammal doesn't necessarily mean it can overlap with a fossil - it could all be soft tissue characters.
-However, in general our datasets appear to contain at least one hard character for each living species so it shouldn't influence our results.
+However, in general our datasets appear to contain at least some hard character for each living species so it shouldn't influence our results.
 But yes it is an important consideration for future work and we thank the reviewer for pointing this out (we misunderstood this point in the first round of revisions).
-% NC: We did misunderstand this first go round.
-This is why I don't like you to rush.
-You need to learn how to very carefully read a reviewers letter and understand the criticisms.
-Remember one paragraph of comments might have several things to fix.
 
 % NC: Need to add a line or two for this in the discussion. Or maybe rephrase the above and we can just stick with the clarifications I've added to the methods and intro.
+% TO DO
 
 % NC: I've changed a few bits in the text too
+% TO DO
 
@@
 We agree with this comment (and Referee 3 comment 2) and modified the paragraph discussing this caveat as follows:
-``Matrices containing anatomical characters were much more common before the advent of molecular phylogenetics and therefore are also more likely to be unavailable in a reusable format.
-This might include some bias in Total Evidence analyses.
-Nonetheless, these matrices are also likely to differ from more recent ones in terms of their underlying definition of homology and their coding practices (see [19]).
-Additionally, many recent morphological matrices reuse living taxa from previous matrices (see ESM1).''
+``Although this will bias our results towards lower coverage we do not think this bias will be large, as many recent morphological matrices reuse living taxa characters from older matrices (see ESM1), so many of these data will be present in our analyses.
+Additionally, these older matrices are likely to differ from more recent ones in terms of their underlying definition of homology and their coding practices (see [19]).
+Therefore, care needs to be taken when deciding how to include these older matrices.
+''
+
 lines @@@.
-
-%TG: Here's how I went with the section, basically, I acknowledge referee 1's points but add Peters ones just after:
-%TG: - Yes, we specifically ignored old matrices because definition of accesible data (see above)
-%TG: - Yes, we do agree it could include some bias
-%TG: - But these matrices are probably not so good anyway (Peter's comment)
-%TG: - And anyway it doesn't solve the problem because the living taxa used in these "old" matrices are reused in new ones anyway (or at least that's what I observed in primates)
-%TG: Personally I doubt the problem is due to data availability but way more to the problem we underlined in the abstract: palaeontologist don't wanna waste their time with living species (and I don't specially blame them).
 
@@
 % Make sure you also write a brief explanation here as well as editting the text. Perhaps add a line like:
-
+% TO DO
 We are unable to properly quantify the bias introduced by ignoring older matrices without going out and collecting them all exhaustively and re-running our analyses.
@@
 % TO DO (see Referee 1 comment 2)
-
 %NC: Also need to add a line sayong we've now mentioned this in the discussion and add the bit here.

### 2016-03-31 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Revision 2 round 2

@@
 We downloaded all discrete morphological matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{morphobank.org} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}).
-We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material (ESM) for details).
+We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material 1 (ESM1) for details).
 In total, we downloaded 286 matrices containing 5228 unique operational taxonomic units (OTUs). 
 We used OTUs rather than species because entries in the matrices ranged from species to families. 
-We standardised the taxonomy as described in the ESM and excluded OTUs that were not present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal} to remove fossil species.
+We standardised the taxonomy as described in the ESM1 and excluded OTUs that were not present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal} to remove fossil species.
 This resulted in 1601 unique OTUs from 286 matrices.
@@
 
-% NC: The ref for our paper looks weird - it has a url, probably fix that and check the reference is correct so we get the citation.
-% Ref 10 - there's an 03 after the date. Not sure what that refers to.
 % NC: Ref 16 - is KPg mising a - ?
-% NC: Ref 17 has a missing space
 

### 2016-03-31 Manuscript/BiologyLetters/response2.tex
Revision 2 round 2

@@
 We emphasised the importance of skeletal characters in the introduction as follows:
-``A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared between both living and fossil taxa (i.e.
-hard tissue characters such as skeletal features).''
+\textit{``A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared between both living and fossil taxa (i.e.
+hard tissue characters such as skeletal features).''}
 lines @@@.
 
-We also added a comment in the discussion to the effect that having data for a living mammal doesn't necessarily mean it can overlap with a fossil - it could all be soft tissue characters.
+We also added a comment in the discussion to the effect that having data for living mammals doesn't necessarily mean it can overlap with a fossil - it could all be soft tissue characters for example.
 However, in general our datasets appear to contain at least some hard character for each living species so it shouldn't influence our results.
@@
 
-% NC: Need to add a line or two for this in the discussion. Or maybe rephrase the above and we can just stick with the clarifications I've added to the methods and intro.
-% TO DO
-
-% NC: I've changed a few bits in the text too
-% TO DO
+We add the following sentence to the discussion:
+\textit{``Additionally, our analyses may underestimate the problem as we do not distiguish between soft and hard tissue characters; if a living taxon has only soft tissue coded anatomical characters then it will not have overlapping data with fossils that only have hard tissues preserved.''}
+lines @@@.
 
@@
 We agree with this comment (and Referee 3 comment 2) and modified the paragraph discussing this caveat as follows:
-``Although this will bias our results towards lower coverage we do not think this bias will be large, as many recent morphological matrices reuse living taxa characters from older matrices (see ESM1), so many of these data will be present in our analyses.
+\textit{``Although this will bias our results towards lower coverage we do not think this bias will be large, as many recent morphological matrices reuse living taxa characters from older matrices (see ESM1), so many of these data will be present in our analyses.
 Additionally, these older matrices are likely to differ from more recent ones in terms of their underlying definition of homology and their coding practices (see [19]).
-Therefore, care needs to be taken when deciding how to include these older matrices.
-''
+Therefore, care needs to be taken when deciding how to include these older matrices.''} lines @@@.
 
-lines @@@.
-
-% NC: I've editted this a bit to get more closely at what the reviewer wants (it thinks OUR analysis might be biased, and is correct).
-% Make sure you also write a brief explanation here as well as editting the text. Perhaps add a line like:
-% TO DO
-We are unable to properly quantify the bias introduced by ignoring older matrices without going out and collecting them all exhaustively and re-running our analyses.
-This would take a substantial amount of time not available at present.
+Since we did not collect these ``older'' matrices, we are unable to properly quantify the bias introduced by ignoring them.
+To test it we should rerun the whole analysis from the data collection to include them and to see the difference between the present data and the ones including these ``older'' matrices.
+This would really interesting but would also take a substantial amount of time not available at present.
 
@@
 I'd like to see it moderated a bit in the final version of the manuscript.}}
+
 Following Referee's 1 and 4's comments, we removed this paragraph.
@@
 \item{\textcolor{blue}{line 34: spelling: ccombines = combines}}
+
 We fixed this typo.
@@
 \item{\textcolor{blue}{lines 44 - 52. I'd suggest rewording this a little bit as lines 48 - 52 are a little awkward. May I suggest the following:
-"We might expect the total evidence method to perform poorly when there is little overlap between coded anatomical characters in living and fossil taxa.
+``We might expect the total evidence method to perform poorly when there is little overlap between coded anatomical characters in living and fossil taxa.
 This is because fossil taxa cannot be correctly placed within a clade of living species if none of its members have been coded for morphological characters.
-Furthermore, simulations show that fossils are more likely to be placed in clades for which more characters have been coded, regardless of whether this is the correct clade [8]."}}
+Furthermore, simulations show that fossils are more likely to be placed in clades for which more characters have been coded, regardless of whether this is the correct clade [8].''}}
+
 We replaced the section with Referee 2's suggestion.
@@
 \item{\textcolor{blue}{line 152: ... a carnivoran fossil is unlikely to be placed in Herpestidae because ....}}
+
 We fixed this typo.
 
-\item{\textcolor{blue}{line 159: I'd replace "It is worth noting" with "We acknowledge"}}
+\item{\textcolor{blue}{line 159: I'd replace ``It is worth noting'' with``We acknowledge''}}
+
 We changed the start of the sentence as suggested.
@@
 \item{\textcolor{blue}{line 161: word order: provided easily accessible matrices}}
+
 We changed the word order.
@@
 \item{\textcolor{blue}{The new title is an improvement.
-However, I think that it could be altered still further: would something like: "Assessment of available anatomical characters for linking living mammals to fossil taxa in phylogenetic analyses"?
-That is the main thrust of the paper, after all, and it also avoids having someone simply think: "Who cares?
-I have molecules" and moving on to the next article.}}
+However, I think that it could be altered still further: would something like: ``Assessment of available anatomical characters for linking living mammals to fossil taxa in phylogenetic analyses``"''?
+That is the main thrust of the paper, after all, and it also avoids having someone simply think: ``Who cares?
+I have molecules'' and moving on to the next article.}}
+
 We changed the title as suggested by Referee 3.
 
-\item{\textcolor{blue}{Lines 159 - 160: "It is worth noting, however, that our analysis did not include all the matrices containing anatomical characters ever published."
+\item{\textcolor{blue}{Lines 159 - 160: ``It is worth noting, however, that our analysis did not include all the matrices containing anatomical characters ever published''.
 One thing to note is that a lot of the older analyses, which tended to be morphological, were done in the early days of modern phylogenetics, and those are the matrices that are most apt to not have pdf or html versions.
-So, there might be a bias against the "pioneering" studies.
+So, there might be a bias against the ``pioneering``"'' studies.
 However, many of those studies will be a quarter of a century old, and the homology concepts and (especially!) coding protocols being used back then almost certainly differ from now.
-Also, a lot of those matrices were "cleaned" to remove homoplasy: this was not attempted fraud, but due to the fact that computer runs that are done in hours today took weeks and weeks back then, and actually storing all of the trees that might come out of "messy" matrices was beyond the old floppy drives.
+Also, a lot of those matrices were ``cleaned``"'' to remove homoplasy: this was not attempted fraud, but due to the fact that computer runs that are done in hours today took weeks and weeks back then, and actually storing all of the trees that might come out of ``messy'' matrices was beyond the old floppy drives.
 That is one reason why I rarely use matrices published before the early 1990's in my own meta-analyses.
-So, if those matrices do exist and can be computerized, then I think that they should be viewed as "starting points" for coding taxa: they almost certainly will need updating.}}
+So, if those matrices do exist and can be computerized, then I think that they should be viewed as ``starting points``"'' for coding taxa: they almost certainly will need updating.}}
 
-% TO DO (see Referee 1 comment 2)
-%NC: Also need to add a line sayong we've now mentioned this in the discussion and add the bit here.
+We dealt with this comment above and mentioned this in the discussion (see Referee 1, comment 2).
 
@@
 
-\item{\textcolor{blue}{In particular I agree with Referee 1 regarding the continued use of "cladistic" and the genus-level issue (many palaeontologists are now focused on the species-level and certainly many mammalian genera are not monotypic - this is more of an issue with dinosaur workers), and Referee 3 (Pete Wagner's) title suggestion is - I think - more likely to attract readers.}}
+\item{\textcolor{blue}{In particular I agree with Referee 1 regarding the continued use of ``cladistic``"'' and the genus-level issue (many palaeontologists are now focused on the species-level and certainly many mammalian genera are not monotypic - this is more of an issue with dinosaur workers), and Referee 3 (Pete Wagner's) title suggestion is - I think - more likely to attract readers.}}
+
 We dealt with these suggestions (see respectively Referee 1 comment 1, Referee 1 comment 3 and Referee 3 comment 1).
 
-\item{\textcolor{blue}{on L174 ("but also to any researcher focusing understanding macroevolutionary patterns and processes"), this seems to be missing an "on".}}
+\item{\textcolor{blue}{on L174 ("but also to any researcher focusing understanding macroevolutionary patterns and processes"), this seems to be missing an ``on``"''.}}
+
 We fixed this typo.

### 2016-04-04 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Final edits after response 2

@@
 \medskip
-\noindent{*\bf Corresponding author.
-Current address: } \textit{t.guillerme@imperial.ac.uk}\\  % NC: You need to add your current address to the corresponding author bit I think.
+\noindent{*\bf Corresponding author.} \textit{t.guillerme@imperial.ac.uk}\\  % NC: You need to add your current address to the corresponding author bit I think.
 TG: i.e.
@@
 That'll eat a shit load of words for not much no?
+% NC: OK, leave it for now, but you are supposed to use a current address for corresponding author if you have moved.
+It's leftover from the days people would write you a letter to ask about your paper!
+Hopefully Biology Letters won't notice :)
 \vspace{1in}
@@
 11 of 28 mammalian orders have \textless 25\% species with available characters; this has implications for the accurate placement of fossils, although the issue is less pronounced at higher taxonomic levels.
-In most orders, species with available characters are randomly distributed across the phylogeny, that may reduce the impact of the problem.
+In most orders, species with available characters are randomly distributed across the phylogeny, which may reduce the impact of the problem. % NC: Sorry can't decide if it is a that or a which :)
 We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence phylogenies. 
@@
 
-\noindent Key words: Total Evidence method, phylogenetic clustering, discrete morphological matrix, extinct, topology\\
+\noindent Key words: Total Evidence method, phylogenetic clustering, discrete morphological matrix, extinct, topology.\\
 
@@
 \cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a chronogram with living and fossil taxa at the tips.
-A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared between both living and fossil taxa (i.e.
+A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared among both living and fossil taxa (i.e.
 hard tissue characters such as skeletal features).
-Sections of this data can be difficult, or impossible, to collect for every taxon in the analysis.
+% NC: across or among is correct.
+Between is only for two things being compared.
+Sections of these data can be difficult, or impossible, to collect for every taxon in the analysis.
 For example, fossils rarely have molecular data and incomplete fossil preservation may reduce the number of anatomical characters available.
-Additionally, it has become less common to collect anatomical characters for living taxa when molecular data is available (e.g.
+Additionally, it has become less common to collect anatomical characters for living taxa when molecular data are available (e.g.
 in \cite{slaterphylogenetic2013}, only 13\% of living taxa have coded anatomical characters).
@@
 However, it is unclear how many coded anatomical characters are actually available for living taxa, i.e.
-already coded from museum specimens and deposited in phylogenetic matrices accessible online, and how this data is distributed across clades.
-Intuitively, most people assume this data has already been collected, but empirical analyses suggest otherwise (e.g. in \cite{ronquista2012,slaterphylogenetic2013,beckancient2014}).
-To investigate this further, we assess the number of available coded anatomical characters for living mammals to determine whether enough data exists to build reliable Total Evidence phylogenies.
+already coded from museum specimens and deposited in phylogenetic matrices accessible online, and how these data are distributed across clades.
+Intuitively, most people assume these data have already been collected, but empirical analyses suggest otherwise (e.g. in \cite{ronquista2012,slaterphylogenetic2013,beckancient2014}).
+To investigate this further, we assess the number of available coded anatomical characters for living mammals to determine whether enough data exist to build reliable Total Evidence phylogenies.
 We also determine whether the characters are phylogenetically overdispersed or clustered across mammalian orders.
-%TG: a point to be made here maybe is that we do NOT look at hard tissue characters, rather just the presence of any characters (which is already poor enough)?
-%TG: on the other hand, we might not need to emphasize that to much since, anyway, it is clearly explained in the methods (i.e. we tool all cladistic data from matrices)
-% NC: If someone makes a point about it in a review it means it isn't explained clearly enough. I think adding a note to the methods and/or discussion is needed.
+
 %---------------------------------------------
@@
 %---------------------------------------------
-% NC: From here on in we need to replace the word cladistic. I don't like the "anatomical characters" as it's a bit long. How about character matrices? You decide.
-% TO DO
 \section{Materials and Methods}
@@
 We do not distinguish between soft and hard characters, but the majority of matrices contain at least some hard tissue characters. 
-% NC: True?
-This is to deal with the skeletal bit.
-Put into response letter.
-TG: probably true.
-Didn't check all matrices but because of their data's nature (i.e.
-collected by anatomists back in the days and paleontologists nowadays) I'll bet you there's "at least some".
 We consider orders with \textless 25\% of living taxa with available anatomical characters as having low data coverage, and orders with \textgreater 75\% of living taxa with available anatomical characters as having high data coverage.
@@
 At the genus-level, one order had low coverage and 15 had high coverage, and at the family-level, no orders had low coverage and 25 had high coverage (Table\ref{Table_results}).
-
-%TG: former results:
-% 22 of 28 orders have low coverage (\textless 25\% species with available anatomical characters) and six have high coverage (\textgreater 75\% species with available anatomical characters) at the species-level.
-% At the genus-level, three orders have low coverage and 12 have high coverage, and at the family-level, no orders have low coverage and 23 have high coverage (Table\ref{Table_results}).
-
-%TG: don't clean up this caption please!
-%\caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels.
-The left vertical bar represents low coverage (\textless 25\%); the right vertical bar represents high coverage (\textgreater 75\%).
-Negative Net Relatedness Index (NRI) and Nearest Taxon Index (NTI) values indicate phylogenetic overdispersion; positive values indicate phylogenetic clustering.
-Significant NRI or NTI values are in bold.
-*p \textless 0.05; **p \textless 0.01; ***p \textless 0.001.} \\
-%TG: don't clean up this caption please!
-
-% NC: Can you check the caption we used doesn't have cladistic in it? I have a feeling it does...
-% TO DO
 
@@
 Additionally, our analyses may underestimate the problem as we do not distiguish between soft and hard tissue characters; if a living taxon has only soft tissue coded anatomical characters then it will not have overlapping data with fossils that only have hard tissues preserved.
-%NC: This deals with reviewer 1.
-What do you think?
-TG: Yap, excellent.
-
-% This is important in practice because of the slight discrepancy between neontological and palaeontological species concepts.
-% While neontological species are described using morphology, genes, distribution, etc.; palaeontological species can be based only on morphological/anatomical, spatial and temporal data (e.g.
-\cite{ni2013oldest}).
-% Therefore, many palaeontological studies use the genus (or monospecific genera) as their smallest OTU (e.g.
-\cite{ni2013oldest,O'Leary08022013}), so data availability at the genus-level in living mammals should be our primary concern when building phylogenies of living and fossil taxa.
-%
-%
-%TG: My first feeling was to just remove the last sentence but, when reading the previous one (which I like), I'm not sure if this sentence can be removed without removing the two previous ones on the species concept differences.
-What do you think?
-Anyway, it would then maybe just be better to dump the three sentences an finish this paragraph at "Therefore we expect difficulties in placing fossils at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels."
-%TG: On the other hand removing that gives us space for discussing the data sampling problem below
-%
-% NC: Yeah I think this is the correct thing to do here.
 
@@
 Then we expect two major biases: first, fossils will not be placed within a clade containing no hard tissue data, and second, fossils will have higher probability of being placed within the most sampled clade by chance.
-% NC: I added hard tissue here
 Our results suggest that this may be problematic at the genus-level in Carnivora, Chiroptera and Soricomorpha. 
@@
 This is analogous to the problem of long-branch attraction/short branch repulsion, as one can think of Herpestidae as having zero-length branches for anatomical characters, and Canidae and Ursidae having long branches and thus ``attracting'' fossil placements.
-%TG: changed family names here to be consistent (going with always latin).
 
@@
 
-% NC: I editted this as the reviewer is concerned that our analyses might be biased, not total evidence in general
 Despite the absence of good morphological/anatomical data coverage for living mammals, the Total Evidence method still seems to be the most promising way of combining living and fossil species in macroevolutionary analyses.
-Following the recommendations in \cite{GuillermeCooper}, we must code anatomical characters for as many living species as possible. 
+Following the recommendations in \cite{GuillermeCooper}, we should code anatomical characters for as many living species as possible. 
 Fortunately, mammal specimens are usually readily available in natural history collections, therefore, we propose increased effort into coding anatomical characters from living species, possibly by engaging in collaborative data collection projects.
@@
 
-% NC: Ref 16 - is KPg mising a - ?
+% NC: Ref 5 - J Evol Biol
+% NC: Ref 8. Still that weird "Available from" bit in the ref to our paper. This is not needed and wasting words. Plus looks odd.
+% NC: Ref 12 has p. 
+% NC: Ref 16 - Now this says KP-g which is wrong! K-Pg or KPg (not sure what it says in the paper title)
 

### 2016-04-04 Manuscript/BiologyLetters/response2.tex

@@
 We emphasised the importance of skeletal characters in the introduction as follows:
-\textit{``A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared between both living and fossil taxa (i.e.
+\textit{``A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared among both living and fossil taxa (i.e.
 hard tissue characters such as skeletal features).''}
 lines @@@.
+% NC: Add line numbers
 
@@
 lines @@@.
-
+% NC: Add line numbers
 \item{\textcolor{blue}{I'm also concerned that my original comments about the biases in character sets included still aren't adequately addressed; they now acknowledge that they didn't include any character sets that weren't available in print, but that doesn't address what kind of difference it would make to their results.
@@
 
-We agree with this comment (and Referee 3 comment 2) and modified the paragraph discussing this caveat as follows:
+% NC: I've rewritten this to make it clearer what you have done and to stop it reading like "yeah we know there is a bias but we are too lazy to fix it".
+Now it says "yeah we know there is a bias but we are too lazy to fix it, but we also think it isn't a big bias and would add more bias anyway".
+A much better justification.
+We acknowledge this comment (and Referee 3 comment 2) and have now made more explicit reference to this bias. 
+However, because we did not collect these matrices, we are unable to properly quantify the bias introduced by excluding them.
+To test this we would need to collect these data and then rerun the analysis.
+This would really interesting but would also take a substantial amount of time to collect the data, which is not available at present. 
+Additionally we do not believe this bias will be as large as the reviewer fears.
+Many recent matrices reuse characters from older matrices, thus it is likely most of the data from major previous studies are present in our analyses.
+We also share the concerns of Peter Wagner that using ``old'' matrices may introduce a different kind of bias into phylogenies, as definitions of homology and characters have changed over time, making the kind of matrix we are unlikely to find with our search protocol harder to incorporate into Total Evidence analyses.
+We therefore modified the paragraph discussing this bias as follows:
 \textit{``Although this will bias our results towards lower coverage we do not think this bias will be large, as many recent morphological matrices reuse living taxa characters from older matrices (see ESM1), so many of these data will be present in our analyses.
@@
 Therefore, care needs to be taken when deciding how to include these older matrices.''} lines @@@.
-
-Since we did not collect these ``older'' matrices, we are unable to properly quantify the bias introduced by ignoring them.
-To test it we should rerun the whole analysis from the data collection to include them and to see the difference between the present data and the ones including these ``older'' matrices.
-This would really interesting but would also take a substantial amount of time not available at present.
-
+% NC: Add line numbers
 \item{\textcolor{blue}{I don't agree with the assertion that the genus level should be our primary concern; more and more studies are building species-level trees with fossil taxa.

### 2016-04-04 Manuscript/response.tex
Cleaned up manuscript file

--- 2016-02-13 Manuscript/BiologyLetters/response.tex
@@
 \section{Reviewer 1:}
+
 \begin{enumerate}
@@
 I need evidence to support this cutoff; as it stands, I'm concerned that you're missing some relevant material.}}
-% NC: It's not just the first 10 results per search term is it?
-I thought you went through the first 10 pages of results?
-Clarify this.
-TG: No I collected the 20 first results.
-It's mentioned down there.
 
@@
 We are aware however that we might have missed some specific studies (see Reviewer 2 comment 2) but are confident that our protocol gives a good overall picture of the data availability and that the three repositories used before the Google Scholar search (MorphoBank, Graeme Lloyd's and Ross Mounce's) contained most of the available data.
-In fact, as mentioned in ESM 1, part 1 ``among the 660 papers [collected from the Google Scholar search], only 50 contained a total of 425 extra living OTUs on a total of 4950 OTUs''.
+In fact, as mentioned in ESM 1, part 1 ``in the 660 papers [collected from the Google Scholar search], only 50 contained extra living OTUs and only contributed 425 OTUs to our total of 4950 OTUs''.
 
@@
 \begin{enumerate}
-\item{\textcolor{blue}{The authors state on page 3 lines 82-90 that they avoided matrices with ``few characters'' due to lower probability of overlapping characters with other matrices and so only selected matrices containing >100 characters.
+\item{\textcolor{blue}{The authors state on page 3 lines 82-90 that they avoided matrices with ``few characters'' due to lower probability of overlapping characters with other matrices and so only selected matrices containing $>$100 characters.
 First, a lack of overlapping characters doesn’t seem justifiable, at least to my mind, as a reason to exclude a matrix if the ultimate goal is to infer the placement of extinct taxa relative to extant taxa.
@@
 Looking at Figure 1B, again the clade I'm most familiar with, representation for carnivorans appears very poor.
-Notwithstanding the issues discussed above regarding matrices with < 100 characters, there are definitely matrices out there that fulfill the authors' requirements that are not represented.
+Notwithstanding the issues discussed above regarding matrices with $<$100 characters, there are definitely matrices out there that fulfill the authors' requirements that are not represented.
 For example, a 349 character matrix with near complete sampling at the species is available for Viverridae, Prionodontidae, and Eupleridae in Gaubert et al (2005 Syst Biol.).
@@
 In fact, the matrix mentioned as Appendix 2 in the paper is not present in the article and neither of the TreeBase accession numbers (``S1255'', ``M2193'') nor the provided URL (``www.systematicbiology.org'') links to the matrix.
-% TG: I don't know if that's a good justification but basically, it's impossible to get this specific matrix.
-% NC: Hmmm it's tricky as I said above.
-If you were really building a tree you would find this kind of data somehow.
-Also wouldn't this not appear because you used the 2010 cutoff?
-I think tbh Graham has some good points - you did this the quick and dirty way rather than the thorough way someone would do it if they were constructing a tree.
 
@@
 We added this analysis to the ESM 2 (Supplementary results - 1) showing, as suggested, that mammalian morphological phylogeneticists have focused their energies on resolving the higher level relationships among mammalian lineages.
-% We also added a sentence mentioning these results in the manuscript:
-
-% ``It is also encouraging to note that the sampling effort is consistent with diversity at each taxonomic level and there is therefore no direct bias in sampling at any taxonomic level (ESM 2, Figure 1).
-% '' p.@ lines@@
-
-% NC: OK this is not what Graham meant.
-I think what he is referring to is that when you have more people working on a group (effort) you are likely to get more taxa with coded data.
-It's like in taxonomy where groups with more taxonomists (eg primates) have more species than groups with fewer (eg rodents).
-I think what he'd want is a regression of number of papers (effort) against percentage of coded species/OTUs with each order being a datapoint.
 
@@
 \item{\textcolor{blue}{My biggest criticism is a semantic one: Morphological data $\neq$ cladistic data.}}
-We agree with this reviewers semantinc point and changed ``cladistic'' and ``morphological'' data to ``anatomical characters'' throughout the manuscript.
+We agree with this reviewers semantic point and changed the first mention of ``cladistic'' data to ``discrete morphological/anatomical data'' (p.2 line 38) and then to ``anatomical characters'' throughout the manuscript.
 
@@
 Instead the fossil will have a high probability of being placed on a branch that contains many anatomical characters such as within the Canidae or Ursidae (Figure 1B). 
-This is analogous to the problem of long-branch attraction/short branch repulsion, as one can think of herpestids as having zero-length branches for anatomical characters, and canids and ursids having long branches and thus ``attracting'' fossil placements.''
+This is analogous to the problem of long-branch attraction/short branch repulsion, as one can think of herpestids as having zero-length branches for anatomical characters, and canids and ursids having long branches and thus ``attracting'' fossil placements.'' p.5 lines 154-160.
 
@@
 
-\section{Reviewer 4:}
+\section{Reviewer 4 (Graeme Lloyd):}
 \begin{enumerate}
@@
 \item{\textcolor{blue}{Reviewer 2 (Graham Slater)'s complaint are - in my view - the most substantial.
-I agree that the >100 characters cut off is not as strongly justified as it might be.
+I agree that the $>$100 characters cut off is not as strongly justified as it might be.
 However, as the authors have run their protocol without this threshold this is easily fixed by at least briefly mentioning how this affects their results in the main text (not relegating it to the ESM where no actual discussion is provided).
@@
 \item{\textcolor{blue}{Line 104 - ``NTI 15'' should probably be ``NTI [15]''.}}
-%TO DO IN THE WORD MANUSCRIPT! 
 We fixed this typo.
@@
 
-% Editorial office comments to authors:
-% Please ensure that you include; 
-% *An author contributions section which fulfils all *four* of our criteria given here https://royalsociety.org/journals/ethics-policies/openness/, including a sentence stating all authors agree to be held accountable for the content therein and approve the final version of the manuscript
-%TG: Ok, we I'm not sure here.
-Is it just an automatic message: did all the author contributed well enough to be author (yes!); or is it that we have to modify the author contribution list to the exact wordings they put in this URL (seems odd though).
-
-% *A funding statement which is the same as that given in ScholarOne
-%TG: Yes, is matching exactly.
-
-% *Please also confirm in your cover letter whether all the figures are your own or whether permission has been obtained for their use
-%TG: TO DO (yes)
-% *Please upload your original figure files as eps, tiff or jpeg files, rather than PDFs
-%TG: TO DO (easy)
-% *If you have any images that can be used to promote your article on social media (should this be accepted) please upload them as a supplementary file
-%TG: what about the "phylogeny" with the duck and the pig?
-% *Please confirm whether your manuscript should be available via open access, as we note that this may be a requirement of your funder.
-%TG: TO DO (yes)
-
-%TG: just fix/add all this in this document
-%TG: Make sure word limit is kept to 2500 +/- 5%
-
-Thomas Guillerme (t.guillerme@imperial.ac.uk)\\
+Thomas Guillerme (t.guillerme@imperial.ac.uk),\\
 Natalie Cooper

### 2016-04-04 Manuscript/Missing_mammals-draft.tex
color testings
Cleaned up manuscript file

@@
 
-\noindent{\Large \bf Morphological data is lacking for living mammals}
-%Or: Missing cladistic data in living mammals - but I prefer the first one emphasizing on the availability of it. The data is not missing per se, it's just not coded!
-% NC: sometimes the better title is one that answers the question rather than giving the topic. e.g., "Morphological data is lacking for living species" or similar
-% TG: maybe worth considering the special issue as well? "Putting fossils in tree" so maybe something like "There is not enough morphological data for living mammals to branch fossils in the trees."
-% TG: or some poetic... "Living and dead mammals: morphological data is lacking in living for combined analysis"
+\noindent{\Large \bf Assessment of available anatomical characters for linking living mammals to fossil taxa in phylogenetic analyses}
+
 \bigskip
@@
 \medskip
-\noindent{*\bf Corresponding author.} \textit{Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.}\\  
+\noindent{*\bf Corresponding author.} \textit{t.guillerme@imperial.ac.uk}\\ 
 \vspace{1in}
 
-%Line numbering REMOVE FOR WORDCOUNT
-%\modulolinenumbers[1]
-%\linenumbers
+%Line numbering
+\modulolinenumbers[1]
+\linenumbers
 
@@
 %---------------------------------------------
-% NC: Time to sort this out!
 \newpage
-\begin{abstract} % 200 words, all keywords written in abstract if possible. TG: That's 195 words for now.
-Combining living and fossil in the same analysis data is crucial for studying changes in global biodiversity through time.
-One method allowing to combine this data is the Total Evidence method that uses both molecular data for living species and morphological data for both living and fossil species.
-With this method, a good overlap of morphological data between living and fossil taxa is crucial for accurately inferring the phylogenies’ topology.
+\begin{abstract}
+\section{abstract}
 
-Since the advent of DNA, molecular data has become easily and widely available.
-However, despite two centuries of morphological studies, scientists using and generating such data mainly focus on palaeontological data.
-Therefore, there is a gap in our knowledge of neontological morphological data even in well studied groups such as mammals.
+Analyses of living and fossil taxa are crucial for understanding biodiversity through time.
+The Total Evidence method allows living and fossil taxa to be combined in phylogenies, using molecular data for living taxa and morphological data for living and fossil taxa.
+With this method, substantial overlap of coded anatomical characters among living and fossil taxa is vital for accurately inferring topology.
+However, although molecular data for living species are widely available, scientists generating morphological data mainly focus on fossils.
+Therefore, there are fewer coded anatomical characters in living taxa, even in well-studied groups like mammals.
 
-In this study, we quantify the morphological data available for living mammal taxa.
-We then analyse the structure of the available data by testing if it is clustered or evenly spread across the phylogeny.
-We found that 78\% of mammalian orders have less than 25\% data available at the species level.
-However, we found that the available is often randomly distributed among these orders apart from six of them where the data is clustered
+We investigated the number of coded anatomical characters available in phylogenetic matrices for living mammals and how these were phylogenetically distributed across orders.
+11 of 28 mammalian orders have \textless 25\% species with available characters; this has implications for the accurate placement of fossils, although the issue is less pronounced at higher taxonomic levels.
+In most orders, species with available characters are randomly distributed across the phylogeny, which may reduce the impact of the problem.
+We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence phylogenies. 
 \end{abstract}
 
-%\noindent Key words: Total Evidence method, data structure, phylogenetic, fossil, topology\\ Excluded from word count.
+\noindent Key words: Total Evidence method, phylogenetic clustering, discrete morphological matrix, extinct, topology.\\
 
@@
 \section{Introduction}
-There is an increasing consensus among evolutionary biologists that studying both living and fossil taxa is essential for fully understanding macroevolutionary patterns and processes \cite{slaterunifying2013,fritzdiversity2013,Wood01032013}.
-For example, including both living and fossil taxa in evolutionary studies can improve the accuracy of timing diversification events (e.g.
-\cite{ronquista2012}, our understanding of relationships among lineages (e.g.
-\cite{beckancient2014}, and our ability to infer biogeographical patterns through time (e.g.
-\cite{Meseguer01032015}.
+There is an increasing consensus among biologists that studying both living and fossil taxa is essential for fully understanding macroevolutionary patterns and processes \cite{slaterunifying2013,fritzdiversity2013}.
 To perform such analyses it is necessary to combine living and fossil taxa in phylogenetic trees.
-One increasingly popular method, the Total Evidence method \cite{eernissetaxonomic1993,ronquista2012}, combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix (e.g.
-\cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014,Meseguer01032015}, producing a phylogeny with living and fossil taxa at the tips.
-These phylogenies can be dated using methods such as tip-dating \cite{ronquista2012,Wood01032013} and incorporated into macroevolutionary studies (e.g.
-\cite{ronquista2012,Wood01032013,slaterphylogenetic2013}.
+One increasingly popular method, the Total Evidence method \cite{ronquista2012}, combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix that can then be used with the tip-dating method (e.g.
+\cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a chronogram with living and fossil taxa at the tips.
+A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared among both living and fossil taxa (i.e.
+hard tissue characters such as skeletal features).
+Sections of these data can be difficult, or impossible, to collect for every taxon in the analysis.
+For example, fossils rarely have molecular data and incomplete fossil preservation may reduce the number of anatomical characters available.
+Additionally, it has become less common to collect anatomical characters for living taxa when molecular data are available (e.g.
+in \cite{slaterphylogenetic2013}, only 13\% of living taxa have coded anatomical characters).
+Unfortunately this missing data can lead to errors in phylogenetic inference.
+We might expect the Total Evidence method to perform poorly when there is little overlap between coded anatomical characters in living and fossil taxa, because fossil taxa cannot be correctly placed within a clade of living species with no coded characters.
+Furthermore, simulations show that fossils are more likely to be placed in clades for which more characters have been coded, regardless of whether this is the correct clade \cite{GuillermeCooper}.
 
-A downside of the Total Evidence method is that it requires a lot of data.
-One must collect molecular data for living taxa and morphological data for both living and fossil taxa; two types of data that require fairly different technical skills (e.g.
-molecular sequencing \textit{vs.} anatomical description).
-Additionally, large chunks of this data can be difficult, or even impossible, to collect for every taxon present in the analysis.
-For example, fossils very rarely have molecular data and incomplete fossil preservation (e.g.
-soft \textit{vs.} hard tissues) may restrict the amount of morphological data available \cite{sansomfossilization2013}.
-Additionally, since the molecular phylogenetics revolution, it has become less common to collect morphological characters for living taxa when molecular data is available (e.g.
-in \cite{slaterphylogenetic2013}, only 13\% of the 169 living taxa have coded morphological data).
-Unfortunately this missing data can lead to errors in phylogenetic inference; in fact, simulations show that the ability of the Total Evidence method to recover the correct phylogenetic topology decreases when there is a low overlap between morphological data in the living and fossil taxa \cite{GuillermeCooper}, regardless the overall amount of morphological data available for the fossils (or the amount of molecular data available for the living species).
-The effect of missing data on topology is greatest when living taxa have few morphological data.
-This is because (1) a fossil cannot branch in the correct clade if there is no overlapping morphological data in the clade; and (2) a fossil has a higher probability of branching within a clade with more morphological data available for living taxa, regardless of whether this is the correct clade or not \cite{GuillermeCooper}.
+The issues above highlight that it is crucial to have sufficient coded anatomical characters available for living taxa in a clade before using the Total Evidence approach.
+However, it is unclear how many coded anatomical characters are actually available for living taxa, i.e.
+already coded from museum specimens and deposited in phylogenetic matrices accessible online, and how these data are distributed across clades.
+Intuitively, most people assume these data have already been collected, but empirical analyses suggest otherwise (e.g. in \cite{ronquista2012,slaterphylogenetic2013,beckancient2014}).
+To investigate this further, we assess the number of available coded anatomical characters for living mammals to determine whether enough data exist to build reliable Total Evidence phylogenies.
+We also determine whether the characters are phylogenetically overdispersed or clustered across mammalian orders.
 
-The issues above highlight that it is crucial to have sufficient morphological data for living taxa in a clade before using a Total Evidence approach.
-However, it is unclear how much morphological data for living taxa is actually available (i.e.
-already coded from museum specimens and deposited in phylogenetic matrices accessible online), and how this data is distributed across clades.
-Intuitively, most people assume this kind of data has already been collected, but empirical data suggest otherwise (e.g. in \cite{ronquista2012,slaterphylogenetic2013,beckancient2014}).
-To investigate this further, we assess the amount of available morphological data for living mammals to determine whether sufficient data exists to build reliable Total Evidence phylogenies in this group.
-We collected cladistic data (i.e.
-discrete morphological characters used in phylogenetics) from 286 phylogenetic matrices available online and measured the proportion of cladistic data available for each mammalian order.
-%Additionally, if the available data is randomly distributed across clades, the topological errors should be less extreme than if the data is biased towards particulars groups.
-Additionally, because missing morphological data in living species can influence tree topology as described above, %NC: Might still need to fix this sentence.
-"Topological caveats above" was too unclear.
-we determined whether the available cladistic data was phylogenetically overdispersed or clustered in the mammalian orders where data was missing. 
-We find that available morphological data for living mammals is scarce but generally randomly distributed across phylogenies. 
-We recommend that efforts be made to collect and share more cladistic data for living species to improve the accuracy of Total Evidence phylogenies.
-
-% NC: If there are word count problems we can ditch the last two sentences.
 %---------------------------------------------
@@
 \subsection{Data collection and standardisation}
-We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases (accessed 10th of June 2015): Morphobank (\url{http://www.morphobank.org/} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{https://github.com/rossmounce/cladistic-data}).
-We also performed a systematic Google Scholar search (accessed 11th of June 2015) for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
-In total, we downloaded 286 matrices containing a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique. % Number of unique OTUs (including fossils)
-In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardised and ranged from specific individual specimen names (i.e.
-the name of a collection item) to the family-level.
-Where possible, we considered OTUs at their lowest valid taxonomic level (i.e. species) but some OTUs were only valid at a higher taxonomic level (e.g. genus or family).
-Therefore for some orders, we sampled more genera than species (Table \ref{Table_morpho_taxa_proportion}).
-
-To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H.
-sapiens} was transformed to \textit{Homo sapiens}).
-We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\url{https://paleobiodb.org/}).
-For OTUs that did not appear in these three sources, we first decomposed the name (i.e.
-\textit{Homo sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
-Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; see Supplementary Material for more details).
-
-The number of characters in each matrix ranged from 6 to 4541.
-Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having characters that overlap with those of other matrices \cite{wagner2000} and (2) they are more likely to contain a higher proportion of specific characters that are not-applicable across large clades (e.g.
-``antler ramifications'' is a character that is only applicable to Cervidae not all mammals \cite{Brazeau2011}).
-Therefore we selected only matrices containing \textgreater 100 characters for each OTU.
-This threshold was chosen to correspond with the number of characters used in \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
-Note that results of analyses with no character threshold are available in Supplementary Material. 
-After removing all matrices with \textless 100 characters, we retained 1074 unique living mammal OTUs from 126 matrices for our analyses. % 1601 unique living OTUs for 286 matrices (no threshold)
+We downloaded all discrete morphological matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{morphobank.org} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}).
+We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material 1 (ESM1) for details).
+In total, we downloaded 286 matrices containing 5228 unique operational taxonomic units (OTUs). 
+We used OTUs rather than species because entries in the matrices ranged from species to families. 
+We standardised the taxonomy as described in the ESM1 and excluded OTUs that were not present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal} to remove fossil species.
+This resulted in 1601 unique OTUs from 286 matrices.
 
 \subsection{Data availability and distribution}
-To assess the availability of cladistic data for each mammalian order, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species.
-%We used these different taxonomic levels because some clades are well covered at the family- or genus-level, but poorly covered at the species-level. % NC: Probably obvious?
-We consider orders with \textless 25\% of living taxa with cladistic data as having poor data coverage (``low'' coverage), and orders with \textgreater 75\% of living taxa with cladistic data as having good data coverage (hereafter ``high'' coverage).
+To assess the availability of coded anatomical characters for each mammalian order and across mammals, we calculated the percentage of OTUs with coded anatomical characters at three different taxonomic levels: family, genus and species.
+We do not distinguish between soft and hard characters, but the majority of matrices contain at least some hard tissue characters. 
+We consider orders with \textless 25\% of living taxa with available anatomical characters as having low data coverage, and orders with \textgreater 75\% of living taxa with available anatomical characters as having high data coverage.
 
-For orders with \textless 100\% cladistic data coverage at any taxonomic level, we investigated whether the available cladistic data was (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index (NTI; \cite{webb2002phylogenies}) and the Net Relatedness Index (NRI; \cite{webb2002phylogenies}).
-NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to clustering or overdispersion across the whole phylogeny \cite{Cooper2008}. 
+For each order and for all mammals, we investigated whether the available coded anatomical characters were (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index (NTI; \cite{webb2002phylogenies}) and the Net Relatedness Index (NRI; \cite{webb2002phylogenies}).
+NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to them across the whole phylogeny \cite{Cooper2008}. 
 Both metrics were calculated using the \texttt{picante} package in R \cite{picante,R}.
 
-NTI \cite{webb2002phylogenies} is based on mean nearest neighbour distance (MNND) and is calculated as follows:
+NTI \cite{webb2002phylogenies} is based on mean nearest neighbour distance ($MNND$) and is calculated as follows:
   \begin{equation}
-    NTI=-\left(\frac{\overline{MNND}_{obs}-\overline{MNND}_{n}}{\sigma(MNND_{n})}\right)
+    \text{NTI}=-\left(\frac{\overline{MNND}_{obs}-\overline{MNND}_{n}}{\sigma(MNND_{n})}\right)
   \end{equation}
-where $\overline{MNND}_{obs}$ is the observed mean distance between each of $n$ taxa with cladistic data and its nearest neighbour with cladistic data in the phylogeny, 
-$\overline{MNND}_{n}$ is the mean of 1000 mean MNND between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random MNND values.
-NRI is similar but is based on mean phylogenetic distance (MPD) as follows:
+where $\overline{MNND}_{obs}$ is the observed mean sum of the branch lengths between each of $n$ taxa with available coded anatomical characters and its nearest neighbour with available coded anatomical characters in the phylogeny,
+$\overline{MNND}_{n}$ is the mean of 1000 $MNND$ between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random $MNND$ values.
+NRI is calculated in the same way, but using the mean phylogenetic distance ($MPD$):
   \begin{equation}
-    NRI=-\left(\frac{\overline{MPD}_{obs}-\overline{MPD}_{n}}{\sigma(MPD_{n})}\right)
+    \text{NRI}=-\left(\frac{\overline{MPD}_{obs}-\overline{MPD}_{n}}{\sigma(MPD_{n})}\right)
   \end{equation}
-where $\overline{MPD}_{obs}$ is the observed mean phylogenetic distance of the tree containing only the $n$ taxa with cladistic data, $\overline{MPD}_{n}$ is the expected random MPD for $n$ taxa estimated by calculating the MPD from $n$ taxa randomly drawn from the phylogeny and repeated 1000 times, and $\sigma(MPD_{n})$ is the standard deviation of the 1000 random MPD values.
-% NC: This can be shortened a bit but it's only gonna save you 50 words max.
-Negative NTI and NRI values show that the focal taxa are more overdispersed across the phylogeny than expected by chance, and positive values reflect significant clustering.
+where $\overline{MPD}_{obs}$ is the observed mean phylogenetic branch length of the tree containing only the $n$ taxa with available coded anatomical characters.
+Negative NTI and NRI values show that the focal taxa are more overdispersed across the phylogeny than expected by chance, and positive values reflect clustering.
 
-We calculated NTI and NRI values for each mammalian order separately, at each different taxonomic level. 
-For each analysis our focal taxa were those with available cladistic data at that taxonomic level and the phylogeny was the phylogeny of the order pruned from \cite{BinindaEmonds}.
+We calculated NTI and NRI values for all mammals or each mammalian order separately, at each different taxonomic level. 
+For each analysis our focal taxa were those with available coded anatomical characters at that taxonomic level and the phylogeny was the order pruned from \cite{BinindaEmonds}.
 
@@
 \section{Results}
-Across the 126 cladistic matrices we extracted, 22 out of 28 mammalian orders have low coverage (\textless 25\% of species with cladistic data) and six have high coverage (\textgreater 75\% of species with cladistic data) at the species-level.
-At the genus-level, three orders have low coverage and 12 have high coverage, and at the family-level, no orders have low coverage and 23 have high coverage (Table \ref{Table_morpho_taxa_proportion}).
+Across mammals, species coverage was low (\textless 25\% species with available coded anatomical characters) but family coverage was high (\textgreater 75\% families with available coded anatomical characters).
+For each order, 11 out of 28 had low coverage and seven had high coverage at the species-level.
+At the genus-level, one order had low coverage and 15 had high coverage, and at the family-level, no orders had low coverage and 25 had high coverage (Table\ref{Table_results}).
 
 \input{Table_results}
-% NC: Can probably cut words from the table legend too
 
-Among the mammalian orders containing OTUs with no available cladistic data, only six orders had significantly clustered data (Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha at both species- and genus-level and Afrosoricida and Pholidota at the species-level only) and no order had significantly overdispersed data at any taxonomic level (Table~\ref{Table_data_structure}).
+Across mammals, taxa with available coded anatomical characters were significantly clustered using NTI at the species- and genus-level.
+For each order, only seven showed significant clustering (Cetartiodactyla, Cingulata, Pilosa and Rodentia at the species-level and Carnivora, Chiroptera and Soricomorpha at both species- and genus-level) and none showed significant overdispersion (Table~\ref{Table_results}).
 
-Two contrasting results are shown in Figure \ref{Figure_example_coverage} with randomly distributed OTUs with cladistic data in Primates (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered OTUs with cladistic data in Carnivora (mainly Canidae; Figure \ref{Figure_example_coverage}B).
+Figure \ref{Figure_example_coverage} shows randomly distributed OTUs with available coded anatomical characters in Primates (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered OTUs with available coded anatomical characters in Carnivora (mainly Canidae and Urisdae but no Herpestidae; Figure \ref{Figure_example_coverage}B).
 
@@
 \centering
-    \includegraphics[width=1\textwidth]{example_coverage.pdf}
-\caption{Phylogenetic distribution of species with available cladistic data across two mammalian orders (A: Primates; B: Carnivora).
-Edges are colored in grey when no cladistic data is available for a species and in blue when data is available.}
+   \includegraphics[width=1\textwidth]{figure1.pdf}
+\caption{Phylogenetic distribution of species with available coded anatomical characters across two orders (A: Primates; B: Carnivora).
+Blue branches indicate species with available coded anatomical characters.}
 \label{Figure_example_coverage}
@@
 Our results show that although phylogenetic relationships among living mammals are well-resolved (e.g.
-\cite{BinindaEmonds,meredithimpacts2011}) , most of the data used to build these phylogenies is molecular, and very little cladistic data is available for living mammals compared to fossil mammals (e.g.
+\cite{BinindaEmonds,meredithimpacts2011}), most of the data used to build these phylogenies is molecular, and few coded anatomical characters are available for living mammals compared to fossils (e.g.
 \cite{O'Leary08022013,ni2013oldest}).
-This has implications for building Total Evidence phylogenies containing both living and fossil mammals, as without sufficient cladistic data for living species, fossil placements in these trees are very uncertain \cite{GuillermeCooper}.
-Cladistic data coverage in living mammals varies across taxonomic levels and in its phylogenetic distribution.
-Higher taxonomic levels are always better sampled than lower ones and within these taxonomic levels, the available data is mostly randomly distributed across the phylogeny, apart from in six orders).
+This has implications for building Total Evidence phylogenies, as without sufficient overlapping anatomical characters for living and fossil species, fossil placements in these trees may be unreliable \cite{GuillermeCooper}.
 
-The number of living mammalian taxa with no available cladistic data was surprisingly high at the species-level: only six out of 28 orders have a high coverage of taxa with available cladistic data (and two of the 28 orders are monospecific!).
-This high coverage threshold of 75\% of taxa with available cladistic data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
-Beyond this threshold, there is considerable displacement of wildcard taxa (\textit{sensu} \cite{kearneyfragmentary2002}) and decreases in clade conservation \cite{GuillermeCooper}.
-Therefore we expect a high probability of topological artefacts for the placement of fossil taxa at the species-level in most mammalian orders.
-However, data coverage seems to be less of an issue at higher taxonomic levels (i.e. genus- and family-level).
-This point is important from a practical point of view because of the slight discrepancy between the neontological and palaeontological concept of species.
-While neontological species are described using morphology, genetic distance, spatial distribution and even behaviour, palaeontological species can be based only on morphological, spatial and temporal data (e.g.
-\cite{ni2013oldest}).
-Because of this, most palaeontological studies are using the genus as their smallest OTU (e.g. \cite{ni2013oldest,O'Leary08022013}).
-Thus data availability at the genus-level in living mammals should be our primary concern when aiming to build phylogenies of living and fossil taxa.
+The number of living mammalian OTUs with available coded anatomical characters was surprisingly low at the species-level: only 17\%.
+Only seven out of 28 orders have a high coverage of taxa with available coded anatomical characters.
+This high coverage threshold of 75\% of taxa with available characters represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
+Beyond this threshold, there is considerable displacement of wildcard taxa and decreased clade conservation \cite{GuillermeCooper}.
+Therefore we expect difficulties in placing fossils at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels.
+Additionally, our analyses may underestimate the problem as we do not distiguish between soft and hard tissue characters; if a living taxon has only soft tissue coded anatomical characters then it will not have overlapping data with fossils that only have hard tissues preserved.
 
-When only a few species with cladistic data are available, the ideal scenario is for them to be phylogenetically overdispersed (i.e.
-that there is data for at least every sub-clade) to maximize the possibilities of a fossil branching from the right clade.
-The second best scenario is that species with cladistic data are randomly distributed across the phylogeny. 
-In this scenario we expect no special bias in the placement of the fossil \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with cladistic data were randomly distributed across the phylogeny of each order.
-The worst case scenario for fossil placement is that species with cladistic data are phylogenetically clustered. 
-In this situation we expect two major biases to occur: first, the fossil will not be able to branch within a clade containing no data, and second, the fossil will have a higher probability, at random, of branching within the clade containing most of the available data.
-This means that fossils with uncertain phylogenetic affinities (\textit{incertae sedis}) will have a higher probability of branching within the most sampled clade just by chance.
-Our results suggest that this may be an issue, at the genus-level, in Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha. 
-For example, a Carnivora fossil will be unable to branch in the Herpestidae that has no species with cladistic data, and will also have more chance to branch, randomly, within the Canidae clade than any other clade in Carnivora (Figure \ref{Figure_example_coverage}B).
-Thus, in Total Evidence trees, placements of some carnivoran fossils should considered with caution. 
-% NC: Trying not to piss off GraHam - really it's weird stuff appearing in Carnivora that we are concerned by, not all Carnivora. TG: I think it's mainly Graham's work that focuses on canids.
-In this study, we treated all cladistic matrices as equal in a similar way to molecular matrices. 
-For example, if matrix A contained 100 characters for taxa X and Y, and matrix B contained 50 different characters for taxa X and Z, we assumed that both matrices can be combined in a supermatrix containing 150 independent characters for taxon X, 100 for taxon Y and 50 for taxon Z.
-Unfortunately, cladistic data cannot always be treated in this way because some characters may overlap.
-For example, if matrix A has a character coding for the shape of a particular morphological feature and matrix B has a character coding for the presence of this same morphological feature and a second character coding for its shape, then these three characters are non-independent compound characters \cite{Brazeau2011}.
-However, in reasonably sized matrices (\textgreater 100 characters; \cite{GuillermeCooper,harrisonamong-character2014}) it is more likely that a number of characters are consistently conserved among the different matrices and thus easily combinable.
-For example, within the Primate cladistic literature, the character \textit{p7} - the size of the $4^{th}$ lower premolar paraconid - has been used consistently for \textgreater 15 years (e.g.
-\cite{ross1998phylogenetic,marivaux2005anthropoi,ni2013oldest} and can therefore be combined among the matrices.
-A conservative approach to avoid compound characters would be to select only the most recent matrix for each group, but this would result in the loss of a lot of data.
+When few species have available coded anatomical characters, the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion) to maximize the possibilities of a fossil being placed in the correct clade.
+The second best scenario is that species with available characters are randomly distributed across the phylogeny. 
+Here we expect no bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with available coded anatomical characters were randomly distributed across the phylogeny.
+The worst case scenario for fossil placement is that species with available characters are phylogenetically clustered. 
+Then we expect two major biases: first, fossils will not be placed within a clade containing no hard tissue data, and second, fossils will have higher probability of being placed within the most sampled clade by chance.
+Our results suggest that this may be problematic at the genus-level in Carnivora, Chiroptera and Soricomorpha. 
+For example, a carnivoran fossil is unlikely to be placed in Herpestidae because they have no coded anatomical characters available.
+Instead the fossil will have a high probability of being placed on a branch that contains many anatomical characters such as within the Canidae or Ursidae (Figure \ref{Figure_example_coverage}B). 
+This is analogous to the problem of long-branch attraction/short branch repulsion, as one can think of Herpestidae as having zero-length branches for anatomical characters, and Canidae and Ursidae having long branches and thus ``attracting'' fossil placements.
 
-Despite the absence of good cladistic data coverage for living mammals, the Total Evidence methods still seems to be the most promising way of combining living and fossil data for macroevolutionary analyses.
-Following the recommendations in \cite{GuillermeCooper}, we need to code cladistic characters for as many living species possible. 
-Fortunately, data for living mammals is usually readily available in natural history collections, therefore, we propose that an increased effort be put into coding morphological characters from living species, possibly by engaging in collaborative data collection projects through web portals such as \textit{Morphobank} \cite{morphobank}.
-% NC: Still feels like it needs one final sentence to round it all off. Many something like below, but less crappy:
-Such an effort would be valuable not only to phylogeneticists, but also to any researcher focusing understanding macroevolutionary patterns and processes.%interested in the diversity of life on Earth.
+We acknowledge, however, that our analysis does not include all matrices containing anatomical characters ever published.
+Instead, our data collection procedure focused on including studies that provided easily accessible matrices, i.e.
+we did not include matrices that are only available in books, non-reusable formats (e.g.
+an image of the matrix), or matrices available only upon request from the authors.
+Matrices containing anatomical characters were more common before the advent of molecular phylogenetics, but these matrices are also more likely to be unavailable in a reusable format, thus will be missing from our analyses.
+Although this will bias our results towards lower coverage we do not think this bias will be large, as many recent morphological matrices reuse living taxa characters from older matrices (see ESM1), so many of these data will be present in our analyses.
+Additionally, these older matrices are likely to differ from more recent ones in terms of their underlying definition of homology and their coding practices (see \cite{Brazeau2011}).
+Therefore, care needs to be taken when deciding how to include these older matrices.
+
+Despite the absence of good morphological/anatomical data coverage for living mammals, the Total Evidence method still seems to be the most promising way of combining living and fossil species in macroevolutionary analyses.
+Following the recommendations in \cite{GuillermeCooper}, we should code anatomical characters for as many living species as possible. 
+Fortunately, mammal specimens are usually readily available in natural history collections, therefore, we propose increased effort into coding anatomical characters from living species, possibly by engaging in collaborative data collection projects.
+Such efforts would be valuable not only to phylogeneticists, but also to any researcher focusing on understanding macroevolutionary patterns and processes.
 
@@
 \section{Ethics statement}
+N/A
 \section{Data accessibility statement}
-All data and analysis code is available on GitHub (\url{https://github.com/TGuillerme/Missing_living_mammals}).
+All data and code are available on GitHub (\url{https://github.com/TGuillerme/Missing_living_mammals}).
 \section{Authors' Contributions}
-T.G.
-and N.C conceived and designed the experiments.
-T.G.
-performed the experiments and analysed the data.
-T.G.
-and N.C.
-contributed to the writing of the manuscript.
-All authors approved the final version of the manuscript.
+TG and NC designed the study. TG analysed the data. TG and NC wrote the the manuscript.
 \section{Competing Interests}
 We have no competing interests.
-\section{Acknowledgements}
-David Bapst, Graeme Lloyd, Nick Matzke and April Wright.
+\section{Acknowledgments}
+We thank David Bapst, Graeme Lloyd, Nick Matzke, April Wright, Graham Slater, Peter Wagner and one anonymous reviewer.
 \section{Funding statement}
-This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).
+This was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal: 321696).
+
+\section{References}
 

### 2016-04-04 Manuscript/Table_results.tex
Cleaned up manuscript file

@@
 % latex table generated in R 3.2.3 by xtable 1.8-0 package
-% Mon Feb  8 17:25:14 2016
+% Mon Feb  8 22:09:15 2016
 \begin{longtable}{lL{1.8cm}C{2cm}lcc}
-\caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels.
-The coverage represents the proportion of taxa with available morphological data.
-The left vertical bar represents 25\% of available data (``low'' coverage if \textless 25\%); The right vertical bar represents 75\% of available data (``high'' coverage if \textgreater 75\%).
-When the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) are negative, taxa are more phylogenetically dispersed than expected by chance; when NRI or NTI are positive, taxa are more phylogenetically clustered by expected by chance.
-Significant NRI or NTI are highlighted in bold.
-One star (*) represents a p-value between 0.05 and 0.005; two starts between 0.005 and 0.0005 and three stars a p-value less than 0.0005.} \\
+\caption{Number of taxa with available discrete morphological data for mammalian orders at three
+taxonomic levels.
+The left vertical bar represents low coverage (\textless 25\%; coloured in blue); the right vertical bar represents high coverage (\textgreater 75\%; coloured in orange).
+Negative Net Relatedness Index (NRI) and Nearest Taxon Index (NTI) values indicate phylogenetic overdispersion; positive values indicate phylogenetic clustering.
+Significant NRI or NTI values are in bold.
+*p \textless 0.05; **p \textless 0.01; ***p \textless 0.001.
+} \\
+
   \hline
@@
   \hline
+  Mammalia (class) & family & 129/148 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar40.pdf} & -1.19 & 1.09 \\ 
+  \textbf{Mammalia (class)} & \textbf{genus} & \textbf{517/1186} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar41.pdf} & \textbf{-5.19} & \textbf{3.71**} \\ 
+  \textbf{Mammalia (class)} & \textbf{species} & \textbf{847/5017} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar42.pdf} & \textbf{-7.75} & \textbf{3.54**} \\ 
   Afrosoricida & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar1.pdf} &   &   \\ 
   Afrosoricida & genus & 17/17 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar2.pdf} &   &   \\ 
-  Afrosoricida & species & 23/42 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar3.pdf} & 1.75 & 1.08 \\ 
-  Carnivora & family & 14/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar4.pdf} & 0.63 & 0.6 \\ 
-  \textbf{Carnivora} & \textbf{genus} & \textbf{54/125} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar5.pdf} & \textbf{4.81**} & \textbf{1.78*} \\ 
-  \textbf{Carnivora} & \textbf{species} & \textbf{76/283} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar6.pdf} & \textbf{7.66**} & \textbf{0.85} \\ 
+  Afrosoricida & species & 23/42 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar3.pdf} & 1.52 & 1.1 \\ 
+  Carnivora & family & 14/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar4.pdf} & 0.65 & 0.55 \\ 
+  \textbf{Carnivora} & \textbf{genus} & \textbf{52/125} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar5.pdf} & \textbf{4.27**} & \textbf{1.26} \\ 
+  \textbf{Carnivora} & \textbf{species} & \textbf{75/283} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar6.pdf} & \textbf{7.24**} & \textbf{0.8} \\ 
   Cetartiodactyla & family & 21/21 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar7.pdf} &   &   \\ 
-  Cetartiodactyla & genus & 100/128 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar8.pdf} & 0.85 & 0.94 \\ 
-  \textbf{Cetartiodactyla} & \textbf{species} & \textbf{171/310} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar9.pdf} & \textbf{1.92*} & \textbf{-0.46} \\ 
-  Chiroptera & family & 15/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar10.pdf} & -0.28 & 0.56 \\ 
-  \textbf{Chiroptera} & \textbf{genus} & \textbf{93/202} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar11.pdf} & \textbf{13.47**} & \textbf{1.1} \\ 
-  \textbf{Chiroptera} & \textbf{species} & \textbf{215/1053} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar12.pdf} & \textbf{8.82**} & \textbf{1.22} \\ 
+  Cetartiodactyla & genus & 97/128 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar8.pdf} & 0.7 & 1.28 \\ 
+  \textbf{Cetartiodactyla} & \textbf{species} & \textbf{169/310} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar9.pdf} & \textbf{1.82*} & \textbf{-0.24} \\ 
+  Chiroptera & family & 15/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar10.pdf} & -0.23 & 0.61 \\ 
+  \textbf{Chiroptera} & \textbf{genus} & \textbf{92/202} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar11.pdf} & \textbf{13.07**} & \textbf{0.99} \\ 
+  \textbf{Chiroptera} & \textbf{species} & \textbf{214/1053} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar12.pdf} & \textbf{9.21**} & \textbf{1.27} \\ 
   Cingulata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar13.pdf} &   &   \\ 
-  Cingulata & genus & 8/9 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar14.pdf} & 1.51 & -1.57 \\ 
-  \textbf{Cingulata} & \textbf{species} & \textbf{9/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar15.pdf} & \textbf{1.9*} & \textbf{0.11} \\ 
+  Cingulata & genus & 8/9 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar14.pdf} & 1.48 & -1.54 \\ 
+  \textbf{Cingulata} & \textbf{species} & \textbf{9/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar15.pdf} & \textbf{2.06*} & \textbf{0.2} \\ 
   Dasyuromorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar16.pdf} &   &   \\ 
-  Dasyuromorphia & genus & 8/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar17.pdf} & -0.75 & -1.07 \\ 
-  Dasyuromorphia & species & 9/64 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar18.pdf} & -0.88 & -0.34 \\ 
+  Dasyuromorphia & genus & 8/22 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar17.pdf} & -0.78 & -1.06 \\ 
+  Dasyuromorphia & species & 9/64 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar18.pdf} & -0.86 & -0.37 \\ 
   Dermoptera & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar19.pdf} &   &   \\ 
@@
   Lagomorpha & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar34.pdf} &   &   \\ 
-  Lagomorpha & genus & 5/12 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar35.pdf} & -1.06 & -0.95 \\ 
-  Lagomorpha & species & 12/86 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar36.pdf} & -0.62 & -1.88 \\ 
+  Lagomorpha & genus & 5/12 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar35.pdf} & -0.95 & -0.94 \\ 
+  Lagomorpha & species & 12/86 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar36.pdf} & -0.62 & -1.96 \\ 
   Macroscelidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar37.pdf} &   &   \\ 
   Macroscelidea & genus & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar38.pdf} &   &   \\ 
-  Macroscelidea & species & 12/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar39.pdf} & -1.3 & -1.06 \\ 
-  Microbiotheria & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar40.pdf} &   &   \\ 
-  Microbiotheria & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar41.pdf} &   &   \\ 
-  Microbiotheria & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar42.pdf} &   &   \\ 
-  Monotremata & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar43.pdf} &   &   \\ 
-  Monotremata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar44.pdf} & -0.72 & -0.69 \\ 
-  Monotremata & species & 2/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar45.pdf} & -0.97 & -0.97 \\ 
-  Notoryctemorphia & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar46.pdf} &   &   \\ 
-  Notoryctemorphia & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar47.pdf} &   &   \\ 
-  Notoryctemorphia & species & 0/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar48.pdf} &   &   \\ 
-  Paucituberculata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar49.pdf} &   &   \\ 
-  Paucituberculata & genus & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar50.pdf} &   &   \\ 
-  Paucituberculata & species & 5/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar51.pdf} &   &   \\ 
-  Peramelemorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar52.pdf} &   &   \\ 
-  Peramelemorphia & genus & 7/7 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar53.pdf} &   &   \\ 
-  Peramelemorphia & species & 16/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar54.pdf} & -0.13 & 0.97 \\ 
-  Perissodactyla & family & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar55.pdf} &   &   \\ 
-  Perissodactyla & genus & 6/6 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar56.pdf} &   &   \\ 
-  Perissodactyla & species & 10/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar57.pdf} & -0.07 & -2.63 \\ 
-  Pholidota & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar58.pdf} &   &   \\ 
-  Pholidota & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar59.pdf} &   &   \\ 
-  Pholidota & species & 4/8 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar60.pdf} & 1.18 & 0.94 \\ 
-  Pilosa & family & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar61.pdf} & 1.87 & 2 \\ 
-  Pilosa & genus & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar62.pdf} & -0.96 & 0.36 \\ 
-  \textbf{Pilosa} & \textbf{species} & \textbf{5/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar63.pdf} & \textbf{1.28} & \textbf{2.38**} \\ 
-  Primates & family & 15/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar64.pdf} &   &   \\ 
-  Primates & genus & 48/68 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar65.pdf} & -0.35 & -1.33 \\ 
-  Primates & species & 64/351 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar66.pdf} & -0.67 & -1.27 \\ 
-  Proboscidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar67.pdf} &   &   \\ 
-  Proboscidea & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar68.pdf} &   &   \\ 
-  Proboscidea & species & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar69.pdf} & -0.69 & -0.69 \\ 
-  Rodentia & family & 18/32 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar70.pdf} & 0.66 & -0.98 \\ 
-  Rodentia & genus & 82/450 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar71.pdf} & -1.66 & 1.55 \\ 
-  \textbf{Rodentia} & \textbf{species} & \textbf{90/2094} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar72.pdf} & \textbf{2.76**} & \textbf{2.34**} \\ 
-  Scandentia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar73.pdf} &   &   \\ 
-  Scandentia & genus & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar74.pdf} & -0.74 & -0.74 \\ 
-  Scandentia & species & 3/20 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar75.pdf} & -1.88 & -0.84 \\ 
-  Sirenia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar76.pdf} &   &   \\ 
-  Sirenia & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar77.pdf} &   &   \\ 
-  Sirenia & species & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar78.pdf} &   &   \\ 
-  Soricomorpha & family & 3/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar79.pdf} & -0.98 & -0.99 \\ 
-  \textbf{Soricomorpha} & \textbf{genus} & \textbf{19/43} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar80.pdf} & \textbf{7.11**} & \textbf{2.59**} \\ 
-  \textbf{Soricomorpha} & \textbf{species} & \textbf{21/392} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar81.pdf} & \textbf{10.65**} & \textbf{3.56**} \\ 
-  Tubulidentata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar82.pdf} &   &   \\ 
-  Tubulidentata & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar83.pdf} &   &   \\ 
-  Tubulidentata & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar84.pdf} &   &   \\ 
+  Macroscelidea & species & 12/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar39.pdf} & -1.24 & -1.2 \\ 
+  Microbiotheria & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar43.pdf} &   &   \\ 
+  Microbiotheria & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar44.pdf} &   &   \\ 
+  Microbiotheria & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar45.pdf} &   &   \\ 
+  Monotremata & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar46.pdf} &   &   \\ 
+  Monotremata & genus & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar47.pdf} & -0.68 & -0.69 \\ 
+  Monotremata & species & 2/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar48.pdf} & -1.01 & -1 \\ 
+  Notoryctemorphia & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar49.pdf} &   &   \\ 
+  Notoryctemorphia & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar50.pdf} &   &   \\ 
+  Notoryctemorphia & species & 0/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar51.pdf} &   &   \\ 
+  Paucituberculata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar52.pdf} &   &   \\ 
+  Paucituberculata & genus & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar53.pdf} &   &   \\ 
+  Paucituberculata & species & 5/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar54.pdf} &   &   \\ 
+  Peramelemorphia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar55.pdf} &   &   \\ 
+  Peramelemorphia & genus & 7/7 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar56.pdf} &   &   \\ 
+  Peramelemorphia & species & 16/18 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar57.pdf} & -0.14 & 0.91 \\ 
+  Perissodactyla & family & 3/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar58.pdf} &   &   \\ 
+  Perissodactyla & genus & 6/6 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar59.pdf} &   &   \\ 
+  Perissodactyla & species & 10/16 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar60.pdf} & -0.1 & -2.77 \\ 
+  Pholidota & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar61.pdf} &   &   \\ 
+  Pholidota & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar62.pdf} &   &   \\ 
+  Pholidota & species & 4/8 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar63.pdf} & 1.14 & 0.97 \\ 
+  Pilosa & family & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar64.pdf} & 2.01 & 1.96 \\ 
+  Pilosa & genus & 4/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar65.pdf} & -0.91 & 0.36 \\ 
+  \textbf{Pilosa} & \textbf{species} & \textbf{5/29} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar66.pdf} & \textbf{1.18} & \textbf{2.35**} \\ 
+  Primates & family & 15/15 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar67.pdf} &   &   \\ 
+  Primates & genus & 48/68 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar68.pdf} & -0.37 & -1.39 \\ 
+  Primates & species & 64/351 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar69.pdf} & -0.66 & -1.4 \\ 
+  Proboscidea & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar70.pdf} &   &   \\ 
+  Proboscidea & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar71.pdf} &   &   \\ 
+  Proboscidea & species & 2/3 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar72.pdf} & -0.67 & -0.72 \\ 
+  Rodentia & family & 18/32 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar73.pdf} & 0.66 & -0.95 \\ 
+  \textbf{Rodentia} & \textbf{genus} & \textbf{82/450} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar74.pdf} & \textbf{-1.81} & \textbf{1.7*} \\ 
+  \textbf{Rodentia} & \textbf{species} & \textbf{90/2094} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar75.pdf} & \textbf{2.66**} & \textbf{2.36**} \\ 
+  Scandentia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar76.pdf} &   &   \\ 
+  Scandentia & genus & 2/5 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar77.pdf} & -0.77 & -0.76 \\ 
+  Scandentia & species & 3/20 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar78.pdf} & -2 & -0.8 \\ 
+  Sirenia & family & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar79.pdf} &   &   \\ 
+  Sirenia & genus & 2/2 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar80.pdf} &   &   \\ 
+  Sirenia & species & 4/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar81.pdf} &   &   \\ 
+  Soricomorpha & family & 3/4 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar82.pdf} & -0.98 & -0.97 \\ 
+  \textbf{Soricomorpha} & \textbf{genus} & \textbf{19/43} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar83.pdf} & \textbf{7.07**} & \textbf{2.64**} \\ 
+  \textbf{Soricomorpha} & \textbf{species} & \textbf{21/392} & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar84.pdf} & \textbf{10.17**} & \textbf{3.36**} \\ 
+  Tubulidentata & family & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar85.pdf} &   &   \\ 
+  Tubulidentata & genus & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar86.pdf} &   &   \\ 
+  Tubulidentata & species & 1/1 & \includegraphics[width=0.20\linewidth, height=0.05\linewidth]{Table_figures/bar87.pdf} &   &   \\ 
    \hline

### 2016-04-04 Manuscript/mediaSummary.tex
Cleaned up manuscript file

\documentclass{article}

\begin{document}

Scientists use family trees to understand changes in biodiversity through time.
But the trees often only use living species, ignoring extinct groups like mammoths.Therefore we must place living and extinct species in the same trees.
We can do so using the DNA from living species and anatomical data from living and extinct species
However, we found that few living mammals have enough available anatomical data to determine how they are related to extinct species
We suggest that we must collect more anatomical data from living species in museum collections to efficiently create family trees of living and extinct species.

\end{document}

### 2016-04-04 Manuscript/response2.tex
Cleaned up manuscript file

--- 2016-04-04 Manuscript/BiologyLetters/response2.tex
@@
 \textit{``A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared among both living and fossil taxa (i.e.
-hard tissue characters such as skeletal features).''}
-lines @@@.
-% NC: Add line numbers
+hard tissue characters such as skeletal features).''} lines 37-39.
 
@@
 We add the following sentence to the discussion:
-\textit{``Additionally, our analyses may underestimate the problem as we do not distiguish between soft and hard tissue characters; if a living taxon has only soft tissue coded anatomical characters then it will not have overlapping data with fossils that only have hard tissues preserved.''}
-lines @@@.
-% NC: Add line numbers
+\textit{``Additionally, our analyses may underestimate the problem as we do not distiguish between soft and hard tissue characters; if a living taxon has only soft tissue coded anatomical characters then it will not have overlapping data with fossils that only have hard tissues preserved.''} lines 134-137.
 \item{\textcolor{blue}{I'm also concerned that my original comments about the biases in character sets included still aren't adequately addressed; they now acknowledge that they didn't include any character sets that weren't available in print, but that doesn't address what kind of difference it would make to their results.
@@
 
-% NC: I've rewritten this to make it clearer what you have done and to stop it reading like "yeah we know there is a bias but we are too lazy to fix it".
-Now it says "yeah we know there is a bias but we are too lazy to fix it, but we also think it isn't a big bias and would add more bias anyway".
-A much better justification.
 We acknowledge this comment (and Referee 3 comment 2) and have now made more explicit reference to this bias. 
@@
 Additionally we do not believe this bias will be as large as the reviewer fears.
-Many recent matrices reuse characters from older matrices, thus it is likely most of the data from major previous studies are present in our analyses.
+Many recent matrices reuse taxa and characters from older matrices, thus it is likely most of the data from major previous studies are present in our analyses.
 We also share the concerns of Peter Wagner that using ``old'' matrices may introduce a different kind of bias into phylogenies, as definitions of homology and characters have changed over time, making the kind of matrix we are unlikely to find with our search protocol harder to incorporate into Total Evidence analyses.
@@
 Additionally, these older matrices are likely to differ from more recent ones in terms of their underlying definition of homology and their coding practices (see [19]).
-Therefore, care needs to be taken when deciding how to include these older matrices.''} lines @@@.
-% NC: Add line numbers
+Therefore, care needs to be taken when deciding how to include these older matrices.''} lines 163-168.
 \item{\textcolor{blue}{I don't agree with the assertion that the genus level should be our primary concern; more and more studies are building species-level trees with fossil taxa.

### 2016-04-04 Manuscript/CiteChange.sh

@@ -1,79 +1,0 @@
-#!/usr/bin/sh
-##########################
-#Swap the citation types between sysbiol or vancouver for a tex file
-##########################
-#SYNTAX:
-#sh CiteChange.sh <tex>
-#with:
-#<tex> being the tex file
-#########################
-#version 0.2
-#----
-#guillert(at)tcd.ie - 23/06/2015
-##########################
-
-#Input value
-tex=$1
-
-#Check which is the actual cite format
-if grep 'bibliographystyle' ${tex} | grep 'vancouver' > /dev/null
-then
-    #Change from vancouver to sysbio style
-    echo "Changing bibliography style from 'vancouver' to 'sysbio'."
-    sed 's/\\bibliographystyle{vancouver}/\\bibliographystyle{sysbio}/g' $tex > ${tex}.tmp
-
-    #rename the tex file
-    mv ${tex}.tmp ${tex}
-
-    #enable natbib
-    header=$(grep -n "documentclass" $tex | sed 's/\:\\documentclass.*//g')
-    sed ''"$header"'s/$/\'$'\n\\\usepackage\{natbib\}/' ${tex} > ${tex}.tmp
-
-    #rename the tex file
-    mv ${tex}.tmp ${tex}
-
-    #change all the \cite into \citep
-    sed 's/\\cite/\\citep/g' ${tex} > ${tex}.tmp 
-
-    #rename the tex file
-    mv ${tex}.tmp ${tex}
-
-    #correct the e.g. (1)
-    sed 's/(e.g. \\citep/\\citep[e.g.][]/g' ${tex} | sed 's/})/}/g' > ${tex}.tmp 
-
-    #rename the tex file
-    mv ${tex}.tmp ${tex}
-
-    #Done!
-    echo "Remember to add the 'sysbio.bst' file to the current directory."
-
-else
-    #Change from sysbio to vancouver style
-    echo "Changing bibliography style from 'sysbio' to 'vancouver':"
-    sed 's/\\bibliographystyle{sysbio}/\\bibliographystyle{vancouver}/g' $tex > ${tex}.tmp
-
-    #rename the tex file
-    mv ${tex}.tmp ${tex}
-
-    #disable natbib
-    sed 's/\\usepackage{natbib}//g' ${tex} > ${tex}.tmp
-
-    #rename the tex file
-    mv ${tex}.tmp ${tex}
-
-    #change all the \citep into \cite
[…]

### 2016-04-04 Manuscript/Table_LaTeX/Table_results_BW.tex
Fixing BW table legend

--- 2016-02-14 Manuscript/BiologyLetters/Table_LaTeX/Table_results_BW.tex
@@
 \begin{longtable}{lL{1.8cm}C{2cm}lcc}
-\caption{Number of taxa with available cladistic data for mammalian orders at three
-taxonomic levels.
-The left vertical bar represents low coverage (\textless 25\%; coloured in blue); the right vertical bar represents high coverage (\textgreater 75\%; coloured in orange).
+\caption{Number of taxa with available discrete morphological data for mammalian orders at three taxonomic levels.
+The left vertical bar represents low coverage (\textless 25\%; coloured in dark grey); the right vertical bar represents high coverage (\textgreater 75\%; coloured in light grey).
 Negative Net Relatedness Index (NRI) and Nearest Taxon Index (NTI) values indicate phylogenetic overdispersion; positive values indicate phylogenetic clustering.

### 2016-04-07 Functions/table.results.R
Update table.results.R

@@ -10,7 +10,7 @@
 #<environement> table environement
 #<col.threshold> a vector of colours that must be the length of threshold + 1
 ##########################
-#environement(at)tcd.ie - 05/06/2015
+#guillert(at)tcd.ie - 05/06/2015
 ##########################
 
 table.result<-function(results, metric, threshold=c(25,75), save.path, file.save, caption, environement, col.threshold=c("grey", "grey", "grey")) {
